### 2024-09-10 paper/main.tex
adding latex file for the paper

\documentclass[draft]{aa}

% \usepackage[draft]{graphicx}
% \usepackage{txfonts}
\usepackage{amsmath}

\title{Palomar 5 Gaps}

\subtitle{Globular clusters as hole punchers}

\author{S. Ferrone
       \inst{1,2}
       \and
       P. Di Matteo\inst{2}
       \and
       M. Montuori\inst{1}
       }

\institute{Dipartimento di Fisica, Universit\`a di Roma ``La Sapienza'',
           Piazza Aldo Moro\\
           \email{salvatore.ferrone@uniroma1.it}
      \and
          Paris Observatory. Paris Sciences et Lettres\\
          \email{c.ptolemy@hipparch.uheaven.space}
          \thanks{The university of heaven temporarily does not
                  accept e-mails}
          }

\date{Born in 1996; Accepted in 2018}

\begin{document}

\abstract
  {Holy moly artichokey}
 % aims heading (mandatory)
  {ok}
 % methods heading (mandatory)
  {good}
 % results heading (mandatory)
  {alright}
 % conclusions heading (optional), leave it empty if necessary
  {done}

\maketitle
\section{Introduction}

  The presence of dark matter has substantial observational evidence on scales larger than individual galaxies. The aim of this work is to add more context to aid in the search for evidence of dark matter on the sub-galactic scale and within the Milky Way. Initially proposed by [Author Name] in [Year], the concept of dark matter emerged to explain the anomalously high velocities of stars in distant dwarf galaxies. These velocities exceed what would be expected given the visible mass exerting gravitational forces on them. Furthermore, dark matter's influence extends to cosmological scales, as evidenced by the bending of light trajectories within cosmic web filaments due to the mass of dark matter. Despite significant advancements, understanding the true nature of dark matter remains elusive, with the scientific community continually seeking evidence of its existence beyond its gravitational effects on cosmic and extragalactic scales. A current focus is the detection of dark matter at sub-galactic scales, such as through the study of stellar streams.

  Stellar streams, which are long, thin structures formed from the tidal disruption of globular clusters or dwarf galaxies as they orbit a host galaxy, offer a promising avenue for probing the distribution of dark matter. Tidal forces arise because these stellar systems are extended objects, leading to a differential gravitational pull across their structure. As a result, stars farther from the galactic center lag behind the cluster, while those closer are pulled away. This process stretches the cluster in the radial direction, creating two tidal tails that trace the orbit of the cluster. Theoretical predictions of this phenomenon have long existed, but it was not until the early 1995 that streams were observed for the first time \citet{carl_j_grillmair_globular_1995}. Then, \citet{odenkirchen_detection_2001} discovered that stars beyond the tidal radius of the Palomar 5 cluster shared the same spectroscopic characteristics as those within it.

  Over the past two decades, both theoretical and numerical studies have demonstrated the utility of stellar streams in constraining the gravitational field of host galaxies, particularly the Milky Way. Galactic dynamics, governed by Poisson's equation, allow the decomposition of a galaxy into components such as the central bulge, disk, and dark matter halo. The dark matter halo, often modeled as spherical, is the largest component and plays a critical role in the observed constant rotational velocity of galaxies. However, determining whether the halo might be an oblate spheroid could provide insight into the self-interaction properties of dark matter. Numerical studies, such as those by \citet{varghese2011stellar} who explored the extent to which stellar streams can constrain the galactic potential under varying levels of observational data. Meanwhile, Bonaca et al. approached stellar streams from an information-theoretic perspective, identifying the orbits and configurations that yield the most information about the galactic potential.

  The release of Gaia Data Release 2 (DR2) and DR3 significantly advanced the observational study of stellar streams. The StreamFinder algorithm has identified 90 stellar streams, with some of these candidates being followed up by other instruments and surveys. For instance, streams like GD-1 and Palomar 5 have approximately 1,000 to 3,000 currently associated members, although there is an observational bias toward brighter stars, such as red giants.

  The goal of this study is to contribute to the ongoing effort to use stellar streams as probes for dark matter on sub-galactic scales. According to the cold dark matter model, which is supported by observations, galaxies form hierarchically, leading to a mass distribution of dark matter subhalos that follows a power-law. This suggests that numerous small dark matter clumps should exist within galaxies. \citet{rodrigo_ibata_uncovering_2002} suggested that the presence of said dark matter subhaloes dark matter subhalos can influence the stellar streams by diffussing their distribution of orbital elements. \citet{r_g_carlberg_pal_2012} extended this idea by noting that subhaloes could have flybys and that leave gaps in stellar streams and subsquently explored the statistical distribution the number of gaps gaps given a probabilistic model of the rate of encounters given based on the expected distribution of dark matter subhalos. \citet{bonaca_spur_2018} took a more direct observational approach, identifying an underdensity in the GD-1 stream that could not be explained by known objects like globular clusters and was inconsistent with the mass-size relationship predicted by the Lambda-CDM model.

  In this study, we aim to investigate the potential influence of globular clusters on stellar streams, specifically examining whether galactic globular clusters can produce gaps similar to those expected from dark matter subhalos. If globular clusters do create significant gaps, this must be considered when using stellar streams to infer the local presence of dark matter. Our focus is on the Palomar 5 stream, which has been well-studied and has observable gaps, a high membership count, and favorable visibility due to its position above the galactic disk. In the following sections, we describe the modeling of this stream, the gravitational interactions with the galaxy and other globular clusters, and the statistical analysis of the perturbations affecting the stream.

\section{Methods}

  \subsection{Numerical Methodology}
    Our numerical methodology follows the approach outlined in Ferrone et al. (2023), which we briefly summarize here for completeness. We begin by extracting the phase space properties (positions, velocities) as well as the masses and half-mass radii of 165 globular clusters from the dataset provided by Baumgardt and Vasiliev et al. The dataset includes the clusters' masses, radial velocities, distances, proper motions in RA and DEC, and their associated covariance terms.

    To generate initial conditions, we treat the five-dimensional phase space variables (position and velocity components) for each globular cluster as Gaussian distributions. Using the \texttt{np.random.gauss} function, we generate 50 samples per cluster, producing 50 sets of initial conditions for the entire globular cluster system. These samples are stored in data files for later use.

    Next, we convert the initial conditions into a galactocentric reference frame. The local standard of rest was computed using XXX, with the Sun's peculiar velocity set to XXX, and the Sun's distance from the Galactic center taken as XXX. These transformations were performed using the \texttt{astropy} package.

    For the galactic potential, we employed the static model proposed by Pouliasis et al. (2017), which consists of a superposition of a thin disk, thick disk, and a dark matter halo. This model is assumed to be time-independent over the course of our simulations.

    The primary methodological departure from Ferrone et al. (2023) is that now we include globular cluster interactions. Instead of treating the clusters as point masses that only experience the gravitational potential of the galaxy, we now compute N-body interactions between the clusters during the backward integration over 5 billion years. Although these interactions were included, we observed no significant exchanges of orbital energy or momentum between clusters. The resulting orbits were saved to data files for subsequent analysis.

    At the computed positions 5 Gyr ago, each globular cluster was replaced by a system of 100,000 particles, sampled from a Plummer distribution to represent the stellar content of the cluster. We then integrated the evolution of these particles forward in time to the present day. During this forward integration, each star particle feels the gravitational influence of the galaxy and all globular clusters, but not other star particles. We do not perform direct N-body simulations between star particles; instead, at each time step, we compute the positions of all globular clusters and apply their forces to the star particles.

    The integration was performed using a leapfrog integrator. The timestep was chosen to be sufficiently small to ensure both the conservation of energy and that the integration error over 5 Gyr was negligible compared to the scale radius of the globular clusters. Further details on the timestep selection are discussed in the appendix.

    During the integration, intermediate snapshots were saved to facilitate the analysis of stellar streams and the effects of cluster impacts. Specifically, for each of the 50 realizations of the Palomar 5 stream, we saved 5,000 intermediate timesteps, resulting in approximately XX GB of data ($N_p \times N_s \times 7 \times \text{float}$).

  \subsection{Gap detection}

    At the end of the simulations, some gaps in the stellar stream are so prominent that they can be identified visually without any sophisticated quantitative analysis. However, upon closer inspection, more subtle gaps emerge, highlighting the need for quantitative methods to systematically detect and assess these features.

    We initially sought an automated, quantitative approach from signal and image processing that could reliably detect all gaps and evaluate their significance. Unfortunately, despite extensive efforts, an ideal method remained elusive. As a result, we opted for a simpler approach: a combination of visual inspection aided by comparative analysis.

    To facilitate this process, we ran a series of \texttt{vanilla} simulations. In these control simulations, we used the same 50 sets of initial conditions but excluded interactions between globular clusters and omitted the gravitational influence of the clusters on the star particles within the stream. By comparing the morphological differences between these control simulations and the full simulations (which included cluster interactions), we could identify and assess the impact of globular clusters on the stream structure more effectively.

    Comparisons between the \texttt{vanilla} and full simulations  were performed in the tail coordinate system. This system allows for consistent morphological comparisons by aligning the stream with the cluster's orbit. The coordinate system is based on the work of Dehnen et al. (2004) and is detailed in Appendix~\ref{appendix:TailCoordinates}.

    Briefly, the x'coordinate represents the position of a particle along the orbit relative to the globular cluster—whether it is ahead or behind the cluster. The y'coordinate measures the particle's distance within the orbital plane, where positive values indicate that the particle is farther from the Galactic center and negative values indicate that it is closer.

    \subsubsection*{Density maps and profiles}
      For each of the 50 simulations, we compare the final snapshots between the \texttt{vanilla} and full simulations in the tail coordinate system. This is done by creating 2D density maps of the streams, followed by marginalizing over the y'-coordinate to produce 1D density profiles along the x'-axis. These comparisons are illustrated in the results section (see Fig.~XXX).

      To construct the 1D density profiles, we first bin the data using the  $\sqrt{N}$-rule for determining the number of bins. After histogramming the 1D profiles, we apply a boxcar smoothing technique, where each central value is replaced by the median of the surrounding XX adjacent data points, corresponding to a smoothing length of approximately 300 parsecs. This procedure reduces high-frequency noise and smooths the profiles, as shown in Fig.~XX, where the center of mass has been effectively removed.

      With the smoothed 1D profiles in hand, we proceed to compare them by identifying regions where the full simulations (GCGC) show significant underdensity compared to the \texttt{vanilla} simulations. To achieve this, we compute the difference between the logarithms of the binned counts for each profile and search for points that fall more than two standard deviations below the mean. However, we only consider regions that surpass a predefined significance threshold, where the significance is determined by the uncertainties in the binned counts governed by Poisson statistics multiplied by a factor of 50. This noise floor was chosen to eliminate the low-density regions near Palomar 5, where gaps are more likely caused by epicyclic oscillations rather than impacts from globular clusters. The identified gap candidates are marked and labeled on the plots.

      This method, however, is not without limitations, particularly when detecting smaller gaps. As discussed in Erkal et al. (2015), a gap's size is not necessarily indicative of a weak impact but could result from a more recent interaction. Additionally, because our streams have finite width, some gaps may be oriented obliquely relative to the stream axis. In these cases, marginalizing over y' can dilute the gap's signal, making it difficult to detect. An example of this is the small gap around 5 kpc, which, as we will demonstrate later, is caused by the interaction with NGC 2808 and appears in most simulations.

    \subsubsection*{From Suspect to Culprit} \label{SuspectToCulprit}

      A key question we seek to answer is: from the gaps present at the end of the simulation, who caused them? To address this, we examine the evolution of stream density over time. Instead of using the x'coordinate, we introduce $\tau$, which represents time rather than distance. Specifically, $\tau$ indicates how long it will take for a cluster to reach, or how long ago it passed, a given point on its orbit. This is advantageous because the growth of the stream is approximately linear in $\tau$ whereas in physical space, streams on eccentric orbits expand and contract depending on the orbital phase.

      \citet{sanders2016dynamics} extended the analysis of Erkal et al. (2015), demonstrating that action-angle variables provide a useful coordinate system for analyzing stream evolution, as actions are conserved and the angles associated with the stream's growth evolve linearly over time. Although we became aware of this work only after completing our analysis, we note that $\tau$ is a suitable approximation and behaves similarly to the angle variable corresponding to the azimuthal action: $\tau \approx \theta_{\phi,i} - \theta_{\phi,\text{GC}}$

      The core of our analysis is presented in Fig.~XXX. The bottom panel shows the evolution of the stream density over time. To avoid extreme low-density regions at the stream's edges, we applied a density threshold to focus on the more significant areas of the stream. Next, we modeled Palomar 5's orbit as a proxy for its stream and sampled points along the orbit to measure the gravitational force exerted by other globular clusters. This force data is displayed in the top panel of Fig.~XXX, showing how the total gravitational acceleration on Palomar 5's stream evolves over its length throughout the simulation.

      We then used \texttt{scipy}'s \texttt{ndimage} package to identify the top five local maxima in the data space of gravitational acceleration $\vec{g}$ as a function of time $t$ and the stream coordinate $\tau$. Once these local maxima were identified, we iterated over the contributions of individual globular clusters to determine which cluster contributed the most to each peak in $\vec{g}$. Each significant peak was labeled with the corresponding globular cluster.

      Afterward, we cross-referenced these peaks with the previously identified gaps in the stream. For large gaps resulting from strong interactions, we observed that after an impact, a low-density wake is left behind in the ($t,\tau$) plane.

      The results of this analysis were compiled into a table. If a gap was attributed to a particular globular cluster, it was labeled as \texttt{TRUE}; otherwise, it was marked as \texttt{FALSE}. For a handful of simulations, to double check that the verdict made correctly when passing from suspect to culprit, we re-compute the simulations yet individually adding one globular cluster at a time. As a result, we are able to confirm that singular gaps arrise from the suspected clusters.

  \subsection{Reconstruction of the impact geometry}
    In its simplest form, impact theory, as presented in works such as Binney and Tremaine, John Taylor, and Galaxies, considers two particles: one stationary and the other moving past it. The distance between the two particles at their closest approach is known as the impact parameter. To simplify the analysis, the \textit{impulse approximation} is often employed, which assumes that the velocity of the perturber remains unchanged during the interaction. This assumption greatly simplifies the computation.

    To understand how the impacted particle is perturbed, one needs to compute its change in momentum, which is determined by integrating the force acting on the particle over the duration of the interaction. A useful approximation for this change in momentum is the force at the closest approach multiplied by an estimate of the interaction time:
    \begin{equation} 
      \Delta p \approx \text{Force} \times \text{interaction time} = \frac{GMm}{b^2} \times \frac{b}{v} = \frac{GM}{bv}, 
      \end{equation}
    where $M$ is the mass of the perturber, $m$ is the mass of the particle, $b$ is the impact parameter, $v$ is the relative velocity of the perturber with respect to the particle, and $G$ is the gravitational constant.

    From this analysis, we can conclude that a more massive perturber, passing closer to the particle, and moving more slowly, will have a greater impact. It's important to note that the momentum change is inversely proportional to the velocity of the perturber, which contrasts with the intuition from elastic collisions, such as those between billiard balls, where higher velocities generally result in greater impact.

    Erkal et al. (2015) extended this impact theory to a one-dimensional stream, studying how the momentum change of a particle varies as a function of its distance from the point of impact. In their analysis, the perturber was modeled not as a point mass, but as a Plummer sphere with a characteristic length scale $r_p$. Since the stream is not zero-dimensional and has length, both the parallel and perpendicular components of the velocity must be considered. Consequently, five parameters determine the change in velocity of a given particle: $M$, $r_p$, $b$, $W_\parallel$, and $W_\perp$. Our goal is to constrain these parameters during the impact in our simulations and compare them to close approaches that do not generate gaps, in order to establish the criteria necessary for gap formation.

    To achieve this, we begin by identifying the approximate moments of impact from the most significant clusters, as determined in the previous analysis in Sec.~\ref{SuspectToCulprit}. Then, we refine these estimates. While the orbit serves as a good first-order proxy for the stream, it is insufficient, as the typical offset between the stream and its orbit can be on the order of 200-400 pc—significantly larger than the impact parameters we are investigating.

    In this refinement process, we aim to pinpoint both the exact location of the impact along the stream and the precise moment it occurred. To do so, we fit a third-order parametric polynomial to the stream, using the saved snapshots from our simulations.
    \begin{align}
      x(\tau) &= a_0 + a_1 \tau + a_2 \tau^2 + a_3 \tau^3, \\ 
      y(\tau) &= b_0 + b_1 \tau + b_2 \tau^2 + b_3 \tau^3, \\
      z(\tau) &= c_0 + c_1 \tau + c_2 \tau^2 + c_3 \tau^3,
      \end{align}
    where $x$, $y$, and $z$ represent the parametric line describing the stream in galactocentric coordinates, $\tau$ is the stream coordinate in time as described in the previous section, and is used as the independent variable in this analysis. The coefficients $a_i$, $b_i$, and $c_i$ are the polynomial coefficients. We found that a second-order polynomial was insufficient to capture the curvature along the full length of the stream, with divergence from the star particle positions occurring at the extremes. A third-order polynomial was sufficient and desirable, as it is the lowest order that adequately captures the path of the stream over the entire length under consideration.

    In this analysis, only one side of the stream is considered. For instance, if the impact candidate was in the leading tail, only the star particles with $\tau > 0$ are used to constrain the stream. The polynomial coefficients were determined using the Nelder-Mead algorithm from \texttt{scipy}'s optimization package.

    Since the simulation snapshots were saved at a temporal resolution of 1 Myr—rather than at the integration timestep of $10^4$ years, which would generate an excessive amount of data—we interpolate between snapshots to more precisely estimate the impact geometry. This is achieved by fitting the stream at five timesteps surrounding the approximate impact time, a period of 5 Myr, which sufficiently covers the interaction time. The interaction time can be estimated as $t \approx \frac{100~\text{pc}}{300~\frac{\text{km}}{\text{s}}} \approx 0.3~\text{Myr}$.

    As a result, each polynomial coefficient can now be expressed as a function of time. Consequently, we can parameterize the stream as a function of both simulation time and position along the stream:

    \begin{align}
      x(t,\tau) &= a_0(t) + a_1(t)\tau + a_2(t) \tau^2 + a_3(t)\tau^3, \\ 
      y(t,\tau) &= b_0(t) + b_1(t)\tau + b_2(t) \tau^2 + b_3(t)\tau^3, \\
      z(t,\tau) &= c_0(t) + c_1(t)\tau + c_2(t) \tau^2 + c_3(t)\tau^3.
      \end{align}
    The values of the coefficients as a function of time are obtained through linear interpolation, ensuring that the coefficients at the snapshot times match the values constrained by the simulation data.

    Next, we fit the trajectory of the perturber with a second-order polynomial. With equations for both the stream and the perturber as functions of time, we identify the time and location of impact by minimizing a cost function, defined as the distance between the stream and the perturber:

    \begin{equation} b(t, \tau) = \left\lVert \vec{s}(t, \tau) - \vec{p}(t) \right\rVert, \end{equation}

    where $\vec{s}(t, \tau)$ is the galactocentric position of a point on the stream, $\vec{p}(t)$ is the position of the perturber, and $b$ is the distance between the two. The minimum value of $b$, denoted as $\text{min}(b)$, represents the impact parameter. The minimization is carried out using \texttt{scipy}'s optimization package with the \textit{L-BFGS-B} method, which allows us to constrain $t$ and $\tau$, ensuring no extrapolation occurs.

    Once this minimization is performed, determining the relative velocity becomes straightforward. Since the minimization provides the impact parameter, time of impact, and corresponding value of $\tau$, we can compute the derivatives of the parametric equations at $t_{\text{min}}$ and $\tau_{\text{min}}$. The parallel and perpendicular components of the perturber's velocity relative to the stream are given by:

    \begin{align}
      \Delta \vec{v} &=\vec{v_p} - \vec{v_s} \\
      w_\parallel &= \left(\Delta \vec{v}\right)\cdot \hat{v_s}\\  
      w_\perp &=  \sqrt{\Delta v ^2 - w_\parallel ^ 2}
      \end{align}

    where $\vec{v}_p$ and $\vec{v}_s$ are the velocities of the perturber and the stream, respectively. For each of the 50 simulations, this information was compute for the strongs 5 flybys of a perturber with the stream. Thus, we compute a sample of 250 impacts, and flag those that give way to gaps.

\section{Results}

  I'm gonna go remake the figures before I adapt the text. However, before that, I will say the general outline of the flow of the article. I will show a general result here, which is what we would expect to see on the sky, which is comparable with reality. It's a great first figure becuase it shows the essence of the paper. Then, I pretty much follow the methodology and keep the story the same with just monte-carlo-009. The story is we need to understand how many gaps are present in a given simulation, then we need to identidy who the perturbers are. So step 1 (fig 2) shows our aided eye-test. Fig~3 shows how I tried to find each perturber, and Fig~4 shows a sanity check to reassure me that the suspects are indeed guilty. However, I will say that we picked monte-carlo-009 because it has two obvious gaps and two subtle ones, which aid in communicating our point. A full display of the gaps in our simulations is shown in the appendix (i'll make pair-plots with vanilla and full simulations above and below one another with the gaps labeled. :D )

  After we discuss the magic to make detections, I THEN will discuss some general properties, i.e. is there something special about these perturbers compared to other clusters? The main results is that \textit{you need to enter Pal5's orbital space}, which is not a shocker. If you can hit me, you can't perturb me. NEXT, I want to compare the population of perturbers to all globular clusters based on their internal parameters, so here we do show a tendency. Next, I wanna show how often gaps appear, and thus how many time a given GC pertubered Pal5, and when the gaps occur. I think a conclusion with the histograms in time is

  Lastly, I will try to show something about the impact geometry. I will state that we want to compare the ones that create gaps to those that did not. We wanted to apply what we learned from Carlberg, Erkal and Sanders. The moral of the story will be that, no combination of parameters explaining the geometry cleared demarcated the presence of a gap or not. This can be inline with Sander's expanion of the analysis initiated by Carlberg and extended by Erkal. Namly, a uniformly dense stream on a circular orbit is just not good enough. I think Sanders also said that the streams needs to be sufficiently diffuse, which could make sense given the epicycles.

  \textbf{\textit{Action item:}} I need to make sure I check out real good the papers that investigate the gaps from Palomar 5. I wanna compare this to the known length of Palomar 5, how does this reduce the results, and are there observed gaps in Palomar 5 that are consistent with these simulations?

  \begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{figures/galactic-coords-monte-carlo-009.png}
    \caption[]{A model of the Palomar 5 stream in an axis-symmetric galactic potential plus all other globular clusters. The top plot is a scatter plot of star-particles that escaped the cluster due to tidal forces, the bottom plot shows the marginalized 1D density profile in longitude. Two large gaps are present and are due to the passage of two globular clusters. }
    \end{figure*}

  \begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{figures/monte-carlo-009-pouliasis2017pii-GCNBody-2000-milisigma-50-noisefactor-20-boxcarindexlength-shifted-0.png}
    \caption[]{A comparison between the vanilla and full simulations. This showcases our gap detection method, in which the 2D maps in \textit{tail coordinates} are inspected, and their 1D marginalized density distributions are compared quantitatively. The $x'$ coordinate indicates distance from the center of mass of Pal5 along it's orbit. The two profiles are subtraced in locations where the signal is above the noise threshold, and locations that are more that are underdense by 2-$\sigma$ are highlighted with vertical blue lines, and labeled \textit{gap candidate}. This also showcases the limitation in this 1D profile method since a very skinny and oblique gap is present, whose signal is washed out from marginalizing.}
    \label{fig:profiles}
    \end{figure*}

  \begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{figures/Pal5_pouliasis2017pii-GCNBody_monte-carlo-009_suspects.png}
    \caption[]{This figure demonstrates how we determined which globular clusters were responsible for the gaps. The bottom plot showcases the evolution of the stream density in simulation time. The y-axis is $\tau$, which is a coordinate in units of time that indicates how far ahead or behind it is from a globular cluster. The x-axis is the simulation time, $t$, where 0~s km~kpc$^{-1}$ indicates present time. The density was used to determine a sutible length of the stream as a function of time. This length was then used to extract a piece of Palomar 5's orbit about a given position of Pal 5 in time. This orbital segment is used to approximate the stream. Then, the gravitational force from all other clusers was computed on the orbit. Moments of high acceleration indicate the passage of another cluster. The top 5 strongest passages are labeled with red circles as well as the name of the clusters. }
    \label{fig:profiles}
    \end{figure*}

  \begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{figures/monte-carlo-009-decomposition.png}
    \caption[]{This figure decomposes the gaps found in the full simulation to the responsible perturbers. The pertuber candidates were found in Fig.~\ref{fig:profiles}, and extra simulations were run in which we singularly add the suspects to verify that it was indeed them who created the gaps. The titles explain which simulation is which, and there's also a descrepandy map. (I think I'm going to get rid of the DIFF map)}
    \label{fig:profiles}
    \end{figure*}

  \begin{figure}
    \centering
    \includegraphics[width=1\linewidth]{figures/E_Lz_perturbers.png}
    \includegraphics[width=1\linewidth]{figures/mass_size_plane.png}
    \includegraphics[width=\linewidth]{figures/histogram_impact_time.png}
    \caption{One plot shows the distribution in energy and angular momentum, the next are the inernal proeprties of the clusters, and the last show when the clusters commit their crime, and how many times it happens accross the 50 simulations. }
    \label{fig:mass_size_plane}
    \end{figure}

  Here, I am typing words

\bibliographystyle{aa} % or another style like plain, unsrt, etc.
\bibliography{bibliography} % replace 'references' with the name of your .bib file

\begin{appendix}

  \section{Tail Coordinates} \label{appendix:TailCoordinates}

  \begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/along_orbit_coordinate_system-approved.png}
    \caption{Tail coordinates}
    \label{fig:TailCoordinates}
  \end{figure}

\end{appendix}

\end{document}

### 2024-09-10 figures/decomposition.py
adding old code to make the figures for publication

@@ -1,0 +1,276 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import h5py
+import plotting_functions as PF
+import sys 
+sys.path.append("/obs/sferrone/gc-tidal-loss/code/")
+import StreamOrbitCoords as SOC
+import matplotlib.gridspec as gridspec
+from matplotlib.colors import LogNorm
+
+
+'''
+DECOMPOSITION OF MONTE CARLO 009
+
+'''
+
+############################################################
+############################################################
+##################### GLOBAL VARIABLES #####################
+############################################################
+############################################################
+
+#### PATH VARIABLES
+streampathbase="/scratch2/sferrone/simulations/Streams/"
+orbitpathbase="/scratch2/sferrone/simulations/Orbits/"
+
+##### integration variables 
+NP=100000
+montecarlokey="monte-carlo-009"
+internaldynamics = "isotropic-plummer"
+targetGC="Pal5"
+### the names of the Pal5 orbits that were used for each stream computation
+correspondingOrbits = {
+    "pouliasis2017pii": "pouliasis2017pii",
+    "pouliasis2017pii-GCNBody": "pouliasis2017pii-GCNBody/",
+    "pouliasis2017pii-NGC2808": "pouliasis2017pii-GCNBody/",
+    "pouliasis2017pii-NGC104": "pouliasis2017pii-GCNBody/",
+    "pouliasis2017pii-NGC7078": "pouliasis2017pii-GCNBody/",
+}
+
+###### PARAMS FOR PLOT COMPUTATION ######
+xmax,ymax=10,1
+currenttime=0
+potentials = ["pouliasis2017pii",
+              "pouliasis2017pii-GCNBody",
+              "DIFF",
+              "pouliasis2017pii-NGC2808",
+              "pouliasis2017pii-NGC104",
+              "pouliasis2017pii-NGC7078"]
+Nrows=len(potentials)
+
+
+
+###### PARAMS FOR PLOT ASTHETICS
+xlim,ylim=10,0.5
+diffScatterProperties = {
+    "cmap": "coolwarm_r",
+    "s": 10,
[…]

### 2024-09-10 figures/gallery_of_gaps.py

@@ -1,0 +1,100 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from matplotlib.colors import LogNorm
+import h5py
+import os
+# routines in this folder
+import plotting_functions as PF
+import gap_results as GR
+import sys 
+sys.path.append("/obs/sferrone/gc-tidal-loss/code/")
+import StreamOrbitCoords as SOC
+
+
+def main():
+    # get the parameters
+    NP,internaldynamics,potential,targetGC,xmax,ymax = default_parameters()
+    # set the paths
+    basepath="/scratch2/sferrone/simulations/Streams/"+potential+"/"+targetGC+"/"+str(NP)+"/"
+    orbitpath="/scratch2/sferrone/simulations/Orbits/"+potential+"/"+targetGC+"-orbits.hdf5"
+    # make sure the out directory exists
+    outdir="./gallery_of_gaps/"
+    os.makedirs(outdir,exist_ok=True)
+    # get the perturber dictionary
+    MC_gaps=GR.GetPerturberDictPouliasis2017pii()
+    MCkeys=list(MC_gaps.keys())
+    MCkeys = ["monte-carlo-"+str(i).zfill(3) for i in range(0,50)]
+    # current time is always 0, b/c zero means today
+    currenttime= 0 # started at -5*u.Gyr
+    n_dynamic_times=2 # filtering the orbit by dynamical time
+    # iterate over the monte-carlo-keys
+    
+    scat_params={'s':1,'alpha':0.9,'cmap':'rainbow','norm':LogNorm(vmin=1e-5, vmax=5e-4)}
+    for montecarlokey in MCkeys:
+        # open the stream 
+        streamfilename = basepath+targetGC+"-stream-"+montecarlokey+".hdf5"
+        myStream = h5py.File(streamfilename, 'r')
+        # take a piece of the orbit in order to do the coordinate transform
+        with h5py.File(orbitpath,'r') as target:
+            tGC =target[montecarlokey]['t'][:]
+            xGC =target[montecarlokey]['xt'][:]
+            yGC =target[montecarlokey]['yt'][:]
+            zGC =target[montecarlokey]['zt'][:]
+            vxGC=target[montecarlokey]['vxt'][:]
+            vyGC=target[montecarlokey]['vyt'][:]
+            vzGC=target[montecarlokey]['vzt'][:]
+            
[…]

### 2024-09-10 figures/plotting_functions.py

@@ -1,0 +1,72 @@
+import numpy as np 
+import matplotlib.pyplot as plt
+from matplotlib.colors import LogNorm
+import matplotlib.gridspec as gridspec
+
+def histogramEdges(NP,xmax,ymax):
+    '''
+    given the number of particles, return the number of bins
+    '''
+    nBins=int(np.ceil(np.sqrt(NP)))
+    nYbins=int(np.ceil(nBins/np.sqrt(xmax/ymax)))
+    nXbins=int(np.ceil(nBins*(xmax/ymax)))
+    xedges=np.linspace(-xmax,xmax,nXbins)
+    yedges=np.linspace(-ymax,ymax,nYbins)
+    return xedges,yedges  
+
+
+def getHistogram2d(x,y,xedges,yedges):
+    '''
+    given the x and y coordinates, return the histogram
+    '''
+    H,_,_=np.histogram2d(x,y,bins=(xedges,yedges))
+    H=H.T
+    cond=H>0
+    xcenter,ycenter=0.5*(xedges[1:]+xedges[:-1]),0.5*(yedges[1:]+yedges[:-1])
+    X,Y=np.meshgrid(xcenter,ycenter)
+    X,Y=X[cond],Y[cond]
+    H=H[cond]
+    zsort=np.argsort(H)
+    X,Y,H=X[zsort],Y[zsort],H[zsort]
+    return X,Y,H
+
+
+def getHistogram2dDIFF(x1,y1,x2,y2,xedges,yedges):
+    '''
+    given the x and y coordinates, return the histogram
+    '''
+
+    xcenter,ycenter=0.5*(xedges[1:]+xedges[:-1]),0.5*(yedges[1:]+yedges[:-1])
+    X,Y=np.meshgrid(xcenter,ycenter)
+
+    H1,_,_=np.histogram2d(x1,y1,bins=(xedges,yedges))
+    H1=H1.T
+
+    H2,_,_=np.histogram2d(x2,y2,bins=(xedges,yedges))
+    H2=H2.T
+
+    DIFF = H1 - H2
+
+    cond=np.abs(DIFF)>0
+
+    X,Y,DIFF=X[cond],Y[cond],DIFF[cond]
+        
+    return X,Y,DIFF
+
+
+def plot2dHist(X,Y,H,
+               xlim=10,ylim=0.5,
[…]

### 2024-09-10 paper/main.tex
the decomposition works
added on sky plot

@@
-\documentclass[draft]{aa}
+\documentclass[]{aa}
 
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/galactic-coords-monte-carlo-009.png}
+    \includegraphics[width=\linewidth]{stream_on_sky_Pal5_monte-carlo-009_pouliasis2017pii-GCNBody.png}
     \caption[]{A model of the Palomar 5 stream in an axis-symmetric galactic potential plus all other globular clusters.
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/monte-carlo-009-decomposition.png}
+    \includegraphics[width=\linewidth]{decomposition-monte-carlo-009-with-3-gaps.png}
     \caption[]{This figure decomposes the gaps found in the full simulation to the responsible perturbers.

### 2024-09-10 paper/main.tex
added profile detection

@@
     \centering
-    \includegraphics[width=\linewidth]{figures/monte-carlo-009-pouliasis2017pii-GCNBody-2000-milisigma-50-noisefactor-20-boxcarindexlength-shifted-0.png}
+    \includegraphics[width=\linewidth, trim=20 0 15 0]{monte-carlo-009-pouliasis2017pii-GCNBody-2000-milisigma-50-noisefactor-20-boxcarindexlength-shifted-0.png}
     \caption[]{A comparison between the vanilla and full simulations.

### 2024-09-10 figures/profile_detections.py

@@ -39,6 +39,7 @@ def gspec_double_xy_profile():
     cbar_ax = fig.add_subplot(gs[0:2,1])
     return fig,ax0,ax1,ax2,cbar_ax
 
+
 def doplot(handles,data,properties):
     # unpack each 
     data_append2Dmaps                   = data['append2Dmaps']

### 2024-09-11 stream_analysis/identify_suspects.py
added command to move text for labeing perturbation

@@ -58,10 +58,10 @@ def main(dataparams,hyperparams):
 
 
     fig,axis1,axis2,caxisFOOB,caxisTau = sa.identify_suspects.getfigure()
-    foobNORM,tauNORM,AXIS1,AXIS2 = sa.identify_suspects.properties(time_foob,tau_left,tau_right,NP,montecarlokey)
+    foobNORM,tauNORM,AXIS1,AXIS2,TEXT = sa.identify_suspects.properties(time_foob,tau_left,tau_right,NP,montecarlokey)
     FIGSTUFF=[fig,axis1,axis2,caxisFOOB,caxisTau]
     DATASTUFF=time_foob,tau_foob,convolved_mask,suspects,coordinates,tau_counts,tau_centers,time_stamps,tau_left,tau_right
-    PROPERTIESSTUFF=foobNORM,tauNORM,AXIS1,AXIS2
+    PROPERTIESSTUFF=foobNORM,tauNORM,AXIS1,AXIS2,TEXT
     sa.identify_suspects.doplot(FIGSTUFF,DATASTUFF,PROPERTIESSTUFF)
     fig.savefig(fname,dpi=300) 
     print("Saved to ",fname)
@@ -98,11 +98,13 @@ def properties(time_foob,tau_left,tau_right,NP,montecarlokey):
     ylimits     =   np.max([np.abs(tau_left),np.abs(tau_right)])
     AXIS1       =   {"xlim":[time_foob[0],0],"ylim":[-ylimits,ylimits],"ylabel":"$\\tau$ [s kpc / km]","xlabel":"","title":montecarlokey}
     AXIS2       =   {"xlim":[time_foob[0],0],"ylim":[-ylimits,ylimits],"ylabel":"$\\tau$ [s kpc / km]","xlabel":"$t$ [s kpc / km]",}
-    return foobpcolor,taupcolor,AXIS1,AXIS2
+    TEXT        =   {"color":'red',"ha":'left',"va":'top'}
+    
+    return foobpcolor,taupcolor,AXIS1,AXIS2,TEXT
 
-def doplot(FIGSTUFF,DATASTUFF,PROPERTIESSTUFF):
+def doplot(FIGSTUFF,DATASTUFF,PROPERTIESSTUFF,x_text_shift=-0.1,y_text_shift=0.005):
     fig,axis1,axis2,caxisFOOB,caxisTau=FIGSTUFF
-    foobpcolor,taupcolor,AXIS1,AXIS2=PROPERTIESSTUFF
[…]

### 2024-09-12 paper/main.tex
added force on orbit figure
I now have Lz E
added mass size plane
added impact time to paper

@@
     \centering
-    \includegraphics[width=\linewidth]{figures/Pal5_pouliasis2017pii-GCNBody_monte-carlo-009_suspects.png}
+    \includegraphics[width=\linewidth]{force_on_orbit-monte-carlo-009.png}
     \caption[]{This figure demonstrates how we determined which globular clusters were responsible for the gaps.
@@
     \centering
-    \includegraphics[width=1\linewidth]{figures/E_Lz_perturbers.png}
-    \includegraphics[width=1\linewidth]{figures/mass_size_plane.png}
-    \includegraphics[width=\linewidth]{figures/histogram_impact_time.png}
+    \includegraphics[width=1\linewidth]{E_Lz_perturbers.png}
+    \includegraphics[width=1\linewidth]{mass_size_plane.png}
     \caption{One plot shows the distribution in energy and angular momentum, the next are the inernal proeprties of the clusters, and the last show when the clusters commit their crime, and how many times it happens accross the 50 simulations.
@@
   \end{figure}
+
+  \begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{histogram_impact_time.png}
+    \caption{When the impacts occured}
+    \label{fig:histogram_impact_time}
+  \end{figure*}
 

### 2024-09-16 paper/main.tex
added 1
added 2
3
4
5
updated plots

@@
     \end{figure}
-
   \begin{figure*}
@@
 
-  Here, I am typing words
+  \section{Discussion}
+    $10^5-10^6$. Sotto alone averbbero questo effetto?
 
@@
 
+  \section{Gallery of Gaps}
+
+    \begin{figure*}
+      \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-000.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-001.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-002.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-003.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-004.png}
+      \caption{Gap Gallery}
+      \label{fig:TailCoordinates}
+    \end{figure*}
+
+    \begin{figure*}
+      \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-005.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-006.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-007.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-008.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-009.png}
+      \caption{Gap Gallery}
+      \label{fig:TailCoordinates}
+    \end{figure*}
+
+    \begin{figure*}
+      \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-010.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-011.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-012.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-013.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-014.png}
+      \caption{Gap Gallery}
+      \label{fig:TailCoordinates}
+    \end{figure*}
+
+    \begin{figure*}
+      \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-015.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-016.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-017.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-018.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-019.png}
+      \caption{Gap Gallery}
+      \label{fig:TailCoordinates}
+    \end{figure*}
+
 \end{appendix}

### 2024-09-16 paper/main.tex
added the plots to the text

@@
 
+    \begin{figure*}
+      \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-020.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-021.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-022.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-023.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-024.png}
+      \caption{Gap Gallery}
+      \label{fig:TailCoordinates}
+    \end{figure*}
+
+    \begin{figure*}
+      \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-025.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-026.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-027.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-028.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-029.png}
+      \caption{Gap Gallery}
+      \label{fig:TailCoordinates}
+    \end{figure*}
+
+    \begin{figure*}
+      \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-030.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-031.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-032.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-033.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-034.png}
+      \caption{Gap Gallery}
+      \label{fig:TailCoordinates}
+    \end{figure*}
+
+    \begin{figure*}
+      \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-035.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-036.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-037.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-038.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-039.png}
+      \caption{Gap Gallery}
+      \label{fig:TailCoordinates}
+    \end{figure*}
+
+    \begin{figure*}
+      \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-040.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-041.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-042.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-043.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-044.png}
+      \caption{Gap Gallery}
+      \label{fig:TailCoordinates}
+    \end{figure*}
+
+    \begin{figure*}
+      \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-045.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-046.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-047.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-048.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-049.png}
+      \caption{Gap Gallery}
+      \label{fig:TailCoordinates}
+    \end{figure*}
+
 \end{appendix}

### 2024-09-16 paper/main.tex
trying to solve problems with noise

@@
-\documentclass[]{aa}
+\documentclass[draft]{aa}
 
@@
     Briefly, the x'coordinate represents the position of a particle along the orbit relative to the globular cluster—whether it is ahead or behind the cluster.
-The y'coordinate measures the particle's distance within the orbital plane, where positive values indicate that the particle is farther from the Galactic center and negative values indicate that it is closer.
+The y' coordinate measures the particle's distance within the orbital plane, where positive values indicate that the particle is farther from the Galactic center and negative values indicate that it is closer.
 
@@
       For each of the 50 simulations, we compare the final snapshots between the \texttt{vanilla} and full simulations in the tail coordinate system.
-This is done by creating 2D density maps of the streams, followed by marginalizing over the y'-coordinate to produce 1D density profiles along the x'-axis.
-These comparisons are illustrated in the results section (see Fig.~XXX).
+We inspect these differences by creating 2D density maps of the streams.
+Subsequently, we wish to employ a quantitative method to make detections discover when the \textit{full} simulations are significantly less dense than the \textit{vanilla} simulations.
+To do so, we marginalize over the y'-coordinate to produce 1D density profiles along the x'-axis.
+These comparisons are illustrated in the results section (see Fig.~\ref{fig:profiles}).
 
-      To construct the 1D density profiles, we first bin the data using the  $\sqrt{N}$-rule for determining the number of bins.
+      To construct the 1D density profiles, we first bin the data using the $\sqrt{N}$-rule for determining the number of bins, where N is the number of data points which is 10$^5$.
 After histogramming the 1D profiles, we apply a boxcar smoothing technique, where each central value is replaced by the median of the surrounding XX adjacent data points, corresponding to a smoothing length of approximately 300 parsecs.
-This procedure reduces high-frequency noise and smooths the profiles, as shown in Fig.~XX, where the center of mass has been effectively removed.
+This procedure reduces high-frequency noise and smooths the profiles, as shown in Fig.~\ref{fig:profiles}, where the center of mass has been effectively removed.
 
-      With the smoothed 1D profiles in hand, we proceed to compare them by identifying regions where the full simulations (GCGC) show significant underdensity compared to the \texttt{vanilla} simulations.
-To achieve this, we compute the difference between the logarithms of the binned counts for each profile and search for points that fall more than two standard deviations below the mean.
+      With the smoothed 1D density profiles in hand, we consider the \textit{vanilla} simulations to be truth, and search for locations that are significantly less dense than \textit{vanilla} in a way that is greater than stochastic chance.
+To do so, we first compute the log ratio of the two profiles:
+      \begin{equation}
+        \mathcal{R} = \log\left(\frac{\rho_f}{\rho_v}\right),
+        \end{equation}
+      where $\rho$ is the 1D density, $v$ corresponding to the vanilla simulation, and $f$ corresponding to the full simulation, and $R$ is the log ratio of the two.
+Using the log ratio is important, because the distribution is symmetric and zero centered.
+For example if we only used the ratio, then the codomain would be $\frac{\rho_f}{\rho_v} \in [1,+\inf]$ for $\rho_f > \rho_v$, and $\frac{\rho_f}{\rho_v} \in [0,1]$ for $\rho_f < \rho_v$.
+
+      Yet, we only wish to consider bins with a significant amount of counts.
+We find this threshold by asserting that the signal to noise ratio must be greater than 50.
+The signal is based on the \textit{vanilla} profile, and the noise is the error propogation given poisson uncertainties for the number of counts a given bin:
+      \begin{equation}
+        \mathcal{S}/\mathcal{N} =50= \frac{\log{N}}{\frac{1}{\ln(10)} \sqrt{N}}
+        \end{equation}
+
+      we compute the difference between the logarithms of the binned counts for each profile and search for points that fall more than two standard deviations below the mean.
 However, we only consider regions that surpass a predefined significance threshold, where the significance is determined by the uncertainties in the binned counts governed by Poisson statistics multiplied by a factor of 50.

### 2024-09-16 build/lib/stream_analysis/forceOnOrbit.py

@@ -1,0 +1,277 @@
+"""
+This module was created to help answer the question, which globular cluster is responsible for the perturbation of the stream? 
+    i.e. who caused the gap? 
+This module contains functions for calculating the acceleration on an orbit due to perturbing masses.
+We are using the orbit as a proxy for the force on the stream.
+The philosophy is to downsample so that the output data is smaller. 
+
+
+IN A NUT SHELL:
+
+1. Given a host orbit, estimate the size of the stream by some factor of the dynamical time
+2. Use the points from the orbit ahead and behind the mass to approximate the stellar stream
+3. Down sample the time and tau indices to reduce the data size
+4. Compute the acceleration on the stream due to each of the perturbing masses.
+5. Return 
+    a. The components of the total acceleration from all perturbers
+    b. The magnitude of the acceleration for each perturber at each time step
+    note. This way we can
+        i. see which perturber is responsible for the gap at a given time
+        ii. have the acceleration vectors for the animations I want to make to verify the computation
+
+6. Write the file!
+
+"""
+
+import numpy as np 
+# import tstrippy
+import h5py
+import datetime
+
+### This is the main function that calculates the acceleration on an orbit due to perturbing masses
+def compute_force_from_all_gcs(G,Masses,orbit_perturber,orbit_host,indexing,outputs):
+    """
+    Calculate the acceleration on an orbit due to perturbing masses.
+
+    Parameters
+    ----------
+    G : float
+        Gravitational constant.
+    Masses : array_like
+        Array of masses of the perturbing objects.
+    orbit_perturber : tuple of array_like
+        Tuple containing the x, y, and z coordinates of the perturbing objects' orbits.
+    orbit_host : tuple of array_like
+        Tuple containing the x, y, and z coordinates of the host object's orbit.
+    indexing : tuple of array_like
+        Tuple containing the time indexing and tau indexing arrays.
+    outputs : tuple of array_like
+        Tuple containing the arrays for the x, y, and z components of acceleration and the magnitude of acceleration.
+
+    Returns
+    -------
+    ax : array_like
+        Array of x components of the net acceleration.
[…]

### 2024-09-16 build/lib/stream_analysis/identify_suspects.py

@@ -1,0 +1,204 @@
+"""
+Use the results from:
+    FORCE ON ORBIT
+    STREAM LENGTH
+
+in order to identify the moments the are most likely responsible for the gaps.
+
+Save the top 5 moments in some kind of file
+
+"""
+
+
+# imports
+import stream_analysis as sa
+from gcs import path_handler as ph 
+import h5py
+import numpy as np 
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+import pandas as pd
+from scipy.ndimage import maximum_filter, label, find_objects
+from scipy import ndimage
+
+
+
+
+def main(dataparams,hyperparams):
+    internal_dynamics,montecarlokey,NP,MWpotential,GCname = dataparams
+    threshold,NKEEPS,fname = hyperparams
+
+    fileFOOB = ph.ForceOnOrbit(GCname,MWpotential,montecarlokey)
+    fileTau  = ph.tauDensityMaps(GCname=GCname,MWpotential=MWpotential,montecarlokey=montecarlokey,NP=NP,internal_dynamics=internal_dynamics)
+
+
+    time_foob,tau_foob,mag_total,Perturbers,GCmags = sa.identify_suspects.extract_FOOB(fileFOOB)
+    tau_centers,time_stamps,tau_counts = sa.identify_suspects.extract_tau_stream_density(fileTau)
+
+    # get the envlopes
+    leftindexes,rightindexes=sa.streamLength.get_envelop_indexes(tau_counts,threshold)
+    tau_left,tau_right=sa.streamLength.tau_envelopes(tau_centers,leftindexes,rightindexes)
+
+
+    mask=sa.identify_suspects.build_mask_on_FOOB_from_density((time_stamps,tau_left,tau_right),(time_foob,tau_foob,mag_total))
+    masked_data = mag_total[mask]
+
+    kernel = np.ones(4) / 4  # 5-point moving average
+    convolved_data = ndimage.convolve1d(mag_total, kernel, axis=0, mode='reflect')
+    convolved_mask = np.ma.masked_where(~mask,convolved_data)
+
+
+    coordinates = sa.identify_suspects.get_peaks(convolved_mask)
+
[…]

### 2024-09-16 build/lib/stream_analysis/plotters/binned_density.py

@@ -6,15 +6,20 @@ import numpy as np
 
 
 
-def get_edges(NP,xmax,ymax):
+def get_edges(NP,xlims,ylims):
     '''
     given the number of particles, return the number of bins
     '''
+    ylen = ylims[1]-ylims[0]
+    xlen = xlims[1]-xlims[0]
+    print('xlen',xlen)
+    print('ylen',ylen)
+    
     nBins=int(np.ceil(np.sqrt(NP)))
-    nYbins=int(np.ceil(nBins/np.sqrt(xmax/ymax)))
-    nXbins=int(np.ceil(nBins*(xmax/ymax)))
-    xedges=np.linspace(-xmax,xmax,nXbins)
-    yedges=np.linspace(-ymax,ymax,nYbins)
+    nYbins=int(np.ceil(nBins*np.sqrt(ylen/xlen)))
+    nXbins=int(np.ceil(nBins)*np.sqrt(xlen/ylen))
+    xedges=np.linspace(xlims[0],xlims[1],nXbins)
+    yedges=np.linspace(ylims[0],ylims[1],nYbins)
     return xedges,yedges
 
 def get_2D_density(X,Y,xedges,yedges):

### 2024-09-16 build/lib/stream_analysis/plotters/tailcoordinates.py

@@ -1,0 +1,66 @@
+"""A module for preparing plots of the tail coordinates of a stream.
+
+"""
+
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+
+
+#########################################################################
+############################## PROPERTIES ###############################
+#########################################################################
+def xy_AXIS_properties(xlims=[-10,10],ylims=[-1,1]):
+    return {"xlim":xlims,"ylim":ylims,"xlabel":" x' [kpc]","ylabel":" y' [kpc]"}
+
+def profile_AXIS_properties(xlims=[-10,10]):
+    return  {"yscale":"log","xlabel":" x' [kpc]","ylabel":" N/NP","xlim":xlims}
+
+def density_SCAT_properties(NP,s=3,upfactor=1e2,cmap=mpl.colormaps.get_cmap("rainbow")):
+    cnorm=mpl.colors.LogNorm(vmin=1/NP, vmax=upfactor/NP)
+    return {'s':s,'cmap':cmap,'norm':cnorm}
+
+
+
+############################################################################
+############################## FIGURE LAYOUTS ##############################
+############################################################################
+def gspec_double_xy_profile():
+    fig=plt.figure(figsize=(15,6))
+    gs = fig.add_gridspec(3, 2, height_ratios=[1,1,4],width_ratios=[1, 1/100],wspace=0.05)
+    ax0=fig.add_subplot(gs[0,0])
+    ax1=fig.add_subplot(gs[1,0])
+    ax2=fig.add_subplot(gs[2,0])
+    cbar_ax = fig.add_subplot(gs[0:2,1])
+    return fig,ax0,ax1,ax2,cbar_ax
+
+def gspec_single_xy():
+    fig=plt.figure(figsize=(15,1))
+    gs=fig.add_gridspec(1,2,width_ratios=[1,0.01],wspace=0.02)
+    ax=fig.add_subplot(gs[0,0])
+    cbar_ax=fig.add_subplot(gs[0,1])
+    return fig,ax,cbar_ax
+
+def gspec_xy_and_profile():
+    fig=plt.figure(figsize=(15,4))
+    gs=fig.add_gridspec(2,2,height_ratios=[1,4],width_ratios=[1,0.01],wspace=0.05)
+    axis0 = fig.add_subplot(gs[0,0])
[…]

### 2024-09-16 build/lib/stream_analysis/tailCoordinates.py

@@ -14,6 +14,15 @@ Typical usage:
 
 '''
 import numpy as np 
+import gcs
+
+def tau_gamma(indexes:np.ndarray,yT:np.ndarray,TORB:np.ndarray,XORB:np.ndarray,YORB:np.ndarray,ZORB:np.ndarray,\
+    VXORB:np.ndarray,VYORB:np.ndarray,VZORB:np.ndarray,currenttime:float):
+    lenOrb=TORB.shape[0]
+    _,_,_,rdot,_,_=gcs.misc.cartesian_to_spherical(XORB,YORB,ZORB,VXORB,VYORB,VZORB)
+    tau = TORB[indexes] - currenttime
+    gamma=yT/np.abs(rdot[lenOrb//2])    
+    return tau,gamma
 
 def transform_from_galactico_centric_to_tail_coordinates(xp,yp,zp,vxp,vyp,vzp,tORB,xORB,yORB,zORB,vxORB,vyORB,vzORB,t0=0):
     '''
@@ -174,7 +183,7 @@ def filter_orbit_by_dynamical_time(tORB: np.ndarray, xtORB: np.ndarray, ytORB: n
     todayindex = np.argmin(np.abs(tORB - time_of_interest))
     ttoday = tORB[todayindex]
 
-    Tdyn = rORB[todayindex] / vORB[todayindex]
+    Tdyn = np.median(rORB / vORB)
 
     cond1 = (tORB - ttoday) < (nDynTimes * Tdyn)
     cond2 = (tORB - ttoday) > (-nDynTimes * Tdyn)

### 2024-09-17 build/lib/stream_analysis/__init__.py
updated method to compute noise threshold

@@ -1,2 +1,7 @@
 from . import tailCoordinates
 from . import plotters
+from . import density_profile_gap_detections
+from . import forceOnOrbit
+from . import streamLength
+from . import identify_suspects
+from . import parametric_stream_fitting

### 2024-09-17 build/lib/stream_analysis/density_profile_gap_detections.py

@@ -6,15 +6,21 @@ A module for comparing the density profiles of the control group to the group wi
 """
 import stream_analysis as sa 
 import numpy as np 
-
+from scipy.optimize import fsolve
 ####### COMPUTATIONS
-def log_difference_detection(profiles,noise_factor,sigma):
+def log_difference_detection(profiles,signal_to_noise_ratio,sigma):
     """Return the noise filtered profiles, noise level and the detection candidates.
     """
     CENTERS,CONTROL,COMPARE         =   profiles
-    noiselevel                      =   sa.density_profile_gap_detections.noise_log_counts(CONTROL)
+    # noiselevel                      =   sa.density_profile_gap_detections.noise_log_counts(CONTROL)
+    
+    N_threshold     =   int(np.ceil(fsolve(sa.density_profile_gap_detections.counts_threshold_given_signal_to_noise_ratio,10,args=(signal_to_noise_ratio))))
+    print(f"Threshold for counts is {N_threshold}")
+    noiselevel      =    N_threshold*np.ones_like(CONTROL)
+    
+    # lets do a new noise threshold
     LOGDIFF                         =   np.log10(COMPARE/CONTROL)
-    noise_filt                      =   CONTROL > noise_factor*noiselevel
+    noise_filt                      =   CONTROL > noiselevel
     centers,control,compare,logdiff =   sa.density_profile_gap_detections.apply_filter(noise_filt,[CENTERS,CONTROL,COMPARE,LOGDIFF])
     noise_filtered_profiles         =   centers,control,compare
     ############################
@@ -84,6 +90,11 @@ def apply_filter(myfilter,mylist):
     assert(type(mylist) == list)
     assert(type(myfilter) == np.ndarray)
     return [array[myfilter] for array in mylist]
+
+def counts_threshold_given_signal_to_noise_ratio(N,SNR):
+    """ Given a signal as log10(N), and the error as poisson distribution, return the threshold for the counts.
+    i.e. if we have less than this counts, then our SNR is less than the desired SNR, and maybe we won't consider this data."""
+    return SNR - np.log10(N)*np.sqrt(N)*np.log(10)
 
 def noise_log_counts(counts):
     """Return the noise level of the log counts."""

### 2024-09-17 build/lib/stream_analysis/plotters/binned_density.py

@@ -11,10 +11,7 @@ def get_edges(NP,xlims,ylims):
     given the number of particles, return the number of bins
     '''
     ylen = ylims[1]-ylims[0]
-    xlen = xlims[1]-xlims[0]
-    print('xlen',xlen)
-    print('ylen',ylen)
-    
+    xlen = xlims[1]-xlims[0]    
     nBins=int(np.ceil(np.sqrt(NP)))
     nYbins=int(np.ceil(nBins*np.sqrt(ylen/xlen)))
     nXbins=int(np.ceil(nBins)*np.sqrt(xlen/ylen))

### 2024-09-17 build/lib/stream_analysis/__init__.py
Stop tracking files that are now in .gitignore

@@ -1,7 +1,0 @@
-from . import tailCoordinates
-from . import plotters
-from . import density_profile_gap_detections
-from . import forceOnOrbit
-from . import streamLength
-from . import identify_suspects
-from . import parametric_stream_fitting

### 2024-09-17 build/lib/stream_analysis/density_profile_gap_detections.py

@@ -1,142 +1,0 @@
-"""
-
-A module for comparing the density profiles of the control group to the group with gaps and seeing where the under densities are. Log differences only. 
-
-
-"""
-import stream_analysis as sa 
-import numpy as np 
-from scipy.optimize import fsolve
-####### COMPUTATIONS
-def log_difference_detection(profiles,signal_to_noise_ratio,sigma):
-    """Return the noise filtered profiles, noise level and the detection candidates.
-    """
-    CENTERS,CONTROL,COMPARE         =   profiles
-    # noiselevel                      =   sa.density_profile_gap_detections.noise_log_counts(CONTROL)
-    
-    N_threshold     =   int(np.ceil(fsolve(sa.density_profile_gap_detections.counts_threshold_given_signal_to_noise_ratio,10,args=(signal_to_noise_ratio))))
-    print(f"Threshold for counts is {N_threshold}")
-    noiselevel      =    N_threshold*np.ones_like(CONTROL)
-    
-    # lets do a new noise threshold
-    LOGDIFF                         =   np.log10(COMPARE/CONTROL)
-    noise_filt                      =   CONTROL > noiselevel
-    centers,control,compare,logdiff =   sa.density_profile_gap_detections.apply_filter(noise_filt,[CENTERS,CONTROL,COMPARE,LOGDIFF])
-    noise_filtered_profiles         =   centers,control,compare
-    ############################
-    # zscore filter
-    myz                             =   sa.density_profile_gap_detections.zscores(logdiff)
-    candidates                      =   myz < -sigma
-    return noise_filtered_profiles,noiselevel,candidates
-
-def apply_box_car_multiple_times(signal,box_length,N_apply):
-    outsignal = signal.copy()
-    i=0
-    while i < N_apply:
-        outsignal = sa.density_profile_gap_detections.median_box_car(outsignal,box_length)
-        i+=1
-    return outsignal
-
-
-def clean_up_stream_profiles(x,c0,c1,box_length,N_APPlY,do_cross_correlation):
-    """Return the stream density profiles after (1) projecting, (2) applying a boxcar filter (3) cross correlating for a better match up.
-
-    Parameters
-    ----------
-    box_length : int
-        length of box in indices
-    N_APPlY : int
-        number of times to apply the boxcar filter
-    do_cross_correlation : bool
-        consider cross correlation
-
[…]

### 2024-09-17 build/lib/stream_analysis/forceOnOrbit.py

@@ -1,277 +1,0 @@
-"""
-This module was created to help answer the question, which globular cluster is responsible for the perturbation of the stream? 
-    i.e. who caused the gap? 
-This module contains functions for calculating the acceleration on an orbit due to perturbing masses.
-We are using the orbit as a proxy for the force on the stream.
-The philosophy is to downsample so that the output data is smaller. 
-
-
-IN A NUT SHELL:
-
-1. Given a host orbit, estimate the size of the stream by some factor of the dynamical time
-2. Use the points from the orbit ahead and behind the mass to approximate the stellar stream
-3. Down sample the time and tau indices to reduce the data size
-4. Compute the acceleration on the stream due to each of the perturbing masses.
-5. Return 
-    a. The components of the total acceleration from all perturbers
-    b. The magnitude of the acceleration for each perturber at each time step
-    note. This way we can
-        i. see which perturber is responsible for the gap at a given time
-        ii. have the acceleration vectors for the animations I want to make to verify the computation
-
-6. Write the file!
-
-"""
-
-import numpy as np 
-# import tstrippy
-import h5py
-import datetime
-
-### This is the main function that calculates the acceleration on an orbit due to perturbing masses
-def compute_force_from_all_gcs(G,Masses,orbit_perturber,orbit_host,indexing,outputs):
-    """
-    Calculate the acceleration on an orbit due to perturbing masses.
-
-    Parameters
-    ----------
-    G : float
-        Gravitational constant.
-    Masses : array_like
-        Array of masses of the perturbing objects.
-    orbit_perturber : tuple of array_like
-        Tuple containing the x, y, and z coordinates of the perturbing objects' orbits.
-    orbit_host : tuple of array_like
-        Tuple containing the x, y, and z coordinates of the host object's orbit.
-    indexing : tuple of array_like
-        Tuple containing the time indexing and tau indexing arrays.
-    outputs : tuple of array_like
-        Tuple containing the arrays for the x, y, and z components of acceleration and the magnitude of acceleration.
-
-    Returns
-    -------
-    ax : array_like
-        Array of x components of the net acceleration.
[…]

### 2024-09-17 build/lib/stream_analysis/identify_suspects.py

@@ -1,204 +1,0 @@
-"""
-Use the results from:
-    FORCE ON ORBIT
-    STREAM LENGTH
-
-in order to identify the moments the are most likely responsible for the gaps.
-
-Save the top 5 moments in some kind of file
-
-"""
-
-
-# imports
-import stream_analysis as sa
-from gcs import path_handler as ph 
-import h5py
-import numpy as np 
-import matplotlib.pyplot as plt
-import matplotlib as mpl
-import pandas as pd
-from scipy.ndimage import maximum_filter, label, find_objects
-from scipy import ndimage
-
-
-
-
-def main(dataparams,hyperparams):
-    internal_dynamics,montecarlokey,NP,MWpotential,GCname = dataparams
-    threshold,NKEEPS,fname = hyperparams
-
-    fileFOOB = ph.ForceOnOrbit(GCname,MWpotential,montecarlokey)
-    fileTau  = ph.tauDensityMaps(GCname=GCname,MWpotential=MWpotential,montecarlokey=montecarlokey,NP=NP,internal_dynamics=internal_dynamics)
-
-
-    time_foob,tau_foob,mag_total,Perturbers,GCmags = sa.identify_suspects.extract_FOOB(fileFOOB)
-    tau_centers,time_stamps,tau_counts = sa.identify_suspects.extract_tau_stream_density(fileTau)
-
-    # get the envlopes
-    leftindexes,rightindexes=sa.streamLength.get_envelop_indexes(tau_counts,threshold)
-    tau_left,tau_right=sa.streamLength.tau_envelopes(tau_centers,leftindexes,rightindexes)
-
-
-    mask=sa.identify_suspects.build_mask_on_FOOB_from_density((time_stamps,tau_left,tau_right),(time_foob,tau_foob,mag_total))
-    masked_data = mag_total[mask]
-
-    kernel = np.ones(4) / 4  # 5-point moving average
-    convolved_data = ndimage.convolve1d(mag_total, kernel, axis=0, mode='reflect')
-    convolved_mask = np.ma.masked_where(~mask,convolved_data)
-
-
-    coordinates = sa.identify_suspects.get_peaks(convolved_mask)
-
[…]

### 2024-09-17 build/lib/stream_analysis/parametric_stream_fitting.py

@@ -1,210 +1,0 @@
-"""
-...
-Author: Salvatore Ferrone
-Date: April 2024 (first functioning full version)
-"""
-
-
-
-import numpy as np 
-
-
-
-
-#############################################
-######## Obtaining the full geometry ########
-#############################################
-def get_full_impact_geometry_from_parametrization(
-    t:float,
-    tau:float,
-    coefficient_time_fit_params: np.ndarray,
-    trajectory_coeffs: np.ndarray,
-    time_stamps: np.ndarray,):
-    
-
-    rp  =   perturber_position(t,trajectory_coeffs)
-    vp  =   perturber_velocity(t,trajectory_coeffs)
-    rs  =   moving_parametric_stream(t,tau,coefficient_time_fit_params,time_stamps)
-    vs  =   moving_parametric_stream_velocity(t,tau,coefficient_time_fit_params,time_stamps)
-    
-
-    impact_parameter = np.linalg.norm(rp-rs)
-    bvec    =   get_impact_vector(rs,vs,rp)
-    alpha   =   get_imact_alpha_from_b_vector(bvec)
-    # the magnitude of b_vec should be the impact impact_parameter
-    w_par,w_per=get_parallel_and_perpendicular_velocity_components(vp,vs)
-
-    return impact_parameter,w_par,w_per,alpha
-   
-   
-######################################################################
-########### Obtaining individual Erkal geometry Parameters ###########
-######################################################################
-def get_imact_alpha_from_b_vector(b_vec_tail_basis:np.ndarray):
-    """
-    Returns the impact parameter in the direction of the position vector.
-    The angle is in radians and is the angle up from the the x-axis to the z-axis.
-    """
-    tolerence=2e-2
-    alpha1=np.arccos(b_vec_tail_basis[0]/np.linalg.norm(b_vec_tail_basis))
-    alpha2=np.arcsin(b_vec_tail_basis[2]/np.linalg.norm(b_vec_tail_basis))
-    alpha3=np.arctan2(b_vec_tail_basis[2],b_vec_tail_basis[0])
-    to_degrees=180/np.pi
-    mean_alpha=np.mean([alpha1,alpha2])
-    discrepancy=np.abs(alpha1-alpha2) 
-    if (discrepancy/mean_alpha)>tolerence:
[…]

### 2024-09-17 build/lib/stream_analysis/plotters/__init__.py

@@ -1,2 +1,0 @@
-from . import tailcoordinates
-from . import binned_density

### 2024-09-17 build/lib/stream_analysis/plotters/binned_density.py

@@ -1,48 +1,0 @@
-import numpy as np
-
-#####################################################################
-############### Functions for making the density maps ###############
-#####################################################################
-
-
-
-def get_edges(NP,xlims,ylims):
-    '''
-    given the number of particles, return the number of bins
-    '''
-    ylen = ylims[1]-ylims[0]
-    xlen = xlims[1]-xlims[0]    
-    nBins=int(np.ceil(np.sqrt(NP)))
-    nYbins=int(np.ceil(nBins*np.sqrt(ylen/xlen)))
-    nXbins=int(np.ceil(nBins)*np.sqrt(xlen/ylen))
-    xedges=np.linspace(xlims[0],xlims[1],nXbins)
-    yedges=np.linspace(ylims[0],ylims[1],nYbins)
-    return xedges,yedges
-
-def get_2D_density(X,Y,xedges,yedges):
-    H, xedges, yedges = np.histogram2d(X, Y, bins=(xedges, yedges))
-    xcenters,ycenters = (xedges[:-1] + xedges[1:]) / 2, (yedges[:-1] + yedges[1:]) / 2
-    XX,YY = np.meshgrid(xcenters,ycenters)
-    H = H.T
-    return XX,YY,H
-
-def discard_empty_bins(XX,YY,H):
-    cond = H>0
-    H = H[cond]
-    XX = XX[cond]
-    YY = YY[cond]
-    return XX,YY,H
-
-def short_cut(NP,X,Y,xlims,ylims):
-    xedges,yedges = get_edges(NP,xlims,ylims)
-    XX,YY,H = get_2D_density(X,Y,xedges,yedges)
-    XX,YY,H = discard_empty_bins(XX,YY,H)
-    return XX,YY,H
-
-def normalize_density_by_particle_number(H,NP):
-    return H/NP
-
-def order_by_density(XX,YY,H):
-    order = np.argsort(H)
-    return XX[order],YY[order],H[order]
-

### 2024-09-17 build/lib/stream_analysis/plotters/tailcoordinates.py

@@ -1,66 +1,0 @@
-"""A module for preparing plots of the tail coordinates of a stream.
-
-"""
-
-import matplotlib as mpl
-import matplotlib.pyplot as plt
-
-
-#########################################################################
-############################## PROPERTIES ###############################
-#########################################################################
-def xy_AXIS_properties(xlims=[-10,10],ylims=[-1,1]):
-    return {"xlim":xlims,"ylim":ylims,"xlabel":" x' [kpc]","ylabel":" y' [kpc]"}
-
-def profile_AXIS_properties(xlims=[-10,10]):
-    return  {"yscale":"log","xlabel":" x' [kpc]","ylabel":" N/NP","xlim":xlims}
-
-def density_SCAT_properties(NP,s=3,upfactor=1e2,cmap=mpl.colormaps.get_cmap("rainbow")):
-    cnorm=mpl.colors.LogNorm(vmin=1/NP, vmax=upfactor/NP)
-    return {'s':s,'cmap':cmap,'norm':cnorm}
-
-
-
-############################################################################
-############################## FIGURE LAYOUTS ##############################
-############################################################################
-def gspec_double_xy_profile():
-    fig=plt.figure(figsize=(15,6))
-    gs = fig.add_gridspec(3, 2, height_ratios=[1,1,4],width_ratios=[1, 1/100],wspace=0.05)
-    ax0=fig.add_subplot(gs[0,0])
-    ax1=fig.add_subplot(gs[1,0])
-    ax2=fig.add_subplot(gs[2,0])
-    cbar_ax = fig.add_subplot(gs[0:2,1])
-    return fig,ax0,ax1,ax2,cbar_ax
-
-def gspec_single_xy():
-    fig=plt.figure(figsize=(15,1))
-    gs=fig.add_gridspec(1,2,width_ratios=[1,0.01],wspace=0.02)
-    ax=fig.add_subplot(gs[0,0])
-    cbar_ax=fig.add_subplot(gs[0,1])
-    return fig,ax,cbar_ax
-
-def gspec_xy_and_profile():
-    fig=plt.figure(figsize=(15,4))
-    gs=fig.add_gridspec(2,2,height_ratios=[1,4],width_ratios=[1,0.01],wspace=0.05)
-    axis0 = fig.add_subplot(gs[0,0])
[…]

### 2024-09-17 build/lib/stream_analysis/streamLength.py

@@ -1,170 +1,0 @@
-"""
-This module contains functions that are used to predict how the length of a stream will change in time
-
-NUMERICAL 
-    - dr_RMS_distributions
-        - loads particles from plummer and gets the RMS radial distance
-    - get_two_particles_about_galactocentric_radius
-        - places two particles a distance dr away from the original position along the radial line
-    - perform_integration
-        - performs the integration of the particles
-    - get_time_from_theta
-        - uses the cylindrical angle to find how far ahead or behind in time the leading and trailing particles are from the host
-
-
-DATA BASED
-    -
-
-"""
-
-import numpy as np
-# import gcs
-
-
-# import tstrippy
-def dr_RMS_distributions(xp,yp,zp):
-    rp                  =   np.sqrt(xp**2+yp**2+zp**2) #type: ignore
-    dr                  =   np.sqrt(np.mean(rp**2))
-    return dr
-
-
-# def get_two_particles_about_galactocentric_radius(dr,position):
-#     """
-#     Given a position and a change in radius, this function returns the position and velocity of two particles that are a distance dr away from the original position along the radial line
-#     """
-#     xp0,yp0,zp0,vxp0,vyp0,vzp0 = position
-    
-#     r, theta, phi, rdot, thetadot, phidot = gcs.misc.cartesian_to_spherical(xp0,yp0,zp0,vxp0,vyp0,vzp0)
-    
-#     x0,y0,z0,vx0,vy0,vz0 = gcs.misc.spherical_to_cartesian(r-dr, theta, phi, rdot, thetadot, phidot) #type: ignore
-#     x1,y1,z1,vx1,vy1,vz1 = gcs.misc.spherical_to_cartesian(r+dr, theta, phi, rdot, thetadot, phidot) #type: ignore
-
-#     x,y,z = np.array([x0,x1]),np.array([y0,y1]),np.array([z0,z1])
-#     vx,vy,vz = np.array([vxp0,vxp0]),np.array([vyp0,vyp0]),np.array([vzp0,vzp0])
-
-#     return x,y,z,vx,vy,vz
-
-
-# def perform_integration(staticgalaxy,integrationparameters,initialkinematics,inithostperturber):
-#     Nstep = integrationparameters[2]
-#     nparticles = len(initialkinematics[0])
-#     integrator=tstrippy.integrator
-#     integrator.setstaticgalaxy(*staticgalaxy)
[…]

### 2024-09-17 build/lib/stream_analysis/tailCoordinates.py

@@ -1,201 +1,0 @@
-'''
-A module to calculate the coordinates of the particles in the stream relative to the orbit of the GC
-
-Typical usage:
-    import StreamOrbitCoords as SOC
-    # get the coordinates of the particles in the orbit frame
-
-    indexes0=SOC.GetClosestOrbitalIndex(\
-                WGC0[0],WGC0[1],WGC0[2],WP0[0],WP0[1],WP0[2])
-    xprimeOrbi0=SOC.get_xprimeOrbit(\
-                tORB0,WGC0[0],WGC0[1],WGC0[2],t0=currenttime0.value)
-    WP0f=SOC.numbaTransformGalcentricToOrbit(\
-                WP0,WGC0,xprimeOrbi0,indexes0,WP0f)
-
-'''
-import numpy as np 
-import gcs
-
-def tau_gamma(indexes:np.ndarray,yT:np.ndarray,TORB:np.ndarray,XORB:np.ndarray,YORB:np.ndarray,ZORB:np.ndarray,\
-    VXORB:np.ndarray,VYORB:np.ndarray,VZORB:np.ndarray,currenttime:float):
-    lenOrb=TORB.shape[0]
-    _,_,_,rdot,_,_=gcs.misc.cartesian_to_spherical(XORB,YORB,ZORB,VXORB,VYORB,VZORB)
-    tau = TORB[indexes] - currenttime
-    gamma=yT/np.abs(rdot[lenOrb//2])    
-    return tau,gamma
-
-def transform_from_galactico_centric_to_tail_coordinates(xp,yp,zp,vxp,vyp,vzp,tORB,xORB,yORB,zORB,vxORB,vyORB,vzORB,t0=0):
-    '''
-    transform the galcentric coordinates of the particles to the orbit coordinates
-    '''
-    indexes=get_closests_orbital_index(xORB,yORB,zORB,xp,yp,zp)
-    xprimeOrbi=get_orbital_path_length(tORB,xORB,yORB,zORB,t0=t0)
-    xprimeP=np.zeros(xp.shape)
-    yprimeP=np.zeros(xp.shape)
-    zprimeP=np.zeros(xp.shape)
-    vxprimeP=np.zeros(xp.shape)
-    vyprimeP=np.zeros(xp.shape)
-    vzprimeP=np.zeros(xp.shape)
-    for i in range(xp.shape[0]):
-        POSpart=np.array([xp[i],yp[i],zp[i]])
[…]

### 2024-09-17 paper/main.tex
updated impact time graph

@@
-\documentclass[draft]{aa}
+\documentclass{aa}
 
@@
     \centering
-    \includegraphics[width=\linewidth, trim=20 0 15 0]{monte-carlo-009-pouliasis2017pii-GCNBody-2000-milisigma-50-noisefactor-20-boxcarindexlength-shifted-0.png}
+    \includegraphics[width=\linewidth, trim=20 0 15 0]{monte-carlo-009-pouliasis2017pii-GCNBody-2000-milisigma-5-noisefactor-20-boxcarindexlength-shifted-0.png}
     \caption[]{A comparison between the vanilla and full simulations.

### 2024-09-17 paper/main.tex
updated equations for collision and 1D profile

@@
-\documentclass{aa}
+\documentclass[draft]{aa}
 
@@
       To construct the 1D density profiles, we first bin the data using the $\sqrt{N}$-rule for determining the number of bins, where N is the number of data points which is 10$^5$.
-After histogramming the 1D profiles, we apply a boxcar smoothing technique, where each central value is replaced by the median of the surrounding XX adjacent data points, corresponding to a smoothing length of approximately 300 parsecs.
+After histogramming the 1D profiles, we apply a boxcar smoothing technique, where each central value is replaced by the median of the surrounding XX adjacent data points, corresponding to a smoothing length of approximately XX parsecs.
 This procedure reduces high-frequency noise and smooths the profiles, as shown in Fig.~\ref{fig:profiles}, where the center of mass has been effectively removed.
@@
       With the smoothed 1D density profiles in hand, we consider the \textit{vanilla} simulations to be truth, and search for locations that are significantly less dense than \textit{vanilla} in a way that is greater than stochastic chance.
-To do so, we first compute the log ratio of the two profiles:
+To do so, we first only consider bins that are significant, to do this we impose a signal to noise ratio threshold, $\mathcal{SNR}$, of $\mathcal{SNR} = 5$.
+We assert that the singal is the log of the counts per bin of the vanilla 1D density profile, and we propogate the errors accordingly assert that the error in the counts is poisson distribution.
+As such, we are able to find a counts threshold, and discard any bins with out at least this number of counts.
+This $N$ is given by the transandental equation:
       \begin{equation}
-        \mathcal{R} = \log\left(\frac{\rho_f}{\rho_v}\right),
+        \mathcal{SNR} = \ln(10) \log_{10}\left(N_v\right)\sqrt{N_v}. 
         \end{equation}
-      where $\rho$ is the 1D density, $v$ corresponding to the vanilla simulation, and $f$ corresponding to the full simulation, and $R$ is the log ratio of the two.
-Using the log ratio is important, because the distribution is symmetric and zero centered.
-For example if we only used the ratio, then the codomain would be $\frac{\rho_f}{\rho_v} \in [1,+\inf]$ for $\rho_f > \rho_v$, and $\frac{\rho_f}{\rho_v} \in [0,1]$ for $\rho_f < \rho_v$.
+      By setting a threshold of $\mathcal{S/R} = 5$, we solve for the equation by using \texttt{scipy.optimize.fsolve} and find that N must be greater than 7.
 
-      Yet, we only wish to consider bins with a significant amount of counts.
-We find this threshold by asserting that the signal to noise ratio must be greater than 50.
-The signal is based on the \textit{vanilla} profile, and the noise is the error propogation given poisson uncertainties for the number of counts a given bin:
+      After discarding the insignificant bins, we compute the log ratio of the counts between the vanilla and full simulations: 
       \begin{equation} 
-        \mathcal{S}/\mathcal{N} =50= \frac{\log{N}}{\frac{1}{\ln(10)} \sqrt{N}}
+        \mathcal{R}_i = \log_{10}\left(\frac{N_{f,i}}{N_{v,i}}\right),
         \end{equation}
-
-      we compute the difference between the logarithms of the binned counts for each profile and search for points that fall more than two standard deviations below the mean.
-However, we only consider regions that surpass a predefined significance threshold, where the significance is determined by the uncertainties in the binned counts governed by Poisson statistics multiplied by a factor of 50.
-This noise floor was chosen to eliminate the low-density regions near Palomar 5, where gaps are more likely caused by epicyclic oscillations rather than impacts from globular clusters.
-The identified gap candidates are marked and labeled on the plots.
+      where $\mathcal{R}$ is the log ratio, $N_f$ are the number of counts from the full simulation, and $i$ is the index for a given bin.
+Subsequently, we inspect the resultant distribution of $\mathcal{R}_i$.
+This distribution would be gaussian if the differences between the density profiles are due to the sum of stochastic processes similar in magnitude, as in the central limit theorem.
+Therefore, we flag all locations that are underdense by two standard deviations.
 
@@
 In these cases, marginalizing over y' can dilute the gap's signal, making it difficult to detect.
-An example of this is the small gap around 5 kpc, which, as we will demonstrate later, is caused by the interaction with NGC 2808 and appears in most simulations.
+As will be discussed in the results, this is the case for most of the gaps caused by NGC2808.
+Thus, studying the differences between the two 1D density profiles is an \textit{aid} to assist in visual inspect.
 
@@
 To do so, we fit a third-order parametric polynomial to the stream, using the saved snapshots from our simulations.
-    \begin{align}
-      x(\tau) &= a_0 + a_1 \tau + a_2 \tau^2 + a_3 \tau^3, \\ 
-      y(\tau) &= b_0 + b_1 \tau + b_2 \tau^2 + b_3 \tau^3, \\
-      z(\tau) &= c_0 + c_1 \tau + c_2 \tau^2 + c_3 \tau^3,
-      \end{align}
+
+    \begin{equation}
+      \vec{s}(\tau) = 
+      \left\{
+        \begin{aligned}
+          x(\tau) &= a_0 + a_1 \tau + a_2 \tau^2 + a_3 \tau^3 \\ 
+          y(\tau) &= b_0 + b_1 \tau + b_2 \tau^2 + b_3 \tau^3 \\
+          z(\tau) &= c_0 + c_1 \tau + c_2 \tau^2 + c_3 \tau^3
+        \end{aligned}
+      \right.
+  \end{equation}  
     where $x$, $y$, and $z$ represent the parametric line describing the stream in galactocentric coordinates, $\tau$ is the stream coordinate in time as described in the previous section, and is used as the independent variable in this analysis.
@@
 
-    \begin{align}
-      x(t,\tau) &= a_0(t) + a_1(t)\tau + a_2(t) \tau^2 + a_3(t)\tau^3, \\ 
-      y(t,\tau) &= b_0(t) + b_1(t)\tau + b_2(t) \tau^2 + b_3(t)\tau^3, \\
+    \begin{equation}
+      \vec{s}(t,\tau) = 
+      \left\{
+      \begin{aligned}
+        x(t,\tau) &= a_0(t) + a_1(t)\tau + a_2(t) \tau^2 + a_3(t)\tau^3 \\ 
+        y(t,\tau) &= b_0(t) + b_1(t)\tau + b_2(t) \tau^2 + b_3(t)\tau^3 \\
         z(t,\tau) &= c_0(t) + c_1(t)\tau + c_2(t) \tau^2 + c_3(t)\tau^3.
-      \end{align}
+        \end{aligned}
+      \right.
+    \end{equation}
     The values of the coefficients as a function of time are obtained through linear interpolation, ensuring that the coefficients at the snapshot times match the values constrained by the simulation data.
@@
 
-    \begin{align}
+    \begin{equation}
+        \begin{aligned}
           \Delta \vec{v} &=\vec{v_p} - \vec{v_s} \\
@@
           w_\perp &=  \sqrt{\Delta v ^2 - w_\parallel ^ 2}
-      \end{align}
+          \end{aligned}
+        \end{equation}
 

### 2024-09-17 paper/main.tex
passed the text to chatGPT for edits

@@
 
-    \subsubsection*{Density maps and profiles}
-      For each of the 50 simulations, we compare the final snapshots between the \texttt{vanilla} and full simulations in the tail coordinate system.
-We inspect these differences by creating 2D density maps of the streams.
-Subsequently, we wish to employ a quantitative method to make detections discover when the \textit{full} simulations are significantly less dense than the \textit{vanilla} simulations.
-To do so, we marginalize over the y'-coordinate to produce 1D density profiles along the x'-axis.
-These comparisons are illustrated in the results section (see Fig.~\ref{fig:profiles}).
+    \subsubsection*{Density Maps and Profiles}
+    For each of the 50 simulations, we compare the final snapshots between the \texttt{vanilla} and \textit{full} simulations using the tail coordinate system.
+We inspect these differences by generating 2D density maps of the streams.
+Additionally, we marginalize over the $y'$-coordinate to produce 1D density profiles along the $x'$-axis.
+These comparisons are shown in the results section (see Fig.~\ref{fig:profiles}).
 
-      To construct the 1D density profiles, we first bin the data using the $\sqrt{N}$-rule for determining the number of bins, where N is the number of data points which is 10$^5$.
-After histogramming the 1D profiles, we apply a boxcar smoothing technique, where each central value is replaced by the median of the surrounding XX adjacent data points, corresponding to a smoothing length of approximately XX parsecs.
-This procedure reduces high-frequency noise and smooths the profiles, as shown in Fig.~\ref{fig:profiles}, where the center of mass has been effectively removed.
+    To construct the 1D density profiles, we first bin the data using the $\sqrt{N}$ rule, where $N$ is the number of data points ($N = 10^5$).
+After binning the 1D profiles, we apply a boxcar smoothing technique.
+Each central value is replaced by the median of the surrounding XX adjacent data points, corresponding to a smoothing length of approximately XX parsecs.
+This procedure reduces high-frequency noise and smooths the profiles, effectively removing the center of mass (see Fig.~\ref{fig:profiles}).
 
-      With the smoothed 1D density profiles in hand, we consider the \textit{vanilla} simulations to be truth, and search for locations that are significantly less dense than \textit{vanilla} in a way that is greater than stochastic chance.
-To do so, we first only consider bins that are significant, to do this we impose a signal to noise ratio threshold, $\mathcal{SNR}$, of $\mathcal{SNR} = 5$.
-We assert that the singal is the log of the counts per bin of the vanilla 1D density profile, and we propogate the errors accordingly assert that the error in the counts is poisson distribution.
-As such, we are able to find a counts threshold, and discard any bins with out at least this number of counts.
-This $N$ is given by the transandental equation:
+    With the smoothed 1D density profiles in hand, we treat the \textit{vanilla} simulations as the reference model and search for regions in the \textit{full} simulations that are significantly underdense compared to \textit{vanilla}, surpassing stochastic fluctuations.
+To accomplish this, we impose a signal-to-noise ratio threshold, $\mathcal{SNR}$.
+The signal is defined as the log of the counts per bin from the \textit{vanilla} 1D density profile, and errors are propagated assuming a Poisson distribution.
+We then compute a threshold for the number of counts in the \textit{vanilla} simulation, $N_v$, using the transcendental equation:
     \begin{equation}
-        \mathcal{SNR} = \ln(10) \log_{10}\left(N_v\right)\sqrt{N_v}. 
+        \mathcal{SNR} = \ln(10) \log_{10}\left(N_v\right) \sqrt{N_v}.
     \end{equation}
-      By setting a threshold of $\mathcal{S/R} = 5$, we solve for the equation by using \texttt{scipy.optimize.fsolve} and find that N must be greater than 7.
-
-      After discarding the insignificant bins, we compute the log ratio of the counts between the vanilla and full simulations: 
+    By setting $\mathcal{SNR} = 5$, we solve for $N$ using \texttt{scipy.optimize.fsolve}, finding that $N$ must be greater than 7.
+After discarding insignificant bins (i.e., those with counts below the threshold), we compute the log ratio of the counts between the \textit{vanilla} and full simulations:
     \begin{equation}
@@
     \end{equation}
-      where $\mathcal{R}$ is the log ratio, $N_f$ are the number of counts from the full simulation, and $i$ is the index for a given bin.
-Subsequently, we inspect the resultant distribution of $\mathcal{R}_i$.
-This distribution would be gaussian if the differences between the density profiles are due to the sum of stochastic processes similar in magnitude, as in the central limit theorem.
-Therefore, we flag all locations that are underdense by two standard deviations.
+    where $\mathcal{R}_i$ is the log ratio, $N_{f,i}$ are the counts from the full simulation, and $N_{v,i}$ are the counts from the \textit{vanilla} simulation for each bin $i$.
+We then analyze the distribution of $\mathcal{R}_i$.
+If the differences between the density profiles are primarily due to stochastic processes of similar magnitude, this distribution should resemble a Gaussian as expected from the central limit theorem.
+Thus, we flag all regions where the density is underdense by more than two standard deviations, which should highlight regions whose underdensity is unlikely to be the result of the sum of stochastic processess but rather the passage of another globular cluster.
 
-      This method, however, is not without limitations, particularly when detecting smaller gaps.
-As discussed in Erkal et al.
-(2015), a gap's size is not necessarily indicative of a weak impact but could result from a more recent interaction.
-Additionally, because our streams have finite width, some gaps may be oriented obliquely relative to the stream axis.
-In these cases, marginalizing over y' can dilute the gap's signal, making it difficult to detect.
-As will be discussed in the results, this is the case for most of the gaps caused by NGC2808.
-Thus, studying the differences between the two 1D density profiles is an \textit{aid} to assist in visual inspect.
+    This method, however, has its limitations, especially when detecting smaller gaps.
+As outlined by Erkal et al.
+(2015), a gap's size may not indicate a weak impact, but rather a more recent one.
+Additionally, since our streams have finite width, some gaps might be oblique to the stream axis.
+In such cases, marginalizing over $y'$ can dilute the gap's signal, making it harder to detect.
+This limitation is particularly evident in gaps caused by NGC2808, as discussed in the results.
+Therefore, this quantitative analysis serves as an \textit{aid} to visual inspection rather than a complete substitute for it.
 

### 2024-09-17 paper/main.tex
added figure links

@@
 
-      The core of our analysis is presented in Fig.~XXX.
+      The core of our analysis is presented in Fig.~\ref{fig:force-on-orbit}.
 The bottom panel shows the evolution of the stream density over time.
@@
 Next, we modeled Palomar 5's orbit as a proxy for its stream and sampled points along the orbit to measure the gravitational force exerted by other globular clusters.
-This force data is displayed in the top panel of Fig.~XXX, showing how the total gravitational acceleration on Palomar 5's stream evolves over its length throughout the simulation.
+This force data is displayed in the top panel of Fig.~\ref{fig:force-on-orbit}, showing how the total gravitational acceleration on Palomar 5's stream evolves over its length throughout the simulation.
 
@@
 For a handful of simulations, to double check that the verdict made correctly when passing from suspect to culprit, we re-compute the simulations yet individually adding one globular cluster at a time.
-As a result, we are able to confirm that singular gaps arrise from the suspected clusters.
+As a result, we are able to confirm that singular gaps arrise from the suspected clusters, an example of which is shown in Fig.~\ref{fig:decomposition}.
 
@@
 }
-    \label{fig:profiles}
+    \label{fig:force-on-orbit}
     \end{figure*}
@@
 (I think I'm going to get rid of the DIFF map)}
-    \label{fig:profiles}
+    \label{fig:decomposition}
     \end{figure*}

### 2024-09-18 paper/main.tex
added impact statistics!
added 6

--- 2024-09-17 paper/main.tex
@@
 
-\title{Palomar 5 Gaps}
+\title{Palomar 5 Gaps 6}
 
@@
           \email{c.ptolemy@hipparch.uheaven.space}
-          \thanks{The university of heaven temporarily does not
-                  accept e-mails}
+          \thanks{The }
           }
@@
     Briefly, the x'coordinate represents the position of a particle along the orbit relative to the globular cluster—whether it is ahead or behind the cluster.
-The y' coordinate measures the particle's distance within the orbital plane, where positive values indicate that the particle is farther from the Galactic center and negative values indicate that it is closer.
+The y'coordinate measures the particle's distance within the orbital plane, where positive values indicate that the particle is farther from the Galactic center and negative values indicate that it is closer.
 
@@
       For each of the 50 simulations, we compare the final snapshots between the \texttt{vanilla} and full simulations in the tail coordinate system.
-We inspect these differences by creating 2D density maps of the streams.
-Subsequently, we wish to employ a quantitative method to make detections discover when the \textit{full} simulations are significantly less dense than the \textit{vanilla} simulations.
-To do so, we marginalize over the y'-coordinate to produce 1D density profiles along the x'-axis.
-These comparisons are illustrated in the results section (see Fig.~\ref{fig:profiles}).
+This is done by creating 2D density maps of the streams, followed by marginalizing over the y'-coordinate to produce 1D density profiles along the x'-axis.
+These comparisons are illustrated in the results section (see Fig.~XXX).
 
-      To construct the 1D density profiles, we first bin the data using the $\sqrt{N}$-rule for determining the number of bins, where N is the number of data points which is 10$^5$.
-After histogramming the 1D profiles, we apply a boxcar smoothing technique, where each central value is replaced by the median of the surrounding XX adjacent data points, corresponding to a smoothing length of approximately XX parsecs.
-This procedure reduces high-frequency noise and smooths the profiles, as shown in Fig.~\ref{fig:profiles}, where the center of mass has been effectively removed.
+      To construct the 1D density profiles, we first bin the data using the  $\sqrt{N}$-rule for determining the number of bins.
+After histogramming the 1D profiles, we apply a boxcar smoothing technique, where each central value is replaced by the median of the surrounding XX adjacent data points, corresponding to a smoothing length of approximately 300 parsecs.
+This procedure reduces high-frequency noise and smooths the profiles, as shown in Fig.~XX, where the center of mass has been effectively removed.
 
-      With the smoothed 1D density profiles in hand, we consider the \textit{vanilla} simulations to be truth, and search for locations that are significantly less dense than \textit{vanilla} in a way that is greater than stochastic chance.
-To do so, we first only consider bins that are significant, to do this we impose a signal to noise ratio threshold, $\mathcal{SNR}$, of $\mathcal{SNR} = 5$.
-We assert that the singal is the log of the counts per bin of the vanilla 1D density profile, and we propogate the errors accordingly assert that the error in the counts is poisson distribution.
-As such, we are able to find a counts threshold, and discard any bins with out at least this number of counts.
-This $N$ is given by the transandental equation:
-      \begin{equation}
-        \mathcal{SNR} = \ln(10) \log_{10}\left(N_v\right)\sqrt{N_v}. 
-        \end{equation}
-      By setting a threshold of $\mathcal{S/R} = 5$, we solve for the equation by using \texttt{scipy.optimize.fsolve} and find that N must be greater than 7.
-
-      After discarding the insignificant bins, we compute the log ratio of the counts between the vanilla and full simulations: 
-      \begin{equation} 
-        \mathcal{R}_i = \log_{10}\left(\frac{N_{f,i}}{N_{v,i}}\right),
-        \end{equation}
-      where $\mathcal{R}$ is the log ratio, $N_f$ are the number of counts from the full simulation, and $i$ is the index for a given bin.
-Subsequently, we inspect the resultant distribution of $\mathcal{R}_i$.
-This distribution would be gaussian if the differences between the density profiles are due to the sum of stochastic processes similar in magnitude, as in the central limit theorem.
-Therefore, we flag all locations that are underdense by two standard deviations.
+      With the smoothed 1D profiles in hand, we proceed to compare them by identifying regions where the full simulations (GCGC) show significant underdensity compared to the \texttt{vanilla} simulations.
+To achieve this, we compute the difference between the logarithms of the binned counts for each profile and search for points that fall more than two standard deviations below the mean.
+However, we only consider regions that surpass a predefined significance threshold, where the significance is determined by the uncertainties in the binned counts governed by Poisson statistics multiplied by a factor of 50.
+This noise floor was chosen to eliminate the low-density regions near Palomar 5, where gaps are more likely caused by epicyclic oscillations rather than impacts from globular clusters.
+The identified gap candidates are marked and labeled on the plots.
 
@@
 In these cases, marginalizing over y' can dilute the gap's signal, making it difficult to detect.
-As will be discussed in the results, this is the case for most of the gaps caused by NGC2808.
-Thus, studying the differences between the two 1D density profiles is an \textit{aid} to assist in visual inspect.
+An example of this is the small gap around 5 kpc, which, as we will demonstrate later, is caused by the interaction with NGC 2808 and appears in most simulations.
 
@@
 To do so, we fit a third-order parametric polynomial to the stream, using the saved snapshots from our simulations.
-
-    \begin{equation}
-      \vec{s}(\tau) = 
-      \left\{
-        \begin{aligned}
-          x(\tau) &= a_0 + a_1 \tau + a_2 \tau^2 + a_3 \tau^3 \\ 
-          y(\tau) &= b_0 + b_1 \tau + b_2 \tau^2 + b_3 \tau^3 \\
-          z(\tau) &= c_0 + c_1 \tau + c_2 \tau^2 + c_3 \tau^3
-        \end{aligned}
-      \right.
-  \end{equation}  
+    \begin{align}
+      x(\tau) &= a_0 + a_1 \tau + a_2 \tau^2 + a_3 \tau^3, \\ 
+      y(\tau) &= b_0 + b_1 \tau + b_2 \tau^2 + b_3 \tau^3, \\
+      z(\tau) &= c_0 + c_1 \tau + c_2 \tau^2 + c_3 \tau^3,
+      \end{align}
     where $x$, $y$, and $z$ represent the parametric line describing the stream in galactocentric coordinates, $\tau$ is the stream coordinate in time as described in the previous section, and is used as the independent variable in this analysis.
@@
 
-    \begin{equation}
-      \vec{s}(t,\tau) = 
-      \left\{
-      \begin{aligned}
-        x(t,\tau) &= a_0(t) + a_1(t)\tau + a_2(t) \tau^2 + a_3(t)\tau^3 \\ 
-        y(t,\tau) &= b_0(t) + b_1(t)\tau + b_2(t) \tau^2 + b_3(t)\tau^3 \\
+    \begin{align}
+      x(t,\tau) &= a_0(t) + a_1(t)\tau + a_2(t) \tau^2 + a_3(t)\tau^3, \\ 
+      y(t,\tau) &= b_0(t) + b_1(t)\tau + b_2(t) \tau^2 + b_3(t)\tau^3, \\
       z(t,\tau) &= c_0(t) + c_1(t)\tau + c_2(t) \tau^2 + c_3(t)\tau^3.
-        \end{aligned}
-      \right.
-    \end{equation}
+      \end{align}
     The values of the coefficients as a function of time are obtained through linear interpolation, ensuring that the coefficients at the snapshot times match the values constrained by the simulation data.
@@
 
-    \begin{equation}
-        \begin{aligned}
+    \begin{align}
       \Delta \vec{v} &=\vec{v_p} - \vec{v_s} \\
@@
       w_\perp &=  \sqrt{\Delta v ^2 - w_\parallel ^ 2}
-          \end{aligned}
-        \end{equation}
+      \end{align}
 
@@
     \centering
-    \includegraphics[width=\linewidth, trim=20 0 15 0]{monte-carlo-009-pouliasis2017pii-GCNBody-2000-milisigma-5-noisefactor-20-boxcarindexlength-shifted-0.png}
+    \includegraphics[width=\linewidth, trim=20 0 15 0]{monte-carlo-009-pouliasis2017pii-GCNBody-2000-milisigma-50-noisefactor-20-boxcarindexlength-shifted-0.png}
     \caption[]{A comparison between the vanilla and full simulations.
@@
   \end{figure}
+
   \begin{figure*}
@@
 
-  \section{Discussion}
-    $10^5-10^6$. Sotto alone averbbero questo effetto?
+  Here, I am typing words
 
@@
 
-  \section{Gallery of Gaps}
-
-    \begin{figure*}
-      \centering
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-000.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-001.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-002.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-003.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-004.png}
-      \caption{Gap Gallery}
-      \label{fig:TailCoordinates}
-    \end{figure*}
-
-    \begin{figure*}
-      \centering
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-005.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-006.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-007.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-008.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-009.png}
-      \caption{Gap Gallery}
-      \label{fig:TailCoordinates}
-    \end{figure*}
-
-    \begin{figure*}
-      \centering
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-010.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-011.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-012.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-013.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-014.png}
-      \caption{Gap Gallery}
-      \label{fig:TailCoordinates}
-    \end{figure*}
-
-    \begin{figure*}
-      \centering
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-015.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-016.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-017.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-018.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-019.png}
-      \caption{Gap Gallery}
-      \label{fig:TailCoordinates}
-    \end{figure*}
-
-    \begin{figure*}
-      \centering
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-020.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-021.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-022.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-023.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-024.png}
-      \caption{Gap Gallery}
-      \label{fig:TailCoordinates}
-    \end{figure*}
-
-    \begin{figure*}
-      \centering
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-025.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-026.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-027.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-028.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-029.png}
-      \caption{Gap Gallery}
-      \label{fig:TailCoordinates}
-    \end{figure*}
-
-    \begin{figure*}
-      \centering
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-030.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-031.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-032.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-033.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-034.png}
-      \caption{Gap Gallery}
-      \label{fig:TailCoordinates}
-    \end{figure*}
-
-    \begin{figure*}
-      \centering
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-035.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-036.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-037.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-038.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-039.png}
-      \caption{Gap Gallery}
-      \label{fig:TailCoordinates}
-    \end{figure*}
-
-    \begin{figure*}
-      \centering
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-040.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-041.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-042.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-043.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-044.png}
-      \caption{Gap Gallery}
-      \label{fig:TailCoordinates}
-    \end{figure*}
-
-    \begin{figure*}
-      \centering
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-045.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-046.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-047.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-048.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-049.png}
-      \caption{Gap Gallery}
-      \label{fig:TailCoordinates}
-    \end{figure*}
-
 \end{appendix}

### 2024-09-19 paper/main.tex
added change in momentum

--- 2024-09-17 paper/main.tex
@@
   The goal of this study is to contribute to the ongoing effort to use stellar streams as probes for dark matter on sub-galactic scales.
-According to the cold dark matter model, which is supported by observations, galaxies form hierarchically, leading to a mass distribution of dark matter subhalos that follows a power-law.
-This suggests that numerous small dark matter clumps should exist within galaxies.
+According to the cold dark matter model, which is supported by observations, galaxies form hierarchically.
+As shown in Springel et al 2008 in the suite of Acquarious simulations, $\lambda$-CDM predicts a distribution of masses of dark matter subhalos that follows a power-law.
+The authors also demonstrate that subhaloes should exist within other haloes.
+To my knowledge, thanks to the introduction of Bonaca et al 2019 who pointed this paper out to me, the smallest dark matter sub halo detected to date was found within an enstien ring from gravitational lensing by Vigetti et al 20XX.
+
+  Over the years, people began to think of the role that subhalos can have on effecting stellar streams.
 \citet{rodrigo_ibata_uncovering_2002} suggested that the presence of said dark matter subhaloes dark matter subhalos can influence the stellar streams by diffussing their distribution of orbital elements.
@@
 Instead of treating the clusters as point masses that only experience the gravitational potential of the galaxy, we now compute N-body interactions between the clusters during the backward integration over 5 billion years.
+The equation of motion for the globular clusters is thus:
+    \begin{equation}
+      \ddot{\vec{r}}_i = -\nabla \Phi(R_i,z_i) + \left.\sum_{j\neq i} \frac{Gm_j}{\left(|\vec{r}_j - \vec{r}_i|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j - \vec{r}_i\right).
+      \end{equation}
     Although these interactions were included, we observed no significant exchanges of orbital energy or momentum between clusters.
@@
 We do not perform direct N-body simulations between star particles; instead, at each time step, we compute the positions of all globular clusters and apply their forces to the star particles.
+The equation of motion is thus:
+    \begin{equation}
+      \ddot{\vec{r}}_p = -\nabla \Phi(R_p,z_p) + \left.\sum_{j} \frac{Gm_j}{\left(|\vec{r}_j(t) - \vec{r}_p|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j(t)- \vec{r}_p\right)
+      \end{equation}
 
@@
         \begin{aligned}
-          \Delta \vec{v} &=\vec{v_p} - \vec{v_s} \\
-          w_\parallel &= \left(\Delta \vec{v}\right)\cdot \hat{v_s}\\  
+          \delta \vec{v} &=\vec{v_p} - \vec{v_s} \\
+          w_\parallel &= \left(\delta \vec{v}\right)\cdot \hat{v_s}\\  
           w_\perp &=  \sqrt{\Delta v ^2 - w_\parallel ^ 2}
@@
     \includegraphics[width=1\linewidth]{mass_size_plane.png}
+    \includegraphics[width=1\linewidth]{impact_geometry_statistics_deltaP.png}
     \caption{One plot shows the distribution in energy and angular momentum, the next are the inernal proeprties of the clusters, and the last show when the clusters commit their crime, and how many times it happens accross the 50 simulations.
@@
   \end{figure}
+
+  \section{Impact Geometry Statistics}
+
+  \begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{impact_geometry_statistics.png}
+    \caption{Impact geometry statistics}
+    \label{fig:impact_geometry_statistics}    
+    \end{figure*}
 

### 2024-09-19 paper/main.tex
small edits, replacing XXXs

@@
 Our focus is on the Palomar 5 stream, which has been well-studied and has observable gaps, a high membership count, and favorable visibility due to its position above the galactic disk.
+We show a result from one our simulations in \ref{fig:stream_on_sky}.
 In the following sections, we describe the modeling of this stream, the gravitational interactions with the galaxy and other globular clusters, and the statistical analysis of the perturbations affecting the stream.
+
+  \begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{stream_on_sky_Pal5_monte-carlo-009_pouliasis2017pii-GCNBody.png}
+    \caption[]{A model of the Palomar 5 stream in an axis-symmetric galactic potential plus all other globular clusters.
+The top plot is a scatter plot of star-particles that escaped the cluster due to tidal forces, the bottom plot shows the marginalized 1D density profile in longitude.
+Two large gaps are present and are due to the passage of two globular clusters.
+}
+    \label{fig:stream_on_sky}
+    \end{figure*}
 
@@
 
-    To generate initial conditions, we treat the five-dimensional phase space variables (position and velocity components) for each globular cluster as Gaussian distributions.
+    To generate initial conditions, we treat the four phase space variables with uncertainties (distance, line-of-sight velocities, and proper motions) and mass for each globular cluster as five-dimensional Gaussian distributions.
 Using the \texttt{np.random.gauss} function, we generate 50 samples per cluster, producing 50 sets of initial conditions for the entire globular cluster system.
@@
 
-    For the galactic potential, we employed the static model proposed by Pouliasis et al.
+    For the galactic potential, we employed the second static model proposed by Pouliasis et al.
 (2017), which consists of a superposition of a thin disk, thick disk, and a dark matter halo.
@@
       \ddot{\vec{r}}_i = -\nabla \Phi(R_i,z_i) + \left.\sum_{j\neq i} \frac{Gm_j}{\left(|\vec{r}_j - \vec{r}_i|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j - \vec{r}_i\right).
-      \end{equation}
+      \end{equation}\label{eq:GCNBody}
     Although these interactions were included, we observed no significant exchanges of orbital energy or momentum between clusters.
 The resulting orbits were saved to data files for subsequent analysis.
+We integrate clusters into the past by 5 Gyr.
 
-    At the computed positions 5 Gyr ago, each globular cluster was replaced by a system of 100,000 particles, sampled from a Plummer distribution to represent the stellar content of the cluster.
+    At the computed positions 5 Gyr ago, each globular cluster was gifted a system of 100,000 particles, sampled from a Plummer distribution to represent the stellar content of the cluster.
 We then integrated the evolution of these particles forward in time to the present day.
 During this forward integration, each star particle feels the gravitational influence of the galaxy and all globular clusters, but not other star particles.
-We do not perform direct N-body simulations between star particles; instead, at each time step, we compute the positions of all globular clusters and apply their forces to the star particles.
+We do not perform direct N-body simulations between star particles as in eq.~\ref{eq:GCNBody}; instead, at each time step, we compute the positions of all globular clusters and apply their forces to the star particles.
 The equation of motion is thus:
@@
     \centering
-    \includegraphics[width=\linewidth]{stream_on_sky_Pal5_monte-carlo-009_pouliasis2017pii-GCNBody.png}
-    \caption[]{A model of the Palomar 5 stream in an axis-symmetric galactic potential plus all other globular clusters.
-The top plot is a scatter plot of star-particles that escaped the cluster due to tidal forces, the bottom plot shows the marginalized 1D density profile in longitude.
-Two large gaps are present and are due to the passage of two globular clusters.
-}
-    \end{figure*}
-
-  \begin{figure*}
-    \centering
     \includegraphics[width=\linewidth, trim=20 0 15 0]{monte-carlo-009-pouliasis2017pii-GCNBody-2000-milisigma-5-noisefactor-20-boxcarindexlength-shifted-0.png}

### 2024-09-20 paper/main.tex
tightened up the methods

@@
 
-    Briefly, the x'coordinate represents the position of a particle along the orbit relative to the globular cluster—whether it is ahead or behind the cluster.
+    Briefly, the x' coordinate represents the position of a particle along the orbit relative to the globular cluster—whether it is ahead or behind the cluster.
 The y' coordinate measures the particle's distance within the orbital plane, where positive values indicate that the particle is farther from the Galactic center and negative values indicate that it is closer.
@@
           \mathcal{SNR} = \ln(10) \log_{10}\left(N_v\right) \sqrt{N_v}.
-        \end{equation}
+        \end{equation} \label{eq:density_threshold}
       By setting $\mathcal{SNR} = 5$, we solve for $N$ using \texttt{scipy.optimize.fsolve}, finding that $N$ must be greater than 7.
@@
 (2015), demonstrating that action-angle variables provide a useful coordinate system for analyzing stream evolution, as actions are conserved and the angles associated with the stream's growth evolve linearly over time.
-Although we became aware of this work only after completing our analysis, we note that $\tau$ is a suitable approximation and behaves similarly to the angle variable corresponding to the azimuthal action: $\tau \approx \theta_{\phi,i} - \theta_{\phi,\text{GC}}$
+Although we became aware of this work only after completing our analysis, we note that $\tau$ is a suitable approximation and behaves similarly to the angle variable corresponding to the azimuthal action: $\tau \approx \theta_{\phi,i} - \theta_{\phi,\text{GC}}$.
 
@@
 The bottom panel shows the evolution of the stream density over time.
-To avoid extreme low-density regions at the stream's edges, we applied a density threshold to focus on the more significant areas of the stream.
+To avoid extreme low-density regions at the stream's edges, we applied the same density threshold as from eq.~\ref{eq:density_threshold} to focus on the more significant areas of the stream.
 Next, we modeled Palomar 5's orbit as a proxy for its stream and sampled points along the orbit to measure the gravitational force exerted by other globular clusters.
@@
     From this analysis, we can conclude that a more massive perturber, passing closer to the particle, and moving more slowly, will have a greater impact.
-It's important to note that the momentum change is inversely proportional to the velocity of the perturber, which contrasts with the intuition from elastic collisions, such as those between billiard balls, where higher velocities generally result in greater impact.
+It's important to note that the momentum change is inversely proportional to the velocity of the perturber, which contrasts with the intuition from elastic collisions, such as those between billiard balls, where higher velocities result in greater impacts.
 
     Erkal et al.
-(2015) extended this impact theory to a one-dimensional stream, studying how the momentum change of a particle varies as a function of its distance from the point of impact.
-In their analysis, the perturber was modeled not as a point mass, but as a Plummer sphere with a characteristic length scale $r_p$.
+(2015) extended this impact theory from one point mass impacting another point mass, to studying how an extended body impacts as stream.
+This quantifies the change in momentum of a given particle as a function of its distance from the point of greatest impact.
+In their analysis, the perturber was modeled as a Plummer sphere.
 Since the stream is not zero-dimensional and has length, both the parallel and perpendicular components of the velocity must be considered.
@@
     In this refinement process, we aim to pinpoint both the exact location of the impact along the stream and the precise moment it occurred.
-To do so, we fit a third-order parametric polynomial to the stream, using the saved snapshots from our simulations.
-
+To do so, we fit a third-order parametric polynomial to the stream, using the saved snapshots from our simulations:
     \begin{equation}
@@
 Consequently, we can parameterize the stream as a function of both simulation time and position along the stream:
-
     \begin{equation}
@@
 With equations for both the stream and the perturber as functions of time, we identify the time and location of impact by minimizing a cost function, defined as the distance between the stream and the perturber:
-
-    \begin{equation} b(t, \tau) = \left\lVert \vec{s}(t, \tau) - \vec{p}(t) \right\rVert, \end{equation}
-
+    \begin{equation} 
+      b(t, \tau) = \left\lVert \vec{s}(t, \tau) - \vec{p}(t) \right\rVert, 
+      \end{equation}
     where $\vec{s}(t, \tau)$ is the galactocentric position of a point on the stream, $\vec{p}(t)$ is the position of the perturber, and $b$ is the distance between the two.
@@
 The parallel and perpendicular components of the perturber's velocity relative to the stream are given by:
-
     \begin{equation}
@@
       \end{equation}
-
     where $\vec{v}_p$ and $\vec{v}_s$ are the velocities of the perturber and the stream, respectively.

### 2024-09-20 paper/main.tex
starting the results

@@
 
-  \subsection{Gap detection}
+  \subsection{Gap detection} \label{sec:gap_detection}
 
@@
 These comparisons are shown in the results section (see Fig.~\ref{fig:profiles}).
+
+      \begin{figure*}
+        \centering
+        \includegraphics[width=\linewidth, trim=20 0 15 0]{monte-carlo-009-pouliasis2017pii-GCNBody-2000-milisigma-5-noisefactor-20-boxcarindexlength-shifted-0.png}
+        \caption{A comparison between the density maps and profiles of the \textit{full} and \textit{vanilla} simulations is presented.
+The locations where the full simulation is less dense than the vanilla simulation by 2-$\sigma$ is highlighted by blue vertical bars.
+Bins below the noise threshold are not considered when measuring the differences between the two 1D profiles.
+The 1D profiles have been smoothed with a median box-car filter.}
+        \label{fig:profiles}
+        \end{figure*}
 
@@
       \Delta p \approx \text{Force} \times \text{interaction time} = \frac{GMm}{b^2} \times \frac{b}{v} = \frac{GM}{bv}, 
-      \end{equation}
+      \end{equation} \label{eq:change_in_momentum}
     where $M$ is the mass of the perturber, $m$ is the mass of the particle, $b$ is the impact parameter, $v$ is the relative velocity of the perturber with respect to the particle, and $G$ is the gravitational constant.
@@
 
+  One of the 50 simulations is shown in Fig.~\ref{fig:stream_on_sky}--which we selected for its clear and prominent gaps which is visibile when considering the scatter plot of the stars on the sky and becomes even more apparent when marginalizing over latitutde to see it's 1D density over longitude.
+
+  We then transformed the stream to tail coordinates for the both the \textit{vanilla} and \textit{full} simulations in Fig.~\ref{fig:profiles}.
+As described in section~\ref{sec:gap_detection}, we locate regions on the stream where the full simulations are significantly less dense than the vanilla simulations.
+Notice the \textit{Tiny Gap} labeled in the middle plot.
+This signal is not present in the 1D profile since it is too small in size and at an oblique angle, thus is lost when marginalizing.
+This process was followed for all 50 simulations to identify possible gaps.
+We find that small gaps are identifiable by eye rather well, where the 1D profiles aid in finding subtle yet extended underdensities.
+All simulations and labeled gaps are presented in the Appendix~\ref{appendix:gallery_of_gaps}.
+The profile method was useful in identifying the gaps in sampling 002, 006, 008, 029, 030.
+
   I'm gonna go remake the figures before I adapt the text.
@@
 I wanna compare this to the known length of Palomar 5, how does this reduce the results, and are there observed gaps in Palomar 5 that are consistent with these simulations?
-
-  \begin{figure*}
-    \centering
-    \includegraphics[width=\linewidth, trim=20 0 15 0]{monte-carlo-009-pouliasis2017pii-GCNBody-2000-milisigma-5-noisefactor-20-boxcarindexlength-shifted-0.png}
-    \caption[]{A comparison between the vanilla and full simulations.
-This showcases our gap detection method, in which the 2D maps in \textit{tail coordinates} are inspected, and their 1D marginalized density distributions are compared quantitatively.
-The $x'$ coordinate indicates distance from the center of mass of Pal5 along it's orbit.
-The two profiles are subtraced in locations where the signal is above the noise threshold, and locations that are more that are underdense by 2-$\sigma$ are highlighted with vertical blue lines, and labeled \textit{gap candidate}.
-This also showcases the limitation in this 1D profile method since a very skinny and oblique gap is present, whose signal is washed out from marginalizing.}
-    \label{fig:profiles}
-    \end{figure*}
 
@@
 
-  \section{Gallery of Gaps}
+  \section{Gallery of Gaps} \label{appendix:gallery_of_gaps}
 

### 2024-09-23 paper/main.tex
did an edit of the introduction

@@
 
-  The presence of dark matter has substantial observational evidence on scales larger than individual galaxies.
-The aim of this work is to add more context to aid in the search for evidence of dark matter on the sub-galactic scale and within the Milky Way.
-Initially proposed by [Author Name] in [Year], the concept of dark matter emerged to explain the anomalously high velocities of stars in distant dwarf galaxies.
-These velocities exceed what would be expected given the visible mass exerting gravitational forces on them.
-Furthermore, dark matter's influence extends to cosmological scales, as evidenced by the bending of light trajectories within cosmic web filaments due to the mass of dark matter.
-Despite significant advancements, understanding the true nature of dark matter remains elusive, with the scientific community continually seeking evidence of its existence beyond its gravitational effects on cosmic and extragalactic scales.
-A current focus is the detection of dark matter at sub-galactic scales, such as through the study of stellar streams.
+  The existence of dark matter is supported by substantial observational evidence on scales larger than individual galaxies.
+This work aims to contribute to the search for dark matter on sub-galactic scales, specifically within the Milky Way.
+First proposed by [Author Name] in [Year], the concept of dark matter arose to explain the unusually high velocities of stars in dwarf galaxies—velocities that exceed those predicted by the visible mass exerting gravitational forces.
+Dark matter's influence extends to cosmological scales as well, as demonstrated by the bending of light within cosmic web filaments due to the unseen mass.
+Despite significant progress, the true nature of dark matter remains elusive, and scientists continue to search for evidence of its existence beyond its gravitational effects on cosmic and extragalactic scales.
+A key focus today is detecting dark matter on sub-galactic scales, such as through the study of stellar streams.
 
-  Stellar streams, which are long, thin structures formed from the tidal disruption of globular clusters or dwarf galaxies as they orbit a host galaxy, offer a promising avenue for probing the distribution of dark matter.
-Tidal forces arise because these stellar systems are extended objects, leading to a differential gravitational pull across their structure.
-As a result, stars farther from the galactic center lag behind the cluster, while those closer are pulled away.
-This process stretches the cluster in the radial direction, creating two tidal tails that trace the orbit of the cluster.
-Theoretical predictions of this phenomenon have long existed, but it was not until the early 1995 that streams were observed for the first time \citet{carl_j_grillmair_globular_1995}.
-Then, \citet{odenkirchen_detection_2001} discovered that stars beyond the tidal radius of the Palomar 5 cluster shared the same spectroscopic characteristics as those within it.
+  Stellar streams, long and thin structures formed by the tidal disruption of globular clusters or dwarf galaxies orbiting a host galaxy, present a promising method for probing dark matter distribution.
+These tidal forces arise due to differential gravitational pulls across extended objects, causing stars farther from the galactic center to lag behind while those closer are pulled away.
+This stretching creates two tidal tails that trace the cluster’s orbit.
+Theoretical predictions of this phenomenon existed long before the first observation in 1995 by \citet{carl_j_grillmair_globular_1995}, who detected stars beyond the tidal radii of globular clusters.
+Subsequent work by \citet{odenkirchen_detection_2001} further demonstrated that Palomar 5’s stars not only extended beyond the tidal radius but also formed distinct tidal tails.
 
-  Over the past two decades, both theoretical and numerical studies have demonstrated the utility of stellar streams in constraining the gravitational field of host galaxies, particularly the Milky Way.
-Galactic dynamics, governed by Poisson's equation, allow the decomposition of a galaxy into components such as the central bulge, disk, and dark matter halo.
-The dark matter halo, often modeled as spherical, is the largest component and plays a critical role in the observed constant rotational velocity of galaxies.
-However, determining whether the halo might be an oblate spheroid could provide insight into the self-interaction properties of dark matter.
-Numerical studies, such as those by \citet{varghese2011stellar} who explored the extent to which stellar streams can constrain the galactic potential under varying levels of observational data.
+  Over the past two decades, theoretical and numerical studies have shown that stellar streams can constrain the gravitational field of host galaxies, particularly the Milky Way.
+By applying Poisson’s equation, the density distribution and gravitational potential can be separated into different components, including the dark matter halo.
+This halo, often modeled as spherical, is the largest component and crucial to explaining the constant rotational velocity of galaxies.
+For example, \citet{varghese2011stellar} demonstrated how stellar streams can infer the mass and scale parameters of haloes using various amounts of observational data, from basic right ascension and declination to full six-dimensional phase space information.
 Meanwhile, Bonaca et al.
-approached stellar streams from an information-theoretic perspective, identifying the orbits and configurations that yield the most information about the galactic potential.
+employed an information-theoretic approach, identifying the orbits and configurations of stellar streams that yield the most information about the galactic potential.
 
-  The release of Gaia Data Release 2 (DR2) and DR3 significantly advanced the observational study of stellar streams.
-The StreamFinder algorithm has identified 90 stellar streams, with some of these candidates being followed up by other instruments and surveys.
-For instance, streams like GD-1 and Palomar 5 have approximately 1,000 to 3,000 currently associated members, although there is an observational bias toward brighter stars, such as red giants.
+  The Gaia Data Release 2 (DR2) and DR3 revolutionized the observational study of stellar streams, enabling significant advances in the field.
+Malhan et al.
+developed the StreamFinder algorithm, which identified 90 stellar streams within the Milky Way, some of which have been followed up by other instruments and surveys.
+The best-observed streams, such as GD-1 and Palomar 5, contain between 1,000 and 3,000 stars.
+Leveraging this dataset, Ibata et al.
+(2024) constructed the most accurate model of the Milky Way to date based on stellar stream data.
 
-  The goal of this study is to contribute to the ongoing effort to use stellar streams as probes for dark matter on sub-galactic scales.
-According to the cold dark matter model, which is supported by observations, galaxies form hierarchically.
-As shown in Springel et al 2008 in the suite of Acquarious simulations, $\lambda$-CDM predicts a distribution of masses of dark matter subhalos that follows a power-law.
-The authors also demonstrate that subhaloes should exist within other haloes.
-To my knowledge, thanks to the introduction of Bonaca et al 2019 who pointed this paper out to me, the smallest dark matter sub halo detected to date was found within an enstien ring from gravitational lensing by Vigetti et al 20XX.
+  In this study, we build on this body of work by investigating another important use case for stellar streams: probing dark matter on sub-galactic scales.
+According to simulations by Springel et al.
+(2008), $\Lambda$-Cold Dark Matter ($\Lambda$CDM) predicts that galaxies grow hierarchically, with dark matter clumps forming at a wide range of masses and sizes.
+These clumps, or \textit{subhalos}, are predicted to follow a mass distribution with a power-law slope slightly shallower than -2.0.
+To date, the smallest observed dark matter halo was detected through gravitational lensing in an Einstein ring, as described by Vingetti et al.
+(20XX).
+However, some models predict that dark matter clumps could exist down to the mass of a planet like Jupiter (I hear this at conferences, is it true?
+where are the papers that suggest this?).
 
-  Over the years, people began to think of the role that subhalos can have on effecting stellar streams.
-\citet{rodrigo_ibata_uncovering_2002} suggested that the presence of said dark matter subhaloes dark matter subhalos can influence the stellar streams by diffussing their distribution of orbital elements.
-\citet{r_g_carlberg_pal_2012} extended this idea by noting that subhaloes could have flybys and that leave gaps in stellar streams and subsquently explored the statistical distribution the number of gaps gaps given a probabilistic model of the rate of encounters given based on the expected distribution of dark matter subhalos.
-\citet{bonaca_spur_2018} took a more direct observational approach, identifying an underdensity in the GD-1 stream that could not be explained by known objects like globular clusters and was inconsistent with the mass-size relationship predicted by the Lambda-CDM model.
+  \citet{rodrigo_ibata_uncovering_2002} first suggested that dark matter subhalos could influence stellar streams by diffusing their orbital elements.
+Later, \citet{r_g_carlberg_pal_2012} expanded this idea, proposing that subhalos might create gaps in stellar streams during flyby encounters, where a subhalo approaches closely enough to a segment of a stream and significantly changes the orbits of the closest stars.
+This theory was further explored through a probabilistic model of encounter rates based on the expected subhalo distribution.
+\citet{bonaca_spur_2018} provided observational evidence for this idea, identifying an underdensity in the GD-1 stream that could not be explained by known objects, such as globular clusters, and, interestingly enough, was inconsistent with the $\Lambda$CDM mass-size relationship from M et al 20XX.
 
-  In this study, we aim to investigate the potential influence of globular clusters on stellar streams, specifically examining whether galactic globular clusters can produce gaps similar to those expected from dark matter subhalos.
-If globular clusters do create significant gaps, this must be considered when using stellar streams to infer the local presence of dark matter.
-Our focus is on the Palomar 5 stream, which has been well-studied and has observable gaps, a high membership count, and favorable visibility due to its position above the galactic disk.
-We show a result from one our simulations in \ref{fig:stream_on_sky}.
-In the following sections, we describe the modeling of this stream, the gravitational interactions with the galaxy and other globular clusters, and the statistical analysis of the perturbations affecting the stream.
+  Our study focuses on the influence of globular clusters on stellar streams, specifically examining whether globular clusters could produce gaps similar to those expected from dark matter subhalos.
+If globular clusters do indeed create such gaps, this factor must be considered when using stellar streams to detect dark matter.
+An additional and unique aspect of our study is that we focus on modeling real stellar streams and not just pure simulated streams.
+We focus on the Palomar 5 stream, which is well-studied, has observable gaps (bonaca et al 20XX), and is highly visible due to its position above the galactic disk.
+We present a result from one of our simulations in \ref{fig:stream_on_sky}, and the following sections describe our modeling of the stream, the gravitational interactions with the galaxy and other globular clusters, and the statistical analysis of perturbations affecting the stream.
 

### 2024-09-23 paper/main.tex
updated the numerical methodology

@@
     \includegraphics[width=\linewidth]{stream_on_sky_Pal5_monte-carlo-009_pouliasis2017pii-GCNBody.png}
-    \caption[]{A model of the Palomar 5 stream in an axis-symmetric galactic potential plus all other globular clusters.
+    \caption{An instsance of the Palomar 5 stream created modeling the cluster as a plummer sphere within an axis-symmetric galactic potential plus 164 other galactic globular clusters.
 The top plot is a scatter plot of star-particles that escaped the cluster due to tidal forces, the bottom plot shows the marginalized 1D density profile in longitude.
-Two large gaps are present and are due to the passage of two globular clusters.
-}
+Two large gaps are present and are due to the passage of two globular clusters.}
     \label{fig:stream_on_sky}
@@
 (2023), which we briefly summarize here for completeness.
-We begin by extracting the phase space properties (positions, velocities) as well as the masses and half-mass radii of 165 globular clusters from the dataset provided by Baumgardt and Vasiliev et al.
-The dataset includes the clusters' masses, radial velocities, distances, proper motions in RA and DEC, and their associated covariance terms.
+We begin by extracting the phase space properties (positions, velocities) as well as the masses and half-mass radii of 165 globular clusters from the galactic globular cluster catalog by Baumgardt and Vasiliev et al.
+We handle the uncertainties through a monte-carlo approach.
+The catalog provides uncertainties for distances, radial velocities, proper motions, and mass, and a covariance term between the proper motions.
+We repeate the whole experiment 50 times, for each of which we vary the initial conditions of each cluster by sampling its five-dimensional gaussian uncertaintiy distribution.
+However, for the first simulation, which we name \texttt{Sampling 000}, we use the most probable values for the initial conditions.
 
-    To generate initial conditions, we treat the four phase space variables with uncertainties (distance, line-of-sight velocities, and proper motions) and mass for each globular cluster as five-dimensional Gaussian distributions.
-Using the \texttt{np.random.gauss} function, we generate 50 samples per cluster, producing 50 sets of initial conditions for the entire globular cluster system.
-These samples are stored in data files for later use.
-
-    Next, we convert the initial conditions into a galactocentric reference frame.
-The local standard of rest was computed using XXX, with the Sun's peculiar velocity set to XXX, and the Sun's distance from the Galactic center taken as XXX.
+    We convert the initial conditions from sky coordinates into a galactocentric reference frame.
+We use the local standard of rest and the peculiary motion of the sun as reported by Sch\"onrich 2010.
+The circular velocity at the solar radius was set to: $v_{\text{LSR}} = 240$~km~s$^{-1}$.
+The peculiar velocity of the sun was set to $(U_\odot, V_\odot, W_\odot)=(11.1, 12.24, 7.25)$~km~s$^{-1}$.
+The position of the Sun was set to $(x_\odot,y_\odot,z_\odot) = (-8.34,0,0.027)$~kpc, vertical position above the disk was taken from Chen et al 2001 and the galacto-radial distance was taken from Reid et al 2014.
 These transformations were performed using the \texttt{astropy} package.
+(for the next paper, I would also like to sample these uncertainties)
 
@@
 (2017), which consists of a superposition of a thin disk, thick disk, and a dark matter halo.
-This model is assumed to be time-independent over the course of our simulations.
+The values are provided in Table 1 of Ferrone et al 2023.
+This model is time-independent over the course of our simulations.
 
@@
     Although these interactions were included, we observed no significant exchanges of orbital energy or momentum between clusters.
-The resulting orbits were saved to data files for subsequent analysis.
-We integrate clusters into the past by 5 Gyr.
 
     At the computed positions 5 Gyr ago, each globular cluster was gifted a system of 100,000 particles, sampled from a Plummer distribution to represent the stellar content of the cluster.
+For eaching sampling of the initial conditions, a different sampling of the positions of the 100,000 star particles was performed.
+This is becuase, we sampled the uncertainties on the mass in this experiment contraty to Ferrone et al 2023.
+
     We then integrated the evolution of these particles forward in time to the present day.
 During this forward integration, each star particle feels the gravitational influence of the galaxy and all globular clusters, but not other star particles.
-We do not perform direct N-body simulations between star particles as in eq.~\ref{eq:GCNBody}; instead, at each time step, we compute the positions of all globular clusters and apply their forces to the star particles.
+We do not perform direct N-body simulations between star particles as in eq.~\ref{eq:GCNBody}; instead, at each time step, we load the positions of all globular clusters, as found from the initial integration, and apply their forces to the star particles.
 The equation of motion is thus:
     \begin{equation}
-      \ddot{\vec{r}}_p = -\nabla \Phi(R_p,z_p) + \left.\sum_{j} \frac{Gm_j}{\left(|\vec{r}_j(t) - \vec{r}_p|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j(t)- \vec{r}_p\right)
+      \ddot{\vec{r}}_p = -\nabla \Phi(R_p,z_p) + \left.\sum_{j} \frac{Gm_j}{\left(|\vec{r}_j(t) - \vec{r}_p|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j(t)- \vec{r}_p\right).
       \end{equation}
@@
     During the integration, intermediate snapshots were saved to facilitate the analysis of stellar streams and the effects of cluster impacts.
-Specifically, for each of the 50 realizations of the Palomar 5 stream, we saved 5,000 intermediate timesteps, resulting in approximately XX GB of data ($N_p \times N_s \times 7 \times \text{float}$).
+Specifically, for each of the 50 realizations of the Palomar 5 stream, we saved 5,000 intermediate timesteps which is a temporal resolution of 1 million years.
+To provide an estimate of the size of our simulations, we have $N_p=100 000$, $N_\textrm{time stamps} = 5000$, $N_\textrm{phase}=6$, $N_\textrm{sampling}=50$, and using the data type as a single precision floating point number $\textrm{Single}=4$~bytes results in approximately 600 GB of data.
 

### 2024-09-23 paper/main.tex
touched up detections

@@
     \subsubsection*{Density Maps and Profiles}
-      For each of the 50 simulations, we compare the final snapshots between the \texttt{vanilla} and \textit{full} simulations using the tail coordinate system.
+      For each of the 50 simulations, we compare the final snapshots between the \texttt{vanilla} and \texttt{full} simulations using the tail coordinate system.
 We inspect these differences by generating 2D density maps of the streams.
@@
         \includegraphics[width=\linewidth, trim=20 0 15 0]{monte-carlo-009-pouliasis2017pii-GCNBody-2000-milisigma-5-noisefactor-20-boxcarindexlength-shifted-0.png}
-        \caption{A comparison between the density maps and profiles of the \textit{full} and \textit{vanilla} simulations is presented.
+        \caption{A comparison between the density maps and profiles of the \texttt{full} and \texttt{vanilla} simulations is presented.
 The locations where the full simulation is less dense than the vanilla simulation by 2-$\sigma$ is highlighted by blue vertical bars.
@@
 
-      To construct the 1D density profiles, we first bin the data using the $\sqrt{N}$ rule, where $N$ is the number of data points ($N = 10^5$).
+      To construct the 1D density profiles, we first bin the data using the $\sqrt{N}$ rule, where $N$ is the number of data points ($N_p = 10^5$).
 After binning the 1D profiles, we apply a boxcar smoothing technique.
 Each central value is replaced by the median of the surrounding XX adjacent data points, corresponding to a smoothing length of approximately XX parsecs.
-This procedure reduces high-frequency noise and smooths the profiles, effectively removing the center of mass (see Fig.~\ref{fig:profiles}).
+This procedure reduces high-frequency noise and smooths the profiles.
+For instance, notice the absence of a high mass peak indicating for the center of mass in the bottom panel of Fig.~\ref{fig:profiles}.
 
-      With the smoothed 1D density profiles in hand, we treat the \textit{vanilla} simulations as the reference model and search for regions in the \textit{full} simulations that are significantly underdense compared to \textit{vanilla}, surpassing stochastic fluctuations.
+      With the smoothed 1D density profiles in hand, we treat the \texttt{vanilla} simulations as the reference model and search for regions in the \texttt{full} simulations that are significantly underdense compared to \texttt{vanilla}, surpassing stochastic fluctuations.
 To accomplish this, we impose a signal-to-noise ratio threshold, $\mathcal{SNR}$.
-The signal is defined as the log of the counts per bin from the \textit{vanilla} 1D density profile, and errors are propagated assuming a Poisson distribution.
-We then compute a threshold for the number of counts in the \textit{vanilla} simulation, $N_v$, using the transcendental equation:
+The signal is defined as the log of the counts per bin from the \texttt{vanilla} 1D density profile, and errors are propagated assuming a Poisson distribution.
+We then compute a threshold for the number of counts in the \texttt{vanilla} simulation, $N_v$, using the transcendental equation:
       \begin{equation}
@@
       By setting $\mathcal{SNR} = 5$, we solve for $N$ using \texttt{scipy.optimize.fsolve}, finding that $N$ must be greater than 7.
-After discarding insignificant bins (i.e., those with counts below the threshold), we compute the log ratio of the counts between the \textit{vanilla} and full simulations:
+After discarding insignificant bins (i.e., those with counts below the threshold), we compute the log ratio of the counts between the \texttt{vanilla} and full simulations:
       \begin{equation}
@@
         \end{equation}
-      where $\mathcal{R}_i$ is the log ratio, $N_{f,i}$ are the counts from the full simulation, and $N_{v,i}$ are the counts from the \textit{vanilla} simulation for each bin $i$.
+      where $\mathcal{R}_i$ is the log ratio, $N_{f,i}$ are the counts from the full simulation, and $N_{v,i}$ are the counts from the \texttt{vanilla} simulation for each bin $i$.
 We then analyze the distribution of $\mathcal{R}_i$.
@@
 (2015), a gap's size may not indicate a weak impact, but rather a more recent one.
-Additionally, since our streams have finite width, some gaps might be oblique to the stream axis.
-In such cases, marginalizing over $y'$ can dilute the gap's signal, making it harder to detect.
+Additionally, since our streams have finite width, some gaps are oblique with respect to the stream axis.
+In such cases, marginalizing over $y'$ erases the gap's signal, making it impossible to detect in a 1D profile.
 This limitation is particularly evident in gaps caused by NGC2808, as discussed in the results.
 Therefore, this quantitative analysis serves as an \textit{aid} to visual inspection rather than a complete substitute for it.
+This method helps particularly with large subtle gaps that the eye does not notice in the 2D maps.
 
@@
 
-  We then transformed the stream to tail coordinates for the both the \textit{vanilla} and \textit{full} simulations in Fig.~\ref{fig:profiles}.
+  We then transformed the stream to tail coordinates for the both the \texttt{vanilla} and \texttt{full} simulations in Fig.~\ref{fig:profiles}.
 As described in section~\ref{sec:gap_detection}, we locate regions on the stream where the full simulations are significantly less dense than the vanilla simulations.

### 2024-09-23 paper/main.tex
added box car smoothing information

@@
 After binning the 1D profiles, we apply a boxcar smoothing technique.
-Each central value is replaced by the median of the surrounding XX adjacent data points, corresponding to a smoothing length of approximately XX parsecs.
+The density profiles are scanned.
+At each bin, a selected number of adjacent data points from both sides are placed in a list.
+The bin is then replaced with the median.
+We choose to use 10 adjacent points per side, which corresponding to a smoothing length of approximately 1~kpc.
 This procedure reduces high-frequency noise and smooths the profiles.
@@
       With the smoothed 1D density profiles in hand, we treat the \texttt{vanilla} simulations as the reference model and search for regions in the \texttt{full} simulations that are significantly underdense compared to \texttt{vanilla}, surpassing stochastic fluctuations.
-To accomplish this, we impose a signal-to-noise ratio threshold, $\mathcal{SNR}$.
+To accomplish this, we first impose a signal-to-noise ratio threshold, $\mathcal{SNR}$.
 The signal is defined as the log of the counts per bin from the \texttt{vanilla} 1D density profile, and errors are propagated assuming a Poisson distribution.

### 2024-09-23 paper/main.tex
editted captions and such

@@
 These tidal forces arise due to differential gravitational pulls across extended objects, causing stars farther from the galactic center to lag behind while those closer are pulled away.
-This stretching creates two tidal tails that trace the cluster’s orbit.
+This stretching creates two tidal tails that trace the cluster's orbit.
 Theoretical predictions of this phenomenon existed long before the first observation in 1995 by \citet{carl_j_grillmair_globular_1995}, who detected stars beyond the tidal radii of globular clusters.
-Subsequent work by \citet{odenkirchen_detection_2001} further demonstrated that Palomar 5’s stars not only extended beyond the tidal radius but also formed distinct tidal tails.
+Subsequent work by \citet{odenkirchen_detection_2001} further demonstrated that Palomar 5's stars not only extended beyond the tidal radius but also formed distinct tidal tails.
 
   Over the past two decades, theoretical and numerical studies have shown that stellar streams can constrain the gravitational field of host galaxies, particularly the Milky Way.
-By applying Poisson’s equation, the density distribution and gravitational potential can be separated into different components, including the dark matter halo.
+By applying Poisson's equation, the density distribution and gravitational potential can be separated into different components, including the dark matter halo.
 This halo, often modeled as spherical, is the largest component and crucial to explaining the constant rotational velocity of galaxies.
@@
     Our numerical methodology follows the approach outlined in Ferrone et al.
-(2023), which we briefly summarize here for completeness.
+(2023), which we present here for completeness.
 We begin by extracting the phase space properties (positions, velocities) as well as the masses and half-mass radii of 165 globular clusters from the galactic globular cluster catalog by Baumgardt and Vasiliev et al.
 We handle the uncertainties through a monte-carlo approach.
-The catalog provides uncertainties for distances, radial velocities, proper motions, and mass, and a covariance term between the proper motions.
+The catalog provides uncertainties for mass, distances, line of sight velocities, proper motions, with a covariance term between the proper motions.
 We repeate the whole experiment 50 times, for each of which we vary the initial conditions of each cluster by sampling its five-dimensional gaussian uncertaintiy distribution.
@@
 (2017), which consists of a superposition of a thin disk, thick disk, and a dark matter halo.
-The values are provided in Table 1 of Ferrone et al 2023.
+The masses and scale lengths are provided in Table 1 of Ferrone et al 2023.
 This model is time-independent over the course of our simulations.
@@
       \ddot{\vec{r}}_p = -\nabla \Phi(R_p,z_p) + \left.\sum_{j} \frac{Gm_j}{\left(|\vec{r}_j(t) - \vec{r}_p|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j(t)- \vec{r}_p\right).
-      \end{equation}
+      \end{equation} \label{eq:equation_of_motion_particle}
 
@@
     During the integration, intermediate snapshots were saved to facilitate the analysis of stellar streams and the effects of cluster impacts.
-Specifically, for each of the 50 realizations of the Palomar 5 stream, we saved 5,000 intermediate timesteps which is a temporal resolution of 1 million years.
-To provide an estimate of the size of our simulations, we have $N_p=100 000$, $N_\textrm{time stamps} = 5000$, $N_\textrm{phase}=6$, $N_\textrm{sampling}=50$, and using the data type as a single precision floating point number $\textrm{Single}=4$~bytes results in approximately 600 GB of data.
+Specifically, for each of the 50 realizations of the Palomar 5 stream, we saved $5000$ intermediate timesteps which is a temporal resolution of 1 million years.
+To provide an estimate of the size of our simulations, we have $N_p=100 000$, $N_\textrm{time stamps} = 5 000$, $N_\textrm{phase}=6$, $N_\textrm{sampling}=50$, and using single precision floating point numbers, $\textrm{Single}=4$~bytes, results in approximately 600 GB of data.
 
@@
 As outlined by Erkal et al.
-(2015), a gap's size may not indicate a weak impact, but rather a more recent one.
+(2015), a small gap is indicative of a weak impact, but rather a more recent one.
+This is because gap growth is a dispersion phenomenon.
 Additionally, since our streams have finite width, some gaps are oblique with respect to the stream axis.
 In such cases, marginalizing over $y'$ erases the gap's signal, making it impossible to detect in a 1D profile.
-This limitation is particularly evident in gaps caused by NGC2808, as discussed in the results.
+This limitation is particularly evident in gaps caused by NGC~2808, as discussed in the results.
 Therefore, this quantitative analysis serves as an \textit{aid} to visual inspection rather than a complete substitute for it.
@@
 To address this, we examine the evolution of stream density over time.
-Instead of using the x'coordinate, we introduce $\tau$, which represents time rather than distance.
+Instead of using the x'-coordinate, we introduce $\tau$, which represents time rather than distance.
 Specifically, $\tau$ indicates how long it will take for a cluster to reach, or how long ago it passed, a given point on its orbit.
@@
 
+      \begin{figure*}
+        \centering
+        \includegraphics[width=\linewidth]{force_on_orbit-monte-carlo-009.png}
+        \caption{This figure demonstrates how we determined which globular clusters were responsible for the gaps.
+The y-axis is $\tau$, which is a coordinate in units of time that indicates how far ahead or behind it is from a globular cluster.
+The x-axis is the simulation time, $t$, where 0~s km~kpc$^{-1}$ indicates present time.
+The bottom plot showcases the evolution of the stream density in simulation time.
+The density was used to determine a sutible length of the stream as a function of time.
+This length was then used to extract a piece of Palomar 5's orbit about a given position of Pal 5 in time.
+This orbital segment is used to approximate the stream.
+Then, the gravitational force from all other clusers was computed on the orbit.
+This is shown in the top plot, where the gravitational force is in measured in acceleration and is given in integration units: km$^2$~kpc$^{-1}$~s$^{-2}$.
+Moments of high acceleration indicate the passage of another cluster.
+The top 5 strongest passages are labeled with red circles as well as the name of the clusters.
+The example shown in this plot is the same simulation as Fig.~\ref{fig:stream_on_sky}.}
+        \label{fig:force-on-orbit}
+        \end{figure*}
+
       We then used \texttt{scipy}'s \texttt{ndimage} package to identify the top five local maxima in the data space of gravitational acceleration $\vec{g}$ as a function of time $t$ and the stream coordinate $\tau$.
+TheThis is done by first smoothing the image by taking a 5 point moving average kernel.
+Secondly, we use maximum filter to locate coordinates in the $t,\tau$ data plane who are local maxima to at least 10-adjacent data points.
+These locations are then ordered and the top five strongest interactions are saved.
 Once these local maxima were identified, we iterated over the contributions of individual globular clusters to determine which cluster contributed the most to each peak in $\vec{g}$.
@@
 
-      Afterward, we cross-referenced these peaks with the previously identified gaps in the stream.
-For large gaps resulting from strong interactions, we observed that after an impact, a low-density wake is left behind in the ($t,\tau$) plane.
+      Afterward, we cross-referenced these peaks with the locations of the gaps identified by studying the density maps and profiles from Fig.~\ref{fig:profiles}.
+For large gaps resulting from strong interactions, we observed that after an impact, a low-density wake is left behind in the ($t,\tau$) plane, which can be seen corresponding to the impacts of NGC~104 and NGC~7808 in the bottom panel of Fig.~\ref{fig:force-on-orbit}.
 
@@
 
+      \begin{figure*}
+        \centering
+        \includegraphics[width=\linewidth]{decomposition-monte-carlo-009-with-3-gaps.png}
+        \caption{This figure decomposes the gaps shown in the middle panel of Fig.~\ref{fig:profiles} into the individual perturbuers whose cause them, as identified from Fig.~\ref{fig:force-on-orbit}.
+These simulations only include the suspected globular clusters in one-at-a-time.}
+        \label{fig:decomposition}
+        \end{figure*}
+
   \subsection{Reconstruction of the impact geometry}
-    In its simplest form, impact theory, as presented in works such as Binney and Tremaine, John Taylor, and Galaxies, considers two particles: one stationary and the other moving past it.
+
+    With the perturbers identified, we perform statistics in the pursuit of understanding what conditions are necessary for a globular cluster and a stream to have in order for a gap to form.
+We turn to impact theory, which in its simplest form as presented in works such as Binney and Tremaine, John Taylor, and Galaxiesbook, considers two particles: one stationary and the other moving past it.
 The distance between the two particles at their closest approach is known as the impact parameter.
@@
 Then, we refine these estimates.
-While the orbit serves as a good first-order proxy for the stream, it is insufficient, as the typical offset between the stream and its orbit can be on the order of 200-400 pc—significantly larger than the impact parameters we are investigating.
+While the orbit serves as a good first-order proxy for the stream, it is insufficient, as the typical offset between the stream and its orbit can be on the order of 200-400 pc—--significantly larger than the impact parameters we are investigating.
 
@@
       \end{equation}  
-    where $x$, $y$, and $z$ represent the parametric line describing the stream in galactocentric coordinates, $\tau$ is the stream coordinate in time as described in the previous section, and is used as the independent variable in this analysis.
+    where $x$, $y$, and $z$ represent the parametric line describing the stream in galactocentric coordinates, $\tau$ is the stream coordinate in time as described in the previous section, and is used as the independent variable to parameterize the motion.
 The coefficients $a_i$, $b_i$, and $c_i$ are the polynomial coefficients.
@@
     In this analysis, only one side of the stream is considered.
-For instance, if the impact candidate was in the leading tail, only the star particles with $\tau > 0$ are used to constrain the stream.
-The polynomial coefficients were determined using the Nelder-Mead algorithm from \texttt{scipy}'s optimization package.
+For instance, if the impact candidate was in the leading tail, only the star particles with $\tau > 0$ are used to constrain the stream track.
+The polynomial coefficients were determined using the through a minimization method using Nelder-Mead algorithm from \texttt{scipy}'s optimization package.
 
-    Since the simulation snapshots were saved at a temporal resolution of 1 Myr—rather than at the integration timestep of $10^4$ years, which would generate an excessive amount of data—we interpolate between snapshots to more precisely estimate the impact geometry.
+    Since the simulation snapshots were saved at a temporal resolution of 1 Myr--—rather than at the integration timestep, which would have generated an excessive amount of data—--we interpolate between snapshots to more precisely estimate the impact geometry.
 This is achieved by fitting the stream at five timesteps surrounding the approximate impact time, a period of 5 Myr, which sufficiently covers the interaction time.
@@
 I wanna compare this to the known length of Palomar 5, how does this reduce the results, and are there observed gaps in Palomar 5 that are consistent with these simulations?
-
-  \begin{figure*}
-    \centering
-    \includegraphics[width=\linewidth]{force_on_orbit-monte-carlo-009.png}
-    \caption[]{This figure demonstrates how we determined which globular clusters were responsible for the gaps.
-The bottom plot showcases the evolution of the stream density in simulation time.
-The y-axis is $\tau$, which is a coordinate in units of time that indicates how far ahead or behind it is from a globular cluster.
-The x-axis is the simulation time, $t$, where 0~s km~kpc$^{-1}$ indicates present time.
-The density was used to determine a sutible length of the stream as a function of time.
-This length was then used to extract a piece of Palomar 5's orbit about a given position of Pal 5 in time.
-This orbital segment is used to approximate the stream.
-Then, the gravitational force from all other clusers was computed on the orbit.
-Moments of high acceleration indicate the passage of another cluster.
-The top 5 strongest passages are labeled with red circles as well as the name of the clusters.
-}
-    \label{fig:force-on-orbit}
-    \end{figure*}
-
-  \begin{figure*}
-    \centering
-    \includegraphics[width=\linewidth]{decomposition-monte-carlo-009-with-3-gaps.png}
-    \caption[]{This figure decomposes the gaps found in the full simulation to the responsible perturbers.
-The pertuber candidates were found in Fig.~\ref{fig:profiles}, and extra simulations were run in which we singularly add the suspects to verify that it was indeed them who created the gaps.
-The titles explain which simulation is which, and there's also a descrepandy map.
-(I think I'm going to get rid of the DIFF map)}
-    \label{fig:decomposition}
-    \end{figure*}
 

### 2024-09-23 paper/main.tex
laid out the results

@@
 
-  One of the 50 simulations is shown in Fig.~\ref{fig:stream_on_sky}--which we selected for its clear and prominent gaps which is visibile when considering the scatter plot of the stars on the sky and becomes even more apparent when marginalizing over latitutde to see it's 1D density over longitude.
+  One of the 50 simulations is shown in Fig.~\ref{fig:stream_on_sky}--which we selected for its clear and prominent gaps which are visibile in the scatter plot of the stars on the sky and becomes even more apparent when marginalizing over latitutde to see it's 1D density over longitude.
 
@@
 All simulations and labeled gaps are presented in the Appendix~\ref{appendix:gallery_of_gaps}.
-The profile method was useful in identifying the gaps in sampling 002, 006, 008, 029, 030.
+The profile method was useful in identifying the subtle underdense gaps of Samplings: 000, 002, 006, 008, 029, 030.
 
-  I'm gonna go remake the figures before I adapt the text.
-However, before that, I will say the general outline of the flow of the article.
-I will show a general result here, which is what we would expect to see on the sky, which is comparable with reality.
-It's a great first figure becuase it shows the essence of the paper.
-Then, I pretty much follow the methodology and keep the story the same with just monte-carlo-009.
-The story is we need to understand how many gaps are present in a given simulation, then we need to identidy who the perturbers are.
-So step 1 (fig 2) shows our aided eye-test.
-Fig~3 shows how I tried to find each perturber, and Fig~4 shows a sanity check to reassure me that the suspects are indeed guilty.
-However, I will say that we picked monte-carlo-009 because it has two obvious gaps and two subtle ones, which aid in communicating our point.
-A full display of the gaps in our simulations is shown in the appendix (i'll make pair-plots with vanilla and full simulations above and below one another with the gaps labeled.
-:D )
+  Besides gaps, a handful of simulations show the entire stream laterally shifted in tail coordinates.
+The most extreme case was observed in Sampling 006 and is presented in the appendix XXX.
+This of course hampered our 1D profile difference analysis whose based the assumption that any difference in density would be due to the passage of other globular clusters and thus under densities.
+To combat this, we performed a cosmetique data reduction where I moved the \texttt{Full} 1D profile to best overlap with \texttt{Vanilla} by computing the correlation integral.
+Since this is not the focus of this study, we did not investigate further.
+We conjecture that the translation may arrise from a resonance phenomenon, most likely coming from Omega Centauri.
+Perhaps, in some orbital configurations Palomar 5 and Omega Centauri pass each other in regular intervals that can accelerate the globular cluster differently from the stream.
+
+  The clusters that impacted Palomar 5 in our simulations are presented in Fig.~\ref{fig:histogram_impact_time}, along with when they impacted the stream. An imediate result is
+
+      - It takes a while for the stream to grow to hit a significant length
+      - A lot of them are from NGC2808. This is is because it is very recent. 
+      - Of the recent hits, the gaps like NGC2808 have widths as small as 100 parsecs.
+      - of the morphology, there are some gaps that have "bites" removed from them. Like that of NGC6333 in Sampling 033. 
+      - I think it is important to keep in mind that many of said gaps would not be detectable with current
+
+  \begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{histogram_impact_time.png}
+    \caption{When the impacts occured}
+    \label{fig:histogram_impact_time}
+    \end{figure*}
+
+    in terms of statistics: 
+    - We observe that only clusters above 1e5 cause gaps
+
+    - to no surprise, you have to come within Palomar 5's orbital space
+
+    - trying to separate to the components of the velocity, Mass, impact parameter is not fruitful. It's better just to look at the change in momentum as given in the EQ.
+
+    - Notice that the distributions between the gap and no gap are significantly different, yet also have significant overlap.
+Perhaps these 3 parameters don't tell the whole story but also there's the factor identifies in Sanders, that I don't understand yet, that shows that a gap will not be identifiable.
+
+  \begin{figure}
+    \centering
+    \includegraphics[width=1\linewidth]{E_Lz_perturbers.png}
+    \includegraphics[width=1\linewidth]{mass_size_plane.png}
+    \includegraphics[width=1\linewidth]{impact_geometry_statistics_deltaP.png}
+    \caption{One plot shows the distribution in energy and angular momentum, the next are the inernal proeprties of the clusters, and the last show when the clusters commit their crime, and how many times it happens accross the 50 simulations.
+}
+    \label{fig:mass_size_plane}
+    \end{figure}
 
@@
 
-  \begin{figure}
-    \centering
-    \includegraphics[width=1\linewidth]{E_Lz_perturbers.png}
-    \includegraphics[width=1\linewidth]{mass_size_plane.png}
-    \includegraphics[width=1\linewidth]{impact_geometry_statistics_deltaP.png}
-    \caption{One plot shows the distribution in energy and angular momentum, the next are the inernal proeprties of the clusters, and the last show when the clusters commit their crime, and how many times it happens accross the 50 simulations.
-}
-    \label{fig:mass_size_plane}
-    \end{figure}
-  \begin{figure*}
-    \centering
-    \includegraphics[width=\linewidth]{histogram_impact_time.png}
-    \caption{When the impacts occured}
-    \label{fig:histogram_impact_time}
-    \end{figure*}
+  \section{Discussion}
+    How many of these could be observable?
+What's the main result?
+in 5 billion years of evolution, perhaps you'd only expect to 1.4 gaps, reaching up to maximum three?
+However, its possible that we dont see gaps at all for sure.
 
-  \section{Discussion}
-    $10^5-10^6$. Sotto alone averbbero questo effetto?
+    What about drawing attention to future studies?
+
+    Doing a complete census of all globular clusters?
+
+    Doing a complete census of dark matter subhaloes?
+
+    How many gaps do you expect to see in Palomar 5? Bonaca said there are about 2 of them.
 

### 2024-09-26 stream_analysis/identify_suspects.py
added variable number of peaks to overplot

@@ -110,7 +110,7 @@ def doplot(FIGSTUFF,DATASTUFF,PROPERTIESSTUFF,x_text_shift=-0.1,y_text_shift=0.0
     
     im=axis1.pcolorfast(time_foob,tau_foob,convolved_mask.T,**foobpcolor)
     cbarfoob=fig.colorbar(im,cax=caxisFOOB,label="Force on Orbit")
-    for i in range(5):
+    for i in range(len(coordinates)):
         x,y=coordinates[i]
         axis1.scatter(time_foob[x], tau_foob[y], marker="o", s=75,edgecolor="red",facecolors='none')
         axis1.text(x=time_foob[x]+x_text_shift,y=tau_foob[y]+y_text_shift,s=str(i)+" "+suspects[i],**TEXT)

### 2024-09-26 jobs/identify_suspects/main.py
many changes

@@ -47,8 +47,8 @@ def main(dataparams,hyperparams):
         return None
 
     if os.path.exists(outfilename):
-        print(outfilename, " already exists")
-        return None 
+        print(outfilename, " already exists, overwriting")
+        # return None 
 
 
     time_foob,tau_foob,mag_total,Perturbers,GCmags = sa.identify_suspects.extract_FOOB(fileFOOB)
@@ -68,22 +68,24 @@ def main(dataparams,hyperparams):
 
 
     coordinates = sa.identify_suspects.get_peaks(convolved_mask)
+    coordinates = [coordinates[i] for i in range(NKEEPS) ]
 
     time,tau,mag,convolved,suspects=sa.identify_suspects.build_output_data(coordinates,tau_foob,time_foob,mag_total,convolved_data,GCmags,Perturbers,NKEEPS)
     dframe =pd.DataFrame({'time':time,'tau':tau,'mag':mag,'convolved':convolved,'suspects':suspects})
     dframe.to_csv(outfilename,index=False)
     print("Saved to ",outfilename)
 
-
-    fig,axis1,axis2,caxisFOOB,caxisTau = sa.identify_suspects.getfigure()
-    foobNORM,tauNORM,AXIS1,AXIS2 = sa.identify_suspects.properties(time_foob,tau_left,tau_right,NP,montecarlokey)
-    FIGSTUFF=[fig,axis1,axis2,caxisFOOB,caxisTau]
-    DATASTUFF=time_foob,tau_foob,convolved_mask,suspects,coordinates,tau_counts,tau_centers,time_stamps,tau_left,tau_right
-    PROPERTIESSTUFF=foobNORM,tauNORM,AXIS1,AXIS2
-    sa.identify_suspects.doplot(FIGSTUFF,DATASTUFF,PROPERTIESSTUFF)
-    fig.savefig(fname,dpi=300) 
-    print("Saved to ",fname)
-    plt.close(fig)
+    doplot=False
+    if doplot:
+        fig,axis1,axis2,caxisFOOB,caxisTau = sa.identify_suspects.getfigure()
+        foobNORM,tauNORM,AXIS1,AXIS2,TEXT = sa.identify_suspects.properties(time_foob,tau_left,tau_right,NP,montecarlokey)
+        FIGSTUFF=[fig,axis1,axis2,caxisFOOB,caxisTau]
[…]

### 2024-09-26 jobs/parameterize_stream_impact/parameterize_and_compute_impact.py

@@ -270,11 +270,11 @@ if __name__=="__main__":
     GCname = "Pal5"
     NP = int(1e5)
     MWpotential = "pouliasis2017pii-GCNBody"
-    montecarlokey="monte-carlo-009"
+    montecarlokey="monte-carlo-014"
     internal_dynamics = "isotropic-plummer"
 
     # hyper params
-    targetnumber        = 2
+    targetnumber        = 6
     n_adjacent_points   = 5
     n_stamps            = 2*n_adjacent_points+1
     nDynTimes           = 2

### 2024-09-26 stream_analysis/identify_suspects.py

@@ -110,6 +110,10 @@ def doplot(FIGSTUFF,DATASTUFF,PROPERTIESSTUFF,x_text_shift=-0.1,y_text_shift=0.0
     
     im=axis1.pcolorfast(time_foob,tau_foob,convolved_mask.T,**foobpcolor)
     cbarfoob=fig.colorbar(im,cax=caxisFOOB,label="Force on Orbit")
+    print("LEN",len(coordinates))
+    print("len(time_foob)",time_foob.shape)
+    print("len(tau_foob)",tau_foob.shape)
+    print("len(suspects)",len(suspects))
     for i in range(len(coordinates)):
         x,y=coordinates[i]
         axis1.scatter(time_foob[x], tau_foob[y], marker="o", s=75,edgecolor="red",facecolors='none')

### 2024-09-26 jobs/parameterize_stream_impact/mp_parameterize_and_compute_impact.py
added more top hitting clusters

@@ -7,8 +7,8 @@ import parameterize_and_compute_impact as pci  #type: ignore
 def loop_over_targets(dataparams,hyperparams):
     GCname,NP,MWpotential,montecarlokey,internal_dynamics = dataparams
     _,n_adjacent_points,nDynTimes,width_factor,streampolyorder,perturberPolyOrder = hyperparams
-
-    for i in range(5):
+    NKEEPS = 8
+    for i in range(NKEEPS):
         dataparams  =   (GCname,NP,MWpotential,montecarlokey,internal_dynamics)
         hyperparams =   (i,n_adjacent_points,nDynTimes,width_factor,streampolyorder,perturberPolyOrder)
         pci.main(dataparams,hyperparams)
@@ -28,6 +28,7 @@ if __name__=="__main__":
     width_factor        = 10
     streampolyorder     = 2
     perturberPolyOrder  = 2
+    NKEEPS              = 8
 
 
     hyperparams =   (0,n_adjacent_points,nDynTimes,width_factor,streampolyorder,perturberPolyOrder)    

### 2024-09-27 jobs/identify_suspects/assemble_into_csv.py
made into its own file and not notebook

@@ -1,0 +1,52 @@
+"""
+Takes the impact geometry that was saved in a bunch of HDF5 files and assembles them into a single CSV file.
+"""
+
+from gcs import path_handler as ph
+import pandas as pd
+import h5py
+import numpy as np
+import os 
+
+
+if __name__ == '__main__':
+    # Load the data
+    GCname = "Pal5"
+    NP = int(1e5)
+    MWpotential = "pouliasis2017pii-GCNBody"
+    internal_dynamics = "isotropic-plummer"
+
+
+    outfilename = GCname + "-perturber-suspects.csv"
+
+    outpath=ph.paths['simulations'] + "ImpactGeometry/" + MWpotential + "/" 
+
+    row_list = []
+    for j in range(50):
+        montecarlokey = "monte-carlo-"+str(j).zfill(3)
+        fnameSuspects   =   ph.PerturberSuspects(GCname,MWpotential,montecarlokey)
+        suspects = pd.read_csv(fnameSuspects)
+        for i in range(len(suspects['suspects'])):
+            target_number = i
+            suspect=suspects['suspects'][target_number]
+            infile=ph.ImpactGeometry(GCname,MWpotential,montecarlokey,suspect,target_number)
+            if not os.path.exists(infile):
+                print("File not found: ",infile)
+                continue
+            with h5py.File(infile, 'r') as myfile:
+                impact_parameter=myfile['geometry']['impact_parameter'][()]
+                w_par=np.linalg.norm(myfile['geometry']['w_par'][()])
+                w_per=np.linalg.norm(myfile['geometry']['w_per'][()])
+                v_rel = np.sqrt(w_par**2 + w_per**2)
+                alpha=myfile['geometry']['alpha'][()]
+                Mass=myfile['geometry']['Mass'][()]
+                rplum=myfile['geometry']['rplum'][()]
+                T=myfile['geometry']['T'][()]
+                tau=myfile['geometry']['tau'][()]
+                gap_flag = False
+                profile_density = myfile['geometry']['profile_density'][()]
+                data_slice = {'impact_parameter':impact_parameter,'w_par':w_par,'w_per':w_per,'v_rel':v_rel,'alpha':alpha,'Mass':Mass,'rplum':rplum,'T':T,'tau':tau,"suspect":suspect,"montecarlokey":montecarlokey,"target_number":target_number,"gap_flag":gap_flag,"profile_density":profile_density}
[…]

### 2024-09-27 jobs/identify_suspects/assemble_into_csv.py
changed column order

@@ -45,7 +45,7 @@ if __name__ == '__main__':
                 tau=myfile['geometry']['tau'][()]
                 gap_flag = False
                 profile_density = myfile['geometry']['profile_density'][()]
-                data_slice = {'impact_parameter':impact_parameter,'w_par':w_par,'w_per':w_per,'v_rel':v_rel,'alpha':alpha,'Mass':Mass,'rplum':rplum,'T':T,'tau':tau,"suspect":suspect,"montecarlokey":montecarlokey,"target_number":target_number,"gap_flag":gap_flag,"profile_density":profile_density}
+                data_slice = {'impact_parameter':impact_parameter,'w_par':w_par,'w_per':w_per,'v_rel':v_rel,'alpha':alpha,'Mass':Mass,'rplum':rplum,'T':T,'tau':tau,"profile_density":profile_density,"suspect":suspect,"montecarlokey":montecarlokey,"target_number":target_number,"gap_flag":gap_flag}
             row_list.append(data_slice)
     df = pd.DataFrame(row_list)
     df.to_csv(outpath + outfilename,index=False)

### 2024-09-27 figures/decomposition.py
deleted this legacy file

@@ -1,276 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-import h5py
-import plotting_functions as PF
-import sys 
-sys.path.append("/obs/sferrone/gc-tidal-loss/code/")
-import StreamOrbitCoords as SOC
-import matplotlib.gridspec as gridspec
-from matplotlib.colors import LogNorm
-
-
-'''
-DECOMPOSITION OF MONTE CARLO 009
-
-'''
-
-############################################################
-############################################################
-##################### GLOBAL VARIABLES #####################
-############################################################
-############################################################
-
-#### PATH VARIABLES
-streampathbase="/scratch2/sferrone/simulations/Streams/"
-orbitpathbase="/scratch2/sferrone/simulations/Orbits/"
-
-##### integration variables 
-NP=100000
-montecarlokey="monte-carlo-009"
-internaldynamics = "isotropic-plummer"
-targetGC="Pal5"
-### the names of the Pal5 orbits that were used for each stream computation
-correspondingOrbits = {
-    "pouliasis2017pii": "pouliasis2017pii",
-    "pouliasis2017pii-GCNBody": "pouliasis2017pii-GCNBody/",
-    "pouliasis2017pii-NGC2808": "pouliasis2017pii-GCNBody/",
-    "pouliasis2017pii-NGC104": "pouliasis2017pii-GCNBody/",
-    "pouliasis2017pii-NGC7078": "pouliasis2017pii-GCNBody/",
-}
-
-###### PARAMS FOR PLOT COMPUTATION ######
-xmax,ymax=10,1
-currenttime=0
-potentials = ["pouliasis2017pii",
-              "pouliasis2017pii-GCNBody",
-              "DIFF",
-              "pouliasis2017pii-NGC2808",
-              "pouliasis2017pii-NGC104",
-              "pouliasis2017pii-NGC7078"]
-Nrows=len(potentials)
-
-
-
-###### PARAMS FOR PLOT ASTHETICS
-xlim,ylim=10,0.5
-diffScatterProperties = {
-    "cmap": "coolwarm_r",
-    "s": 10,
[…]

### 2024-09-27 figures/gallery_of_gaps.py

@@ -1,100 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-from matplotlib.colors import LogNorm
-import h5py
-import os
-# routines in this folder
-import plotting_functions as PF
-import gap_results as GR
-import sys 
-sys.path.append("/obs/sferrone/gc-tidal-loss/code/")
-import StreamOrbitCoords as SOC
-
-
-def main():
-    # get the parameters
-    NP,internaldynamics,potential,targetGC,xmax,ymax = default_parameters()
-    # set the paths
-    basepath="/scratch2/sferrone/simulations/Streams/"+potential+"/"+targetGC+"/"+str(NP)+"/"
-    orbitpath="/scratch2/sferrone/simulations/Orbits/"+potential+"/"+targetGC+"-orbits.hdf5"
-    # make sure the out directory exists
-    outdir="./gallery_of_gaps/"
-    os.makedirs(outdir,exist_ok=True)
-    # get the perturber dictionary
-    MC_gaps=GR.GetPerturberDictPouliasis2017pii()
-    MCkeys=list(MC_gaps.keys())
-    MCkeys = ["monte-carlo-"+str(i).zfill(3) for i in range(0,50)]
-    # current time is always 0, b/c zero means today
-    currenttime= 0 # started at -5*u.Gyr
-    n_dynamic_times=2 # filtering the orbit by dynamical time
-    # iterate over the monte-carlo-keys
-    
-    scat_params={'s':1,'alpha':0.9,'cmap':'rainbow','norm':LogNorm(vmin=1e-5, vmax=5e-4)}
-    for montecarlokey in MCkeys:
-        # open the stream 
-        streamfilename = basepath+targetGC+"-stream-"+montecarlokey+".hdf5"
-        myStream = h5py.File(streamfilename, 'r')
-        # take a piece of the orbit in order to do the coordinate transform
-        with h5py.File(orbitpath,'r') as target:
-            tGC =target[montecarlokey]['t'][:]
-            xGC =target[montecarlokey]['xt'][:]
-            yGC =target[montecarlokey]['yt'][:]
-            zGC =target[montecarlokey]['zt'][:]
-            vxGC=target[montecarlokey]['vxt'][:]
-            vyGC=target[montecarlokey]['vyt'][:]
-            vzGC=target[montecarlokey]['vzt'][:]
-            
[…]

### 2024-09-27 paper/main.tex
made tau profiles better, changed color in figs

@@
 All simulations and labeled gaps are presented in the Appendix~\ref{appendix:gallery_of_gaps}.
-The profile method was useful in identifying the subtle underdense gaps of Samplings: 000, 002, 006, 008, 029, 030.
+The profile method was useful in identifying the subtle underdense gaps of Samplings: 000, 002, 006, 024, 028, 029, 030, 044, \& 048.
 
   Besides gaps, a handful of simulations show the entire stream laterally shifted in tail coordinates.
-The most extreme case was observed in Sampling 006 and is presented in the appendix XXX.
-This of course hampered our 1D profile difference analysis whose based the assumption that any difference in density would be due to the passage of other globular clusters and thus under densities.
+The most extreme case was observed in Sampling 006 and is presented in the Appendix~\ref{appendix:gallery_of_gaps}.
+This hampered our 1D profile difference analysis whose based the assumption that any difference in density would be due to the passage of other globular clusters and thus under densities.
 To combat this, we performed a cosmetique data reduction where I moved the \texttt{Full} 1D profile to best overlap with \texttt{Vanilla} by computing the correlation integral.
@@
 
-  The clusters that impacted Palomar 5 in our simulations are presented in Fig.~\ref{fig:histogram_impact_time}, along with when they impacted the stream. An imediate result is
-
-      - It takes a while for the stream to grow to hit a significant length
-      - A lot of them are from NGC2808. This is is because it is very recent. 
-      - Of the recent hits, the gaps like NGC2808 have widths as small as 100 parsecs.
-      - of the morphology, there are some gaps that have "bites" removed from them. Like that of NGC6333 in Sampling 033. 
-      - I think it is important to keep in mind that many of said gaps would not be detectable with current
+  The clusters that impacted Palomar 5's stream in our simulations are presented in Fig.~\ref{fig:histogram_impact_time}, along with when they impacted the stream.
+First, there are few impacts that happen early on in the simulation.
+This is not surprising, as it the tidal tails are smaller and thus there is less volume to be perturbed.
+Next, is that 44/50 simulations contain a visible gap from NGC2808, all of the encounters occuring about 200~Myr ago.
+This is so much more than any other peak that we broke the scale of the histogram.
+The reason is that the uncertainty in the precise position of the clusters along their orbits grows in look back time.
+Since this encounter is recent, most of the sampled initial conditions give rise to a significant enough impact in the stream that causes a gap.
 
@@
 
-    in terms of statistics: 
-    - We observe that only clusters above 1e5 cause gaps
+  To speak on the morphological of gaps, the passage of globular clusters has not effected the width of the streams, i.e.
+they have not made them dynamically \textit{warmer}.
+The flybys gently impart a change in momentum on a local patch of particles and general pushes them to the side.
+As identified in Erkal et al 2015, the width of a gap is determined by how long ago the impact occured, while the depth of a gap is more indicative of how hard it was hit.
+
+  The quantities of the gap that perhap define how much of a gap it is, would be the width and the depth.
+We do not attempt to quantify this in this study.
+However, there are very few simulations that have both wide and deep gaps.
+These are the most prominent gaps and come from NGC104 and NGC7078 on Sampling 009, NGC104 on Sampling 027, NGC7089 on Samplings 035, 049.
+
+  We desire to search for some generalizable characterisitcs of the perturbers to see what contrasts them to globular clusters that do not impact the Pal5 Stream.
+We present three different graphs in Fig.~\ref{fig:mass_size_plane}.
+The top plot shows the angular momentm and energy plane.
+The
 
@@
     \includegraphics[width=1\linewidth]{impact_geometry_statistics_deltaP.png}
-    \caption{One plot shows the distribution in energy and angular momentum, the next are the inernal proeprties of the clusters, and the last show when the clusters commit their crime, and how many times it happens accross the 50 simulations.
-}
+    \caption{Characterisitcs of clusters that cause gaps compared to those that do not.
+\textbf{Top:} shows the energy-angular momentum space of the globular clusters present in the simulations.
+Notice that we include all initial conditions that were sampled, therefore there are 50 $\times$ 165 data points.
+The clusters of a given simulation that impacted Palomar 5 are given with a Red marker, while non-gap causing clusters are given in gray.
+\textbf{Middle:} The mass-size plane.
+Uncertainties on the mass are indicated, while no uncertainties on the radius were reported.
+\textbf{Bottom:} the imparted momentum on the star-particles from a given flyby as given in Eq.~\ref{eq:change_in_momentum}.
+The top 5 strongest flybys from each simulation are included in this dataset, those that cause gaps are flagged.}
     \label{fig:mass_size_plane}
@@
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-000.png}
+      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-000.png}
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-001.png}
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-002.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-003.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-004.png}
+      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-002.png}
       \caption{Gap Gallery}
@@
       \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-003.png}      
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-004.png}
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-005.png}
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-006.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-007.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-008.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-009.png}
+      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-006.png}
       \caption{Gap Gallery}
@@
       \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-007.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-008.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-009.png}      
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-010.png}
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-011.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-012.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-013.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-014.png}
       \caption{Gap Gallery}
@@
       \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-012.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-013.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-014.png}      
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-015.png}
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-016.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-017.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-018.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-019.png}
       \caption{Gap Gallery}
@@
       \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-017.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-018.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-019.png}      
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-020.png}
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-021.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-022.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-023.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-024.png}
       \caption{Gap Gallery}
@@
       \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-022.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-023.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-024.png}   
+      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-024.png}
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-025.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-026.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-027.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-028.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-029.png}
       \caption{Gap Gallery}
@@
       \centering
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-030.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-031.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-032.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-033.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-034.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-026.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-027.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-028.png}
+      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-028.png}
       \caption{Gap Gallery}
@@
       \centering      
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-035.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-036.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-037.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-038.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-039.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-029.png}    
+      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-029.png}  
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-030.png}
+      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-030.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-031.png}
       \caption{Gap Gallery}
@@
       \centering
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-040.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-041.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-042.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-043.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-044.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-032.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-033.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-034.png}      
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-035.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-036.png}      
       \caption{Gap Gallery}
@@
       \centering
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-045.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-037.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-038.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-039.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-040.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-041.png}      
+      \caption{Gap Gallery}
+      \label{fig:TailCoordinates}
+    \end{figure*}        
+    \begin{figure*}
+      \centering
+
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-042.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-043.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-044.png}
+      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-044.png}
+      \caption{Gap Gallery}
+      \label{fig:TailCoordinates}
+    \end{figure*}
+
+    \begin{figure*}
+      \centering
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-046.png}
@@
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-048.png}
+      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-048.png}
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-049.png}

### 2024-09-27 paper/main.tex
more edits

@@
     \begin{equation} 
-      \Delta p \approx \text{Force} \times \text{interaction time} = \frac{GMm}{b^2} \times \frac{b}{v} = \frac{GM}{bv}, 
+      \Delta p \approx \text{Force} \times \text{interaction time} = \frac{GMm}{b^2} \times 2\frac{b}{v} = 2\frac{GM}{bv}, 
       \end{equation} \label{eq:change_in_momentum}
@@
 This process was followed for all 50 simulations to identify possible gaps.
-We find that small gaps are identifiable by eye rather well, where the 1D profiles aid in finding subtle yet extended underdensities.
+We find that small gaps are identifiable by eye rather well in the 2D density maps, where the 1D profiles aid in finding subtle yet extended underdensities.
 All simulations and labeled gaps are presented in the Appendix~\ref{appendix:gallery_of_gaps}.

### 2024-09-27 paper/main.tex
more results writing

@@
 
-    \subsubsection*{From Suspect to Culprit} \label{SuspectToCulprit}
+    \subsubsection*{From Suspect to Culprit} \label{sec:SuspectToCulprit}
 
@@
 
-    To achieve this, we begin by identifying the approximate moments of impact from the most significant clusters, as determined in the previous analysis in Sec.~\ref{SuspectToCulprit}.
+    To achieve this, we begin by identifying the approximate moments of impact from the most significant clusters, as determined in the previous analysis in Sec.~\ref{sec:SuspectToCulprit}.
 Then, we refine these estimates.
@@
 
-  One of the 50 simulations is shown in Fig.~\ref{fig:stream_on_sky}--which we selected for its clear and prominent gaps which are visibile in the scatter plot of the stars on the sky and becomes even more apparent when marginalizing over latitutde to see it's 1D density over longitude.
+  \subsection*{Overview}
+  This experiment follows a monte-carlo scheme, where we sample the initial conditions of the globular clusters' masses and kinematics based on their corresponding uncertainties and then repeat the experiment multiple times.
+We choose to perform the simulation 50 times to balance computational cost with a high enough number to allow us to explore a range of possible outcomes.
+The uncertainties on the globular clusters are large enough that over the course of the 5 Gyr integration period, they could be anywhere allowed in their phase space given their initial energy and angular momenta.
+We inspected all simulations upon their completion and present one of them in in Fig.~\ref{fig:stream_on_sky}--which we selected for its clear and prominent gaps.
+Two of these gaps are visible in galactic coordinates and become even more apparent when marginalizing over latitutde to view it's 1D profile density over longitude.
 
-  We then transformed the stream to tail coordinates for the both the \texttt{vanilla} and \texttt{full} simulations in Fig.~\ref{fig:profiles}.
+  For each completed simulation, we follow the scheme laid out in the methods.
+I.e.
+we simulate the creation of the Palomar 5 stream for 50 different samplings of the initial conditions, with two different cases.
+The \textit{Full} simulation that includes the influence of all globualr clusters on the star particles representing Palomar 5 and its stream, as well as the \textit{vanilla} simulations in which the star-particles feel only the galaxy and Palomar 5.
+We do this for we attempt to quantify the number of gaps from globular clusters in the limit of maximum information.
+I.e.
+we are not limited by contamination of field stars, and since we are computing our own simulations, we know what a gapless stream should look like.
+
+  This comparative analysis is shown in Fig.~\refP{fig:profiles}.
+We visually compare the 2D density map in tail coordinates as well as their marginalized 1D density profiles along the length of Palomar 5's orbit.
 As described in section~\ref{sec:gap_detection}, we locate regions on the stream where the full simulations are significantly less dense than the vanilla simulations.
-Notice the \textit{Tiny Gap} labeled in the middle plot.
-This signal is not present in the 1D profile since it is too small in size and at an oblique angle, thus is lost when marginalizing.
-This process was followed for all 50 simulations to identify possible gaps.
-We find that small gaps are identifiable by eye rather well in the 2D density maps, where the 1D profiles aid in finding subtle yet extended underdensities.
+We find that many streams as visible by eye.
+However, some subtle under densities only became apparent when comparing the 1D density profiles.
 All simulations and labeled gaps are presented in the Appendix~\ref{appendix:gallery_of_gaps}.
-The profile method was useful in identifying the subtle underdense gaps of Samplings: 000, 002, 006, 024, 028, 029, 030, 044, \& 048.
 
-  Besides gaps, a handful of simulations show the entire stream laterally shifted in tail coordinates.
-The most extreme case was observed in Sampling 006 and is presented in the Appendix~\ref{appendix:gallery_of_gaps}.
-This hampered our 1D profile difference analysis whose based the assumption that any difference in density would be due to the passage of other globular clusters and thus under densities.
-To combat this, we performed a cosmetique data reduction where I moved the \texttt{Full} 1D profile to best overlap with \texttt{Vanilla} by computing the correlation integral.
-Since this is not the focus of this study, we did not investigate further.
-We conjecture that the translation may arrise from a resonance phenomenon, most likely coming from Omega Centauri.
-Perhaps, in some orbital configurations Palomar 5 and Omega Centauri pass each other in regular intervals that can accelerate the globular cluster differently from the stream.
+  Subsequently, we followed the analysis presented in section~\ref{sec:SuspectToCulprit} to identify which globualr clusters were responsible for the gaps.
+This method used the orbit of Palomar 5 as a proxy for the stream.
+The gravitational force from all other globular clusters as computed at segments of the orbit about the position of Palomar 5 for the entire simulation.
+The top 5 strongest flybys were recorded to identify which cluster was responsible.
+This is shown in Fig.~\ref{fig:force-on-orbit}, where the top panel shows the force from the globular clusters on the orbit, while the bottom panel shows the density of the stream.
+A density threshold was imparted to quantify the length of the stream.
+The location and time of impact was cross referenced to the location of the gaps to identify which cluster was responsible.
+
+  Fig.~\ref{fig:force-on-orbit} contains some interesting information.
+Notice the periodic ribbions of force in the $(t,\tau)$ plane.
+This is due to pericenter passages where Palomar 5 is getting closer to the center of mass of the globular cluster system.
+Additionally, for the impacts of NGC104 and NGC7078, wakes can be observed in the density map.
+Another important aspect, is that the strongest peak in gravitational force does not necessarily create a gap.
+Notice how NGC5272, which was labeled with 0 to indicate that it has the greatest local maxima does not have a gap.
+This is manifold, for instance, the force needs to be modulated with the time of impact because the  important quantity that determins gap creation.
+Additionally, there is an offset of about 200~pc between the stream at the orbit, as seen when viewing the stream in tail coordinates, so peaks upon the orbit are good proxy's for the stream but are not definitive.
+We found that the top five greatest impacts accounted for all gaps, except for Sampling~014 as shown in Fig.~\ref{fig:gallery3}, whose gap from NGC~6584 corresponded to the 7th peak.
+
+ For some cases, we performed a double check by launching extra simulations only including one suspect at a time to confirm that this suspect indeed caused a gap in question, an example is shown in Fig.~\ref{fig:decomposition}.
+Note that the gap from NGC~7078 was \textit{hit twice}.
+
+ \subsection*{Gap Morphology}
+
+ \subsection*{Perturber Statistics}
+  With this analysis complete, we are able to study the population of globular clusters that could perturb Palomar 5's stream.
 
@@
     \end{figure*}
-
-  To speak on the morphological of gaps, the passage of globular clusters has not effected the width of the streams, i.e.
-they have not made them dynamically \textit{warmer}.
-The flybys gently impart a change in momentum on a local patch of particles and general pushes them to the side.
-As identified in Erkal et al 2015, the width of a gap is determined by how long ago the impact occured, while the depth of a gap is more indicative of how hard it was hit.
-
-  The quantities of the gap that perhap define how much of a gap it is, would be the width and the depth.
-We do not attempt to quantify this in this study.
-However, there are very few simulations that have both wide and deep gaps.
-These are the most prominent gaps and come from NGC104 and NGC7078 on Sampling 009, NGC104 on Sampling 027, NGC7089 on Samplings 035, 049.
 
@@
       \end{figure}
-
-  After we discuss the magic to make detections, I THEN will discuss some general properties, i.e.
-is there something special about these perturbers compared to other clusters?
-The main results is that \textit{you need to enter Pal5's orbital space}, which is not a shocker.
-If you can hit me, you can't perturb me.
-NEXT, I want to compare the population of perturbers to all globular clusters based on their internal parameters, so here we do show a tendency.
-Next, I wanna show how often gaps appear, and thus how many time a given GC pertubered Pal5, and when the gaps occur.
-I think a conclusion with the histograms in time is
-
-  Lastly, I will try to show something about the impact geometry.
-I will state that we want to compare the ones that create gaps to those that did not.
-We wanted to apply what we learned from Carlberg, Erkal and Sanders.
-The moral of the story will be that, no combination of parameters explaining the geometry cleared demarcated the presence of a gap or not.
-This can be inline with Sander's expanion of the analysis initiated by Carlberg and extended by Erkal.
-Namly, a uniformly dense stream on a circular orbit is just not good enough.
-I think Sanders also said that the streams needs to be sufficiently diffuse, which could make sense given the epicycles.
 
@@
       \caption{Gap Gallery}
-      \label{fig:TailCoordinates}
+      \label{fig:gallery0}
       \end{figure*}
@@
       \caption{Gap Gallery}
-      \label{fig:TailCoordinates}
+      \label{fig:gallery1}
       \end{figure*}
@@
       \caption{Gap Gallery}
-      \label{fig:TailCoordinates}
+      \label{fig:gallery2}
       \end{figure*}
@@
       \caption{Gap Gallery}
-      \label{fig:TailCoordinates}
+      \label{fig:gallery3}
       \end{figure*}
@@
       \caption{Gap Gallery}
-      \label{fig:TailCoordinates}
+      \label{fig:gallery4}
       \end{figure*}
@@
       \caption{Gap Gallery}
-      \label{fig:TailCoordinates}
+      \label{fig:gallery5}
       \end{figure*}
@@
       \caption{Gap Gallery}
-      \label{fig:TailCoordinates}
+      \label{fig:gallery6}
       \end{figure*}
@@
       \caption{Gap Gallery}
-      \label{fig:TailCoordinates}
+      \label{fig:gallery7}
       \end{figure*}
@@
       \caption{Gap Gallery}
-      \label{fig:TailCoordinates}
+      \label{fig:gallery8}
       \end{figure*}
@@
       \caption{Gap Gallery}
-      \label{fig:TailCoordinates}
+      \label{fig:gallery9}
     \end{figure*}
+
     \begin{figure*}
       \centering
-
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-042.png}
@@
       \includegraphics[width=\linewidth]{tau-profile-monte-carlo-044.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-045.png}
       \caption{Gap Gallery}
-      \label{fig:TailCoordinates}
+      \label{fig:gallery010}
     \end{figure*}
@@
       \caption{Gap Gallery}
-      \label{fig:TailCoordinates}
+      \label{fig:gallery11}
     \end{figure*}

### 2024-09-30 paper/main.tex
edited the overview

@@
   \subsection*{Overview}
-  This experiment follows a monte-carlo scheme, where we sample the initial conditions of the globular clusters' masses and kinematics based on their corresponding uncertainties and then repeat the experiment multiple times.
-We choose to perform the simulation 50 times to balance computational cost with a high enough number to allow us to explore a range of possible outcomes.
-The uncertainties on the globular clusters are large enough that over the course of the 5 Gyr integration period, they could be anywhere allowed in their phase space given their initial energy and angular momenta.
+  This experiment follows a monte-carlo scheme, where we sample the initial conditions of the globular clusters' masses and kinematics based on their reported observational uncertainties and repeat the experiment multiple times.
+We choose to perform the simulation 50 times to balance computational and exploring a range of possible outcomes.
+The uncertainties on the globular clusters are large enough that over the course of the 5 Gyr integration period, they could be anywhere in their phase space given their initial energy and angular momentum.
 We inspected all simulations upon their completion and present one of them in in Fig.~\ref{fig:stream_on_sky}--which we selected for its clear and prominent gaps.
@@
 I.e.
-we simulate the creation of the Palomar 5 stream for 50 different samplings of the initial conditions, with two different cases.
-The \textit{Full} simulation that includes the influence of all globualr clusters on the star particles representing Palomar 5 and its stream, as well as the \textit{vanilla} simulations in which the star-particles feel only the galaxy and Palomar 5.
-We do this for we attempt to quantify the number of gaps from globular clusters in the limit of maximum information.
-I.e.
-we are not limited by contamination of field stars, and since we are computing our own simulations, we know what a gapless stream should look like.
+we simulate the creation of the Palomar 5 stream for 50 different samplings of the initial conditions in two different environments.
+The first are the \texttt{vanilla} simulations in which the star-particles feel only the galaxy and Palomar 5.
+On the other hand, the \texttt{full} simulations includes the influence of all globualr clusters on the star-particles.
 
-  This comparative analysis is shown in Fig.~\refP{fig:profiles}.
+  By performing a comparative analysis, we are exploring the phenomenon of the influence from all globular clusters on stellar streams in a regieme of far more complete information respect to reality.
+In this experiment, we do not deal with differentiating stream and field stars, and above all else in reality we do not have a reference reality without the effect of other globular clusters as a point of comparison.
+Additionally, we avoid the difficulty of attempting to define a gap in its own right.
+Erkal et al 2015 was able to do this pretty well by studying two parameters: the width of the gaps and the ratio of the minimum density to the density of the gaps' shoulders.
+However, this becomes impossible when departing from the idealized case of uniformly dense \& one-dimensional streams on circular orbits.
+Sanders et al 2016 extended this in action-angle space which allowed for dropping the constraint of circular orbits.
+However, I need to revist the article again because I believe they also explored non uniformly dense streams.
+Nonetheless, we avoid this all together by doing our comparative analysis.
+Since we are numerical modellers, we can play god and run nearly the exact same experiment by including or not including the globular clusters.
+Therefore, it's much easier to identify what comes from the clusters or not.
+
+  This comparative analysis is shown in Fig.~\ref{fig:profiles}.
 We visually compare the 2D density map in tail coordinates as well as their marginalized 1D density profiles along the length of Palomar 5's orbit.
 As described in section~\ref{sec:gap_detection}, we locate regions on the stream where the full simulations are significantly less dense than the vanilla simulations.
-We find that many streams as visible by eye.
+We find that many gaps are visible by eye.
 However, some subtle under densities only became apparent when comparing the 1D density profiles.
@@
 This method used the orbit of Palomar 5 as a proxy for the stream.
-The gravitational force from all other globular clusters as computed at segments of the orbit about the position of Palomar 5 for the entire simulation.
-The top 5 strongest flybys were recorded to identify which cluster was responsible.
+This is contrary to the typical case where one uses the stream to approximate the orbit--but $\approx$ works both ways.
+The gravitational force from all other globular clusters was computed at segments of the orbit about the position of Palomar 5 for the entire simulation.
+The top 5 strongest flybys were identified and the corresponding perturbers are labeled as \textit{suspects}.
 This is shown in Fig.~\ref{fig:force-on-orbit}, where the top panel shows the force from the globular clusters on the orbit, while the bottom panel shows the density of the stream.
@@
 The location and time of impact was cross referenced to the location of the gaps to identify which cluster was responsible.
+I.e.
+if the gap was observed on the leading tail, we searched for suspects with $\tau>0$.
 
-  Fig.~\ref{fig:force-on-orbit} contains some interesting information.
+  The top panel of Fig.~\ref{fig:force-on-orbit} contains some interesting information.
 Notice the periodic ribbions of force in the $(t,\tau)$ plane.
 This is due to pericenter passages where Palomar 5 is getting closer to the center of mass of the globular cluster system.
-Additionally, for the impacts of NGC104 and NGC7078, wakes can be observed in the density map.
+Additionally, for the impacts of NGC104 and NGC7078, wakes can be observed in the density map---evidence that signifies their guilt.
 Another important aspect, is that the strongest peak in gravitational force does not necessarily create a gap.
-Notice how NGC5272, which was labeled with 0 to indicate that it has the greatest local maxima does not have a gap.
-This is manifold, for instance, the force needs to be modulated with the time of impact because the  important quantity that determins gap creation.
-Additionally, there is an offset of about 200~pc between the stream at the orbit, as seen when viewing the stream in tail coordinates, so peaks upon the orbit are good proxy's for the stream but are not definitive.
+Notice how NGC~5272, which was labeled with 0 to indicate that it has the greatest force of all local maxima, does not provoke a gap.
+The reason for this is manifold.
+First, remember that the important quantity is \textit{the change in momentum}.
+Thus, a strong force applied for a brief peroid of time may have no effect on the stream!
+Additionally, there is an offset of about 200~pc between the stream at the orbit, as seen when viewing the stream in tail coordinates.
+This demonstrates that the orbit are good proxy's for the stream but are not a one for one substitute.
 We found that the top five greatest impacts accounted for all gaps, except for Sampling~014 as shown in Fig.~\ref{fig:gallery3}, whose gap from NGC~6584 corresponded to the 7th peak.
@@
  For some cases, we performed a double check by launching extra simulations only including one suspect at a time to confirm that this suspect indeed caused a gap in question, an example is shown in Fig.~\ref{fig:decomposition}.
-Note that the gap from NGC~7078 was \textit{hit twice}.
+Note that the gap from NGC~7078 was \textit{hit} by NGC~7078 \textit{twice}.
 

### 2024-09-30 paper/main.tex
first pass at the morphology

@@
  \subsection*{Gap Morphology}
+  We study the structure of the streams that have been impact by the globular clusters.
+The gallery of gaps is shown in the appendix, which includes the results from all of the simulations we considered in this experiment.
+
+  In general, how did the passage of these globular clusters effect the stream?
+well.
+A lot of this has been studied quite well and presented in Sanders et al 2016.
+They even consdiered different types of impactors.
+I.e.
+impactors that had different density profiles.
+They compared Hernquist, Plummer, NFW, and a truncated NFW.
+There are some minor differences between the impactors however, the conclusion is still the same.
+Impacts from small and dense perturbers such as globular clusters only effect the stream locally.
+
+  Our results are consistent with those shown previously.
+Perturber crossings only perturb the stream locally.
+Global characterisitcs of the stream do not change.
+For instance, the stream is not warmer of thicker.
+Other types of interactions have been demonstrated to do this.
+For example THIS AUTHOR showed that XXX and do YYY.
+As laid out in Erkal et al 2015, perturbers pull particles from the sides towards the impact point.
+These particles eventually move beyond the impact point and leave an under dense center and then pile up on the sides leaving an overdensity on the shoulders.
+
+  Interestingly, most of our gaps occur for the leading tail.
+8 on the trailing tail.
+21 excluding NGC~2808 are on the leading tail.
+We have to exclude NGC~2808 because the simulations are clearly correlated with one another, so trials are not independent.
+So is 8/29 consistent with a fair coin toss?
+In anycase, we argue this may be a coincidence since being hit should only depend on the size and location of the stream.
+The main difference between the tails is a $\approx$ 300-400 pc offset in $r$, yet this is negligible.
+
+  All of the gaps are oblique in tail coordinates.
+This is an artifact of using tail coordinates and arrises because we project the star-particles onto the nearest point of the orbit.
+(this is really difficult to explain with words and super simple with a picture.
+The picture being that we have three evenly spaced parallel lines that are close together.
+Then, a third point that is far away.
+)
+
+  Do our gaps grow linearly in time? what about this square root stuff?
+
+  Lastly, since the streams have width, there are some instances where the stream is not significantly impact across this width.
+For instance, the gaps from NGC~5284 in Sampling 019 and NGC~6333 in Sampling 033 are more akin to bites than gaps.
 

### 2024-09-30 paper/main.tex
updated spacing
updating the statistics

@@
 Next, is that 44/50 simulations contain a visible gap from NGC2808, all of the encounters occuring about 200~Myr ago.
-This is so much more than any other peak that we broke the scale of the histogram.
+This is so much more than any other peak that we broke the scale of the histogram so that this peak would not dwarf the significance of all others.
 The reason is that the uncertainty in the precise position of the clusters along their orbits grows in look back time.
@@
       \includegraphics[width=\linewidth]{histogram_impact_time.png}
-      \caption{When the impacts occured}
+      \caption{The time of impact for each flyby that caused a gap.
+Each cluster is given a consistent and unique color.
+The scale of the histogram is broken between 8 and 42 since the gaps from NGC~2808 are in a league of their own with respect to other clusters.}
       \label{fig:histogram_impact_time}
@@
 The top plot shows the angular momentm and energy plane.
-The
+To no surprise, the globular clusters that perturb Palomar 5 are within a band of energy enclosed the energy of Palomar 5.
+This indicates that a cluster's orbital space needs to overlap with Palomar 5's.
+There appears not to be a preference in angular momentum.
+Note that in this panel all initial conditions for the clusters are plotted.
+Clusters of gray dots indicate the same cluster.
+However, for the least bound milky way clusters, the uncertainties in distances and motions allow for a wide range of energies, prograde and retrograde orbits.
+These are Crater, Palomar X and Eridanus.
+This is consistent with Vasiliev et al 2021, yet not very significant for this study since they do not influence Palomar 5 and could have been ignored.
 
-    - to no surprise, you have to come within Palomar 5's orbital space
+    Next, in the middle panel of Fig.~\ref{fig:mass_size_plane} we view the mass size plane. The
 

### 2024-09-30 paper/main.tex
compiled the whole thang
passed the overview to chatgpt

@@
   \subsection*{Overview}
-    This experiment follows a monte-carlo scheme, where we sample the initial conditions of the globular clusters' masses and kinematics based on their reported observational uncertainties and repeat the experiment multiple times.
-We choose to perform the simulation 50 times to balance computational and exploring a range of possible outcomes.
-The uncertainties on the globular clusters are large enough that over the course of the 5 Gyr integration period, they could be anywhere in their phase space given their initial energy and angular momentum.
-We inspected all simulations upon their completion and present one of them in in Fig.~\ref{fig:stream_on_sky}--which we selected for its clear and prominent gaps.
-Two of these gaps are visible in galactic coordinates and become even more apparent when marginalizing over latitutde to view it's 1D profile density over longitude.
+    This experiment follows a Monte Carlo scheme, where we sample the initial conditions of the globular clusters' masses and kinematics based on their reported observational uncertainties.
+We repeat the experiment multiple times, running 50 simulations to balance computational cost with the exploration of possible outcomes.
+The uncertainties in the globular clusters' parameters are significant enough that, over the 5 Gyr integration period, they could occupy various regions of their phase space given their initial energy and angular momentum.
+After inspecting all simulations, we present one (Fig.~\ref{fig:stream_on_sky}) chosen for its clear and prominent gaps.
+Two of these gaps are visible in galactic coordinates, and they become even more apparent when we marginalize over latitude to view the 1D profile density along the longitude.
 
-    For each completed simulation, we follow the scheme laid out in the methods.
-I.e.
-we simulate the creation of the Palomar 5 stream for 50 different samplings of the initial conditions in two different environments.
-The first are the \texttt{vanilla} simulations in which the star-particles feel only the galaxy and Palomar 5.
-On the other hand, the \texttt{full} simulations includes the influence of all globualr clusters on the star-particles.
+    For each simulation, we follow the methods outlined earlier.
+We simulate the creation of the Palomar 5 stream for 50 different samplings of the initial conditions in two environments.
+In the \texttt{vanilla} simulations, the star particles only feel the gravitational influence of the galaxy and Palomar 5.
+In contrast, the \texttt{full} simulations include the effects of all globular clusters on the star particles.
 
-    By performing a comparative analysis, we are exploring the phenomenon of the influence from all globular clusters on stellar streams in a regieme of far more complete information respect to reality.
-In this experiment, we do not deal with differentiating stream and field stars, and above all else in reality we do not have a reference reality without the effect of other globular clusters as a point of comparison.
-Additionally, we avoid the difficulty of attempting to define a gap in its own right.
-Erkal et al 2015 was able to do this pretty well by studying two parameters: the width of the gaps and the ratio of the minimum density to the density of the gaps' shoulders.
-However, this becomes impossible when departing from the idealized case of uniformly dense \& one-dimensional streams on circular orbits.
-Sanders et al 2016 extended this in action-angle space which allowed for dropping the constraint of circular orbits.
-However, I need to revist the article again because I believe they also explored non uniformly dense streams.
-Nonetheless, we avoid this all together by doing our comparative analysis.
-Since we are numerical modellers, we can play god and run nearly the exact same experiment by including or not including the globular clusters.
-Therefore, it's much easier to identify what comes from the clusters or not.
+    Through a comparative analysis, we explore the influence of the globular clusters on stellar streams under conditions of greater clarity than reality offers.
+In this controlled setting, we avoid challenges like distinguishing stream stars from field stars and do not need to define gaps independently.
+As numerical modelers, we can "play god" and run the same experiment with or without globular clusters, making it easier to isolate their effects.
 
-    This comparative analysis is shown in Fig.~\ref{fig:profiles}.
-We visually compare the 2D density map in tail coordinates as well as their marginalized 1D density profiles along the length of Palomar 5's orbit.
-As described in section~\ref{sec:gap_detection}, we locate regions on the stream where the full simulations are significantly less dense than the vanilla simulations.
-We find that many gaps are visible by eye.
-However, some subtle under densities only became apparent when comparing the 1D density profiles.
-All simulations and labeled gaps are presented in the Appendix~\ref{appendix:gallery_of_gaps}.
+    This analysis is illustrated in Fig.~\ref{fig:profiles} we identify regions where the full simulations show significantly lower densities than the vanilla simulations.
+Many gaps are clearly visible, while some subtler under-densities only appear when examining the 1D profiles.
+All simulations and labeled gaps are presented in Appendix~\ref{appendix:gallery_of_gaps}
 
-    Subsequently, we followed the analysis presented in section~\ref{sec:SuspectToCulprit} to identify which globualr clusters were responsible for the gaps.
-This method used the orbit of Palomar 5 as a proxy for the stream.
-This is contrary to the typical case where one uses the stream to approximate the orbit--but $\approx$ works both ways.
-The gravitational force from all other globular clusters was computed at segments of the orbit about the position of Palomar 5 for the entire simulation.
-The top 5 strongest flybys were identified and the corresponding perturbers are labeled as \textit{suspects}.
-This is shown in Fig.~\ref{fig:force-on-orbit}, where the top panel shows the force from the globular clusters on the orbit, while the bottom panel shows the density of the stream.
-A density threshold was imparted to quantify the length of the stream.
-The location and time of impact was cross referenced to the location of the gaps to identify which cluster was responsible.
-I.e.
-if the gap was observed on the leading tail, we searched for suspects with $\tau>0$.
+    Next, we followed the method in section~\ref{sec:SuspectToCulprit} to identify which globular clusters caused the gaps.
+This approach uses the orbit of Palomar 5 as a proxy for the stream.
+We computed the gravitational force from all globular clusters at various segments along Palomar 5's orbit for the entire simulation.
+The five strongest flybys were identified, and the corresponding clusters labeled as "suspects." Fig.~\ref{fig:force-on-orbit} shows this: the top panel displays the force exerted by the globular clusters on Palomar 5's orbit, while the bottom panel shows the stream density.
+A density threshold was imposed to quantify the stream's length.
+We cross-referenced the impact locations with the gaps to determine which cluster was responsible.
+For example, gaps in the leading tail correspond to suspects with $\tau>0$.
 
-    The top panel of Fig.~\ref{fig:force-on-orbit} contains some interesting information.
-Notice the periodic ribbions of force in the $(t,\tau)$ plane.
-This is due to pericenter passages where Palomar 5 is getting closer to the center of mass of the globular cluster system.
-Additionally, for the impacts of NGC104 and NGC7078, wakes can be observed in the density map---evidence that signifies their guilt.
-Another important aspect, is that the strongest peak in gravitational force does not necessarily create a gap.
-Notice how NGC~5272, which was labeled with 0 to indicate that it has the greatest force of all local maxima, does not provoke a gap.
-The reason for this is manifold.
-First, remember that the important quantity is \textit{the change in momentum}.
-Thus, a strong force applied for a brief peroid of time may have no effect on the stream!
-Additionally, there is an offset of about 200~pc between the stream at the orbit, as seen when viewing the stream in tail coordinates.
-This demonstrates that the orbit are good proxy's for the stream but are not a one for one substitute.
-We found that the top five greatest impacts accounted for all gaps, except for Sampling~014 as shown in Fig.~\ref{fig:gallery3}, whose gap from NGC~6584 corresponded to the 7th peak.
-
-    For some cases, we performed a double check by launching extra simulations only including one suspect at a time to confirm that this suspect indeed caused a gap in question, an example is shown in Fig.~\ref{fig:decomposition}.
-Note that the gap from NGC~7078 was \textit{hit} by NGC~7078 \textit{twice}.
+    The top panel of Fig.\ref{fig:force-on-orbit} reveals interesting periodic patterns in the $(t,\tau)$ plane.
+This is due to pericenter passages when Palomar 5 approaches the center of mass of the globular cluster system.
+Notably, for the collisions from NGC104 and NGC7078, we observe wakes in the density map, which confirm their role in creating gaps.
+However, the strongest peak in gravitational force does not always result in a gap.
+For example, NGC~5272 exerts the greatest force, but it does not cause a gap.
+This is because the change in momentum—not just the force—matters.
+A brief, strong force may have little effect on the stream, especially with an offset of about 200 pc between the stream and the orbit in tail coordinates.
+Despite this, the five strongest impacts account for all gaps except in Sampling-014 (Fig.\ref{fig:gallery3}), where a gap caused by NGC~6584 was linked to the 7th strongest peak.
+In some cases, we ran additional simulations with only a single suspect to confirm its role in creating a gap.
+An example of this is shown in Fig.\ref{fig:decomposition}.
 

### 2024-09-30 paper/main.tex
touched up gap morphology

@@
   \subsection*{Gap Morphology}
-    We study the structure of the streams that have been impact by the globular clusters.
-The gallery of gaps is shown in the appendix, which includes the results from all of the simulations we considered in this experiment.
+    We analyze the structure of streams impacted by globular clusters. The gallery of gaps, found in the appendix, includes results from all the simulations considered in this experiment.
 
-    In general, how did the passage of these globular clusters effect the stream?
-well.
-A lot of this has been studied quite well and presented in Sanders et al 2016.
-They even consdiered different types of impactors.
-I.e.
-impactors that had different density profiles.
-They compared Hernquist, Plummer, NFW, and a truncated NFW.
-There are some minor differences between the impactors however, the conclusion is still the same.
-Impacts from small and dense perturbers such as globular clusters only effect the stream locally.
+    How do globular cluster passages affect the stream?
+Much of this has been well studied, as discussed in Sanders et al.
+(2016).
+They even considered different types of impactors, such as those with varying density profiles (e.g., Hernquist, Plummer, NFW, and truncated NFW).
+While there are some minor differences between these impactors, the overall conclusion remains the same: impacts from small, dense perturbers, such as globular clusters, only affect the stream locally.
 
-    Our results are consistent with those shown previously.
-Perturber crossings only perturb the stream locally.
-Global characterisitcs of the stream do not change.
-For instance, the stream is not warmer of thicker.
-Other types of interactions have been demonstrated to do this.
-For example THIS AUTHOR showed that XXX and do YYY.
-As laid out in Erkal et al 2015, perturbers pull particles from the sides towards the impact point.
-These particles eventually move beyond the impact point and leave an under dense center and then pile up on the sides leaving an overdensity on the shoulders.
+    Our results align with previous findings—perturber crossings cause local perturbations in the stream without altering its global characteristics.
+For instance, the stream does not become warmer or thicker, changes that have been linked to other types of interactions.
+For example, [AUTHOR] demonstrated that XXX can do YYY.
+As described by Erkal et al.
+(2015), perturbers pull particles from the sides toward the impact point, creating an under-dense region at the center of the gap, with overdensities piling up at the shoulders.
 
-    Interestingly, most of our gaps occur for the leading tail.
-8 on the trailing tail.
-21 excluding NGC~2808 are on the leading tail.
-We have to exclude NGC~2808 because the simulations are clearly correlated with one another, so trials are not independent.
-So is 8/29 consistent with a fair coin toss?
-In anycase, we argue this may be a coincidence since being hit should only depend on the size and location of the stream.
-The main difference between the tails is a $\approx$ 300-400 pc offset in $r$, yet this is negligible.
+    Interestingly, most of the gaps we observe occur in the leading tail.
+However, we believe this to be coincidental.
+Stars escape from the globular cluster through the two Lagrange points at equal rates (cite?).
+The lengths of the tails are also equal, with only a small difference of approximately 400 pc in radial distance from the galactic center.
+This difference is negligible compared to the orbital radius of about 10 kpc.
+Therefore, we expect the probability of a gap appearing in the leading or trailing tail to be equal, and the presence of one gap should not influence the likelihood of another.
+
+    Yet, the observation of only 8 gaps in the trailing tail out of 73 total gaps is inconsistent with this expectation.
+However, 44 of these gaps are associated with NGC2808.
+As shown in Fig.\ref{fig:histogram_impact_time}, all of these gaps result from the same recent encounter.
+We thus discard NGC~2808 for the current analysis, as its gaps violate the assumption of independent events.
+After excluding these, we are left with 21 gaps in the leading tail and 8 in the trailing tail across the remaining simulations.
+The probability of observing 8 or fewer gaps in 29 trials, assuming each has a 50\% chance, is approximately 1.2\%.
+While this outcome is unlikely, it is not impossible.
 

### 2024-10-02 paper/main.tex
moved some methods to appendix

--- 2024-09-30 paper/main.tex
@@
 We focus on the Palomar 5 stream, which is well-studied, has observable gaps (bonaca et al 20XX), and is highly visible due to its position above the galactic disk.
-We present a result from one of our simulations in \ref{fig:stream_on_sky}, and the following sections describe our modeling of the stream, the gravitational interactions with the galaxy and other globular clusters, and the statistical analysis of perturbations affecting the stream.
+We present a result from one of our simulations in Fig.~\ref{fig:stream_on_sky}, and the following sections describe our modeling of the stream, the gravitational interactions with the galaxy and other globular clusters, and the statistical analysis of perturbations affecting the stream.
 
@@
 We handle the uncertainties through a monte-carlo approach.
-The catalog provides uncertainties for mass, distances, line of sight velocities, proper motions, with a covariance term between the proper motions.
+The catalog provides uncertainties for mass, distances, line of sight velocities, proper motions and includes a covariance term between the proper motions.
 We repeate the whole experiment 50 times, for each of which we vary the initial conditions of each cluster by sampling its five-dimensional gaussian uncertaintiy distribution.
@@
 Instead of treating the clusters as point masses that only experience the gravitational potential of the galaxy, we now compute N-body interactions between the clusters during the backward integration over 5 billion years.
-The equation of motion for the globular clusters is thus:
+The equations of motion for the globular cluster system are thus:
     \begin{equation}
-      \ddot{\vec{r}}_i = -\nabla \Phi(R_i,z_i) + \left.\sum_{j\neq i} \frac{Gm_j}{\left(|\vec{r}_j - \vec{r}_i|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j - \vec{r}_i\right).
-      \end{equation}\label{eq:GCNBody}
-    Although these interactions were included, we observed no significant exchanges of orbital energy or momentum between clusters.
+      \ddot{\vec{r}}_i = -\nabla \Phi(R_i,z_i) + \left.\sum_{j\neq i} \frac{Gm_j}{\left(|\vec{r}_j - \vec{r}_i|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j - \vec{r}_i\right),
+    \end{equation}\label{eq:GCNBody} where $\vec{r}$ indicates the galacto-centric position, the index $i$ indicates a globular cluster of interest with $R_i,z_i$ being its galactocentric positions in cylindrical coordinates.
+The index $j$ indicates the other globular cluster that are summed over where: $m_j$ is its mass, $b_j$ is its plummer scale radius, and $r_j$ is it's galacto-centric position.
+$\Phi$ is the axis-symmteric potential of the model used for the Milky Way.
 
@@
     \begin{equation}
-      \ddot{\vec{r}}_p = -\nabla \Phi(R_p,z_p) + \left.\sum_{j} \frac{Gm_j}{\left(|\vec{r}_j(t) - \vec{r}_p|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j(t)- \vec{r}_p\right).
-      \end{equation} \label{eq:equation_of_motion_particle}
+      \ddot{\vec{r}}_p = -\nabla \Phi(R_p,z_p) + \left.\sum_{j} \frac{Gm_j}{\left(|\vec{r}_j(t) - \vec{r}_p|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j(t)- \vec{r}_p\right),
+      \end{equation} \label{eq:equation_of_motion_particle} where the index $p$ represents one of the 100,000 particles of interest.
 
@@
 Specifically, for each of the 50 realizations of the Palomar 5 stream, we saved $5000$ intermediate timesteps which is a temporal resolution of 1 million years.
-To provide an estimate of the size of our simulations, we have $N_p=100 000$, $N_\textrm{time stamps} = 5 000$, $N_\textrm{phase}=6$, $N_\textrm{sampling}=50$, and using single precision floating point numbers, $\textrm{Single}=4$~bytes, results in approximately 600 GB of data.
+To provide an estimate of the size of our simulations, we have $N_p=100 000$ number of particles, $N_\textrm{time stamps} = 5 000$ time steps, $N_\textrm{phase}=6$ phase space coordinates, $N_\textrm{sampling}=50$ monte-carlo samplings, and using single precision floating point numbers, $\textrm{Single}=4$~bytes.
+The product of the aforementioned terms results in approximately: $N_p=100 000 \times N_\textrm{time stamps}\times N_\textrm{phase}\times N_\textrm{sampling} 4 \textrm{bytes}\approx$ 600 GB of data.
 
-  \subsection{Gap detection} \label{sec:gap_detection}
+  \subsection{Gap detection and Characterization} \label{sec:gap_detection}
 
@@
 
-    To facilitate this process, we ran a series of \texttt{vanilla} simulations.
-In these control simulations, we used the same 50 sets of initial conditions but excluded interactions between globular clusters and omitted the gravitational influence of the clusters on the star particles within the stream.
-By comparing the morphological differences between these control simulations and the full simulations (which included cluster interactions), we could identify and assess the impact of globular clusters on the stream structure more effectively.
+    To facilitate this process, we ran a series of \texttt{reference} simulations.
+In these control simulations, we used the same 50 sets of initial conditions but excluded interactions between globular clusters and omitted the gravitational influence of the clusters on the star particles within the stream--save for the host cluster Palomar 5.
+By comparing the morphological differences between these control simulations and the \texttt{full} simulations (which included cluster interactions), we could identify and assess the impact of globular clusters on the stream structure more effectively.
 
-    Comparisons between the \texttt{vanilla} and full simulations were performed in the tail coordinate system.
+    Comparisons between the \texttt{reference} and \texttt{full} simulations were performed in the tail coordinate system.
 This system allows for consistent morphological comparisons by aligning the stream with the cluster's orbit.
@@
 (2004) and is detailed in Appendix~\ref{appendix:TailCoordinates}.
-
 Briefly, the x' coordinate represents the position of a particle along the orbit relative to the globular cluster—whether it is ahead or behind the cluster.
@@
 
+    More details on how the gaps and corresponding perturbing globular clusters were identified is provided in Appendix.~XXX.
+A first order approach of understanding the significance of an impact one needs to compute its change in momentum, which is determined by integrating the force acting on the particle over the duration of the interaction.
+A useful approximation for this change in momentum is the force at the closest approach multiplied by an estimate of the interaction time:
+    \begin{equation} 
+      \Delta v \approx \text{Force} \times \text{interaction time} = \frac{GM}{b^2} \times 2\frac{b}{v} = 2\frac{GM}{bv}, 
+      \end{equation} \label{eq:change_in_momentum}
+    where $M$ is the mass of the perturber, $m$ is the mass of the particle, $b$ is the impact parameter, $v$ is the relative velocity of the perturber with respect to the particle, and $G$ is the gravitational constant.
+We compute this for the top five cloestest approaches of all globular clusters on an orbital segment about Palomar 5 that approximates the stream through out each of the 50 simulations.
+In the end, we have a sample of 250 impacts and compare the impact geometry of those that cause gaps to those that do not.
+
+\section{Results}
+
+  \subsection*{Overview}
+
+    Although these interactions were included, we observed no significant exchanges of orbital energy or momentum between clusters.
+
+    This experiment follows a monte-carlo scheme, where we sample the initial conditions of the globular clusters' masses and kinematics based on their corresponding uncertainties and then repeat the experiment multiple times.
+We choose to perform the simulation 50 times to balance computational cost with a high enough number to allow us to explore a range of possible outcomes.
+The uncertainties on the globular clusters are large enough that over the course of the 5 Gyr integration period, they could be anywhere allowed in their phase space given their initial energy and angular momenta.
+We inspected all simulations upon their completion and present one of them in in Fig.~\ref{fig:stream_on_sky}--which we selected for its clear and prominent gaps.
+Two of these gaps are visible in galactic coordinates and become even more apparent when marginalizing over latitutde to view it's 1D profile density over longitude.
+
+    For each completed simulation, we follow the scheme laid out in the methods.
+I.e.
+we simulate the creation of the Palomar 5 stream for 50 different samplings of the initial conditions, with two different cases.
+The \textit{Full} simulation that includes the influence of all globualr clusters on the star particles representing Palomar 5 and its stream, as well as the \textit{vanilla} simulations in which the star-particles feel only the galaxy and Palomar 5.
+We do this for we attempt to quantify the number of gaps from globular clusters in the limit of maximum information.
+I.e.
+we are not limited by contamination of field stars, and since we are computing our own simulations, we know what a gapless stream should look like.
+
+    This comparative analysis is shown in Fig.~\ref{fig:profiles}.
+We visually compare the 2D density map in tail coordinates as well as their marginalized 1D density profiles along the length of Palomar 5's orbit.
+As described in section~\ref{sec:gap_detection}, we locate regions on the stream where the full simulations are significantly less dense than the vanilla simulations.
+We find that many streams as visible by eye.
+However, some subtle under densities only became apparent when comparing the 1D density profiles.
+All simulations and labeled gaps are presented in the Appendix~\ref{appendix:gallery_of_gaps}.
+
+    Subsequently, we followed the analysis presented in section~\ref{sec:SuspectToCulprit} to identify which globualr clusters were responsible for the gaps.
+This method used the orbit of Palomar 5 as a proxy for the stream.
+The gravitational force from all other globular clusters as computed at segments of the orbit about the position of Palomar 5 for the entire simulation.
+The top 5 strongest flybys were recorded to identify which cluster was responsible.
+This is shown in Fig.~\ref{fig:force-on-orbit}, where the top panel shows the force from the globular clusters on the orbit, while the bottom panel shows the density of the stream.
+A density threshold was imparted to quantify the length of the stream.
+The location and time of impact was cross referenced to the location of the gaps to identify which cluster was responsible.
+
+    Fig.~\ref{fig:force-on-orbit} contains some interesting information.
+Notice the periodic ribbions of force in the $(t,\tau)$ plane.
+This is due to pericenter passages where Palomar 5 is getting closer to the center of mass of the globular cluster system.
+Additionally, for the impacts of NGC104 and NGC7078, wakes can be observed in the density map.
+Another important aspect, is that the strongest peak in gravitational force does not necessarily create a gap.
+Notice how NGC5272, which was labeled with 0 to indicate that it has the greatest local maxima does not have a gap.
+This is manifold, for instance, the force needs to be modulated with the time of impact because the  important quantity that determins gap creation.
+Additionally, there is an offset of about 200~pc between the stream at the orbit, as seen when viewing the stream in tail coordinates, so peaks upon the orbit are good proxy's for the stream but are not definitive.
+We found that the top five greatest impacts accounted for all gaps, except for Sampling~014 as shown in Fig.~\ref{fig:gallery3}, whose gap from NGC~6584 corresponded to the 7th peak.
+
+    For some cases, we performed a double check by launching extra simulations only including one suspect at a time to confirm that this suspect indeed caused a gap in question, an example is shown in Fig.~\ref{fig:decomposition}.
+Note that the gap from NGC~7078 was \textit{hit twice}.
+
+    \begin{figure*}
+      \centering
+      \includegraphics[width=\linewidth]{decomposition-monte-carlo-009-with-3-gaps.png}
+      \caption{This figure decomposes the gaps shown in the middle panel of Fig.~\ref{fig:profiles} into the individual perturbuers whose cause them, as identified from Fig.~\ref{fig:force-on-orbit}.
+These simulations only include the suspected globular clusters in one-at-a-time.}
+      \label{fig:decomposition}
+    \end{figure*}
+
+ \subsection*{Gap Morphology}
+
+ \subsection*{Perturber Statistics}
+  With this analysis complete, we are able to study the population of globular clusters that could perturb Palomar 5's stream.
+
+  The clusters that impacted Palomar 5's stream in our simulations are presented in Fig.~\ref{fig:histogram_impact_time}, along with when they impacted the stream.
+First, there are few impacts that happen early on in the simulation.
+This is not surprising, as it the tidal tails are smaller and thus there is less volume to be perturbed.
+Next, is that 44/50 simulations contain a visible gap from NGC2808, all of the encounters occuring about 200~Myr ago.
+This is so much more than any other peak that we broke the scale of the histogram.
+The reason is that the uncertainty in the precise position of the clusters along their orbits grows in look back time.
+Since this encounter is recent, most of the sampled initial conditions give rise to a significant enough impact in the stream that causes a gap.
+
+  \begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{histogram_impact_time.png}
+    \caption{When the impacts occured}
+    \label{fig:histogram_impact_time}
+    \end{figure*}
+
+    We desire to search for some generalizable characterisitcs of the perturbers to see what contrasts them to globular clusters that do not impact the Pal5 Stream.
+We present three different graphs in Fig.~\ref{fig:mass_size_plane}.
+The top plot shows the angular momentm and energy plane.
+The
+
+    - to no surprise, you have to come within Palomar 5's orbital space
+
+    - trying to separate to the components of the velocity, Mass, impact parameter is not fruitful. It's better just to look at the change in momentum as given in the EQ.
+
+    - Notice that the distributions between the gap and no gap are significantly different, yet also have significant overlap.
+Perhaps these 3 parameters don't tell the whole story but also there's the factor identifies in Sanders, that I don't understand yet, that shows that a gap will not be identifiable.
+
+    \begin{figure}
+      \centering
+      \includegraphics[width=1\linewidth]{E_Lz_perturbers.png}
+      \includegraphics[width=1\linewidth]{mass_size_plane.png}
+      \includegraphics[width=1\linewidth]{impact_geometry_statistics_deltaP.png}
+      \caption{Characterisitcs of clusters that cause gaps compared to those that do not.
+\textbf{Top:} shows the energy-angular momentum space of the globular clusters present in the simulations.
+Notice that we include all initial conditions that were sampled, therefore there are 50 $\times$ 165 data points.
+The clusters of a given simulation that impacted Palomar 5 are given with a Red marker, while non-gap causing clusters are given in gray.
+\textbf{Middle:} The mass-size plane.
+Uncertainties on the mass are indicated, while no uncertainties on the radius were reported.
+\textbf{Bottom:} the imparted momentum on the star-particles from a given flyby as given in Eq.~\ref{eq:change_in_momentum}.
+The top 5 strongest flybys from each simulation are included in this dataset, those that cause gaps are flagged.}
+      \label{fig:mass_size_plane}
+      \end{figure}
+
+    \textbf{\textit{Action item:}} I need to make sure I check out real good the papers that investigate the gaps from Palomar 5.
+I wanna compare this to the known length of Palomar 5, how does this reduce the results, and are there observed gaps in Palomar 5 that are consistent with these simulations?
+
+  \section{Discussion}
+    How many of these could be observable?
+What's the main result?
+in 5 billion years of evolution, perhaps you'd only expect to 1.4 gaps, reaching up to maximum three?
+However, its possible that we dont see gaps at all for sure.
+
+    What about drawing attention to future studies?
+
+    Doing a complete census of all globular clusters?
+
+    Doing a complete census of dark matter subhaloes?
+
+    How many gaps do you expect to see in Palomar 5? Bonaca said there are about 2 of them.
+
+\bibliographystyle{aa} % or another style like plain, unsrt, etc.
+\bibliography{bibliography} % replace 'references' with the name of your .bib file
+
+\begin{appendix}
+
+  \section{Tail Coordinates} \label{appendix:TailCoordinates}
+
+  \begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{figures/along_orbit_coordinate_system-approved.png}
+    \caption{Tail coordinates}
+    \label{fig:TailCoordinates}
+  \end{figure}
+
+  \section{Gap detection and characterization methods}
     \subsubsection*{Density Maps and Profiles}
@@
 
-      \begin{figure*}
-        \centering
-        \includegraphics[width=\linewidth]{decomposition-monte-carlo-009-with-3-gaps.png}
-        \caption{This figure decomposes the gaps shown in the middle panel of Fig.~\ref{fig:profiles} into the individual perturbuers whose cause them, as identified from Fig.~\ref{fig:force-on-orbit}.
-These simulations only include the suspected globular clusters in one-at-a-time.}
-        \label{fig:decomposition}
-        \end{figure*}
-
     \subsection{Reconstruction of the impact geometry}
@@
 Thus, we compute a sample of 250 impacts, and flag those that give way to gaps.
-
-\section{Results}
-
-  \subsection*{Overview}
-    This experiment follows a monte-carlo scheme, where we sample the initial conditions of the globular clusters' masses and kinematics based on their reported observational uncertainties and repeat the experiment multiple times.
-We choose to perform the simulation 50 times to balance computational and exploring a range of possible outcomes.
-The uncertainties on the globular clusters are large enough that over the course of the 5 Gyr integration period, they could be anywhere in their phase space given their initial energy and angular momentum.
-We inspected all simulations upon their completion and present one of them in in Fig.~\ref{fig:stream_on_sky}--which we selected for its clear and prominent gaps.
-Two of these gaps are visible in galactic coordinates and become even more apparent when marginalizing over latitutde to view it's 1D profile density over longitude.
-
-    For each completed simulation, we follow the scheme laid out in the methods.
-I.e.
-we simulate the creation of the Palomar 5 stream for 50 different samplings of the initial conditions in two different environments.
-The first are the \texttt{vanilla} simulations in which the star-particles feel only the galaxy and Palomar 5.
-On the other hand, the \texttt{full} simulations includes the influence of all globualr clusters on the star-particles.
-
-    By performing a comparative analysis, we are exploring the phenomenon of the influence from all globular clusters on stellar streams in a regieme of far more complete information respect to reality.
-In this experiment, we do not deal with differentiating stream and field stars, and above all else in reality we do not have a reference reality without the effect of other globular clusters as a point of comparison.
-Additionally, we avoid the difficulty of attempting to define a gap in its own right.
-Erkal et al 2015 was able to do this pretty well by studying two parameters: the width of the gaps and the ratio of the minimum density to the density of the gaps' shoulders.
-However, this becomes impossible when departing from the idealized case of uniformly dense \& one-dimensional streams on circular orbits.
-Sanders et al 2016 extended this in action-angle space which allowed for dropping the constraint of circular orbits.
-However, I need to revist the article again because I believe they also explored non uniformly dense streams.
-Nonetheless, we avoid this all together by doing our comparative analysis.
-Since we are numerical modellers, we can play god and run nearly the exact same experiment by including or not including the globular clusters.
-Therefore, it's much easier to identify what comes from the clusters or not.
-
-    This comparative analysis is shown in Fig.~\ref{fig:profiles}.
-We visually compare the 2D density map in tail coordinates as well as their marginalized 1D density profiles along the length of Palomar 5's orbit.
-As described in section~\ref{sec:gap_detection}, we locate regions on the stream where the full simulations are significantly less dense than the vanilla simulations.
-We find that many gaps are visible by eye.
-However, some subtle under densities only became apparent when comparing the 1D density profiles.
-All simulations and labeled gaps are presented in the Appendix~\ref{appendix:gallery_of_gaps}.
-
-    Subsequently, we followed the analysis presented in section~\ref{sec:SuspectToCulprit} to identify which globualr clusters were responsible for the gaps.
-This method used the orbit of Palomar 5 as a proxy for the stream.
-This is contrary to the typical case where one uses the stream to approximate the orbit--but $\approx$ works both ways.
-The gravitational force from all other globular clusters was computed at segments of the orbit about the position of Palomar 5 for the entire simulation.
-The top 5 strongest flybys were identified and the corresponding perturbers are labeled as \textit{suspects}.
-This is shown in Fig.~\ref{fig:force-on-orbit}, where the top panel shows the force from the globular clusters on the orbit, while the bottom panel shows the density of the stream.
-A density threshold was imparted to quantify the length of the stream.
-The location and time of impact was cross referenced to the location of the gaps to identify which cluster was responsible.
-I.e.
-if the gap was observed on the leading tail, we searched for suspects with $\tau>0$.
-
-    The top panel of Fig.~\ref{fig:force-on-orbit} contains some interesting information.
-Notice the periodic ribbions of force in the $(t,\tau)$ plane.
-This is due to pericenter passages where Palomar 5 is getting closer to the center of mass of the globular cluster system.
-Additionally, for the impacts of NGC104 and NGC7078, wakes can be observed in the density map---evidence that signifies their guilt.
-Another important aspect, is that the strongest peak in gravitational force does not necessarily create a gap.
-Notice how NGC~5272, which was labeled with 0 to indicate that it has the greatest force of all local maxima, does not provoke a gap.
-The reason for this is manifold.
-First, remember that the important quantity is \textit{the change in momentum}.
-Thus, a strong force applied for a brief peroid of time may have no effect on the stream!
-Additionally, there is an offset of about 200~pc between the stream at the orbit, as seen when viewing the stream in tail coordinates.
-This demonstrates that the orbit are good proxy's for the stream but are not a one for one substitute.
-We found that the top five greatest impacts accounted for all gaps, except for Sampling~014 as shown in Fig.~\ref{fig:gallery3}, whose gap from NGC~6584 corresponded to the 7th peak.
-
-    For some cases, we performed a double check by launching extra simulations only including one suspect at a time to confirm that this suspect indeed caused a gap in question, an example is shown in Fig.~\ref{fig:decomposition}.
-Note that the gap from NGC~7078 was \textit{hit} by NGC~7078 \textit{twice}.
-
-  \subsection*{Gap Morphology}
-    We study the structure of the streams that have been impact by the globular clusters.
-The gallery of gaps is shown in the appendix, which includes the results from all of the simulations we considered in this experiment.
-
-    In general, how did the passage of these globular clusters effect the stream?
-well.
-A lot of this has been studied quite well and presented in Sanders et al 2016.
-They even consdiered different types of impactors.
-I.e.
-impactors that had different density profiles.
-They compared Hernquist, Plummer, NFW, and a truncated NFW.
-There are some minor differences between the impactors however, the conclusion is still the same.
-Impacts from small and dense perturbers such as globular clusters only effect the stream locally.
-
-    Our results are consistent with those shown previously.
-Perturber crossings only perturb the stream locally.
-Global characterisitcs of the stream do not change.
-For instance, the stream is not warmer of thicker.
-Other types of interactions have been demonstrated to do this.
-For example THIS AUTHOR showed that XXX and do YYY.
-As laid out in Erkal et al 2015, perturbers pull particles from the sides towards the impact point.
-These particles eventually move beyond the impact point and leave an under dense center and then pile up on the sides leaving an overdensity on the shoulders.
-
-    Interestingly, most of our gaps occur for the leading tail.
-8 on the trailing tail.
-21 excluding NGC~2808 are on the leading tail.
-We have to exclude NGC~2808 because the simulations are clearly correlated with one another, so trials are not independent.
-So is 8/29 consistent with a fair coin toss?
-In anycase, we argue this may be a coincidence since being hit should only depend on the size and location of the stream.
-The main difference between the tails is a $\approx$ 300-400 pc offset in $r$, yet this is negligible.
-
-    All of the gaps are oblique in tail coordinates.
-This is an artifact of using tail coordinates and arrises because we project the star-particles onto the nearest point of the orbit.
-(this is really difficult to explain with words and super simple with a picture.
-The picture being that we have three evenly spaced parallel lines that are close together.
-Then, a third point that is far away.
-)
-
-    Do our gaps grow linearly in time? what about this square root stuff?
-
-    Lastly, since the streams have width, there are some instances where the stream is not significantly impact across this width.
-For instance, the gaps from NGC~5284 in Sampling 019 and NGC~6333 in Sampling 033 are more akin to bites than gaps.
-
-  \subsection*{Perturber Statistics}
-    With this analysis complete, we are able to study the population of globular clusters that could perturb Palomar 5's stream.
-
-    The clusters that impacted Palomar 5's stream in our simulations are presented in Fig.~\ref{fig:histogram_impact_time}, along with when they impacted the stream.
-First, there are few impacts that happen early on in the simulation.
-This is not surprising, as it the tidal tails are smaller and thus there is less volume to be perturbed.
-Next, is that 44/50 simulations contain a visible gap from NGC2808, all of the encounters occuring about 200~Myr ago.
-This is so much more than any other peak that we broke the scale of the histogram so that this peak would not dwarf the significance of all others.
-The reason is that the uncertainty in the precise position of the clusters along their orbits grows in look back time.
-Since this encounter is recent, most of the sampled initial conditions give rise to a significant enough impact in the stream that causes a gap.
-
-    \begin{figure*}
-      \centering
-      \includegraphics[width=\linewidth]{histogram_impact_time.png}
-      \caption{The time of impact for each flyby that caused a gap.
-Each cluster is given a consistent and unique color.
-The scale of the histogram is broken between 8 and 42 since the gaps from NGC~2808 are in a league of their own with respect to other clusters.}
-      \label{fig:histogram_impact_time}
-      \end{figure*}
-
-    We desire to search for some generalizable characterisitcs of the perturbers to see what contrasts them to globular clusters that do not impact the Pal5 Stream.
-We present three different graphs in Fig.~\ref{fig:mass_size_plane}.
-The top plot shows the angular momentm and energy plane.
-To no surprise, the globular clusters that perturb Palomar 5 are within a band of energy enclosed the energy of Palomar 5.
-This indicates that a cluster's orbital space needs to overlap with Palomar 5's.
-There appears not to be a preference in angular momentum.
-Note that in this panel all initial conditions for the clusters are plotted.
-Clusters of gray dots indicate the same cluster.
-However, for the least bound milky way clusters, the uncertainties in distances and motions allow for a wide range of energies, prograde and retrograde orbits.
-These are Crater, Palomar X and Eridanus.
-This is consistent with Vasiliev et al 2021, yet not very significant for this study since they do not influence Palomar 5 and could have been ignored.
-
-    Next, in the middle panel of Fig.~\ref{fig:mass_size_plane} we view the mass size plane. The
-
-    - trying to separate to the components of the velocity, Mass, impact parameter is not fruitful. It's better just to look at the change in momentum as given in the EQ.
-
-    - Notice that the distributions between the gap and no gap are significantly different, yet also have significant overlap.
-Perhaps these 3 parameters don't tell the whole story but also there's the factor identifies in Sanders, that I don't understand yet, that shows that a gap will not be identifiable.
-
-    \begin{figure}
-      \centering
-      \includegraphics[width=1\linewidth]{E_Lz_perturbers.png}
-      \includegraphics[width=1\linewidth]{mass_size_plane.png}
-      \includegraphics[width=1\linewidth]{impact_geometry_statistics_deltaP.png}
-      \caption{Characterisitcs of clusters that cause gaps compared to those that do not.
-\textbf{Top:} shows the energy-angular momentum space of the globular clusters present in the simulations.
-Notice that we include all initial conditions that were sampled, therefore there are 50 $\times$ 165 data points.
-The clusters of a given simulation that impacted Palomar 5 are given with a Red marker, while non-gap causing clusters are given in gray.
-\textbf{Middle:} The mass-size plane.
-Uncertainties on the mass are indicated, while no uncertainties on the radius were reported.
-\textbf{Bottom:} the imparted momentum on the star-particles from a given flyby as given in Eq.~\ref{eq:change_in_momentum}.
-The top 5 strongest flybys from each simulation are included in this dataset, those that cause gaps are flagged.}
-      \label{fig:mass_size_plane}
-    \end{figure}
-
-    \textbf{\textit{Action item:}} I need to make sure I check out real good the papers that investigate the gaps from Palomar 5.
-I wanna compare this to the known length of Palomar 5, how does this reduce the results, and are there observed gaps in Palomar 5 that are consistent with these simulations?
-
-    \section{Discussion}
-    How many of these could be observable?
-What's the main result?
-in 5 billion years of evolution, perhaps you'd only expect to 1.4 gaps, reaching up to maximum three?
-However, its possible that we dont see gaps at all for sure.
-
-    What about drawing attention to future studies?
-
-    Doing a complete census of all globular clusters?
-
-    Doing a complete census of dark matter subhaloes?
-
-    How many gaps do you expect to see in Palomar 5? Bonaca said there are about 2 of them.
-
-\bibliographystyle{aa} % or another style like plain, unsrt, etc.
-\bibliography{bibliography} % replace 'references' with the name of your .bib file
-
-\begin{appendix}
-
-  \section{Tail Coordinates} \label{appendix:TailCoordinates}
-
-  \begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{figures/along_orbit_coordinate_system-approved.png}
-    \caption{Tail coordinates}
-    \label{fig:TailCoordinates}
-  \end{figure}
-
-  \section{Impact Geometry Statistics}
 

### 2024-10-02 paper/main.tex
proof read the methods

@@
 Specifically, for each of the 50 realizations of the Palomar 5 stream, we saved $5000$ intermediate timesteps which is a temporal resolution of 1 million years.
-To provide an estimate of the size of our simulations, we have $N_p=100 000$ number of particles, $N_\textrm{time stamps} = 5 000$ time steps, $N_\textrm{phase}=6$ phase space coordinates, $N_\textrm{sampling}=50$ monte-carlo samplings, and using single precision floating point numbers, $\textrm{Single}=4$~bytes.
-The product of the aforementioned terms results in approximately: $N_p=100 000 \times N_\textrm{time stamps}\times N_\textrm{phase}\times N_\textrm{sampling} 4 \textrm{bytes}\approx$ 600 GB of data.
+To provide an estimate of the size of our simulations, we have $N_p=100 000$ particles, $N_\textrm{time stamps} = 5 000$ time steps, $N_\textrm{phase}=6$ phase space coordinates, $N_\textrm{sampling}=50$ monte-carlo samplings, and using single precision floating point numbers, $\textrm{Single}=4$~bytes.
+The product of the aforementioned terms results in approximately: $N_p \times N_{\textrm{time stamps}}\times N_{\textrm{phase}}\times N_{\textrm{sampling}} \times 4 \textrm{bytes}\approx$ 600 GB of data.
 
-  \subsection{Gap detection and Characterization} \label{sec:gap_detection}
+  \subsection{Gap Detection and Characterization} \label{sec:gap_methods}
 
@@
 
-    To facilitate this process, we ran a series of \texttt{reference} simulations.
-In these control simulations, we used the same 50 sets of initial conditions but excluded interactions between globular clusters and omitted the gravitational influence of the clusters on the star particles within the stream--save for the host cluster Palomar 5.
-By comparing the morphological differences between these control simulations and the \texttt{full} simulations (which included cluster interactions), we could identify and assess the impact of globular clusters on the stream structure more effectively.
+    To facilitate this process, we conducted a series of \texttt{reference} simulations.
+These control simulations used the same 50 sets of initial conditions but excluded interactions with globular clusters, except for the host cluster, Palomar 5.
+In other words, the gravitational influence of all other clusters on the star particles within the stream was omitted.
+By comparing the morphological differences between these control simulations and the \texttt{full} simulations (which included cluster interactions), we could more effectively identify and assess the impact of globular clusters on the stream's structure.
+Details on gap identification are provided in Appendix~\ref{sec:gap_detection}, and the identification of corresponding perturbing globular clusters is detailed in Appendix~\ref{sec:Perturber_Identification}.
 
@@
 
-    More details on how the gaps and corresponding perturbing globular clusters were identified is provided in Appendix.~XXX.
-A first order approach of understanding the significance of an impact one needs to compute its change in momentum, which is determined by integrating the force acting on the particle over the duration of the interaction.
-A useful approximation for this change in momentum is the force at the closest approach multiplied by an estimate of the interaction time:
-    \begin{equation} 
-      \Delta v \approx \text{Force} \times \text{interaction time} = \frac{GM}{b^2} \times 2\frac{b}{v} = 2\frac{GM}{bv}, 
-      \end{equation} \label{eq:change_in_momentum}
-    where $M$ is the mass of the perturber, $m$ is the mass of the particle, $b$ is the impact parameter, $v$ is the relative velocity of the perturber with respect to the particle, and $G$ is the gravitational constant.
-We compute this for the top five cloestest approaches of all globular clusters on an orbital segment about Palomar 5 that approximates the stream through out each of the 50 simulations.
-In the end, we have a sample of 250 impacts and compare the impact geometry of those that cause gaps to those that do not.
+    Once the gaps and perturbers were identified, we sought to characterize the properties of the globular clusters that caused gaps and differentiate them from those that did not.
+Carlberg (2012) initiated the theory of gap creation in stellar streams due to gravitational flybys.
+This analysis was extended both analytically and numerically by Erkal et al.
+(2015), who studied uniform-density streams on circular orbits.
+Sanders et al.
+(2016) presented the most comprehensive theory, generalizing Erkal et al.'s work to include impacts on streams of varying densities on inclined and eccentric orbits using an action-angle approach.
+
+    The significance of an impact is measured by the change in momentum imparted to the stream stars by the perturber.
+To first order, we follow the developments from collision theory as laid out in Galactic Dynamics (Binney \& Tremaine) and similar works by John Taylor.
+In the simplest case, the perturber is considered unperturbed during the interaction, and thus has constant velocity.
+We can consider the most basic scenario of one point mass impacting another.
+A useful approximation for the change in momentum is the force at the closest approach multiplied by an estimate of the interaction time: \begin{equation} \Delta v \approx \text{Force} \times \text{interaction time} = \frac{GM}{b^2} \times 2\frac{b}{v} = 2\frac{GM}{bv}, \end{equation} where $M$ is the mass of the perturber, $b$ is the impact parameter, $v$ is the relative velocity of the perturber with respect to the particle, and $G$ is the gravitational constant.
+We compute this for the five closest approaches of all globular clusters throughout each of the 50 simulations.
+Details on how this computation was performed are given in Appendix~\ref{sec:reconstruction}.
+Ultimately, we obtained a sample of 250 impacts and compared the impact geometry of those that caused gaps with those that did not.
 
@@
 
-    Although these interactions were included, we observed no significant exchanges of orbital energy or momentum between clusters.
+    Firstly, to our knowledge this is the first time any one tried to study the interactions between the globular clusters.
+By including N~body forces on the orbits with the globular clusters, we did not observe any significant changes to orbital properties such as energy or angular momentum between any of the globular clusters in any of the simulations.
+This is an expected result given their small sizes and vast distances between them.
 
-    This experiment follows a monte-carlo scheme, where we sample the initial conditions of the globular clusters' masses and kinematics based on their corresponding uncertainties and then repeat the experiment multiple times.
-We choose to perform the simulation 50 times to balance computational cost with a high enough number to allow us to explore a range of possible outcomes.
-The uncertainties on the globular clusters are large enough that over the course of the 5 Gyr integration period, they could be anywhere allowed in their phase space given their initial energy and angular momenta.
    We inspected all simulations upon their completion and present one of them in in Fig.~\ref{fig:stream_on_sky}--which we selected for its clear and prominent gaps.
@@
 The \textit{Full} simulation that includes the influence of all globualr clusters on the star particles representing Palomar 5 and its stream, as well as the \textit{vanilla} simulations in which the star-particles feel only the galaxy and Palomar 5.
-We do this for we attempt to quantify the number of gaps from globular clusters in the limit of maximum information.
-I.e.
-we are not limited by contamination of field stars, and since we are computing our own simulations, we know what a gapless stream should look like.
 
@@
 We visually compare the 2D density map in tail coordinates as well as their marginalized 1D density profiles along the length of Palomar 5's orbit.
-As described in section~\ref{sec:gap_detection}, we locate regions on the stream where the full simulations are significantly less dense than the vanilla simulations.
+As described in section~XXX, we locate regions on the stream where the full simulations are significantly less dense than the vanilla simulations.
 We find that many streams as visible by eye.
@@
 
-    Subsequently, we followed the analysis presented in section~\ref{sec:SuspectToCulprit} to identify which globualr clusters were responsible for the gaps.
+    Subsequently, we followed the analysis presented in section~\ref{sec:Perturber_Identification} to identify which globualr clusters were responsible for the gaps.
 This method used the orbit of Palomar 5 as a proxy for the stream.
@@
 
+    We do this for we attempt to quantify the number of gaps from globular clusters in the limit of maximum information.
+I.e.
+we are not limited by contamination of field stars, and since we are computing our own simulations, we know what a gapless stream should look like.
+
     What about drawing attention to future studies?
@@
 
-  \section{Gap detection and characterization methods}
-    \subsubsection*{Density Maps and Profiles}
+  \section{Gap Detection} \label{sec:gap_detection}
     For each of the 50 simulations, we compare the final snapshots between the \texttt{vanilla} and \texttt{full} simulations using the tail coordinate system.
@@
 
-    \subsubsection*{From Suspect to Culprit} \label{sec:SuspectToCulprit}
+    \section{Perturber Identification} \label{sec:Perturber_Identification}
 
@@
 
-    \subsection{Reconstruction of the impact geometry}
+    \section{Reconstruction of the impact geometry} \label{sec:reconstruction}
 
@@
 
-      To achieve this, we begin by identifying the approximate moments of impact from the most significant clusters, as determined in the previous analysis in Sec.~\ref{sec:SuspectToCulprit}.
+      To achieve this, we begin by identifying the approximate moments of impact from the most significant clusters, as determined in the previous analysis in Sec.~\ref{sec:Perturber_Identification}.
 Then, we refine these estimates.

### 2024-10-02 paper/main.tex
reducing the results a lot

@@
 
-    For each completed simulation, we follow the scheme laid out in the methods.
-I.e.
-we simulate the creation of the Palomar 5 stream for 50 different samplings of the initial conditions, with two different cases.
-The \textit{Full} simulation that includes the influence of all globualr clusters on the star particles representing Palomar 5 and its stream, as well as the \textit{vanilla} simulations in which the star-particles feel only the galaxy and Palomar 5.
-
-    This comparative analysis is shown in Fig.~\ref{fig:profiles}.
-We visually compare the 2D density map in tail coordinates as well as their marginalized 1D density profiles along the length of Palomar 5's orbit.
-As described in section~XXX, we locate regions on the stream where the full simulations are significantly less dense than the vanilla simulations.
-We find that many streams as visible by eye.
-However, some subtle under densities only became apparent when comparing the 1D density profiles.
-All simulations and labeled gaps are presented in the Appendix~\ref{appendix:gallery_of_gaps}.
-
-    Subsequently, we followed the analysis presented in section~\ref{sec:Perturber_Identification} to identify which globualr clusters were responsible for the gaps.
-This method used the orbit of Palomar 5 as a proxy for the stream.
-The gravitational force from all other globular clusters as computed at segments of the orbit about the position of Palomar 5 for the entire simulation.
-The top 5 strongest flybys were recorded to identify which cluster was responsible.
-This is shown in Fig.~\ref{fig:force-on-orbit}, where the top panel shows the force from the globular clusters on the orbit, while the bottom panel shows the density of the stream.
-A density threshold was imparted to quantify the length of the stream.
-The location and time of impact was cross referenced to the location of the gaps to identify which cluster was responsible.
-
-    Fig.~\ref{fig:force-on-orbit} contains some interesting information.
-Notice the periodic ribbions of force in the $(t,\tau)$ plane.
-This is due to pericenter passages where Palomar 5 is getting closer to the center of mass of the globular cluster system.
-Additionally, for the impacts of NGC104 and NGC7078, wakes can be observed in the density map.
-Another important aspect, is that the strongest peak in gravitational force does not necessarily create a gap.
-Notice how NGC5272, which was labeled with 0 to indicate that it has the greatest local maxima does not have a gap.
-This is manifold, for instance, the force needs to be modulated with the time of impact because the  important quantity that determins gap creation.
-Additionally, there is an offset of about 200~pc between the stream at the orbit, as seen when viewing the stream in tail coordinates, so peaks upon the orbit are good proxy's for the stream but are not definitive.
-We found that the top five greatest impacts accounted for all gaps, except for Sampling~014 as shown in Fig.~\ref{fig:gallery3}, whose gap from NGC~6584 corresponded to the 7th peak.
-
-    For some cases, we performed a double check by launching extra simulations only including one suspect at a time to confirm that this suspect indeed caused a gap in question, an example is shown in Fig.~\ref{fig:decomposition}.
-Note that the gap from NGC~7078 was \textit{hit twice}.
+    We apply the gap detection methods described in Appendix.~\ref{sec:gap_detection} and the perturber identification methods from Appendix.~\ref{sec:Perturber_Identification}.
+We present the results from all the simulations in Appendix.~\ref{sec:gallery_of_gaps}.
+Since we identify which perturber is responsible for a given gap, we launch the more simulations in which only one globular cluster at a time is included to isolate specific gaps.
+We present this in Fig.~\ref{fig:decomposition}.
 
@@
 
+      Fig.~\ref{fig:force-on-orbit} contains some interesting information.
+Notice the periodic ribbions of force in the $(t,\tau)$ plane.
+This is due to pericenter passages where Palomar 5 is getting closer to the center of mass of the globular cluster system.
+Additionally, for the impacts of NGC104 and NGC7078, wakes can be observed in the density map.
+Another important aspect, is that the strongest peak in gravitational force does not necessarily create a gap.
+Notice how NGC5272, which was labeled with 0 to indicate that it has the greatest local maxima does not have a gap.
+This is manifold, for instance, the force needs to be modulated with the time of impact because the  important quantity that determins gap creation.
+Additionally, there is an offset of about 200~pc between the stream at the orbit, as seen when viewing the stream in tail coordinates, so peaks upon the orbit are good proxy's for the stream but are not definitive.
+We found that the top five greatest impacts accounted for all gaps, except for Sampling~014 as shown in Fig.~\ref{fig:gallery3}, whose gap from NGC~6584 corresponded to the 7th peak.
+
       The results of this analysis were compiled into a table.
@@
 
-  \section{Gallery of Gaps} \label{appendix:gallery_of_gaps}
+  \section{Gallery of Gaps} \label{sec:gallery_of_gaps}
 

### 2024-10-02 paper/main.tex
rewrote overview

@@
 
+    \begin{figure*}
+      \centering
+      \includegraphics[width=\linewidth]{decomposition-monte-carlo-009-with-3-gaps.png}
+      \caption{This figure decomposes the gaps shown in the middle panel of Fig.~\ref{fig:profiles} into the individual perturbuers whose cause them, as identified from Fig.~\ref{fig:force-on-orbit}.
+These simulations only include the suspected globular clusters in one-at-a-time.}
+      \label{fig:decomposition}
+    \end{figure*}
+
     Firstly, to our knowledge this is the first time any one tried to study the interactions between the globular clusters.
@@
 
-    \begin{figure*}
-      \centering
-      \includegraphics[width=\linewidth]{decomposition-monte-carlo-009-with-3-gaps.png}
-      \caption{This figure decomposes the gaps shown in the middle panel of Fig.~\ref{fig:profiles} into the individual perturbuers whose cause them, as identified from Fig.~\ref{fig:force-on-orbit}.
-These simulations only include the suspected globular clusters in one-at-a-time.}
-      \label{fig:decomposition}
-    \end{figure*}
-
- \subsection*{Gap Morphology}
-
- \subsection*{Perturber Statistics}
-  With this analysis complete, we are able to study the population of globular clusters that could perturb Palomar 5's stream.
-
-  The clusters that impacted Palomar 5's stream in our simulations are presented in Fig.~\ref{fig:histogram_impact_time}, along with when they impacted the stream.
-First, there are few impacts that happen early on in the simulation.
-This is not surprising, as it the tidal tails are smaller and thus there is less volume to be perturbed.
-Next, is that 44/50 simulations contain a visible gap from NGC2808, all of the encounters occuring about 200~Myr ago.
-This is so much more than any other peak that we broke the scale of the histogram.
-The reason is that the uncertainty in the precise position of the clusters along their orbits grows in look back time.
-Since this encounter is recent, most of the sampled initial conditions give rise to a significant enough impact in the stream that causes a gap.
+    We present a summary of the number of gaps found in the simulation of the gallery of gaps from Appendix.~\ref{sec:gallery_of_gaps} in Fig.~\ref{fig:histogram_impact_time}, where the number of impacts and the time of impact is reported.
 
@@
       \includegraphics[width=\linewidth]{histogram_impact_time.png}
-    \caption{When the impacts occured}
+      \caption{The time of impact for all gap causing flybys summed over all 50 simulations.
+Each perturbing cluster is label and is color consistent.
+Time is presented in integration units.
+Note that 1~s~kpc~km~$^{-1}$ is roughly 1~Gyr.
+Also, the y-axis is broken to show the total number of encounters from NGC~2808 yet not dwarf the scale from the other interactions.}
       \label{fig:histogram_impact_time}
       \end{figure*}
+
+    A striking result from the simulations is that more than half of the gaps are caused by NGC~2808.
+Upon inspecting the figures in the appendix, we notice that these gaps are of similar size and occur in nearly the same position in the leading tail across multiple simulations.
+The size of a gap is a function of time, and recent encounters leave limited time for the gap to grow.
+As shown in Fig.~\ref{fig:histogram_impact_time}, the gaps from NGC~2808 do occur at about 200 Myr ago.
+It is no coincidence that a recent fly occurs in most of the simulations.
+As we look further back in time, present-day uncertainties allow for much broader variation in the phase space positions of the globular clusters.
+However, the uncertainties associated with NGC 2808 and Palomar 5 are sufficiently low that the recent flyby of NGC 2808 occurs in 44 out of 50 simulated cases.
+The flybys do not happen exactly the same way in each simulation, the range in contrast of the gaps shown in the appendix illustrates that some encounters were closer, leading to more significant impacts.
+
+    Next, we notice that most of the gaps occur on the leading tail.
+However, we believe this to be a coincidence.
+Stars escape from the globular cluster at the lagrange points at roughly equivalent rates.
+The two substantial differences are the orbital phase, and that the leading tail is about ~400~pc closers to the galactic center than the trailing tail.
+However, this is peanuts and negligible compared to the orbital radius of 8-12~kpc.
+Therefore, we expect the probability of a gap appearing in the leading or trailing tail to be equal and the presence of one gap should not influence the likelihood of another.
+
+    Yet, the observation of only 8 gaps in the trailing tail out of 73 total gaps is inconsistent with this expectation.
+However, 44 of these gaps are associated with NGC~2808.
+As shown in Fig.\ref{fig:histogram_impact_time}, all of these gaps result from the same recent encounter.
+We thus discard NGC~2808 for the current analysis, as its gaps violate the assumption of independent events.
+After excluding these, we are left with 21 gaps in the leading tail and 8 in the trailing tail across the remaining simulations.
+By integrating the binomial distribution, we compute that the probability of having at least 8 successes or fewer given 29 trials is approximately 1.2\%.
+While this outcome is unlikely it is not impossible.
+
+    Overall, in 50 simulations we observe 73 gaps.
+The distribution of number of simulations with a give number of gaps is zero: 4, one: 25, two: 16, three: 4, four:1.
+If we regard NGC2808 as a fluke, then we are stil left with 29 gaps.
+Thus, in our simulations we expect [0.6-1.5] gaps.
+
+ \subsection*{Perturber Statistics}
 

### 2024-10-02 paper/main.tex
passed over results with GPT

@@
 
-    Firstly, to our knowledge this is the first time any one tried to study the interactions between the globular clusters.
-By including N~body forces on the orbits with the globular clusters, we did not observe any significant changes to orbital properties such as energy or angular momentum between any of the globular clusters in any of the simulations.
-This is an expected result given their small sizes and vast distances between them.
+    To our knowledge, this is the first study to investigate interactions between globular clusters in this context.
+We included N-body forces on the orbits with the globular clusters, yet did not observe significant changes to orbital properties, such as energy or angular momentum, between any of the globular clusters in any simulation.
+This result is expected given the small sizes and vast distances between the clusters.
 
-    We inspected all simulations upon their completion and present one of them in in Fig.~\ref{fig:stream_on_sky}--which we selected for its clear and prominent gaps.
-Two of these gaps are visible in galactic coordinates and become even more apparent when marginalizing over latitutde to view it's 1D profile density over longitude.
+    We inspected all simulations upon their completion and present one particularly clear example shown in Fig.~\ref{fig:stream_on_sky} which we selected for its prominent gaps.
+Two of these gaps are visible in galactic coordinates and become even more apparent when marginalizing over latitude to view their 1D density profile over longitude.
 
-    We apply the gap detection methods described in Appendix.~\ref{sec:gap_detection} and the perturber identification methods from Appendix.~\ref{sec:Perturber_Identification}.
-We present the results from all the simulations in Appendix.~\ref{sec:gallery_of_gaps}.
-Since we identify which perturber is responsible for a given gap, we launch the more simulations in which only one globular cluster at a time is included to isolate specific gaps.
-We present this in Fig.~\ref{fig:decomposition}.
+    We applied the gap detection methods described in Appendix~\ref{sec:gap_detection}, along with the perturber identification methods from Appendix~\ref{sec:Perturber_Identification}.
+Results from all simulations are presented in Appendix~\ref{sec:gallery_of_gaps}.
+Once we identified the perturbers responsible for each gap, we conducted additional simulations to verify correct identification of the perturbers.
+These additional contain only the suspected clusters and were launched one-at-a-time to isolate the effects of a specific clusters.
+These results are shown in Fig.~\ref{fig:decomposition}.
 
@@
 Each perturbing cluster is label and is color consistent.
-Time is presented in integration units.
-Note that 1~s~kpc~km~$^{-1}$ is roughly 1~Gyr.
-Also, the y-axis is broken to show the total number of encounters from NGC~2808 yet not dwarf the scale from the other interactions.}
+The time axis is given in simulation units, with 1~s~kpc~km~$^{-1}$ corresponding to roughly 1~Gyr.
+Note that the y-axis is broken to accommodate the large number of encounters from NGC2808, without overshadowing the other interactions.}
       \label{fig:histogram_impact_time}
@@
 
-    A striking result from the simulations is that more than half of the gaps are caused by NGC~2808.
-Upon inspecting the figures in the appendix, we notice that these gaps are of similar size and occur in nearly the same position in the leading tail across multiple simulations.
-The size of a gap is a function of time, and recent encounters leave limited time for the gap to grow.
-As shown in Fig.~\ref{fig:histogram_impact_time}, the gaps from NGC~2808 do occur at about 200 Myr ago.
-It is no coincidence that a recent fly occurs in most of the simulations.
-As we look further back in time, present-day uncertainties allow for much broader variation in the phase space positions of the globular clusters.
-However, the uncertainties associated with NGC 2808 and Palomar 5 are sufficiently low that the recent flyby of NGC 2808 occurs in 44 out of 50 simulated cases.
-The flybys do not happen exactly the same way in each simulation, the range in contrast of the gaps shown in the appendix illustrates that some encounters were closer, leading to more significant impacts.
+    One striking result is that more than half of the gaps observed are caused by NGC2808.
+Upon examining the figures in the appendix, we note that these gaps are of similar size and occur in almost the same position in the leading tail across most simulations.
+This is no coincidence, as the size of a gap is a function of time, and recent encounters allow only limited time for the gap to develop.
+As shown in Fig.\ref{fig:histogram_impact_time}, the gaps caused by NGC2808 result from a flyby that occurred approximately 200 Myr ago.
+It is expected that a recent flyby appears in most simulations, as uncertainties grow when we look further back in time, allowing broader variation in the phase space positions of the clusters.
+However, due to relatively low uncertainties associated with NGC2808 and Palomar 5 compared to the integration time, the recent flyby of NGC~2808 occurs in 44 out of 50 simulations.
+Although the flybys do not occur identically in each simulation, the range in gap contrast, shown in the appendix, indicates that closer encounters lead to more significant impacts.
 
-    Next, we notice that most of the gaps occur on the leading tail.
-However, we believe this to be a coincidence.
-Stars escape from the globular cluster at the lagrange points at roughly equivalent rates.
-The two substantial differences are the orbital phase, and that the leading tail is about ~400~pc closers to the galactic center than the trailing tail.
-However, this is peanuts and negligible compared to the orbital radius of 8-12~kpc.
-Therefore, we expect the probability of a gap appearing in the leading or trailing tail to be equal and the presence of one gap should not influence the likelihood of another.
+    Additionally, we observe that most gaps occur in the leading tail.
+We believe this to be a coincidence, as stars escape from the globular cluster's Lagrange points at roughly equal rates.
+The two main differences are the orbital phase and that the leading tail is about 400~pc closer to the Galactic center than the trailing tail.
+However, this difference is negligible compared to the orbital radius of 8-12kpc.
+As such, we expect the likelihood of a gap appearing in either tail to be equal, with the presence of one gap not influencing the probability of another.
 
@@
 
-    Overall, in 50 simulations we observe 73 gaps.
-The distribution of number of simulations with a give number of gaps is zero: 4, one: 25, two: 16, three: 4, four:1.
-If we regard NGC2808 as a fluke, then we are stil left with 29 gaps.
-Thus, in our simulations we expect [0.6-1.5] gaps.
+    Overall, across 50 simulations, we observe 73 gaps.
+The distribution of simulations with a given number of gaps is as follows: 0 gaps in 4 simulations, 1 gap in 25 simulations, 2 gaps in 16 simulations, 3 gaps in 4 simulations, and 4 gaps in 1 simulation.
+Excluding NGC~2808, we are left with 29 gaps, leading to an expected range of 0.6–1.5 gaps per simulation.
 

### 2024-10-04 paper/main.tex
retouch the overview again

@@
 These results are shown in Fig.~\ref{fig:decomposition}.
-
-    We present a summary of the number of gaps found in the simulation of the gallery of gaps from Appendix.~\ref{sec:gallery_of_gaps} in Fig.~\ref{fig:histogram_impact_time}, where the number of impacts and the time of impact is reported.
+A summary of which perturber caused a gap and when is presented in Fig.~\ref{fig:histogram_impact_time}.
 
@@
 
-    One striking result is that more than half of the gaps observed are caused by NGC2808.
-Upon examining the figures in the appendix, we note that these gaps are of similar size and occur in almost the same position in the leading tail across most simulations.
-This is no coincidence, as the size of a gap is a function of time, and recent encounters allow only limited time for the gap to develop.
-As shown in Fig.\ref{fig:histogram_impact_time}, the gaps caused by NGC2808 result from a flyby that occurred approximately 200 Myr ago.
-It is expected that a recent flyby appears in most simulations, as uncertainties grow when we look further back in time, allowing broader variation in the phase space positions of the clusters.
-However, due to relatively low uncertainties associated with NGC2808 and Palomar 5 compared to the integration time, the recent flyby of NGC~2808 occurs in 44 out of 50 simulations.
-Although the flybys do not occur identically in each simulation, the range in gap contrast, shown in the appendix, indicates that closer encounters lead to more significant impacts.
+    One striking result is that more than half of the observed gaps are caused by NGC 2808, which had a flyby about 200 Myr ago.
+In fact, 44 of the 50 simulations show a gap from this interaction.
+Upon examining their morphology (Appendix~\ref{sec:gallery_of_gaps}), we observe that these gaps are consistently similar in size and location within the leading tail across most simulations.
+This is no coincidence.
+As demonstrated by Sanders and Erkal, gap creation is a dispersion phenomenon, where stars affected by a perturber acquire slightly different orbital frequencies, causing them to gradually spread apart over time.
+Thus, smaller gaps are indicative of more recent encounters.
 
-    Additionally, we observe that most gaps occur in the leading tail.
-We believe this to be a coincidence, as stars escape from the globular cluster's Lagrange points at roughly equal rates.
-The two main differences are the orbital phase and that the leading tail is about 400~pc closer to the Galactic center than the trailing tail.
+    It is also expected that a recent flyby would appear in most simulations, as uncertainties in a cluster's precise location increase as we look further back in time.
+In general, after enough time passes, different orbital solutions \textit{phase mix}, meaning a body could be found anywhere in its phase space given its Hamiltonian and initial conditions.
+The characteristic time scale is proportional to its orbital period.
+However, since the encounter between NGC~2808 and Palomar~5's stream occurred less than one orbital period ago for either cluster, the orbital solutions remain relatively similar, causing the flyby to appear consistently in most simulations.
+
+    Next, we observe that most gaps occur in the leading tail.
+We believe this to be a coincidence, as we expect there to be no physical motivator to cause more gaps to occur in one tail than another.
+First, stars escape from the globular cluster's L1 and L2 Lagrange points at roughly equal rates, leading the same size and number of stars per tail.
+The two main differences each tail is that the particles find themselves at different locations on the orbital phase at a given point and time.
+The second is that the leading tail is about 400~pc closer to the Galactic center than the trailing tail.
 However, this difference is negligible compared to the orbital radius of 8-12kpc.
@@
 While this outcome is unlikely it is not impossible.
+Additionally, by studying Fig.~\ref{fig:histogram_impact_time}, we notice that, although in much smaller numbers compared to NGC~2808, other pertubers cause gaps at the same time accross different simulations.
+These could also be violating the assumption of independent events.
 
@@
 The distribution of simulations with a given number of gaps is as follows: 0 gaps in 4 simulations, 1 gap in 25 simulations, 2 gaps in 16 simulations, 3 gaps in 4 simulations, and 4 gaps in 1 simulation.
-Excluding NGC~2808, we are left with 29 gaps, leading to an expected range of 0.6–1.5 gaps per simulation.
+Excluding NGC~2808, we are left with 29 gaps, leading to an expected range of 0.6-1.5 gaps per simulation.
 

### 2024-10-04 paper/main.tex
making colors of perturbers consistent

@@
 In general, after enough time passes, different orbital solutions \textit{phase mix}, meaning a body could be found anywhere in its phase space given its Hamiltonian and initial conditions.
-The characteristic time scale is proportional to its orbital period.
+The characteristic phase-mixing time scale is proportional to its orbital period.
 However, since the encounter between NGC~2808 and Palomar~5's stream occurred less than one orbital period ago for either cluster, the orbital solutions remain relatively similar, causing the flyby to appear consistently in most simulations.
@@
 After excluding these, we are left with 21 gaps in the leading tail and 8 in the trailing tail across the remaining simulations.
-By integrating the binomial distribution, we compute that the probability of having at least 8 successes or fewer given 29 trials is approximately 1.2\%.
+By integrating the binomial distribution, we compute that the probability of having 8 successes or fewer given 29 trials is approximately 1.2\%.
 While this outcome is unlikely it is not impossible.
@@
 
-    \begin{figure}
+    \begin{figure*}
       \centering
-      \includegraphics[width=1\linewidth]{E_Lz_perturbers.png}
-      \includegraphics[width=1\linewidth]{mass_size_plane.png}
-      \includegraphics[width=1\linewidth]{impact_geometry_statistics_deltaP.png}
+      \includegraphics[width=0.45\linewidth]{E_Lz_perturbers.png}
+      \includegraphics[width=0.45\linewidth]{mass_size_plane.png}
+
+      \includegraphics[width=0.45\linewidth]{impact_geometry_statistics_deltaP.png}
+      \includegraphics[width=0.45\linewidth]{impact_stats_phase_space.png}
       \caption{Characterisitcs of clusters that cause gaps compared to those that do not.
@@
       \label{fig:mass_size_plane}
-      \end{figure}
+      \end{figure*}
 

### 2024-10-11 paper/main.tex
added paolas edits

@@
 
+\newcommand{\paola}[1]{\textcolor{magenta}{{#1}}}
+
 \begin{document}
@@
 
-  The existence of dark matter is supported by substantial observational evidence on scales larger than individual galaxies.
-This work aims to contribute to the search for dark matter on sub-galactic scales, specifically within the Milky Way.
-First proposed by [Author Name] in [Year], the concept of dark matter arose to explain the unusually high velocities of stars in dwarf galaxies—velocities that exceed those predicted by the visible mass exerting gravitational forces.
-Dark matter's influence extends to cosmological scales as well, as demonstrated by the bending of light within cosmic web filaments due to the unseen mass.
-Despite significant progress, the true nature of dark matter remains elusive, and scientists continue to search for evidence of its existence beyond its gravitational effects on cosmic and extragalactic scales.
-A key focus today is detecting dark matter on sub-galactic scales, such as through the study of stellar streams.
+  \paola{Postulated about a century ago, to explain the large velocity dispersions of galaxies inside clusters \citep[see][]{zwicky33}\footnote{The English translation of the original paper is available at \citet{zwicky_EN}} and the rotation curves of disk galaxies as well \citep{freeman70, rubin70}, different  astronomical measurements conducted over the last century point to the existence of dark matter, from temperature fluctuations in the Cosmic Microwave Background  to weak and strong gravitational lensing \citep[see, among many, ][]{planck20, mandelbaum06, hsueh20}.
+Dark matter constitutes the dominant ($\sim$~85\%) contributor to matter in the Universe \citep{planck20}, and shapes the building-up of its structures over time \citep{padmanabhan93, bertschinger98, springel05}.}
 
-  Stellar streams, long and thin structures formed by the tidal disruption of globular clusters or dwarf galaxies orbiting a host galaxy, present a promising method for probing dark matter distribution.
+  Despite significant progress, the true nature of dark matter remains elusive, and scientists continue to search for evidence of its existence beyond its gravitational effects on cosmic and galactic scales \citep[see][ for a review]{arbey21}.
+A key focus today is detecting dark matter on sub-galactic scales, because at these scales different cosmological scenarios predict different lumpiness of the halos \citep{madau08, zavala19}.
+\paola{One of the most promising tools to set limit to the halo lumpiness, or granularity, and hence to the number and mass density of the so-called sub-halos, is provided by the study of} stellar streams.
+
+  Stellar streams, several-kpc long structures formed by the tidal disruption of globular clusters or dwarf galaxies orbiting a host galaxy, present \paola{indeed} a promising method for probing dark matter distribution.
 These tidal forces arise due to differential gravitational pulls across extended objects, causing stars farther from the galactic center to lag behind while those closer are pulled away.
-This stretching creates two tidal tails that trace the cluster's orbit.
-Theoretical predictions of this phenomenon existed long before the first observation in 1995 by \citet{carl_j_grillmair_globular_1995}, who detected stars beyond the tidal radii of globular clusters.
-Subsequent work by \citet{odenkirchen_detection_2001} further demonstrated that Palomar 5's stars not only extended beyond the tidal radius but also formed distinct tidal tails.
+This stretching creates two tidal tails that trace the cluster's orbit, \paola{unless in the closest vicinity to the object \citep{montuori07}}.
+Numerical predictions of this phenomenon exist \paola{since already the '70s of the last century \citep[see, for example][]{keenan75}, well}  before the first detections of \paola{stellar streams around some Galactic globular clusters made by}  \citet{carl_j_grillmair_globular_1995}, \paola{and the nearly contemporary discovery of the Sagittarius stellar stream, the closest (to the Sun) example of a stream emerging from a dwarf satellite currently interacting with the Milky Way \citep{mateo98, newberg02, majewski03, martinez04}.} \paola{Subsequent works \citep[see][]{lehmann97, testa00, leon00, siegel01, lee03} extended Grillmair's findings to other globular cluster streams, until the discovery made by \citet{odenkirchen01,odenkirchen02, odenkirchen03} of long and thin tails outside the Palomar~5 globular cluster, which were shown to be extended for more than $10^\circ$ degrees in the sky.
+These early works stimulated a vigorous search for stellar streams in the Galaxy in the following years \citep{grillmair06a, grillmair06b, belokurov06, grillmair09}, which has now entered a new dimension thanks to the Gaia astrometric mission \citep{prousti16}, which allows to look for these structures coupling photometry, astrometry and (when possible) spectroscopic  information.
+Currently about 20 GCs have been found to possess streams around them \citep[see][ for a summary of these discoveries]{mateu23}, while some streams do not have any clear association with existing GCs, pointing to the early presence of a population of clusters  now completely dissolved in the Galactic field \citep{ibata24, martin22, wan20}.}
 
-  Over the past two decades, theoretical and numerical studies have shown that stellar streams can constrain the gravitational field of host galaxies, particularly the Milky Way.
-By applying Poisson's equation, the density distribution and gravitational potential can be separated into different components, including the dark matter halo.
-This halo, often modeled as spherical, is the largest component and crucial to explaining the constant rotational velocity of galaxies.
-For example, \citet{varghese2011stellar} demonstrated how stellar streams can infer the mass and scale parameters of haloes using various amounts of observational data, from basic right ascension and declination to full six-dimensional phase space information.
-Meanwhile, Bonaca et al.
-employed an information-theoretic approach, identifying the orbits and configurations of stellar streams that yield the most information about the galactic potential.
+  \paola{One of the interests in studying stellar streams, is that they } can constrain the gravitational field of their host galaxies, particularly the Milky Way.
+For example, \citet{varghese2011stellar} demonstrated how stellar streams can infer the mass and scale parameters of \paola{dark matter} haloes using various amounts of observational data, from basic right ascension and declination to full six-dimensional phase space information.
+\paola{The use of single streams for constraining the Galactic potential has led, however, to some ambiguous and not converging results over time.
+For example, \citet{law10} made use of the Sagittarius stream to infer that the dark matter halo of our Galaxy has a triaxial shape \citep[but see also][]{helmi04, johnston05, law05}; \citet{bovy16} concluded that the dark matter halo of our Galaxy is nearly spherical at the radii of the Palomar~5 and GD~1 streams.
+} Meanwhile, \citet{bonaca14} employed an information-theoretic approach, identifying the orbits and configurations of stellar streams that yield the most information about the galactic potential.
 
-  The Gaia Data Release 2 (DR2) and DR3 revolutionized the observational study of stellar streams, enabling significant advances in the field.
+  \paola{As previously stated, }  the Gaia mission, and its Data Releases 2 and 3 \citep{gaia18, gaia23}, revolutionized the observational study of stellar streams, \paola{unraveling a population of stellar streams much more numerous than that previously known \citep[see, for example][]{malhan18}}, enabling significant advances in the field.
 Malhan et al.
@@
 The best-observed streams, such as GD-1 and Palomar 5, contain between 1,000 and 3,000 stars.
-Leveraging this dataset, Ibata et al.
-(2024) constructed the most accurate model of the Milky Way to date based on stellar stream data.
+Leveraging this dataset, \citet{ibata24} constructed the most accurate model of the Milky Way to date based on a set of about 90 streams.
 
@@
 These clumps, or \textit{subhalos}, are predicted to follow a mass distribution with a power-law slope slightly shallower than -2.0.
-To date, the smallest observed dark matter halo was detected through gravitational lensing in an Einstein ring, as described by Vingetti et al.
-(20XX).
-However, some models predict that dark matter clumps could exist down to the mass of a planet like Jupiter (I hear this at conferences, is it true?
-where are the papers that suggest this?).
+To date, the smallest observed dark matter halo was detected through gravitational lensing in an Einstein ring, as described by \citet{vegetti12}.
+However, some models predict that dark matter clumps could exist down to the mass of Earth-like planets at least \citep[see][ and discussion in \citet{amorisco21}]{green05, wang20}.
 

### 2024-10-11 paper/main.tex
rewrote the edit

@@
 
-  \paola{Postulated about a century ago, to explain the large velocity dispersions of galaxies inside clusters \citep[see][]{zwicky33}\footnote{The English translation of the original paper is available at \citet{zwicky_EN}} and the rotation curves of disk galaxies as well \citep{freeman70, rubin70}, different  astronomical measurements conducted over the last century point to the existence of dark matter, from temperature fluctuations in the Cosmic Microwave Background  to weak and strong gravitational lensing \citep[see, among many, ][]{planck20, mandelbaum06, hsueh20}.
+  % \paola{Postulated about a century ago, to explain the large velocity dispersions of galaxies inside clusters \citep[see][]{zwicky33}\footnote{The English translation of the original paper is available at \citet{zwicky_EN}} and the rotation curves of disk galaxies as well \citep{freeman70, rubin70}, different  astronomical measurements conducted over the last century point to the existence of dark matter, from temperature fluctuations in the Cosmic Microwave Background  to weak and strong gravitational lensing \citep[see, among many, ][]{planck20, mandelbaum06, hsueh20}.
 Dark matter constitutes the dominant ($\sim$~85\%) contributor to matter in the Universe \citep{planck20}, and shapes the building-up of its structures over time \citep{padmanabhan93, bertschinger98, springel05}.}
+
+  Over a century ago \citet{zwicky33}\footnote{The English translation of the original paper is available at \citet{zwicky_EN}} postulated the existence of dark matter to explain the excessively large velocity dispersions of galaxies within their galaxy clusters.
+Since this time, mounting evidence has been assembled to support this claim.
+\citet{freeman70} and \citet{rubin70} provided some of the first measurements of the rotation curves of disk galaxies that are too fast moving to be driven by ordinary matter alone.
+The temperature fluctuations in the Cosmic Microwave Background radation are well explained by the presence of dark matter, which is well studied and documented by the Planck Mission, a review of which is presented by \citet{planck20}.
+The location and mass of dark matter on the large scale can be inferred and spatially mapped through strong and weak gravitational lensing, as demonstrated in \citet{planck20, mandelbaum06, hsueh20}.
+Dark matter contributes $\sim$~85~\% of the matter in the universe \citep{planck20} and is crucial for driving the hierarchical growth of ordinary in the universe.
+See \citet{padmanabhan93} for XXX, \citet{bertschinger98} for XXX and \citet{springel05} for XXX.
 

### 2024-10-18 paper/main.tex
updated the opening paragraph
resarch rabbit and zotero bib's DONT WORK. Double author. Changing all to ADS. Big pain. Much sad

--- 2024-10-04 paper/main.tex
@@
 
+\newcommand{\paola}[1]{\textcolor{magenta}{{#1}}}
+\newcommand{\salvatore}[1]{\textcolor{green}{{#1}}}
+
 \begin{document}
@@
 
-  The existence of dark matter is supported by substantial observational evidence on scales larger than individual galaxies.
-This work aims to contribute to the search for dark matter on sub-galactic scales, specifically within the Milky Way.
-First proposed by [Author Name] in [Year], the concept of dark matter arose to explain the unusually high velocities of stars in dwarf galaxies—velocities that exceed those predicted by the visible mass exerting gravitational forces.
-Dark matter's influence extends to cosmological scales as well, as demonstrated by the bending of light within cosmic web filaments due to the unseen mass.
-Despite significant progress, the true nature of dark matter remains elusive, and scientists continue to search for evidence of its existence beyond its gravitational effects on cosmic and extragalactic scales.
-A key focus today is detecting dark matter on sub-galactic scales, such as through the study of stellar streams.
+  \salvatore{I couldn't find the bib tex for zwicky's original paper}
+  Over a century ago, \citet{1933AcHPh...6..110Z} postulated the existence of dark matter to explain the excessively large velocity dispersions of galaxies within their galaxy clusters.
+Since then, mounting evidence has been assembled to support this claim.
+\citet{1970IAUS...38..351F,1970ApJ...161..802F} elaborated on the dynamics of disk galaxies and briefly acknoledged in the appendix of his paper that the rotation curves of NGC 300 and M~33 cannot be fully explained by disk models and that there must be the same amount of mass or more that is undetected and beyond the extend of the galaxies to explain their rotation curves.
+Concurrently \citet{1970ApJ...159..379R} studied the rotation of Andromenda and observed it to be flat and not decreasing as one would expect from a potential with more mass concentration in the center.
+Eventually, astronomers returned to Zwicky's non-luminous matter to explain these fast-rotating disks, as discussed in \citet{1983SciAm.248f..96R}'s review.
 
-  Stellar streams, long and thin structures formed by the tidal disruption of globular clusters or dwarf galaxies orbiting a host galaxy, present a promising method for probing dark matter distribution.
+  \salvatore{I see why we cite freeman70 now.
+there are two comments in the appendix stating that there is as much mass beyond the optical limit as there is within it for two different galaxoes.
+NGC 300 and M 33.
+It's not clear to me why we cite Rubin 70.
+The rotation curve is flat, to me implies a logarithmic potential and hence we need dark matter.
+However, I don't get any impression of missing matter when I read the text.}
+
+  Beyond the galactic scale, dark matter is essential for explaining many cosmological phenomena.
+For instance, the Planck mission \citep{2020A&A...641A...1P} was designed to study the Cosmic Microwave Background (CMB), whose spectrum closely follows that of a black body with a peak temperature in the millimeter wavelength range, corresponding to approximately 2.725~K.
+However, there are slight temperature variations on the order of one part in ten thousand.
+Notably, the colder regions of the CMB temperature map correlate with whole-sky gravitational lensing maps \citep{2020A&A...641A...8P}.
+This is because photons lose energy as they escape from gravitational potentials.
+Additionally, galaxy clusters reside within these deep gravitational wells.
+The $\Lambda$ Cold Dark Matter model accounts for these observations almost perfectly, proposing that dark matter exists and that its stochastic quantum fluctuations led to density variations.
+These variations propagated as the universe expanded, shaping the large-scale structure and distribution of galaxies we observe today.
+
+  As laid out in \citet{2021PrPNP.11903865A}'s review, despite significant progress the true nature of dark matter remains elusive and scientists continue to search for evidence of its existence beyond its gravitational effects on cosmic and galactic scales.
+A key focus today is detecting dark matter on sub-galactic scales, because at these scales different cosmological scenarios predict different lumpiness of the halos.
+For instance, \citet{2008ApJ...679.1260M} compared the Via Lactea simulations to some that they made themselves and showed that you can have a large variation for the number of dark matter subhalos within a galaxy.
+\citet{2019Galax...7...81Z} provides a review for the different proposed mechanisms of forming dark matter haloes from perturbation theory and deviations from a uniformly dense medium and also disccses techniques and limitations for modelign such physical phenomena.
+\citet{2019Galax...7...81Z} rightly states that until particles are discovered $\Lambda$-CDM remains a hypothesis.
+Since the true nature of dark matter is elusive.
+The possibilities for sub-galactic behaviour and structure of dark matter are vaste.
+However, as proposed by Carlberg in 2012, a most promising tool to set limits on the number and mass density of the sub-halos, are stellar streams.
+
+  Stellar streams, several-kpc long structures formed by the tidal disruption of globular clusters or dwarf galaxies orbiting a host galaxy.
 These tidal forces arise due to differential gravitational pulls across extended objects, causing stars farther from the galactic center to lag behind while those closer are pulled away.
-This stretching creates two tidal tails that trace the cluster's orbit.
-Theoretical predictions of this phenomenon existed long before the first observation in 1995 by \citet{carl_j_grillmair_globular_1995}, who detected stars beyond the tidal radii of globular clusters.
-Subsequent work by \citet{odenkirchen_detection_2001} further demonstrated that Palomar 5's stars not only extended beyond the tidal radius but also formed distinct tidal tails.
+This stretching creates two tidal tails that trace the cluster's orbit, unless in the closest vicinity to the object \citep{montuori07}.
+Numerical predictions of this phenomenon exist since the 1970s \citep[see, for example][]{keenan75}.
+These predictions occured well before the first detections which were made by \citet{carl_j_grillmair_globular_1995} around some Galactic globular clusters.
+Interestingly, this work was a nearly contemporary discovery of the Sagittarius stellar stream by \citet{ibata1994} which is the closest example of a stream emerging from a dwarf satellite currently interacting with the Milky Way.
+Subsequent studies\footnote{\citep[see][for more subsequent observation detections of tidal debris, i.e.
+globular cluster stars beyond the tidal radius.]{lehmann97, testa00, leon00, siegel01, lee03}} extended Grillmair's findings to other globular cluster streams, until the discovery made by \citet{2001ApJ...548L.165O,odenkirchen02, odenkirchen03} of long and thin tails outside the Palomar~5 globular cluster, which were shown to be extended for more than $10^\circ$ degrees in the sky.
+These early works stimulated a vigorous search for stellar streams in the Galaxy in the following years \citep{grillmair06a, grillmair06b, belokurov06, grillmair09}.
+The review of Bonaca properly asserts that we have envolved into a new area of stellar stream astronomy Gaia astrometric mission \citep{prousti16}.
 
-  Over the past two decades, theoretical and numerical studies have shown that stellar streams can constrain the gravitational field of host galaxies, particularly the Milky Way.
-By applying Poisson's equation, the density distribution and gravitational potential can be separated into different components, including the dark matter halo.
-This halo, often modeled as spherical, is the largest component and crucial to explaining the constant rotational velocity of galaxies.
-For example, \citet{varghese2011stellar} demonstrated how stellar streams can infer the mass and scale parameters of haloes using various amounts of observational data, from basic right ascension and declination to full six-dimensional phase space information.
-Meanwhile, Bonaca et al.
-employed an information-theoretic approach, identifying the orbits and configurations of stellar streams that yield the most information about the galactic potential.
+  Gaia's characterization of billions of stars in the Milky Way allows us to look for these structures coupling photometry, astrometry and spectroscopy.
+The development and deployment of the Streamfinder algorithm has led to the detection of about 90 stellar streams in the Milky Way.
+\citet{mateu23} provides a summary of these streams and tabulates that about 20 are associated to modern day globular clusters.
+Thus, currently there are about 70 progenitor-less streams and about 145 stream-less clusteres.
+The solo-streams can be explained by a larger globular cluster population whose memebers have been disolved due to two body relaxation and tidal stripping.
+For instance, Z Wan 2020 reports that the Phenoix stream is dynamically consistent from a globular cluster \footnote{see bovy for a description on the difference between dwarf and cluster streams} yet far more metal poor than any cluster ever observed, galactic or otherwise.
+This implies that the stream is from a completely disolved globular cluster which comes from a population of clusters that have not survived until modern times.
+Nicolas Martin in 2022 showed that many very metal poor streams exist in the Milky Way.
+On the other hand, it is possible that some globular clusters do not loose a significant number of stars to create tidal tails.
+For instance, Vitral and Bouldrini demonstrated that clusters existing in dark matter subhaloes can place the tidal radius far beyond their half mass radii and sheild stars from escape.
+Another factor is indeed detectability.
+Additionally, there are over XXX cluster in the internal regional of the galaxy, a region overwhich stream finder was not applied since the high stellar density implicates a computational cost that is currently far to expensive.
+Also, Balbinot demonstrates that internal dynamics preferentially kicks out small stars, which have lower luminosities and can make detection more difficult.
 
-  The Gaia Data Release 2 (DR2) and DR3 revolutionized the observational study of stellar streams, enabling significant advances in the field.
-Malhan et al.
-developed the StreamFinder algorithm, which identified 90 stellar streams within the Milky Way, some of which have been followed up by other instruments and surveys.
-The best-observed streams, such as GD-1 and Palomar 5, contain between 1,000 and 3,000 stars.
-Leveraging this dataset, Ibata et al.
-(2024) constructed the most accurate model of the Milky Way to date based on stellar stream data.
+  One of the interests in studying stellar streams, is that they can constrain the gravitational field of their host galaxies, particularly the Milky Way.
+For example, \citet{varghese2011stellar} demonstrated how stellar streams can infer the mass and scale parameters of \paola{dark matter} haloes using various amounts of observational data, from basic right ascension and declination to full six-dimensional phase space information.
+\citet{bonaca14} reviewed this concept from an information-theoretic point of view, identifying which orbits and configurations of stellar streams that yield the most information about the galactic potential.
+However, the use of single streams for constraining the Galactic potential has led, however, to some ambiguous and not converging results over time.
+For example, \citet{law10} made use of the Sagittarius stream to infer that the dark matter halo of our Galaxy has a triaxial shape \citep[but see also][]{helmi04, johnston05, law05}; \citet{bovy16} concluded that the dark matter halo of our Galaxy is nearly spherical at the radii of the Palomar~5 and GD~1 streams.
+Nonetheless, a recent goal of stellar stream astronomy was realized by \citet{ibata24} constructed the most accurate model of the Milky Way to date based on a set of about 90 streams.
 
@@
 These clumps, or \textit{subhalos}, are predicted to follow a mass distribution with a power-law slope slightly shallower than -2.0.
-To date, the smallest observed dark matter halo was detected through gravitational lensing in an Einstein ring, as described by Vingetti et al.
-(20XX).
-However, some models predict that dark matter clumps could exist down to the mass of a planet like Jupiter (I hear this at conferences, is it true?
-where are the papers that suggest this?).
+To date, the smallest observed dark matter halo was detected through gravitational lensing in an Einstein ring, as described by \citet{vegetti12}.
+However, some models predict that dark matter clumps could exist down to the mass of Earth-like planets at least \citep[see][ and discussion in \citet{amorisco21}]{green05, wang20}.
 
   \citet{rodrigo_ibata_uncovering_2002} first suggested that dark matter subhalos could influence stellar streams by diffusing their orbital elements.
-Later, \citet{r_g_carlberg_pal_2012} expanded this idea, proposing that subhalos might create gaps in stellar streams during flyby encounters, where a subhalo approaches closely enough to a segment of a stream and significantly changes the orbits of the closest stars.
-This theory was further explored through a probabilistic model of encounter rates based on the expected subhalo distribution.
+Later, \citet{r_g_carlberg_pal_2012} expanded this idea, proposing that subhalos could create gaps in stellar streams during flyby encounters, where a subhalo approaches closely enough to a segment of a stream and significantly changes the orbits of the closest stars.
+Carlberg 2013 further explored this idea through a probabilistic model of encounter rates based on the expected subhalo distribution.
 \citet{bonaca_spur_2018} provided observational evidence for this idea, identifying an underdensity in the GD-1 stream that could not be explained by known objects, such as globular clusters, and, interestingly enough, was inconsistent with the $\Lambda$CDM mass-size relationship from M et al 20XX.

### 2024-10-21 paper/main.tex
the bib should work now

@@
 These tidal forces arise due to differential gravitational pulls across extended objects, causing stars farther from the galactic center to lag behind while those closer are pulled away.
-This stretching creates two tidal tails that trace the cluster's orbit, unless in the closest vicinity to the object \citep{montuori07}.
-Numerical predictions of this phenomenon exist since the 1970s \citep[see, for example][]{keenan75}.
-These predictions occured well before the first detections which were made by \citet{carl_j_grillmair_globular_1995} around some Galactic globular clusters.
-Interestingly, this work was a nearly contemporary discovery of the Sagittarius stellar stream by \citet{ibata1994} which is the closest example of a stream emerging from a dwarf satellite currently interacting with the Milky Way.
+This stretching creates two tidal tails that trace the cluster's orbit, unless in the closest vicinity to the object \citep{2007ApJ...659.1212M}.
+Numerical predictions of this phenomenon exist since the 1970s \citep[see, for example][]{1975AJ.....80..290K}.
+These predictions occured well before the first detections which were made by \citet{1995AJ....109.2553G} around some Galactic globular clusters.
+Interestingly, this work was a nearly contemporary discovery of the Sagittarius stellar stream by \citet{1994Natur.370..194I} which is the closest example of a stream emerging from a dwarf satellite currently interacting with the Milky Way.
 Subsequent studies\footnote{\citep[see][for more subsequent observation detections of tidal debris, i.e.
-globular cluster stars beyond the tidal radius.]{lehmann97, testa00, leon00, siegel01, lee03}} extended Grillmair's findings to other globular cluster streams, until the discovery made by \citet{2001ApJ...548L.165O,odenkirchen02, odenkirchen03} of long and thin tails outside the Palomar~5 globular cluster, which were shown to be extended for more than $10^\circ$ degrees in the sky.
-These early works stimulated a vigorous search for stellar streams in the Galaxy in the following years \citep{grillmair06a, grillmair06b, belokurov06, grillmair09}.
-The review of Bonaca properly asserts that we have envolved into a new area of stellar stream astronomy Gaia astrometric mission \citep{prousti16}.
+globular cluster stars beyond the tidal radius.]{1997A&A...320..776L, 2000A&A...356..127T, 2000A&A...359..907L, 2001AAS...19910906S, 2003AJ....126..815L}} extended Grillmair's findings to other globular cluster streams, until the discovery made by \citet{2001ApJ...548L.165O,2002AAS...200.1001O, 2003AJ....126.2385O} of long and thin tails outside the Palomar~5 globular cluster, which were shown to be extended for more than $10^\circ$ degrees in the sky.
+These early works stimulated a vigorous search for stellar streams in the Galaxy in the following years \citep{2006ApJ...643L..17G, 2006ApJ...637L..29B, 2006ApJ...637L..29B, 2009ApJ...693.1118G}.
+The review of Bonaca properly asserts that we have envolved into a new area of stellar stream astronomy Gaia astrometric mission \citep{2016A&A...595A...1G}.
 
@@
 The development and deployment of the Streamfinder algorithm has led to the detection of about 90 stellar streams in the Milky Way.
-\citet{mateu23} provides a summary of these streams and tabulates that about 20 are associated to modern day globular clusters.
+\citet{2023MNRAS.520.5225M} provides a summary of these streams and tabulates that about 20 are associated to modern day globular clusters.
 Thus, currently there are about 70 progenitor-less streams and about 145 stream-less clusteres.
@@
   One of the interests in studying stellar streams, is that they can constrain the gravitational field of their host galaxies, particularly the Milky Way.
-For example, \citet{varghese2011stellar} demonstrated how stellar streams can infer the mass and scale parameters of \paola{dark matter} haloes using various amounts of observational data, from basic right ascension and declination to full six-dimensional phase space information.
-\citet{bonaca14} reviewed this concept from an information-theoretic point of view, identifying which orbits and configurations of stellar streams that yield the most information about the galactic potential.
+For example, \citet{2011MNRAS.417..198V} demonstrated how stellar streams can infer the mass and scale parameters of \paola{dark matter} haloes using various amounts of observational data, from basic right ascension and declination to full six-dimensional phase space information.
+\citet{2018ApJ...867..101B} reviewed this concept from an information-theoretic point of view, identifying which orbits and configurations of stellar streams that yield the most information about the galactic potential.
 However, the use of single streams for constraining the Galactic potential has led, however, to some ambiguous and not converging results over time.
-For example, \citet{law10} made use of the Sagittarius stream to infer that the dark matter halo of our Galaxy has a triaxial shape \citep[but see also][]{helmi04, johnston05, law05}; \citet{bovy16} concluded that the dark matter halo of our Galaxy is nearly spherical at the radii of the Palomar~5 and GD~1 streams.
-Nonetheless, a recent goal of stellar stream astronomy was realized by \citet{ibata24} constructed the most accurate model of the Milky Way to date based on a set of about 90 streams.
+For example, \citet{2010ApJ...718.1128L} made use of the Sagittarius stream to infer that the dark matter halo of our Galaxy has a triaxial shape \citep[but see also][]{2004MNRAS.351..643H, johnston05, 2005ApJ...619..807L}; \citet{2016ApJ...833...31B} concluded that the dark matter halo of our Galaxy is nearly spherical at the radii of the Palomar~5 and GD~1 streams.
+Nonetheless, a recent goal of stellar stream astronomy was realized by \citet{2024ApJ...967...89I} constructed the most accurate model of the Milky Way to date based on a set of about 90 streams.
 
@@
 These clumps, or \textit{subhalos}, are predicted to follow a mass distribution with a power-law slope slightly shallower than -2.0.
-To date, the smallest observed dark matter halo was detected through gravitational lensing in an Einstein ring, as described by \citet{vegetti12}.
-However, some models predict that dark matter clumps could exist down to the mass of Earth-like planets at least \citep[see][ and discussion in \citet{amorisco21}]{green05, wang20}.
+To date, the smallest observed dark matter halo was detected through gravitational lensing in an Einstein ring, as described by \citet{2012Natur.481..341V}.
+However, some models predict that dark matter clumps could exist down to the mass of Earth-like planets at least \citep[see][ and discussion in \citet{2021arXiv211101148A}]{2005JCAP...08..003G, wang20}.
 
-  \citet{rodrigo_ibata_uncovering_2002} first suggested that dark matter subhalos could influence stellar streams by diffusing their orbital elements.
-Later, \citet{r_g_carlberg_pal_2012} expanded this idea, proposing that subhalos could create gaps in stellar streams during flyby encounters, where a subhalo approaches closely enough to a segment of a stream and significantly changes the orbits of the closest stars.
+  \citet{2002MNRAS.332..915I} first suggested that dark matter subhalos could influence stellar streams by diffusing their orbital elements.
+Later, \citet{2012ApJ...748...20C} expanded this idea, proposing that subhalos could create gaps in stellar streams during flyby encounters, where a subhalo approaches closely enough to a segment of a stream and significantly changes the orbits of the closest stars.
 Carlberg 2013 further explored this idea through a probabilistic model of encounter rates based on the expected subhalo distribution.
-\citet{bonaca_spur_2018} provided observational evidence for this idea, identifying an underdensity in the GD-1 stream that could not be explained by known objects, such as globular clusters, and, interestingly enough, was inconsistent with the $\Lambda$CDM mass-size relationship from M et al 20XX.
+\citet{2019ApJ...880...38B} provided observational evidence for this idea, identifying an underdensity in the GD-1 stream that could not be explained by known objects, such as globular clusters, and, interestingly enough, was inconsistent with the $\Lambda$CDM mass-size relationship from M et al 20XX.
 
@@
 An additional and unique aspect of our study is that we focus on modeling real stellar streams and not just pure simulated streams.
-We focus on the Palomar 5 stream, which is well-studied, has observable gaps (bonaca et al 20XX), and is highly visible due to its position above the galactic disk.
+We focus on the Palomar 5 stream, which is well-studied, has observable gaps \citep{2020ApJ...889...70B}, and is highly visible due to its position above the galactic disk.
 We present a result from one of our simulations in Fig.~\ref{fig:stream_on_sky}, and the following sections describe our modeling of the stream, the gravitational interactions with the galaxy and other globular clusters, and the statistical analysis of perturbations affecting the stream.
@@
 
-      \citet{sanders2016dynamics} extended the analysis of Erkal et al.
-(2015), demonstrating that action-angle variables provide a useful coordinate system for analyzing stream evolution, as actions are conserved and the angles associated with the stream's growth evolve linearly over time.
+      \citet{2016MNRAS.457.3817S} extended the analysis of \citet{2015MNRAS.450.1136E}, demonstrating that action-angle variables provide a useful coordinate system for analyzing stream evolution, as actions are conserved and the angles associated with the stream's growth evolve linearly over time.
 Although we became aware of this work only after completing our analysis, we note that $\tau$ is a suitable approximation and behaves similarly to the angle variable corresponding to the azimuthal action: $\tau \approx \theta_{\phi,i} - \theta_{\phi,\text{GC}}$.

### 2024-10-21 paper/main.tex
bib is growing!

@@
 \newcommand{\salvatore}[1]{\textcolor{green}{{#1}}}
+\newcommand{\marcook}[1]{\textcolor{green}{{#1}}}
+\newcommand{\marcoko}[1]{\textcolor{red}{{#1}}}
+\newcommand{\qmar}[1]{\textcolor{brown}{{#1}}}
 
@@
 
-  As laid out in \citet{2021PrPNP.11903865A}'s review, despite significant progress the true nature of dark matter remains elusive and scientists continue to search for evidence of its existence beyond its gravitational effects on cosmic and galactic scales.
+  As laid out in \citet{2021PrPNP.11903865A}'s review, despite significant progress, the true nature of dark matter remains elusive and scientists continue to search for evidence of its existence beyond its gravitational effects on cosmic and galactic scales.
 A key focus today is detecting dark matter on sub-galactic scales, because at these scales different cosmological scenarios predict different lumpiness of the halos.
@@
   Stellar streams, several-kpc long structures formed by the tidal disruption of globular clusters or dwarf galaxies orbiting a host galaxy.
+They are one of the evidences for the hierarchical nature of the galaxy formation as predicted by the $\Lambda CDM$ model \citep{white91}.
 These tidal forces arise due to differential gravitational pulls across extended objects, causing stars farther from the galactic center to lag behind while those closer are pulled away.
@@
 The solo-streams can be explained by a larger globular cluster population whose memebers have been disolved due to two body relaxation and tidal stripping.
-For instance, Z Wan 2020 reports that the Phenoix stream is dynamically consistent from a globular cluster \footnote{see bovy for a description on the difference between dwarf and cluster streams} yet far more metal poor than any cluster ever observed, galactic or otherwise.
+For instance, Z Wan 2020(???) reports that the Phenoix stream is dynamically consistent from a globular cluster yet far more metal poor than any cluster ever observed, galactic or otherwise.
 This implies that the stream is from a completely disolved globular cluster which comes from a population of clusters that have not survived until modern times.
-Nicolas Martin in 2022 showed that many very metal poor streams exist in the Milky Way.
+
+  \citet{2022MNRAS.516.5331M} showed that many very metal poor streams exist in the Milky Way.
 On the other hand, it is possible that some globular clusters do not loose a significant number of stars to create tidal tails.
-For instance, Vitral and Bouldrini demonstrated that clusters existing in dark matter subhaloes can place the tidal radius far beyond their half mass radii and sheild stars from escape.
+For instance, \citet{2022A&A...667A.112V} demonstrated that clusters existing in dark matter subhaloes can place the tidal radius far beyond their half mass radii and sheild stars from escape.
 Another factor is indeed detectability.
-Additionally, there are over XXX cluster in the internal regional of the galaxy, a region overwhich stream finder was not applied since the high stellar density implicates a computational cost that is currently far to expensive.
-Also, Balbinot demonstrates that internal dynamics preferentially kicks out small stars, which have lower luminosities and can make detection more difficult.
+Additionally, there are mant cluster in the internal regional of the galaxy, a region overwhich \texttt{streamfinder} was not applied since the high stellar density implicates a computational cost that is currently far to expensive.
+Also, \citet{2018MNRAS.474.2479B} demonstrates that internal dynamics preferentially kicks out small stars first, which have lower luminosities.
+Earlier escape times places larger distances to the cluster and make detection more difficult.
 
   One of the interests in studying stellar streams, is that they can constrain the gravitational field of their host galaxies, particularly the Milky Way.
-For example, \citet{2011MNRAS.417..198V} demonstrated how stellar streams can infer the mass and scale parameters of \paola{dark matter} haloes using various amounts of observational data, from basic right ascension and declination to full six-dimensional phase space information.
+ompared to the measurement of the HI rotation curve, stellar streams offer the opportunity to investigate the potential of the host galaxy over a wide range of distances, reaching the outermost regions of the halo.
+For example, \citet{2011MNRAS.417..198V} demonstrated how stellar streams can infer the mass and scale parameters of dark matter haloes using various amounts of observational data, from basic right ascension and declination to full six-dimensional phase space information.
 \citet{2018ApJ...867..101B} reviewed this concept from an information-theoretic point of view, identifying which orbits and configurations of stellar streams that yield the most information about the galactic potential.
@@
 These clumps, or \textit{subhalos}, are predicted to follow a mass distribution with a power-law slope slightly shallower than -2.0.
-To date, the smallest observed dark matter halo was detected through gravitational lensing in an Einstein ring, as described by \citet{2012Natur.481..341V}.
+To date, the smallest observed dark matter halo was detected through gravitational lensing in an Einstein ring and had a mass of $10^8$ solar masses, as described by \citet{2012Natur.481..341V}.
 However, some models predict that dark matter clumps could exist down to the mass of Earth-like planets at least \citep[see][ and discussion in \citet{2021arXiv211101148A}]{2005JCAP...08..003G, wang20}.
@@
 Carlberg 2013 further explored this idea through a probabilistic model of encounter rates based on the expected subhalo distribution.
-\citet{2019ApJ...880...38B} provided observational evidence for this idea, identifying an underdensity in the GD-1 stream that could not be explained by known objects, such as globular clusters, and, interestingly enough, was inconsistent with the $\Lambda$CDM mass-size relationship from M et al 20XX.
+\citet{2019ApJ...880...38B} provided observational evidence for this idea, identifying an underdensity in the GD-1 stream that could not be explained by known objects, such as globular clusters, and, interestingly enough, was inconsistent with the $\Lambda$CDM mass-size relationship presented in \cite{2017MNRAS.466.4974M}.
 
@@
 If globular clusters do indeed create such gaps, this factor must be considered when using stellar streams to detect dark matter.
-An additional and unique aspect of our study is that we focus on modeling real stellar streams and not just pure simulated streams.
+An additional and unique aspect of our study is that we focus on modeling real stellar streams.
 We focus on the Palomar 5 stream, which is well-studied, has observable gaps \citep{2020ApJ...889...70B}, and is highly visible due to its position above the galactic disk.
@@
 We begin by extracting the phase space properties (positions, velocities) as well as the masses and half-mass radii of 165 globular clusters from the galactic globular cluster catalog by Baumgardt and Vasiliev et al.
-We handle the uncertainties through a monte-carlo approach.
-The catalog provides uncertainties for mass, distances, line of sight velocities, proper motions and includes a covariance term between the proper motions.
-We repeate the whole experiment 50 times, for each of which we vary the initial conditions of each cluster by sampling its five-dimensional gaussian uncertaintiy distribution.
-However, for the first simulation, which we name \texttt{Sampling 000}, we use the most probable values for the initial conditions.
+We handle the uncertainties on the measured phase space parameters through a monte-carlo approach.
+The catalog provides uncertainties for mass, distances, line of sight velocities, proper motions and includes a covariance term between the proper motions---uncertainties for the right ascension, declination, and the characterisitcs radii are not available.
+We repeate simulation the creation of Palomar 5's tidail tails 50 times, for each of which we vary the initial conditions of each cluster by sampling its five-dimensional gaussian uncertaintiy distribution.
+For the first simulation, which we name \texttt{Sampling 000}, we use the most probable values for the initial conditions.
 
@@
 The position of the Sun was set to $(x_\odot,y_\odot,z_\odot) = (-8.34,0,0.027)$~kpc, vertical position above the disk was taken from Chen et al 2001 and the galacto-radial distance was taken from Reid et al 2014.
-These transformations were performed using the \texttt{astropy} package.
-(for the next paper, I would also like to sample these uncertainties)
+These transformations were performed using the \texttt{astropy} \citep{2013A&A...558A..33A}.
 
-    For the galactic potential, we employed the second static model proposed by Pouliasis et al.
-(2017), which consists of a superposition of a thin disk, thick disk, and a dark matter halo.
-The masses and scale lengths are provided in Table 1 of Ferrone et al 2023.
-This model is time-independent over the course of our simulations.
+    For the galactic potential, we used the second static model from \citet{2017A&A...598A..66P}, a superposition of a thin disk, thick disk, and dark matter halo, with masses and scale lengths provided in Table 1 of \citet{2023A&A...673A..44F}.
+This model is time-independent throughout our simulations.
+To balance data volume and computation time, we use only one galactic potential, which should suffice.
+\citet{2021MNRAS.505.5978V} found that only a few outer globular clusters are strongly affected by different potential models, with kinematic uncertainties being the dominant factor in orbital solution differences.
+Similarly, \citet{2024MNRAS.528.5189G} generated a globular cluster mass-loss catalog using seven different potential models and found their debris distributions to be model-independent.
+While the clusters' positions may depend on the model, we assert that interaction rates and stream formation are unaffected by the choice of galactic model.
 
-    The primary methodological departure from Ferrone et al.
-(2023) is that now we include globular cluster interactions.
+    The primary methodological departure from \citet{2023A&A...673A..44F} is that now we include globular cluster interactions.
 Instead of treating the clusters as point masses that only experience the gravitational potential of the galaxy, we now compute N-body interactions between the clusters during the backward integration over 5 billion years.
@@
 
-    At the computed positions 5 Gyr ago, each globular cluster was gifted a system of 100,000 particles, sampled from a Plummer distribution to represent the stellar content of the cluster.
-For eaching sampling of the initial conditions, a different sampling of the positions of the 100,000 star particles was performed.
-This is becuase, we sampled the uncertainties on the mass in this experiment contraty to Ferrone et al 2023.
+    At their computed positions 5 Gyr ago, each globular cluster was assigned 100,000 particles sampled from a Plummer distribution to represent its stellar content.
+For each set of initial conditions, the particle positions were resampled to account for uncertainties in mass, which is an improvement from \citet{2023A&A...673A..44F}.
+Since globular clusters are older than 8 Gyr, their most massive stars have already gone supernova, leaving primarily low-mass stars, typically under one solar mass (CITATION?).
+Given that clusters usually have masses between $10^5-10^6$ solar masses, 100~000 particles is sufficient to represent the distribution of tidal debris.
 
@@
 
-    The integration was performed using a leapfrog integrator.
-The timestep was chosen to be sufficiently small to ensure both the conservation of energy and that the integration error over 5 Gyr was negligible compared to the scale radius of the globular clusters.
+    The integration was performed using a leapfrog integrator (add citation).
+The timestep size was chosen to be based on two critera.
+First, to ensure the conservation of energy to XXX.
+The second, is the drift in position due to numerical error over 5 Gyr was negligible compared to the scale radius of the globular clusters, which is about 5~pc.
 Further details on the timestep selection are discussed in the appendix.
@@
     At the end of the simulations, some gaps in the stellar stream are so prominent that they can be identified visually without any sophisticated quantitative analysis.
-However, upon closer inspection, more subtle gaps emerge, highlighting the need for quantitative methods to systematically detect and assess these features.
+The analysis of the morphology of the tidal tails of Palomar~5 is carried out by comparing simulations with globular clusters (called \texttt{full}) and without them (called \texttt{reference}).
+A preliminary evaluation is performed visually, followed by a quantitative analysis of the differences between the density profiles in the two mentioned cases.
 
-    We initially sought an automated, quantitative approach from signal and image processing that could reliably detect all gaps and evaluate their significance.
+    % We initially sought an automated, quantitative approach from signal and image processing that could reliably detect all gaps and evaluate their significance.
 Unfortunately, despite extensive efforts, an ideal method remained elusive.
@@
 
-    To facilitate this process, we conducted a series of \texttt{reference} simulations.
-These control simulations used the same 50 sets of initial conditions but excluded interactions with globular clusters, except for the host cluster, Palomar 5.
+    The \texttt{reference} simulations used the same 50 sets of initial conditions but excluded interactions with globular clusters, except for the host cluster, Palomar 5.
 In other words, the gravitational influence of all other clusters on the star particles within the stream was omitted.
-By comparing the morphological differences between these control simulations and the \texttt{full} simulations (which included cluster interactions), we could more effectively identify and assess the impact of globular clusters on the stream's structure.
+By comparing the morphological differences between these \texttt{reference} simulations and the \texttt{full} simulations, we could more effectively identify and assess the impact of globular clusters on the stream's structure.
 Details on gap identification are provided in Appendix~\ref{sec:gap_detection}, and the identification of corresponding perturbing globular clusters is detailed in Appendix~\ref{sec:Perturber_Identification}.
@@
 This system allows for consistent morphological comparisons by aligning the stream with the cluster's orbit.
-The coordinate system is based on the work of Dehnen et al.
-(2004) and is detailed in Appendix~\ref{appendix:TailCoordinates}.
+The coordinate system is based on the work of \citet{2004AJ....127.2753D} and is detailed in Appendix~\ref{appendix:TailCoordinates}.
 Briefly, the x' coordinate represents the position of a particle along the orbit relative to the globular cluster—whether it is ahead or behind the cluster.
@@
     Once the gaps and perturbers were identified, we sought to characterize the properties of the globular clusters that caused gaps and differentiate them from those that did not.
-Carlberg (2012) initiated the theory of gap creation in stellar streams due to gravitational flybys.
-This analysis was extended both analytically and numerically by Erkal et al.
-(2015), who studied uniform-density streams on circular orbits.
-Sanders et al.
-(2016) presented the most comprehensive theory, generalizing Erkal et al.'s work to include impacts on streams of varying densities on inclined and eccentric orbits using an action-angle approach.
+\citet{2013ApJ...775...90C} initiated the theory of gap creation in stellar streams due to gravitational flybys.
+This analysis was extended both analytically and numerically by \citet{2015MNRAS.450.1136E}, who studied uniform-density streams on circular orbits.
+\citet{2016MNRAS.457.3817S} presented the most comprehensive theory, generalizing Erkal et al.'s work to include impacts on streams of varying densities on inclined and eccentric orbits using an action-angle approach.
 
     The significance of an impact is measured by the change in momentum imparted to the stream stars by the perturber.
-To first order, we follow the developments from collision theory as laid out in Galactic Dynamics (Binney \& Tremaine) and similar works by John Taylor.
-In the simplest case, the perturber is considered unperturbed during the interaction, and thus has constant velocity.
-We can consider the most basic scenario of one point mass impacting another.
 A useful approximation for the change in momentum is the force at the closest approach multiplied by an estimate of the interaction time: \begin{equation} \Delta v \approx \text{Force} \times \text{interaction time} = \frac{GM}{b^2} \times 2\frac{b}{v} = 2\frac{GM}{bv}, \end{equation} where $M$ is the mass of the perturber, $b$ is the impact parameter, $v$ is the relative velocity of the perturber with respect to the particle, and $G$ is the gravitational constant.
@@
 
-    One striking result is that more than half of the observed gaps are caused by NGC 2808, which had a flyby about 200 Myr ago.
-In fact, 44 of the 50 simulations show a gap from this interaction.
-Upon examining their morphology (Appendix~\ref{sec:gallery_of_gaps}), we observe that these gaps are consistently similar in size and location within the leading tail across most simulations.
-This is no coincidence.
-As demonstrated by Sanders and Erkal, gap creation is a dispersion phenomenon, where stars affected by a perturber acquire slightly different orbital frequencies, causing them to gradually spread apart over time.
-Thus, smaller gaps are indicative of more recent encounters.
+    NGC~2808 impacts Palomar 5's stream about 200 Myr ago in 44 out of 50 simulations, creating a small gap at a similar position in each case.
+The gap size aligns with gap formation as a dispersion process, where stars gradually spread apart due to slight differences in orbital frequencies.
+Smaller gaps indicate more recent encounters.
+Since this interaction occurred less than one orbital period ago, the orbital solutions remain similar, producing consistent results across simulations.
 
-    It is also expected that a recent flyby would appear in most simulations, as uncertainties in a cluster's precise location increase as we look further back in time.
-In general, after enough time passes, different orbital solutions \textit{phase mix}, meaning a body could be found anywhere in its phase space given its Hamiltonian and initial conditions.
-The characteristic phase-mixing time scale is proportional to its orbital period.
-However, since the encounter between NGC~2808 and Palomar~5's stream occurred less than one orbital period ago for either cluster, the orbital solutions remain relatively similar, causing the flyby to appear consistently in most simulations.
-
-    Next, we observe that most gaps occur in the leading tail.
-We believe this to be a coincidence, as we expect there to be no physical motivator to cause more gaps to occur in one tail than another.
-First, stars escape from the globular cluster's L1 and L2 Lagrange points at roughly equal rates, leading the same size and number of stars per tail.
-The two main differences each tail is that the particles find themselves at different locations on the orbital phase at a given point and time.
-The second is that the leading tail is about 400~pc closer to the Galactic center than the trailing tail.
-However, this difference is negligible compared to the orbital radius of 8-12kpc.
-As such, we expect the likelihood of a gap appearing in either tail to be equal, with the presence of one gap not influencing the probability of another.
-
-    Yet, the observation of only 8 gaps in the trailing tail out of 73 total gaps is inconsistent with this expectation.
-However, 44 of these gaps are associated with NGC~2808.
-As shown in Fig.\ref{fig:histogram_impact_time}, all of these gaps result from the same recent encounter.
-We thus discard NGC~2808 for the current analysis, as its gaps violate the assumption of independent events.
-After excluding these, we are left with 21 gaps in the leading tail and 8 in the trailing tail across the remaining simulations.
-By integrating the binomial distribution, we compute that the probability of having 8 successes or fewer given 29 trials is approximately 1.2\%.
-While this outcome is unlikely it is not impossible.
-Additionally, by studying Fig.~\ref{fig:histogram_impact_time}, we notice that, although in much smaller numbers compared to NGC~2808, other pertubers cause gaps at the same time accross different simulations.
-These could also be violating the assumption of independent events.
+    Most gaps appear in the leading tail, which we consider coincidental, as stars escape the cluster's L1 and L2 points at similar rates, and the small distance difference between tails' galactocentric distance is negligible.
+We expect gaps to be equally likely in both tails.
+However, only 8 of the 73 gaps occur in the trailing tail, with 44 linked to NGC2808's flyby.
+After removing these, there are 21 gaps in the leading tail and 8 in the trailing.
+By the binomial distribution, the probability of observing up to 8 successes in 29 trials, given a 50\% chance of success, is 1.2\%—unlikely, but possible.
+Additionally, other perturbers impact at consistent times, which may also violate the assumption of independent events, as seen with NGC2808.
 
@@
       With the perturbers identified, we perform statistics in the pursuit of understanding what conditions are necessary for a globular cluster and a stream to have in order for a gap to form.
-We turn to impact theory, which in its simplest form as presented in works such as Binney and Tremaine, John Taylor, and Galaxiesbook, considers two particles: one stationary and the other moving past it.
+We turn to impact theory, which in its simplest form as presented in works such as \citet{2008gady.book.....B}, John Taylor, and Galaxiesbook, considers two particles: one stationary and the other moving past it.
 The distance between the two particles at their closest approach is known as the impact parameter.

### 2024-10-21 paper/main.tex
more citations and little changes

@@
 % \usepackage[draft]{graphicx}
-% \usepackage{txfonts}
 \usepackage{amsmath}
+\usepackage{hyperref}
+\usepackage{xcolor}
 
@@
 
+\definecolor{darkgreen}{rgb}{0.0, 0.5, 0.0}
+
 \newcommand{\paola}[1]{\textcolor{magenta}{{#1}}}
-\newcommand{\salvatore}[1]{\textcolor{green}{{#1}}}
+\newcommand{\salvatore}[1]{\textcolor{darkgreen}{{#1}}}
 \newcommand{\marcook}[1]{\textcolor{green}{{#1}}}
@@
 
-  \salvatore{I couldn't find the bib tex for zwicky's original paper}
   Over a century ago, \citet{1933AcHPh...6..110Z} postulated the existence of dark matter to explain the excessively large velocity dispersions of galaxies within their galaxy clusters.
@@
 Eventually, astronomers returned to Zwicky's non-luminous matter to explain these fast-rotating disks, as discussed in \citet{1983SciAm.248f..96R}'s review.
-
 \salvatore{I see why we cite freeman70 now.
@@
 However, there are slight temperature variations on the order of one part in ten thousand.
-Notably, the colder regions of the CMB temperature map correlate with whole-sky gravitational lensing maps \citep{2020A&A...641A...8P}.
-This is because photons lose energy as they escape from gravitational potentials.
-Additionally, galaxy clusters reside within these deep gravitational wells.
-The $\Lambda$ Cold Dark Matter model accounts for these observations almost perfectly, proposing that dark matter exists and that its stochastic quantum fluctuations led to density variations.
+Notably, the colder regions of the CMB temperature map correlate with whole-sky gravitational lensing maps \citep{2020A&A...641A...8P}--since the photons lost energy as they escaped the gravitational wells.
+The $\Lambda$ Cold Dark Matter ($\Lambda$-CDM) model accounts for these observations almost perfectly, proposing that dark matter exists and that its distribution originated from stochastic quantum fluctuations within the first second of the universe led to density variations.
 These variations propagated as the universe expanded, shaping the large-scale structure and distribution of galaxies we observe today.
@@
 A key focus today is detecting dark matter on sub-galactic scales, because at these scales different cosmological scenarios predict different lumpiness of the halos.
-For instance, \citet{2008ApJ...679.1260M} compared the Via Lactea simulations to some that they made themselves and showed that you can have a large variation for the number of dark matter subhalos within a galaxy.
+For instance, \citet{2008ApJ...679.1260M} computed cosmological simulations and compared the Via Lactea simulations \citep{2008ApJ...679.1260M} and demonstrated (\salvatore{I don't quite understand Madau 2008.
+The main result is that based on the assumptions you can produce a variance in the number of subhaloes that are consistent with global properties of a milky-way like galaxy?
+I need to discuss this with another human}).
 \citet{2019Galax...7...81Z} provides a review for the different proposed mechanisms of forming dark matter haloes from perturbation theory and deviations from a uniformly dense medium and also disccses techniques and limitations for modelign such physical phenomena.
@@
 (2023), which we present here for completeness.
-We begin by extracting the phase space properties (positions, velocities) as well as the masses and half-mass radii of 165 globular clusters from the galactic globular cluster catalog by Baumgardt and Vasiliev et al.
-We handle the uncertainties on the measured phase space parameters through a monte-carlo approach.
+We begin by extracting the phase space properties (positions, velocities) as well as the masses and half-mass radii of 165 globular clusters from the galactic globular cluster catalog by \cite{2021MNRAS.505.5957B}.\footnote{The Baumgardt catalog has been assembled across a series of works, see: \cite{2020PASA...37...46B,2019MNRAS.482.5138B,2018MNRAS.478.1520B} as well.
+The catalog can be found on the World Wide Web at \href{https://people.smp.uq.edu.au/HolgerBaumgardt/globular/}{https://people.smp.uq.edu.au/HolgerBaumgardt/globular/}.} We handle the uncertainties on the measured phase space parameters through a monte-carlo approach.
 The catalog provides uncertainties for mass, distances, line of sight velocities, proper motions and includes a covariance term between the proper motions---uncertainties for the right ascension, declination, and the characterisitcs radii are not available.
@@
 The position of the Sun was set to $(x_\odot,y_\odot,z_\odot) = (-8.34,0,0.027)$~kpc, vertical position above the disk was taken from Chen et al 2001 and the galacto-radial distance was taken from Reid et al 2014.
-These transformations were performed using the \texttt{astropy} \citep{2013A&A...558A..33A}.
+These transformations were performed using \texttt{astropy} \citep{2013A&A...558A..33A}.
 
@@
 For each set of initial conditions, the particle positions were resampled to account for uncertainties in mass, which is an improvement from \citet{2023A&A...673A..44F}.
-Since globular clusters are older than 8 Gyr, their most massive stars have already gone supernova, leaving primarily low-mass stars, typically under one solar mass (CITATION?).
+Since globular clusters are older than 8 Gyr, their most massive stars have already gone supernova, leaving primarily low-mass stars, typically under one solar mass (\@\salvatore{Citation?
+Paolo knows this stuff well.}).
 Given that clusters usually have masses between $10^5-10^6$ solar masses, 100~000 particles is sufficient to represent the distribution of tidal debris.
@@
 
-    The integration was performed using a leapfrog integrator (add citation).
+    The integration was performed using a leapfrog integrator.
 The timestep size was chosen to be based on two critera.
@@
 
-    Overall, across 50 simulations, we observe 73 gaps.
-The distribution of simulations with a given number of gaps is as follows: 0 gaps in 4 simulations, 1 gap in 25 simulations, 2 gaps in 16 simulations, 3 gaps in 4 simulations, and 4 gaps in 1 simulation.
-Excluding NGC~2808, we are left with 29 gaps, leading to an expected range of 0.6-1.5 gaps per simulation.
+    Overall, across 50 simulations, we observe 73 gaps. The distribution of simulations with a given number of gaps is as follows: 
+    \begin{itemize}
+      \item 0 gaps in 4 simulations,
+      \item 1 gap in 25 simulations,
+      \item 2 gaps in 16 simulations,
+      \item 3 gaps in 4 simulations,
+      \item 4 gaps in 1 simulation. 
+    \end{itemize} Excluding NGC~2808, we are left with 29 gaps, leading to an expected range of 0.6-1.5 gaps per simulation.
 

### 2024-10-21 paper/main.tex
tabularized some data

@@
 Specifically, for each of the 50 realizations of the Palomar 5 stream, we saved $5000$ intermediate timesteps which is a temporal resolution of 1 million years.
-To provide an estimate of the size of our simulations, we have $N_p=100 000$ particles, $N_\textrm{time stamps} = 5 000$ time steps, $N_\textrm{phase}=6$ phase space coordinates, $N_\textrm{sampling}=50$ monte-carlo samplings, and using single precision floating point numbers, $\textrm{Single}=4$~bytes.
-The product of the aforementioned terms results in approximately: $N_p \times N_{\textrm{time stamps}}\times N_{\textrm{phase}}\times N_{\textrm{sampling}} \times 4 \textrm{bytes}\approx$ 600 GB of data.
+To provide an estimate of the size of our simulations, we have:
+    \begin{tabular}{c|c|c|c}
+      $N_p$ & $N_{\textrm{ts}}$ & $N_{\textrm{phase}}$ & $N_{\textrm{sampling}}$  \\
+      \hline
+      $100000$ & $5000$ & $6$ & $50$   \\
+    \end{tabular} 
+    where $N_p$  is the number of particles, $N_{\textrm{ts}}$ is the number of time-steps saved, $N_{\textrm{phase}}$ is the number of phase space coordinates, and $N_{\textrm{sampling}}$ is the number of monte-carlo samplings of the initial conditions.
+Using single precision floating point numbers, the size of our simulations is approximately:
+    \begin{equation*} 
+      N_p \times N_{\textrm{ts}}\times N_{\textrm{phase}}\times N_{\textrm{sampling}} \times 4~\textrm{bytes}\approx 600~\textrm{Gb}.
+    \end{equation*}
 
@@
 
-    Overall, across 50 simulations, we observe 73 gaps. The distribution of simulations with a given number of gaps is as follows: 
-    \begin{itemize}
-      \item 0 gaps in 4 simulations,
-      \item 1 gap in 25 simulations,
-      \item 2 gaps in 16 simulations,
-      \item 3 gaps in 4 simulations,
-      \item 4 gaps in 1 simulation. 
-    \end{itemize} Excluding NGC~2808, we are left with 29 gaps, leading to an expected range of 0.6-1.5 gaps per simulation.
+    Overall, across 50 simulations, we observe 73 gaps. The distribution of simulations with a given number of gaps is as follows: \begin{tabular}{|l|c|c|c|c|c|}
+      \hline
+      Number of Gaps & 0 & 1 & 2 & 3 & 4 \\
+      \hline
+      Number of Sims. & 4 & 25 & 16 & 4 & 1 \\
+      \hline
+    \end{tabular} Excluding NGC~2808, we are left with 29 gaps, leading to an expected range of 0.6-1.5 gaps per simulation.
 

### 2024-10-21 paper/main.tex
figure caption

@@
     The significance of an impact is measured by the change in momentum imparted to the stream stars by the perturber.
-A useful approximation for the change in momentum is the force at the closest approach multiplied by an estimate of the interaction time: \begin{equation} \Delta v \approx \text{Force} \times \text{interaction time} = \frac{GM}{b^2} \times 2\frac{b}{v} = 2\frac{GM}{bv}, \end{equation} where $M$ is the mass of the perturber, $b$ is the impact parameter, $v$ is the relative velocity of the perturber with respect to the particle, and $G$ is the gravitational constant.
+A useful approximation for the change in momentum is the force at the closest approach multiplied by an estimate of the interaction time: \begin{equation} \label{eq:change_in_momentum} \Delta v \approx \text{Force} \times \text{interaction time} = \frac{GM}{b^2} \times 2\frac{b}{v} = 2\frac{GM}{bv}, \end{equation} where $M$ is the mass of the perturber, $b$ is the impact parameter, $v$ is the relative velocity of the perturber with respect to the particle, and $G$ is the gravitational constant.
 We compute this for the five closest approaches of all globular clusters throughout each of the 50 simulations.
@@
     We desire to search for some generalizable characterisitcs of the perturbers to see what contrasts them to globular clusters that do not impact the Pal5 Stream.
-We present three different graphs in Fig.~\ref{fig:mass_size_plane}.
-The top plot shows the angular momentm and energy plane.
-The
+We present a suite of figures to answer for this end in Fig.~\ref{fig:mass_size_plane}.
+Note that the colors of the markers and histogram bins are
 
@@
       \includegraphics[width=0.45\linewidth]{impact_stats_phase_space.png}
-      \caption{Characterisitcs of clusters that cause gaps compared to those that do not.
-\textbf{Top:} shows the energy-angular momentum space of the globular clusters present in the simulations.
-Notice that we include all initial conditions that were sampled, therefore there are 50 $\times$ 165 data points.
-The clusters of a given simulation that impacted Palomar 5 are given with a Red marker, while non-gap causing clusters are given in gray.
-\textbf{Middle:} The mass-size plane.
-Uncertainties on the mass are indicated, while no uncertainties on the radius were reported.
-\textbf{Bottom:} the imparted momentum on the star-particles from a given flyby as given in Eq.~\ref{eq:change_in_momentum}.
-The top 5 strongest flybys from each simulation are included in this dataset, those that cause gaps are flagged.}
+      \caption{\textbf{Characterisitcs of gap causing clusters}--\textit{Note} that the colors of the markers and histogram bars are consistent with the clusters from \ref{fig:histogram_impact_time}.\\ \textbf{Top Left:} Energy-angular momentum space of the globular clusters in the simulations, with 50~$\times$~165 data points representing all sampled initial conditions.
+Clusters impacting Palomar 5 are shown with colored markers; non-gap-causing clusters are gray.
+Palomar 5 is marked with a white symbol and black edge.
+\\ \textbf{Top Right:} The mass-size plane.
+Uncertainties on the mass are indicated with vertical bars.
+\\ \textbf{Bottom Left:} The distribution of imparted momentum from a cluster flyby given by Eq.~\ref{eq:change_in_momentum}.
+The data set includes the top 5 strongest flybys from each simulation.
+Those that cause gaps are colored while those that do not are left in black.
+\\ \textbf{Bottom Right:} Palomar 5's orbit in the meridional plane.
+The grayscale represents all 50 stacked orbits, with $\mathcal{P}$ indicating the probability of  Palomar~5's position, normalized to $\mathcal{P}_\textrm{max}=1$.
+Markers show the perturber's position during each impact.}
       \label{fig:mass_size_plane}
@@
         \Delta p \approx \text{Force} \times \text{interaction time} = \frac{GMm}{b^2} \times 2\frac{b}{v} = 2\frac{GM}{bv}, 
-        \end{equation} \label{eq:change_in_momentum}
+        \end{equation} 
       where $M$ is the mass of the perturber, $m$ is the mass of the particle, $b$ is the impact parameter, $v$ is the relative velocity of the perturber with respect to the particle, and $G$ is the gravitational constant.

### 2024-10-21 paper/main.tex
once again made caption better

@@
 We present a suite of figures to answer for this end in Fig.~\ref{fig:mass_size_plane}.
-Note that the colors of the markers and histogram bins are
-
-    - to no surprise, you have to come within Palomar 5's orbital space
-
-    - trying to separate to the components of the velocity, Mass, impact parameter is not fruitful. It's better just to look at the change in momentum as given in the EQ.
-
-    - Notice that the distributions between the gap and no gap are significantly different, yet also have significant overlap.
-Perhaps these 3 parameters don't tell the whole story but also there's the factor identifies in Sanders, that I don't understand yet, that shows that a gap will not be identifiable.
 
@@
       \includegraphics[width=0.45\linewidth]{impact_stats_phase_space.png}
-      \caption{\textbf{Characterisitcs of gap causing clusters}--\textit{Note} that the colors of the markers and histogram bars are consistent with the clusters from \ref{fig:histogram_impact_time}.\\ \textbf{Top Left:} Energy-angular momentum space of the globular clusters in the simulations, with 50~$\times$~165 data points representing all sampled initial conditions.
+      \caption{\textbf{Characterisitcs of gap causing clusters}--\textit{Note} that the colors of the markers and histogram bars are consistent with Fig.~\ref{fig:histogram_impact_time}.\\ \textbf{Top Left:} Energy-angular momentum space of the globular clusters in the simulations, with 50~$\times$~165 data points representing all sampled initial conditions.
 Clusters impacting Palomar 5 are shown with colored markers; non-gap-causing clusters are gray.
 Palomar 5 is marked with a white symbol and black edge.
-\\ \textbf{Top Right:} The mass-size plane.
+The circular velocity-curve is drawn with a light-pink dashed curve.\\ \textbf{Top Right:} The mass-size plane.
 Uncertainties on the mass are indicated with vertical bars.
@@
       \end{figure*}
+
+      From the top left panel of Fig.~\ref{fig:mass_size_plane}, the immediate result is that all clusters that impact Palomar~5's stream are within a band of orbital energies from about $-2\times10^2~\textrm{km/s}^2$ to $-1\times10^2~\textrm{km/s}^2$.
+This is a non-surprising result indicating that in order for a cluster to perturb Palomar 5's stream, it must enter Palomar~5's orbital space.
+Additionally, we reproduce the results of \citet{2021MNRAS.505.5978V}, who found a wide range of possible orbital solutions for the least bound clusters, some being retrograde and some prograde.
+For this potential model, only 3 samplings of a globular cluster had positive energies.
+
+      - to no surprise, you have to come within Palomar 5's orbital space
+
+      - trying to separate to the components of the velocity, Mass, impact parameter is not fruitful. It's better just to look at the change in momentum as given in the EQ.
+
+      - Notice that the distributions between the gap and no gap are significantly different, yet also have significant overlap.
+Perhaps these 3 parameters don't tell the whole story but also there's the factor identifies in Sanders, that I don't understand yet, that shows that a gap will not be identifiable.
 

### 2024-10-21 paper/main.tex
I'm doing it

@@
 
-      From the top left panel of Fig.~\ref{fig:mass_size_plane}, the immediate result is that all clusters that impact Palomar~5's stream are within a band of orbital energies from about $-2\times10^2~\textrm{km/s}^2$ to $-1\times10^2~\textrm{km/s}^2$.
-This is a non-surprising result indicating that in order for a cluster to perturb Palomar 5's stream, it must enter Palomar~5's orbital space.
+      From the top left panel of Fig.~\ref{fig:mass_size_plane}, the immediate result is that all clusters that impact Palomar~5's stream are within a band of orbital energies from about $-2~\textrm{ to }~-1 \times10^5~\textrm{km}^2\textrm{s}^{-2}$.
+This is non-surprising result indicates that in order for a cluster to perturb Palomar 5's stream, it must enter Palomar~5's orbital space.
 Additionally, we reproduce the results of \citet{2021MNRAS.505.5978V}, who found a wide range of possible orbital solutions for the least bound clusters, some being retrograde and some prograde.
@@
 
-      - to no surprise, you have to come within Palomar 5's orbital space
-
-      - trying to separate to the components of the velocity, Mass, impact parameter is not fruitful. It's better just to look at the change in momentum as given in the EQ.
-
-      - Notice that the distributions between the gap and no gap are significantly different, yet also have significant overlap.
-Perhaps these 3 parameters don't tell the whole story but also there's the factor identifies in Sanders, that I don't understand yet, that shows that a gap will not be identifiable.
+      From the top left panel of Fig.~\ref{fig:mass_size_plane} shows the mass-size plane of the perturbers against those that do not provoke gaps.
+We find a lower limit of under $10^5~\textrm{M}_{\odot}$ for provoking gaps.
+Mass and size are not the full picture, since provoking a gap is given by how much momentum is injected into the stream by Eq.~\ref{eq:change_in_momentum}.
+However, this emperical lower limit is a practical result since the relative velocity and impact parameter are known values since they happen so long ago in the past.
+The bottom left panel shows a result which is more true to the nature of the physics of the problem.
+By tracking the closest approaches and inspecting their approximate imparted momentum, we find that the gap-creating impacts have systematically higher imparted momenta that those that do not create gaps.
+However, there is significant overlap.
+Of course, the three parameter law from ~\ref{eq:change_in_momentum} is a simplified image.
+\citet{2015MNRAS.450.1136E} extended this to include the characterisitc size of the perturber and separated the relative velocity into perpendicular and parallel components to the stream an subsequently quantified how the parameters effect gap evolution.
+However, when studying the impact geometry at this level we found no satisfying combination of the parameters that clearly discriminated between the gap and no gap case.
+An extension this is shown in Fig.~\ref{fig:impact_geometry_statistics}, the only additional insights being that there is an observed maximum impact parameter which is about ~100~parsecs.
+This is roughly on the order of magnitude of the stream width.
+Additionally, as \citet{2016MNRAS.457.3817S} demonstrated, the dispersion of the action-frequencies of the star-particles being impacted plays an important factor.
+Stream regions that are well separated by orbital frequency accept gaps easier while the regions of the stream that have a wide range of frequencies (those closests to the cluster) erase the history of the impact easier.
+In essence, an additional criteria is that the imparted change in frequency must be greater than the range in frequencies present in the stream area being hit.
 

### 2024-10-21 paper/main.tex
computed linear dependence

@@
 In essence, an additional criteria is that the imparted change in frequency must be greater than the range in frequencies present in the stream area being hit.
+
+      From the bottom right panel we observe where the impacts happen in the meridional plane lain over Palomar 5's orbit.
+We see that there is a trend with $r$, the 3D galactocentric radius.
+The number of counts follows this equation: $N = -2.5 r + 50$ and has a pearson coefficient of -0.86, so quite strong I must say.
 

### 2024-10-22 paper/main.tex
updating appendix to be consistent

@@
        M. Montuori\inst{1}
+       \and
+       A. Mastrobuono-Battisti
        }
@@
 Using single precision floating point numbers, the size of our simulations is approximately:
-    \begin{equation*} 
+    \begin{equation} \label{eq:data_volume_estimate}
       N_p \times N_{\textrm{ts}}\times N_{\textrm{phase}}\times N_{\textrm{sampling}} \times 4~\textrm{bytes}\approx 600~\textrm{Gb}.
-    \end{equation*}
+    \end{equation}
 
@@
 This system allows for consistent morphological comparisons by aligning the stream with the cluster's orbit.
-The coordinate system is based on the work of \citet{2004AJ....127.2753D} and is detailed in Appendix~\ref{appendix:TailCoordinates}.
+The coordinate system is based on the work of \citet{2004AJ....127.2753D} and is detailed in Fig.~\ref{appendix:TailCoordinates}.
 Briefly, the x' coordinate represents the position of a particle along the orbit relative to the globular cluster—whether it is ahead or behind the cluster.
@@
 
-  \section{Tail Coordinates} \label{appendix:TailCoordinates}
-
+  \section{Gap Detection} \label{sec:gap_detection}
+    For each of the 50 simulations, we compare the final snapshots between the \texttt{reference} and \texttt{full} simulations using the tail coordinate system, as shown in Fig.~\ref{fig:TailCoordinates}.
+The vanilla simulations were only saved at the final timestamps, thus from Eq.\ref{eq:data_volume_estimate} where $N_ts$=1, leading to a data volumne of about one hundred messily megabytes.
+We inspect these differences by generating 2D density maps of the streams.
+Additionally, we marginalize over the $y'$-coordinate to produce 1D density profiles along the $x'$-axis.
+These comparisons are shown in the results section (see Fig.~\ref{fig:profiles}).
     \begin{figure}
@@
 
-  \section{Gap Detection} \label{sec:gap_detection}
-    For each of the 50 simulations, we compare the final snapshots between the \texttt{vanilla} and \texttt{full} simulations using the tail coordinate system.
-We inspect these differences by generating 2D density maps of the streams.
-Additionally, we marginalize over the $y'$-coordinate to produce 1D density profiles along the $x'$-axis.
-These comparisons are shown in the results section (see Fig.~\ref{fig:profiles}).
-
     \begin{figure*}
@@
       \includegraphics[width=\linewidth, trim=20 0 15 0]{monte-carlo-009-pouliasis2017pii-GCNBody-2000-milisigma-5-noisefactor-20-boxcarindexlength-shifted-0.png}
-      \caption{A comparison between the density maps and profiles of the \texttt{full} and \texttt{vanilla} simulations is presented.
+      \caption{A comparison between the density maps and profiles of the \texttt{full} and \texttt{reference} simulations is presented.
 The locations where the full simulation is less dense than the vanilla simulation by 2-$\sigma$ is highlighted by blue vertical bars.
@@
 
-    To construct the 1D density profiles, we first bin the data using the $\sqrt{N}$ rule, where $N$ is the number of data points ($N_p = 10^5$).
+    To construct the 1D density profiles, we bin the data using the $\sqrt{N}$ rule, where $N$ is the number of data points ($N_p = 10^5$).
 After binning the 1D profiles, we apply a boxcar smoothing technique.
 The density profiles are scanned.
+This raw profile density curve is then smoothed with the median box-car filter.
 At each bin, a selected number of adjacent data points from both sides are placed in a list.
@@
 
-    With the smoothed 1D density profiles in hand, we treat the \texttt{vanilla} simulations as the reference model and search for regions in the \texttt{full} simulations that are significantly underdense compared to \texttt{vanilla}, surpassing stochastic fluctuations.
+    With the smoothed 1D density profiles in hand, we search for regions where the \texttt{full} simulations are significantly underdense compared to \texttt{reference} simulations, surpassing stochastic fluctuations.
 To accomplish this, we first impose a signal-to-noise ratio threshold, $\mathcal{SNR}$.
@@
     This method, however, has its limitations, especially when detecting smaller gaps.
-As outlined by Erkal et al.
-(2015), a small gap is indicative of a weak impact, but rather a more recent one.
+As outlined by \citet{2015MNRAS.450.1136E}, a small gap is not indicative of a weak impact, but a recent one.
 This is because gap growth is a dispersion phenomenon.
@@
 This method helps particularly with large subtle gaps that the eye does not notice in the 2D maps.
+These profiles are included in the Appenix~\ref{sec:gallery_of_gaps}.
 
@@
 To avoid extreme low-density regions at the stream's edges, we applied the same density threshold as from eq.~\ref{eq:density_threshold} to focus on the more significant areas of the stream.
-Next, we modeled Palomar 5's orbit as a proxy for its stream and sampled points along the orbit to measure the gravitational force exerted by other globular clusters.
+Next, we modeled Palomar~5's orbit as a proxy for its stream and sampled points along the orbit to measure the gravitational force exerted by other globular clusters.
 This force data is displayed in the top panel of Fig.~\ref{fig:force-on-orbit}, showing how the total gravitational acceleration on Palomar 5's stream evolves over its length throughout the simulation.
@@
 The bottom plot showcases the evolution of the stream density in simulation time.
-The density was used to determine a sutible length of the stream as a function of time.
-This length was then used to extract a piece of Palomar 5's orbit about a given position of Pal 5 in time.
+The density was used to determine a sutible length of the stream at a given function.
+This length was then used to extract a piece of Palomar 5's orbit.
 This orbital segment is used to approximate the stream.
@@
 
-      We then used \texttt{scipy}'s \texttt{ndimage} package to identify the top five local maxima in the data space of gravitational acceleration $\vec{g}$ as a function of time $t$ and the stream coordinate $\tau$.
-TheThis is done by first smoothing the image by taking a 5 point moving average kernel.
+      We then used \texttt{scipy}'s \texttt{ndimage} \citep{2020NatMe..17..261V} package to identify the top five local maxima in the data space of gravitational acceleration $\vec{g}$ as a function of time $t$ and the stream coordinate $\tau$.
+This is done by first smoothing the image by taking a 5 point moving average kernel.
 Secondly, we use maximum filter to locate coordinates in the $t,\tau$ data plane who are local maxima to at least 10-adjacent data points.
@@
 Notice the periodic ribbions of force in the $(t,\tau)$ plane.
-This is due to pericenter passages where Palomar 5 is getting closer to the center of mass of the globular cluster system.
-Additionally, for the impacts of NGC104 and NGC7078, wakes can be observed in the density map.
+This is due to pericenter passages where Palomar~5 is getting closer to the center of mass of the globular cluster system.
+Additionally, for the impacts of NGC~104 and NG~C7078, wakes can be observed in the density map.
 Another important aspect, is that the strongest peak in gravitational force does not necessarily create a gap.
-Notice how NGC5272, which was labeled with 0 to indicate that it has the greatest local maxima does not have a gap.
-This is manifold, for instance, the force needs to be modulated with the time of impact because the  important quantity that determins gap creation.
+Notice how NGC~5272, which was labeled with 0 to indicate that it has the greatest local maxima does not have a gap.
+The reason for this is manifold, for instance, the force needs to be modulated with the time since the change in momentum is the determining factor and not the peak magnitude of the force.
 Additionally, there is an offset of about 200~pc between the stream at the orbit, as seen when viewing the stream in tail coordinates, so peaks upon the orbit are good proxy's for the stream but are not definitive.
@@
       With the perturbers identified, we perform statistics in the pursuit of understanding what conditions are necessary for a globular cluster and a stream to have in order for a gap to form.
-We turn to impact theory, which in its simplest form as presented in works such as \citet{2008gady.book.....B}, John Taylor, and Galaxiesbook, considers two particles: one stationary and the other moving past it.
+We turn to impact theory, which in its simplest form is presented in works such as \citet{2008gady.book.....B}.Consider two particles: one stationary and the other moving past it.
 The distance between the two particles at their closest approach is known as the impact parameter.
-To simplify the analysis, the \textit{impulse approximation} is often employed, which assumes that the velocity of the perturber remains unchanged during the interaction.
-This assumption greatly simplifies the computation.
+To simplify the analysis, the \textit{impulse approximation} is employed, which assumes that the velocity of the perturber remains unchanged during the interaction.
+This assumption simplifies the computation.
 
@@
 
-      Erkal et al.
-(2015) extended this impact theory from one point mass impacting another point mass, to studying how an extended body impacts as stream.
+      \citet{2015MNRAS.450.1136E} extended this impact theory from one point mass impacting another point mass, to studying how an extended body impacts as stream.
 This quantifies the change in momentum of a given particle as a function of its distance from the point of greatest impact.
-In their analysis, the perturber was modeled as a Plummer sphere.
-Since the stream is not zero-dimensional and has length, both the parallel and perpendicular components of the velocity must be considered.
-Consequently, five parameters determine the change in velocity of a given particle: $M$, $r_p$, $b$, $W_\parallel$, and $W_\perp$.
-Our goal is to constrain these parameters during the impact in our simulations and compare them to close approaches that do not generate gaps, in order to establish the criteria necessary for gap formation.
+In their analysis, the perturber was modeled as a Plummer sphere, like in our simulations.
+Since the stream is not zero-dimensional and has length, both the parallel and perpendicular components of the pertuber's velocity must be considered.
+Consequently, five parameters determine the change in velocity of a given stream particle: $M$, $r_p$, $b$, $W_\parallel$, and $W_\perp$.
+We set out is to constrain these parameters during the impact in our simulations and compare them to close approaches that do not generate gaps, in order to establish the criteria necessary for gap formation.
 
       To achieve this, we begin by identifying the approximate moments of impact from the most significant clusters, as determined in the previous analysis in Sec.~\ref{sec:Perturber_Identification}.
-Then, we refine these estimates.
-While the orbit serves as a good first-order proxy for the stream, it is insufficient, as the typical offset between the stream and its orbit can be on the order of 200-400 pc—--significantly larger than the impact parameters we are investigating.
-
-      In this refinement process, we aim to pinpoint both the exact location of the impact along the stream and the precise moment it occurred.
+Then, we refine these estimates in oredr to to pinpoint both the exact location of the impact along the stream and the precise moment it occurred.
 To do so, we fit a third-order parametric polynomial to the stream, using the saved snapshots from our simulations:
@@
         \end{equation}  
-      where $x$, $y$, and $z$ represent the parametric line describing the stream in galactocentric coordinates, $\tau$ is the stream coordinate in time as described in the previous section, and is used as the independent variable to parameterize the motion.
+      where $x$, $y$, and $z$ represent the parametric line describing the stream in galactocentric coordinates, $\tau$ is the stream coordinate in time as described in the Appendix~\ref{sec:Perturber_Identification}, and is used as the independent variable to parameterize the position along the stream.
 The coefficients $a_i$, $b_i$, and $c_i$ are the polynomial coefficients.
-We found that a second-order polynomial was insufficient to capture the curvature along the full length of the stream, with divergence from the star particle positions occurring at the extremes.
+We found that a second-order polynomial was insufficient to capture the curvature along the full length of the stream, with divergence at the ends of the tails.
 A third-order polynomial was sufficient and desirable, as it is the lowest order that adequately captures the path of the stream over the entire length under consideration.
@@
 The minimum value of $b$, denoted as $\text{min}(b)$, represents the impact parameter.
-The minimization is carried out using \texttt{scipy}'s optimization package with the \textit{L-BFGS-B} method, which allows us to constrain $t$ and $\tau$, ensuring no extrapolation occurs.
+The minimization is carried out using \texttt{scipy}'s optimization package with the \textit{L-BFGS-B} method, which allows us to place bounds on $t$ and $\tau$, ensuring no extrapolation occurs.
 
@@
 
+      From visual inspection of the five-impact defining quantifies ($M$, $r_p$, $b$, $W_\parallel$, and $W_\perp$) did not provide an obvious discrimination between flybys that provoke gaps from those that do not.
+From the clutter, we do not present this graph and instead present Fig.~\ref{fig:impact_geometry_statistics} in which we use the total relative velocity instead of separating between the parallel and perpendicular components.
+Also, we ommit the characterisitc radius which did not seem to correlate strongly with results.
+This is likely due to the small range of characterisitic radii of the globular clusters, as shown in the top right panel of Fig.~\ref{fig:mass_size_plane}.
+Perhaps this factor could be more influential for a dark matter subhalo population where the sizes can vary grately.
       \begin{figure*}
@@
         \end{figure*}
+      While Fig.~\ref{fig:impact_geometry_statistics} shows us that in isolation Mass, relative velocity, or impact parameter from Eq.~\ref{eq:change_in_momentum} is insufficient to predict the presence of a gap or not, one interesting result does appear.
+There are not gaps from interactions with impact parameters at a distance greater than 200~pc.
+We notice that the widths of the stream are on the order of 200~pc, or have a radial thickness of 100~pc.
+This result becomes more obvious when looking at the plane of $b$-$M$, where the near vertical line at a mass of about 7$\times 10^5 M_\odot$ indicates the simulations with NGC~2808.
+The simulations with impact parameters greater than 200~pc do not provide gaps where ever encounter under this indeed does.
 

### 2024-10-22 paper/main.tex
draft mode to full compile
proper implementation of the tail coordiantes figure

@@
 This system allows for consistent morphological comparisons by aligning the stream with the cluster's orbit.
-The coordinate system is based on the work of \citet{2004AJ....127.2753D} and is detailed in Fig.~\ref{appendix:TailCoordinates}.
+The coordinate system is based on the work of \citet{2004AJ....127.2753D} and is shown in Fig.~\ref{fig:TailCoordinates}.
 Briefly, the x' coordinate represents the position of a particle along the orbit relative to the globular cluster—whether it is ahead or behind the cluster.
@@
       \centering
-      \includegraphics[width=\linewidth]{figures/along_orbit_coordinate_system-approved.png}
+      \includegraphics[width=\linewidth]{along_orbit_coordinate_system-approved.png}
       \caption{Tail coordinates}

### 2024-10-22 paper/main.tex
editted perturber stats appendix

@@
       \includegraphics[width=\linewidth]{along_orbit_coordinate_system-approved.png}
-      \caption{Tail coordinates}
+      \caption{Tail coordinates. \textbf{I want to update this}. }
       \label{fig:TailCoordinates}
@@
 
-      From visual inspection of the five-impact defining quantifies ($M$, $r_p$, $b$, $W_\parallel$, and $W_\perp$) did not provide an obvious discrimination between flybys that provoke gaps from those that do not.
-From the clutter, we do not present this graph and instead present Fig.~\ref{fig:impact_geometry_statistics} in which we use the total relative velocity instead of separating between the parallel and perpendicular components.
-Also, we ommit the characterisitc radius which did not seem to correlate strongly with results.
-This is likely due to the small range of characterisitic radii of the globular clusters, as shown in the top right panel of Fig.~\ref{fig:mass_size_plane}.
-Perhaps this factor could be more influential for a dark matter subhalo population where the sizes can vary grately.
+      Visual inspection of the five key impact parameters ($M$, $r_p$, $b$, $W_\parallel$, and $W_\perp$) did not reveal a clear distinction between flybys that create gaps and those that do not.
+Due to the clutter, we omit this graph and instead present Fig.\ref{fig:impact_geometry_statistics}, which uses the total relative velocity rather than separating parallel and perpendicular components.
+We also exclude the characteristic radius, as it showed little correlation with the results, likely due to the narrow range of globular cluster radii (see Fig.\ref{fig:mass_size_plane}).
+This factor might be more significant for dark matter subhalos, where size variation is greater.
+
       \begin{figure*}
@@
         \includegraphics[width=\linewidth]{impact_geometry_statistics.png}
-        \caption{Impact geometry statistics}
+        \caption{The distribution and relationship between the basic impact variables from Eq.\ref{eq:change_in_momentum} and the close encounter that cause impacts, shown in red against those that do not, shown in blue.
+}
         \label{fig:impact_geometry_statistics}    
         \end{figure*}
-      While Fig.~\ref{fig:impact_geometry_statistics} shows us that in isolation Mass, relative velocity, or impact parameter from Eq.~\ref{eq:change_in_momentum} is insufficient to predict the presence of a gap or not, one interesting result does appear.
-There are not gaps from interactions with impact parameters at a distance greater than 200~pc.
-We notice that the widths of the stream are on the order of 200~pc, or have a radial thickness of 100~pc.
-This result becomes more obvious when looking at the plane of $b$-$M$, where the near vertical line at a mass of about 7$\times 10^5 M_\odot$ indicates the simulations with NGC~2808.
-The simulations with impact parameters greater than 200~pc do not provide gaps where ever encounter under this indeed does.
+
+      While Fig.\ref{fig:impact_geometry_statistics} demonstrates that mass, relative velocity, or impact parameter alone (from Eq.\ref{eqeq:change_in_momentum}) cannot predict gap formation, one interesting result emerges: no gaps are created from interactions with impact parameters greater than 200~pc.
+The stream widths are roughly 200~pc, with a radial thickness of about 100~pc.
+This finding is even more evident when examining the $b$-$M$ plane.
+A series of perturbers at roughly $\sim7 \times 10^5 M_\odot$ highlights NGC~2808's flybys, where all encounters with impact parameters under 200~pc result in gaps, while those beyond this distance do not.
 

### 2024-10-22 paper/main.tex
cleaned the discussion on the perturber stats

@@
 
-    We desire to search for some generalizable characterisitcs of the perturbers to see what contrasts them to globular clusters that do not impact the Pal5 Stream.
-We present a suite of figures to answer for this end in Fig.~\ref{fig:mass_size_plane}.
+    We aim to identify general characteristics of perturbers that distinguish them from globular clusters that do not affect the Palomar 5 stream.
+Fig.~\ref{fig:mass_size_plane} presents several analyses to explore this.
 
@@
 
-      From the top left panel of Fig.~\ref{fig:mass_size_plane}, the immediate result is that all clusters that impact Palomar~5's stream are within a band of orbital energies from about $-2~\textrm{ to }~-1 \times10^5~\textrm{km}^2\textrm{s}^{-2}$.
-This is non-surprising result indicates that in order for a cluster to perturb Palomar 5's stream, it must enter Palomar~5's orbital space.
-Additionally, we reproduce the results of \citet{2021MNRAS.505.5978V}, who found a wide range of possible orbital solutions for the least bound clusters, some being retrograde and some prograde.
-For this potential model, only 3 samplings of a globular cluster had positive energies.
+      The top-left panel shows that all clusters impacting the stream fall within a specific range of orbital energies, between $-2$ and $-1 \times10^5~\textrm{km}^2\textrm{s}^{-2}$.
+This is expected, as only clusters within Palomar 5's orbital space can interact with the stream.
+Our results are consistent with the findings from \citet{2021MNRAS.505.5978V}, who identified a wide range of orbital solutions for the least bound clusters, including both retrograde and prograde orbits.
+In our potential model, only three clusters samplings had positive orbital energies.
 
-      From the top left panel of Fig.~\ref{fig:mass_size_plane} shows the mass-size plane of the perturbers against those that do not provoke gaps.
-We find a lower limit of under $10^5~\textrm{M}_{\odot}$ for provoking gaps.
-Mass and size are not the full picture, since provoking a gap is given by how much momentum is injected into the stream by Eq.~\ref{eq:change_in_momentum}.
-However, this emperical lower limit is a practical result since the relative velocity and impact parameter are known values since they happen so long ago in the past.
-The bottom left panel shows a result which is more true to the nature of the physics of the problem.
-By tracking the closest approaches and inspecting their approximate imparted momentum, we find that the gap-creating impacts have systematically higher imparted momenta that those that do not create gaps.
-However, there is significant overlap.
-Of course, the three parameter law from ~\ref{eq:change_in_momentum} is a simplified image.
-\citet{2015MNRAS.450.1136E} extended this to include the characterisitc size of the perturber and separated the relative velocity into perpendicular and parallel components to the stream an subsequently quantified how the parameters effect gap evolution.
-However, when studying the impact geometry at this level we found no satisfying combination of the parameters that clearly discriminated between the gap and no gap case.
-An extension this is shown in Fig.~\ref{fig:impact_geometry_statistics}, the only additional insights being that there is an observed maximum impact parameter which is about ~100~parsecs.
-This is roughly on the order of magnitude of the stream width.
-Additionally, as \citet{2016MNRAS.457.3817S} demonstrated, the dispersion of the action-frequencies of the star-particles being impacted plays an important factor.
-Stream regions that are well separated by orbital frequency accept gaps easier while the regions of the stream that have a wide range of frequencies (those closests to the cluster) erase the history of the impact easier.
-In essence, an additional criteria is that the imparted change in frequency must be greater than the range in frequencies present in the stream area being hit.
+      The mass-size relation, shown in the top-right panel, reveals a lower limit of $10^5~\textrm{M}_{\odot}$ for clusters that provoke gaps.
+While mass and size alone are insufficient predictors, this empirical limit is useful because relative velocity and impact parameters are historical values.
+The bottom-left panel offers a more physically meaningful result: gap-creating impacts tend to impart more momentum into the stream than non-gap impacts, though with some overlap.
+While \citet{2015MNRAS.450.1136E} extended the three-parameter model (Eq.~\ref{eq:change_in_momentum}) to include the size of the perturber and the perpendicular and parallel components of relative velocity, we found no clear combination of these factors that distinguishes gap from non-gap cases.
 
-      From the bottom right panel we observe where the impacts happen in the meridional plane lain over Palomar 5's orbit.
-We see that there is a trend with $r$, the 3D galactocentric radius.
-The number of counts follows this equation: $N = -2.5 r + 50$ and has a pearson coefficient of -0.86, so quite strong I must say.
+      Further insights are shown in Fig.\ref{fig:impact_geometry_statistics}, highlighting a maximum impact parameter of around 200~pc, roughly the width of the stream.
+Additionally, as \citet{2016MNRAS.457.3817S} demonstrated, the dispersion of action-frequencies in the stream plays a role.
+Regions of the stream with well-separated frequencies are more susceptible to gap formation, while those with a wide frequency range (near the cluster) tend to erase the history of impacts.
+Thus, for a gap to form, the imparted change in frequency must exceed the range of frequencies in the impacted region.
+Thus, the strong flybys that occured close to the cluster and were thus inconsequential for gap formation.
+
+      The bottom-right panel reveals a strong negative correlation between the location of impacts and the galactocentric radius, $r$.
+The number of gap creating impacts follows the relation $N = -2.5r + 50$, with a Pearson coefficient of -0.86.
+This is a non-obvious result.
+While there are more clusters near the galactic center, orbits naturally spend more time near their apocenters, and the lower relative velocities in these regions should favor gap creation.
+However, this result suggests that cluster galactic number density outweighs these factors in determining the number of gaps.
 

### 2024-10-22 paper/main.tex
started paragraph for carlberg gap creation rate quantity

@@
 
-    \textbf{\textit{Action item:}} I need to make sure I check out real good the papers that investigate the gaps from Palomar 5.
-I wanna compare this to the known length of Palomar 5, how does this reduce the results, and are there observed gaps in Palomar 5 that are consistent with these simulations?
+      Lastly, \citet{2012ApJ...748...20C} conceptualized the number of gaps with a statistical quantity, the rate at which gaps are created per unit stream-length.
+To put our results in these terms, the global cluster population has a gap creation rate on Palomar 5 of XXX.
 

### 2024-10-22 paper/main.tex
starting gap creation rate calculation

@@
       Lastly, \citet{2012ApJ...748...20C} conceptualized the number of gaps with a statistical quantity, the rate at which gaps are created per unit stream-length.
-To put our results in these terms, the global cluster population has a gap creation rate on Palomar 5 of XXX.
+For our simulations, this is given by:\begin{equation}
+      \mathcal{R}_{\textrm{Pal 5}} =  \frac{1}{T}\int_{0}^T l^{-1}(t) \sum_i \delta(t-t_i) dt,\end{equation}where $T$ is the total integration time, $l(t)$ is the length of the stream, and $\sum_i \delta(t-t_i)$ sums over the number of gap occurances, where the index $i$ iterates over the number of gaps in a given simulation since $\delta$ is the diract delta function.
+This can be simplified to:\begin{equation}\mathcal{R}_{\textrm{Pal 5}} =  \frac{1}{T} \sum_i \frac{1}{l (t_i)}.
+\end{equation}
 

### 2024-10-23 paper/main.tex
draft mode
included gap creation rate and itemized the discussion

--- 2024-10-22 paper/main.tex
@@
 
-      Lastly, \citet{2012ApJ...748...20C} conceptualized the number of gaps with a statistical quantity, the rate at which gaps are created per unit stream-length.
-To put our results in these terms, the global cluster population has a gap creation rate on Palomar 5 of XXX.
+      Lastly, \citet{2012ApJ...748...20C} conceptualized the number of gaps using a statistical quantity: the rate at which gaps are created per unit stream length.
+For our simulations, this is given by: \begin{equation} \label{eq:gap_creation_rate}
+      \mathcal{R}_{\textrm{Pal 5}} =  \frac{1}{T}\int_{0}^T l^{-1}(t) \sum_i \delta(t-t_i) dt,\end{equation}where $T$ is the total integration time, $l(t)$ is the length of the stream, and $\sum_i \delta(t-t_i)$ sums over the gap occurrences, with $i$ indexing over the number of gaps in a given simulation.
+Here, $\delta$ represents the diract delta function.
+This expression can be simplified to:\begin{equation}\mathcal{R}_{\textrm{Pal 5}} =  \frac{1}{T} \sum_i \frac{1}{l (t_i)}.
+\end{equation}
+      This computation allows us to analyze the distribution of gap creation rates across all simulations.
+The results are shown in Fig.~\ref{fig:TailCoordinates}.
+This is roughly consistent with a simple estimate of the average gap creation rate: with 72 gaps over 5 Gyr of integration time for a stream about 20 kpc in length, the naive estimate is approximately 0.014~[km~s$^{-1}$~kpc$^{-2}$].
+This is about double the weighted mean gap creation rate of 0.009~[km~s$^{-1}$~kpc$^{-2}$].
+The naive estimate is higher because it does not account for the growth of the stream over time, unlike Eq.\ref{eq:gap_creation_rate}, where the likelihood of impacting a shorter stream is lower.
+
+      \begin{figure}
+        \centering
+        \includegraphics[width=\linewidth]{gap_creation_rate.png}
+        \caption{The distribution of the number of gaps normalized over the total integration time and unit stream length, as described by Eq.\ref{eq:gap_creation_rate}. }
+        \label{fig:TailCoordinates}
+      \end{figure}
 
 \section{Discussion}
-    How many of these could be observable?
-What's the main result?
-in 5 billion years of evolution, perhaps you'd only expect to 1.4 gaps, reaching up to maximum three?
-However, its possible that we dont see gaps at all for sure.
 
-    We do this for we attempt to quantify the number of gaps from globular clusters in the limit of maximum information.
-I.e.
-we are not limited by contamination of field stars, and since we are computing our own simulations, we know what a gapless stream should look like.
-
-    What about drawing attention to future studies?
-
-    Doing a complete census of all globular clusters?
-
-    Doing a complete census of dark matter subhaloes?
-
-    How many gaps do you expect to see in Palomar 5? Bonaca said there are about 2 of them.
+    \begin{itemize}
+      \item Would these gaps be observable? We are in the max info limit, what about observational difficulties?
+      \item the detected track of Palomar 5 is much shorter than my simulations. Does this mean cluster gaps sould excedingly rare?
+      \item What about Carlberg's Pal 5 gap paper? 
+      \item What about Bonaca's Pal 5 gap paper?
+      \item future works?
+      \item bars substructure?
+      \item What else?
+      \begin{itemize}
+        \item Doing a complete census of all globular clusters?
+        \item doing a compelte census of all observed streams? 
+        \item comparing to dark matter subhaloes?
+        \end{itemize}
+      \end{itemize}
 

### 2024-10-28 paper/main.tex
more discussion points
discussion and trying to handle Pal 5's mass

@@
 
-    At their computed positions 5 Gyr ago, each globular cluster was assigned 100,000 particles sampled from a Plummer distribution to represent its stellar content.
-For each set of initial conditions, the particle positions were resampled to account for uncertainties in mass, which is an improvement from \citet{2023A&A...673A..44F}.
-Since globular clusters are older than 8 Gyr, their most massive stars have already gone supernova, leaving primarily low-mass stars, typically under one solar mass (\@\salvatore{Citation?
-Paolo knows this stuff well.}).
-Given that clusters usually have masses between $10^5-10^6$ solar masses, 100~000 particles is sufficient to represent the distribution of tidal debris.
+    At their computed positions 5 Gyr ago, Palomar 5 was assigned 100,000 particles sampled from a Plummer distribution to represent its stellar content.
+For each set of initial conditions, the particle were resampled based on uncertainties in mass, which is an improvement from \citet{2023A&A...673A..44F}.
+
+    We use a mass estimate for Palomar 5 taken from \citet{2021MNRAS.505.5957B}, $1.34\pm0.24\times10^{4}M_{\odot}$--which is a contested value.
+
+    \salvatore{I'm a bit lost on trying to reconcile the differences between Baumgardt's estiamte and \citet{2017ApJ...842..120I}'s estimate in sections 5.2 and 5.3.
+It looks like 5.2 is based on counts and then supplementing the completness function with dNdM \citet{2001AJ....122.3231G} and demonstrating that Salpeter would overestimate the mass.
+I think 5.3 is showing that an MCMC was used to add more context to the estimate from 5.3.}
+
+    \salvatore{\textbf{If Salpeter over estimates the number of low mass stars}, then wouldn't this appear in the velocity dispersion profile?
+or wait, is \citet{2017ApJ...842..120I} fitting the density for the King, supplementing missing stars by a Mass function from \citet{2001AJ....122.3231G}, and finding the total mass this way?
+i.e.
+no velocity dispersion was investigated?
+I feel like viewing the velocity dispersion directly would provide a good estiamte for the dynamical mass, no?}
+
+    \salvatore{I am lost in Baumgardt's papers trying to flush out how a MF was used to provide their mass estimate. It looks like this information is within Sollima 2017. Need to revisit this. }
+
+    \salvatore{What I say will depend.
+I think using \citet{2021MNRAS.505.5957B}'s mass estimate is still ok because it is similar in order of magnitude to cluster + stream from \citet{2017ApJ...842..120I}.
+}
 
@@
 A preliminary evaluation is performed visually, followed by a quantitative analysis of the differences between the density profiles in the two mentioned cases.
-
-    % We initially sought an automated, quantitative approach from signal and image processing that could reliably detect all gaps and evaluate their significance.
-Unfortunately, despite extensive efforts, an ideal method remained elusive.
-As a result, we opted for a simpler approach: a combination of visual inspection aided by comparative analysis.
 
@@
 
+  In this work, we quantified the number of gaps expected in Palomar 5's stream from the 165 Galactic globular clusters using a Monte Carlo-style experiment.
+We performed 50 iterations, varying the orbits based on the uncertainties in their observed proper motions, radial velocities, and distances.
+Our findings are as follows:
+
+  \begin{itemize}
+    \item \textbf{Orbital Overlap}: clusters with orbital energies and angular momenta that overlap with Palomar 5's orbit can impact its stream.
+    \item \textbf{Mass Threshold}: Perturbing clusters have masses above $10^5 M_\odot$; no perturber has a mass smaller than this.
+    \item \textbf{Galactic radius}: Impacts preferentially occur at shorter galactic radii.
+    \item \textbf{Momentum and Frequency Change}: Gap creation is linked to the momentum imparted to the stream during the flyby, yet not entirely.
+As showcased by \citet{2016MNRAS.457.3817S}, the change in orbital frequencies induced by the flyby must be greater than the spread in frequencies present in the stream.
+Some impacts imparted significant momentum changes that did not lead to gap growth for this reason.
+  \end{itemize}
+
+  Finally, we present a gap creation rate as conceptualized by \citet{2012ApJ...748...20C}, which is the number of gaps normalized to the age and length of the stream.
+This rate is: 0.009~[km~s$^{-1}$~kpc$^{-2}$], or 0.01~[Gyr$^{-1}$~kpc$^{-1}$].
+
+  Now, \textit{is this consistent with reality?} Can globular clusters explain the observed gaps in Palomar 5?
+Currently, the results on the number of gaps in the stream are debated.
+\citet{2012ApJ...760...75C} and \citet{2020ApJ...889...70B} attempted to quantify the stream's density along its width, isolating the stream from background and foreground contamination.
+\citet{2020ApJ...889...70B} reported the existence of two potential gaps, whereas \citet{2012ApJ...760...75C} identified five gaps with a 99~\% detection confidence, leading to a gap creation rate of 0.17~Gyr$^{-1}$~kpc$^{-1}$.
+If true, this rate is too high to be explained by Galactic globular clusters.
+
+  On the other hand, for a stream length of 10~kpc reported from \citet{2020ApJ...889...70B} and given our gap creation rate, in 5~Gyr we would expect to see about 0.5 gaps, in 10~Gyr's we would expect 1.
+Therefore, finding two gaps from globular clusters would not be impossible.
+
     \begin{itemize}
@@
       \item the detected track of Palomar 5 is much shorter than my simulations. Does this mean cluster gaps sould excedingly rare?
-      \item What about Carlberg's Pal 5 gap paper? 
-      \item What about Bonaca's Pal 5 gap paper?
       \item future works?
       \item bars substructure?
-      \item What else?
+      \item link to tstrippy? open source
+      \item add the new simulations to etidal public database?
+      \item the new proposed instrument and survey by Conroy and Bonaca to get improve stream data? https://via-project.org/. How to cite or include in the discussion?
+      \item \textbf{What else?}
+      \item Future potential works
       \begin{itemize}
+        \item mock observations to build confusion matrix for gaps true/false positive/neagtive detections. 
         \item Doing a complete census of all globular clusters?

### 2024-10-28 paper/main.tex
first draft of the discussion

@@
     \includegraphics[width=\linewidth]{stream_on_sky_Pal5_monte-carlo-009_pouliasis2017pii-GCNBody.png}
-    \caption{An instsance of the Palomar 5 stream created modeling the cluster as a plummer sphere within an axis-symmetric galactic potential plus 164 other galactic globular clusters.
+    \caption{An instsance of the Palomar 5 stream created modeling the cluster as a Plummer sphere within an axis-symmetric galactic potential plus 164 other galactic globular clusters.
 The top plot is a scatter plot of star-particles that escaped the cluster due to tidal forces, the bottom plot shows the marginalized 1D density profile in longitude.
@@
     \end{equation}\label{eq:GCNBody} where $\vec{r}$ indicates the galacto-centric position, the index $i$ indicates a globular cluster of interest with $R_i,z_i$ being its galactocentric positions in cylindrical coordinates.
-The index $j$ indicates the other globular cluster that are summed over where: $m_j$ is its mass, $b_j$ is its plummer scale radius, and $r_j$ is it's galacto-centric position.
+The index $j$ indicates the other globular cluster that are summed over where: $m_j$ is its mass, $b_j$ is its Plummer scale radius, and $r_j$ is it's galacto-centric position.
 $\Phi$ is the axis-symmteric potential of the model used for the Milky Way.
@@
 
-    We then integrated the evolution of these particles forward in time to the present day.
+    We then integrate the evolution of these particles forward in time to the present day.
 During this forward integration, each star particle feels the gravitational influence of the galaxy and all globular clusters, but not other star particles.
-We do not perform direct N-body simulations between star particles as in eq.~\ref{eq:GCNBody}; instead, at each time step, we load the positions of all globular clusters, as found from the initial integration, and apply their forces to the star particles.
+We do not perform direct N-body simulations between star particles as in eq.~\ref{eq:GCNBody}; in other words, we are using test-particle and solving the restricted three body problem.
+At each time step, we load the positions of all globular clusters, as found from the initial integration, and apply their forces to the star particles.
 The equation of motion is thus:
@@
 
-    \begin{itemize}
-      \item Would these gaps be observable? We are in the max info limit, what about observational difficulties?
-      \item the detected track of Palomar 5 is much shorter than my simulations. Does this mean cluster gaps sould excedingly rare?
-      \item future works?
-      \item bars substructure?
-      \item link to tstrippy? open source
-      \item add the new simulations to etidal public database?
-      \item the new proposed instrument and survey by Conroy and Bonaca to get improve stream data? https://via-project.org/. How to cite or include in the discussion?
-      \item \textbf{What else?}
-      \item Future potential works
-      \begin{itemize}
-        \item mock observations to build confusion matrix for gaps true/false positive/neagtive detections. 
-        \item Doing a complete census of all globular clusters?
-        \item doing a compelte census of all observed streams? 
-        \item comparing to dark matter subhaloes?
-        \end{itemize}
-      \end{itemize}
+  In the future, I would like to investigate stream detections in a full mock observation experiment and build a confusion matrix for the number of true/false positivies/negatives for the stream gaps from globular cluster and dark matter subhaloes.
+Like this, we will be able to understand what data quality is necessary to make such distinctions and what signs in the streams from substructure will be recoverable.
+This could support the https://via-project.org/ mission by Bonaca and Conroy.
+Additionally, I would like to expand this work to include all gap creation rates for all observed stellar streams.
+
+\section*{Data Availability}
+
+  The code used to produce the data in these simulations is publically available at https://github.com/salvatore-ferrone/tstrippy.
+Examples on the codes installation and use are provided at https://tstrippy.readthedocs.io.
+The simulations will also be made publically available at etidal-project.obspm.fr.
 

### 2024-10-28 paper/main.tex
fixed up comment on NGC2808 impact on number of gaps per simulation

@@
       \hline
-    \end{tabular} Excluding NGC~2808, we are left with 29 gaps, leading to an expected range of 0.6-1.5 gaps per simulation.
+    \end{tabular} If we consider NGC 2808 to be an outlier and exclude it from the experiment, we observe an average of 0.6 gaps per simulation.
+Including all observed gaps results in an average of 1.5 gaps per simulation.
 

### 2024-10-28 paper/main.tex
trying to beef up motivation for this work

@@
   One of the interests in studying stellar streams, is that they can constrain the gravitational field of their host galaxies, particularly the Milky Way.
-ompared to the measurement of the HI rotation curve, stellar streams offer the opportunity to investigate the potential of the host galaxy over a wide range of distances, reaching the outermost regions of the halo.
+Compared to the measurement of the HI rotation curve, stellar streams offer the opportunity to investigate the potential of the host galaxy over a wide range of distances, reaching the outermost regions of the halo.
 For example, \citet{2011MNRAS.417..198V} demonstrated how stellar streams can infer the mass and scale parameters of dark matter haloes using various amounts of observational data, from basic right ascension and declination to full six-dimensional phase space information.
@@
 
-  Our study focuses on the influence of globular clusters on stellar streams, specifically examining whether globular clusters could produce gaps similar to those expected from dark matter subhalos.
-If globular clusters do indeed create such gaps, this factor must be considered when using stellar streams to detect dark matter.
-An additional and unique aspect of our study is that we focus on modeling real stellar streams.
-We focus on the Palomar 5 stream, which is well-studied, has observable gaps \citep{2020ApJ...889...70B}, and is highly visible due to its position above the galactic disk.
-We present a result from one of our simulations in Fig.~\ref{fig:stream_on_sky}, and the following sections describe our modeling of the stream, the gravitational interactions with the galaxy and other globular clusters, and the statistical analysis of perturbations affecting the stream.
+  Previous studies have predominantly focused on the creation of gaps in stellar streams due to dark matter subhalo passages, with some exceptions that considered giant molecular clouds.
+If globular clusters can indeed create such gaps, this factor must be taken into account when using stellar streams to infer the population of dark matter subhaloes.
+In our study, we focus on Palomar 5, one of the most extensively studied and well-constrained globular cluster streams, known for its ease of observation.
+We present a result from one of our simulations in Fig.
+\ref{fig:stream_on_sky}.
+The following sections detail our modeling of the stream, its gravitational interactions with the galaxy and other globular clusters, and the statistical analysis of perturbations impacting the stream.
 

### 2024-10-28 paper/main.tex
another touch up

@@
 If globular clusters can indeed create such gaps, this factor must be taken into account when using stellar streams to infer the population of dark matter subhaloes.
-In our study, we focus on Palomar 5, one of the most extensively studied and well-constrained globular cluster streams, known for its ease of observation.
+In our study, we will fill this gap in the literature by simulating the creation of gaps in the Palomar 5~stream from the galactic globular clusters.
+We choose Palomar~5 since it is one of the most extensively studied and well-constrained globular cluster streams, known for its ease of observation.
 We present a result from one of our simulations in Fig.

### 2024-10-28 paper/main.tex
no draft mode
added tables

@@
 Specifically, for each of the 50 realizations of the Palomar 5 stream, we saved $5000$ intermediate timesteps which is a temporal resolution of 1 million years.
-To provide an estimate of the size of our simulations, we have:
-    \begin{tabular}{c|c|c|c}
-      $N_p$ & $N_{\textrm{ts}}$ & $N_{\textrm{phase}}$ & $N_{\textrm{sampling}}$  \\
-      \hline
-      $100000$ & $5000$ & $6$ & $50$   \\
-    \end{tabular} 
-    where $N_p$  is the number of particles, $N_{\textrm{ts}}$ is the number of time-steps saved, $N_{\textrm{phase}}$ is the number of phase space coordinates, and $N_{\textrm{sampling}}$ is the number of monte-carlo samplings of the initial conditions.
+We provide the parameters that specify our data volumne in Table~\ref{tab:data_volume}.
 Using single precision floating point numbers, the size of our simulations is approximately:
@@
     \end{equation}
+
+    \begin{table}[h]
+      \centering
+      \begin{tabular}{|c|c|c|c|}
+          \hline
+          $N_p$ & $N_{\textrm{ts}}$ & $N_{\textrm{phase}}$ & $N_{\textrm{sampling}}$ \\
+          \hline
+          $100000$ & $5000$ & $6$ & $50$ \\
+          \hline
+      \end{tabular}
+      \caption{The parameters determining the data volume of this experiment.
+where $N_p$  is the number of particles, $N_{\textrm{ts}}$ is the number of time-steps saved, $N_{\textrm{phase}}$ is the number of phase space coordinates, and $N_{\textrm{sampling}}$ is the number of monte-carlo samplings of the initial conditions.}
+      \label{tab:data_volume}
+  \end{table}
 
@@
 
-    Overall, across 50 simulations, we observe 73 gaps. The distribution of simulations with a given number of gaps is as follows: \begin{tabular}{|l|c|c|c|c|c|}
+    Overall, across 50 simulations, we observe 73 gaps.
+The distribution of simulations with a given number of gaps is provided in Table~\ref{table:gap_distribution}.
+If we consider NGC 2808 to be an outlier and exclude it from the experiment, we observe an average of 0.6 gaps per simulation.
+Including all observed gaps results in an average of 1.5 gaps per simulation.
+
+    \begin{table}[h]
+      \centering
+        \begin{tabular}{|l|c|c|c|c|c|}
           \hline
@@
           \hline
-    \end{tabular} If we consider NGC 2808 to be an outlier and exclude it from the experiment, we observe an average of 0.6 gaps per simulation.
-Including all observed gaps results in an average of 1.5 gaps per simulation.
+        \end{tabular}
+      \caption{The gap occurrences.}\label{table:gap_distribution}
+    \end{table}
 

### 2024-10-28 paper/main.tex
compiled without draft mde
eqeq typo

@@
 
-      While Fig.\ref{fig:impact_geometry_statistics} demonstrates that mass, relative velocity, or impact parameter alone (from Eq.\ref{eqeq:change_in_momentum}) cannot predict gap formation, one interesting result emerges: no gaps are created from interactions with impact parameters greater than 200~pc.
+      While Fig.\ref{fig:impact_geometry_statistics} demonstrates that mass, relative velocity, or impact parameter alone (from Eq.\ref{eq:change_in_momentum}) cannot predict gap formation, one interesting result emerges: no gaps are created from interactions with impact parameters greater than 200~pc.
 The stream widths are roughly 200~pc, with a radial thickness of about 100~pc.

### 2024-10-30 paper/main.tex
added comments post disucssion with Paola

@@
 I need to discuss this with another human}).
-\citet{2019Galax...7...81Z} provides a review for the different proposed mechanisms of forming dark matter haloes from perturbation theory and deviations from a uniformly dense medium and also disccses techniques and limitations for modelign such physical phenomena.
-\citet{2019Galax...7...81Z} rightly states that until particles are discovered $\Lambda$-CDM remains a hypothesis.
-Since the true nature of dark matter is elusive.
-The possibilities for sub-galactic behaviour and structure of dark matter are vaste.
+\citet{2019Galax...7...81Z} provides a review for the different proposed mechanisms of forming dark matter haloes from perturbation theory and deviations from a uniformly dense medium and also disccses techniques and limitations for modeling such physical phenomena.
+\salvatore{OK.
+The most important point from Zavala to highlight is there information baked within a population of dark matter haloes, cioè fundamental parameters.
+Only certain fundamental parameters of dark matter could give a certain population.
+Having information thus on the population of streams will give information on the nature of dark matter.} \citet{2019Galax...7...81Z} rightly states that until particles are discovered $\Lambda$-CDM remains a hypothesis.
+Since the true nature of dark matter is elusive the possibilities for sub-galactic behaviour and structure of dark matter are vaste.
 However, as proposed by Carlberg in 2012, a most promising tool to set limits on the number and mass density of the sub-halos, are stellar streams.
@@
 The solo-streams can be explained by a larger globular cluster population whose memebers have been disolved due to two body relaxation and tidal stripping.
-For instance, Z Wan 2020(???) reports that the Phenoix stream is dynamically consistent from a globular cluster yet far more metal poor than any cluster ever observed, galactic or otherwise.
+For instance, Z Wan 2020( can't find the citation???) reports that the Phenoix stream is dynamically consistent from a globular cluster yet far more metal poor than any cluster ever observed, galactic or otherwise.
 This implies that the stream is from a completely disolved globular cluster which comes from a population of clusters that have not survived until modern times.
@@
 I think using \citet{2021MNRAS.505.5957B}'s mass estimate is still ok because it is similar in order of magnitude to cluster + stream from \citet{2017ApJ...842..120I}.
+}
+
+    \salvatore{What's the point?
+In the end were more interested in the streams for probing the galaxy.
+Having a more accurate model of the internal dynamics Palomar 5 would change the luminosity per square degree.
 }
@@
 \section{Discussion}
+
+  \salvatore{Model limits? lets discuss them here.}
 

### 2024-10-31 paper/main.tex
Implemented alessandra's comments

@@
 Thus, currently there are about 70 progenitor-less streams and about 145 stream-less clusteres.
-The solo-streams can be explained by a larger globular cluster population whose memebers have been disolved due to two body relaxation and tidal stripping.
+The solo-streams can be explained by a larger globular cluster population whose members have been disolved due to two body relaxation and tidal stripping.
 For instance, Z Wan 2020( can't find the citation???) reports that the Phenoix stream is dynamically consistent from a globular cluster yet far more metal poor than any cluster ever observed, galactic or otherwise.
@@
 Another factor is indeed detectability.
-Additionally, there are mant cluster in the internal regional of the galaxy, a region overwhich \texttt{streamfinder} was not applied since the high stellar density implicates a computational cost that is currently far to expensive.
+Additionally, there are many cluster in the internal regional of the galaxy, a region overwhich \texttt{streamfinder} was not applied since the high stellar density implicates a computational cost that is currently far to expensive.
 Also, \citet{2018MNRAS.474.2479B} demonstrates that internal dynamics preferentially kicks out small stars first, which have lower luminosities.
@@
   Previous studies have predominantly focused on the creation of gaps in stellar streams due to dark matter subhalo passages, with some exceptions that considered giant molecular clouds.
+To our knowledge, no study has yet been perfomed investigating the possibility of globular clusters greating gaps in streams.
 If globular clusters can indeed create such gaps, this factor must be taken into account when using stellar streams to infer the population of dark matter subhaloes.
@@
 The following sections detail our modeling of the stream, its gravitational interactions with the galaxy and other globular clusters, and the statistical analysis of perturbations impacting the stream.
+
+  \salvatore{Alessandra dice che va messo molto di più in evidenza il fatto che nessuno sa se le predizioni cosmologiche sono validi sulla scala galactica riguarda ai sotto aloni.
+.
+Forse va cambiato l'ordine del discorso.
+To give more focus to the intro, I would start introducing GCs, their streams and their use as DM probes.
+Then you can say something about dark matter, the fact that we don't it's distribution (which makes streas so important) and that we don't know if there are the predicted subhalos.
+You can continue saying that gap in GCs have been used to put constraints on these subhalo number but...
+and then you continue with the literature, what is missing and what you did.}
 
@@
 This model is time-independent throughout our simulations.
-To balance data volume and computation time, we use only one galactic potential, which should suffice.
+\salvatore{This model satisfies a series of observational constraints such as local solar stellar density, the galactic rotation curve, etc, similarly to other galactic models such as Bovy2014, McMillian.} However, To balance data volume and computation time, we use only one galactic potential, which should suffice.
 \citet{2021MNRAS.505.5978V} found that only a few outer globular clusters are strongly affected by different potential models, with kinematic uncertainties being the dominant factor in orbital solution differences.
-Similarly, \citet{2024MNRAS.528.5189G} generated a globular cluster mass-loss catalog using seven different potential models and found their debris distributions to be model-independent.
+Similarly, \citet{2024MNRAS.528.5189G} generated a globular cluster mass-loss catalog using seven different potential models and found their debris distributions to be model-independent, similarly to \citet{2023A&A...673A..44F}.
 While the clusters' positions may depend on the model, we assert that interaction rates and stream formation are unaffected by the choice of galactic model.
 
-    The primary methodological departure from \citet{2023A&A...673A..44F} is that now we include globular cluster interactions.
-Instead of treating the clusters as point masses that only experience the gravitational potential of the galaxy, we now compute N-body interactions between the clusters during the backward integration over 5 billion years.
+    There are two primary methodological departures from \citet{2023A&A...673A..44F}.
+Previously, the clusters only experienced the galaxy.
+Now, we include N-body forces between the clusters.
+Secondly, the clusters are not point masses but are softened as Plummer spheres each with their own scale radii.
 The equations of motion for the globular cluster system are thus:
@@
 $\Phi$ is the axis-symmteric potential of the model used for the Milky Way.
+Note that the mass and scale radii are constant in these simulations and we do not take into account the internal evolution of the globular clusters.
 
-    At their computed positions 5 Gyr ago, Palomar 5 was assigned 100,000 particles sampled from a Plummer distribution to represent its stellar content.
-For each set of initial conditions, the particle were resampled based on uncertainties in mass, which is an improvement from \citet{2023A&A...673A..44F}.
+    At the end of this integration we have positions and velocities of the entire globular cluster system 5 Gyr ago.
+We sample 100,000 particles from a Plummer distribution using the same total mass and characteristic radius for the orbital integration.
+We repeate this sampling for each of the 50 Monte-Carlo initial conditions.
+Considering the mass uncertainties is an improvement from \citet{2023A&A...673A..44F} where the most probable mass was used for each Monte-Carlo sampling.
 
-    We use a mass estimate for Palomar 5 taken from \citet{2021MNRAS.505.5957B}, $1.34\pm0.24\times10^{4}M_{\odot}$--which is a contested value.
+    Regarding Palomar 5's mass, we use a mass estimate for Palomar 5 taken from \citet{2021MNRAS.505.5957B}, $1.34\pm0.24\times10^{4}M_{\odot}$--which is a contested value.
 
@@
 During this forward integration, each star particle feels the gravitational influence of the galaxy and all globular clusters, but not other star particles.
-We do not perform direct N-body simulations between star particles as in eq.~\ref{eq:GCNBody}; in other words, we are using test-particle and solving the restricted three body problem.
+We do not perform direct N-body simulations between star particles as we did for globular clusters.
+In other words, we are using test-particle and solving the restricted three body problem.
+The particles do not exert/feel gravitational forces upon/from one another.
 At each time step, we load the positions of all globular clusters, as found from the initial integration, and apply their forces to the star particles.
@@
       \ddot{\vec{r}}_p = -\nabla \Phi(R_p,z_p) + \left.\sum_{j} \frac{Gm_j}{\left(|\vec{r}_j(t) - \vec{r}_p|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j(t)- \vec{r}_p\right),
-      \end{equation} \label{eq:equation_of_motion_particle} where the index $p$ represents one of the 100,000 particles of interest.
+      \end{equation} \label{eq:equation_of_motion_particle} where the index $p$ represents one of the 100,000 particles of interest, and $j$ indexes over the globular clusters as in Eq.~\ref{eq:GCNBody}
 
@@
     The mass-size relation, shown in the top-right panel, reveals a lower limit of $10^5~\textrm{M}_{\odot}$ for clusters that provoke gaps.
-While mass and size alone are insufficient predictors, this empirical limit is useful because relative velocity and impact parameters are historical values.
+While mass and size alone are insufficient predictors, this empirical limit is useful because relative velocity and impact parameters are inacessible.
+Concentration or the density profile of the perturber has a must less significant effect respect to the total mass and subsequent evolution of the gap as shown in Fig.~3 of \citet{2016MNRAS.457.3817S}.
 The bottom-left panel offers a more physically meaningful result: gap-creating impacts tend to impart more momentum into the stream than non-gap impacts, though with some overlap.
@@
 Additionally, as \citet{2016MNRAS.457.3817S} demonstrated, the dispersion of action-frequencies in the stream plays a role.
+The frequencies being the rate at which a particle completes a cycle in phase space.
+For instance, the frequency corresponding to the azimuthal action, $J_\phi$, is $\dot\theta_\phi = - \frac{\mathcal{\partial H}}{\partial J_\phi}$ and where $\theta_\phi$ is the angle describing the position of the particle phase space between momentum ($p_\phi$) and position ($\phi$) axes---where $\phi$ is the azimuhtal angle between the x-y axes in physical space.
 Regions of the stream with well-separated frequencies are more susceptible to gap formation, while those with a wide frequency range (near the cluster) tend to erase the history of impacts.
@@
     \item \textbf{Mass Threshold}: Perturbing clusters have masses above $10^5 M_\odot$; no perturber has a mass smaller than this.
-    \item \textbf{Galactic radius}: Impacts preferentially occur at shorter galactic radii.
+    \item \textbf{Galactic radius}: Impact occurrences negatively correlates with galacto-centric distance.
     \item \textbf{Momentum and Frequency Change}: Gap creation is linked to the momentum imparted to the stream during the flyby, yet not entirely.
@@
 
-  On the other hand, for a stream length of 10~kpc reported from \citet{2020ApJ...889...70B} and given our gap creation rate, in 5~Gyr we would expect to see about 0.5 gaps, in 10~Gyr's we would expect 1.
+  On the other hand, to provide a crude estiamte, consider  a stream length of 10~kpc reported from \citet{2020ApJ...889...70B} and given our gap creation rate.
+In 5~Gyr we would expect to see about 0.5 gaps, in 10~Gyr's we would expect 1.
 Therefore, finding two gaps from globular clusters would not be impossible.

### 2024-11-07 paper/main.tex
adding edits

@@
 \section{Introduction}
-
   Over a century ago, \citet{1933AcHPh...6..110Z} postulated the existence of dark matter to explain the excessively large velocity dispersions of galaxies within their galaxy clusters.
@@
 \citet{1970IAUS...38..351F,1970ApJ...161..802F} elaborated on the dynamics of disk galaxies and briefly acknoledged in the appendix of his paper that the rotation curves of NGC 300 and M~33 cannot be fully explained by disk models and that there must be the same amount of mass or more that is undetected and beyond the extend of the galaxies to explain their rotation curves.
-Concurrently \citet{1970ApJ...159..379R} studied the rotation of Andromenda and observed it to be flat and not decreasing as one would expect from a potential with more mass concentration in the center.
+Concurrently \citet{1970ApJ...159..379R} studied the rotation of Andromenda and observed it to be flat outside its effective radius and not decreasing as one would expect from a potential with more mass concentration in the center.
+
   Eventually, astronomers returned to Zwicky's non-luminous matter to explain these fast-rotating disks, as discussed in \citet{1983SciAm.248f..96R}'s review.
-\salvatore{I see why we cite freeman70 now.
-there are two comments in the appendix stating that there is as much mass beyond the optical limit as there is within it for two different galaxoes.
-NGC 300 and M 33.
-It's not clear to me why we cite Rubin 70.
-The rotation curve is flat, to me implies a logarithmic potential and hence we need dark matter.
-However, I don't get any impression of missing matter when I read the text.}
-
 Beyond the galactic scale, dark matter is essential for explaining many cosmological phenomena.
@@
 A key focus today is detecting dark matter on sub-galactic scales, because at these scales different cosmological scenarios predict different lumpiness of the halos.
-For instance, \citet{2008ApJ...679.1260M} computed cosmological simulations and compared the Via Lactea simulations \citep{2008ApJ...679.1260M} and demonstrated (\salvatore{I don't quite understand Madau 2008.
-The main result is that based on the assumptions you can produce a variance in the number of subhaloes that are consistent with global properties of a milky-way like galaxy?
-I need to discuss this with another human}).
-\citet{2019Galax...7...81Z} provides a review for the different proposed mechanisms of forming dark matter haloes from perturbation theory and deviations from a uniformly dense medium and also disccses techniques and limitations for modeling such physical phenomena.
-\salvatore{OK.
-The most important point from Zavala to highlight is there information baked within a population of dark matter haloes, cioè fundamental parameters.
-Only certain fundamental parameters of dark matter could give a certain population.
-Having information thus on the population of streams will give information on the nature of dark matter.} \citet{2019Galax...7...81Z} rightly states that until particles are discovered $\Lambda$-CDM remains a hypothesis.
+
+  \citet{2019Galax...7...81Z} provides a review for the different proposed mechanisms of forming dark matter haloes theoretically and how to model them pratically, showcasing how different assumptions for the particle behavior of dark matter influences the mass and size distribution of sub haloes.
+The authors rightly state that until particles are discovered $\Lambda$-CDM remains a hypothesis.
 Since the true nature of dark matter is elusive the possibilities for sub-galactic behaviour and structure of dark matter are vaste.
-However, as proposed by Carlberg in 2012, a most promising tool to set limits on the number and mass density of the sub-halos, are stellar streams.
+Obtaining constraints on the distirbution of dark matter subhaloes in turn will place constraints on the particle nature of dark matter.
+As proposed by \citet{2012ApJ...748...20C}, a most promising tool to set limits on the number and mass density of the sub-halos, are stellar streams.
 
   Stellar streams, several-kpc long structures formed by the tidal disruption of globular clusters or dwarf galaxies orbiting a host galaxy.
-They are one of the evidences for the hierarchical nature of the galaxy formation as predicted by the $\Lambda CDM$ model \citep{white91}.
+They are one of the evidences for the hierarchical nature of the galaxy formation as predicted by the $\Lambda CDM$ model \citep{1991ApJ...379...52W}.
 These tidal forces arise due to differential gravitational pulls across extended objects, causing stars farther from the galactic center to lag behind while those closer are pulled away.
@@
 These early works stimulated a vigorous search for stellar streams in the Galaxy in the following years \citep{2006ApJ...643L..17G, 2006ApJ...637L..29B, 2006ApJ...637L..29B, 2009ApJ...693.1118G}.
-The review of Bonaca properly asserts that we have envolved into a new area of stellar stream astronomy Gaia astrometric mission \citep{2016A&A...595A...1G}.
+The review of \citet{2024arXiv240519410B} properly asserts that we have envolved into a new area of stellar stream astronomy Gaia astrometric mission \citep{2016A&A...595A...1G}.
 
   Gaia's characterization of billions of stars in the Milky Way allows us to look for these structures coupling photometry, astrometry and spectroscopy.
-The development and deployment of the Streamfinder algorithm has led to the detection of about 90 stellar streams in the Milky Way.
+The development and deployment of the Streamfinder algorithmm, first presented in \citet{2018MNRAS.477.4063M} has led to the detection of about 90 stellar streams in the Milky Way \citet{2024ApJ...967...89I}.
 \citet{2023MNRAS.520.5225M} provides a summary of these streams and tabulates that about 20 are associated to modern day globular clusters.
-Thus, currently there are about 70 progenitor-less streams and about 145 stream-less clusteres.
+One of the interests in studying stellar streams, is that they can constrain the gravitational field of their host galaxies, particularly the Milky Way.
+Compared to the measurement of the HI rotation curve, stellar streams offer the opportunity to investigate the potential of the host galaxy over a wide range of distances, reaching the outermost regions of the halo.
+For example, \citet{2011MNRAS.417..198V} demonstrated how stellar streams can infer the mass and scale parameters of dark matter haloes using various amounts of observational data, from basic right ascension and declination to full six-dimensional phase space information.
+\citet{2018ApJ...867..101B} reviewed this concept from an information-theoretic point of view, identifying which orbits and configurations of stellar streams that yield the most information about the galactic potential.
+However, the use of single streams for constraining the Galactic potential has led, however, to some ambiguous and not converging results over time.
+For example, \citet{2010ApJ...718.1128L} made use of the Sagittarius stream to infer that the dark matter halo of our Galaxy has a triaxial shape \citep[but see also][]{2004MNRAS.351..643H, johnston05, 2005ApJ...619..807L}; \citet{2016ApJ...833...31B} concluded that the dark matter halo of our Galaxy is nearly spherical at the radii of the Palomar~5 and GD~1 streams.
+Nonetheless, a recent goal of stellar stream astronomy was realized by \citet{2024ApJ...967...89I} constructed the most accurate model of the Milky Way to date based on a set of about 20 streams.
+
+  % \paola{the following sentences seem to me a bit out of the focus of the discussion.
+It is interesting to say that a number of streams do not have a GC counterpart, but I would'nt go too much into the details, bcs you risk to bring the reader into another topic, that of the detection/undetection of streams, that of the metal-poor streams, etc.
+I would reduce what follows, and start again from "One of the interests in studying stellar streams,.." }Thus, currently there are about 70 progenitor-less streams and about 145 stream-less clusteres.
 The solo-streams can be explained by a larger globular cluster population whose members have been disolved due to two body relaxation and tidal stripping.
@@
 
-  \citet{2022MNRAS.516.5331M} showed that many very metal poor streams exist in the Milky Way.
+  % \citet{2022MNRAS.516.5331M} showed that many very metal poor streams exist in the Milky Way.
 On the other hand, it is possible that some globular clusters do not loose a significant number of stars to create tidal tails.
@@
 
-  One of the interests in studying stellar streams, is that they can constrain the gravitational field of their host galaxies, particularly the Milky Way.
-Compared to the measurement of the HI rotation curve, stellar streams offer the opportunity to investigate the potential of the host galaxy over a wide range of distances, reaching the outermost regions of the halo.
-For example, \citet{2011MNRAS.417..198V} demonstrated how stellar streams can infer the mass and scale parameters of dark matter haloes using various amounts of observational data, from basic right ascension and declination to full six-dimensional phase space information.
-\citet{2018ApJ...867..101B} reviewed this concept from an information-theoretic point of view, identifying which orbits and configurations of stellar streams that yield the most information about the galactic potential.
-However, the use of single streams for constraining the Galactic potential has led, however, to some ambiguous and not converging results over time.
-For example, \citet{2010ApJ...718.1128L} made use of the Sagittarius stream to infer that the dark matter halo of our Galaxy has a triaxial shape \citep[but see also][]{2004MNRAS.351..643H, johnston05, 2005ApJ...619..807L}; \citet{2016ApJ...833...31B} concluded that the dark matter halo of our Galaxy is nearly spherical at the radii of the Palomar~5 and GD~1 streams.
-Nonetheless, a recent goal of stellar stream astronomy was realized by \citet{2024ApJ...967...89I} constructed the most accurate model of the Milky Way to date based on a set of about 90 streams.
-
   In this study, we build on this body of work by investigating another important use case for stellar streams: probing dark matter on sub-galactic scales.
-According to simulations by Springel et al.
-(2008), $\Lambda$-Cold Dark Matter ($\Lambda$CDM) predicts that galaxies grow hierarchically, with dark matter clumps forming at a wide range of masses and sizes.
+According to simulations by \citet{2008MNRAS.391.1685S}, $\Lambda$-Cold Dark Matter ($\Lambda$CDM) predicts that galaxies grow hierarchically, with dark matter clumps forming at a wide range of masses and sizes.
 These clumps, or \textit{subhalos}, are predicted to follow a mass distribution with a power-law slope slightly shallower than -2.0.
@@
   Previous studies have predominantly focused on the creation of gaps in stellar streams due to dark matter subhalo passages, with some exceptions that considered giant molecular clouds.
-To our knowledge, no study has yet been perfomed investigating the possibility of globular clusters greating gaps in streams.
-If globular clusters can indeed create such gaps, this factor must be taken into account when using stellar streams to infer the population of dark matter subhaloes.
-In our study, we will fill this gap in the literature by simulating the creation of gaps in the Palomar 5~stream from the galactic globular clusters.
-We choose Palomar~5 since it is one of the most extensively studied and well-constrained globular cluster streams, known for its ease of observation.
+CITE(AMORISCO 2016) CITE(Pearson et al 2017 BAR ON PAL 5) To our knowledge, no study has yet been performed investigating the gravitational impact of globular clusters on stellar streams.
+Since clusters are dense stellar systems, distributed mainly in the Galactic halo, they too should in principle contribute to generating perturbation on stellar streams.
+If globular clusters can indeed create inhomogeneities and gaps in streams, this factor must be taken into account when using stellar streams to infer the population of dark matter subhaloes.
+
+  In our study, we will fill this gap in the literature by studying the gravitational perturbations induced by the whole set of Galactic globular clusters on Palomar 5 stream.
+We choose Palomar~5 since it is one of the most extensively studied and well-constrained globular cluster streams, known for its ease of observation and very long tidal tails.
 We present a result from one of our simulations in Fig.
@@
 The following sections detail our modeling of the stream, its gravitational interactions with the galaxy and other globular clusters, and the statistical analysis of perturbations impacting the stream.
-
-  \salvatore{Alessandra dice che va messo molto di più in evidenza il fatto che nessuno sa se le predizioni cosmologiche sono validi sulla scala galactica riguarda ai sotto aloni.
-.
-Forse va cambiato l'ordine del discorso.
-To give more focus to the intro, I would start introducing GCs, their streams and their use as DM probes.
-Then you can say something about dark matter, the fact that we don't it's distribution (which makes streas so important) and that we don't know if there are the predicted subhalos.
-You can continue saying that gap in GCs have been used to put constraints on these subhalo number but...
-and then you continue with the literature, what is missing and what you did.}
 

### 2024-11-07 paper/main.tex
moved to the end
added pearson and amorisco citations

@@
 \abstract
-  {Holy moly artichokey}
+  {ok}
  % aims heading (mandatory)
@@
  % methods heading (mandatory)
-  {good}
+  {ok}
  % results heading (mandatory)
-  {alright}
+  {ok}
  % conclusions heading (optional), leave it empty if necessary
@@
 Since then, mounting evidence has been assembled to support this claim.
-\citet{1970IAUS...38..351F,1970ApJ...161..802F} elaborated on the dynamics of disk galaxies and briefly acknoledged in the appendix of his paper that the rotation curves of NGC 300 and M~33 cannot be fully explained by disk models and that there must be the same amount of mass or more that is undetected and beyond the extend of the galaxies to explain their rotation curves.
-Concurrently \citet{1970ApJ...159..379R} studied the rotation of Andromenda and observed it to be flat outside its effective radius and not decreasing as one would expect from a potential with more mass concentration in the center.
+\citet{1970IAUS...38..351F,1970ApJ...161..802F} elaborated on the dynamics of disk galaxies and briefly acknoledged in the appendix of his paper that the rotation curves of NGC~300 and M~33 cannot be fully explained by disk models and that there must be the same amount of undetected mass as luminous mass and beyond the extent of the galaxies to explain their rotation curves.
+Concurrently \citet{1970ApJ...159..379R} studied the rotation of Andromenda and observed it to be flat outside its effective radius---not decreasing as one would expect from a gravitational field from a centrally concentrated mass distribution.
 
@@
 
-  \citet{2019Galax...7...81Z} provides a review for the different proposed mechanisms of forming dark matter haloes theoretically and how to model them pratically, showcasing how different assumptions for the particle behavior of dark matter influences the mass and size distribution of sub haloes.
+  \citet{2019Galax...7...81Z} provide a review for the different proposed mechanisms of forming dark matter haloes theoretically and how to model them pratically, showcasing how different assumptions for the fundamental particle behavior of dark matter propogates to the mass and size distribution of sub haloes on the galactic scale.
 The authors rightly state that until particles are discovered $\Lambda$-CDM remains a hypothesis.
@@
   Stellar streams, several-kpc long structures formed by the tidal disruption of globular clusters or dwarf galaxies orbiting a host galaxy.
-They are one of the evidences for the hierarchical nature of the galaxy formation as predicted by the $\Lambda CDM$ model \citep{1991ApJ...379...52W}.
+They are one of the many pieces of evidence supporting the hierarchical nature of the galaxy formation as predicted by the $\Lambda CDM$ model \citep{1991ApJ...379...52W}.
 These tidal forces arise due to differential gravitational pulls across extended objects, causing stars farther from the galactic center to lag behind while those closer are pulled away.
@@
 Subsequent studies\footnote{\citep[see][for more subsequent observation detections of tidal debris, i.e.
-globular cluster stars beyond the tidal radius.]{1997A&A...320..776L, 2000A&A...356..127T, 2000A&A...359..907L, 2001AAS...19910906S, 2003AJ....126..815L}} extended Grillmair's findings to other globular cluster streams, until the discovery made by \citet{2001ApJ...548L.165O,2002AAS...200.1001O, 2003AJ....126.2385O} of long and thin tails outside the Palomar~5 globular cluster, which were shown to be extended for more than $10^\circ$ degrees in the sky.
+globular cluster stars beyond the tidal radius.]{1997A&A...320..776L, 2000A&A...356..127T, 2000A&A...359..907L, 2001AAS...19910906S, 2003AJ....126..815L}} extended Grillmair's findings to other globular cluster streams, until the discovery made by \citet{2001ApJ...548L.165O,2002AAS...200.1001O, 2003AJ....126.2385O} of long and thin tails outside the Palomar~5 globular cluster, which were shown to be extended for more than $10^\circ$ on the sky.
 These early works stimulated a vigorous search for stellar streams in the Galaxy in the following years \citep{2006ApJ...643L..17G, 2006ApJ...637L..29B, 2006ApJ...637L..29B, 2009ApJ...693.1118G}.
@@
   In this study, we build on this body of work by investigating another important use case for stellar streams: probing dark matter on sub-galactic scales.
-According to simulations by \citet{2008MNRAS.391.1685S}, $\Lambda$-Cold Dark Matter ($\Lambda$CDM) predicts that galaxies grow hierarchically, with dark matter clumps forming at a wide range of masses and sizes.
+According to simulations by \citet{2008MNRAS.391.1685S}, $\Lambda$-CDM predicts that galaxies grow hierarchically, with dark matter clumps forming at a wide range of masses and sizes.
 These clumps, or \textit{subhalos}, are predicted to follow a mass distribution with a power-law slope slightly shallower than -2.0.
@@
   \citet{2002MNRAS.332..915I} first suggested that dark matter subhalos could influence stellar streams by diffusing their orbital elements.
-Later, \citet{2012ApJ...748...20C} expanded this idea, proposing that subhalos could create gaps in stellar streams during flyby encounters, where a subhalo approaches closely enough to a segment of a stream and significantly changes the orbits of the closest stars.
-Carlberg 2013 further explored this idea through a probabilistic model of encounter rates based on the expected subhalo distribution.
-\citet{2019ApJ...880...38B} provided observational evidence for this idea, identifying an underdensity in the GD-1 stream that could not be explained by known objects, such as globular clusters, and, interestingly enough, was inconsistent with the $\Lambda$CDM mass-size relationship presented in \cite{2017MNRAS.466.4974M}.
+Later, \citet{2012ApJ...748...20C} expanded this idea, proposing that subhalos could create gaps in stellar streams during flyby encounters by passing closely enough to a segment of a stream and significantly changing the orbits of the closest stars.
+In this work, \citet{2012ApJ...748...20C} further explored this idea through a probabilistic model of encounter rates based on the expected subhalo distribution.
+\citet{2019ApJ...880...38B} provided potential observational evidence for such a gap, identifying an underdensity in the GD-1 stream that could not be explained by known objects, such as globular clusters.
+Interestingly enough the expected mass and size of the perturber causing the gap was inconsistent with the $\Lambda$CDM mass-size relationship presented in \cite{2017MNRAS.466.4974M}.
 
-  Previous studies have predominantly focused on the creation of gaps in stellar streams due to dark matter subhalo passages, with some exceptions that considered giant molecular clouds.
-CITE(AMORISCO 2016) CITE(Pearson et al 2017 BAR ON PAL 5) To our knowledge, no study has yet been performed investigating the gravitational impact of globular clusters on stellar streams.
+  Previous studies have predominantly focused on the creation of gaps in stellar streams due to dark matter subhalo passages, with some exceptions.
+\citet{2016MNRAS.463L..17A} studied the possibility of disc giant molecular clouds gausing gaps in cluster streams.
+\citet{2017NatAs...1..633P} explored the effects of a galactic bar causing gaps and perturbing other aspects of a stellar stream.
+To our knowledge, no study has yet been performed investigating the gravitational impact of globular clusters on stellar streams.
 Since clusters are dense stellar systems, distributed mainly in the Galactic halo, they too should in principle contribute to generating perturbation on stellar streams.
@@
 
-    \salvatore{I'm a bit lost on trying to reconcile the differences between Baumgardt's estiamte and \citet{2017ApJ...842..120I}'s estimate in sections 5.2 and 5.3.
-It looks like 5.2 is based on counts and then supplementing the completness function with dNdM \citet{2001AJ....122.3231G} and demonstrating that Salpeter would overestimate the mass.
-I think 5.3 is showing that an MCMC was used to add more context to the estimate from 5.3.}
-
-    \salvatore{\textbf{If Salpeter over estimates the number of low mass stars}, then wouldn't this appear in the velocity dispersion profile?
-or wait, is \citet{2017ApJ...842..120I} fitting the density for the King, supplementing missing stars by a Mass function from \citet{2001AJ....122.3231G}, and finding the total mass this way?
-i.e.
-no velocity dispersion was investigated?
-I feel like viewing the velocity dispersion directly would provide a good estiamte for the dynamical mass, no?}
-
-    \salvatore{I am lost in Baumgardt's papers trying to flush out how a MF was used to provide their mass estimate. It looks like this information is within Sollima 2017. Need to revisit this. }
-
-    \salvatore{What I say will depend.
-I think using \citet{2021MNRAS.505.5957B}'s mass estimate is still ok because it is similar in order of magnitude to cluster + stream from \citet{2017ApJ...842..120I}.
-}
-
-    \salvatore{What's the point?
-In the end were more interested in the streams for probing the galaxy.
-Having a more accurate model of the internal dynamics Palomar 5 would change the luminosity per square degree.
-}
-
     We then integrate the evolution of these particles forward in time to the present day.
@@
 
+\section{Results}
+
   \subsection{Gap Detection and Characterization} \label{sec:gap_methods}
@@
 Ultimately, we obtained a sample of 250 impacts and compared the impact geometry of those that caused gaps with those that did not.
-
-\section{Results}
 
@@
   \salvatore{Model limits? lets discuss them here.}
+
+  \salvatore{I'm a bit lost on trying to reconcile the differences between Baumgardt's estiamte and \citet{2017ApJ...842..120I}'s estimate in sections 5.2 and 5.3.
+It looks like 5.2 is based on counts and then supplementing the completness function with dNdM \citet{2001AJ....122.3231G} and demonstrating that Salpeter would overestimate the mass.
+I think 5.3 is showing that an MCMC was used to add more context to the estimate from 5.3.}
+
+  \salvatore{\textbf{If Salpeter over estimates the number of low mass stars}, then wouldn't this appear in the velocity dispersion profile?
+or wait, is \citet{2017ApJ...842..120I} fitting the density for the King, supplementing missing stars by a Mass function from \citet{2001AJ....122.3231G}, and finding the total mass this way?
+i.e.
+no velocity dispersion was investigated?
+I feel like viewing the velocity dispersion directly would provide a good estiamte for the dynamical mass, no?}
+
+  \salvatore{I am lost in Baumgardt's papers trying to flush out how a MF was used to provide their mass estimate. It looks like this information is within Sollima 2017. Need to revisit this. }
+
+  \salvatore{What I say will depend.
+I think using \citet{2021MNRAS.505.5957B}'s mass estimate is still ok because it is similar in order of magnitude to cluster + stream from \citet{2017ApJ...842..120I}.
+}
+
+  \salvatore{What's the point?
+In the end were more interested in the streams for probing the galaxy.
+Having a more accurate model of the internal dynamics Palomar 5 would change the luminosity per square degree.
+}
 

### 2024-11-07 paper/main.tex
edited methods

@@
   \subsection{Numerical Methodology}
-    Our numerical methodology follows the approach outlined in Ferrone et al.
-(2023), which we present here for completeness.
+    Our numerical methodology follows the approach outlined in \citet{2023A&A...673A..44F}, which we present here for completeness.
 We begin by extracting the phase space properties (positions, velocities) as well as the masses and half-mass radii of 165 globular clusters from the galactic globular cluster catalog by \cite{2021MNRAS.505.5957B}.\footnote{The Baumgardt catalog has been assembled across a series of works, see: \cite{2020PASA...37...46B,2019MNRAS.482.5138B,2018MNRAS.478.1520B} as well.
 The catalog can be found on the World Wide Web at \href{https://people.smp.uq.edu.au/HolgerBaumgardt/globular/}{https://people.smp.uq.edu.au/HolgerBaumgardt/globular/}.} We handle the uncertainties on the measured phase space parameters through a monte-carlo approach.
-The catalog provides uncertainties for mass, distances, line of sight velocities, proper motions and includes a covariance term between the proper motions---uncertainties for the right ascension, declination, and the characterisitcs radii are not available.
-We repeate simulation the creation of Palomar 5's tidail tails 50 times, for each of which we vary the initial conditions of each cluster by sampling its five-dimensional gaussian uncertaintiy distribution.
+The catalog provides uncertainties for mass, distances, line of sight velocities, proper motions and includes a covariance term between the proper motions---uncertainties for the right ascension and declination are negligible and the characterisitcs radii are not available.
+We sample the these five-dimensional gaussians for each cluster 50 times.
 For the first simulation, which we name \texttt{Sampling 000}, we use the most probable values for the initial conditions.
@@
 The peculiar velocity of the sun was set to $(U_\odot, V_\odot, W_\odot)=(11.1, 12.24, 7.25)$~km~s$^{-1}$.
-The position of the Sun was set to $(x_\odot,y_\odot,z_\odot) = (-8.34,0,0.027)$~kpc, vertical position above the disk was taken from Chen et al 2001 and the galacto-radial distance was taken from Reid et al 2014.
+The position of the Sun was set to $(x_\odot,y_\odot,z_\odot) = (-8.34,0,0.027)$~kpc, vertical position above the disk was taken from \citet{2001ApJ...553..184C} and the galacto-radial distance was taken from \citet{2004ApJ...616..872R}.
 These transformations were performed using \texttt{astropy} \citep{2013A&A...558A..33A}.
 
-    For the galactic potential, we used the second static model from \citet{2017A&A...598A..66P}, a superposition of a thin disk, thick disk, and dark matter halo, with masses and scale lengths provided in Table 1 of \citet{2023A&A...673A..44F}.
+    For the galactic potential, we used the second static model from \citet{2017A&A...598A..66P}, a superposition of a thin disk, thick disk, and dark matter halo, with masses and scale lengths provided in Table~1 of \citet{2023A&A...673A..44F}.
 This model is time-independent throughout our simulations.
-\salvatore{This model satisfies a series of observational constraints such as local solar stellar density, the galactic rotation curve, etc, similarly to other galactic models such as Bovy2014, McMillian.} However, To balance data volume and computation time, we use only one galactic potential, which should suffice.
+This model satisfies a series of observational constraints such as local solar stellar density, the galactic rotation curve, etc, similarly to other galactic models such as Bovy2014, McMillian.
+However, To balance data volume and computation time, we use only one galactic potential, which should suffice.
 \citet{2021MNRAS.505.5978V} found that only a few outer globular clusters are strongly affected by different potential models, with kinematic uncertainties being the dominant factor in orbital solution differences.
@@
 
-    There are two primary methodological departures from \citet{2023A&A...673A..44F}.
-Previously, the clusters only experienced the galaxy.
-Now, we include N-body forces between the clusters.
-Secondly, the clusters are not point masses but are softened as Plummer spheres each with their own scale radii.
+    The primary methodological departure from \citet{2023A&A...673A..44F} is that in the first integration step, when we integrate the globular clusters backward by five billion years, we compute N-body forces between them.
+Specifically, the clusters are not point masses but are softened as Plummer spheres each with their own scale radii.
 The equations of motion for the globular cluster system are thus:
@@
 We repeate this sampling for each of the 50 Monte-Carlo initial conditions.
-Considering the mass uncertainties is an improvement from \citet{2023A&A...673A..44F} where the most probable mass was used for each Monte-Carlo sampling.
+Considering the mass uncertainties is an improvement from \citet{2023A&A...673A..44F} where only the most probable mass was used for each Monte-Carlo sampling.
 
-    Regarding Palomar 5's mass, we use a mass estimate for Palomar 5 taken from \citet{2021MNRAS.505.5957B}, $1.34\pm0.24\times10^{4}M_{\odot}$--which is a contested value.
-
-    We then integrate the evolution of these particles forward in time to the present day.
+    This marks the begining of the second step of our procedure where we then integrate the evolution of these particles forward in time to the present day.
 During this forward integration, each star particle feels the gravitational influence of the galaxy and all globular clusters, but not other star particles.
@@
 
-    The integration was performed using a leapfrog integrator.
-The timestep size was chosen to be based on two critera.
-First, to ensure the conservation of energy to XXX.
-The second, is the drift in position due to numerical error over 5 Gyr was negligible compared to the scale radius of the globular clusters, which is about 5~pc.
-Further details on the timestep selection are discussed in the appendix.
+    The integration was conducted using a leapfrog integrator, chosen for its ability to preserve phase-space volume and conserve the Hamiltonian with each integration step.
+This method is preferable to, for instance, a Runge-Kutta scheme, which can introduce non-physical and significant numerical errors in systems that require long-term stability and energy conservation.
+One drawback of the leapfrog integrator is that it requires a uniform timestep across the entire system, leading to unnecessary computations for slower-moving particles compared to adaptive timestep methods.
+However, energy conservation and phase-space volume preservation are paramount when modeling stellar streams.
+The timestep was therefore set small enough to ensure energy conservation for all particles within the host globular cluster, a value that depends on the system and scales inversely with its concentration.
 
@@
   \end{table}
+
+  This procedure was conducted on Palomar 5 with 50 different samplings of the initial conditions, producing 50 distinct realizations of the cluster's orbits.
+For these simulations, a timestep of 10,000 years was selected to maintain adequate energy conservation and enable \textit{reverse integrability} of the globular cluster system.
+By reverse integrability, we mean the integrator's capability to track the cluster backward in time and then re-integrate it forward along the same trajectory.
+The drift in the final position after forward integration, compared to the initial position from the backward integration, was consistently at least two orders of magnitude smaller than the Plummer scale radius used for Palomar 5.
+This high precision ensures that no fictitious evil numerical forces influenced the system, preventing any artificial mass loss or retention of star particles.
 
@@
 It looks like 5.2 is based on counts and then supplementing the completness function with dNdM \citet{2001AJ....122.3231G} and demonstrating that Salpeter would overestimate the mass.
-I think 5.3 is showing that an MCMC was used to add more context to the estimate from 5.3.}
+I think 5.3 is showing that an MCMC was used to add more context to the estimate from 5.3.}  Regarding Palomar 5's mass, we use a mass estimate for Palomar 5 taken from \citet{2021MNRAS.505.5957B}, $1.34\pm0.24\times10^{4}M_{\odot}$--which is a contested value.
 

### 2024-11-18 paper/main.tex
integrated paola's intro rewrite

@@
 
-\title{Palomar 5 Gaps}
+%\title{Gaps in stellar streams as result of globular cluster close passages: the case of Palomar 5}
+%\title{The impact of globular clusters close passages on stellar streams: the case of Pal5}
+%\title{Globular clusters fly-bys and gap creations in stellar streams.}
+\title{Gaps in stellar streams as a result of globular cluster fly-bys}
 
-\subtitle{Globular clusters as hole punchers}
+\subtitle{The case of Palomar~5}
 
@@
          \and
+       M. Montuori\inst{1}
+       \and
        P. Di Matteo\inst{2}
        \and
-       M. Montuori\inst{1}
+       A. Mastrobuono-Battisti
        \and
-       A. Mastrobuono-Battisti
+       et al.
        }
@@
 \abstract
-  {ok}
+  {Thin stellar streams, such as those resulting from the tidal disruption of globular clusters, have long been known and used as probes of the gravitational potential of our Galaxy, both its visible and dark contents.
+In particular, the presence of under-density regions, or gaps, along these streams is commonly interpreted as being due to the close passage of dark matter sub-haloes.
+}
  % aims heading (mandatory)
-  {ok}
+  {In this work, we investigate the perturbations induced on streams by the passage of dense stellar systems, such as globular clusters themselves, to test the possibility that they too may cause the formation of gaps.
+In particular, we focus on the study of the stream of Palomar 5, a well-known globular cluster in the Galactic halo, which has particularly long tidal tails.
+}
  % methods heading (mandatory)
-  {ok}
+  {For this purpose, we used a particle-test code to simulate the dynamic evolution of Palomar~5 and its tidal tails when Pal~5 is subject to the gravitational action of the Galaxy and its whole system of globular clusters.}
  % results heading (mandatory)
-  {ok}
+  {Our study shows that the tails of Pal~5 are strongly perturbed by the close passage of other clusters, in particular of NGC~2808, NGC~7078, NGC~104, and that these perturbations induce the formation of gaps in the tails.}
  % conclusions heading (optional), leave it empty if necessary
-  {done}
+  {This result suggests that baryonic dense stellar systems such as globular clusters are capable of causing gaps in streams, as already demonstrated in the case of galaxy bar-induced perturbations and giant molecular clouds, and reinforce the idea that the effect of these components should be fully addressed when interpreting the morphology and density structures of observed stellar streams.}
 
@@
 \section{Introduction}
+
   Over a century ago, \citet{1933AcHPh...6..110Z} postulated the existence of dark matter to explain the excessively large velocity dispersions of galaxies within their galaxy clusters.
 Since then, mounting evidence has been assembled to support this claim.
-\citet{1970IAUS...38..351F,1970ApJ...161..802F} elaborated on the dynamics of disk galaxies and briefly acknoledged in the appendix of his paper that the rotation curves of NGC~300 and M~33 cannot be fully explained by disk models and that there must be the same amount of undetected mass as luminous mass and beyond the extent of the galaxies to explain their rotation curves.
-Concurrently \citet{1970ApJ...159..379R} studied the rotation of Andromenda and observed it to be flat outside its effective radius---not decreasing as one would expect from a gravitational field from a centrally concentrated mass distribution.
+\citet{1970IAUS...38..351F,1970ApJ...161..802F} elaborated on the dynamics of disk galaxies and briefly acknowledged in the appendix of his paper that the rotation curves of NGC~300 and M~33 cannot be fully explained by disk models and that there must be the same amount of mass or more that is undetected and beyond the extend of the galaxies to explain their kinematics.
+Concurrently \citet{1970ApJ...159..379R} studied the rotation of Andromeda and observed it to be flat---not decreasing as one would expect from a gravitational potential with more mass concentration in the center.
+Eventually, astronomers returned to Zwicky's non-luminous matter to explain these fast-rotating disks, as discussed in \citet{1983SciAm.248f..96R}'s review.
 
-  Eventually, astronomers returned to Zwicky's non-luminous matter to explain these fast-rotating disks, as discussed in \citet{1983SciAm.248f..96R}'s review.
   Beyond the galactic scale, dark matter is essential for explaining many cosmological phenomena.
@@
 
-  Stellar streams, several-kpc long structures formed by the tidal disruption of globular clusters or dwarf galaxies orbiting a host galaxy.
-They are one of the many pieces of evidence supporting the hierarchical nature of the galaxy formation as predicted by the $\Lambda CDM$ model \citep{1991ApJ...379...52W}.
+  Stellar streams are several-kpc long structures formed by the tidal disruption of globular clusters or dwarf galaxies orbiting a host galaxy.
 These tidal forces arise due to differential gravitational pulls across extended objects, causing stars farther from the galactic center to lag behind while those closer are pulled away.
@@
 These predictions occured well before the first detections which were made by \citet{1995AJ....109.2553G} around some Galactic globular clusters.
+
   Interestingly, this work was a nearly contemporary discovery of the Sagittarius stellar stream by \citet{1994Natur.370..194I} which is the closest example of a stream emerging from a dwarf satellite currently interacting with the Milky Way.
@@
 globular cluster stars beyond the tidal radius.]{1997A&A...320..776L, 2000A&A...356..127T, 2000A&A...359..907L, 2001AAS...19910906S, 2003AJ....126..815L}} extended Grillmair's findings to other globular cluster streams, until the discovery made by \citet{2001ApJ...548L.165O,2002AAS...200.1001O, 2003AJ....126.2385O} of long and thin tails outside the Palomar~5 globular cluster, which were shown to be extended for more than $10^\circ$ on the sky.
-These early works stimulated a vigorous search for stellar streams in the Galaxy in the following years \citep{2006ApJ...643L..17G, 2006ApJ...637L..29B, 2006ApJ...637L..29B, 2009ApJ...693.1118G}.
-The review of \citet{2024arXiv240519410B} properly asserts that we have envolved into a new area of stellar stream astronomy Gaia astrometric mission \citep{2016A&A...595A...1G}.
+The discovery of its prominent tails stimulated a vigorous and successful search in the following years.
+New streams were discovered, mostly taking advantage of SDSS data, but also Pan-STARRS and ATLAS  \citep{2006ApJ...643L..17G, 2006ApJ...637L..29B, 2009ApJ...693.1118G, 2012ApJ...760L...6B, 2013ApJ...769L..23G, 2014ApJ...790L..10G, 2015ApJ...812L..26G, 2014MNRAS.443L..84B, 2016MNRAS.463.1759B, 2017ApJ...847..119G, 2014MNRAS.442L..85K}.
 
-  Gaia's characterization of billions of stars in the Milky Way allows us to look for these structures coupling photometry, astrometry and spectroscopy.
-The development and deployment of the Streamfinder algorithmm, first presented in \citet{2018MNRAS.477.4063M} has led to the detection of about 90 stellar streams in the Milky Way \citet{2024ApJ...967...89I}.
-\citet{2023MNRAS.520.5225M} provides a summary of these streams and tabulates that about 20 are associated to modern day globular clusters.
-One of the interests in studying stellar streams, is that they can constrain the gravitational field of their host galaxies, particularly the Milky Way.
+  It is with the publication of the data from the Gaia astrometric mission \citep{2016A&A...595A...1G} that a new era for stellar stream astronomy has started  \citep{2024arXiv240519410B}.
+Gaia's characterization of billions of stars in the Milky Way allows us to look for these structures coupling photometry, astrometry and, for the brightest stars, spectroscopy.
+The possibility given by Gaia to track stars with coherent movements, over the entirety of the sky, has led to the discovery of dozens of new streams, as Slidr, Sylgr, Ylgr, Fimbulthul, Svol, Gjoll, Leiptr, Phlegeton and Fjorm \citep{2018MNRAS.481.3442M,  2018ApJ...865...85I, 2019ApJ...872..152I}, this latter having an extension of more than 150 degrees in the sky, that is more than five times greater (in projection) than that of Pal~5, according to the most recent estimates provided by \citet{2020MNRAS.493.4978S}.
+The possibility to couple Gaia data with spectroscopic surveys, as for example $S^5$ \citep{2019MNRAS.490.3508L},  has extended the study of stellar streams beyond the quantification of their orbital properties, to a full chemical characterization \citep{2020AJ....160..181J, 2021ApJ...911..149L, 2022ApJ...928...30L, 2024MNRAS.529.2413U}.
+About a hundred stellar streams are currently known in our Galaxy, about 20 of which are associated to known Galactic globular clusters  \citep[see][for a compilation]{2023MNRAS.520.5225M}.
+
+  One of the interests in studying stellar streams is that they can constrain the gravitational field of their host galaxies, particularly the Milky Way.
 Compared to the measurement of the HI rotation curve, stellar streams offer the opportunity to investigate the potential of the host galaxy over a wide range of distances, reaching the outermost regions of the halo.
 For example, \citet{2011MNRAS.417..198V} demonstrated how stellar streams can infer the mass and scale parameters of dark matter haloes using various amounts of observational data, from basic right ascension and declination to full six-dimensional phase space information.
-\citet{2018ApJ...867..101B} reviewed this concept from an information-theoretic point of view, identifying which orbits and configurations of stellar streams that yield the most information about the galactic potential.
-However, the use of single streams for constraining the Galactic potential has led, however, to some ambiguous and not converging results over time.
-For example, \citet{2010ApJ...718.1128L} made use of the Sagittarius stream to infer that the dark matter halo of our Galaxy has a triaxial shape \citep[but see also][]{2004MNRAS.351..643H, johnston05, 2005ApJ...619..807L}; \citet{2016ApJ...833...31B} concluded that the dark matter halo of our Galaxy is nearly spherical at the radii of the Palomar~5 and GD~1 streams.
-Nonetheless, a recent goal of stellar stream astronomy was realized by \citet{2024ApJ...967...89I} constructed the most accurate model of the Milky Way to date based on a set of about 20 streams.
-
-  % \paola{the following sentences seem to me a bit out of the focus of the discussion.
-It is interesting to say that a number of streams do not have a GC counterpart, but I would'nt go too much into the details, bcs you risk to bring the reader into another topic, that of the detection/undetection of streams, that of the metal-poor streams, etc.
-I would reduce what follows, and start again from "One of the interests in studying stellar streams,.." }Thus, currently there are about 70 progenitor-less streams and about 145 stream-less clusteres.
-The solo-streams can be explained by a larger globular cluster population whose members have been disolved due to two body relaxation and tidal stripping.
-For instance, Z Wan 2020( can't find the citation???) reports that the Phenoix stream is dynamically consistent from a globular cluster yet far more metal poor than any cluster ever observed, galactic or otherwise.
-This implies that the stream is from a completely disolved globular cluster which comes from a population of clusters that have not survived until modern times.
+\citet{2018ApJ...867..101B} reviewed this concept from an information-theoretic point of view, identifying which orbits and configurations of stellar streams yield the most information about the galactic potential.
+However, the use of single streams for constraining the potential led to some ambiguous and not converging results over time.
+For example, \citet{2010ApJ...718.1128L} made use of the Sagittarius stream to infer that the dark matter halo of our Galaxy has a triaxial shape \citep[but see also][]{2004MNRAS.351..643H, 2005ApJ...619..800J, 2005ApJ...619..807L}; \citet{2016ApJ...833...31B} concluded that the dark matter halo of our Galaxy is nearly spherical at the distances of the Palomar~5 and GD~1 streams.
+Probably only by using the set of known streams can the degeneracy of the solution be eliminated.
+This approach has been recently used by \citet{2024ApJ...967...89I} who constructed the most accurate model of the Milky Way.
 
@@
 
-  In this study, we build on this body of work by investigating another important use case for stellar streams: probing dark matter on sub-galactic scales.
+  Beyond the global (visible and dark) mass distribution, as we anticipated earlier in this Introduction, streams can be used also to infer the granularity of the dark matter, that is the mass and density of the sub-halos populating our Galaxy.
 According to simulations by \citet{2008MNRAS.391.1685S}, $\Lambda$-CDM predicts that galaxies grow hierarchically, with dark matter clumps forming at a wide range of masses and sizes.
@@
   \citet{2002MNRAS.332..915I} first suggested that dark matter subhalos could influence stellar streams by diffusing their orbital elements.
-Later, \citet{2012ApJ...748...20C} expanded this idea, proposing that subhalos could create gaps in stellar streams during flyby encounters by passing closely enough to a segment of a stream and significantly changing the orbits of the closest stars.
-In this work, \citet{2012ApJ...748...20C} further explored this idea through a probabilistic model of encounter rates based on the expected subhalo distribution.
-\citet{2019ApJ...880...38B} provided potential observational evidence for such a gap, identifying an underdensity in the GD-1 stream that could not be explained by known objects, such as globular clusters.
-Interestingly enough the expected mass and size of the perturber causing the gap was inconsistent with the $\Lambda$CDM mass-size relationship presented in \cite{2017MNRAS.466.4974M}.
+Later, \citet{2012ApJ...748...20C} expanded this idea, proposing that subhalos could create gaps in stellar streams during flyby encounters, where a subhalo approaches closely enough to a segment of a stream and significantly changes the orbits of the closest stars.
+\citet{2012ApJ...760...75C, 2013ApJ...768..171C} applied this idea and looked for the presence of gaps in the known Palomar~5 and GD~1 streams, concluding that the density variations found in their streams were consistent with expectations from $\Lambda$~CDM models.
+\citet{2019ApJ...880...38B} provided further observational evidence for this idea, identifying a gap and a spur in the GD-1 stream that they could not explain by known objects, such as globular clusters, and which they suggested was due to the close passage of a dark matter sub-halo.
+Interestingly enough, however, the recovered properties of this sub-halo (mass and size) were inconsistent with the $\Lambda$CDM mass-size relationship presented in \cite{2017MNRAS.466.4974M}.
+The above-mentioned works are only a few of a much more extensive literature that has explored the impact of dark matter sub-haloes in simulated streams \citep{2016ApJ...828L..10H, 2021MNRAS.507.1999H, 2021JCAP...10..043B, 2024arXiv240402953H, 2024arXiv241021174N}, or searched for their traces in observed streams \citep{2016MNRAS.460.2711T, 2017MNRAS.470...60E, 2020ApJ...889...70B, 2020ApJ...892L..37B}.
 
-  Previous studies have predominantly focused on the creation of gaps in stellar streams due to dark matter subhalo passages, with some exceptions.
-\citet{2016MNRAS.463L..17A} studied the possibility of disc giant molecular clouds gausing gaps in cluster streams.
-\citet{2017NatAs...1..633P} explored the effects of a galactic bar causing gaps and perturbing other aspects of a stellar stream.
-To our knowledge, no study has yet been performed investigating the gravitational impact of globular clusters on stellar streams.
-Since clusters are dense stellar systems, distributed mainly in the Galactic halo, they too should in principle contribute to generating perturbation on stellar streams.
-If globular clusters can indeed create inhomogeneities and gaps in streams, this factor must be taken into account when using stellar streams to infer the population of dark matter subhaloes.
+  In contrast, a more limited number of works have been interested in understanding whether other structures, e.g.
+baryonics, can cause variations in the density of streams and gaps that can be assimilated and confused with those produced by dark matter sub-haloes.
+Among these works, it is worth mentioning the results of \citet{2017NatAs...1..633P} that suggests how the presence of the bar at the center of the Galaxy can perturb the characteristics of a stream such as Pal~5 and generate gaps along its tail.
+That the Galactic bar could have an influence on streams morphology was also discussed by \citet{2016MNRAS.460..497H} and \citet{2016ApJ...824..104P}, in the case of the Ophiuchus stream \citep{2014MNRAS.443L..84B}.
+Beside the Galactic bar, also giant molecular clouds can produce gaps in stellar streams, as shown by \citet{2016MNRAS.463L..17A}.
+All these works thus seem to indicate that baryonic structures can play an important role in tail morphology.
+In this context,  an extensive numerical study, specifically focused on modelling the tails of Palomar~5, under the influence of the Galactic bar, spiral arms, giant molecular clouds and globular clusters, has been realised by \citet{2019MNRAS.484.2009B}, who concluded that both the influence of the bar and that of the giant molecular clouds can leave imprints in Palomar~5 tidal tails similar to those left by dark matter sub-halos, while the effect of globular clusters was found to be negligible.
+A few additional studies have specifically investigated the effect of globular clusters on stellar streams.
+Beside the work by \citet{2017MNRAS.470...60E}, who concluded that density variations, and in particular the dips, they found in Palomar~5 tails could not be due to the passage of globular clusters (but not that this conclusion was not based on a numerical modelling of this effect), more recently \citet{2022ApJ...941..129D} have examined the possibility that the gaps found in the GD-1 stellar stream could be due to the close passage of globular clusters, concluding that this scenario is highly unlikely.
+These first works  seem thus to suggest that the impact of globular clusters on stellar streams is negligible, and this is somewhat surprising, especially for streams of clusters as Palomar~5, which are confined in the inner 20~kpc of the Galaxy, where many other globular clusters do also orbit.
 
-  In our study, we will fill this gap in the literature by studying the gravitational perturbations induced by the whole set of Galactic globular clusters on Palomar 5 stream.
-We choose Palomar~5 since it is one of the most extensively studied and well-constrained globular cluster streams, known for its ease of observation and very long tidal tails.
-We present a result from one of our simulations in Fig.
-\ref{fig:stream_on_sky}.
-The following sections detail our modeling of the stream, its gravitational interactions with the galaxy and other globular clusters, and the statistical analysis of perturbations impacting the stream.
+  In the present study we wish to fill this gap in the literature by reinvestigating this issue and quantify the importance of the effects of passing globular clusters in the vicinity of streams, to understand whether these systems can also be effective - and how often - in altering the distribution of stars in the tails and producing under-dense regions or gaps.
+To this end, on the following pages we will present the results of simulations of the streams of Pal5, subject to the gravitational interaction with the set of 165 Galactic globular clusters for which positions and velocities are known to date, and for which orbits can therefore be reconstructed.
+We chose to simulate Pal~5 because it is a typical halo cluster with extended tails, and because it is a cluster for which the effect of baryonic structures on its stream has already been studied.
+As we will show, and in tension with previous claims, the close passage of other clusters with the stream of Pal~5 is not a rare event, and indeed, in the 50 simulations we ran, we find the formation of numerous gaps, for an average of 1.5 gaps per simulation, generated by 18 different clusters among the whole system of Galactic globular clusters.
 

### 2024-11-18 paper/main.tex
editted the introduction again

@@
 Notably, the colder regions of the CMB temperature map correlate with whole-sky gravitational lensing maps \citep{2020A&A...641A...8P}--since the photons lost energy as they escaped the gravitational wells.
-The $\Lambda$ Cold Dark Matter ($\Lambda$-CDM) model accounts for these observations almost perfectly, proposing that dark matter exists and that its distribution originated from stochastic quantum fluctuations within the first second of the universe led to density variations.
+The $\Lambda$~Cold Dark Matter ($\Lambda$~CDM) model accounts for these observations almost perfectly, proposing that dark matter exists and that its distribution originated from stochastic quantum fluctuations within the first second of the universe led to density variations.
 These variations propagated as the universe expanded, shaping the large-scale structure and distribution of galaxies we observe today.
@@
   \citet{2019Galax...7...81Z} provide a review for the different proposed mechanisms of forming dark matter haloes theoretically and how to model them pratically, showcasing how different assumptions for the fundamental particle behavior of dark matter propogates to the mass and size distribution of sub haloes on the galactic scale.
-The authors rightly state that until particles are discovered $\Lambda$-CDM remains a hypothesis.
+The authors rightly state that until particles are discovered $\Lambda$~CDM remains a hypothesis.
 Since the true nature of dark matter is elusive the possibilities for sub-galactic behaviour and structure of dark matter are vaste.
@@
 These predictions occured well before the first detections which were made by \citet{1995AJ....109.2553G} around some Galactic globular clusters.
+Interestingly, this work was nearly contemporary discovery of the Sagittarius stellar stream by \citet{1994Natur.370..194I} which is the closest example of a stream emerging from a dwarf satellite currently interacting with the Milky Way.
 
-  Interestingly, this work was a nearly contemporary discovery of the Sagittarius stellar stream by \citet{1994Natur.370..194I} which is the closest example of a stream emerging from a dwarf satellite currently interacting with the Milky Way.
-Subsequent studies\footnote{\citep[see][for more subsequent observation detections of tidal debris, i.e.
-globular cluster stars beyond the tidal radius.]{1997A&A...320..776L, 2000A&A...356..127T, 2000A&A...359..907L, 2001AAS...19910906S, 2003AJ....126..815L}} extended Grillmair's findings to other globular cluster streams, until the discovery made by \citet{2001ApJ...548L.165O,2002AAS...200.1001O, 2003AJ....126.2385O} of long and thin tails outside the Palomar~5 globular cluster, which were shown to be extended for more than $10^\circ$ on the sky.
+  Subsequent studies\footnote{For more subsequent observation detections of tidal debris, i.e.
+globular cluster stars beyond the tidal radius see the works of: \citet{1997A&A...320..776L, 2000A&A...356..127T, 2000A&A...359..907L, 2001AAS...19910906S, 2003AJ....126..815L}} extended Grillmair's findings to other globular cluster streams, but were often limited to the detections of stars still close to the cluster tidal radius, until the discovery made by \citet{2001ApJ...548L.165O,2002AAS...200.1001O, 2003AJ....126.2385O} of long and thin tails outside the Palomar~5 globular cluster.
+With a mass of $1.34\pm 0.24 \times 10^4 M_{\sun}$, Palomar~5 is one of the least massive globular clusters in the Galaxy.\citet{2003AJ....126.2385O} showed that its tails contain more mass than the cluster itself.
+The works of \citet{2006ApJ...641L..37G} and \citet{2015MNRAS.446.3297K} showed the tails to be extended for more than $20^\circ$ degrees in the sky.
 The discovery of its prominent tails stimulated a vigorous and successful search in the following years.
@@
 Gaia's characterization of billions of stars in the Milky Way allows us to look for these structures coupling photometry, astrometry and, for the brightest stars, spectroscopy.
-The possibility given by Gaia to track stars with coherent movements, over the entirety of the sky, has led to the discovery of dozens of new streams, as Slidr, Sylgr, Ylgr, Fimbulthul, Svol, Gjoll, Leiptr, Phlegeton and Fjorm \citep{2018MNRAS.481.3442M,  2018ApJ...865...85I, 2019ApJ...872..152I}, this latter having an extension of more than 150 degrees in the sky, that is more than five times greater (in projection) than that of Pal~5, according to the most recent estimates provided by \citet{2020MNRAS.493.4978S}.
-The possibility to couple Gaia data with spectroscopic surveys, as for example $S^5$ \citep{2019MNRAS.490.3508L},  has extended the study of stellar streams beyond the quantification of their orbital properties, to a full chemical characterization \citep{2020AJ....160..181J, 2021ApJ...911..149L, 2022ApJ...928...30L, 2024MNRAS.529.2413U}.
-About a hundred stellar streams are currently known in our Galaxy, about 20 of which are associated to known Galactic globular clusters  \citep[see][for a compilation]{2023MNRAS.520.5225M}.
+The possibility given by Gaia to track stars with coherent movements, over the entirety of the sky, has led to the discovery of dozens of new streams.
+For example, \citet{2018MNRAS.477.4063M} developed the \texttt{streamfinder} algorithm and applied it across a series works of \citep{2018MNRAS.481.3442M,  2018ApJ...865...85I, 2019ApJ...872..152I} to discover the series of streams: Slidr, Sylgr, Ylgr, Fimbulthul, Svol, Gjoll, Leiptr, Phlegeton and Fjorm.
+According to the most recent estimates provided by \citet{2020MNRAS.493.4978S} Fjorm in particular has an extension of more than 150 degrees on the sky which has a projected length more than five times greater than that of Pal~5.
+
+  The possibility to couple Gaia data with spectroscopic surveys, has extended the study of stellar streams beyond the quantification of their orbital properties, to a full chemical characterization.
+For example \citet{2019MNRAS.490.3508L} presents $S^5$, and is one example of a cross-catalogue among other works \citep{2020AJ....160..181J, 2021ApJ...911..149L, 2022ApJ...928...30L, 2024MNRAS.529.2413U}.
+Currently, there are about a hundred stellar streams are currently known in our Galaxy.
+\citet{2023MNRAS.520.5225M} provides a compulation of their tracks on the sky.
+Interestingly, only about 20 streams associated to known Galactic globular clusters.
 
@@
 To date, the smallest observed dark matter halo was detected through gravitational lensing in an Einstein ring and had a mass of $10^8$ solar masses, as described by \citet{2012Natur.481..341V}.
-However, some models predict that dark matter clumps could exist down to the mass of Earth-like planets at least \citep[see][ and discussion in \citet{2021arXiv211101148A}]{2005JCAP...08..003G, wang20}.
+However, some models predict that dark matter clumps could exist down to the mass of Earth-like planets at least \citep[see][ and discussion in \citet{2021arXiv211101148A} ]{2005JCAP...08..003G, wang20}.
 
@@
 \citet{2019ApJ...880...38B} provided further observational evidence for this idea, identifying a gap and a spur in the GD-1 stream that they could not explain by known objects, such as globular clusters, and which they suggested was due to the close passage of a dark matter sub-halo.
-Interestingly enough, however, the recovered properties of this sub-halo (mass and size) were inconsistent with the $\Lambda$CDM mass-size relationship presented in \cite{2017MNRAS.466.4974M}.
+Interestingly enough, however, the recovered properties of this sub-halo (mass and size) were inconsistent with the $\Lambda$~CDM mass-size relationship presented in \cite{2017MNRAS.466.4974M}.
 The above-mentioned works are only a few of a much more extensive literature that has explored the impact of dark matter sub-haloes in simulated streams \citep{2016ApJ...828L..10H, 2021MNRAS.507.1999H, 2021JCAP...10..043B, 2024arXiv240402953H, 2024arXiv241021174N}, or searched for their traces in observed streams \citep{2016MNRAS.460.2711T, 2017MNRAS.470...60E, 2020ApJ...889...70B, 2020ApJ...892L..37B}.
@@
   In contrast, a more limited number of works have been interested in understanding whether other structures, e.g.
-baryonics, can cause variations in the density of streams and gaps that can be assimilated and confused with those produced by dark matter sub-haloes.
+from baryonic matter, can cause variations in the density of streams and gaps that can be assimilated and confused with those produced by dark matter sub-haloes.
 Among these works, it is worth mentioning the results of \citet{2017NatAs...1..633P} that suggests how the presence of the bar at the center of the Galaxy can perturb the characteristics of a stream such as Pal~5 and generate gaps along its tail.
@@
 A few additional studies have specifically investigated the effect of globular clusters on stellar streams.
-Beside the work by \citet{2017MNRAS.470...60E}, who concluded that density variations, and in particular the dips, they found in Palomar~5 tails could not be due to the passage of globular clusters (but not that this conclusion was not based on a numerical modelling of this effect), more recently \citet{2022ApJ...941..129D} have examined the possibility that the gaps found in the GD-1 stellar stream could be due to the close passage of globular clusters, concluding that this scenario is highly unlikely.
-These first works  seem thus to suggest that the impact of globular clusters on stellar streams is negligible, and this is somewhat surprising, especially for streams of clusters as Palomar~5, which are confined in the inner 20~kpc of the Galaxy, where many other globular clusters do also orbit.
+Beside the work by \citet{2017MNRAS.470...60E}, who concluded that density variations, and in particular the dips, they found in Palomar~5 tails could not be due to the passage of globular clusters.
+However, note that this conclusion was not based on a numerical modelling of this effect.
+More recently \citet{2022ApJ...941..129D} have examined the possibility that the gaps found in the GD-1 stellar stream could be due to the close passage of globular clusters, concluding that this scenario is highly unlikely.
+These first works seem to suggest that the impact of globular clusters on stellar streams is negligible, and this is somewhat surprising, especially for streams of clusters as Palomar~5, which are confined in the inner 20~kpc of the Galaxy, where many other globular clusters do also orbit.
 
@@
 While mass and size alone are insufficient predictors, this empirical limit is useful because relative velocity and impact parameters are inacessible.
-Concentration or the density profile of the perturber has a must less significant effect respect to the total mass and subsequent evolution of the gap as shown in Fig.~3 of \citet{2016MNRAS.457.3817S}.
+Concentration or the density profile of the perturber has a much less significant effect respect to the total mass and subsequent evolution of the gap as shown in Fig.~3 of \citet{2016MNRAS.457.3817S}.
 The bottom-left panel offers a more physically meaningful result: gap-creating impacts tend to impart more momentum into the stream than non-gap impacts, though with some overlap.

### 2024-11-18 paper/main.tex
finished implemented paola's comments for the introduction
merged methods

@@
     Our numerical methodology follows the approach outlined in \citet{2023A&A...673A..44F}, which we present here for completeness.
-We begin by extracting the phase space properties (positions, velocities) as well as the masses and half-mass radii of 165 globular clusters from the galactic globular cluster catalog by \cite{2021MNRAS.505.5957B}.\footnote{The Baumgardt catalog has been assembled across a series of works, see: \cite{2020PASA...37...46B,2019MNRAS.482.5138B,2018MNRAS.478.1520B} as well.
-The catalog can be found on the World Wide Web at \href{https://people.smp.uq.edu.au/HolgerBaumgardt/globular/}{https://people.smp.uq.edu.au/HolgerBaumgardt/globular/}.} We handle the uncertainties on the measured phase space parameters through a monte-carlo approach.
-The catalog provides uncertainties for mass, distances, line of sight velocities, proper motions and includes a covariance term between the proper motions---uncertainties for the right ascension and declination are negligible and the characterisitcs radii are not available.
-We sample the these five-dimensional gaussians for each cluster 50 times.
-For the first simulation, which we name \texttt{Sampling 000}, we use the most probable values for the initial conditions.
-
-    We convert the initial conditions from sky coordinates into a galactocentric reference frame.
-We use the local standard of rest and the peculiary motion of the sun as reported by Sch\"onrich 2010.
-The circular velocity at the solar radius was set to: $v_{\text{LSR}} = 240$~km~s$^{-1}$.
-The peculiar velocity of the sun was set to $(U_\odot, V_\odot, W_\odot)=(11.1, 12.24, 7.25)$~km~s$^{-1}$.
-The position of the Sun was set to $(x_\odot,y_\odot,z_\odot) = (-8.34,0,0.027)$~kpc, vertical position above the disk was taken from \citet{2001ApJ...553..184C} and the galacto-radial distance was taken from \citet{2004ApJ...616..872R}.
+We begin by extracting positions in the sky, proper motions, line-of-sight velocities and distances, as well as masses and half-mass radii, of 165 globular clusters from the Galactic globular cluster catalog by \cite{2021MNRAS.505.5957B}.\footnote{The Baumgardt catalog has been assembled across a series of works, see: \cite{2020PASA...37...46B,2019MNRAS.482.5138B,2018MNRAS.478.1520B}.
+The catalog can be found on the World Wide Web at \href{https://people.smp.uq.edu.au/HolgerBaumgardt/globular/}{https://people.smp.uq.edu.au/HolgerBaumgardt/globular/}.} We then convert the initial conditions from sky coordinates into a Galactocentric reference frame, by adopting a velocity for the local standard of rest of $v_{\text{LSR}} = 240$~km~s$^{-1}$ and a peculiary velocity of the Sun equal to $(U_\odot, V_\odot, W_\odot)=(11.1, 12.24, 7.25)$~km~s$^{-1}$, as reported by \citet{2012MNRAS.427..274S}.
+The position of the Sun was set to $(x_\odot,y_\odot,z_\odot) = (-8.34,0,0.027)$~kpc, vertical position above the disk was taken from \citet{2001ApJ...553..184C} and the distance of the Sun to the Galactic center was taken from \citet{2014ApJ...783..130R}.
 These transformations were performed using \texttt{astropy} \citep{2013A&A...558A..33A}.
 
-    For the galactic potential, we used the second static model from \citet{2017A&A...598A..66P}, a superposition of a thin disk, thick disk, and dark matter halo, with masses and scale lengths provided in Table~1 of \citet{2023A&A...673A..44F}.
+    For the galactic potential, we used the second model from \citet{2017A&A...598A..66P}, a superposition of a thin disk, thick disk, and dark matter halo, with masses and scale lengths provided in Table~1 of \citet{2023A&A...673A..44F}.
 This model is time-independent throughout our simulations.
@@
 
-    The primary methodological departure from \citet{2023A&A...673A..44F} is that in the first integration step, when we integrate the globular clusters backward by five billion years, we compute N-body forces between them.
-Specifically, the clusters are not point masses but are softened as Plummer spheres each with their own scale radii.
-The equations of motion for the globular cluster system are thus:
+    \subsubsection*{ \texttt{Full} simulations}
+    There is a primary methodological departure from \citet{2023A&A...673A..44F}.
+In that paper, globular clusters were let evolved under the gravitational effect of the Galaxy only, while now we include also the effect of all the other Galactic globular clusters, by taking into account the $O(N^2)$ interactions among them.
+To quantify these interactions, all clusters are represented by Plummer spheres, each with mass and half-mass radius as reported in the Baumgardt catalogue.
+In the following of this paper, we refer to these simulations, where the globular clusters interactions are taken into account as \texttt{full} simulations.
+\\
+    For these simulations, we proceed into two steps:
+    \begin{enumerate}
+      \item First, starting from the Galactocentric positions and velocities of all 165 Galactic globular clusters, we integrate their orbits back in time for 5~Gyr under the influence of the Galaxy itself, and their mutual influence.
+In the backward integration, the equations of motion for the globular cluster system are thus:
       \begin{equation}
         \ddot{\vec{r}}_i = -\nabla \Phi(R_i,z_i) + \left.\sum_{j\neq i} \frac{Gm_j}{\left(|\vec{r}_j - \vec{r}_i|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j - \vec{r}_i\right),
-    \end{equation}\label{eq:GCNBody} where $\vec{r}$ indicates the galacto-centric position, the index $i$ indicates a globular cluster of interest with $R_i,z_i$ being its galactocentric positions in cylindrical coordinates.
-The index $j$ indicates the other globular cluster that are summed over where: $m_j$ is its mass, $b_j$ is its Plummer scale radius, and $r_j$ is it's galacto-centric position.
-$\Phi$ is the axis-symmteric potential of the model used for the Milky Way.
-Note that the mass and scale radii are constant in these simulations and we do not take into account the internal evolution of the globular clusters.
+      \end{equation}\label{eq:GCNBody} where $\vec{r}$ indicates the galactocentric position, the index $i$ indicates the globular cluster of interest with $R_i,z_i$ being its galactocentric positions in cylindrical coordinates; the index $j$ indicates the other 164 globular clusters that are summed over, where $m_j$ is the mass of the j-th cluster in the sample, $b_j$ is its Plummer scale radius, and $\vec{r_j}$ is its galactocentric position.
+$\Phi$ represents the Galactic smooth potential \citep[][Model~II, in the present case]{2017A&A...598A..66P}.
+Note that the masses and sizes of the globular clusters are kept constant in these simulations and not allowed to vary with time.
+This implies that  we do not take into account in these models the internal evolution of the globular clusters, nor their response to the Galactic tidal field.
 
-    At the end of this integration we have positions and velocities of the entire globular cluster system 5 Gyr ago.
-We sample 100,000 particles from a Plummer distribution using the same total mass and characteristic radius for the orbital integration.
-We repeate this sampling for each of the 50 Monte-Carlo initial conditions.
-Considering the mass uncertainties is an improvement from \citet{2023A&A...673A..44F} where only the most probable mass was used for each Monte-Carlo sampling.
-
-    This marks the begining of the second step of our procedure where we then integrate the evolution of these particles forward in time to the present day.
-During this forward integration, each star particle feels the gravitational influence of the galaxy and all globular clusters, but not other star particles.
-We do not perform direct N-body simulations between star particles as we did for globular clusters.
-In other words, we are using test-particle and solving the restricted three body problem.
-The particles do not exert/feel gravitational forces upon/from one another.
-At each time step, we load the positions of all globular clusters, as found from the initial integration, and apply their forces to the star particles.
-The equation of motion is thus:
+      \item Once the positions and velocities of the entire globular cluster system, as to 5~Gyr ago, are found, we sample Palomar~5 with  100,000 particles which follow a Plummer distribution of mass and half-mass radius as in the Baumgardt catalogue.
+We then integrate the evolution of these particles forward in time to the present day, by taking into account that each particle feels the gravitational potential of the Galaxy, that of its host cluster, and that of all the other clusters in the Galaxy.
+Note that we do not take into account self-gravity among particles, that is in these simulations, particles are simple seismographs of the gravitational field, but do not exert any influence on it.
+The equation of motion of a generic particle among the 100,000 which populate Pal~5 is thus:
     \begin{equation}
@@
 
-    The integration was conducted using a leapfrog integrator, chosen for its ability to preserve phase-space volume and conserve the Hamiltonian with each integration step.
-This method is preferable to, for instance, a Runge-Kutta scheme, which can introduce non-physical and significant numerical errors in systems that require long-term stability and energy conservation.
-One drawback of the leapfrog integrator is that it requires a uniform timestep across the entire system, leading to unnecessary computations for slower-moving particles compared to adaptive timestep methods.
-However, energy conservation and phase-space volume preservation are paramount when modeling stellar streams.
-The timestep was therefore set small enough to ensure energy conservation for all particles within the host globular cluster, a value that depends on the system and scales inversely with its concentration.
+    \end{enumerate}
+
+    The procedure described so far has been repeated 50 times, each time generating a new set of initial conditions, given the uncertainties on proper motions, line-of-sight velocities, distances to the Sun, and masses, of all clusters, as reported in the Baumgardt catalogue.
+We handle these uncertainties  through a Monte-carlo approach, by sampling them with gaussian distribution, and by taking into account the covariance term between the proper motions.
+For the first simulation, we use the most probable values for the initial conditions.
 
@@
 
-  This procedure was conducted on Palomar 5 with 50 different samplings of the initial conditions, producing 50 distinct realizations of the cluster's orbits.
+  \subsubsection*{ \texttt{Reference} simulations}
+
+    To quantify the impact of globular cluster passages to the density of Pal~5 stream, we performed a second set of simulations, which we refer to as the \texttt{reference} simulations in this paper.
+These \texttt{reference} simulations use the same 50 sets of initial conditions as of the \texttt{full} simulations, the same galactic potential but exclude mutual interactions between globular clusters.
+The approach adopted for this second set of simulations is thus equivalent to that adopted already in \citet{2023A&A...673A..44F}.
+In Eq.~\ref{eq:GCNBody} only the gradient of the Galactic potential  is considered.
+In Eq.~\ref{eq:equation_of_motion_particle}, of the second term on the right-side of the equation, only the interaction of Pal~5 particles with Pal~5 itself are considered.
+In other words, sum iterates over only one globular cluster, the host.
+All the interactions with the other clusters in the Galaxy are indeed omitted.
+
+  \subsubsection*{Numerical Stability}
+
+    The integration was conducted using a leapfrog integrator, chosen for its ability to preserve phase-space volume and conserve the Hamiltonian with each integration step.
+This method is preferable to, for instance, a Runge-Kutta scheme, which can introduce non-physical and significant numerical errors in systems that require long-term stability and energy conservation.
+One drawback of the leapfrog integrator is that it requires a uniform timestep across the entire system, leading to unnecessary computations for slower-moving particles compared to adaptive timestep methods.
+However, energy conservation and phase-space volume preservation are paramount when modeling stellar streams.
+The timestep was therefore set small enough to ensure energy conservation for all particles within the host globular cluster, a value that depends on the system and scales inversely with its concentration.
+
     For these simulations, a timestep of 10,000 years was selected to maintain adequate energy conservation and enable \textit{reverse integrability} of the globular cluster system.

### 2024-11-19 paper/main.tex
tweeks to methods

@@
 The top plot is a scatter plot of star-particles that escaped the cluster due to tidal forces, the bottom plot shows the marginalized 1D density profile in longitude.
-Two large gaps are present and are due to the passage of two globular clusters.}
+Two large gaps are present and are due to the passage of two globular clusters.
+$N_p$ indicates the number of particles in a bin while $N$ is the total number of particles which is 100,000.}
     \label{fig:stream_on_sky}
@@
     There is a primary methodological departure from \citet{2023A&A...673A..44F}.
-In that paper, globular clusters were let evolved under the gravitational effect of the Galaxy only, while now we include also the effect of all the other Galactic globular clusters, by taking into account the $O(N^2)$ interactions among them.
+In that paper, globular clusters were let evolved under the gravitational effect of the Galaxy only, while now we include also the effect of all the other Galactic globular clusters, by taking into account the direct N-body interactions among them.
 To quantify these interactions, all clusters are represented by Plummer spheres, each with mass and half-mass radius as reported in the Baumgardt catalogue.
@@
 We then integrate the evolution of these particles forward in time to the present day, by taking into account that each particle feels the gravitational potential of the Galaxy, that of its host cluster, and that of all the other clusters in the Galaxy.
-Note that we do not take into account self-gravity among particles, that is in these simulations, particles are simple seismographs of the gravitational field, but do not exert any influence on it.
+Note that we do not take into account self-gravity among particles.
+The particles expereince the gravitational field yet do not contribute to it.
 The equation of motion of a generic particle among the 100,000 which populate Pal~5 is thus:

### 2024-11-19 paper/main.tex
numerical stability time step discussion

@@
 This method is preferable to, for instance, a Runge-Kutta scheme, which can introduce non-physical and significant numerical errors in systems that require long-term stability and energy conservation.
-One drawback of the leapfrog integrator is that it requires a uniform timestep across the entire system, leading to unnecessary computations for slower-moving particles compared to adaptive timestep methods.
+One drawback of the leapfrog integrator is that it requires a uniform timestep across the entire computation, leading to unnecessary computations for a particle after it escapes form the host cluster.
 However, energy conservation and phase-space volume preservation are paramount when modeling stellar streams.
-The timestep was therefore set small enough to ensure energy conservation for all particles within the host globular cluster, a value that depends on the system and scales inversely with its concentration.
+The time step was therefore set to be small enough to conserve energy for the most interior particles within the cluster---ensuring that a higher mass loss did not arise from numerical error.
+For these simulations, a timestep of 10,000 years was selected to maintain adequate energy conservation.
 
-    For these simulations, a timestep of 10,000 years was selected to maintain adequate energy conservation and enable \textit{reverse integrability} of the globular cluster system.
+    We also checked the \textit{reverse integrability} of the globular cluster system.
 By reverse integrability, we mean the integrator's capability to track the cluster backward in time and then re-integrate it forward along the same trajectory.

### 2024-11-19 paper/main.tex
swapped results and conclusions for Paola's rewrite

@@
 
-  \subsection{Gap Detection and Characterization} \label{sec:gap_methods}
-
-    At the end of the simulations, some gaps in the stellar stream are so prominent that they can be identified visually without any sophisticated quantitative analysis.
-The analysis of the morphology of the tidal tails of Palomar~5 is carried out by comparing simulations with globular clusters (called \texttt{full}) and without them (called \texttt{reference}).
-A preliminary evaluation is performed visually, followed by a quantitative analysis of the differences between the density profiles in the two mentioned cases.
-
-    The \texttt{reference} simulations used the same 50 sets of initial conditions but excluded interactions with globular clusters, except for the host cluster, Palomar 5.
-In other words, the gravitational influence of all other clusters on the star particles within the stream was omitted.
-By comparing the morphological differences between these \texttt{reference} simulations and the \texttt{full} simulations, we could more effectively identify and assess the impact of globular clusters on the stream's structure.
-Details on gap identification are provided in Appendix~\ref{sec:gap_detection}, and the identification of corresponding perturbing globular clusters is detailed in Appendix~\ref{sec:Perturber_Identification}.
-
-    Comparisons between the \texttt{reference} and \texttt{full} simulations were performed in the tail coordinate system.
-This system allows for consistent morphological comparisons by aligning the stream with the cluster's orbit.
-The coordinate system is based on the work of \citet{2004AJ....127.2753D} and is shown in Fig.~\ref{fig:TailCoordinates}.
-Briefly, the x' coordinate represents the position of a particle along the orbit relative to the globular cluster—whether it is ahead or behind the cluster.
-The y' coordinate measures the particle's distance within the orbital plane, where positive values indicate that the particle is farther from the Galactic center and negative values indicate that it is closer.
-
-    Once the gaps and perturbers were identified, we sought to characterize the properties of the globular clusters that caused gaps and differentiate them from those that did not.
-\citet{2013ApJ...775...90C} initiated the theory of gap creation in stellar streams due to gravitational flybys.
-This analysis was extended both analytically and numerically by \citet{2015MNRAS.450.1136E}, who studied uniform-density streams on circular orbits.
-\citet{2016MNRAS.457.3817S} presented the most comprehensive theory, generalizing Erkal et al.'s work to include impacts on streams of varying densities on inclined and eccentric orbits using an action-angle approach.
-
-    The significance of an impact is measured by the change in momentum imparted to the stream stars by the perturber.
-A useful approximation for the change in momentum is the force at the closest approach multiplied by an estimate of the interaction time: \begin{equation} \label{eq:change_in_momentum} \Delta v \approx \text{Force} \times \text{interaction time} = \frac{GM}{b^2} \times 2\frac{b}{v} = 2\frac{GM}{bv}, \end{equation} where $M$ is the mass of the perturber, $b$ is the impact parameter, $v$ is the relative velocity of the perturber with respect to the particle, and $G$ is the gravitational constant.
-We compute this for the five closest approaches of all globular clusters throughout each of the 50 simulations.
-Details on how this computation was performed are given in Appendix~\ref{sec:reconstruction}.
-Ultimately, we obtained a sample of 250 impacts and compared the impact geometry of those that caused gaps with those that did not.
-
-  \subsection*{Overview}
+  \subsection{Overview}
 
@@
       \end{figure*}
-
-    To our knowledge, this is the first study to investigate interactions between globular clusters in this context.
-We included N-body forces on the orbits with the globular clusters, yet did not observe significant changes to orbital properties, such as energy or angular momentum, between any of the globular clusters in any simulation.
-This result is expected given the small sizes and vast distances between the clusters.
-
-    We inspected all simulations upon their completion and present one particularly clear example shown in Fig.~\ref{fig:stream_on_sky} which we selected for its prominent gaps.
-Two of these gaps are visible in galactic coordinates and become even more apparent when marginalizing over latitude to view their 1D density profile over longitude.
-
-    We applied the gap detection methods described in Appendix~\ref{sec:gap_detection}, along with the perturber identification methods from Appendix~\ref{sec:Perturber_Identification}.
-Results from all simulations are presented in Appendix~\ref{sec:gallery_of_gaps}.
-Once we identified the perturbers responsible for each gap, we conducted additional simulations to verify correct identification of the perturbers.
-These additional contain only the suspected clusters and were launched one-at-a-time to isolate the effects of a specific clusters.
-These results are shown in Fig.~\ref{fig:decomposition}.
-A summary of which perturber caused a gap and when is presented in Fig.~\ref{fig:histogram_impact_time}.
 
@@
 
-    NGC~2808 impacts Palomar 5's stream about 200 Myr ago in 44 out of 50 simulations, creating a small gap at a similar position in each case.
-The gap size aligns with gap formation as a dispersion process, where stars gradually spread apart due to slight differences in orbital frequencies.
-Smaller gaps indicate more recent encounters.
-Since this interaction occurred less than one orbital period ago, the orbital solutions remain similar, producing consistent results across simulations.
+    To our knowledge, this is the first study to investigate the gravitational interactions between all Galactic globular clusters.
+While we did not  observe significant changes to orbital properties of the clusters, when subject to their mutual interactions -- a result which is  expected given the small sizes and vast distances between the clusters -- we do see clear effects of the presence of their gravitational field in the properties of Pal~5 stream.
+A particularly clear example of this effect is shown in Fig.~\ref{fig:stream_on_sky} which we selected for its prominent gaps.
+Two of these gaps are clearly visible in galactic coordinates and become even more apparent when marginalizing over latitude to reconstruct the 1D density profile of the stream, as a function of longitude.
 
-    Most gaps appear in the leading tail, which we consider coincidental, as stars escape the cluster's L1 and L2 points at similar rates, and the small distance difference between tails' galactocentric distance is negligible.
+    To make a more quantitative comparison between the \texttt{reference} and \texttt{full} simulations, we adopted to work in the tail coordinate system, in which  the stream is aligned with the cluster's orbit.
+The definition of this coordinate system is based on the work of \citet{2004AJ....127.2753D} and is shown in Fig.~\ref{fig:TailCoordinates}.
+Briefly, in this system the x' coordinate represents the position of a particle along the orbit relative to the globular cluster—whether it is ahead (positive x' ) or behind (negative x') the cluster;  the y' coordinate measures the particle's distance within the orbital plane, where positive values indicate that the particle is farther from the Galactic center and negative values indicate that it is closer.
+Fig.~\ref{fig:decomposition} shows a comparison between one of the 50 realizations of Palomar~5 stream taking into account the gravitational interactions with all other globular clusters in the Galaxy (top panel) and omitting them (bottom panel).
+This comparison clearly shows the presence of two wide ($\sim$ 100~pc and $\sim$ 1~kpc) gaps in the leading tail and of a more subtle underdensity at x'$\sim 5$~kpc (we refer the reader to Appendix~\ref{sec:gap_detection} for a detailed description of the underdensities and gaps detection method).
+To find out which globular clusters were responsible of creating these gaps, and at what time the close passages between these clusters and the Palomar~5 streams took place we estimated the gravitational acceleration along the orbit of Pal~5 and represented it in the ($t$, $\tau$) space, where $t$ is the simulation time, from 5~Gyr in the past to the present day, and $\tau$ is indicates how long it will take for Pal~5 to reach, or how  long ago it passed, a given point on its orbit.
+The use of $\tau$ is advantageous because the growth of the stream is approximately linear in $\tau$ whereas in physical space, streams on eccentric orbits expand and contract depending on the orbital phase.
+Adopting this time space, and reporting the gravitational acceleration along Pal~5 orbit in this space, the identification of the globular clusters which produced the perturbation, and the time at which it occurred becomes straightforward (we refer the reader to Appendix~\ref{sec:Perturber_Identification} for a in-depth description of the procedure).
+In this way, we could identify that the clusters responsible for creating gaps in the simulation of Palomar~5 stream reported in Fig.\ref{fig:decomposition} are NGC~2808, NGC~7078 and NGC~104, and that their close passages occurred 200~Myr, \paola{xx and xx Gyr ago}, respectively.
+To further probe that the three clusters above are responsible of producing the gaps observed in this simulation, we conducted additional experiments by including only the perturbation of each of these three clusters at a time, and neglecting the gravitational perturbations of all the other clusters in the Galaxy.
+The results of these simulations are reported in the middle panels of Fig.~\ref{fig:decomposition} and clearly shows that NGC~2808, NGC~7078 and NGC~104 are the clusters responsible of creating the underdense regions observed in the leading tails of Palomar~5.
+It is worth noting that the times at which the passages of these clusters occurred, according to our analysis, are in agreement with the observed width of the corresponding gaps: the encounter with NGC~2808 being very recent (only 200~Myr ago) its induced gap is still very thin, because it takes time for a perturbation to grow into an extended gap, as it is the case for those induced by the passages of  NGC~7078 and NGC~104, which occurred in earlier times.
+It is also interesting to emphasize that gaps as thin as those generated by the passage of NGC~2808, 200~Myr ago, can be detected by working in the tail coordinate system and by making a differential analysis (\texttt{full} versus \texttt{reference} simulations): they are so thin that they cannot be directly identified in galactic coordinates (see Fig.~\ref{fig:stream_on_sky}).\\
+
+    The analysis presented in Fig.~\ref{fig:decomposition} has been repeated for the whole set of simulations, and is reported in Appendix~\ref{sec:gallery_of_gaps} for completeness.
+From this analysis we can derive a statistical view of: 1) the number of gaps generated on Pal~5 stream by the whole system of Galactic globular clusters; 2) the clusters which generated these gaps and the time history of these perturbations.
+From this, we can then quantify: 3) the properties of the perturbers (their masses, sizes and their orbital parameters) as well as 4) the impact geometry of the encounters, which allows us to understand which encounters are more favorable to generate gaps in Pal~5 stream.
+In the following, we will present results of this analysis, by starting from addressing points 1) and 2).
+
+  \subsection{The history and statistics of gap creations in Palomar~5 stream}\label{sect:history}
+
+    By applying the methodology described above to the whole set of simulations, we can reconstruct the history of close passages of Galactic globular clusters to Palomar~5 stream, in the last 5~Gyr, which -- we recall the reader -- is the time interval investigated in our simulations.
+The results of this analysis are reported in Fig.~\ref{fig:histogram_impact_time} and in Table~xx \paola{I suggest to report the results of Fig 3 also in a Table, in which we report for each GC which produced a gap, the time of close passage and the total number of gaps it generated.} In total, we report the finding of 73 gaps in our 50 simulations, that is 1.5 gaps/simulation, on average, created by 18 different perturbers.
+NGC~2808 impacts Palomar 5's stream about 200 Myr ago in 44 out of 50 simulations, creating a small gap at a similar position to the one reported in Fig.~\ref{fig:decomposition}  in each case.
+Since this interaction occurred less than one orbital period ago, the orbital solutions remain similar, producing consistent results across all simulations, within the uncertainties.
+For encounters which took place at earlier times, the clusters responsible of creating the gaps may vary -- that is, gaps are not created at the same time in all simulations and the parameters of a specific encounter between a  globular cluster and Palomar~5 stream may vary significantly from a simulation to another with the turning back of time, so that in some cases a cluster may produce a gap and in some others the perturbation is not strong or close enough to produce any detectable signature in the Palomar~5 stream.
+In total, beside NGC~2808, we report the creation of 24 gaps in Pal~5 tails, over the last 3~Gyr generated by \paola{15: check this value, I derived it by eye} different clusters, which corresponds to a rate of 0.8 gaps/100~Myr (note that, including NGC~2808, this rate increases to 29.5 gaps/100~Myr).
+At even earlier times, the probability of creating gaps in Palomar~5 streams diminishes significantly, but one must take into account that at these times the tails of Pal~5 have not yet had time to develop sufficiently, and the probability of close encounters with other clusters consequently decreases.
+Overall, across 50 simulations, as previously stated, we observe 73 gaps.
+The distribution of simulations with a given number of gaps is provided in Table~\ref{table:gap_distribution}.
+If we consider NGC~2808 to be an outlier and exclude it from the experiment, we observe an average of 0.6 gaps per simulation.
+Including all observed gaps results in an average of 1.5 gaps per simulation.
+Interestingly, most gaps appear in the leading tail, which we consider coincidental \paola{I am still surprised by the high occurrence of gaps in the leading tail, but I amm still not able to come up with an explanation for this.
+I mean, is it really coincidental?
+I propose to leave the discussion as it is for the moment.
+We will add more, if we come up with an other possible explanation}, as stars escape the cluster's L1 and L2 points at similar rates, and the small distance difference between tails' galactocentric distance is negligible.
 We expect gaps to be equally likely in both tails.
-However, only 8 of the 73 gaps occur in the trailing tail, with 44 linked to NGC2808's flyby.
+However, only 8 of the 73 gaps occur in the trailing tail, with 44 linked to NGC~2808's flyby.
 After removing these, there are 21 gaps in the leading tail and 8 in the trailing.
 By the binomial distribution, the probability of observing up to 8 successes in 29 trials, given a 50\% chance of success, is 1.2\%—unlikely, but possible.
-Additionally, other perturbers impact at consistent times, which may also violate the assumption of independent events, as seen with NGC2808.
-
-    Overall, across 50 simulations, we observe 73 gaps.
-The distribution of simulations with a given number of gaps is provided in Table~\ref{table:gap_distribution}.
-If we consider NGC 2808 to be an outlier and exclude it from the experiment, we observe an average of 0.6 gaps per simulation.
-Including all observed gaps results in an average of 1.5 gaps per simulation.
+Additionally, other perturbers impact at consistent times, which may also violate the assumption of independent events, as seen with NGC~2808.
 
@@
         \end{tabular}
-      \caption{The gap occurrences.}\label{table:gap_distribution}
+        \vspace{0.5cm}
+      \caption{Occurence of gaps in Palomar~5 streams,  in our simulations.
+More specifically, the table reports the number of simulations (second row) for which a given number of gaps (first row) is found.
+}\label{table:gap_distribution}
     \end{table}
 
- \subsection*{Perturber Statistics}
+    Following  \citet{2012ApJ...748...20C} we can also estimate the rate at which gaps are created per unit stream length.
+For our simulations, this rate is given by: \begin{equation} \label{eq:gap_creation_rate} \mathcal{R}_{\textrm{Pal 5}} =  \frac{1}{T}\int_{0}^T l^{-1}(t) \sum_i \delta(t-t_i) dt,\end{equation}where $T$ is the total integration time, $l(t)$ is the length of the stream, and $\sum_i \delta(t-t_i)$ sums over the gap occurrences, with $i$ indexing over the number of gaps in a given simulation.
+Here, $\delta$ represents the Dirac delta function.
+This expression can be simplified to:\begin{equation}\mathcal{R}_{\textrm{Pal 5}} =  \frac{1}{T} \sum_i \frac{1}{l (t_i)}.
+\end{equation}
 
-    We aim to identify general characteristics of perturbers that distinguish them from globular clusters that do not affect the Palomar 5 stream.
-Fig.~\ref{fig:mass_size_plane} presents several analyses to explore this.
+    This computation allows us to analyze the distribution of gap creation rates across all simulations.
+The results are shown in Fig.~\ref{fig:gapcreationrate}.
+This is roughly consistent with a simple estimate of the average gap creation rate: with 72\paola{73?} gaps over 5 Gyr of integration time for a stream about 20 kpc in length, the naive estimate is approximately 0.015~km~s$^{-1}$~kpc$^{-2} = 0.015~\rm{Gyr^{-1}kpc^{-1}}$ .
+This is about double the weighted mean gap creation rate of 0.009~km~s$^{-1}$~kpc$^{-2}$.
+The naive estimate is higher because it does not account for the growth of the stream over time, unlike Eq.\ref{eq:gap_creation_rate}, where the likelihood of impacting a shorter stream is lower.
+
+    \begin{figure}
+      \centering
+      \includegraphics[width=\linewidth]{gap_creation_rate.png}
+      \caption{The distribution of the number of gaps normalized over the total integration time and unit stream length, as described by Eq.\ref{eq:gap_creation_rate} for the whole set of 50 \texttt{full} simulations.
+}
+      \label{fig:gapcreationrate}
+    \end{figure}
+
+  \subsection{Impact geometry and parameters of the perturbers}\label{sect:geometry}
+
+    \begin{figure*}
+      \centering
+      \includegraphics[width=\linewidth]{impact_geometry_statistics.png}
+      \caption{The distribution and relationship between the basic impact variables from Eq.\ref{eq:change_in_momentum} and the close encounter that cause impacts, shown in red against those that do not, shown in blue.
+}
+      \label{fig:impact_geometry_statistics}    
+    \end{figure*}
+
+    With the perturbers identified, we perform statistics in the pursuit of understanding what conditions are necessary for a globular cluster  in order to form a gap on Pal~5 stream.
+We turn to impact theory, which in its simplest form is presented in works such as \citet{2008gady.book.....B}.
+Consider two particles: one stationary and the other moving past it.
+The distance between the two particles at their closest approach is known as the impact parameter.
+To simplify the analysis, the \textit{impulse approximation} is employed, which assumes that the velocity of the perturber remains unchanged during the interaction.
+This assumption simplifies the computation.
+
+    To understand how the impacted particle is perturbed, one needs to compute its change in momentum, which is determined by integrating the force acting on the particle over the duration of the interaction.
+A useful approximation for this change in momentum \paola{(per unit mass)} is the force at the closest approach multiplied by an estimate of the interaction time:
+    \begin{equation} \label{eq:change_in_momentum} \Delta p \approx \text{Force} \times \text{interaction time} = \frac{GMm}{b^2} \times 2\frac{b}{v} = 2\frac{GM}{bv}, \end{equation}where $M$ is the mass of the perturber, $m$ is the mass of the particle, $b$ is the impact parameter, $v$ is the relative velocity of the perturber with respect to the particle, and $G$ is the gravitational constant.
+\paola{ATTENTION : in the last formula there is no dependence on $m$, as  for a change in momentum per unit mass (hence my comment few lines before).
+If you want to keep it as momentum per unit mass, that is a change in velocity, you should also remove $m$ from the other expression (i.e.
+GMm)}
+
+    From this analysis, we can conclude that a more massive perturber, passing closer to the particle, and moving more slowly, will have a greater impact.
+It is important to note that the momentum change is inversely proportional to the velocity of the perturber, which contrasts with the intuition from elastic collisions, such as those between billiard balls, where higher velocities result in greater impacts.
+
+    \citet{2015MNRAS.450.1136E} extended this impact theory from one point mass impacting another point mass, to studying how an extended body impacts a stream.
+This quantifies the change in momentum of a given particle as a function of its distance from the point of greatest impact.
+In their analysis, the perturber was modeled as a Plummer sphere, like in our simulations.
+Since the stream is not zero-dimensional and has length, both the parallel and perpendicular components of the pertuber velocity must be considered.
+Consequently, five parameters determine the change in velocity of a given stream particle: $M$, $r_p$, $b$, $W_\parallel$, and $W_\perp$.
+As detailed in Appendix~\ref{sec:reconstruction}, we calculated these parameters for all our \texttt{full} simulations, by selecting -- for each of them --  the strongest 5 flybys of a perturber with the Pal~5 stream.
+Thus, we compute a total of  250 impacts, and flag those that give way to gaps.
+
+    Visual inspection of the five key impact parameters ($M$, $r_p$, $b$, $W_\parallel$, and $W_\perp$) did not reveal a clear distinction between flybys that create gaps and those that do not.
+Due to the clutter, we omit this graph and instead present...
+In Fig.\ref{fig:impact_geometry_statistics} we report the result of this analysis, by showing the distribution of these impact variables for fly-bys which give rise to gaps and compare it to the distributions of fly-bys which do not lead to gap formation.
+Note that in this figure, we show the total relative velocity rather than separating parallel and perpendicular components, because from a visual inspection, we could not find any specific trends when the two velocity components were plotted separately.
+We also exclude the characteristic cluster radius, as it showed little correlation with the results, likely due to the narrow range of globular cluster radii (see Fig.\ref{fig:mass_size_plane}).
+This factor might be more significant for dark matter subhalos, where size variation is greater.
+
+    While Fig.~\ref{fig:impact_geometry_statistics} demonstrates that mass, relative velocity, or impact parameter alone (from Eq.~\ref{eq:change_in_momentum}) cannot predict gap formation, one interesting result emerges: no gaps are created from interactions with impact parameters greater than 200~pc.
+The stream widths are roughly 200~pc, with a radial thickness\paola{what do you mean by radial thickness?} of about 100~pc.
+This finding is even more evident when examining the $b$-$M$ plane.
+A series of perturbers at roughly $\sim7 \times 10^5 M_\odot$ highlights NGC~2808's flybys, where all encounters with impact parameters under 200~pc result in gaps, while those beyond this distance do not.
+In other words, massive globulars with masses greater than $10^6 \rm{M}_\odot$ are not effective in producing gaps if they pass close to Pal~5 stream with an impact parameter $b > 200$~pc.
+It is interesting to note that for values of the impact parameter below this threshold, even fast encounters ($\delta v > 300$~km/s) can produce gaps.
+\paola{Attention: there is something I do not understand in Fig 5: the range of $\delta v$ values seems not the same in the first row and in the last row (left panel).
+Is there anything to clarify/correct?}
+
+    \begin{figure}
+      \centering
+      \includegraphics[width=1\linewidth]{impact_geometry_statistics_deltaP.png}
+      \caption{The distribution of imparted  change in momentum (per unit mass) from a cluster flyby given by Eq.~\ref{eq:change_in_momentum}.
+The data set includes the top 5 strongest flybys from each simulation.
+Those that cause gaps are colored while those that do not are left in black.
+Note that the meaning of the colors is the same as in Fig.~\ref{fig:histogram_impact_time}.}
+      \label{fig:deltap}
+    \end{figure}
 
@@
       \includegraphics[width=0.45\linewidth]{mass_size_plane.png}
-
-      \includegraphics[width=0.45\linewidth]{impact_geometry_statistics_deltaP.png}
       \includegraphics[width=0.45\linewidth]{impact_stats_phase_space.png}
-      \caption{\textbf{Characterisitcs of gap causing clusters}--\textit{Note} that the colors of the markers and histogram bars are consistent with Fig.~\ref{fig:histogram_impact_time}.\\ \textbf{Top Left:} Energy-angular momentum space of the globular clusters in the simulations, with 50~$\times$~165 data points representing all sampled initial conditions.
-Clusters impacting Palomar 5 are shown with colored markers; non-gap-causing clusters are gray.
+      \caption{\textit{Characterisitcs of gap causing clusters} \textit{Top Left:} Energy-angular momentum space of the globular clusters in the simulations, with 50~$\times$~165 data points representing all sampled initial conditions.
+Clusters impacting Palomar 5 are shown with colored markers; non-gap-causing clusters are shown in gray.
 Palomar 5 is marked with a white symbol and black edge.
-The circular velocity-curve is drawn with a light-pink dashed curve.\\ \textbf{Top Right:} The mass-size plane.
-Uncertainties on the mass are indicated with vertical bars.
-\\ \textbf{Bottom Left:} The distribution of imparted momentum from a cluster flyby given by Eq.~\ref{eq:change_in_momentum}.
-The data set includes the top 5 strongest flybys from each simulation.
-Those that cause gaps are colored while those that do not are left in black.
-\\ \textbf{Bottom Right:} Palomar 5's orbit in the meridional plane.
+The circular velocity-curve is drawn with a light-pink dashed curve.
+\textit{Top Right:} Mass-size plane of the globular clusters in the simulations, with uncertainties on the masses indicated with vertical bars.
+gap-causing clusters are indicated with colored points, whereas Pal~5 is indicated by a white symbol with a black edge.
+\textit{Bottom Right:} Palomar 5's orbit in the meridional plane.
 The grayscale represents all 50 stacked orbits, with $\mathcal{P}$ indicating the probability of  Palomar~5's position, normalized to $\mathcal{P}_\textrm{max}=1$.
-Markers show the perturber's position during each impact.}
+Markers show the perturber's position during each impact.
+In all panels the colors of the markers and histogram bars are consistent with Fig.~\ref{fig:histogram_impact_time}.\paola{can you add also a panel with Pal~5 orbit and impacts in the (x-y) plane?}}
       \label{fig:mass_size_plane}
@@
 
-    The top-left panel shows that all clusters impacting the stream fall within a specific range of orbital energies, between $-2$ and $-1 \times10^5~\textrm{km}^2\textrm{s}^{-2}$.
-This is expected, as only clusters within Palomar 5's orbital space can interact with the stream.
-Our results are consistent with the findings from \citet{2021MNRAS.505.5978V}, who identified a wide range of orbital solutions for the least bound clusters, including both retrograde and prograde orbits.
+  Once all the key impact parameters estimated, we can use Eq~\ref{eq:change_in_momentum} and calculate $\Delta p$, the change in momentum (per unit mass) imparted by a cluster flyby on Pal~5 stream.
+The result  is shown in Fig.~\ref{fig:deltap}, where the distribution of imparted in change in momentum is shown for all impacts which produce a gap, and compared to those that do not produce one.
+On average, encounters which lead to gap creations impart a change in momentum on stream particles which is a factor of 10 higher than that of encounters which do not form gaps (but with some overlap in the low-velocity tail).
+Interestingly, changes in momentum which lead to gap creations extend over a large range in velocities (there is a factor of about 100 between the smallest and largest changes), with NGC~2808 (orange color in the histogram) imparting changes in the velocity of stream particles which redistribute over the whole range of $\Delta p$.
+\paola{I suggest to use $\Delta p$ instead of $\Delta v$ in the xlabel of Fig.~\ref{fig:deltap}, to avoid confusion with $\delta v$ of Fig.\ref{fig:impact_geometry_statistics}.}\\
+
+  In addition to characterizing the parameters governing cluster encounters with the stream, since we know which clusters have produced gaps on the tail of Pal5, we can also verify their orbital and structural properties.
+This is what is shown in Fig~\ref{fig:mass_size_plane}, where we first show the mass and size (i.e.
+half-mass radius) of the clusters that cause gaps on the tails of Pal5, dividing them from those that do not.
+As can be seen, no cluster with mass below $10^5 \rm{M}_\odot$ causes gaps on Pal~5 stream, and all perturbers, except xx \paola{add which cluster has a $r_h > 10$pc} have a  half-mass radius below 10~pc.
+Even more interesting is their distribution in the E-Lz plane, which shows that the clusters that cause gaps are on both direct and retrograde orbits (respectively, negative and positive values of $L_z$), but all confined to an energy interval between $-2$ and $-1 \times10^5~\textrm{km}^2\textrm{s}^{-2}$.
+This is because only clusters within Palomar 5's orbital space can interact with the stream: clusters with higher orbital energies tend to have larger apocenter than that of Pal~5 and thus spend most of their time away from Palomar 5's orbital volume.
+Finally,  it is worth commenting on the location of impacts with the stream, whether they occur when Pal~5 is close to its pericenter or not.
+As it can be seen from Fig.xx (bottom panel) encounters can occur at all orbital phases of Pal~5, when it is close to pericenter, but also very far away from it, at the outskirts of its orbital space.
+However, when taken all together, the location of gap creating impacts shows a strong negative correlation  with the galactocentric radius $r$, with the number $N$ of encounters favorable to gap formation  going as $N = -2.5r + 50$ (with a Pearson coefficient of -0.86).
+While there are more clusters near the galactic center, orbits naturally spend more time near their apocenters, and the lower relative velocities in these regions should favor gap creation.
+However, this result suggests that cluster galactic number density outweighs these factors in determining the number of gaps.
+
+  %Our results are consistent with the findings from \citet{2021MNRAS.505.5978V}, who identified a wide range of orbital solutions for the least bound clusters, including both retrograde and prograde orbits.
 In our potential model, only three clusters samplings had positive orbital energies.
 
-    The mass-size relation, shown in the top-right panel, reveals a lower limit of $10^5~\textrm{M}_{\odot}$ for clusters that provoke gaps.
-While mass and size alone are insufficient predictors, this empirical limit is useful because relative velocity and impact parameters are inacessible.
-Concentration or the density profile of the perturber has a much less significant effect respect to the total mass and subsequent evolution of the gap as shown in Fig.~3 of \citet{2016MNRAS.457.3817S}.
-The bottom-left panel offers a more physically meaningful result: gap-creating impacts tend to impart more momentum into the stream than non-gap impacts, though with some overlap.
-While \citet{2015MNRAS.450.1136E} extended the three-parameter model (Eq.~\ref{eq:change_in_momentum}) to include the size of the perturber and the perpendicular and parallel components of relative velocity, we found no clear combination of these factors that distinguishes gap from non-gap cases.
+\section{Discussion}
 
-    Further insights are shown in Fig.\ref{fig:impact_geometry_statistics}, highlighting a maximum impact parameter of around 200~pc, roughly the width of the stream.
-Additionally, as \citet{2016MNRAS.457.3817S} demonstrated, the dispersion of action-frequencies in the stream plays a role.
-The frequencies being the rate at which a particle completes a cycle in phase space.
+  The simulations presented in this paper suggest that, in the last 5 Gyr of evolution, Palomar~5 should have experienced close encounters with other Galactic globular clusters, some of which are able to create gaps -- even a few kpc wide -- in its tails.
+Unfortunately, none of the predicted gaps is located in regions of the tails that have been observed, and this is especially true for the leading tail, which result shorter than the trailing tail in the observations \citep[see][for a possible explanation of this feature]{2017NatAs...1..633P}.
+Currently, the presence -- or not -- of gaps in the observed portion of Pal~5 tails is debated: \citet{2016ApJ...819....1I} found no statistically significant gap in Pal~5 tails, while \citet{2017MNRAS.470...60E}, analysing the same dataset as \citet{2016ApJ...819....1I}, suggested the presence of a few dips and gaps in the tails, at angular distances between $2^\circ$ and $9^\circ$ from the cluster center \citep[see also][]{ 2020ApJ...889...70B}.
+We cannot reproduce any of these features with clusters flybys in our simulations, since all our gaps are found at much larger angular and spatial distance from the cluster itself.
+There may be a few reasons for the absence of gaps in portion of the tails closer to the cluster center.
+First, as \citet{2016MNRAS.457.3817S} demonstrated, the dispersion of action-frequencies in the stream plays a role.
 For instance, the frequency corresponding to the azimuthal action, $J_\phi$, is $\dot\theta_\phi = - \frac{\mathcal{\partial H}}{\partial J_\phi}$ and where $\theta_\phi$ is the angle describing the position of the particle phase space between momentum ($p_\phi$) and position ($\phi$) axes---where $\phi$ is the azimuhtal angle between the x-y axes in physical space.
@@
 Thus, for a gap to form, the imparted change in frequency must exceed the range of frequencies in the impacted region.
-Thus, the strong flybys that occured close to the cluster and were thus inconsequential for gap formation.
+Thus, the strong flybys that occurred close to the cluster were inconsequential for gap formation.
+Another possible explanation is the different thickness of the simulated tails at different distances from the center of the cluster.
+At distances of less than about 3~kpc from Pal~5, the tails are very thin, with a typical thickness of \paola{xx} pc.
+In order to cause gaps in these regions, the clusters would have to pass close to the stream, with impact parameters similar to the thickness itself.
+Note that the thinness of the tails at these distances is probably a direct consequence of the initial parameters we chose for the simulation.
+We have assumed for Pal5, 5 Gyr years ago, the same internal parameters (mass and size) that the cluster has today.
+With such parameters, after 5 Gyr of evolution, our N-body system has lost most of its mass, and therefore the part of the tails closest to the cluster itself, whose density depends mainly on the most recent mass loss \citep[see, for example, Fig.~A.3 in][]{2012A&A...546L...7M}, is necessarily very thin because the simulated cluster has essentially no more mass to lose.
+This last point has also a consequence in the gap creation rate, which we derived at the end of Sect.~\ref{sect:geometry}: with only 70\% of the tail  ($\sim$ 14~kpc over 20~kpc, excluding the innermost $\pm 3$~kpc from the cluster center) really suitable for forming gaps, the gap creation rates are about 50$\%$ higher than those estimated in the aforementioned section, where values have been derived taking into account the full tail extent.
+This gap creation rate would be still not high enough to reproduce the  number of gaps in Palomar~5 tails, as reported by \citet{2012ApJ...760...75C}.
+In their study, they indeed suggested the presence of  five gaps with a 99~\% detection confidence, leading to a gap creation rate of 0.17~Gyr$^{-1}$~kpc$^{-1}$.
+If true, this rate would be too high to be explained by Galactic globular clusters only.
 
-    The bottom-right panel reveals a strong negative correlation between the location of impacts and the galactocentric radius, $r$.
-The number of gap creating impacts follows the relation $N = -2.5r + 50$, with a Pearson coefficient of -0.86.
-This is a non-obvious result.
-While there are more clusters near the galactic center, orbits naturally spend more time near their apocenters, and the lower relative velocities in these regions should favor gap creation.
-However, this result suggests that cluster galactic number density outweighs these factors in determining the number of gaps.
+  We note that our results seem to be in tension with the conclusions of \citet{2019MNRAS.484.2009B}, who presented a numerical study of Palomar~5 tails, orbiting a Milky Way-like potential, where both dark matter sub-halos and baryonic sub-structures (Galactic bar, spiral arms, giant molecular clouds, globular clusters) were taken into account to quantify the importance of these latter in density variations in Pal~5 streams.
+From their study, they derived that the impact of globular clusters was marginal in gap creation.
+It is difficult to give an exact explanation of the reasons for this difference, because the approach followed by  \citet{2019MNRAS.484.2009B} is extremely similar to ours.
+We would point out, however, that the Galactic potential in which the integration of the orbit of Pal5 and all other galactic clusters was made in their study is the \texttt{MWPotential2014} \citep{2015ApJS..216...29B}.
+As discussed by \citet{2022ApJ...926..107M}, this potential has a halo of dark matter that is probably too light, because 20\% of the Galactic globular clusters turn out to have unbound orbits in it (note that this is not the case for the potential we have adopted, see e.g.
+Fig.~\ref{fig:mass_size_plane}).
+Even for clusters that turn out to be bound in the  \texttt{MWPotential2014} potential, the fact that the dark matter halo is (possibly) too light implies that the orbital volume occupied by the clusters is larger, thus reducing the probability of close impacts with the Pal~5 stream.
+We tentatively propose this as an explanation of the differences found, but a more systematic comparison between \citet{2019MNRAS.484.2009B}'s results and ours should be made to understand the origin of this difference.
+\\
 
-    Lastly, \citet{2012ApJ...748...20C} conceptualized the number of gaps using a statistical quantity: the rate at which gaps are created per unit stream length.
-For our simulations, this is given by: \begin{equation} \label{eq:gap_creation_rate}
-    \mathcal{R}_{\textrm{Pal 5}} =  \frac{1}{T}\int_{0}^T l^{-1}(t) \sum_i \delta(t-t_i) dt,\end{equation}where $T$ is the total integration time, $l(t)$ is the length of the stream, and $\sum_i \delta(t-t_i)$ sums over the gap occurrences, with $i$ indexing over the number of gaps in a given simulation.
-Here, $\delta$ represents the diract delta function.
-This expression can be simplified to:\begin{equation}\mathcal{R}_{\textrm{Pal 5}} =  \frac{1}{T} \sum_i \frac{1}{l (t_i)}.
-\end{equation}
-    This computation allows us to analyze the distribution of gap creation rates across all simulations.
-The results are shown in Fig.~\ref{fig:TailCoordinates}.
-This is roughly consistent with a simple estimate of the average gap creation rate: with 72 gaps over 5 Gyr of integration time for a stream about 20 kpc in length, the naive estimate is approximately 0.014~[km~s$^{-1}$~kpc$^{-2}$].
-This is about double the weighted mean gap creation rate of 0.009~[km~s$^{-1}$~kpc$^{-2}$].
-The naive estimate is higher because it does not account for the growth of the stream over time, unlike Eq.\ref{eq:gap_creation_rate}, where the likelihood of impacting a shorter stream is lower.
+  The impact of globular clusters close interactions with stellar streams has been also the object of another recent paper, by \citet{2022ApJ...941..129D}, who concluded that the chance for GD-1 gaps, which have been reported in a number of works \citep[see, for example, ][]{} to be produced by globular clusters is very low.
+This result is not necessarily in contradiction with ours, since GD-1 has a pericenter which is almost twice that of Palomar~5 \citep[see, for example][]{2019MNRAS.486.2995M}.
+As we have discussed in Sect.~\ref{sect:geometry}, the number, $N$, of close encounters which lead to gap creation is anti-correlated with the distance $r$ to the Galactic center, which implies that even simply applying the same radial dependence of $N$ as that found for Pal~5, but simply changing the pericenter value from $6$~kpc (roughly the value of Pal~5 pericenter \paola{can you check this value in your models?}), to 14~kpc (the closest distance of GD-1 to the Galactic center) reduced the number of gap-favorable impacts of more than a factor of 2.
+We, however, note that in their work \citet{2022ApJ...941..129D} preselect the  globular clusters which could have experienced a close encounter with the GD-1 stream, by selecting only clusters which pass at a distance less than 0.5 kpc from the stream, having a relative velocity smaller than  300 km/s.
+As shown in our Fig.~\ref{fig:impact_geometry_statistics}, bottom-left panel, in the case of Palomar~5 this choice would lead to exclude most of the encounters favorable to gap creation, which turn out to have relative velocities above 300 km/s.
+It would be interesting to repeat a similar study as the one made by  \citet{2022ApJ...941..129D} for GD-1, this time imposing no selection on possible candidate clusters.\\
 
-      \begin{figure}
-        \centering
-        \includegraphics[width=\linewidth]{gap_creation_rate.png}
-        \caption{The distribution of the number of gaps normalized over the total integration time and unit stream length, as described by Eq.\ref{eq:gap_creation_rate}. }
-        \label{fig:TailCoordinates}
-      \end{figure}
+  As already suggested in previous works which have studied the impact of baryonic structures on Palomar~5 tails \citep{2017NatAs...1..633P, 2019MNRAS.484.2009B}, it is possible that this cluster lies in a region of the phase-space which is not favorable to distinguish gaps created by dark matter subhalos from gaps created by baryonic structures, as the Galactic bar and giant molecular clouds.
+With our study, we show that close encounters with globular clusters constitute  a further element that confuses a simple interpretation of the observed Palomar~5 stream gaps (if any).
+Other clusters and streams in the orbital energy range of Palomar~5 may suffer of the same difficulty.
+Going to lower orbital energies make the situation even worst, because the effect of the Galactic bar, giant molecular clouds and interactions with globular clusters becomes even more efficient.
+In the innermost regions of the Galaxy, where the density of dark matter sub-halos is expected to be maximal, globular cluster streams are intrinsically more difficult to be found, in addition to the fact that the dynamic times in these regions become so small that the stars lost from globular clusters do not redistribute themselves for the most part into thin structures \citep[see][]{2023A&A...673A..44F}.
+It is thus probably only at larger distances from the Galactic center than those spanned by Palomar~5 orbit that the impact of dark matter sub-halos may become dominant, but it should be borne in mind that at these distances the number density of sub-halos also decreases.
+In the future, it will be interesting to apply the type of study conducted here to the whole set of clusters for which streams have been found to quantify the regime in which they are found, whether favourable to the creation of gaps from baryonic structures or not.
 
-\section{Discussion}
-
-  \salvatore{Model limits? lets discuss them here.}
-
-  \salvatore{I'm a bit lost on trying to reconcile the differences between Baumgardt's estiamte and \citet{2017ApJ...842..120I}'s estimate in sections 5.2 and 5.3.
-It looks like 5.2 is based on counts and then supplementing the completness function with dNdM \citet{2001AJ....122.3231G} and demonstrating that Salpeter would overestimate the mass.
-I think 5.3 is showing that an MCMC was used to add more context to the estimate from 5.3.}  Regarding Palomar 5's mass, we use a mass estimate for Palomar 5 taken from \citet{2021MNRAS.505.5957B}, $1.34\pm0.24\times10^{4}M_{\odot}$--which is a contested value.
-
-  \salvatore{\textbf{If Salpeter over estimates the number of low mass stars}, then wouldn't this appear in the velocity dispersion profile?
-or wait, is \citet{2017ApJ...842..120I} fitting the density for the King, supplementing missing stars by a Mass function from \citet{2001AJ....122.3231G}, and finding the total mass this way?
-i.e.
-no velocity dispersion was investigated?
-I feel like viewing the velocity dispersion directly would provide a good estiamte for the dynamical mass, no?}
-
-  \salvatore{I am lost in Baumgardt's papers trying to flush out how a MF was used to provide their mass estimate. It looks like this information is within Sollima 2017. Need to revisit this. }
-
-  \salvatore{What I say will depend.
-I think using \citet{2021MNRAS.505.5957B}'s mass estimate is still ok because it is similar in order of magnitude to cluster + stream from \citet{2017ApJ...842..120I}.
-}
-
-  \salvatore{What's the point?
-In the end were more interested in the streams for probing the galaxy.
-Having a more accurate model of the internal dynamics Palomar 5 would change the luminosity per square degree.
-}
-
-  In this work, we quantified the number of gaps expected in Palomar 5's stream from the 165 Galactic globular clusters using a Monte Carlo-style experiment.
-We performed 50 iterations, varying the orbits based on the uncertainties in their observed proper motions, radial velocities, and distances.
-Our findings are as follows:
-
-  \begin{itemize}
-    \item \textbf{Orbital Overlap}: clusters with orbital energies and angular momenta that overlap with Palomar 5's orbit can impact its stream.
-    \item \textbf{Mass Threshold}: Perturbing clusters have masses above $10^5 M_\odot$; no perturber has a mass smaller than this.
-    \item \textbf{Galactic radius}: Impact occurrences negatively correlates with galacto-centric distance.
-    \item \textbf{Momentum and Frequency Change}: Gap creation is linked to the momentum imparted to the stream during the flyby, yet not entirely.
-As showcased by \citet{2016MNRAS.457.3817S}, the change in orbital frequencies induced by the flyby must be greater than the spread in frequencies present in the stream.
-Some impacts imparted significant momentum changes that did not lead to gap growth for this reason.
-  \end{itemize}
-
-  Finally, we present a gap creation rate as conceptualized by \citet{2012ApJ...748...20C}, which is the number of gaps normalized to the age and length of the stream.
-This rate is: 0.009~[km~s$^{-1}$~kpc$^{-2}$], or 0.01~[Gyr$^{-1}$~kpc$^{-1}$].
-
-  Now, \textit{is this consistent with reality?} Can globular clusters explain the observed gaps in Palomar 5?
-Currently, the results on the number of gaps in the stream are debated.
-\citet{2012ApJ...760...75C} and \citet{2020ApJ...889...70B} attempted to quantify the stream's density along its width, isolating the stream from background and foreground contamination.
-\citet{2020ApJ...889...70B} reported the existence of two potential gaps, whereas \citet{2012ApJ...760...75C} identified five gaps with a 99~\% detection confidence, leading to a gap creation rate of 0.17~Gyr$^{-1}$~kpc$^{-1}$.
-If true, this rate is too high to be explained by Galactic globular clusters.
-
-  On the other hand, to provide a crude estiamte, consider  a stream length of 10~kpc reported from \citet{2020ApJ...889...70B} and given our gap creation rate.
-In 5~Gyr we would expect to see about 0.5 gaps, in 10~Gyr's we would expect 1.
-Therefore, finding two gaps from globular clusters would not be impossible.
-
-  In the future, I would like to investigate stream detections in a full mock observation experiment and build a confusion matrix for the number of true/false positivies/negatives for the stream gaps from globular cluster and dark matter subhaloes.
-Like this, we will be able to understand what data quality is necessary to make such distinctions and what signs in the streams from substructure will be recoverable.
-This could support the https://via-project.org/ mission by Bonaca and Conroy.
-Additionally, I would like to expand this work to include all gap creation rates for all observed stellar streams.
+\section{Conclusions}
+  \paola{to be written}
 

### 2024-11-19 paper/main.tex
ingested appendix changes

@@
 
-      With the perturbers identified, we perform statistics in the pursuit of understanding what conditions are necessary for a globular cluster and a stream to have in order for a gap to form.
-We turn to impact theory, which in its simplest form is presented in works such as \citet{2008gady.book.....B}.Consider two particles: one stationary and the other moving past it.
-The distance between the two particles at their closest approach is known as the impact parameter.
-To simplify the analysis, the \textit{impulse approximation} is employed, which assumes that the velocity of the perturber remains unchanged during the interaction.
-This assumption simplifies the computation.
-
-      To understand how the impacted particle is perturbed, one needs to compute its change in momentum, which is determined by integrating the force acting on the particle over the duration of the interaction.
-A useful approximation for this change in momentum is the force at the closest approach multiplied by an estimate of the interaction time:
-      \begin{equation} 
-        \Delta p \approx \text{Force} \times \text{interaction time} = \frac{GMm}{b^2} \times 2\frac{b}{v} = 2\frac{GM}{bv}, 
-        \end{equation} 
-      where $M$ is the mass of the perturber, $m$ is the mass of the particle, $b$ is the impact parameter, $v$ is the relative velocity of the perturber with respect to the particle, and $G$ is the gravitational constant.
-
-      From this analysis, we can conclude that a more massive perturber, passing closer to the particle, and moving more slowly, will have a greater impact.
-It's important to note that the momentum change is inversely proportional to the velocity of the perturber, which contrasts with the intuition from elastic collisions, such as those between billiard balls, where higher velocities result in greater impacts.
-
-      \citet{2015MNRAS.450.1136E} extended this impact theory from one point mass impacting another point mass, to studying how an extended body impacts as stream.
-This quantifies the change in momentum of a given particle as a function of its distance from the point of greatest impact.
-In their analysis, the perturber was modeled as a Plummer sphere, like in our simulations.
-Since the stream is not zero-dimensional and has length, both the parallel and perpendicular components of the pertuber's velocity must be considered.
-Consequently, five parameters determine the change in velocity of a given stream particle: $M$, $r_p$, $b$, $W_\parallel$, and $W_\perp$.
-We set out is to constrain these parameters during the impact in our simulations and compare them to close approaches that do not generate gaps, in order to establish the criteria necessary for gap formation.
-
+    As discussed in Sect~\ref{sect:geometry},  five parameters determine the change in velocity of a given stream particle: $M$, $r_p$, $b$, $W_\parallel$, and $W_\perp$.
+In the following, we describe how we estimated these parameters during impacts in our simulations.
     To achieve this, we begin by identifying the approximate moments of impact from the most significant clusters, as determined in the previous analysis in Sec.~\ref{sec:Perturber_Identification}.
@@
 Thus, we compute a sample of 250 impacts, and flag those that give way to gaps.
-
-      Visual inspection of the five key impact parameters ($M$, $r_p$, $b$, $W_\parallel$, and $W_\perp$) did not reveal a clear distinction between flybys that create gaps and those that do not.
-Due to the clutter, we omit this graph and instead present Fig.\ref{fig:impact_geometry_statistics}, which uses the total relative velocity rather than separating parallel and perpendicular components.
-We also exclude the characteristic radius, as it showed little correlation with the results, likely due to the narrow range of globular cluster radii (see Fig.\ref{fig:mass_size_plane}).
-This factor might be more significant for dark matter subhalos, where size variation is greater.
-
-      \begin{figure*}
-        \centering
-        \includegraphics[width=\linewidth]{impact_geometry_statistics.png}
-        \caption{The distribution and relationship between the basic impact variables from Eq.\ref{eq:change_in_momentum} and the close encounter that cause impacts, shown in red against those that do not, shown in blue.
-}
-        \label{fig:impact_geometry_statistics}    
-        \end{figure*}
-
-      While Fig.\ref{fig:impact_geometry_statistics} demonstrates that mass, relative velocity, or impact parameter alone (from Eq.\ref{eq:change_in_momentum}) cannot predict gap formation, one interesting result emerges: no gaps are created from interactions with impact parameters greater than 200~pc.
-The stream widths are roughly 200~pc, with a radial thickness of about 100~pc.
-This finding is even more evident when examining the $b$-$M$ plane.
-A series of perturbers at roughly $\sim7 \times 10^5 M_\odot$ highlights NGC~2808's flybys, where all encounters with impact parameters under 200~pc result in gaps, while those beyond this distance do not.
+We refer the reader to Sect.~\ref{sect:geometry} for the presentation and discussion of the results.
 

### 2024-11-19 paper/main.tex
editted overview

@@
 
-    To make a more quantitative comparison between the \texttt{reference} and \texttt{full} simulations, we adopted to work in the tail coordinate system, in which  the stream is aligned with the cluster's orbit.
+    To make a more quantitative comparison between the \texttt{reference} and \texttt{full} simulations, we work in the tail coordinate system, in which  the stream is aligned with the cluster's orbit.
 The definition of this coordinate system is based on the work of \citet{2004AJ....127.2753D} and is shown in Fig.~\ref{fig:TailCoordinates}.
-Briefly, in this system the x' coordinate represents the position of a particle along the orbit relative to the globular cluster—whether it is ahead (positive x' ) or behind (negative x') the cluster;  the y' coordinate measures the particle's distance within the orbital plane, where positive values indicate that the particle is farther from the Galactic center and negative values indicate that it is closer.
+Briefly, in this system the $x_{\textrm{tail}}$ coordinate represents the position of a particle along the orbit relative to the globular cluster—whether it is ahead (positive $x_{\textrm{tail}}$) or behind (negative $x_{\textrm{tail}}$) the cluster;  the $y_{\textrm{tail}}$ coordinate measures the particle's distance within the orbital plane, where positive values indicate that the particle is farther from the Galactic center and negative values indicate that it is closer.
 Fig.~\ref{fig:decomposition} shows a comparison between one of the 50 realizations of Palomar~5 stream taking into account the gravitational interactions with all other globular clusters in the Galaxy (top panel) and omitting them (bottom panel).
-This comparison clearly shows the presence of two wide ($\sim$ 100~pc and $\sim$ 1~kpc) gaps in the leading tail and of a more subtle underdensity at x'$\sim 5$~kpc (we refer the reader to Appendix~\ref{sec:gap_detection} for a detailed description of the underdensities and gaps detection method).
-To find out which globular clusters were responsible of creating these gaps, and at what time the close passages between these clusters and the Palomar~5 streams took place we estimated the gravitational acceleration along the orbit of Pal~5 and represented it in the ($t$, $\tau$) space, where $t$ is the simulation time, from 5~Gyr in the past to the present day, and $\tau$ is indicates how long it will take for Pal~5 to reach, or how  long ago it passed, a given point on its orbit.
-The use of $\tau$ is advantageous because the growth of the stream is approximately linear in $\tau$ whereas in physical space, streams on eccentric orbits expand and contract depending on the orbital phase.
-Adopting this time space, and reporting the gravitational acceleration along Pal~5 orbit in this space, the identification of the globular clusters which produced the perturbation, and the time at which it occurred becomes straightforward (we refer the reader to Appendix~\ref{sec:Perturber_Identification} for a in-depth description of the procedure).
-In this way, we could identify that the clusters responsible for creating gaps in the simulation of Palomar~5 stream reported in Fig.\ref{fig:decomposition} are NGC~2808, NGC~7078 and NGC~104, and that their close passages occurred 200~Myr, \paola{xx and xx Gyr ago}, respectively.
+This comparison clearly shows the presence of two wide ($\sim$ 100~pc and $\sim$ 1~kpc) gaps in the leading tail and of a more subtle underdensity at $x_{\textrm{tail}}\sim 5$~kpc (we refer the reader to Appendix~\ref{sec:gap_detection} for a detailed description of the underdensities and gaps detection method).
+To find out which globular clusters were responsible of creating these gaps, and at what time the close passages took place, we estimated the gravitational acceleration along the orbit of Pal~5 and represented it in the ($t$, $\tau$) space.
+$t$ is the simulation time and $\tau$ is indicates how long it will take for Pal~5 to reach, or how  long ago it passed, a given point on its orbit.
+The use of $\tau$ is advantageous because the growth of the stream is approximately linear in $\tau$.
+On the other hand, in physical space streams on eccentric orbits indeed grow in time yet the growth rate is modulated with periodic expansion and contraction depending on the orbital phase.
+Adopting this time space, and reporting the gravitational acceleration along Pal~5 orbit in this space, the identification of the globular clusters which produced the perturbation, and the time at which it occurred becomes straightforwar.
+We refer the reader to Appendix~\ref{sec:Perturber_Identification} for an in-depth description of the procedure.
+In this way, we could identify that the clusters responsible for creating gaps in the simulation of Palomar~5 stream reported in Fig.\ref{fig:decomposition} are NGC~2808, NGC~7078 and NGC~104, and that their close passages occurred 200~Myr, -1.9 and -2.1~Gyr ago, respectively.
 To further probe that the three clusters above are responsible of producing the gaps observed in this simulation, we conducted additional experiments by including only the perturbation of each of these three clusters at a time, and neglecting the gravitational perturbations of all the other clusters in the Galaxy.
@@
 It is worth noting that the times at which the passages of these clusters occurred, according to our analysis, are in agreement with the observed width of the corresponding gaps: the encounter with NGC~2808 being very recent (only 200~Myr ago) its induced gap is still very thin, because it takes time for a perturbation to grow into an extended gap, as it is the case for those induced by the passages of  NGC~7078 and NGC~104, which occurred in earlier times.
-It is also interesting to emphasize that gaps as thin as those generated by the passage of NGC~2808, 200~Myr ago, can be detected by working in the tail coordinate system and by making a differential analysis (\texttt{full} versus \texttt{reference} simulations): they are so thin that they cannot be directly identified in galactic coordinates (see Fig.~\ref{fig:stream_on_sky}).\\
+It is also interesting to emphasize that gaps as thin as those generated by the passage of NGC~2808, 200~Myr ago, can be detected by working in the tail coordinate system and by making a comparative analysis (\texttt{full} versus \texttt{reference} simulations): they are so thin that they cannot be directly identified in galactic coordinates (see Fig.~\ref{fig:stream_on_sky}).\\
 

### 2024-11-19 paper/main.tex
editted history and statistics

@@
 
-    By applying the methodology described above to the whole set of simulations, we can reconstruct the history of close passages of Galactic globular clusters to Palomar~5 stream, in the last 5~Gyr, which -- we recall the reader -- is the time interval investigated in our simulations.
-The results of this analysis are reported in Fig.~\ref{fig:histogram_impact_time} and in Table~xx \paola{I suggest to report the results of Fig 3 also in a Table, in which we report for each GC which produced a gap, the time of close passage and the total number of gaps it generated.} In total, we report the finding of 73 gaps in our 50 simulations, that is 1.5 gaps/simulation, on average, created by 18 different perturbers.
+    By applying the methodology described above to the whole set of simulations, we can reconstruct the history of close passages of Galactic globular clusters to Palomar~5's stream, in the last 5~Gyr, which -- we recall the reader -- is the time interval investigated in our simulations.
+The results of this analysis are reported in Fig.~\ref{fig:histogram_impact_time} and in Table~xx \paola{I suggest to report the results of Fig 3 also in a Table, in which we report for each GC which produced a gap, the time of close passage and the total number of gaps it generated.}.
 NGC~2808 impacts Palomar 5's stream about 200 Myr ago in 44 out of 50 simulations, creating a small gap at a similar position to the one reported in Fig.~\ref{fig:decomposition} in each case.
-Since this interaction occurred less than one orbital period ago, the orbital solutions remain similar, producing consistent results across all simulations, within the uncertainties.
-For encounters which took place at earlier times, the clusters responsible of creating the gaps may vary -- that is, gaps are not created at the same time in all simulations and the parameters of a specific encounter between a  globular cluster and Palomar~5 stream may vary significantly from a simulation to another with the turning back of time, so that in some cases a cluster may produce a gap and in some others the perturbation is not strong or close enough to produce any detectable signature in the Palomar~5 stream.
-In total, beside NGC~2808, we report the creation of 24 gaps in Pal~5 tails, over the last 3~Gyr generated by \paola{15: check this value, I derived it by eye} different clusters, which corresponds to a rate of 0.8 gaps/100~Myr (note that, including NGC~2808, this rate increases to 29.5 gaps/100~Myr).
-At even earlier times, the probability of creating gaps in Palomar~5 streams diminishes significantly, but one must take into account that at these times the tails of Pal~5 have not yet had time to develop sufficiently, and the probability of close encounters with other clusters consequently decreases.
-Overall, across 50 simulations, as previously stated, we observe 73 gaps.
+Since this interaction occurred less than one orbital period ago, the uncertainties lead to orbital solutions remain that similar and thus producing consistent results across all simulations.
+However, as we continue to turn back time further, the uncertainties in the initial conditions allow the for the various orbital solutions to diverge from one another.
+Thus, in one configuration a cluster can impact the stream at a given time and yet at the same moment in a different set of initial conditions, it could be on the other side of the galaxy.
+
+    In total, we report the finding of 73 gaps in our 50 simulations which is on average 1.5 gaps/simulation and were created by 18 different perturbers.
 The distribution of simulations with a given number of gaps is provided in Table~\ref{table:gap_distribution}.
@@
 Including all observed gaps results in an average of 1.5 gaps per simulation.
-Interestingly, most gaps appear in the leading tail, which we consider coincidental \paola{I am still surprised by the high occurrence of gaps in the leading tail, but I amm still not able to come up with an explanation for this.
-I mean, is it really coincidental?
-I propose to leave the discussion as it is for the moment.
-We will add more, if we come up with an other possible explanation}, as stars escape the cluster's L1 and L2 points at similar rates, and the small distance difference between tails' galactocentric distance is negligible.
-We expect gaps to be equally likely in both tails.
-However, only 8 of the 73 gaps occur in the trailing tail, with 44 linked to NGC~2808's flyby.
-After removing these, there are 21 gaps in the leading tail and 8 in the trailing.
-By the binomial distribution, the probability of observing up to 8 successes in 29 trials, given a 50\% chance of success, is 1.2\%—unlikely, but possible.
-Additionally, other perturbers impact at consistent times, which may also violate the assumption of independent events, as seen with NGC~2808.
 
@@
 
-    Following  \citet{2012ApJ...748...20C} we can also estimate the rate at which gaps are created per unit stream length.
+    We need a more sophisticated satistic that can be compared to other simulations and turn to the gap creation rate developed by \citet{2012ApJ...748...20C}.
+This is the number of gaps per time and normalized by the length of the stream.
 For our simulations, this rate is given by: \begin{equation} \label{eq:gap_creation_rate} \mathcal{R}_{\textrm{Pal 5}} =  \frac{1}{T}\int_{0}^T l^{-1}(t) \sum_i \delta(t-t_i) dt,\end{equation}where $T$ is the total integration time, $l(t)$ is the length of the stream, and $\sum_i \delta(t-t_i)$ sums over the gap occurrences, with $i$ indexing over the number of gaps in a given simulation.
@@
     This computation allows us to analyze the distribution of gap creation rates across all simulations.
+Notice that since the gaps \textit{add in parallel}, naturally gaps that occur at earlier times when the stream was shorter are weighted higher than those that occur when the stream is longer.
 The results are shown in Fig.~\ref{fig:gapcreationrate}.
-This is roughly consistent with a simple estimate of the average gap creation rate: with 72\paola{73?} gaps over 5 Gyr of integration time for a stream about 20 kpc in length, the naive estimate is approximately 0.015~km~s$^{-1}$~kpc$^{-2} = 0.015~\rm{Gyr^{-1}kpc^{-1}}$ .
+This is roughly consistent with a simple estimate of the average gap creation rate: with 73 gaps over 5 Gyr of integration time for a stream about 20 kpc in length, the naive estimate is approximately 0.015~km~s$^{-1}$~kpc$^{-2} = 0.015~\rm{Gyr^{-1}kpc^{-1}}$.
 This is about double the weighted mean gap creation rate of 0.009~km~s$^{-1}$~kpc$^{-2}$.
-The naive estimate is higher because it does not account for the growth of the stream over time, unlike Eq.\ref{eq:gap_creation_rate}, where the likelihood of impacting a shorter stream is lower.
+The naive estimate is higher because it does not account for the growth of the stream over time, unlike Eq.\ref{eq:gap_creation_rate}.
 
@@
     \end{figure}
+
+    Lastly, we note that of the 73 observed gaps only 8 are in the trailing tail and the rest are in the leading tail.
+This is a surprising result.
+Apriori, since the number of star-particles escape at similar rates from the L1 and L2 Lagrange points, each tail is of similar length and density.
+The main difference between the two tails is that the leading tail is closer to the galactic center than the trailing tail by about 400~pc.
+Since the lengths are equal, and the offset between the tails is small compared to the galactocentric distance of about 10 kpc, we expected the number of gaps in each tail to be more or less the same.
+We can compute the the probability of observing the unequal occurances though the binomail distirbution.
+First, the since the 44 gaps linked to NGC~2808 are the same fly-by, they are not independent events.
+We remove them from this consideration.
+This leaves 21 gaps in the leading tail and 8 in the trailing.
+The probability of observing up to 8 successes in 29 trials, given a 50\% chance of success, is 1.2\%--—unlikely, but possible.
+Additionally, other perturbers impact at consistent times, which may also violate the assumption of independent events, as seen with NGC~2808.
 

### 2024-11-20 paper/main.tex
changes to impact geometry

@@
 A useful approximation for this change in momentum \paola{(per unit mass)} is the force at the closest approach multiplied by an estimate of the interaction time:
-    \begin{equation} \label{eq:change_in_momentum} \Delta p \approx \text{Force} \times \text{interaction time} = \frac{GMm}{b^2} \times 2\frac{b}{v} = 2\frac{GM}{bv}, \end{equation}where $M$ is the mass of the perturber, $m$ is the mass of the particle, $b$ is the impact parameter, $v$ is the relative velocity of the perturber with respect to the particle, and $G$ is the gravitational constant.
-\paola{ATTENTION : in the last formula there is no dependence on $m$, as  for a change in momentum per unit mass (hence my comment few lines before).
-If you want to keep it as momentum per unit mass, that is a change in velocity, you should also remove $m$ from the other expression (i.e.
-GMm)}
+    \begin{equation} \label{eq:change_in_momentum} \Delta p \approx \text{Force} \times \text{interaction time} = \frac{GM}{b^2} \times 2\frac{b}{\delta v} = 2\frac{GM}{b \delta v}, \end{equation}where $M$ is the mass of the perturber, $b$ is the impact parameter, $\delta v$ is the relative velocity of the perturber with respect to the particle, and $G$ is the gravitational constant.
 
@@
     Visual inspection of the five key impact parameters ($M$, $r_p$, $b$, $W_\parallel$, and $W_\perp$) did not reveal a clear distinction between flybys that create gaps and those that do not.
-Due to the clutter, we omit this graph and instead present...
-In Fig.\ref{fig:impact_geometry_statistics} we report the result of this analysis, by showing the distribution of these impact variables for fly-bys which give rise to gaps and compare it to the distributions of fly-bys which do not lead to gap formation.
+Therefore, we only present the the quantities from Eq.~\ref{eq:change_in_momentum} in Fig.~\ref{fig:impact_geometry_statistics}.
     Note that in this figure, we show the total relative velocity rather than separating parallel and perpendicular components, because from a visual inspection, we could not find any specific trends when the two velocity components were plotted separately.
@@
 
-    While Fig.~\ref{fig:impact_geometry_statistics} demonstrates that mass, relative velocity, or impact parameter alone (from Eq.~\ref{eq:change_in_momentum}) cannot predict gap formation, one interesting result emerges: no gaps are created from interactions with impact parameters greater than 200~pc.
-The stream widths are roughly 200~pc, with a radial thickness\paola{what do you mean by radial thickness?} of about 100~pc.
+    While Fig.~\ref{fig:impact_geometry_statistics} demonstrates that mass, relative velocity, or impact parameter alone cannot predict gap formation, one interesting result emerges: no gaps are created from interactions with impact parameters greater than 200~pc.
+The stream widths are roughly 200~pc.
 This finding is even more evident when examining the $b$-$M$ plane.
@@
 It is interesting to note that for values of the impact parameter below this threshold, even fast encounters ($\delta v > 300$~km/s) can produce gaps.
-\paola{Attention: there is something I do not understand in Fig 5: the range of $\delta v$ values seems not the same in the first row and in the last row (left panel).
-Is there anything to clarify/correct?}
 
@@
 Interestingly, changes in momentum which lead to gap creations extend over a large range in velocities (there is a factor of about 100 between the smallest and largest changes), with NGC~2808 (orange color in the histogram) imparting changes in the velocity of stream particles which redistribute over the whole range of $\Delta p$.
-\paola{I suggest to use $\Delta p$ instead of $\Delta v$ in the xlabel of Fig.~\ref{fig:deltap}, to avoid confusion with $\delta v$ of Fig.\ref{fig:impact_geometry_statistics}.}\\
 
-  In addition to characterizing the parameters governing cluster encounters with the stream, since we know which clusters have produced gaps on the tail of Pal5, we can also verify their orbital and structural properties.
+    In addition to characterizing the parameters governing cluster encounters with the stream, since we know which clusters have produced gaps on the tail of Palomar~5, we can also verify their orbital and structural properties.
 This is what is shown in Fig~\ref{fig:mass_size_plane}, where we first show the mass and size (i.e.
 half-mass radius) of the clusters that cause gaps on the tails of Pal5, dividing them from those that do not.
-As can be seen, no cluster with mass below $10^5 \rm{M}_\odot$ causes gaps on Pal~5 stream, and all perturbers, except xx \paola{add which cluster has a $r_h > 10$pc} have a  half-mass radius below 10~pc.
+As can be seen, no cluster with mass below $10^5 \rm{M}_\odot$ causes gaps on Pal~5 stream, and all perturbers, except xx \paola{add which cluster has a $r_h > 10$~pc} have a half-mass radius below 10~pc.
 Even more interesting is their distribution in the E-Lz plane, which shows that the clusters that cause gaps are on both direct and retrograde orbits (respectively, negative and positive values of $L_z$), but all confined to an energy interval between $-2$ and $-1 \times10^5~\textrm{km}^2\textrm{s}^{-2}$.
@@
 Finally, it is worth commenting on the location of impacts with the stream, whether they occur when Pal~5 is close to its pericenter or not.
-As it can be seen from Fig.xx (bottom panel) encounters can occur at all orbital phases of Pal~5, when it is close to pericenter, but also very far away from it, at the outskirts of its orbital space.
+As it can be seen from Fig.\ref{fig:mass_size_plane} (bottom panel) encounters can occur at all orbital phases of Pal~5, when it is close to pericenter, but also very far away from it, at the outskirts of its orbital space.
 However, when taken all together, the location of gap creating impacts shows a strong negative correlation  with the galactocentric radius $r$, with the number $N$ of encounters favorable to gap formation  going as $N = -2.5r + 50$ (with a Pearson coefficient of -0.86).

### 2024-11-25 paper/main.tex
affiliations and little edits

@@
 
-\author{S. Ferrone
+\author{Salvatore Ferrone
        \inst{1,2}
          \and
-       M. Montuori\inst{1}
+       Marco Montuori\inst{1}
        \and
-       P. Di Matteo\inst{2}
+       Paola Di Matteo\inst{2}
        \and
-       A. Mastrobuono-Battisti
+       Alessandra Mastrobuono-Battisti \inst{3}
        \and
-       et al.
+       Rodrigo Ibata \inst{4}
+       \and 
+       Paolo Bianchini \inst{4}
+       \and
+       Sergey Khopserskov \inst{2,5}
+       \and
+       Nicolas Leclerc \inst{2}
+       \and 
+       Clement Hottier \inst{2}
+       \and 
+       Eliot Stein  \inst{2}
+       \and
+       David Valls-Gabaud \inst{2}
+       \and
+       Owain N. Snaith \inst{6}
        }
 
-\institute{Dipartimento di Fisica, Universit\`a di Roma ``La Sapienza'',
+\institute{
+          Dipartimento di Fisica, Universit\`a di Roma ``La Sapienza'',
           Piazza Aldo Moro\\
@@
       \and
-          Paris Observatory. Paris Sciences et Lettres\\
-          \email{c.ptolemy@hipparch.uheaven.space}
-          \thanks{The university of heaven temporarily does not
-                  accept e-mails}
+          Paris Observatory. Paris Sciences et Lettres
+      \and
+        Dipartimento di Fisica e Astronomia. ``Galileo Gallilei'' Università di Padova, Vicolo dell'Osservatorio 3 Padova 35122, Italy.
+      \and
+        Universit\'e de Strasbourg, CNRS, Observatoire astronomique de Strasbourg, UMR 7550, F-67000 Strasbourg, France
+      \and
+        Leibniz-Institut für Astrophysik Potsdam (AIP), An der Sternwarte 16, 14482 Potsdam, Germany
+      \and
+        Exeter
+      % \thanks{}
       }
 
-\date{Born in 1996; Accepted in 2018}
+      \date{Born in 1996; Accepted in 2025}
 
@@
 globular cluster stars beyond the tidal radius see the works of: \citet{1997A&A...320..776L, 2000A&A...356..127T, 2000A&A...359..907L, 2001AAS...19910906S, 2003AJ....126..815L}} extended Grillmair's findings to other globular cluster streams, but were often limited to the detections of stars still close to the cluster tidal radius, until the discovery made by \citet{2001ApJ...548L.165O,2002AAS...200.1001O, 2003AJ....126.2385O} of long and thin tails outside the Palomar~5 globular cluster.
-With a mass of $1.34\pm 0.24 \times 10^4 M_{\sun}$, Palomar~5 is one of the least massive globular clusters in the Galaxy.\citet{2003AJ....126.2385O} showed that its tails contain more mass than the cluster itself.
+With a mass of $1.34\pm 0.24 \times 10^4 M_{\sun}$, Palomar~5 is one of the least massive globular clusters in the Galaxy.
+\citet{2003AJ....126.2385O} showed that its tails contain more mass than the cluster itself.
 The works of \citet{2006ApJ...641L..37G} and \citet{2015MNRAS.446.3297K} showed the tails to be extended for more than $20^\circ$ degrees in the sky.
@@
     There is a primary methodological departure from \citet{2023A&A...673A..44F}.
-In that paper, globular clusters were let evolved under the gravitational effect of the Galaxy only, while now we include also the effect of all the other Galactic globular clusters, by taking into account the direct N-body interactions among them.
+In that paper, globular clusters evolved under the gravitational effect of the Galaxy only, while now we include also the effect of all the other Galactic globular clusters, by taking into account the direct N-body interactions among them.
 To quantify these interactions, all clusters are represented by Plummer spheres, each with mass and half-mass radius as reported in the Baumgardt catalogue.
@@
     To understand how the impacted particle is perturbed, one needs to compute its change in momentum, which is determined by integrating the force acting on the particle over the duration of the interaction.
-A useful approximation for this change in momentum \paola{(per unit mass)} is the force at the closest approach multiplied by an estimate of the interaction time:
+A useful approximation for this change in momentum, per unit mass, is the force at the closest approach multiplied by an estimate of the interaction time:
     \begin{equation} \label{eq:change_in_momentum} \Delta p \approx \text{Force} \times \text{interaction time} = \frac{GM}{b^2} \times 2\frac{b}{\delta v} = 2\frac{GM}{b \delta v}, \end{equation}where $M$ is the mass of the perturber, $b$ is the impact parameter, $\delta v$ is the relative velocity of the perturber with respect to the particle, and $G$ is the gravitational constant.
@@
 half-mass radius) of the clusters that cause gaps on the tails of Pal5, dividing them from those that do not.
-As can be seen, no cluster with mass below $10^5 \rm{M}_\odot$ causes gaps on Pal~5 stream, and all perturbers, except xx \paola{add which cluster has a $r_h > 10$~pc} have a half-mass radius below 10~pc.
+As can be seen, no cluster with mass below $10^5 \rm{M}_\odot$ causes gaps on Pal~5 stream, and all perturbers, except FSR~1758 have a half-mass radius below 10~pc.
 Even more interesting is their distribution in the E-Lz plane, which shows that the clusters that cause gaps are on both direct and retrograde orbits (respectively, negative and positive values of $L_z$), but all confined to an energy interval between $-2$ and $-1 \times10^5~\textrm{km}^2\textrm{s}^{-2}$.
@@
 Another possible explanation is the different thickness of the simulated tails at different distances from the center of the cluster.
-At distances of less than about 3~kpc from Pal~5, the tails are very thin, with a typical thickness of \paola{xx} pc.
+At distances of less than about 3~kpc from Pal~5, the tails are very thin, with a typical thickness being less than 100 pc.
 In order to cause gaps in these regions, the clusters would have to pass close to the stream, with impact parameters similar to the thickness itself.
 Note that the thinness of the tails at these distances is probably a direct consequence of the initial parameters we chose for the simulation.
-We have assumed for Pal5, 5 Gyr years ago, the same internal parameters (mass and size) that the cluster has today.
+We have assumed for Pal~5, 5 Gyr years ago, the same internal parameters (mass and size) that the cluster has today.
 With such parameters, after 5 Gyr of evolution, our N-body system has lost most of its mass, and therefore the part of the tails closest to the cluster itself, whose density depends mainly on the most recent mass loss \citep[see, for example, Fig.~A.3 in][]{2012A&A...546L...7M}, is necessarily very thin because the simulated cluster has essentially no more mass to lose.
@@
 
-  The impact of globular clusters close interactions with stellar streams has been also the object of another recent paper, by \citet{2022ApJ...941..129D}, who concluded that the chance for GD-1 gaps, which have been reported in a number of works \citep[see, for example, ][]{} to be produced by globular clusters is very low.
+  The impact of globular clusters close interactions with stellar streams has also been the object of another recent paper, by \citet{2022ApJ...941..129D}, who concluded that the chance for GD-1 gaps, which have been reported in a number of works \citep[see, for example, ][]{} to be produced by globular clusters is very low.
 This result is not necessarily in contradiction with ours, since GD-1 has a pericenter which is almost twice that of Palomar~5 \citep[see, for example][]{2019MNRAS.486.2995M}.

### 2024-11-25 paper/main.tex
small edits

@@
 This result is not necessarily in contradiction with ours, since GD-1 has a pericenter which is almost twice that of Palomar~5 \citep[see, for example][]{2019MNRAS.486.2995M}.
-As we have discussed in Sect.~\ref{sect:geometry}, the number, $N$, of close encounters which lead to gap creation is anti-correlated with the distance $r$ to the Galactic center, which implies that even simply applying the same radial dependence of $N$ as that found for Pal~5, but simply changing the pericenter value from $6$~kpc (roughly the value of Pal~5 pericenter \paola{can you check this value in your models?}), to 14~kpc (the closest distance of GD-1 to the Galactic center) reduced the number of gap-favorable impacts of more than a factor of 2.
-We, however, note that in their work \citet{2022ApJ...941..129D} preselect the  globular clusters which could have experienced a close encounter with the GD-1 stream, by selecting only clusters which pass at a distance less than 0.5 kpc from the stream, having a relative velocity smaller than  300 km/s.
+As we have discussed in Sect.~\ref{sect:geometry}, the number, $N$, of close encounters which lead to gap creation is anti-correlated with the distance $r$ to the Galactic center.
+If we naively use the same radial dependence of gaps from Palomar 5 for GD-1 by swapping a pericenter from 6~kpc to 16~kpc, we would reduced the number of gap-favorable impacts of more than a factor of 2.
+However, we note that \citet{2022ApJ...941..129D} preselect the globular clusters which could have experienced a close encounter with the GD-1 stream, by selecting only clusters which pass at a distance less than 0.5 kpc from the stream, having a relative velocity smaller than  300 km/s.
 As shown in our Fig.~\ref{fig:impact_geometry_statistics}, bottom-left panel, in the case of Palomar~5 this choice would lead to exclude most of the encounters favorable to gap creation, which turn out to have relative velocities above 300 km/s.
@@
 Other clusters and streams in the orbital energy range of Palomar~5 may suffer of the same difficulty.
-Going to lower orbital energies make the situation even worst, because the effect of the Galactic bar, giant molecular clouds and interactions with globular clusters becomes even more efficient.
+Going to lower orbital energies make the situation even worse, because the effect of the Galactic bar, giant molecular clouds and interactions with globular clusters becomes even more efficient.
 In the innermost regions of the Galaxy, where the density of dark matter sub-halos is expected to be maximal, globular cluster streams are intrinsically more difficult to be found, in addition to the fact that the dynamic times in these regions become so small that the stars lost from globular clusters do not redistribute themselves for the most part into thin structures \citep[see][]{2023A&A...673A..44F}.

### 2024-11-25 paper/main.tex
added a conclusion

@@
 \section{Conclusions}
-  \paola{to be written}
+  Our study demonstrates that globular cluster flybys can produce density gaps in the stellar streams of clusters like Palomar 5, but the occurrence and characteristics of these gaps depend heavily on the dynamical and structural properties of the stream and the Galactic environment.
+While our simulations predict the formation of gaps in Palomar 5's tails, the predicted gaps do not align with the regions of the stream currently observed.
+This discrepancy highlights the challenges in reconciling theoretical predictions with observational data, particularly given the complex interplay between orbital dynamics, stream structure, and the nature of close encounters.
+
+  Several factors contribute to the absence of simulated gaps in the observed portions of Palomar 5's tails, including the stream's varying thickness, the dispersion of action-frequencies near the cluster, and the initial conditions adopted in our simulations.
+Additionally, the gap creation rates predicted by our model fall short of explaining the number of gaps reported in previous observational studies, suggesting that globular cluster interactions alone cannot fully account for the observed stream substructure.
+
+  Moreover, our results reveal potential tensions with previous numerical studies, such as those by \citet{2019MNRAS.484.2009B}, which found a marginal role for globular clusters in gap formation.
+This discrepancy may stem from differences in the adopted Galactic potentials, as our analysis suggests that the mass of the dark matter halo in their study may be underestimated, altering the predicted interaction rates.
+
+  The broader implications of our findings indicate that globular cluster interactions add a layer of complexity to interpreting stellar stream substructures, complicating efforts to distinguish between baryonic and dark matter-induced gaps.
+While Palomar 5's phase-space region may not be ideal for isolating the impact of dark matter subhalos, extending this analysis to other streams, particularly those at larger galactocentric distances, could provide clearer insights.
+
+  Future work should involve a systematic comparison of gap creation rates across various Galactic potentials and extend the scope to a wider range of streams and globular clusters.
+This would help identify regimes where baryonic and dark matter-induced substructures can be disentangled, shedding light on the elusive influence of dark matter subhalos on stellar streams.
 

### 2024-12-02 paper/main.tex
paragraph for the conclusion

@@
 \section{Conclusions}
-  Our study demonstrates that globular cluster flybys can produce density gaps in the stellar streams of clusters like Palomar 5, but the occurrence and characteristics of these gaps depend heavily on the dynamical and structural properties of the stream and the Galactic environment.
-While our simulations predict the formation of gaps in Palomar 5's tails, the predicted gaps do not align with the regions of the stream currently observed.
+  Our study demonstrates that globular cluster flybys can produce density gaps in the stellar streams of Palomar~5, but the occurrence and characteristics of these gaps depend heavily on the dynamical and structural properties of the stream and the Galactic environment.
+While our simulations predict the formation of gaps in Palomar~5's tails, the predicted gaps do not align with the regions of the stream currently observed.
 This discrepancy highlights the challenges in reconciling theoretical predictions with observational data, particularly given the complex interplay between orbital dynamics, stream structure, and the nature of close encounters.

### 2024-12-02 paper/main.tex
added de baur citations

@@
 
-  The impact of globular clusters close interactions with stellar streams has also been the object of another recent paper, by \citet{2022ApJ...941..129D}, who concluded that the chance for GD-1 gaps, which have been reported in a number of works \citep[see, for example, ][]{} to be produced by globular clusters is very low.
+  The impact of globular clusters close interactions with stellar streams has also been the object of another recent paper, by \citet{2022ApJ...941..129D}, who concluded that the chance for GD-1 gaps, which have been reported in a number of works \citep[see, for example, ][]{2019ApJ...880...38B,2018MNRAS.477.1893D,2020AAS...23533607D} to be produced by globular clusters is very low.
 This result is not necessarily in contradiction with ours, since GD-1 has a pericenter which is almost twice that of Palomar~5 \citep[see, for example][]{2019MNRAS.486.2995M}.

### 2024-12-02 paper/main.tex
added xy plane

@@
       \includegraphics[width=0.45\linewidth]{mass_size_plane.png}
+      \includegraphics[width=0.45\linewidth]{impact_stats_phase_space_xy.png}
       \includegraphics[width=0.45\linewidth]{impact_stats_phase_space.png}
@@
 gap-causing clusters are indicated with colored points, whereas Pal~5 is indicated by a white symbol with a black edge.
-\textit{Bottom Right:} Palomar 5's orbit in the meridional plane.
+\textit{Bottom Left:} Palomar 5's orbit in the xy plane.
 The grayscale represents all 50 stacked orbits, with $\mathcal{P}$ indicating the probability of  Palomar~5's position, normalized to $\mathcal{P}_\textrm{max}=1$.
 Markers show the perturber's position during each impact.
-In all panels the colors of the markers and histogram bars are consistent with Fig.~\ref{fig:histogram_impact_time}.\paola{can you add also a panel with Pal~5 orbit and impacts in the (x-y) plane?}}
+The 50 white dots indicates Palomar~5's sampled initial conditions for the current day poisition.
+\textit{Bottom Right:} same as bottom left but in the meridional plane.
+In all panels the colors of the markers and histogram bars are consistent with Fig.~\ref{fig:histogram_impact_time}.}
       \label{fig:mass_size_plane}

### 2024-12-02 paper/main.tex
added the tabulated version of thie Histogram of the number of gaps

@@
     By applying the methodology described above to the whole set of simulations, we can reconstruct the history of close passages of Galactic globular clusters to Palomar~5's stream, in the last 5~Gyr, which -- we recall the reader -- is the time interval investigated in our simulations.
-The results of this analysis are reported in Fig.~\ref{fig:histogram_impact_time} and in Table~xx \paola{I suggest to report the results of Fig 3 also in a Table, in which we report for each GC which produced a gap, the time of close passage and the total number of gaps it generated.}.
+The results of this analysis are reported in Fig.~\ref{fig:histogram_impact_time} and in Table~\ref{tab:gaps_per_perturber}.
 NGC~2808 impacts Palomar 5's stream about 200 Myr ago in 44 out of 50 simulations, creating a small gap at a similar position to the one reported in Fig.~\ref{fig:decomposition} in each case.
@@
 Thus, in one configuration a cluster can impact the stream at a given time and yet at the same moment in a different set of initial conditions, it could be on the other side of the galaxy.
+
+    \begin{table}[h]
+      \centering
+      \begin{tabular}{|ll|ll|ll|}
+      \hline
+      NGC2808 & 44 & NGC7089 & 5 & NGC5272 & 4 \\
+      NGC6584 & 3 & NGC6341 & 2 & NGC6656 & 2 \\
+      NGC104 & 2 & NGC3201 & 1 & NGC5634 & 1 \\
+      NGC5286 & 1 & NGC362 & 1 & NGC5824 & 1 \\
+      NGC6356 & 1 & NGC6333 & 1 & NGC6715 & 1 \\
+      FSR1758 & 1 & NGC7078 & 1 & Pal10 & 1 \\
+      \hline
+      \end{tabular}
+      \caption{The number of gaps created by each perturber across all 50 simulations of Palomar~5's tidal tail. This is the same data Fig.~\ref{fig:histogram_impact_time} but tabulated.}
+      \label{tab:gaps_per_perturber}
+      \end{table}
 

### 2024-12-02 paper/main.tex
modified the conclusion

@@
 \section{Conclusions}
-  Our study demonstrates that globular cluster flybys can produce density gaps in the stellar streams of Palomar~5, but the occurrence and characteristics of these gaps depend heavily on the dynamical and structural properties of the stream and the Galactic environment.
+  Our study demonstrates that globular cluster flybys can produce density gaps in the stellar streams of Palomar~5.
+The occurrence and characteristics of these gaps depend on the dynamical and structural properties of the stream, the mass of the perturber, and the impact parameter.
 While our simulations predict the formation of gaps in Palomar~5's tails, the predicted gaps do not align with the regions of the stream currently observed.
-This discrepancy highlights the challenges in reconciling theoretical predictions with observational data, particularly given the complex interplay between orbital dynamics, stream structure, and the nature of close encounters.
-
 Several factors contribute to the absence of simulated gaps in the observed portions of Palomar 5's tails, including the stream's varying thickness, the dispersion of action-frequencies near the cluster, and the initial conditions adopted in our simulations.
-Additionally, the gap creation rates predicted by our model fall short of explaining the number of gaps reported in previous observational studies, suggesting that globular cluster interactions alone cannot fully account for the observed stream substructure.
-
-  Moreover, our results reveal potential tensions with previous numerical studies, such as those by \citet{2019MNRAS.484.2009B}, which found a marginal role for globular clusters in gap formation.
-This discrepancy may stem from differences in the adopted Galactic potentials, as our analysis suggests that the mass of the dark matter halo in their study may be underestimated, altering the predicted interaction rates.
 
@@
 This would help identify regimes where baryonic and dark matter-induced substructures can be disentangled, shedding light on the elusive influence of dark matter subhalos on stellar streams.
+Additionally, we can continue to add realisitic physics to the galactic enironment such as the merger of the Sagittarius dwarf galaxy and the passages of the large and small magellanic clouds.
 

### 2024-12-04 paper/main.tex
added comment about the model

@@
     \centering
-    \includegraphics[width=\linewidth]{stream_on_sky_Pal5_monte-carlo-009_pouliasis2017pii-GCNBody.png}
-    \caption{An instsance of the Palomar 5 stream created modeling the cluster as a Plummer sphere within an axis-symmetric galactic potential plus 164 other galactic globular clusters.
-The top plot is a scatter plot of star-particles that escaped the cluster due to tidal forces, the bottom plot shows the marginalized 1D density profile in longitude.
+    \includegraphics[width=\linewidth]{paper/stream_on_sky_Pal5_monte-carlo-009_pouliasis2017pii-GCNBody_pouliasis2017pii.png}
+    \caption{Simulated Palomar 5 sstream created by modeling the host cluster as a Plummer sphere disrupting within an axis-symmetric galactic potential plus the gravitational effect of 164 other galactic globular clusters.
+The top panel shows the distribution of star-particles that escaped the cluster due to tidal forces.
+The bottom panel shows the 1D density profile marginalized over longitude.
+The gray fill shows a reference simulation where the same conditions to produce the stream are used but the effect of other globular clusters are not included.
+The large central peak in density are of the particles still bound to Palomar~5.
+$\ell_0,b_0$ are located Palomar~5's cluster's center of mass.
 Two large gaps are present and are due to the passage of two globular clusters.
@@
 Two of these gaps are clearly visible in galactic coordinates and become even more apparent when marginalizing over latitude to reconstruct the 1D density profile of the stream, as a function of longitude.
+
+    Regarding the shape of the density distribution, the central peak corresponds to the still-intact globular cluster whose stars have not yet escaped.
+The density peaks in the stream are of the same order as the cluster itself.
+This is inconsistent with reality, the cluster should be much more populated than the stream's peak density.
+Of course, this discrepency a result of our modeling choices.
+We use the present day mass and radius for Palomar~5 for the whole duration of the simulation.
+A consequence of this is that the system is modeled as being less dense than it should have been.
+In turn, our simulations a strong initial mass which adds to the amplitudes of the profile density peaks of Fig.~\ref{fig:stream_on_sky}.
+This inaccuracy is acceptable for the scope of this work.
+First is that Palomar~5's tails indeed have more mass than the cluster itself.
+\citet{2017ApJ...842..120I} reports that there could be three times as much mass in Palomar 5's tails as the cluster itself.
+Secondly, the exact form of the density distribution is less important than haing a population of particles present that can probe a cluster flyby event.
 

### 2024-12-04 paper/main.tex
typos and commented out intro

@@
 
-  Over a century ago, \citet{1933AcHPh...6..110Z} postulated the existence of dark matter to explain the excessively large velocity dispersions of galaxies within their galaxy clusters.
+  % Over a century ago, \citet{1933AcHPh...6..110Z} postulated the existence of dark matter to explain the excessively large velocity dispersions of galaxies within their galaxy clusters.
 Since then, mounting evidence has been assembled to support this claim.
@@
 
-  Beyond the galactic scale, dark matter is essential for explaining many cosmological phenomena.
+  % Beyond the galactic scale, dark matter is essential for explaining many cosmological phenomena.
 For instance, the Planck mission \citep{2020A&A...641A...1P} was designed to study the Cosmic Microwave Background (CMB), whose spectrum closely follows that of a black body with a peak temperature in the millimeter wavelength range, corresponding to approximately 2.725~K.
@@
 
-  As laid out in \citet{2021PrPNP.11903865A}'s review, despite significant progress, the true nature of dark matter remains elusive and scientists continue to search for evidence of its existence beyond its gravitational effects on cosmic and galactic scales.
+  % As laid out in \citet{2021PrPNP.11903865A}'s review, despite significant progress, the true nature of dark matter remains elusive and scientists continue to search for evidence of its existence beyond its gravitational effects on cosmic and galactic scales.
 A key focus today is detecting dark matter on sub-galactic scales, because at these scales different cosmological scenarios predict different lumpiness of the halos.
 
-  \citet{2019Galax...7...81Z} provide a review for the different proposed mechanisms of forming dark matter haloes theoretically and how to model them pratically, showcasing how different assumptions for the fundamental particle behavior of dark matter propogates to the mass and size distribution of sub haloes on the galactic scale.
+  % \citet{2019Galax...7...81Z} provide a review for the different proposed mechanisms of forming dark matter haloes theoretically and how to model them practically, showcasing how different assumptions for the fundamental particle behavior of dark matter propogates to the mass and size distribution of sub haloes on the galactic scale.
 The authors rightly state that until particles are discovered $\Lambda$~CDM remains a hypothesis.
-Since the true nature of dark matter is elusive the possibilities for sub-galactic behaviour and structure of dark matter are vaste.
+Since the true nature of dark matter is elusive the possibilities for sub-galactic behaviour and structure of dark matter are vast.
 Obtaining constraints on the distirbution of dark matter subhaloes in turn will place constraints on the particle nature of dark matter.
@@
 
-  It is with the publication of the data from the Gaia astrometric mission \citep{2016A&A...595A...1G} that a new era for stellar stream astronomy has started  \citep{2024arXiv240519410B}.
+  \citet{2024arXiv240519410B} provides a review of stellar stream astronomy, which has enetered a new era since the publication of the data from the Gaia astrometric mission \citep{2016A&A...595A...1G}.
 Gaia's characterization of billions of stars in the Milky Way allows us to look for these structures coupling photometry, astrometry and, for the brightest stars, spectroscopy.
@@
 Currently, there are about a hundred stellar streams are currently known in our Galaxy.
-\citet{2023MNRAS.520.5225M} provides a compulation of their tracks on the sky.
+\citet{2023MNRAS.520.5225M} provides a compilation of their tracks on the sky.
 Interestingly, only about 20 streams associated to known Galactic globular clusters.
@@
 To date, the smallest observed dark matter halo was detected through gravitational lensing in an Einstein ring and had a mass of $10^8$ solar masses, as described by \citet{2012Natur.481..341V}.
-However, some models predict that dark matter clumps could exist down to the mass of Earth-like planets at least \citep[see][ and discussion in \citet{2021arXiv211101148A} ]{2005JCAP...08..003G, wang20}.
+However, some models predict that dark matter clumps could exist down to the mass of Earth-like planets at least \citep[see][ and discussion in \citet{2021arXiv211101148A} ]{2005JCAP...08..003G}.
 
@@
 \citet{2019ApJ...880...38B} provided further observational evidence for this idea, identifying a gap and a spur in the GD-1 stream that they could not explain by known objects, such as globular clusters, and which they suggested was due to the close passage of a dark matter sub-halo.
-Interestingly enough, however, the recovered properties of this sub-halo (mass and size) were inconsistent with the $\Lambda$~CDM mass-size relationship presented in \cite{2017MNRAS.466.4974M}.
+Interestingly enough, as stated in \citet{2019ApJ...880...38B}, the recovered properties of this sub-halo (mass and size) were inconsistent with the $\Lambda$~CDM mass-size relationship presented in \cite{2017MNRAS.466.4974M}.
 The above-mentioned works are only a few of a much more extensive literature that has explored the impact of dark matter sub-haloes in simulated streams \citep{2016ApJ...828L..10H, 2021MNRAS.507.1999H, 2021JCAP...10..043B, 2024arXiv240402953H, 2024arXiv241021174N}, or searched for their traces in observed streams \citep{2016MNRAS.460.2711T, 2017MNRAS.470...60E, 2020ApJ...889...70B, 2020ApJ...892L..37B}.
@@
 That the Galactic bar could have an influence on streams morphology was also discussed by \citet{2016MNRAS.460..497H} and \citet{2016ApJ...824..104P}, in the case of the Ophiuchus stream \citep{2014MNRAS.443L..84B}.
-Beside the Galactic bar, also giant molecular clouds can produce gaps in stellar streams, as shown by \citet{2016MNRAS.463L..17A}.
+Besides the Galactic bar, also giant molecular clouds can produce gaps in stellar streams, as shown by \citet{2016MNRAS.463L..17A}.
 All these works thus seem to indicate that baryonic structures can play an important role in tail morphology.
 In this context,  an extensive numerical study, specifically focused on modelling the tails of Palomar~5, under the influence of the Galactic bar, spiral arms, giant molecular clouds and globular clusters, has been realised by \citet{2019MNRAS.484.2009B}, who concluded that both the influence of the bar and that of the giant molecular clouds can leave imprints in Palomar~5 tidal tails similar to those left by dark matter sub-halos, while the effect of globular clusters was found to be negligible.
-A few additional studies have specifically investigated the effect of globular clusters on stellar streams.
-Beside the work by \citet{2017MNRAS.470...60E}, who concluded that density variations, and in particular the dips, they found in Palomar~5 tails could not be due to the passage of globular clusters.
+
+  Few studies have specifically investigated the effect of globular clusters on stellar streams.
+The work of \citet{2017MNRAS.470...60E} concluded that the observed density variations found in Palomar~5 tails could not be due to the passage of globular clusters.
 However, note that this conclusion was not based on a numerical modelling of this effect.
@@
 
-  In the present study we wish to fill this gap in the literature by reinvestigating this issue and quantify the importance of the effects of passing globular clusters in the vicinity of streams, to understand whether these systems can also be effective - and how often - in altering the distribution of stars in the tails and producing under-dense regions or gaps.
-To this end, on the following pages we will present the results of simulations of the streams of Pal5, subject to the gravitational interaction with the set of 165 Galactic globular clusters for which positions and velocities are known to date, and for which orbits can therefore be reconstructed.
-We chose to simulate Pal~5 because it is a typical halo cluster with extended tails, and because it is a cluster for which the effect of baryonic structures on its stream has already been studied.
+  In the present study we wish to fill this gap in the literature of numerically modeling cluster-stream interactions.
+We seak to quantify the importance of the effects of passing globular clusters in the vicinity of streams, to understand whether these systems can also be effective - and how often - in altering the distribution of stars in the tails and producing under-dense regions or gaps.
+To this end, on the following pages we will present the results of simulations of the streams of Palomar~5, subject to the gravitational interaction with the set of 165 Galactic globular clusters for which positions and velocities are known to date, and for which orbits can therefore be reconstructed.
+We chose to simulate Palomar~5 because it is a typical halo cluster with extended tails, and because it is a cluster for which the effect of baryonic structures on its stream has already been studied.
 As we will show, and in tension with previous claims, the close passage of other clusters with the stream of Pal~5 is not a rare event, and indeed, in the 50 simulations we ran, we find the formation of numerous gaps, for an average of 1.5 gaps per simulation, generated by 18 different clusters among the whole system of Galactic globular clusters.
@@
     \centering
-    \includegraphics[width=\linewidth]{paper/stream_on_sky_Pal5_monte-carlo-009_pouliasis2017pii-GCNBody_pouliasis2017pii.png}
+    \includegraphics[width=\linewidth]{stream_on_sky_Pal5_monte-carlo-009_pouliasis2017pii-GCNBody_pouliasis2017pii.png}
     \caption{Simulated Palomar 5 sstream created by modeling the host cluster as a Plummer sphere disrupting within an axis-symmetric galactic potential plus the gravitational effect of 164 other galactic globular clusters.
@@
 This model is time-independent throughout our simulations.
-This model satisfies a series of observational constraints such as local solar stellar density, the galactic rotation curve, etc, similarly to other galactic models such as Bovy2014, McMillian.
+This model satisfies a series of observational constraints such as local solar stellar density, the galactic rotation curve, etc, similarly to other galactic models such as \texttt{MWpotential2014} from \citet{2015ApJS..216...29B} and \texttt{McMillian2017} from \citet{2017MNRAS.465...76M}.
 However, To balance data volume and computation time, we use only one galactic potential, which should suffice.

### 2024-12-04 paper/main.tex
more changes from sergey

@@
 \abstract
+%  menntion that only one galactic model should be sufficient to avoid the criticism
   {Thin stellar streams, such as those resulting from the tidal disruption of globular clusters, have long been known and used as probes of the gravitational potential of our Galaxy, both its visible and dark contents.
@@
 Similarly, \citet{2024MNRAS.528.5189G} generated a globular cluster mass-loss catalog using seven different potential models and found their debris distributions to be model-independent, similarly to \citet{2023A&A...673A..44F}.
-While the clusters' positions may depend on the model, we assert that interaction rates and stream formation are unaffected by the choice of galactic model.
+While the clusters' positions may depend on the model, we assert that interaction rates and stream formation are unaffected by the choice of galactic potential model.
 
@@
 In that paper, globular clusters evolved under the gravitational effect of the Galaxy only, while now we include also the effect of all the other Galactic globular clusters, by taking into account the direct N-body interactions among them.
-To quantify these interactions, all clusters are represented by Plummer spheres, each with mass and half-mass radius as reported in the Baumgardt catalogue.
+To quantify these interactions, all clusters are represented by Plummer spheres, each with mass and half-mass radius as reported in the Baumgardt catalogue \citep{2021MNRAS.505.5957B}.
 In the following of this paper, we refer to these simulations, where the globular clusters interactions are taken into account as \texttt{full} simulations.
@@
       \begin{equation}
-        \ddot{\vec{r}}_i = -\nabla \Phi(R_i,z_i) + \left.\sum_{j\neq i} \frac{Gm_j}{\left(|\vec{r}_j - \vec{r}_i|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j - \vec{r}_i\right),
-      \end{equation}\label{eq:GCNBody} where $\vec{r}$ indicates the galactocentric position, the index $i$ indicates the globular cluster of interest with $R_i,z_i$ being its galactocentric positions in cylindrical coordinates; the index $j$ indicates the other 164 globular clusters that are summed over, where $m_j$ is the mass of the j-th cluster in the sample, $b_j$ is its Plummer scale radius, and $\vec{r_j}$ is its galactocentric position.
+        \ddot{\vec{r}}_i = -\nabla \Phi(R_i,z_i) + \left.\sum_{j\neq i}^{N_{GC}} \frac{Gm_j}{\left(|\vec{r}_j - \vec{r}_i|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j - \vec{r}_i\right),
+      \end{equation}\label{eq:GCNBody} where $\vec{r}$ indicates the galactocentric position vector, the index $i$ indicates the globular cluster of interest with $R_i,z_i$ being its galactocentric positions in cylindrical coordinates; the index $j$ indicates the other globular clusters that are summed over.
+$N_{GC}$ is the total number of globular clusters which in this study is 165, $m_j$ is the mass of the j-th cluster in the sample, $b_j$ is its Plummer scale radius, and $\vec{r_j}$ is its galactocentric position.
 $\Phi$ represents the Galactic smooth potential \citep[][Model~II, in the present case]{2017A&A...598A..66P}.
@@
     \begin{equation}
-      \ddot{\vec{r}}_p = -\nabla \Phi(R_p,z_p) + \left.\sum_{j} \frac{Gm_j}{\left(|\vec{r}_j(t) - \vec{r}_p|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j(t)- \vec{r}_p\right),
-      \end{equation} \label{eq:equation_of_motion_particle} where the index $p$ represents one of the 100,000 particles of interest, and $j$ indexes over the globular clusters as in Eq.~\ref{eq:GCNBody}
+      \ddot{\vec{r}}_p = -\nabla \Phi(R_p,z_p) + \left.\sum_{j}^{N_{GC}} \frac{Gm_j}{\left(|\vec{r}_j(t) - \vec{r}_p|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j(t)- \vec{r}_p\right),
+      \end{equation} \label{eq:equation_of_motion_particle} where the index $p$ represents one of the 100,000 particles of interest, $\vec{r_p}$ being its position, and $j$ indexes over the globular clusters as in Eq.~\ref{eq:GCNBody}.
 
@@
 This is inconsistent with reality, the cluster should be much more populated than the stream's peak density.
-Of course, this discrepency a result of our modeling choices.
+Of course, this discrepancy is a result of our modeling choices.
 We use the present day mass and radius for Palomar~5 for the whole duration of the simulation.
@@
 \citet{2017ApJ...842..120I} reports that there could be three times as much mass in Palomar 5's tails as the cluster itself.
-Secondly, the exact form of the density distribution is less important than haing a population of particles present that can probe a cluster flyby event.
+Secondly, the exact form of the density distribution is less important than having a population of particles present that can probe a cluster flyby event.
 

### 2024-12-05 paper/main.tex
more changes
added Ishchenko citation for 5 gyr orbit integration choice

@@
 
+    Lastly, we select an integration time of 5 Gyr as a compromise between maximizing interaction statistics and adhering to the assumption that the galaxy can be modeled as a time-independent, constant mass distribution.
+\citet{2023A&A...673A.152I} analyzed the orbits of the Galactic cluster population using the same initial conditions as in this work, within five live Milky Way-like potentials from IllustrisTNG \citep{2018MNRAS.473.4077P}.
+They found that in all sampled potentials, orbital changes remain minimal over 5 Gyr, becoming significant only at earlier look-back times when the host galaxy had significantly less mass or was undergoing a merger event.
+
     \subsubsection*{ \texttt{Full} simulations}
@@
       \end{equation} \label{eq:equation_of_motion_particle} where the index $p$ represents one of the 100,000 particles of interest, $\vec{r_p}$ being its position, and $j$ indexes over the globular clusters as in Eq.~\ref{eq:GCNBody}.
+Note in that Eq.~\ref{eq:equation_of_motion_particle} the positions of the globular clusters are time-dependent since they are being loaded during this step and not computed.
 

### 2024-12-05 paper/main.tex
cited the GC-GC interactions from Ishchenko

@@
 These first works seem to suggest that the impact of globular clusters on stellar streams is negligible, and this is somewhat surprising, especially for streams of clusters as Palomar~5, which are confined in the inner 20~kpc of the Galaxy, where many other globular clusters do also orbit.
+Additionally, \citet{2023A&A...678A..69I} showed that globular cluster can even collide with other clusters, which implies that cluster stream collisions should happen much more frequently since streams are far more extended than clusters.
 
@@
 
-    Lastly, we select an integration time of 5 Gyr as a compromise between maximizing interaction statistics and adhering to the assumption that the galaxy can be modeled as a time-independent, constant mass distribution.
+    Lastly, we select an integration time of 5~Gyr as a compromise between maximizing interaction statistics and adhering to the assumption that the galaxy can be modeled as a time-independent, constant mass distribution.
 \citet{2023A&A...673A.152I} analyzed the orbits of the Galactic cluster population using the same initial conditions as in this work, within five live Milky Way-like potentials from IllustrisTNG \citep{2018MNRAS.473.4077P}.
@@
 A consequence of this is that the system is modeled as being less dense than it should have been.
-In turn, our simulations a strong initial mass which adds to the amplitudes of the profile density peaks of Fig.~\ref{fig:stream_on_sky}.
+In turn, our simulations have a strong initial mass which adds to the amplitudes of the profile density peaks of Fig.~\ref{fig:stream_on_sky}.
 This inaccuracy is acceptable for the scope of this work.

### 2024-12-05 paper/main.tex
a series of more small edits from sergey

@@
        Alessandra Mastrobuono-Battisti \inst{3}
+       \and 
+       Misha Haywood \inst{2}
        \and
@@
         \end{equation} \label{eq:equation_of_motion_particle} where the index $p$ represents one of the 100,000 particles of interest, $\vec{r_p}$ being its position, and $j$ indexes over the globular clusters as in Eq.~\ref{eq:GCNBody}.
-Note in that Eq.~\ref{eq:equation_of_motion_particle} the positions of the globular clusters are time-dependent since they are being loaded during this step and not computed.
+Note in that Eq.~\ref{eq:equation_of_motion_particle} the positions of the globular clusters are time-dependent since they are being loaded during this step and not computed, unlike Eq.~\ref{eq:equation_of_motion_particle}.
 
@@
 On the other hand, in physical space streams on eccentric orbits indeed grow in time yet the growth rate is modulated with periodic expansion and contraction depending on the orbital phase.
-Adopting this time space, and reporting the gravitational acceleration along Pal~5 orbit in this space, the identification of the globular clusters which produced the perturbation, and the time at which it occurred becomes straightforwar.
+Adopting this time space, and reporting the gravitational acceleration along Pal~5 orbit in this space, the identification of the globular clusters which produced the perturbation, and the time at which it occurred becomes straight forward.
 We refer the reader to Appendix~\ref{sec:Perturber_Identification} for an in-depth description of the procedure.
@@
     In total, we report the finding of 73 gaps in our 50 simulations which is on average 1.5 gaps/simulation and were created by 18 different perturbers.
-The distribution of simulations with a given number of gaps is provided in Table~\ref{table:gap_distribution}.
+The distribution of number of gaps appearing per simulation is presented in Table~\ref{table:gap_distribution}.
 If we consider NGC~2808 to be an outlier and exclude it from the experiment, we observe an average of 0.6 gaps per simulation.
-Including all observed gaps results in an average of 1.5 gaps per simulation.
 
@@
 
-    We need a more sophisticated satistic that can be compared to other simulations and turn to the gap creation rate developed by \citet{2012ApJ...748...20C}.
-This is the number of gaps per time and normalized by the length of the stream.
+    We need a more sophisticated satistic that can be compared to other simulations and turn to the \textit{gap creation rate} developed by \citet{2012ApJ...748...20C}.
+The gap creation rate is the number of gaps that appear per time and normalized by the length of the stream.
 For our simulations, this rate is given by: \begin{equation} \label{eq:gap_creation_rate} \mathcal{R}_{\textrm{Pal 5}} =  \frac{1}{T}\int_{0}^T l^{-1}(t) \sum_i \delta(t-t_i) dt,\end{equation}where $T$ is the total integration time, $l(t)$ is the length of the stream, and $\sum_i \delta(t-t_i)$ sums over the gap occurrences, with $i$ indexing over the number of gaps in a given simulation.
@@
 The results are shown in Fig.~\ref{fig:gapcreationrate}.
-This is roughly consistent with a simple estimate of the average gap creation rate: with 73 gaps over 5 Gyr of integration time for a stream about 20 kpc in length, the naive estimate is approximately 0.015~km~s$^{-1}$~kpc$^{-2} = 0.015~\rm{Gyr^{-1}kpc^{-1}}$.
+This is roughly consistent with a simple estimate of the average gap creation rate: with 73 gaps over 5 Gyr of integration time for a stream about 20 kpc in length, the naive estimate is approximately 0.015~km~s$^{-1}$~kpc$^{-2} $ (which is roughly equivalent to $0.015~\rm{Gyr^{-1}kpc^{-1}}$).
 This is about double the weighted mean gap creation rate of 0.009~km~s$^{-1}$~kpc$^{-2}$.
@@
 Since the lengths are equal, and the offset between the tails is small compared to the galactocentric distance of about 10 kpc, we expected the number of gaps in each tail to be more or less the same.
-We can compute the the probability of observing the unequal occurances though the binomail distirbution.
+We can compute the probability of observing the unequal occurances though the binomail distirbution.
 First, the since the 44 gaps linked to NGC~2808 are the same fly-by, they are not independent events.
@@
 
-    From this analysis, we can conclude that a more massive perturber, passing closer to the particle, and moving more slowly, will have a greater impact.
+    This equation asserts that a more massive perturber, passing closer to the particle, and moving more slowly, will have a greater impact.
 It is important to note that the momentum change is inversely proportional to the velocity of the perturber, which contrasts with the intuition from elastic collisions, such as those between billiard balls, where higher velocities result in greater impacts.
@@
     Visual inspection of the five key impact parameters ($M$, $r_p$, $b$, $W_\parallel$, and $W_\perp$) did not reveal a clear distinction between flybys that create gaps and those that do not.
-Therefore, we only present the the quantities from Eq.~\ref{eq:change_in_momentum} in Fig.~\ref{fig:impact_geometry_statistics}.
+Therefore, we only present the quantities from Eq.~\ref{eq:change_in_momentum} in Fig.~\ref{fig:impact_geometry_statistics}.
     Note that in this figure, we show the total relative velocity rather than separating parallel and perpendicular components, because from a visual inspection, we could not find any specific trends when the two velocity components were plotted separately.
@@
 \textit{Bottom Right:} same as bottom left but in the meridional plane.
-In all panels the colors of the markers and histogram bars are consistent with Fig.~\ref{fig:histogram_impact_time}.}
+In all panels the colors of the markers and histogram bars correspond to specific perturber as specified in Fig.~\ref{fig:histogram_impact_time}.}
       \label{fig:mass_size_plane}
@@
 However, when taken all together, the location of gap creating impacts shows a strong negative correlation  with the galactocentric radius $r$, with the number $N$ of encounters favorable to gap formation  going as $N = -2.5r + 50$ (with a Pearson coefficient of -0.86).
-While there are more clusters near the galactic center, orbits naturally spend more time near their apocenters, and the lower relative velocities in these regions should favor gap creation.
+While there are more clusters near the galactic center, clusters naturally spend more time near their apocenters, and the lower relative velocities in these regions should favor gap creation.
 However, this result suggests that cluster galactic number density outweighs these factors in determining the number of gaps.
+It should be noted that these results are for Palomar 5 only.
+Streams along various orbits needs to be studied in order for this conclusion to be generalized.
 
@@
 
-  The simulations presented in this paper suggest that, in the last 5 Gyr of evolution, Palomar~5 should have experienced close encounters with other Galactic globular clusters, some of which are able to create gaps -- even a few kpc wide -- in its tails.
+  The simulations presented in this paper suggest that, in the last 5 Gyr of evolution, Palomar~5's stream could have experienced multiple close encounters with other Galactic globular clusters, some of which are able to create gaps -- even a few kpc wide -- in its tails.
+
   Unfortunately, none of the predicted gaps is located in regions of the tails that have been observed, and this is especially true for the leading tail, which result shorter than the trailing tail in the observations \citep[see][for a possible explanation of this feature]{2017NatAs...1..633P}.
@@
 After binning the 1D profiles, we apply a boxcar smoothing technique.
-The density profiles are scanned.
-This raw profile density curve is then smoothed with the median box-car filter.
+This raw profile density curve is smoothed with the median box-car filter.
 At each bin, a selected number of adjacent data points from both sides are placed in a list.
@@
 The bottom plot showcases the evolution of the stream density in simulation time.
-The density was used to determine a sutible length of the stream at a given function.
+The density was used to determine a suitable length of the stream at a given function.
 This length was then used to extract a piece of Palomar 5's orbit.

### 2024-12-05 paper/main.tex
added Misha to author and did spell checker

@@
   {Thin stellar streams, such as those resulting from the tidal disruption of globular clusters, have long been known and used as probes of the gravitational potential of our Galaxy, both its visible and dark contents.
-In particular, the presence of under-density regions, or gaps, along these streams is commonly interpreted as being due to the close passage of dark matter sub-haloes.
+In particular, the presence of under-density regions, or gaps, along these streams is commonly interpreted as being due to the close passage of dark matter sub-halos.
 }
@@
 Numerical predictions of this phenomenon exist since the 1970s \citep[see, for example][]{1975AJ.....80..290K}.
-These predictions occured well before the first detections which were made by \citet{1995AJ....109.2553G} around some Galactic globular clusters.
+These predictions occurred well before the first detections which were made by \citet{1995AJ....109.2553G} around some Galactic globular clusters.
 Interestingly, this work was nearly contemporary discovery of the Sagittarius stellar stream by \citet{1994Natur.370..194I} which is the closest example of a stream emerging from a dwarf satellite currently interacting with the Milky Way.
@@
 
-  \citet{2024arXiv240519410B} provides a review of stellar stream astronomy, which has enetered a new era since the publication of the data from the Gaia astrometric mission \citep{2016A&A...595A...1G}.
+  \citet{2024arXiv240519410B} provides a review of stellar stream astronomy, which has entered a new era since the publication of the data from the Gaia astrometric mission \citep{2016A&A...595A...1G}.
 Gaia's characterization of billions of stars in the Milky Way allows us to look for these structures coupling photometry, astrometry and, for the brightest stars, spectroscopy.
@@
   The possibility to couple Gaia data with spectroscopic surveys, has extended the study of stellar streams beyond the quantification of their orbital properties, to a full chemical characterization.
-For example \citet{2019MNRAS.490.3508L} presents $S^5$, and is one example of a cross-catalogue among other works \citep{2020AJ....160..181J, 2021ApJ...911..149L, 2022ApJ...928...30L, 2024MNRAS.529.2413U}.
+For example \citet{2019MNRAS.490.3508L} presents $S^5$, and is one example of a cross-catalog among other works \citep{2020AJ....160..181J, 2021ApJ...911..149L, 2022ApJ...928...30L, 2024MNRAS.529.2413U}.
 Currently, there are about a hundred stellar streams are currently known in our Galaxy.
@@
 Compared to the measurement of the HI rotation curve, stellar streams offer the opportunity to investigate the potential of the host galaxy over a wide range of distances, reaching the outermost regions of the halo.
-For example, \citet{2011MNRAS.417..198V} demonstrated how stellar streams can infer the mass and scale parameters of dark matter haloes using various amounts of observational data, from basic right ascension and declination to full six-dimensional phase space information.
+For example, \citet{2011MNRAS.417..198V} demonstrated how stellar streams can infer the mass and scale parameters of dark matter halos using various amounts of observational data, from basic right ascension and declination to full six-dimensional phase space information.
 \citet{2018ApJ...867..101B} reviewed this concept from an information-theoretic point of view, identifying which orbits and configurations of stellar streams yield the most information about the galactic potential.
@@
 According to simulations by \citet{2008MNRAS.391.1685S}, $\Lambda$-CDM predicts that galaxies grow hierarchically, with dark matter clumps forming at a wide range of masses and sizes.
-These clumps, or \textit{subhalos}, are predicted to follow a mass distribution with a power-law slope slightly shallower than -2.0.
+These clumps, or \textit{sub-halos}, are predicted to follow a mass distribution with a power-law slope slightly shallower than -2.0.
 To date, the smallest observed dark matter halo was detected through gravitational lensing in an Einstein ring and had a mass of $10^8$ solar masses, as described by \citet{2012Natur.481..341V}.
@@
 
-  \citet{2002MNRAS.332..915I} first suggested that dark matter subhalos could influence stellar streams by diffusing their orbital elements.
-Later, \citet{2012ApJ...748...20C} expanded this idea, proposing that subhalos could create gaps in stellar streams during flyby encounters, where a subhalo approaches closely enough to a segment of a stream and significantly changes the orbits of the closest stars.
+  \citet{2002MNRAS.332..915I} first suggested that dark matter sub-halos could influence stellar streams by diffusing their orbital elements.
+Later, \citet{2012ApJ...748...20C} expanded this idea, proposing that sub-halos could create gaps in stellar streams during flyby encounters, where a sub-halo approaches closely enough to a segment of a stream and significantly changes the orbits of the closest stars.
 \citet{2012ApJ...760...75C, 2013ApJ...768..171C} applied this idea and looked for the presence of gaps in the known Palomar~5 and GD~1 streams, concluding that the density variations found in their streams were consistent with expectations from $\Lambda$~CDM models.
@@
 Interestingly enough, as stated in \citet{2019ApJ...880...38B}, the recovered properties of this sub-halo (mass and size) were inconsistent with the $\Lambda$~CDM mass-size relationship presented in \cite{2017MNRAS.466.4974M}.
-The above-mentioned works are only a few of a much more extensive literature that has explored the impact of dark matter sub-haloes in simulated streams \citep{2016ApJ...828L..10H, 2021MNRAS.507.1999H, 2021JCAP...10..043B, 2024arXiv240402953H, 2024arXiv241021174N}, or searched for their traces in observed streams \citep{2016MNRAS.460.2711T, 2017MNRAS.470...60E, 2020ApJ...889...70B, 2020ApJ...892L..37B}.
+The above-mentioned works are only a few of a much more extensive literature that has explored the impact of dark matter sub-halos in simulated streams \citep{2016ApJ...828L..10H, 2021MNRAS.507.1999H, 2021JCAP...10..043B, 2024arXiv240402953H, 2024arXiv241021174N}, or searched for their traces in observed streams \citep{2016MNRAS.460.2711T, 2017MNRAS.470...60E, 2020ApJ...889...70B, 2020ApJ...892L..37B}.
 
   In contrast, a more limited number of works have been interested in understanding whether other structures, e.g.
-from baryonic matter, can cause variations in the density of streams and gaps that can be assimilated and confused with those produced by dark matter sub-haloes.
+from baryonic matter, can cause variations in the density of streams and gaps that can be assimilated and confused with those produced by dark matter sub-halos.
 Among these works, it is worth mentioning the results of \citet{2017NatAs...1..633P} that suggests how the presence of the bar at the center of the Galaxy can perturb the characteristics of a stream such as Pal~5 and generate gaps along its tail.
@@
 All these works thus seem to indicate that baryonic structures can play an important role in tail morphology.
-In this context,  an extensive numerical study, specifically focused on modelling the tails of Palomar~5, under the influence of the Galactic bar, spiral arms, giant molecular clouds and globular clusters, has been realised by \citet{2019MNRAS.484.2009B}, who concluded that both the influence of the bar and that of the giant molecular clouds can leave imprints in Palomar~5 tidal tails similar to those left by dark matter sub-halos, while the effect of globular clusters was found to be negligible.
+In this context,  an extensive numerical study, specifically focused on modeling the tails of Palomar~5, under the influence of the Galactic bar, spiral arms, giant molecular clouds and globular clusters, has been realized by \citet{2019MNRAS.484.2009B}, who concluded that both the influence of the bar and that of the giant molecular clouds can leave imprints in Palomar~5 tidal tails similar to those left by dark matter sub-halos, while the effect of globular clusters was found to be negligible.
 
@@
 The work of \citet{2017MNRAS.470...60E} concluded that the observed density variations found in Palomar~5 tails could not be due to the passage of globular clusters.
-However, note that this conclusion was not based on a numerical modelling of this effect.
+However, note that this conclusion was not based on a numerical modeling of this effect.
 More recently \citet{2022ApJ...941..129D} have examined the possibility that the gaps found in the GD-1 stellar stream could be due to the close passage of globular clusters, concluding that this scenario is highly unlikely.
@@
   In the present study we wish to fill this gap in the literature of numerically modeling cluster-stream interactions.
-We seak to quantify the importance of the effects of passing globular clusters in the vicinity of streams, to understand whether these systems can also be effective - and how often - in altering the distribution of stars in the tails and producing under-dense regions or gaps.
+We seek to quantify the importance of the effects of passing globular clusters in the vicinity of streams, to understand whether these systems can also be effective - and how often - in altering the distribution of stars in the tails and producing under-dense regions or gaps.
 To this end, on the following pages we will present the results of simulations of the streams of Palomar~5, subject to the gravitational interaction with the set of 165 Galactic globular clusters for which positions and velocities are known to date, and for which orbits can therefore be reconstructed.
@@
     \includegraphics[width=\linewidth]{stream_on_sky_Pal5_monte-carlo-009_pouliasis2017pii-GCNBody_pouliasis2017pii.png}
-    \caption{Simulated Palomar 5 sstream created by modeling the host cluster as a Plummer sphere disrupting within an axis-symmetric galactic potential plus the gravitational effect of 164 other galactic globular clusters.
+    \caption{Simulated Palomar 5 stream created by modeling the host cluster as a Plummer sphere disrupting within an axis-symmetric galactic potential plus the gravitational effect of 164 other galactic globular clusters.
 The top panel shows the distribution of star-particles that escaped the cluster due to tidal forces.
@@
 We begin by extracting positions in the sky, proper motions, line-of-sight velocities and distances, as well as masses and half-mass radii, of 165 globular clusters from the Galactic globular cluster catalog by \cite{2021MNRAS.505.5957B}.\footnote{The Baumgardt catalog has been assembled across a series of works, see: \cite{2020PASA...37...46B,2019MNRAS.482.5138B,2018MNRAS.478.1520B}.
-The catalog can be found on the World Wide Web at \href{https://people.smp.uq.edu.au/HolgerBaumgardt/globular/}{https://people.smp.uq.edu.au/HolgerBaumgardt/globular/}.} We then convert the initial conditions from sky coordinates into a Galactocentric reference frame, by adopting a velocity for the local standard of rest of $v_{\text{LSR}} = 240$~km~s$^{-1}$ and a peculiary velocity of the Sun equal to $(U_\odot, V_\odot, W_\odot)=(11.1, 12.24, 7.25)$~km~s$^{-1}$, as reported by \citet{2012MNRAS.427..274S}.
+The catalog can be found on the World Wide Web at \href{https://people.smp.uq.edu.au/HolgerBaumgardt/globular/}{https://people.smp.uq.edu.au/HolgerBaumgardt/globular/}.} We then convert the initial conditions from sky coordinates into a Galactocentric reference frame, by adopting a velocity for the local standard of rest of $v_{\text{LSR}} = 240$~km~s$^{-1}$ and a peculiar velocity of the Sun equal to $(U_\odot, V_\odot, W_\odot)=(11.1, 12.24, 7.25)$~km~s$^{-1}$, as reported by \citet{2012MNRAS.427..274S}.
 The position of the Sun was set to $(x_\odot,y_\odot,z_\odot) = (-8.34,0,0.027)$~kpc, vertical position above the disk was taken from \citet{2001ApJ...553..184C} and the distance of the Sun to the Galactic center was taken from \citet{2014ApJ...783..130R}.
@@
 In that paper, globular clusters evolved under the gravitational effect of the Galaxy only, while now we include also the effect of all the other Galactic globular clusters, by taking into account the direct N-body interactions among them.
-To quantify these interactions, all clusters are represented by Plummer spheres, each with mass and half-mass radius as reported in the Baumgardt catalogue \citep{2021MNRAS.505.5957B}.
+To quantify these interactions, all clusters are represented by Plummer spheres, each with mass and half-mass radius as reported in the Baumgardt catalog \citep{2021MNRAS.505.5957B}.
 In the following of this paper, we refer to these simulations, where the globular clusters interactions are taken into account as \texttt{full} simulations.
@@
 
-        \item Once the positions and velocities of the entire globular cluster system, as to 5~Gyr ago, are found, we sample Palomar~5 with  100,000 particles which follow a Plummer distribution of mass and half-mass radius as in the Baumgardt catalogue.
+        \item Once the positions and velocities of the entire globular cluster system, as to 5~Gyr ago, are found, we sample Palomar~5 with  100,000 particles which follow a Plummer distribution of mass and half-mass radius as in the Baumgardt catalog.
 We then integrate the evolution of these particles forward in time to the present day, by taking into account that each particle feels the gravitational potential of the Galaxy, that of its host cluster, and that of all the other clusters in the Galaxy.
 Note that we do not take into account self-gravity among particles.
-The particles expereince the gravitational field yet do not contribute to it.
+The particles experience the gravitational field yet do not contribute to it.
 The equation of motion of a generic particle among the 100,000 which populate Pal~5 is thus:
@@
 
-      The procedure described so far has been repeated 50 times, each time generating a new set of initial conditions, given the uncertainties on proper motions, line-of-sight velocities, distances to the Sun, and masses, of all clusters, as reported in the Baumgardt catalogue.
-We handle these uncertainties  through a Monte-carlo approach, by sampling them with gaussian distribution, and by taking into account the covariance term between the proper motions.
+      The procedure described so far has been repeated 50 times, each time generating a new set of initial conditions, given the uncertainties on proper motions, line-of-sight velocities, distances to the Sun, and masses, of all clusters, as reported in the Baumgardt catalog.
+We handle these uncertainties  through a Monte-Carlo approach, by sampling them with Gaussian distribution, and by taking into account the covariance term between the proper motions.
 For the first simulation, we use the most probable values for the initial conditions.
@@
       During the integration, intermediate snapshots were saved to facilitate the analysis of stellar streams and the effects of cluster impacts.
-Specifically, for each of the 50 realizations of the Palomar 5 stream, we saved $5000$ intermediate timesteps which is a temporal resolution of 1 million years.
-We provide the parameters that specify our data volumne in Table~\ref{tab:data_volume}.
+Specifically, for each of the 50 realizations of the Palomar 5 stream, we saved $5000$ intermediate time-steps which is a temporal resolution of 1 million years.
+We provide the parameters that specify our data volume in Table~\ref{tab:data_volume}.
 Using single precision floating point numbers, the size of our simulations is approximately:
@@
       \caption{The parameters determining the data volume of this experiment.
-where $N_p$  is the number of particles, $N_{\textrm{ts}}$ is the number of time-steps saved, $N_{\textrm{phase}}$ is the number of phase space coordinates, and $N_{\textrm{sampling}}$ is the number of monte-carlo samplings of the initial conditions.}
+where $N_p$  is the number of particles, $N_{\textrm{ts}}$ is the number of time-steps saved, $N_{\textrm{phase}}$ is the number of phase space coordinates, and $N_{\textrm{sampling}}$ is the number of Monte-Carlo samplings of the initial conditions.}
       \label{tab:data_volume}
@@
 This method is preferable to, for instance, a Runge-Kutta scheme, which can introduce non-physical and significant numerical errors in systems that require long-term stability and energy conservation.
-One drawback of the leapfrog integrator is that it requires a uniform timestep across the entire computation, leading to unnecessary computations for a particle after it escapes form the host cluster.
+One drawback of the leapfrog integrator is that it requires a uniform time-step across the entire computation, leading to unnecessary computations for a particle after it escapes form the host cluster.
 However, energy conservation and phase-space volume preservation are paramount when modeling stellar streams.
 The time step was therefore set to be small enough to conserve energy for the most interior particles within the cluster---ensuring that a higher mass loss did not arise from numerical error.
-For these simulations, a timestep of 10,000 years was selected to maintain adequate energy conservation.
+For these simulations, a time-step of 10,000 years was selected to maintain adequate energy conservation.
 
@@
       \includegraphics[width=\linewidth]{decomposition-monte-carlo-009-with-3-gaps.png}
-      \caption{This figure decomposes the gaps shown in the middle panel of Fig.~\ref{fig:profiles} into the individual perturbuers whose cause them, as identified from Fig.~\ref{fig:force-on-orbit}.
+      \caption{This figure decomposes the gaps shown in the middle panel of Fig.~\ref{fig:profiles} into the individual perturbers whose cause them, as identified from Fig.~\ref{fig:force-on-orbit}.
 These simulations only include the suspected globular clusters in one-at-a-time.}
@@
 Fig.~\ref{fig:decomposition} shows a comparison between one of the 50 realizations of Palomar~5 stream taking into account the gravitational interactions with all other globular clusters in the Galaxy (top panel) and omitting them (bottom panel).
-This comparison clearly shows the presence of two wide ($\sim$ 100~pc and $\sim$ 1~kpc) gaps in the leading tail and of a more subtle underdensity at $x_{\textrm{tail}}\sim 5$~kpc (we refer the reader to Appendix~\ref{sec:gap_detection} for a detailed description of the underdensities and gaps detection method).
+This comparison clearly shows the presence of two wide ($\sim$ 100~pc and $\sim$ 1~kpc) gaps in the leading tail and of a more subtle under-density at $x_{\textrm{tail}}\sim 5$~kpc (we refer the reader to Appendix~\ref{sec:gap_detection} for a detailed description of the under-densities and gaps detection method).
 To find out which globular clusters were responsible of creating these gaps, and at what time the close passages took place, we estimated the gravitational acceleration along the orbit of Pal~5 and represented it in the ($t$, $\tau$) space.
@@
 To further probe that the three clusters above are responsible of producing the gaps observed in this simulation, we conducted additional experiments by including only the perturbation of each of these three clusters at a time, and neglecting the gravitational perturbations of all the other clusters in the Galaxy.
-The results of these simulations are reported in the middle panels of Fig.~\ref{fig:decomposition} and clearly shows that NGC~2808, NGC~7078 and NGC~104 are the clusters responsible of creating the underdense regions observed in the leading tails of Palomar~5.
+The results of these simulations are reported in the middle panels of Fig.~\ref{fig:decomposition} and clearly shows that NGC~2808, NGC~7078 and NGC~104 are the clusters responsible of creating the under-dense regions observed in the leading tails of Palomar~5.
 It is worth noting that the times at which the passages of these clusters occurred, according to our analysis, are in agreement with the observed width of the corresponding gaps: the encounter with NGC~2808 being very recent (only 200~Myr ago) its induced gap is still very thin, because it takes time for a perturbation to grow into an extended gap, as it is the case for those induced by the passages of  NGC~7078 and NGC~104, which occurred in earlier times.
@@
         \vspace{0.5cm}
-      \caption{Occurence of gaps in Palomar~5 streams,  in our simulations.
+      \caption{Occurrence of gaps in Palomar~5 streams,  in our simulations.
 More specifically, the table reports the number of simulations (second row) for which a given number of gaps (first row) is found.
@@
 
-    We need a more sophisticated satistic that can be compared to other simulations and turn to the \textit{gap creation rate} developed by \citet{2012ApJ...748...20C}.
+    We need a more sophisticated statistic that can be compared to other simulations and turn to the \textit{gap creation rate} developed by \citet{2012ApJ...748...20C}.
 The gap creation rate is the number of gaps that appear per time and normalized by the length of the stream.
@@
 Since the lengths are equal, and the offset between the tails is small compared to the galactocentric distance of about 10 kpc, we expected the number of gaps in each tail to be more or less the same.
-We can compute the probability of observing the unequal occurances though the binomail distirbution.
+We can compute the probability of observing the unequal occurrences though the binomial distribution.
 First, the since the 44 gaps linked to NGC~2808 are the same fly-by, they are not independent events.
@@
 In their analysis, the perturber was modeled as a Plummer sphere, like in our simulations.
-Since the stream is not zero-dimensional and has length, both the parallel and perpendicular components of the pertuber velocity must be considered.
+Since the stream is not zero-dimensional and has length, both the parallel and perpendicular components of the perturber velocity must be considered.
 Consequently, five parameters determine the change in velocity of a given stream particle: $M$, $r_p$, $b$, $W_\parallel$, and $W_\perp$.
@@
 We also exclude the characteristic cluster radius, as it showed little correlation with the results, likely due to the narrow range of globular cluster radii (see Fig.\ref{fig:mass_size_plane}).
-This factor might be more significant for dark matter subhalos, where size variation is greater.
+This factor might be more significant for dark matter sub-halos, where size variation is greater.
 
@@
       \includegraphics[width=0.45\linewidth]{impact_stats_phase_space.png}
-      \caption{\textit{Characterisitcs of gap causing clusters} \textit{Top Left:} Energy-angular momentum space of the globular clusters in the simulations, with 50~$\times$~165 data points representing all sampled initial conditions.
+      \caption{\textit{Characteristics of gap causing clusters} \textit{Top Left:} Energy-angular momentum space of the globular clusters in the simulations, with 50~$\times$~165 data points representing all sampled initial conditions.
 Clusters impacting Palomar 5 are shown with colored markers; non-gap-causing clusters are shown in gray.
@@
 \textit{Bottom Left:} Palomar 5's orbit in the xy plane.
-The grayscale represents all 50 stacked orbits, with $\mathcal{P}$ indicating the probability of  Palomar~5's position, normalized to $\mathcal{P}_\textrm{max}=1$.
+The gray-scale represents all 50 stacked orbits, with $\mathcal{P}$ indicating the probability of  Palomar~5's position, normalized to $\mathcal{P}_\textrm{max}=1$.
 Markers show the perturber's position during each impact.
-The 50 white dots indicates Palomar~5's sampled initial conditions for the current day poisition.
+The 50 white dots indicates Palomar~5's sampled initial conditions for the current day position.
 \textit{Bottom Right:} same as bottom left but in the meridional plane.
@@
   Unfortunately, none of the predicted gaps is located in regions of the tails that have been observed, and this is especially true for the leading tail, which result shorter than the trailing tail in the observations \citep[see][for a possible explanation of this feature]{2017NatAs...1..633P}.
-Currently, the presence -- or not -- of gaps in the observed portion of Pal~5 tails is debated: \citet{2016ApJ...819....1I} found no statistically significant gap in Pal~5 tails, while \citet{2017MNRAS.470...60E}, analysing the same dataset as \citet{2016ApJ...819....1I}, suggested the presence of a few dips and gaps in the tails, at angular distances between $2^\circ$ and $9^\circ$ from the cluster center \citep[see also][]{ 2020ApJ...889...70B}.
+Currently, the presence -- or not -- of gaps in the observed portion of Pal~5 tails is debated: \citet{2016ApJ...819....1I} found no statistically significant gap in Pal~5 tails, while \citet{2017MNRAS.470...60E}, analyzing the same data-set as \citet{2016ApJ...819....1I}, suggested the presence of a few dips and gaps in the tails, at angular distances between $2^\circ$ and $9^\circ$ from the cluster center \citep[see also][]{ 2020ApJ...889...70B}.
 We cannot reproduce any of these features with clusters flybys in our simulations, since all our gaps are found at much larger angular and spatial distance from the cluster itself.
@@
 First, as \citet{2016MNRAS.457.3817S} demonstrated, the dispersion of action-frequencies in the stream plays a role.
-For instance, the frequency corresponding to the azimuthal action, $J_\phi$, is $\dot\theta_\phi = - \frac{\mathcal{\partial H}}{\partial J_\phi}$ and where $\theta_\phi$ is the angle describing the position of the particle phase space between momentum ($p_\phi$) and position ($\phi$) axes---where $\phi$ is the azimuhtal angle between the x-y axes in physical space.
+For instance, the frequency corresponding to the azimuthal action, $J_\phi$, is $\dot\theta_\phi = - \frac{\mathcal{\partial H}}{\partial J_\phi}$ and where $\theta_\phi$ is the angle describing the position of the particle phase space between momentum ($p_\phi$) and position ($\phi$) axes---where $\phi$ is the azimuthal angle between the x-y axes in physical space.
 Regions of the stream with well-separated frequencies are more susceptible to gap formation, while those with a wide frequency range (near the cluster) tend to erase the history of impacts.
@@
 
-  As already suggested in previous works which have studied the impact of baryonic structures on Palomar~5 tails \citep{2017NatAs...1..633P, 2019MNRAS.484.2009B}, it is possible that this cluster lies in a region of the phase-space which is not favorable to distinguish gaps created by dark matter subhalos from gaps created by baryonic structures, as the Galactic bar and giant molecular clouds.
+  As already suggested in previous works which have studied the impact of baryonic structures on Palomar~5 tails \citep{2017NatAs...1..633P, 2019MNRAS.484.2009B}, it is possible that this cluster lies in a region of the phase-space which is not favorable to distinguish gaps created by dark matter sub-halos from gaps created by baryonic structures, as the Galactic bar and giant molecular clouds.
 With our study, we show that close encounters with globular clusters constitute a further element that confuses a simple interpretation of the observed Palomar~5 stream gaps (if any).
@@
 It is thus probably only at larger distances from the Galactic center than those spanned by Palomar~5 orbit that the impact of dark matter sub-halos may become dominant, but it should be borne in mind that at these distances the number density of sub-halos also decreases.
-In the future, it will be interesting to apply the type of study conducted here to the whole set of clusters for which streams have been found to quantify the regime in which they are found, whether favourable to the creation of gaps from baryonic structures or not.
+In the future, it will be interesting to apply the type of study conducted here to the whole set of clusters for which streams have been found to quantify the regime in which they are found, whether favorable to the creation of gaps from baryonic structures or not.
 
@@
   The broader implications of our findings indicate that globular cluster interactions add a layer of complexity to interpreting stellar stream substructures, complicating efforts to distinguish between baryonic and dark matter-induced gaps.
-While Palomar 5's phase-space region may not be ideal for isolating the impact of dark matter subhalos, extending this analysis to other streams, particularly those at larger galactocentric distances, could provide clearer insights.
+While Palomar 5's phase-space region may not be ideal for isolating the impact of dark matter sub-halos, extending this analysis to other streams, particularly those at larger galactocentric distances, could provide clearer insights.
 
   Future work should involve a systematic comparison of gap creation rates across various Galactic potentials and extend the scope to a wider range of streams and globular clusters.
-This would help identify regimes where baryonic and dark matter-induced substructures can be disentangled, shedding light on the elusive influence of dark matter subhalos on stellar streams.
-Additionally, we can continue to add realisitic physics to the galactic enironment such as the merger of the Sagittarius dwarf galaxy and the passages of the large and small magellanic clouds.
+This would help identify regimes where baryonic and dark matter-induced substructures can be disentangled, shedding light on the elusive influence of dark matter sub-halos on stellar streams.
+Additionally, we can continue to add realistic physics to the galactic environment such as the merger of the Sagittarius dwarf galaxy and the passages of the large and small Magellanic clouds.
 
@@
 
-  The code used to produce the data in these simulations is publically available at https://github.com/salvatore-ferrone/tstrippy.
+  The code used to produce the data in these simulations is publicly available at https://github.com/salvatore-ferrone/tstrippy.
 Examples on the codes installation and use are provided at https://tstrippy.readthedocs.io.
-The simulations will also be made publically available at etidal-project.obspm.fr.
+The simulations will also be made publicly available at etidal-project.obspm.fr.
 
@@
     For each of the 50 simulations, we compare the final snapshots between the \texttt{reference} and \texttt{full} simulations using the tail coordinate system, as shown in Fig.~\ref{fig:TailCoordinates}.
-The vanilla simulations were only saved at the final timestamps, thus from Eq.\ref{eq:data_volume_estimate} where $N_ts$=1, leading to a data volumne of about one hundred messily megabytes.
+The vanilla simulations were only saved at the final timestamps, thus from Eq.\ref{eq:data_volume_estimate} where $N_ts$=1, leading to a data volume of about one hundred messily megabytes.
 We inspect these differences by generating 2D density maps of the streams.
@@
 
-    With the smoothed 1D density profiles in hand, we search for regions where the \texttt{full} simulations are significantly underdense compared to \texttt{reference} simulations, surpassing stochastic fluctuations.
+    With the smoothed 1D density profiles in hand, we search for regions where the \texttt{full} simulations are significantly under-dense compared to \texttt{reference} simulations, surpassing stochastic fluctuations.
 To accomplish this, we first impose a signal-to-noise ratio threshold, $\mathcal{SNR}$.
@@
 If the differences between the density profiles are primarily due to stochastic processes of similar magnitude, this distribution should resemble a Gaussian as expected from the central limit theorem.
-Thus, we flag all regions where the density is underdense by more than two standard deviations, which should highlight regions whose underdensity is unlikely to be the result of the sum of stochastic processess but rather the passage of another globular cluster.
+Thus, we flag all regions where the density is under-dense by more than two standard deviations, which should highlight regions whose under-density is unlikely to be the result of the sum of stochastic processes but rather the passage of another globular cluster.
 
@@
 This method helps particularly with large subtle gaps that the eye does not notice in the 2D maps.
-These profiles are included in the Appenix~\ref{sec:gallery_of_gaps}.
+These profiles are included in the Appendix~\ref{sec:gallery_of_gaps}.
 
@@
 The bottom panel shows the evolution of the stream density over time.
-To avoid extreme low-density regions at the stream's edges, we applied the same density threshold as from eq.~\ref{eq:density_threshold} to focus on the more significant areas of the stream.
+To avoid extreme low-density regions at the stream's edges, we applied the same density threshold as from Eq.~\ref{eq:density_threshold} to focus on the more significant areas of the stream.
 Next, we modeled Palomar~5's orbit as a proxy for its stream and sampled points along the orbit to measure the gravitational force exerted by other globular clusters.
@@
 This orbital segment is used to approximate the stream.
-Then, the gravitational force from all other clusers was computed on the orbit.
+Then, the gravitational force from all other clusters was computed on the orbit.
 This is shown in the top plot, where the gravitational force is in measured in acceleration and is given in integration units: km$^2$~kpc$^{-1}$~s$^{-2}$.
@@
     Fig.~\ref{fig:force-on-orbit} contains some interesting information.
-Notice the periodic ribbions of force in the $(t,\tau)$ plane.
+Notice the periodic ribbons of force in the $(t,\tau)$ plane.
 This is due to pericenter passages where Palomar~5 is getting closer to the center of mass of the globular cluster system.
-Additionally, for the impacts of NGC~104 and NG~C7078, wakes can be observed in the density map.
+Additionally, for the impacts of NGC~104 and NGC~C7078, wakes can be observed in the density map.
 Another important aspect, is that the strongest peak in gravitational force does not necessarily create a gap.
@@
 For a handful of simulations, to double check that the verdict made correctly when passing from suspect to culprit, we re-compute the simulations yet individually adding one globular cluster at a time.
-As a result, we are able to confirm that singular gaps arrise from the suspected clusters, an example of which is shown in Fig.~\ref{fig:decomposition}.
+As a result, we are able to confirm that singular gaps arise from the suspected clusters, an example of which is shown in Fig.~\ref{fig:decomposition}.
 
@@
     To achieve this, we begin by identifying the approximate moments of impact from the most significant clusters, as determined in the previous analysis in Sec.~\ref{sec:Perturber_Identification}.
-Then, we refine these estimates in oredr to to pinpoint both the exact location of the impact along the stream and the precise moment it occurred.
+Then, we refine these estimates in order to to pinpoint both the exact location of the impact along the stream and the precise moment it occurred.
 To do so, we fit a third-order parametric polynomial to the stream, using the saved snapshots from our simulations:
@@
 
-      Since the simulation snapshots were saved at a temporal resolution of 1 Myr--—rather than at the integration timestep, which would have generated an excessive amount of data—--we interpolate between snapshots to more precisely estimate the impact geometry.
-This is achieved by fitting the stream at five timesteps surrounding the approximate impact time, a period of 5 Myr, which sufficiently covers the interaction time.
+      Since the simulation snapshots were saved at a temporal resolution of 1 Myr--—rather than at the integration time-step, which would have generated an excessive amount of data—--we interpolate between snapshots to more precisely estimate the impact geometry.
+This is achieved by fitting the stream at five time-steps surrounding the approximate impact time, a period of 5 Myr, which sufficiently covers the interaction time.
 The interaction time can be estimated as $t \approx \frac{100~\text{pc}}{300~\frac{\text{km}}{\text{s}}} \approx 0.3~\text{Myr}$.
@@
       where $\vec{v}_p$ and $\vec{v}_s$ are the velocities of the perturber and the stream, respectively.
-For each of the 50 simulations, this information was compute for the strongs 5 flybys of a perturber with the stream.
+For each of the 50 simulations, this information was compute for the strongest 5 flybys of a perturber with the stream.
 Thus, we compute a sample of 250 impacts, and flag those that give way to gaps.

### 2024-12-08 paper/main.tex
added histogram of "gaps within sight"
moved comment about no gap near GC

@@
 
-\section{Discussion}
+  \subsection{Comparison to observations}
 
-  The simulations presented in this paper suggest that, in the last 5 Gyr of evolution, Palomar~5's stream could have experienced multiple close encounters with other Galactic globular clusters, some of which are able to create gaps -- even a few kpc wide -- in its tails.
+    We briefly compare our simulated gaps to the literature on Paloamr~5.
+In the bottom panel of Fig.~\ref{fig:stream_on_sky}, we compare the tracks of Palomar~5 that were compiled in Galstreams by Fig.~\ref{2023MNRAS.520.5225M}.
+The mid-point positions of the streams do not have the same galacto-centric positions as Palomar~5 from  Baumgardt's catalog and are thus offset when projected into tail coordinates.
+Nonetheless, we note that the lengths of the streams are at max 6~kpc on either side of the globular cluster.
+We apply a threshold of 5.75~kpc on either side to consider the gaps that could occur in currently observed regions of Paloamr~5.
+This is presented in Fig.~\ref{fig:GapsWithinSight}.
+      \begin{figure}
+        \centering
+        \includegraphics[width=\linewidth]{GapWithinSight.png}
+        \caption{The distribution of gaps in absolute distance from the center of mass of Palomar~5's globular cluster.
+A threshold value of $\sim$~5.75~kpc was selected as distance threshold that corresponds to the current observed length.
+Distances beyond this are shaded.
+}
+        \label{fig:GapsWithinSight}
+      \end{figure}
 
-  Unfortunately, none of the predicted gaps is located in regions of the tails that have been observed, and this is especially true for the leading tail, which result shorter than the trailing tail in the observations \citep[see][for a possible explanation of this feature]{2017NatAs...1..633P}.
-Currently, the presence -- or not -- of gaps in the observed portion of Pal~5 tails is debated: \citet{2016ApJ...819....1I} found no statistically significant gap in Pal~5 tails, while \citet{2017MNRAS.470...60E}, analyzing the same data-set as \citet{2016ApJ...819....1I}, suggested the presence of a few dips and gaps in the tails, at angular distances between $2^\circ$ and $9^\circ$ from the cluster center \citep[see also][]{ 2020ApJ...889...70B}.
-We cannot reproduce any of these features with clusters flybys in our simulations, since all our gaps are found at much larger angular and spatial distance from the cluster itself.
+    We observe that there are no gaps near within 3~kpc to the cluster within our simulation.
 There may be a few reasons for the absence of gaps in portion of the tails closer to the cluster center.
@@
 Thus, the strong flybys that occurred close to the cluster were inconsequential for gap formation.
+
     Another possible explanation is the different thickness of the simulated tails at different distances from the center of the cluster.
@@
 We have assumed for Pal~5, 5 Gyr years ago, the same internal parameters (mass and size) that the cluster has today.
-With such parameters, after 5 Gyr of evolution, our N-body system has lost most of its mass, and therefore the part of the tails closest to the cluster itself, whose density depends mainly on the most recent mass loss \citep[see, for example, Fig.~A.3 in][]{2012A&A...546L...7M}, is necessarily very thin because the simulated cluster has essentially no more mass to lose.
+With such parameters, after 5 Gyr of evolution, our system has lost most of its mass, and therefore the part of the tails closest to the cluster itself, whose density depends mainly on the most recent mass loss \citep[see, for example, Fig.~A.3 in][]{2012A&A...546L...7M}, is necessarily very thin because the simulated cluster has essentially no more mass to lose.
 This last point has also a consequence in the gap creation rate, which we derived at the end of Sect.~\ref{sect:geometry}: with only 70\% of the tail  ($\sim$ 14~kpc over 20~kpc, excluding the innermost $\pm 3$~kpc from the cluster center) really suitable for forming gaps, the gap creation rates are about 50$\%$ higher than those estimated in the aforementioned section, where values have been derived taking into account the full tail extent.
@@
 If true, this rate would be too high to be explained by Galactic globular clusters only.
+
+\section{Discussion}
+
+  The simulations presented in this paper suggest that, in the last 5~Gyr of evolution, Palomar~5's stream could have experienced multiple close encounters with other Galactic globular clusters, some of which are able to create gaps -- even a few kpc wide -- in its tails.
+
+  Unfortunately, none of the predicted gaps is located in regions of the tails that have been observed, and this is especially true for the leading tail, which result shorter than the trailing tail in the observations \citep[see][for a possible explanation of this feature]{2017NatAs...1..633P}.
+\textbf{NOT TRUE.
+Good thing I checked.}
+
+  Currently, the presence -- or not -- of gaps in the observed portion of Pal~5 tails is debated: \citet{2016ApJ...819....1I} found no statistically significant gap in Pal~5 tails, while \citet{2017MNRAS.470...60E}, analyzing the same data-set as \citet{2016ApJ...819....1I}, suggested the presence of a few dips and gaps in the tails, at angular distances between $2^\circ$ and $9^\circ$ from the cluster center \citep[see also][]{ 2020ApJ...889...70B}.
+
+  We cannot reproduce any of these features with clusters flybys in our simulations, since all our gaps are found at much larger angular and spatial distance from the cluster itself.
 

### 2024-12-10 paper/main.tex
grammer and consistency edits

@@
 
-  Beyond the global (visible and dark) mass distribution, as we anticipated earlier in this Introduction, streams can be used also to infer the granularity of the dark matter, that is the mass and density of the sub-halos populating our Galaxy.
-According to simulations by \citet{2008MNRAS.391.1685S}, $\Lambda$-CDM predicts that galaxies grow hierarchically, with dark matter clumps forming at a wide range of masses and sizes.
+  Beyond the global visible and dark mass distribution streams can be used also to infer the granularity of the dark matter, that is the mass and density of the sub-halos populating our Galaxy.
+According to simulations by \citet{2008MNRAS.391.1685S}, $\Lambda$~Cold Dark Matter ($\Lambda$~CDM) predicts that galaxies grow hierarchically, with dark matter clumps forming at a wide range of masses and sizes.
 These clumps, or \textit{sub-halos}, are predicted to follow a mass distribution with a power-law slope slightly shallower than -2.0.
@@
 However, To balance data volume and computation time, we use only one galactic potential, which should suffice.
-\citet{2021MNRAS.505.5978V} found that only a few outer globular clusters are strongly affected by different potential models, with kinematic uncertainties being the dominant factor in orbital solution differences.
-Similarly, \citet{2024MNRAS.528.5189G} generated a globular cluster mass-loss catalog using seven different potential models and found their debris distributions to be model-independent, similarly to \citet{2023A&A...673A..44F}.
+\citet{2021MNRAS.505.5978V} found that only a few outer globular clusters are strongly affected by different potential models.
+Generally, kinematic uncertainties are the dominant factor in orbital solution differences per cluster.
+Similarly, \citet{2024MNRAS.528.5189G} generated a globular cluster mass-loss catalog using seven different potential models and found their debris distributions to be rather model-independent, similarly to \citet{2023A&A...673A..44F}.
 While the clusters' positions may depend on the model, we assert that interaction rates and stream formation are unaffected by the choice of galactic potential model.
@@
 To quantify these interactions, all clusters are represented by Plummer spheres, each with mass and half-mass radius as reported in the Baumgardt catalog \citep{2021MNRAS.505.5957B}.
-In the following of this paper, we refer to these simulations, where the globular clusters interactions are taken into account as \texttt{full} simulations.
+For the remainder of this paper, we refer to these simulations, where the globular clusters interactions are taken into account as \texttt{full} simulations.
 \\
@@
       \begin{enumerate}
-        \item First, starting from the Galactocentric positions and velocities of all 165 Galactic globular clusters, we integrate their orbits back in time for 5~Gyr under the influence of the Galaxy itself, and their mutual influence.
-In the backward integration, the equations of motion for the globular cluster system are thus:
+        \item First, starting from the Galactocentric positions and velocities of all 165 Galactic globular clusters, we integrate their orbits back in time for 5~Gyr under the influence of the Galaxy itself and their mutual influence.
+In the backward integration, the system of equations of motion for the globular clusters are thus:
         \begin{equation}
@@
         \end{equation} \label{eq:equation_of_motion_particle} where the index $p$ represents one of the 100,000 particles of interest, $\vec{r_p}$ being its position, and $j$ indexes over the globular clusters as in Eq.~\ref{eq:GCNBody}.
-Note in that Eq.~\ref{eq:equation_of_motion_particle} the positions of the globular clusters are time-dependent since they are being loaded during this step and not computed, unlike Eq.~\ref{eq:equation_of_motion_particle}.
+Note in that Eq.~\ref{eq:equation_of_motion_particle} the positions of the globular clusters are time-dependent since they are being loaded during this step and not computed, unlike Eq.~\ref{eq:GCNBody}.
 
@@
 In Eq.~\ref{eq:GCNBody} only the gradient of the Galactic potential  is considered.
-In Eq.~\ref{eq:equation_of_motion_particle}, of the second term on the right-side of the equation, only the interaction of Pal~5 particles with Pal~5 itself are considered.
+In Eq.~\ref{eq:equation_of_motion_particle}, of the second term on the right-side of the equation, only the influence of Pal~5's Plummer sphere on Pal~5's particles is considered.
 In other words, sum iterates over only one globular cluster, the host.
@@
 
-    To our knowledge, this is the first study to investigate the gravitational interactions between all Galactic globular clusters.
-While we did not  observe significant changes to orbital properties of the clusters, when subject to their mutual interactions -- a result which is  expected given the small sizes and vast distances between the clusters -- we do see clear effects of the presence of their gravitational field in the properties of Pal~5 stream.
+      % To our knowledge, this is the first study to investigate the gravitational interactions between all Galactic globular clusters.
+While we did not  observe significant changes to orbital properties of the clusters, when subject to their mutual interactions -- a result which is  expected given the small sizes and vast distances between the clusters -- we do
+    % This is no longer true after Sergey showed me the papers from Ischenko et al for the globular cluster ineractions
+
+    We see clear effects of the presence of their gravitational field in the properties of Pal~5 stream.
 A particularly clear example of this effect is shown in Fig.~\ref{fig:stream_on_sky} which we selected for its prominent gaps.
@@
 The density peaks in the stream are of the same order as the cluster itself.
-This is inconsistent with reality, the cluster should be much more populated than the stream's peak density.
+This is inconsistent with reality, the cluster's peak density should be higher than the stream's peak density.
 Of course, this discrepancy is a result of our modeling choices.
 We use the present day mass and radius for Palomar~5 for the whole duration of the simulation.
-A consequence of this is that the system is modeled as being less dense than it should have been.
+Consequently the system is modeled as being less dense than it should have been.
 In turn, our simulations have a strong initial mass which adds to the amplitudes of the profile density peaks of Fig.~\ref{fig:stream_on_sky}.
@@
 The definition of this coordinate system is based on the work of \citet{2004AJ....127.2753D} and is shown in Fig.~\ref{fig:TailCoordinates}.
-Briefly, in this system the $x_{\textrm{tail}}$ coordinate represents the position of a particle along the orbit relative to the globular cluster—whether it is ahead (positive $x_{\textrm{tail}}$) or behind (negative $x_{\textrm{tail}}$) the cluster;  the $y_{\textrm{tail}}$ coordinate measures the particle's distance within the orbital plane, where positive values indicate that the particle is farther from the Galactic center and negative values indicate that it is closer.
+Briefly, in this system the $x_{\textrm{tail}}$ coordinate represents the position of a particle along the orbit relative to the globular cluster.
+Positive values of $x_{\textrm{tail}}$ are ahead of the cluster while negative $x_{\textrm{tail}}$ is behind the cluster.
+The $y_{\textrm{tail}}$ coordinate measures the particle's distance within the orbital plane, where positive values indicate that the particle is farther from the Galactic center and negative values indicate that it is closer.
 Fig.~\ref{fig:decomposition} shows a comparison between one of the 50 realizations of Palomar~5 stream taking into account the gravitational interactions with all other globular clusters in the Galaxy (top panel) and omitting them (bottom panel).
 This comparison clearly shows the presence of two wide ($\sim$ 100~pc and $\sim$ 1~kpc) gaps in the leading tail and of a more subtle under-density at $x_{\textrm{tail}}\sim 5$~kpc (we refer the reader to Appendix~\ref{sec:gap_detection} for a detailed description of the under-densities and gaps detection method).
+
     To find out which globular clusters were responsible of creating these gaps, and at what time the close passages took place, we estimated the gravitational acceleration along the orbit of Pal~5 and represented it in the ($t$, $\tau$) space.
@@
 The use of $\tau$ is advantageous because the growth of the stream is approximately linear in $\tau$.
-On the other hand, in physical space streams on eccentric orbits indeed grow in time yet the growth rate is modulated with periodic expansion and contraction depending on the orbital phase.
-Adopting this time space, and reporting the gravitational acceleration along Pal~5 orbit in this space, the identification of the globular clusters which produced the perturbation, and the time at which it occurred becomes straight forward.
+On the other hand, in physical space streams on eccentric orbits indeed grow in time yet the growth rate is modulated with periodic expansion and contraction depending on the orbital phase (see the top panel of Fig.~5 from \citet{2016MNRAS.457.3817S} for an example).
+Adopting this time space, and reporting the gravitational acceleration along Palomar~5 orbit in this space, the identification of the globular clusters which produced the perturbation, and the time at which it occurred becomes straight forward.
 We refer the reader to Appendix~\ref{sec:Perturber_Identification} for an in-depth description of the procedure.
 In this way, we could identify that the clusters responsible for creating gaps in the simulation of Palomar~5 stream reported in Fig.\ref{fig:decomposition} are NGC~2808, NGC~7078 and NGC~104, and that their close passages occurred 200~Myr, -1.9 and -2.1~Gyr ago, respectively.
+
     To further probe that the three clusters above are responsible of producing the gaps observed in this simulation, we conducted additional experiments by including only the perturbation of each of these three clusters at a time, and neglecting the gravitational perturbations of all the other clusters in the Galaxy.
-The results of these simulations are reported in the middle panels of Fig.~\ref{fig:decomposition} and clearly shows that NGC~2808, NGC~7078 and NGC~104 are the clusters responsible of creating the under-dense regions observed in the leading tails of Palomar~5.
+The results of these simulations are reported in the middle panels of Fig.~\ref{fig:decomposition} and clearly shows that NGC~2808, NGC~7078 and NGC~104 are the clusters responsible for creating the under-dense regions observed in the leading tails of Palomar~5.
 It is worth noting that the times at which the passages of these clusters occurred, according to our analysis, are in agreement with the observed width of the corresponding gaps: the encounter with NGC~2808 being very recent (only 200~Myr ago) its induced gap is still very thin, because it takes time for a perturbation to grow into an extended gap, as it is the case for those induced by the passages of  NGC~7078 and NGC~104, which occurred in earlier times.

### 2024-12-10 paper/main.tex
death to vanilla, more small edits to results

@@
 NGC~2808 impacts Palomar 5's stream about 200 Myr ago in 44 out of 50 simulations, creating a small gap at a similar position to the one reported in Fig.~\ref{fig:decomposition} in each case.
-Since this interaction occurred less than one orbital period ago, the uncertainties lead to orbital solutions remain that similar and thus producing consistent results across all simulations.
+Since this interaction occurred less than one orbital period ago, the uncertainties that lead to orbital solutions remain that similar and thus producing consistent results across all simulations.
 However, as we continue to turn back time further, the uncertainties in the initial conditions allow the for the various orbital solutions to diverge from one another.
@@
     This computation allows us to analyze the distribution of gap creation rates across all simulations.
-Notice that since the gaps \textit{add in parallel}, naturally gaps that occur at earlier times when the stream was shorter are weighted higher than those that occur when the stream is longer.
+Notice that since the gap creation rate \textit{adds in parallel}, naturally gaps that occur at earlier times when the stream was shorter are weighted higher than those that occur when the stream is longer.
 The results are shown in Fig.~\ref{fig:gapcreationrate}.
@@
       \includegraphics[width=\linewidth]{impact_geometry_statistics.png}
-      \caption{The distribution and relationship between the basic impact variables from Eq.\ref{eq:change_in_momentum} and the close encounter that cause impacts, shown in red against those that do not, shown in blue.
-}
+      \caption{The distribution and relationship between the impact variables from Eq.\ref{eq:change_in_momentum} for all close fly-bys considered.
+The encounteres that cause gaps are labeled with the same colors as from Fig.~
+      \ref{fig:histogram_impact_time} in the scatter plot, and binned in red for the histograms. The close encounters that did not lead to gaps are shown in grey. }
       \label{fig:impact_geometry_statistics}    
@@
     This equation asserts that a more massive perturber, passing closer to the particle, and moving more slowly, will have a greater impact.
-It is important to note that the momentum change is inversely proportional to the velocity of the perturber, which contrasts with the intuition from elastic collisions, such as those between billiard balls, where higher velocities result in greater impacts.
+It is important to note that the momentum change is inversely proportional to the velocity of the perturber.
+Note that this contrasts with the intuition from elastic collisions such as those between billiard balls where higher velocities result in greater impacts.
 
-    \citet{2015MNRAS.450.1136E} extended this impact theory from one point mass impacting another point mass, to studying how an extended body impacts a stream.
-This quantifies the change in momentum of a given particle as a function of its distance from the point of greatest impact.
+    \citet{2015MNRAS.450.1136E} extended this impact theory from one point mass impacting another to studying how an extended body impacts a stream.
+This quantifies the change in momentum of a given particle as a function of its distance from the point of greatest impact along the stream.
 In their analysis, the perturber was modeled as a Plummer sphere, like in our simulations.
@@
 
-    While Fig.~\ref{fig:impact_geometry_statistics} demonstrates that mass, relative velocity, or impact parameter alone cannot predict gap formation, one interesting result emerges: no gaps are created from interactions with impact parameters greater than 200~pc.
+    While Fig.~\ref{fig:impact_geometry_statistics} demonstrates that mass, relative velocity, or impact parameter alone cannot predict gap formation, one interesting result emerges: no gaps are created from interactions with impact parameters greater than 300~pc.
 The stream widths are roughly 200~pc.
@@
 The circular velocity-curve is drawn with a light-pink dashed curve.
-\textit{Top Right:} Mass-size plane of the globular clusters in the simulations, with uncertainties on the masses indicated with vertical bars.
-gap-causing clusters are indicated with colored points, whereas Pal~5 is indicated by a white symbol with a black edge.
-\textit{Bottom Left:} Palomar 5's orbit in the xy plane.
+\textit{Top Right:} Mass-size plane of the globular clusters in the simulations, with uncertainties on the masses indicated as vertical lines.
+We recall the reader that the globular cluster catalog at this time does not provide uncertainties for clusters' characteristic radii.
+\textit{Bottom Left:} Palomar 5's orbit in the Galactocentric xy plane.
 The gray-scale represents all 50 stacked orbits, with $\mathcal{P}$ indicating the probability of  Palomar~5's position, normalized to $\mathcal{P}_\textrm{max}=1$.
@@
 The 50 white dots indicates Palomar~5's sampled initial conditions for the current day position.
+The colored markers indicate where the pertuber was located when it impact the stream and \textit{not} its present day position.
 \textit{Bottom Right:} same as bottom left but in the meridional plane.
@@
 On average, encounters which lead to gap creations impart a change in momentum on stream particles which is a factor of 10 higher than that of encounters which do not form gaps (but with some overlap in the low-velocity tail).
-Interestingly, changes in momentum which lead to gap creations extend over a large range in velocities (there is a factor of about 100 between the smallest and largest changes), with NGC~2808 (orange color in the histogram) imparting changes in the velocity of stream particles which redistribute over the whole range of $\Delta p$.
+Interestingly, changes in momentum which lead to gap creations extend over a large range in velocities.
+There is a factor of about 100 between the smallest and largest changes with NGC~2808 (orange color in the histogram) imparting changes in the velocity of stream particles which redistribute over the whole range of $\Delta p$.
 
     In addition to characterizing the parameters governing cluster encounters with the stream, since we know which clusters have produced gaps on the tail of Palomar~5, we can also verify their orbital and structural properties.
-This is what is shown in Fig~\ref{fig:mass_size_plane}, where we first show the mass and size (i.e.
-half-mass radius) of the clusters that cause gaps on the tails of Pal5, dividing them from those that do not.
+This is shown in Fig~\ref{fig:mass_size_plane}, where we first show the mass and size (i.e.
+half-mass radius) of the clusters that cause gaps on the tails of Pal~5, dividing them from those that do not.
 As can be seen, no cluster with mass below $10^5 \rm{M}_\odot$ causes gaps on Pal~5 stream, and all perturbers, except FSR~1758 have a half-mass radius below 10~pc.
@@
 This is because only clusters within Palomar 5's orbital space can interact with the stream: clusters with higher orbital energies tend to have larger apocenter than that of Pal~5 and thus spend most of their time away from Palomar 5's orbital volume.
+
     Finally, it is worth commenting on the location of impacts with the stream, whether they occur when Pal~5 is close to its pericenter or not.
-As it can be seen from Fig.\ref{fig:mass_size_plane} (bottom panel) encounters can occur at all orbital phases of Pal~5, when it is close to pericenter, but also very far away from it, at the outskirts of its orbital space.
+As it can be seen from Fig.~\ref{fig:mass_size_plane} (bottom panel) encounters can occur at all orbital phases of Pal~5, when it is close to pericenter, but also very far away from it, at the outskirts of its orbital space.
 However, when taken all together, the location of gap creating impacts shows a strong negative correlation with the galactocentric radius $r$, with the number $N$ of encounters favorable to gap formation  going as $N = -2.5r + 50$ (with a Pearson coefficient of -0.86).
@@
     For each of the 50 simulations, we compare the final snapshots between the \texttt{reference} and \texttt{full} simulations using the tail coordinate system, as shown in Fig.~\ref{fig:TailCoordinates}.
-The vanilla simulations were only saved at the final timestamps, thus from Eq.\ref{eq:data_volume_estimate} where $N_ts$=1, leading to a data volume of about one hundred messily megabytes.
+The \texttt{reference} simulations were only saved at the final timestamps, thus from Eq.\ref{eq:data_volume_estimate} where $N_ts$=1, leading to a data volume of about one hundred messily megabytes.
 We inspect these differences by generating 2D density maps of the streams.
@@
       \caption{A comparison between the density maps and profiles of the \texttt{full} and \texttt{reference} simulations is presented.
-The locations where the full simulation is less dense than the vanilla simulation by 2-$\sigma$ is highlighted by blue vertical bars.
+The locations where the \texttt{full} simulation is less dense than the \texttt{reference} simulation by 2-$\sigma$ is highlighted by blue vertical bars.
 Bins below the noise threshold are not considered when measuring the differences between the two 1D profiles.
@@
 To accomplish this, we first impose a signal-to-noise ratio threshold, $\mathcal{SNR}$.
-The signal is defined as the log of the counts per bin from the \texttt{vanilla} 1D density profile, and errors are propagated assuming a Poisson distribution.
-We then compute a threshold for the number of counts in the \texttt{vanilla} simulation, $N_v$, using the transcendental equation:
+The signal is defined as the log of the counts per bin from the \texttt{reference} 1D density profile, and errors are propagated assuming a Poisson distribution.
+We then compute a threshold for the number of counts in the \texttt{reference} simulation, $N_v$, using the transcendental equation:
     \begin{equation}
@@
     By setting $\mathcal{SNR} = 5$, we solve for $N$ using \texttt{scipy.optimize.fsolve}, finding that $N$ must be greater than 7.
-After discarding insignificant bins (i.e., those with counts below the threshold), we compute the log ratio of the counts between the \texttt{vanilla} and full simulations:
+After discarding insignificant bins (i.e., those with counts below the threshold), we compute the log ratio of the counts between the \texttt{reference} and \texttt{full} simulations:
     \begin{equation}
@@
       \end{equation}
-    where $\mathcal{R}_i$ is the log ratio, $N_{f,i}$ are the counts from the full simulation, and $N_{v,i}$ are the counts from the \texttt{vanilla} simulation for each bin $i$.
+    where $\mathcal{R}_i$ is the log ratio, $N_{f,i}$ are the counts from the \texttt{full} simulation, and $N_{v,i}$ are the counts from the \texttt{reference} simulation for each bin $i$.
 We then analyze the distribution of $\mathcal{R}_i$.

### 2024-12-10 paper/main.tex
trying to contextualize bovik and bovy

@@
   The simulations presented in this paper suggest that, in the last 5~Gyr of evolution, Palomar~5's stream could have experienced multiple close encounters with other Galactic globular clusters, some of which are able to create gaps -- even a few kpc wide -- in its tails.
-
-  Unfortunately, none of the predicted gaps is located in regions of the tails that have been observed, and this is especially true for the leading tail, which result shorter than the trailing tail in the observations \citep[see][for a possible explanation of this feature]{2017NatAs...1..633P}.
-\textbf{NOT TRUE.
-Good thing I checked.}
-
 Currently, the presence -- or not -- of gaps in the observed portion of Pal~5 tails is debated: \citet{2016ApJ...819....1I} found no statistically significant gap in Pal~5 tails, while \citet{2017MNRAS.470...60E}, analyzing the same data-set as \citet{2016ApJ...819....1I}, suggested the presence of a few dips and gaps in the tails, at angular distances between $2^\circ$ and $9^\circ$ from the cluster center \citep[see also][]{ 2020ApJ...889...70B}.
-
-  We cannot reproduce any of these features with clusters flybys in our simulations, since all our gaps are found at much larger angular and spatial distance from the cluster itself.
+While our simulations produced 73 gaps across 50 realizations, only about 22 are beyond the current length of the observed portion of the stream.
+This means on average, we obtain at least one gap from a globular cluster within the past 5~Gyr.
+However, we note that we made no atttempt of creating mock observations to compare a full detection processes of gaps by disentangling field stars from stream stars.
+This is interesting work that is worth pursuing yet beyond the current scope of this work.
+The number of gaps present here is certainly an upper limit of detectable gaps.
 

### 2024-12-11 paper/main.tex
updated conversation about Banik 2017

@@
     We briefly compare our simulated gaps to the literature on Paloamr~5.
-In the bottom panel of Fig.~\ref{fig:stream_on_sky}, we compare the tracks of Palomar~5 that were compiled in Galstreams by Fig.~\ref{2023MNRAS.520.5225M}.
+In the bottom panel of Fig.~\ref{fig:stream_on_sky}, we compare the tracks of Palomar~5 that were compiled in Galstreams by \citet{2023MNRAS.520.5225M}.
 The mid-point positions of the streams do not have the same galacto-centric positions as Palomar~5 from  Baumgardt's catalog and are thus offset when projected into tail coordinates.
@@
 
-  We note that our results seem to be in tension with the conclusions of \citet{2019MNRAS.484.2009B}, who presented a numerical study of Palomar~5 tails, orbiting a Milky Way-like potential, where both dark matter sub-halos and baryonic sub-structures (Galactic bar, spiral arms, giant molecular clouds, globular clusters) were taken into account to quantify the importance of these latter in density variations in Pal~5 streams.
-From their study, they derived that the impact of globular clusters was marginal in gap creation.
-It is difficult to give an exact explanation of the reasons for this difference, because the approach followed by  \citet{2019MNRAS.484.2009B} is extremely similar to ours.
-We would point out, however, that the Galactic potential in which the integration of the orbit of Pal5 and all other galactic clusters was made in their study is the \texttt{MWPotential2014} \citep{2015ApJS..216...29B}.
-As discussed by \citet{2022ApJ...926..107M}, this potential has a halo of dark matter that is probably too light, because 20\% of the Galactic globular clusters turn out to have unbound orbits in it (note that this is not the case for the potential we have adopted, see e.g.
-Fig.~\ref{fig:mass_size_plane}).
-Even for clusters that turn out to be bound in the  \texttt{MWPotential2014} potential, the fact that the dark matter halo is (possibly) too light implies that the orbital volume occupied by the clusters is larger, thus reducing the probability of close impacts with the Pal~5 stream.
-We tentatively propose this as an explanation of the differences found, but a more systematic comparison between \citet{2019MNRAS.484.2009B}'s results and ours should be made to understand the origin of this difference.
+  We note that our results seem to be in tension with the conclusions of \citet{2019MNRAS.484.2009B}, who presented a numerical study of Palomar5 tails, orbiting a Milky Way-like potential, where both dark matter sub-halos and baryonic sub-structures (Galactic bar, spiral arms, giant molecular clouds, globular clusters) were taken into account to quantify the importance of these latter in density variations in Pal5 streams.
+While their methods are extremely similar to ours, their analysis diverges significantly.
+Specifically, \citet{2019MNRAS.484.2009B} focused on examining power spectra, analyzing variations in the one-dimensional stream density in stellar counts along the length of the stream.
+
+  Upon comparison with our simulations, we observe that the power spectra from our \texttt{full} simulations that include globular clusters and the \texttt{reference} simulations that do not, are not significantly different.
+We suggest that this may be due to the signal from only one or two gaps caused by globular clusters not being sufficient to produce notable differences in the overall power spectrum.
+Additionally, \citet{2022ApJ...926..107M} may not have inspected the profiles for individual gaps, which could explain why these features were not reported in their study.
+Further investigations would be needed to confirm this interpretation and to thoroughly assess the impact of globular clusters on stream density profiles.
 \\

### 2024-12-12 paper/main.tex
Figure captions for gallery and decomposition

@@
       \includegraphics[width=\linewidth]{decomposition-monte-carlo-009-with-3-gaps.png}
-      \caption{This figure decomposes the gaps shown in the middle panel of Fig.~\ref{fig:profiles} into the individual perturbers whose cause them, as identified from Fig.~\ref{fig:force-on-orbit}.
-These simulations only include the suspected globular clusters in one-at-a-time.}
+      \caption{Density maps of Palomar~5's stream in the tail coordinate system.
+The colorscale represents normalized particle counts (total: 100,000).
+The top panel shows the \texttt{full} simulation with three gaps on the stream's right-hand side.
+The next three panels depict simulations with identical initial conditions but exclude the gravitational influence of all clusters except those forming a given gap.
+The \texttt{Reference} simulation omits other globular clusters.
+The bottom panel compares Palomar5's observed stream length to the \texttt{Reference} simulation, using the same Monte Carlo realization as Fig.\ref{fig:stream_on_sky} and \texttt{Sampling~009} of Fig.~\ref{fig:gallery2}.}
       \label{fig:decomposition}
@@
   \section{Gallery of Gaps} \label{sec:gallery_of_gaps}
+    Figures~\ref{fig:gallery0}-\ref{fig:gallery11} include all of the realizations of the \texttt{full} simulations of Palomar~5 in tail coordinates.
+If a gap is present, it is indicated by a red arrow with the responsible globular cluster.
+For the gaps that are not easily visible, we show the 1D profiles  showing the significant underdensities compared to the \texttt{reference} simulations.
+For instance, the second and last panels of Fig.~\ref{fig:gallery0}
 
@@
       \includegraphics[width=\linewidth]{tau-profile-monte-carlo-002.png}
-      \caption{Gap Gallery}
+      \caption{First gap gallery. See the text in Appendix~\ref{sec:gallery_of_gaps} for a description.}
       \label{fig:gallery0}
@@
       \includegraphics[width=\linewidth]{tau-profile-monte-carlo-006.png}
-      \caption{Gap Gallery}
+      \caption{Second gap gallery}
       \label{fig:gallery1}
@@
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-011.png}
-      \caption{Gap Gallery}
+      \caption{Third gap gallery}
       \label{fig:gallery2}
@@
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-016.png}
-      \caption{Gap Gallery}
+      \caption{Fourth gap gallery}
       \label{fig:gallery3}
@@
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-021.png}
-      \caption{Gap Gallery}
+      \caption{Fifth gap gallery}
       \label{fig:gallery4}
@@
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-025.png}
-      \caption{Gap Gallery}
+      \caption{Sixth gap gallery}
       \label{fig:gallery5}
@@
       \includegraphics[width=\linewidth]{tau-profile-monte-carlo-028.png}
-      \caption{Gap Gallery}
+      \caption{Seventh gap gallery}
       \label{fig:gallery6}
@@
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-031.png}
-      \caption{Gap Gallery}
+      \caption{Eigth gap gallery}
       \label{fig:gallery7}
@@
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-036.png}      
-      \caption{Gap Gallery}
+      \caption{Ninth gap gallery}
       \label{fig:gallery8}
@@
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-041.png}      
-      \caption{Gap Gallery}
+      \caption{Tenth gap gallery.}
       \label{fig:gallery9}
@@
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-045.png}
-      \caption{Gap Gallery}
+      \caption{Eleventh gap gallery}
       \label{fig:gallery010}
@@
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-049.png}
-      \caption{Gap Gallery}
+      \caption{Twelfth gap gallery}
       \label{fig:gallery11}

### 2024-12-12 paper/main.tex
added discussion about Pearson's mock catalog and GC dissolution

@@
 This means on average, we obtain at least one gap from a globular cluster within the past 5~Gyr.
-However, we note that we made no atttempt of creating mock observations to compare a full detection processes of gaps by disentangling field stars from stream stars.
-This is interesting work that is worth pursuing yet beyond the current scope of this work.
-The number of gaps present here is certainly an upper limit of detectable gaps.
+However, we did not attempt to create mock observations or simulate a full detection process accounting for the challenges of disentangling field stars from stream stars.
+While such an analysis would be valuable, it is beyond the scope of this study.
+
+  Our simulations produce a stream for Palomar~5 that is longer than the currently observed extent.
+In simulations, stream detections are straightforward because we can use reference runs to clearly separate stream particles from the field and compare against a known "true" structure.
+In contrast, observational data are inherently more challenging due to magnitude limits, contamination from field stars, and the lack of a ground truth for comparison.
+These factors suggest that the number of gaps identified in this study could represent an upper limit.
+
+  However, this conclusion is not necessarily accurate.
+Globular clusters lose mass and evaporate over time, leading to an incomplete catalog of perturbers.
+Moreover, the present-day masses used in our simulations are likely lower than the historical masses of these clusters.
+For example, \citet{2024ApJ...976...54P}.
+conducted a study simulating the dissolution of a realistic globular cluster population, demonstrating that the historical population was significantly larger, with masses totaling XXX.
+This implies that more perturbers could have been present in the past, potentially increasing the frequency and number of gaps in Palomar~5's stream.
 

### 2024-12-12 paper/main.tex
updated comment to pearson

@@
 Moreover, the present-day masses used in our simulations are likely lower than the historical masses of these clusters.
-For example, \citet{2024ApJ...976...54P}.
-conducted a study simulating the dissolution of a realistic globular cluster population, demonstrating that the historical population was significantly larger, with masses totaling XXX.
+For example, \citet{2024ApJ...976...54P} conducted a study simulating the dissolution of a realistic globular cluster population, in the aims of identifying how many stellar streams we should expect in the Milky Way and used a mock catalog of globular clusters with an initial amount with masses above 10~$^4 M_\odot$ totaling at about 10,000 clusters.
 This implies that more perturbers could have been present in the past, potentially increasing the frequency and number of gaps in Palomar~5's stream.

### 2024-12-12 paper/main.tex
fixed malhan to banik citation

@@
 We suggest that this may be due to the signal from only one or two gaps caused by globular clusters not being sufficient to produce notable differences in the overall power spectrum.
-Additionally, \citet{2022ApJ...926..107M} may not have inspected the profiles for individual gaps, which could explain why these features were not reported in their study.
+Additionally, \citet{2019MNRAS.484.2009B} may not have inspected the profiles for individual gaps, which could explain why these features were not reported in their study.
 Further investigations would be needed to confirm this interpretation and to thoroughly assess the impact of globular clusters on stream density profiles.

### 2024-12-15 paper/main.tex
updated affiliations
Added Gieles citation

@@
          \and
-       Marco Montuori\inst{1}
+       Marco Montuori\inst{2}
        \and
-       Paola Di Matteo\inst{2}
+       Paola Di Matteo\inst{1}
        \and
@@
        \and 
-       Misha Haywood \inst{2}
+       Misha Haywood \inst{1}
        \and
@@
        \and
-       Sergey Khopserskov \inst{2,5}
+       Sergey Khopserskov \inst{5}
        \and
-       Nicolas Leclerc \inst{2}
+       Nicolas Leclerc \inst{1}
        \and 
-       Clement Hottier \inst{2}
+       Clement Hottier \inst{1}
        \and 
-       Eliot Stein  \inst{2}
+       Eliot Stein  \inst{1}
        \and
-       David Valls-Gabaud \inst{2}
+       David Valls-Gabaud \inst{1}
        \and
@@
       \and
+        Dipartimento di Fisica e Astronomia. ``Galileo Gallilei'' Università di Padova, Vicolo dell'Osservatorio 3 Padova 35122, Italy.
+      \and
         Paris Observatory. Paris Sciences et Lettres
-      \and
-        Dipartimento di Fisica e Astronomia. ``Galileo Gallilei'' Università di Padova, Vicolo dell'Osservatorio 3 Padova 35122, Italy.
       \and
@@
 \section{Methods}
+  The most accurate model of the Palomar~5 stream would involve full modeling of the internal dynamics of the cluster.
+This would mean computing N-Body interactions with a $\mathcal{O}(N^2)$ computation time, stellar evolution, super novae, an initial mass distirbution, treatment of binary star systems, etc.
+\citep[for such an example][see,]{2021NatAs...5..957G}.
+Instead, We opt for solving the restricted-three body problem or particle-test method as we did for \citet{2023A&A...673A..44F}, which we present here for completeness.
 
 \subsection{Numerical Methodology}
-    Our numerical methodology follows the approach outlined in \citet{2023A&A...673A..44F}, which we present here for completeness.
   We begin by extracting positions in the sky, proper motions, line-of-sight velocities and distances, as well as masses and half-mass radii, of 165 globular clusters from the Galactic globular cluster catalog by \cite{2021MNRAS.505.5957B}.\footnote{The Baumgardt catalog has been assembled across a series of works, see: \cite{2020PASA...37...46B,2019MNRAS.482.5138B,2018MNRAS.478.1520B}.

### 2024-12-15 paper/main.tex
added comment about no increase in velocity dispersion

@@
        \and
-       Sergey Khopserskov \inst{5}
+       Sergey Khoperskov \inst{5}
        \and
@@
 For the first simulation, we use the most probable values for the initial conditions.
+The uncertainties are sample for all globular clusters.
+Additionally, for each re-sampling of Palomar~5's mass, we also resample the distribution of the 100,000 star-particles.
 
@@
     The analysis presented in Fig.~\ref{fig:decomposition} has been repeated for the whole set of simulations, and is reported in Appendix~\ref{sec:gallery_of_gaps} for completeness.
+For all of the streams, morphologically speaking, the only changes appear to be the existence of gaps or not.
+We do not observe a thickening of the streams due to an increased velocity dispersion.
+
     From this analysis we can derive a statistical view of: 1) the number of gaps generated on Pal~5 stream by the whole system of Galactic globular clusters; 2) the clusters which generated these gaps and the time history of these perturbations.

### 2024-12-17 paper/main.tex
added alessandra's comments

--- 2024-12-12 paper/main.tex
@@
 
-%\title{Gaps in stellar streams as result of globular cluster close passages: the case of Palomar 5}
+%\title{Gaps in stellar streams as result of globular cluster close passages: the case of Palomar~5}
 %\title{The impact of globular clusters close passages on stellar streams: the case of Pal5}
@@
  % aims heading (mandatory)
-  {In this work, we investigate the perturbations induced on streams by the passage of dense stellar systems, such as globular clusters themselves, to test the possibility that they too may cause the formation of gaps.
-In particular, we focus on the study of the stream of Palomar 5, a well-known globular cluster in the Galactic halo, which has particularly long tidal tails.
+  {In this work, we investigate the perturbations induced on streams by the passage of dense stellar systems, such as globular clusters themselves, to test the possibility that they may cause the formation of gaps as well.
+In particular, we focus on the study of the stream of Palomar~5, a well-known globular cluster in the Galactic halo, which has particularly long tidal tails.
 }
  % methods heading (mandatory)
-  {For this purpose, we used a particle-test code to simulate the dynamic evolution of Palomar~5 and its tidal tails when Pal~5 is subject to the gravitational action of the Galaxy and its whole system of globular clusters.}
+  {For this purpose, we used a particle-test code to simulate the dynamic evolution of Palomar~5 and its tidal tails when Palomar~5 is subject to the gravitational action of the Galaxy and its whole system of globular clusters.}
  % results heading (mandatory)
-  {Our study shows that the tails of Pal~5 are strongly perturbed by the close passage of other clusters, in particular of NGC~2808, NGC~7078, NGC~104, and that these perturbations induce the formation of gaps in the tails.}
+  {Our study shows that the tails of Palomar~5 are strongly perturbed by the close passage of other clusters, in particular of NGC~2808, NGC~7078, NGC~104, and that these perturbations induce the formation of gaps in the tails.}
  % conclusions heading (optional), leave it empty if necessary
@@
 Numerical predictions of this phenomenon exist since the 1970s \citep[see, for example][]{1975AJ.....80..290K}.
-These predictions occurred well before the first detections which were made by \citet{1995AJ....109.2553G} around some Galactic globular clusters.
-Interestingly, this work was nearly contemporary discovery of the Sagittarius stellar stream by \citet{1994Natur.370..194I} which is the closest example of a stream emerging from a dwarf satellite currently interacting with the Milky Way.
+
+  These predictions occurred well before the first detections of Galactic globular cluster tidal tails \citep{1995AJ....109.2553G}.
+Interestingly, \citep{1995AJ....109.2553G}'s detections were made nearly contemporary  with the discovery of the Sagittarius stellar stream by \citet{1994Natur.370..194I} which is the closest example of a stream emerging from a dwarf satellite currently interacting with the Milky Way.
 
@@
 The possibility given by Gaia to track stars with coherent movements, over the entirety of the sky, has led to the discovery of dozens of new streams.
-For example, \citet{2018MNRAS.477.4063M} developed the \texttt{streamfinder} algorithm and applied it across a series works of \citep{2018MNRAS.481.3442M,  2018ApJ...865...85I, 2019ApJ...872..152I} to discover the series of streams: Slidr, Sylgr, Ylgr, Fimbulthul, Svol, Gjoll, Leiptr, Phlegeton and Fjorm.
-According to the most recent estimates provided by \citet{2020MNRAS.493.4978S} Fjorm in particular has an extension of more than 150 degrees on the sky which has a projected length more than five times greater than that of Pal~5.
+For example, \citet{2018MNRAS.477.4063M} developed the \texttt{streamfinder} algorithm and applied it across a series works of \citep{2018MNRAS.481.3442M,  2018ApJ...865...85I, 2019ApJ...872..152I} to discover the series of streams.%: Slidr, Sylgr, Ylgr, Fimbulthul, Svol, Gjoll, Leiptr, Phlegeton and Fjorm.
+According to the most recent estimates provided by \citet{2020MNRAS.493.4978S} Fjorm in particular has an extension of more than 150 degrees on the sky which has a projected length more than five times greater than that of Palomar~5.
 
@@
 \citet{2023MNRAS.520.5225M} provides a compilation of their tracks on the sky.
-Interestingly, only about 20 streams associated to known Galactic globular clusters.
+Interestingly, only about 20 streams are associated to known Galactic globular clusters.
 
@@
 However, the use of single streams for constraining the potential led to some ambiguous and not converging results over time.
-For example, \citet{2010ApJ...718.1128L} made use of the Sagittarius stream to infer that the dark matter halo of our Galaxy has a triaxial shape \citep[but see also][]{2004MNRAS.351..643H, 2005ApJ...619..800J, 2005ApJ...619..807L}; \citet{2016ApJ...833...31B} concluded that the dark matter halo of our Galaxy is nearly spherical at the distances of the Palomar~5 and GD~1 streams.
+For example, \citet{2010ApJ...718.1128L} made use of the Sagittarius stream to infer that the dark matter halo of our Galaxy has a triaxial shape \citep[but see also][]{2004MNRAS.351..643H, 2005ApJ...619..800J, 2005ApJ...619..807L}; \citet{2016ApJ...833...31B} concluded that the dark matter halo of our Galaxy is nearly spherical at the distances of the Palomar~5 and GD-1 streams.
 Probably only by using the set of known streams can the degeneracy of the solution be eliminated.
-This approach has been recently used by \citet{2024ApJ...967...89I} who constructed the most accurate model of the Milky Way.
+Recently, \citet{2024ApJ...967...89I} constructed the most accurate model of the Milky Way to date by applying a Monte Carlo Markov Chain (MCMC) fitting procedure.
+This method identified the set of potential parameters in an axisymmetric model of the Milky Way that best reproduces all observed stellar streams.
+This work represents a seminal achievement and a significant milestone in the field.
 
@@
 These clumps, or \textit{sub-halos}, are predicted to follow a mass distribution with a power-law slope slightly shallower than -2.0.
-To date, the smallest observed dark matter halo was detected through gravitational lensing in an Einstein ring and had a mass of $10^8$ solar masses, as described by \citet{2012Natur.481..341V}.
-However, some models predict that dark matter clumps could exist down to the mass of Earth-like planets at least \citep[see][ and discussion in \citet{2021arXiv211101148A} ]{2005JCAP...08..003G}.
+To date, the smallest observed dark matter halo was detected through gravitational lensing in an Einstein ring with a mass of $10^8$ solar masses by \citet{2012Natur.481..341V}.
+However, some models predict that dark matter clumps could exist down to at least the mass of Earth-like planets at least \citep[see][and discussion in \citet{2021arXiv211101148A}]{2005JCAP...08..003G}.
 
@@
 Later, \citet{2012ApJ...748...20C} expanded this idea, proposing that sub-halos could create gaps in stellar streams during flyby encounters, where a sub-halo approaches closely enough to a segment of a stream and significantly changes the orbits of the closest stars.
-\citet{2012ApJ...760...75C, 2013ApJ...768..171C} applied this idea and looked for the presence of gaps in the known Palomar~5 and GD~1 streams, concluding that the density variations found in their streams were consistent with expectations from $\Lambda$~CDM models.
+\citet{2012ApJ...760...75C, 2013ApJ...768..171C} applied this idea and looked for the presence of gaps in the known Palomar~5 and GD-1 streams, concluding that the density variations found in their streams were consistent with expectations from $\Lambda$~CDM models.
 \citet{2019ApJ...880...38B} provided further observational evidence for this idea, identifying a gap and a spur in the GD-1 stream that they could not explain by known objects, such as globular clusters, and which they suggested was due to the close passage of a dark matter sub-halo.
@@
 from baryonic matter, can cause variations in the density of streams and gaps that can be assimilated and confused with those produced by dark matter sub-halos.
-Among these works, it is worth mentioning the results of \citet{2017NatAs...1..633P} that suggests how the presence of the bar at the center of the Galaxy can perturb the characteristics of a stream such as Pal~5 and generate gaps along its tail.
+Among these works, it is worth mentioning the results of \citet{2017NatAs...1..633P} that suggests how the presence of the bar at the center of the Galaxy can perturb the characteristics of a stream such as Palomar~5 and generate gaps along its tail.
 That the Galactic bar could have an influence on streams morphology was also discussed by \citet{2016MNRAS.460..497H} and \citet{2016ApJ...824..104P}, in the case of the Ophiuchus stream \citep{2014MNRAS.443L..84B}.
@@
 More recently \citet{2022ApJ...941..129D} have examined the possibility that the gaps found in the GD-1 stellar stream could be due to the close passage of globular clusters, concluding that this scenario is highly unlikely.
-These first works seem to suggest that the impact of globular clusters on stellar streams is negligible, and this is somewhat surprising, especially for streams of clusters as Palomar~5, which are confined in the inner 20~kpc of the Galaxy, where many other globular clusters do also orbit.
-Additionally, \citet{2023A&A...678A..69I} showed that globular cluster can even collide with other clusters, which implies that cluster stream collisions should happen much more frequently since streams are far more extended than clusters.
+These first works seem to suggest that the impact of globular clusters on stellar streams is negligible.
+This does not necessarily need to be the general case, especially for streams of clusters as Palomar~5, which are confined in the inner 20~kpc of the Galaxy, where many other globular clusters do also orbit.
+For example, \citet{2023A&A...678A..69I} showed that globular cluster can even collide with other clusters, which implies that cluster stream collisions should happen much more frequently since streams are far more extended than clusters.
 
@@
 We seek to quantify the importance of the effects of passing globular clusters in the vicinity of streams, to understand whether these systems can also be effective - and how often - in altering the distribution of stars in the tails and producing under-dense regions or gaps.
-To this end, on the following pages we will present the results of simulations of the streams of Palomar~5, subject to the gravitational interaction with the set of 165 Galactic globular clusters for which positions and velocities are known to date, and for which orbits can therefore be reconstructed.
+To this end, on the following pages we will present the results of simulations of the streams of Palomar~5, subject to the gravitational interaction with the set of 165 Galactic globular clusters for which positions and velocities are known to date, and for which orbits can therefore be reconstructed \citep{2021MNRAS.505.5957B}.
 We chose to simulate Palomar~5 because it is a typical halo cluster with extended tails, and because it is a cluster for which the effect of baryonic structures on its stream has already been studied.
-As we will show, and in tension with previous claims, the close passage of other clusters with the stream of Pal~5 is not a rare event, and indeed, in the 50 simulations we ran, we find the formation of numerous gaps, for an average of 1.5 gaps per simulation, generated by 18 different clusters among the whole system of Galactic globular clusters.
+As we will show, and in tension with previous claims, the close passage of other clusters with the stream of Palomar~5 is not a rare event, and indeed, in the 50 simulations we ran, we find the formation of numerous gaps, for an average of 1.5 gaps per simulation, generated by 18 different clusters among the whole system of Galactic globular clusters.
 
@@
     \includegraphics[width=\linewidth]{stream_on_sky_Pal5_monte-carlo-009_pouliasis2017pii-GCNBody_pouliasis2017pii.png}
-    \caption{Simulated Palomar 5 stream created by modeling the host cluster as a Plummer sphere disrupting within an axis-symmetric galactic potential plus the gravitational effect of 164 other galactic globular clusters.
+    \caption{Simulated Palomar~5 stream created by modeling the host cluster as a Plummer sphere disrupting within an axis-symmetric galactic potential plus the gravitational effect of 164 other galactic globular clusters.
 The top panel shows the distribution of star-particles that escaped the cluster due to tidal forces.
@@
 This model satisfies a series of observational constraints such as local solar stellar density, the galactic rotation curve, etc, similarly to other galactic models such as \texttt{MWpotential2014} from \citet{2015ApJS..216...29B} and \texttt{McMillian2017} from \citet{2017MNRAS.465...76M}.
-However, To balance data volume and computation time, we use only one galactic potential, which should suffice.
+However, to balance data volume and computation time, we use only one galactic potential, which should suffice.
 \citet{2021MNRAS.505.5978V} found that only a few outer globular clusters are strongly affected by different potential models.
@@
     There is a primary methodological departure from \citet{2023A&A...673A..44F}.
-In that paper, globular clusters evolved under the gravitational effect of the Galaxy only, while now we include also the effect of all the other Galactic globular clusters, by taking into account the direct N-body interactions among them.
+In that paper, globular clusters evolved under the gravitational effect of the Galaxy only, while now we include also the effect of all the other Galactic globular clusters, by taking into account the direct $N$-body interactions among them.
 To quantify these interactions, all clusters are represented by Plummer spheres, each with mass and half-mass radius as reported in the Baumgardt catalog \citep{2021MNRAS.505.5957B}.
@@
 The particles experience the gravitational field yet do not contribute to it.
-The equation of motion of a generic particle among the 100,000 which populate Pal~5 is thus:
+The equation of motion of a generic particle among the 100,000 which populate Palomar~5 is thus:
         \begin{equation}
@@
       During the integration, intermediate snapshots were saved to facilitate the analysis of stellar streams and the effects of cluster impacts.
-Specifically, for each of the 50 realizations of the Palomar 5 stream, we saved $5000$ intermediate time-steps which is a temporal resolution of 1 million years.
+Specifically, for each of the 50 realizations of the Palomar~5 stream, we saved $5000$ intermediate time-steps which is a temporal resolution of 1 million years.
 We provide the parameters that specify our data volume in Table~\ref{tab:data_volume}.
@@
 
-    To quantify the impact of globular cluster passages to the density of Pal~5 stream, we performed a second set of simulations, which we refer to as the \texttt{reference} simulations in this paper.
+    To quantify the impact of globular cluster passages to the density of Palomar~5 stream, we performed a second set of simulations, which we refer to as the \texttt{reference} simulations in this paper.
 These \texttt{reference} simulations use the same 50 sets of initial conditions as of the \texttt{full} simulations, the same galactic potential but exclude mutual interactions between globular clusters.
@@
 In Eq.~\ref{eq:GCNBody} only the gradient of the Galactic potential  is considered.
-In Eq.~\ref{eq:equation_of_motion_particle}, of the second term on the right-side of the equation, only the influence of Pal~5's Plummer sphere on Pal~5's particles is considered.
+In Eq.~\ref{eq:equation_of_motion_particle}, of the second term on the right-side of the equation, only the influence of Palomar~5's Plummer sphere on Palomar~5's particles is considered.
 In other words, sum iterates over only one globular cluster, the host.
@@
 The time step was therefore set to be small enough to conserve energy for the most interior particles within the cluster---ensuring that a higher mass loss did not arise from numerical error.
-For these simulations, a time-step of 10,000 years was selected to maintain adequate energy conservation.
+For these simulations, a time-step of 10,000 years was selected to maintain adequate energy conservation of an median variation of $10^{-12} \frac{\Delta E}{E_0}$.
 
@@
 By reverse integrability, we mean the integrator's capability to track the cluster backward in time and then re-integrate it forward along the same trajectory.
-The drift in the final position after forward integration, compared to the initial position from the backward integration, was consistently at least two orders of magnitude smaller than the Plummer scale radius used for Palomar 5.
+The drift in the final position after forward integration, compared to the initial position from the backward integration, was consistently at least two orders of magnitude smaller than the Plummer scale radius used for Palomar~5.
 This high precision ensures that no fictitious evil numerical forces influenced the system, preventing any artificial mass loss or retention of star particles.
@@
 
-    We see clear effects of the presence of their gravitational field in the properties of Pal~5 stream.
+    We see clear effects of the presence of their gravitational field in the properties of Palomar~5 stream.
 A particularly clear example of this effect is shown in Fig.~\ref{fig:stream_on_sky} which we selected for its prominent gaps.
@@
 First is that Palomar~5's tails indeed have more mass than the cluster itself.
-\citet{2017ApJ...842..120I} reports that there could be three times as much mass in Palomar 5's tails as the cluster itself.
+\citet{2017ApJ...842..120I} reports that there could be three times as much mass in Palomar~5's tails as the cluster itself.
 Secondly, the exact form of the density distribution is less important than having a population of particles present that can probe a cluster flyby event.
@@
 
-    To find out which globular clusters were responsible of creating these gaps, and at what time the close passages took place, we estimated the gravitational acceleration along the orbit of Pal~5 and represented it in the ($t$, $\tau$) space.
-$t$ is the simulation time and $\tau$ is indicates how long it will take for Pal~5 to reach, or how  long ago it passed, a given point on its orbit.
+    To find out which globular clusters were responsible of creating these gaps, and when close passages took place, we estimated the gravitational acceleration along the orbit of Palomar~5 and represented it in the ($t$, $\tau$) space.
+$t$ is the simulation time and $\tau$ is indicates how long it will take for Palomar~5 to reach, or how  long ago it passed, a given point on its orbit.
 The use of $\tau$ is advantageous because the growth of the stream is approximately linear in $\tau$.
@@
     The analysis presented in Fig.~\ref{fig:decomposition} has been repeated for the whole set of simulations, and is reported in Appendix~\ref{sec:gallery_of_gaps} for completeness.
-From this analysis we can derive a statistical view of: 1) the number of gaps generated on Pal~5 stream by the whole system of Galactic globular clusters; 2) the clusters which generated these gaps and the time history of these perturbations.
-From this, we can then quantify: 3) the properties of the perturbers (their masses, sizes and their orbital parameters) as well as 4) the impact geometry of the encounters, which allows us to understand which encounters are more favorable to generate gaps in Pal~5 stream.
+From this analysis we can derive a statistical view of: 1) the number of gaps generated on Palomar~5 stream by the whole system of Galactic globular clusters; 2) the clusters which generated these gaps and the time history of these perturbations.
+From this, we can then quantify: 3) the properties of the perturbers (their masses, sizes and their orbital parameters) as well as 4) the impact geometry of the encounters, which allows us to understand which encounters are more favorable to generate gaps in Palomar~5 stream.
 In the following, we will present results of this analysis, by starting from addressing points 1) and 2).
@@
 The results of this analysis are reported in Fig.~\ref{fig:histogram_impact_time} and in Table~\ref{tab:gaps_per_perturber}.
-NGC~2808 impacts Palomar 5's stream about 200 Myr ago in 44 out of 50 simulations, creating a small gap at a similar position to the one reported in Fig.~\ref{fig:decomposition} in each case.
+NGC~2808 impacts Palomar~5's stream about 200 Myr ago in 44 out of 50 simulations, creating a small gap at a similar position to the one reported in Fig.~\ref{fig:decomposition} in each case.
 Since this interaction occurred less than one orbital period ago, the uncertainties that lead to orbital solutions remain that similar and thus producing consistent results across all simulations.
@@
 This leaves 21 gaps in the leading tail and 8 in the trailing.
-The probability of observing up to 8 successes in 29 trials, given a 50\% chance of success, is 1.2\%--—unlikely, but possible.
+The probability of observing up to 8 successes in 29 trials, given a 50\% chance of success, is 1.2\%--unlikely, but possible.
 Additionally, other perturbers impact at consistent times, which may also violate the assumption of independent events, as seen with NGC~2808.
@@
 
-    With the perturbers identified, we perform statistics in the pursuit of understanding what conditions are necessary for a globular cluster in order to form a gap on Pal~5 stream.
+    With the perturbers identified, we perform statistics in the pursuit of understanding what conditions are necessary for a globular cluster in order to form a gap on Palomar~5 stream.
 We turn to impact theory, which in its simplest form is presented in works such as \citet{2008gady.book.....B}.
@@
 Consequently, five parameters determine the change in velocity of a given stream particle: $M$, $r_p$, $b$, $W_\parallel$, and $W_\perp$.
-As detailed in Appendix~\ref{sec:reconstruction}, we calculated these parameters for all our \texttt{full} simulations, by selecting -- for each of them --  the strongest 5 flybys of a perturber with the Pal~5 stream.
+As detailed in Appendix~\ref{sec:reconstruction}, we calculated these parameters for all our \texttt{full} simulations, by selecting -- for each of them --  the strongest 5 flybys of a perturber with the Palomar~5 stream.
 Thus, we compute a total of  250 impacts, and flag those that give way to gaps.
@@
 A series of perturbers at roughly $\sim7 \times 10^5 M_\odot$ highlights NGC~2808's flybys, where all encounters with impact parameters under 200~pc result in gaps, while those beyond this distance do not.
-In other words, massive globulars with masses greater than $10^6 \rm{M}_\odot$ are not effective in producing gaps if they pass close to Pal~5 stream with an impact parameter $b > 200$~pc.
+In other words, massive globulars with masses greater than $10^6 \rm{M}_\odot$ are not effective in producing gaps if they pass close to Palomar~5 stream with an impact parameter $b > 200$~pc.
 It is interesting to note that for values of the impact parameter below this threshold, even fast encounters ($\delta v > 300$~km/s) can produce gaps.
@@
       \caption{\textit{Characteristics of gap causing clusters} \textit{Top Left:} Energy-angular momentum space of the globular clusters in the simulations, with 50~$\times$~165 data points representing all sampled initial conditions.
-Clusters impacting Palomar 5 are shown with colored markers; non-gap-causing clusters are shown in gray.
-Palomar 5 is marked with a white symbol and black edge.
+Clusters impacting Palomar~5 are shown with colored markers; non-gap-causing clusters are shown in gray.
+Palomar~5 is marked with a white symbol and black edge.
 The circular velocity-curve is drawn with a light-pink dashed curve.
@@
 We recall the reader that the globular cluster catalog at this time does not provide uncertainties for clusters' characteristic radii.
-\textit{Bottom Left:} Palomar 5's orbit in the Galactocentric xy plane.
+\textit{Bottom Left:} Palomar~5's orbit in the Galactocentric xy plane.
 The gray-scale represents all 50 stacked orbits, with $\mathcal{P}$ indicating the probability of  Palomar~5's position, normalized to $\mathcal{P}_\textrm{max}=1$.
@@
 
-    Once all the key impact parameters estimated, we can use Eq~\ref{eq:change_in_momentum} and calculate $\Delta p$, the change in momentum (per unit mass) imparted by a cluster flyby on Pal~5 stream.
+    Once all the key impact parameters estimated, we can use Eq~\ref{eq:change_in_momentum} and calculate $\Delta p$, the change in momentum (per unit mass) imparted by a cluster flyby on Palomar~5 stream.
 The result  is shown in Fig.~\ref{fig:deltap}, where the distribution of imparted in change in momentum is shown for all impacts which produce a gap, and compared to those that do not produce one.
@@
 This is shown in Fig~\ref{fig:mass_size_plane}, where we first show the mass and size (i.e.
-half-mass radius) of the clusters that cause gaps on the tails of Pal~5, dividing them from those that do not.
-As can be seen, no cluster with mass below $10^5 \rm{M}_\odot$ causes gaps on Pal~5 stream, and all perturbers, except FSR~1758 have a half-mass radius below 10~pc.
+half-mass radius) of the clusters that cause gaps on the tails of Palomar~5, dividing them from those that do not.
+As can be seen, no cluster with mass below $10^5 \rm{M}_\odot$ causes gaps on Palomar~5 stream, and all perturbers, except FSR~1758 have a half-mass radius below 10~pc.
 Even more interesting is their distribution in the E-Lz plane, which shows that the clusters that cause gaps are on both direct and retrograde orbits (respectively, negative and positive values of $L_z$), but all confined to an energy interval between $-2$ and $-1 \times10^5~\textrm{km}^2\textrm{s}^{-2}$.
-This is because only clusters within Palomar 5's orbital space can interact with the stream: clusters with higher orbital energies tend to have larger apocenter than that of Pal~5 and thus spend most of their time away from Palomar 5's orbital volume.
+This is because only clusters within Palomar~5's orbital space can interact with the stream: clusters with higher orbital energies tend to have larger apocenter than that of Palomar~5 and thus spend most of their time away from Palomar~5's orbital volume.
 
-    Finally, it is worth commenting on the location of impacts with the stream, whether they occur when Pal~5 is close to its pericenter or not.
-As it can be seen from Fig.~\ref{fig:mass_size_plane} (bottom panel) encounters can occur at all orbital phases of Pal~5, when it is close to pericenter, but also very far away from it, at the outskirts of its orbital space.
+    Finally, it is worth commenting on the location of impacts with the stream, whether they occur when Palomar~5 is close to its pericenter or not.
+As it can be seen from Fig.~\ref{fig:mass_size_plane} (bottom panel) encounters can occur at all orbital phases of Palomar~5, when it is close to pericenter, but also very far away from it, at the outskirts of its orbital space.
 However, when taken all together, the location of gap creating impacts shows a strong negative correlation with the galactocentric radius $r$, with the number $N$ of encounters favorable to gap formation  going as $N = -2.5r + 50$ (with a Pearson coefficient of -0.86).
@@
 However, this result suggests that cluster galactic number density outweighs these factors in determining the number of gaps.
-It should be noted that these results are for Palomar 5 only.
+It should be noted that these results are for Palomar~5 only.
 Streams along various orbits needs to be studied in order for this conclusion to be generalized.
@@
     Another possible explanation is the different thickness of the simulated tails at different distances from the center of the cluster.
-At distances of less than about 3~kpc from Pal~5, the tails are very thin, with a typical thickness being less than 100 pc.
+At distances of less than about 3~kpc from Palomar~5, the tails are very thin, with a typical thickness being less than 100 pc.
 In order to cause gaps in these regions, the clusters would have to pass close to the stream, with impact parameters similar to the thickness itself.
 Note that the thinness of the tails at these distances is probably a direct consequence of the initial parameters we chose for the simulation.
-We have assumed for Pal~5, 5 Gyr years ago, the same internal parameters (mass and size) that the cluster has today.
+We have assumed for Palomar~5, 5 Gyr years ago, the same internal parameters (mass and size) that the cluster has today.
 With such parameters, after 5 Gyr of evolution, our system has lost most of its mass, and therefore the part of the tails closest to the cluster itself, whose density depends mainly on the most recent mass loss \citep[see, for example, Fig.~A.3 in][]{2012A&A...546L...7M}, is necessarily very thin because the simulated cluster has essentially no more mass to lose.
@@
   The simulations presented in this paper suggest that, in the last 5~Gyr of evolution, Palomar~5's stream could have experienced multiple close encounters with other Galactic globular clusters, some of which are able to create gaps -- even a few kpc wide -- in its tails.
-Currently, the presence -- or not -- of gaps in the observed portion of Pal~5 tails is debated: \citet{2016ApJ...819....1I} found no statistically significant gap in Pal~5 tails, while \citet{2017MNRAS.470...60E}, analyzing the same data-set as \citet{2016ApJ...819....1I}, suggested the presence of a few dips and gaps in the tails, at angular distances between $2^\circ$ and $9^\circ$ from the cluster center \citep[see also][]{ 2020ApJ...889...70B}.
+Currently, the presence -- or not -- of gaps in the observed portion of Palomar~5 tails is debated: \citet{2016ApJ...819....1I} found no statistically significant gap in Palomar~5 tails, while \citet{2017MNRAS.470...60E}, analyzing the same data-set as \citet{2016ApJ...819....1I}, suggested the presence of a few dips and gaps in the tails, at angular distances between $2^\circ$ and $9^\circ$ from the cluster center \citep[see also][]{ 2020ApJ...889...70B}.
 While our simulations produced 73 gaps across 50 realizations, only about 22 are beyond the current length of the observed portion of the stream.
@@
 As we have discussed in Sect.~\ref{sect:geometry}, the number, $N$, of close encounters which lead to gap creation is anti-correlated with the distance $r$ to the Galactic center.
-If we naively use the same radial dependence of gaps from Palomar 5 for GD-1 by swapping a pericenter from 6~kpc to 16~kpc, we would reduced the number of gap-favorable impacts of more than a factor of 2.
+If we naively use the same radial dependence of gaps from Palomar~5 for GD-1 by swapping a pericenter from 6~kpc to 16~kpc, we would reduced the number of gap-favorable impacts of more than a factor of 2.
 However, we note that \citet{2022ApJ...941..129D} preselect the globular clusters which could have experienced a close encounter with the GD-1 stream, by selecting only clusters which pass at a distance less than 0.5 kpc from the stream, having a relative velocity smaller than  300 km/s.
@@
 While our simulations predict the formation of gaps in Palomar~5's tails, the predicted gaps do not align with the regions of the stream currently observed.
-Several factors contribute to the absence of simulated gaps in the observed portions of Palomar 5's tails, including the stream's varying thickness, the dispersion of action-frequencies near the cluster, and the initial conditions adopted in our simulations.
+Several factors contribute to the absence of simulated gaps in the observed portions of Palomar~5's tails, including the stream's varying thickness, the dispersion of action-frequencies near the cluster, and the initial conditions adopted in our simulations.
 
   The broader implications of our findings indicate that globular cluster interactions add a layer of complexity to interpreting stellar stream substructures, complicating efforts to distinguish between baryonic and dark matter-induced gaps.
-While Palomar 5's phase-space region may not be ideal for isolating the impact of dark matter sub-halos, extending this analysis to other streams, particularly those at larger galactocentric distances, could provide clearer insights.
+While Palomar~5's phase-space region may not be ideal for isolating the impact of dark matter sub-halos, extending this analysis to other streams, particularly those at larger galactocentric distances, could provide clearer insights.
 
@@
 Next, we modeled Palomar~5's orbit as a proxy for its stream and sampled points along the orbit to measure the gravitational force exerted by other globular clusters.
-This force data is displayed in the top panel of Fig.~\ref{fig:force-on-orbit}, showing how the total gravitational acceleration on Palomar 5's stream evolves over its length throughout the simulation.
+This force data is displayed in the top panel of Fig.~\ref{fig:force-on-orbit}, showing how the total gravitational acceleration on Palomar~5's stream evolves over its length throughout the simulation.
 
@@
 The density was used to determine a suitable length of the stream at a given function.
-This length was then used to extract a piece of Palomar 5's orbit.
+This length was then used to extract a piece of Palomar~5's orbit.
 This orbital segment is used to approximate the stream.

### 2024-12-17 paper/main.tex

@@
 \section{Methods}
+  The most accurate model of the Palomar~5 stream would involve full modeling of the internal dynamics of the cluster.
+This would mean computing N-Body interactions with a $\mathcal{O}(N^2)$ computation time, stellar evolution, super novae, an initial mass distirbution, treatment of binary star systems, etc.
+\citep[for such an example][see,]{2021NatAs...5..957G}.
+Instead, We opt for solving the restricted-three body problem or particle-test method as we did for \citet{2023A&A...673A..44F}, which we present here for completeness.
 
   \subsection{Numerical Methodology}
-    Our numerical methodology follows the approach outlined in \citet{2023A&A...673A..44F}, which we present here for completeness.
     We begin by extracting positions in the sky, proper motions, line-of-sight velocities and distances, as well as masses and half-mass radii, of 165 globular clusters from the Galactic globular cluster catalog by \cite{2021MNRAS.505.5957B}.\footnote{The Baumgardt catalog has been assembled across a series of works, see: \cite{2020PASA...37...46B,2019MNRAS.482.5138B,2018MNRAS.478.1520B}.
@@
     The analysis presented in Fig.~\ref{fig:decomposition} has been repeated for the whole set of simulations, and is reported in Appendix~\ref{sec:gallery_of_gaps} for completeness.
-From this analysis we can derive a statistical view of: 1) the number of gaps generated on Palomar~5 stream by the whole system of Galactic globular clusters; 2) the clusters which generated these gaps and the time history of these perturbations.
-From this, we can then quantify: 3) the properties of the perturbers (their masses, sizes and their orbital parameters) as well as 4) the impact geometry of the encounters, which allows us to understand which encounters are more favorable to generate gaps in Palomar~5 stream.
+For all of the streams, morphologically speaking, the only changes appear to be the existence of gaps or not.
+We do not observe a thickening of the streams due to an increased velocity dispersion.
+
+    From this analysis we can derive a statistical view of: 1) the number of gaps generated on Pal~5 stream by the whole system of Galactic globular clusters; 2) the clusters which generated these gaps and the time history of these perturbations.
+From this, we can then quantify: 3) the properties of the perturbers (their masses, sizes and their orbital parameters) as well as 4) the impact geometry of the encounters, which allows us to understand which encounters are more favorable to generate gaps in Pal~5 stream.
 In the following, we will present results of this analysis, by starting from addressing points 1) and 2).

### 2024-12-18 paper/main.tex
updated track mid point figure in the text
updated text for the observation limit of the gaps

@@
 
-  \citet{2024arXiv240519410B} provides a review of stellar stream astronomy, which has entered a new era since the publication of the data from the Gaia astrometric mission \citep{2016A&A...595A...1G}.
+  \citet{2025NewAR.10001713B} provides a review of stellar stream astronomy, which has entered a new era since the publication of the data from the Gaia astrometric mission \citep{2016A&A...595A...1G}.
 Gaia's characterization of billions of stars in the Milky Way allows us to look for these structures coupling photometry, astrometry and, for the brightest stars, spectroscopy.
@@
       \centering
-      \includegraphics[width=\linewidth]{decomposition-monte-carlo-009-with-3-gaps.png}
+      \includegraphics[width=\linewidth]{decomposition-monte-carlo-009-with-3-gaps-domidpoint-shift.png}
       \caption{Density maps of Palomar~5's stream in the tail coordinate system.
@@
     We briefly compare our simulated gaps to the literature on Paloamr~5.
-In the bottom panel of Fig.~\ref{fig:stream_on_sky}, we compare the tracks of Palomar~5 that were compiled in Galstreams by \citet{2023MNRAS.520.5225M}.
+In the bottom panel of Fig.~\ref{fig:decomposition}, we compare the tracks of Palomar~5 that were compiled in Galstreams by \citet{2023MNRAS.520.5225M}.
 The mid-point positions of the streams do not have the same galacto-centric positions as Palomar~5 from Baumgardt's catalog and are thus offset when projected into tail coordinates.
-Nonetheless, we note that the lengths of the streams are at max 6~kpc on either side of the globular cluster.
-We apply a threshold of 5.75~kpc on either side to consider the gaps that could occur in currently observed regions of Paloamr~5.
+Moreover, since we sample the distances to Palomar~5, the galactocentric position within the Baumgardt catalog varies.
+To combat this, we position the mid-point of the Galstream tracks at the center of mass of the cluster, which allows us to compare the length of the observed tracks to our simulated streams.
+We use the three tracks Pal5-PW19, Pal5-S20, and Pal5-I21 from \citet{2019AJ....158..223P}, \citet{2020MNRAS.495.2222S}, and \citet{2021ApJ...914..123I}, respectively.
+For each track, we find the distance from the cluster in either direction and count the number of gaps that fall within this range.
 This is presented in Fig.~\ref{fig:GapsWithinSight}.
+In the maximum limit many gaps could appear, at a rate of about one per realization.
+However, at the shortest reported stream-legnth only a few gaps occur.
       \begin{figure}
@@
         \includegraphics[width=\linewidth]{GapWithinSight.png}
-        \caption{The distribution of gaps in absolute distance from the center of mass of Palomar~5's globular cluster.
-A threshold value of $\sim$~5.75~kpc was selected as distance threshold that corresponds to the current observed length.
-Distances beyond this are shaded.
+        \caption{The distribution of gaps as a function of absolute distance from the center of mass of Palomar 5's globular cluster.
+The number of gaps within the observable range is evaluated on an individual track basis, with track lengths determined by the minimum and maximum $x'$-coordinates from the bottom panel of Fig.~\ref{fig:decomposition}.
+The ranges for \citet{2019AJ....158..223P}, \citet{2020MNRAS.495.2222S}, and \citet{2021ApJ...914..123I} are shown for their respective dimensions, and the number of gaps within these ranges is indicated in the figure.
+Vertical bars mark the $x' > 0$ limit of the stream--we only show the right hand side instead of both for clarity, as most gaps are located on the right-hand side of the stream in any case.
 }

### 2024-12-18 paper/main.tex
added spaces between number and Fig.\ref

@@
 The \texttt{Reference} simulation omits other globular clusters.
-The bottom panel compares Palomar5's observed stream length to the \texttt{Reference} simulation, using the same Monte Carlo realization as Fig.\ref{fig:stream_on_sky} and \texttt{Sampling~009} of Fig.~\ref{fig:gallery2}.}
+The bottom panel compares Palomar5's observed stream length to the \texttt{Reference} simulation, using the same Monte Carlo realization as Fig.~\ref{fig:stream_on_sky} and \texttt{Sampling~009} of Fig.~\ref{fig:gallery2}.}
       \label{fig:decomposition}
@@
 We refer the reader to Appendix~\ref{sec:Perturber_Identification} for an in-depth description of the procedure.
-In this way, we could identify that the clusters responsible for creating gaps in the simulation of Palomar~5 stream reported in Fig.\ref{fig:decomposition} are NGC~2808, NGC~7078 and NGC~104, and that their close passages occurred 200~Myr, -1.9 and -2.1~Gyr ago, respectively.
+In this way, we could identify that the clusters responsible for creating gaps in the simulation of Palomar~5 stream reported in Fig.~\ref{fig:decomposition} are NGC~2808, NGC~7078 and NGC~104, and that their close passages occurred 200~Myr, -1.9 and -2.1~Gyr ago, respectively.
 
@@
     Note that in this figure, we show the total relative velocity rather than separating parallel and perpendicular components, because from a visual inspection, we could not find any specific trends when the two velocity components were plotted separately.
-We also exclude the characteristic cluster radius, as it showed little correlation with the results, likely due to the narrow range of globular cluster radii (see Fig.\ref{fig:mass_size_plane}).
+We also exclude the characteristic cluster radius, as it showed little correlation with the results, likely due to the narrow range of globular cluster radii (see Fig.~\ref{fig:mass_size_plane}).
 This factor might be more significant for dark matter sub-halos, where size variation is greater.

### 2024-12-19 paper/main.tex
fix affiliation order

@@
 \institute{
-        Dipartimento di Fisica, Universit\`a di Roma ``La Sapienza'',
-          Piazza Aldo Moro\\
-          \email{salvatore.ferrone@uniroma1.it}
+  Paris Observatory. Paris Sciences et Lettres
+  \email{salvatore.ferrone@obspm.fr}
+  \and
+    Dipartimento di Fisica, Universit\`a di Roma ``La Sapienza'',Piazza Aldo Moro\\
   \and
     Dipartimento di Fisica e Astronomia. ``Galileo Gallilei'' Università di Padova, Vicolo dell'Osservatorio 3 Padova 35122, Italy.
-      \and
-        Paris Observatory. Paris Sciences et Lettres
     \and
@@
 For example, \citet{2010ApJ...718.1128L} made use of the Sagittarius stream to infer that the dark matter halo of our Galaxy has a triaxial shape \citep[but see also][]{2004MNRAS.351..643H, 2005ApJ...619..800J, 2005ApJ...619..807L}; \citet{2016ApJ...833...31B} concluded that the dark matter halo of our Galaxy is nearly spherical at the distances of the Palomar~5 and GD-1 streams.
-Probably only by using the set of known streams can the degeneracy of the solution be eliminated.
+The two contrasting results could also show that at one distance is the halo is triaxial and at another spherical showcasing the need for streams at different distances to map the halo shape.
 Recently, \citet{2024ApJ...967...89I} constructed the most accurate model of the Milky Way to date by applying a Monte Carlo Markov Chain (MCMC) fitting procedure.
@@
   Beyond the global visible and dark mass distribution streams can be used also to infer the granularity of the dark matter, that is the mass and density of the sub-halos populating our Galaxy.
-According to simulations by \citet{2008MNRAS.391.1685S}, $\Lambda$~Cold Dark Matter ($\Lambda$~CDM) predicts that galaxies grow hierarchically, with dark matter clumps forming at a wide range of masses and sizes.
+According to simulations by \citet{2008MNRAS.391.1685S}, $\Lambda$~Cold Dark Matter ($\Lambda$CDM) predicts that galaxies grow hierarchically, with dark matter clumps forming at a wide range of masses and sizes.
 These clumps, or \textit{sub-halos}, are predicted to follow a mass distribution with a power-law slope slightly shallower than -2.0.
@@
   The most accurate model of the Palomar~5 stream would involve full modeling of the internal dynamics of the cluster.
-This would mean computing N-Body interactions with a $\mathcal{O}(N^2)$ computation time, stellar evolution, super novae, an initial mass distirbution, treatment of binary star systems, etc.
+This would mean computing $N$-Body interactions with a $\mathcal{O}(N^2)$ computation time, stellar evolution, super novae, an initial mass distirbution, treatment of binary star systems, etc.
 \citep[for such an example][see,]{2021NatAs...5..957G}.

### 2024-12-19 paper/main.tex
alessandra's 3rd round of comments

@@
 \institute{
-  Paris Observatory. Paris Sciences et Lettres
+  Paris Observatory. Paris Sciences et Lettres\\
   \email{salvatore.ferrone@obspm.fr}
   \and
-    Dipartimento di Fisica, Universit\`a di Roma ``La Sapienza'',Piazza Aldo Moro\\
+    Dipartimento di Fisica, Universit\`a di Roma ``La Sapienza'', Piazza Aldo Moro
   \and
@@
 For example, \citet{2010ApJ...718.1128L} made use of the Sagittarius stream to infer that the dark matter halo of our Galaxy has a triaxial shape \citep[but see also][]{2004MNRAS.351..643H, 2005ApJ...619..800J, 2005ApJ...619..807L}; \citet{2016ApJ...833...31B} concluded that the dark matter halo of our Galaxy is nearly spherical at the distances of the Palomar~5 and GD-1 streams.
-The two contrasting results could also show that at one distance is the halo is triaxial and at another spherical showcasing the need for streams at different distances to map the halo shape.
+The two contrasting results could also show that at one distance is the halo is triaxial and at another spherical--highlighting the need for streams at different distances to map the halo shape.
 Recently, \citet{2024ApJ...967...89I} constructed the most accurate model of the Milky Way to date by applying a Monte Carlo Markov Chain (MCMC) fitting procedure.
@@
 This would mean computing $N$-Body interactions with a $\mathcal{O}(N^2)$ computation time, stellar evolution, super novae, an initial mass distirbution, treatment of binary star systems, etc.
-\citep[for such an example][see,]{2021NatAs...5..957G}.
+\citep[for such an example][see,]{2021NatAs...5..957G, 2016MNRAS.458.1450W}.
 Instead, We opt for solving the restricted-three body problem or particle-test method as we did for \citet{2023A&A...673A..44F}, which we present here for completeness.
@@
 \citet{2023A&A...673A.152I} analyzed the orbits of the Galactic cluster population using the same initial conditions as in this work, within five live Milky Way-like potentials from IllustrisTNG \citep{2018MNRAS.473.4077P}.
-They found that in all sampled potentials, orbital changes remain minimal over 5 Gyr, becoming significant only at earlier look-back times when the host galaxy had significantly less mass or was undergoing a merger event.
+They found that in all sampled potentials, orbital changes remain minimal over 5~Gyr, becoming significant only at earlier look-back times when the host galaxy had significantly less mass or was undergoing a merger event.
 
@@
 Note that the masses and sizes of the globular clusters are kept constant in these simulations and not allowed to vary with time.
-This implies that we do not take into account in these models the internal evolution of the globular clusters, nor their response to the Galactic tidal field.
+This equally implies that we do not take into account the internal evolution of the globular clusters in these models.
+We discuss the implications of the modeling limitations in the Discussion.
 
@@
 Note that we do not take into account self-gravity among particles.
-The particles experience the gravitational field yet do not contribute to it.
+The particles experience the gravitational field yet do not contribute to it, which is a common assumption made in Galactic dynamics since the mass of an individual star is negligible compared to the mass of the larger dynamical system.
 The equation of motion of a generic particle among the 100,000 which populate Palomar~5 is thus:
@@
 The time step was therefore set to be small enough to conserve energy for the most interior particles within the cluster---ensuring that a higher mass loss did not arise from numerical error.
-For these simulations, a time-step of 10,000 years was selected to maintain adequate energy conservation of an median variation of $10^{-12} \frac{\Delta E}{E_0}$.
-
-    We also checked the \textit{reverse integrability} of the globular cluster system.
+For these simulations, a time-step of 10,000 years was selected to maintain adequate energy conservation of an median variation of $10^{-12} \frac{\Delta E}{E_0}$, where $E_0$ is a particle's initial energy, and $\Delta E$ is the difference between its final and initial energy.
+
+    We also checked the \textit{reverse integrability} of the globular cluster system for the \texttt{reference} simulations.
 By reverse integrability, we mean the integrator's capability to track the cluster backward in time and then re-integrate it forward along the same trajectory.
-The drift in the final position after forward integration, compared to the initial position from the backward integration, was consistently at least two orders of magnitude smaller than the Plummer scale radius used for Palomar~5.
-This high precision ensures that no fictitious evil numerical forces influenced the system, preventing any artificial mass loss or retention of star particles.
+Integrating point masses in a static axis-symmetric potential conserves $L_z$ and $E$, which create regular periodic and non-chaotic orbits.
+Therefore, any drift would arise from purely numerical error.
+We selected a time stamp for which the drift in the final position after forward integration, compared to the initial position from the backward integration, was consistently at least two orders of magnitude smaller than the Plummer scale radius used for Palomar~5.
+This high precision ensures that no fictitious numerical forces influenced the system, preventing any artificial mass loss or retention of star particles.
 
@@
 The next three panels depict simulations with identical initial conditions but exclude the gravitational influence of all clusters except those forming a given gap.
-The \texttt{Reference} simulation omits other globular clusters.
-The bottom panel compares Palomar5's observed stream length to the \texttt{Reference} simulation, using the same Monte Carlo realization as Fig.~\ref{fig:stream_on_sky} and \texttt{Sampling~009} of Fig.~\ref{fig:gallery2}.}
+The \texttt{Reference} simulation omits the influence from other globular clusters.
+The bottom panel compares Palomar~5's observed stream length to the \texttt{Reference} simulation, using the same Monte Carlo realization as Fig.~\ref{fig:stream_on_sky} and \texttt{Sampling~009} of Fig.~\ref{fig:gallery2}.}
       \label{fig:decomposition}
@@
 Fig.~\ref{fig:decomposition} shows a comparison between one of the 50 realizations of Palomar~5 stream taking into account the gravitational interactions with all other globular clusters in the Galaxy (top panel) and omitting them (bottom panel).
-This comparison clearly shows the presence of two wide ($\sim$ 100~pc and $\sim$ 1~kpc) gaps in the leading tail and of a more subtle under-density at $x_{\textrm{tail}}\sim 5$~kpc (we refer the reader to Appendix~\ref{sec:gap_detection} for a detailed description of the under-densities and gaps detection method).
+This comparison clearly shows the presence of two wide ($\sim$100~pc and $\sim$1~kpc) gaps in the leading tail and of a more subtle under-density at $x_{\textrm{tail}}\sim 5$~kpc (we refer the reader to Appendix~\ref{sec:gap_detection} for a detailed description of the under-densities and gaps detection method).
 
@@
 The use of $\tau$ is advantageous because the growth of the stream is approximately linear in $\tau$.
-On the other hand, in physical space streams on eccentric orbits indeed grow in time yet the growth rate is modulated with periodic expansion and contraction depending on the orbital phase (see the top panel of Fig.~5 from \citet{2016MNRAS.457.3817S} for an example).
+On the other hand, in physical space streams on eccentric orbits indeed grow in time yet the growth rate is modulated with periodic expansion and contraction depending on the orbital phase ( \citep[see the top panel of Fig.~5.][for an example]{2016MNRAS.457.3817S}).
 Adopting this time space, and reporting the gravitational acceleration along Palomar~5 orbit in this space, the identification of the globular clusters which produced the perturbation, and the time at which it occurred becomes straight forward.
@@
 The results are shown in Fig.~\ref{fig:gapcreationrate}.
-This is roughly consistent with a simple estimate of the average gap creation rate: with 73 gaps over 5 Gyr of integration time for a stream about 20 kpc in length, the naive estimate is approximately 0.015~km~s$^{-1}$~kpc$^{-2} $ (which is roughly equivalent to $0.015~\rm{Gyr^{-1}kpc^{-1}}$).
+This is roughly consistent with a simple estimate of the average gap creation rate: with 73 gaps over 5~Gyr of integration time for a stream about 20 kpc in length, the naive estimate is approximately 0.015~km~s$^{-1}$~kpc$^{-2} $ (which is roughly equivalent to $0.015~\rm{Gyr^{-1}kpc^{-1}}$).
 This is about double the weighted mean gap creation rate of 0.009~km~s$^{-1}$~kpc$^{-2}$.
@@
     While Fig.~\ref{fig:impact_geometry_statistics} demonstrates that mass, relative velocity, or impact parameter alone cannot predict gap formation, one interesting result emerges: no gaps are created from interactions with impact parameters greater than 300~pc.
-The stream widths are roughly 200~pc.
+The stream widths are roughly 200~pc, as seen in Appendix~\ref{sec:gallery_of_gaps}.
 This finding is even more evident when examining the $b$-$M$ plane.
-A series of perturbers at roughly $\sim7 \times 10^5 M_\odot$ highlights NGC~2808's flybys, where all encounters with impact parameters under 200~pc result in gaps, while those beyond this distance do not.
-In other words, massive globulars with masses greater than $10^6 \rm{M}_\odot$ are not effective in producing gaps if they pass close to Palomar~5 stream with an impact parameter $b > 200$~pc.
+A series of perturbers at roughly $\sim8 \times 10^5 M_\odot$ highlights NGC~2808's flybys, where all encounters with impact parameters under 200~pc result in gaps, while those beyond this distance do not.
+In other words, even the most massive globulars with masses greater than about $10^6 \rm{M}_\odot$ cannot cause gaps if there impact parameters are greater than about $300$~pc.
 It is interesting to note that for values of the impact parameter below this threshold, even fast encounters ($\delta v > 300$~km/s) can produce gaps.
+Perhaps, this is not surprising since the range of possible relative velocities is much less than mass and impact parameter which vary two and three orders, respectively while the relative velocities only vary by about a factor of three.
 
@@
 
-    Once all the key impact parameters estimated, we can use Eq~\ref{eq:change_in_momentum} and calculate $\Delta p$, the change in momentum (per unit mass) imparted by a cluster flyby on Palomar~5 stream.
+    Once all the key impact parameters estimated, we can use Eq.~\ref{eq:change_in_momentum} and calculate $\Delta p$, the change in momentum (per unit mass) imparted by a cluster flyby on Palomar~5 stream.
 The result  is shown in Fig.~\ref{fig:deltap}, where the distribution of imparted in change in momentum is shown for all impacts which produce a gap, and compared to those that do not produce one.
@@
     Another possible explanation is the different thickness of the simulated tails at different distances from the center of the cluster.
-At distances of less than about 3~kpc from Palomar~5, the tails are very thin, with a typical thickness being less than 100 pc.
+At distances of less than about 3~kpc from Palomar~5, the tails are very thin, with a typical thickness being less than 100~pc.
 In order to cause gaps in these regions, the clusters would have to pass close to the stream, with impact parameters similar to the thickness itself.
 Note that the thinness of the tails at these distances is probably a direct consequence of the initial parameters we chose for the simulation.
-We have assumed for Palomar~5, 5 Gyr years ago, the same internal parameters (mass and size) that the cluster has today.
-With such parameters, after 5 Gyr of evolution, our system has lost most of its mass, and therefore the part of the tails closest to the cluster itself, whose density depends mainly on the most recent mass loss \citep[see, for example, Fig.~A.3 in][]{2012A&A...546L...7M}, is necessarily very thin because the simulated cluster has essentially no more mass to lose.
+We have assumed for Palomar~5, 5~Gyr years ago, the same internal parameters (mass and size) that the cluster has today.
+With such parameters, after 5~Gyr of evolution, our system has lost most of its mass, and therefore the part of the tails closest to the cluster itself, whose density depends mainly on the most recent mass loss \citep[see, for example, Fig.~A.3 in][]{2012A&A...546L...7M}, is necessarily very thin because the simulated cluster has essentially no more mass to lose.
 This last point has also a consequence in the gap creation rate, which we derived at the end of Sect.~\ref{sect:geometry}: with only 70\% of the tail  ($\sim$ 14~kpc over 20~kpc, excluding the innermost $\pm 3$~kpc from the cluster center) really suitable for forming gaps, the gap creation rates are about 50$\%$ higher than those estimated in the aforementioned section, where values have been derived taking into account the full tail extent.
@@
   Our simulations produce a stream for Palomar~5 that is longer than the currently observed extent.
-In simulations, stream detections are straightforward because we can use reference runs to clearly separate stream particles from the field and compare against a known "true" structure.
+In simulations, stream detections are straightforward because we can use reference runs to clearly separate stream particles from the field and compare against a known ``true'' structure.
 In contrast, observational data are inherently more challenging due to magnitude limits, contamination from field stars, and the lack of a ground truth for comparison.

### 2024-12-19 paper/main.tex
typos

@@
 This would mean computing $N$-Body interactions with a $\mathcal{O}(N^2)$ computation time, stellar evolution, super novae, an initial mass distirbution, treatment of binary star systems, etc.
-\citep[for such an example][see,]{2021NatAs...5..957G, 2016MNRAS.458.1450W}.
-Instead, We opt for solving the restricted-three body problem or particle-test method as we did for \citet{2023A&A...673A..44F}, which we present here for completeness.
+\citep[for such an example, see][]{2021NatAs...5..957G, 2016MNRAS.458.1450W}.
+Instead, we opt for solving the restricted-three body problem or particle-test method as we did for \citet{2023A&A...673A..44F}, which we present here for completeness.
 

### 2024-12-19 paper/main.tex
applied aspell!

@@
   The most accurate model of the Palomar~5 stream would involve full modeling of the internal dynamics of the cluster.
-This would mean computing $N$-Body interactions with a $\mathcal{O}(N^2)$ computation time, stellar evolution, super novae, an initial mass distirbution, treatment of binary star systems, etc.
+This would mean computing $N$-Body interactions with a $\mathcal{O}(N^2)$ computation time, stellar evolution, super novae, an initial mass distribution, treatment of binary star systems, etc.
 \citep[for such an example, see][]{2021NatAs...5..957G, 2016MNRAS.458.1450W}.
@@
       \caption{Density maps of Palomar~5's stream in the tail coordinate system.
-The colorscale represents normalized particle counts (total: 100,000).
+The color scale represents normalized particle counts (total: 100,000).
 The top panel shows the \texttt{full} simulation with three gaps on the stream's right-hand side.
@@
       \caption{The distribution and relationship between the impact variables from Eq.\ref{eq:change_in_momentum} for all close fly-bys considered.
-The encounteres that cause gaps are labeled with the same colors as from Fig.~
+The encounters that cause gaps are labeled with the same colors as from Fig.~
       \ref{fig:histogram_impact_time} in the scatter plot, and binned in red for the histograms. The close encounters that did not lead to gaps are shown in grey. }
@@
 The gray-scale represents all 50 stacked orbits, with $\mathcal{P}$ indicating the probability of  Palomar~5's position, normalized to $\mathcal{P}_\textrm{max}=1$.
-Markers show the perturber's position during each impact.
+Markers show the perturbers' positions during each impact.
 The 50 white dots indicates Palomar~5's sampled initial conditions for the current day position.
-The colored markers indicate where the pertuber was located when it impact the stream and \textit{not} its present day position.
+The colored markers indicate where the perturber was located when it impact the stream and \textit{not} its present day position.
 \textit{Bottom Right:} same as bottom left but in the meridional plane.
@@
 
-    We briefly compare our simulated gaps to the literature on Paloamr~5.
+    We briefly compare our simulated gaps to the literature on Palomar~5.
 In the bottom panel of Fig.~\ref{fig:decomposition}, we compare the tracks of Palomar~5 that were compiled in Galstreams by \citet{2023MNRAS.520.5225M}.
@@
 In the maximum limit many gaps could appear, at a rate of about one per realization.
-However, at the shortest reported stream-legnth only a few gaps occur.
+However, at the shortest reported stream-length only a few gaps occur.
       \begin{figure}
@@
 If we naively use the same radial dependence of gaps from Palomar~5 for GD-1 by swapping a pericenter from 6~kpc to 16~kpc, we would reduced the number of gap-favorable impacts of more than a factor of 2.
-However, we note that \citet{2022ApJ...941..129D} preselect the globular clusters which could have experienced a close encounter with the GD-1 stream, by selecting only clusters which pass at a distance less than 0.5 kpc from the stream, having a relative velocity smaller than  300 km/s.
+However, we note that \citet{2022ApJ...941..129D} pre-select the globular clusters which could have experienced a close encounter with the GD-1 stream, by selecting only clusters which pass at a distance less than 0.5 kpc from the stream, having a relative velocity smaller than  300 km/s.
 As shown in our Fig.~\ref{fig:impact_geometry_statistics}, bottom-left panel, in the case of Palomar~5 this choice would lead to exclude most of the encounters favorable to gap creation, which turn out to have relative velocities above 300 km/s.
@@
 If a gap is present, it is indicated by a red arrow with the responsible globular cluster.
-For the gaps that are not easily visible, we show the 1D profiles  showing the significant underdensities compared to the \texttt{reference} simulations.
+For the gaps that are not easily visible, we show the 1D profiles  showing the significant under-densities compared to the \texttt{reference} simulations.
 For instance, the second and last panels of Fig.~\ref{fig:gallery0}
@@
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-031.png}
-      \caption{Eigth gap gallery}
+      \caption{Eighth gap gallery}
       \label{fig:gallery7}

### 2025-01-13 paper/main.tex
some nonsense
recompiled
fixed spaces between Fig and \ref

@@
     \label{fig:stream_on_sky}
-    \end{figure*}
+    \end{figure*}gfitf
 
@@
 This is about double the weighted mean gap creation rate of 0.009~km~s$^{-1}$~kpc$^{-2}$.
-The naive estimate is higher because it does not account for the growth of the stream over time, unlike Eq.\ref{eq:gap_creation_rate}.
+The naive estimate is higher because it does not account for the growth of the stream over time, unlike Eq.~\ref{eq:gap_creation_rate}.
 
@@
       \includegraphics[width=\linewidth]{gap_creation_rate.png}
-      \caption{The distribution of the number of gaps normalized over the total integration time and unit stream length, as described by Eq.\ref{eq:gap_creation_rate} for the whole set of 50 \texttt{full} simulations.
+      \caption{The distribution of the number of gaps normalized over the total integration time and unit stream length, as described by Eq.~\ref{eq:gap_creation_rate} for the whole set of 50 \texttt{full} simulations.
 }
@@
       \includegraphics[width=\linewidth]{impact_geometry_statistics.png}
-      \caption{The distribution and relationship between the impact variables from Eq.\ref{eq:change_in_momentum} for all close fly-bys considered.
+      \caption{The distribution and relationship between the impact variables from Eq.~\ref{eq:change_in_momentum} for all close fly-bys considered.
 The encounters that cause gaps are labeled with the same colors as from Fig.~
@@
     For each of the 50 simulations, we compare the final snapshots between the \texttt{reference} and \texttt{full} simulations using the tail coordinate system, as shown in Fig.~\ref{fig:TailCoordinates}.
-The \texttt{reference} simulations were only saved at the final timestamps, thus from Eq.\ref{eq:data_volume_estimate} where $N_ts$=1, leading to a data volume of about one hundred messily megabytes.
+The \texttt{reference} simulations were only saved at the final timestamps, thus from Eq.~\ref{eq:data_volume_estimate} where $N_ts$=1, leading to a data volume of about one hundred messily megabytes.
 We inspect these differences by generating 2D density maps of the streams.

### 2025-01-13 paper/main.tex
eliminated typos in methods found by Rodrigo

@@
 \abstract
-  %  menntion that only one galactic model should be sufficient to avoid the criticism
+  %  menntion that only one Galactic model should be sufficient to avoid the criticism
   {Thin stellar streams, such as those resulting from the tidal disruption of globular clusters, have long been known and used as probes of the gravitational potential of our Galaxy, both its visible and dark contents.
@@
 
-  % Beyond the galactic scale, dark matter is essential for explaining many cosmological phenomena.
+  % Beyond the Galactic scale, dark matter is essential for explaining many cosmological phenomena.
 For instance, the Planck mission \citep{2020A&A...641A...1P} was designed to study the Cosmic Microwave Background (CMB), whose spectrum closely follows that of a black body with a peak temperature in the millimeter wavelength range, corresponding to approximately 2.725~K.
@@
 
-  % As laid out in \citet{2021PrPNP.11903865A}'s review, despite significant progress, the true nature of dark matter remains elusive and scientists continue to search for evidence of its existence beyond its gravitational effects on cosmic and galactic scales.
-A key focus today is detecting dark matter on sub-galactic scales, because at these scales different cosmological scenarios predict different lumpiness of the halos.
+  % As laid out in \citet{2021PrPNP.11903865A}'s review, despite significant progress, the true nature of dark matter remains elusive and scientists continue to search for evidence of its existence beyond its gravitational effects on cosmic and Galactic scales.
+A key focus today is detecting dark matter on sub-Galactic scales, because at these scales different cosmological scenarios predict different lumpiness of the halos.
 
@@
 For example, \citet{2011MNRAS.417..198V} demonstrated how stellar streams can infer the mass and scale parameters of dark matter halos using various amounts of observational data, from basic right ascension and declination to full six-dimensional phase space information.
-\citet{2018ApJ...867..101B} reviewed this concept from an information-theoretic point of view, identifying which orbits and configurations of stellar streams yield the most information about the galactic potential.
+\citet{2018ApJ...867..101B} reviewed this concept from an information-theoretic point of view, identifying which orbits and configurations of stellar streams yield the most information about the Galactic potential.
 However, the use of single streams for constraining the potential led to some ambiguous and not converging results over time.
@@
 This does not necessarily need to be the general case, especially for streams of clusters as Palomar~5, which are confined in the inner 20~kpc of the Galaxy, where many other globular clusters do also orbit.
-For example, \citet{2023A&A...678A..69I} showed that globular cluster can even collide with other clusters, which implies that cluster stream collisions should happen much more frequently since streams are far more extended than clusters.
+For example, \citet{2023A&A...678A..69I} showed that a globular cluster can even collide with other clusters, which implies that cluster stream collisions should happen much more frequently since streams are far more extended than clusters.
 
@@
     \includegraphics[width=\linewidth]{stream_on_sky_Pal5_monte-carlo-009_pouliasis2017pii-GCNBody_pouliasis2017pii.png}
-    \caption{Simulated Palomar~5 stream created by modeling the host cluster as a Plummer sphere disrupting within an axis-symmetric galactic potential plus the gravitational effect of 164 other galactic globular clusters.
+    \caption{Simulated Palomar~5 stream created by modeling the host cluster as a Plummer sphere disrupting within an axis-symmetric Galactic potential plus the gravitational effect of 164 other galactic globular clusters.
 The top panel shows the distribution of star-particles that escaped the cluster due to tidal forces.
@@
     \label{fig:stream_on_sky}
-    \end{figure*}gfitf
+    \end{figure*}
 
@@
   The most accurate model of the Palomar~5 stream would involve full modeling of the internal dynamics of the cluster.
-This would mean computing $N$-Body interactions with a $\mathcal{O}(N^2)$ computation time, stellar evolution, super novae, an initial mass distribution, treatment of binary star systems, etc.
+This would mean computing $N$-Body interactions with a $\mathcal{O}(N^2)$ computation time, stellar evolution, supernovae, an initial mass distribution, treatment of binary star systems, etc.
 \citep[for such an example, see][]{2021NatAs...5..957G, 2016MNRAS.458.1450W}.
-Instead, we opt for solving the restricted-three body problem or particle-test method as we did for \citet{2023A&A...673A..44F}, which we present here for completeness.
+Instead, we opt for solving the \textit{restricted-three body problem}, or as known as the \texttt{particle-test method} as we did for \citet{2023A&A...673A..44F}, which we describe here for completeness.
 
@@
 
-    For the galactic potential, we used the second model from \citet{2017A&A...598A..66P}, a superposition of a thin disk, thick disk, and dark matter halo, with masses and scale lengths provided in Table~1 of \citet{2023A&A...673A..44F}.
+    For the Galactic potential, we used the second model from \citet{2017A&A...598A..66P}, a superposition of a thin disk, thick disk, and dark matter halo, with masses and scale lengths provided in Table~1 of \citet{2023A&A...673A..44F}.
 This model is time-independent throughout our simulations.
-This model satisfies a series of observational constraints such as local solar stellar density, the galactic rotation curve, etc, similarly to other galactic models such as \texttt{MWpotential2014} from \citet{2015ApJS..216...29B} and \texttt{McMillian2017} from \citet{2017MNRAS.465...76M}.
-However, to balance data volume and computation time, we use only one galactic potential, which should suffice.
+This model satisfies a series of observational constraints such as local solar stellar density, the Galactic rotation curve, etc, similarly to other Galactic models such as \texttt{MWpotential2014} from \citet{2015ApJS..216...29B} and \texttt{McMillian2017} from \citet{2017MNRAS.465...76M}.
+However, to balance data volume and computation time, we use only one Galactic potential, which should suffice.
 \citet{2021MNRAS.505.5978V} found that only a few outer globular clusters are strongly affected by different potential models.
@@
 Similarly, \citet{2024MNRAS.528.5189G} generated a globular cluster mass-loss catalog using seven different potential models and found their debris distributions to be rather model-independent, similarly to \citet{2023A&A...673A..44F}.
-While the clusters' positions may depend on the model, we assert that interaction rates and stream formation are unaffected by the choice of galactic potential model.
+While the clusters' positions may depend on the model, we assert that interaction rates and stream formation are unaffected by the choice of Galactic potential model.
 
@@
 \citet{2023A&A...673A.152I} analyzed the orbits of the Galactic cluster population using the same initial conditions as in this work, within five live Milky Way-like potentials from IllustrisTNG \citep{2018MNRAS.473.4077P}.
-They found that in all sampled potentials, orbital changes remain minimal over 5~Gyr, becoming significant only at earlier look-back times when the host galaxy had significantly less mass or was undergoing a merger event.
+They found that in all sampled potentials, orbital changes remain minimal over 5~Gyr, becoming significant only at earlier look-back times when the host galaxy had either significantly less mass or was undergoing a merger event.
 
@@
           \ddot{\vec{r}}_i = -\nabla \Phi(R_i,z_i) + \left.\sum_{j\neq i}^{N_{GC}} \frac{Gm_j}{\left(|\vec{r}_j - \vec{r}_i|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j - \vec{r}_i\right),
-          \end{equation}\label{eq:GCNBody} where $\vec{r}$ indicates the galactocentric position vector, the index $i$ indicates the globular cluster of interest with $R_i,z_i$ being its galactocentric positions in cylindrical coordinates; the index $j$ indicates the other globular clusters that are summed over.
+          \end{equation}\label{eq:GCNBody} 
+          \noindent where $\vec{r}$ indicates the galactocentric position vector, the index $i$ indicates the globular cluster of interest with $R_i,z_i$ being its galactocentric positions in cylindrical coordinates; the index $j$ indicates the other globular clusters that are summed over.
 $N_{GC}$ is the total number of globular clusters which in this study is 165, $m_j$ is the mass of the j-th cluster in the sample, $b_j$ is its Plummer scale radius, and $\vec{r_j}$ is its galactocentric position.
-$\Phi$ represents the Galactic smooth potential \citep[][Model~II, in the present case]{2017A&A...598A..66P}.
+$\Phi$ represents the same Galactic smooth potential that we discussed previously\citep[][Model~II, in the present case]{2017A&A...598A..66P}.
 Note that the masses and sizes of the globular clusters are kept constant in these simulations and not allowed to vary with time.
@@
 
-        \item Once the positions and velocities of the entire globular cluster system, as to 5~Gyr ago, are found, we sample Palomar~5 with  100,000 particles which follow a Plummer distribution of mass and half-mass radius as in the Baumgardt catalog.
+        \item Once the positions and velocities of the entire globular cluster system 5~Gyr ago are found, we sample Palomar~5 with 100,000 particles from a Plummer distribution taking the mass and half-mass radius from the Baumgardt catalog.
 We then integrate the evolution of these particles forward in time to the present day, by taking into account that each particle feels the gravitational potential of the Galaxy, that of its host cluster, and that of all the other clusters in the Galaxy.
@@
         The procedure described so far has been repeated 50 times, each time generating a new set of initial conditions, given the uncertainties on proper motions, line-of-sight velocities, distances to the Sun, and masses, of all clusters, as reported in the Baumgardt catalog.
-We handle these uncertainties through a Monte-Carlo approach, by sampling them with Gaussian distribution, and by taking into account the covariance term between the proper motions.
+We handle these uncertainties through a Monte-Carlo approach, by sampling them with a Gaussian distribution, and by taking into account the covariance term between the proper motions.
 For the first simulation, we use the most probable values for the initial conditions.
-The uncertainties are sample for all globular clusters.
+The uncertainties are sampled for all globular clusters.
 Additionally, for each re-sampling of Palomar~5's mass, we also resample the distribution of the 100,000 star-particles.
@@
     To quantify the impact of globular cluster passages to the density of Palomar~5 stream, we performed a second set of simulations, which we refer to as the \texttt{reference} simulations in this paper.
-These \texttt{reference} simulations use the same 50 sets of initial conditions as of the \texttt{full} simulations, the same galactic potential but exclude mutual interactions between globular clusters.
+These \texttt{reference} simulations use the same 50 sets of initial conditions as the \texttt{full} simulations, the same Galactic potential but exclude mutual interactions between globular clusters.
 The approach adopted for this second set of simulations is thus equivalent to that adopted already in \citet{2023A&A...673A..44F}.
@@
 The time step was therefore set to be small enough to conserve energy for the most interior particles within the cluster---ensuring that a higher mass loss did not arise from numerical error.
-For these simulations, a time-step of 10,000 years was selected to maintain adequate energy conservation of an median variation of $10^{-12} \frac{\Delta E}{E_0}$, where $E_0$ is a particle's initial energy, and $\Delta E$ is the difference between its final and initial energy.
+For these simulations, a time-step of 10,000 years was selected to maintain adequate energy conservation, with a median variation of $10^{-12} \frac{\Delta E}{E_0}$, where $E_0$ is a particle's initial energy, and $\Delta E$ is the difference between its final and initial energy.
 
@@
 A particularly clear example of this effect is shown in Fig.~\ref{fig:stream_on_sky} which we selected for its prominent gaps.
-Two of these gaps are clearly visible in galactic coordinates and become even more apparent when marginalizing over latitude to reconstruct the 1D density profile of the stream, as a function of longitude.
+Two of these gaps are clearly visible in Galactic coordinates and become even more apparent when marginalizing over latitude to reconstruct the 1D density profile of the stream, as a function of longitude.
 
@@
 It is worth noting that the times at which the passages of these clusters occurred, according to our analysis, are in agreement with the observed width of the corresponding gaps: the encounter with NGC~2808 being very recent (only 200~Myr ago) its induced gap is still very thin, because it takes time for a perturbation to grow into an extended gap, as it is the case for those induced by the passages of  NGC~7078 and NGC~104, which occurred in earlier times.
-It is also interesting to emphasize that gaps as thin as those generated by the passage of NGC~2808, 200~Myr ago, can be detected by working in the tail coordinate system and by making a comparative analysis (\texttt{full} versus \texttt{reference} simulations): they are so thin that they cannot be directly identified in galactic coordinates (see Fig.~\ref{fig:stream_on_sky}).\\
+It is also interesting to emphasize that gaps as thin as those generated by the passage of NGC~2808, 200~Myr ago, can be detected by working in the tail coordinate system and by making a comparative analysis (\texttt{full} versus \texttt{reference} simulations): they are so thin that they cannot be directly identified in Galactic coordinates (see Fig.~\ref{fig:stream_on_sky}).\\
 
@@
 Apriori, since the number of star-particles escape at similar rates from the L1 and L2 Lagrange points, each tail is of similar length and density.
-The main difference between the two tails is that the leading tail is closer to the galactic center than the trailing tail by about 400~pc.
+The main difference between the two tails is that the leading tail is closer to the Galactic center than the trailing tail by about 400~pc.
 Since the lengths are equal, and the offset between the tails is small compared to the galactocentric distance of about 10 kpc, we expected the number of gaps in each tail to be more or less the same.
@@
 However, when taken all together, the location of gap creating impacts shows a strong negative correlation with the galactocentric radius $r$, with the number $N$ of encounters favorable to gap formation  going as $N = -2.5r + 50$ (with a Pearson coefficient of -0.86).
-While there are more clusters near the galactic center, clusters naturally spend more time near their apocenters, and the lower relative velocities in these regions should favor gap creation.
-However, this result suggests that cluster galactic number density outweighs these factors in determining the number of gaps.
+While there are more clusters near the Galactic center, clusters naturally spend more time near their apocenters, and the lower relative velocities in these regions should favor gap creation.
+However, this result suggests that cluster Galactic number density outweighs these factors in determining the number of gaps.
 It should be noted that these results are for Palomar~5 only.
@@
 This would help identify regimes where baryonic and dark matter-induced substructures can be disentangled, shedding light on the elusive influence of dark matter sub-halos on stellar streams.
-Additionally, we can continue to add realistic physics to the galactic environment such as the merger of the Sagittarius dwarf galaxy and the passages of the large and small Magellanic clouds.
+Additionally, we can continue to add realistic physics to the Galactic environment such as the merger of the Sagittarius dwarf galaxy and the passages of the large and small Magellanic clouds.
 

### 2025-01-13 paper/main.tex
rodrigo's corrects are ingested

@@
 
-    We see clear effects of the presence of their gravitational field in the properties of Palomar~5 stream.
+    The presence of the other globular clusters has a clear effect on the properties of Palomar~5 stream.
 A particularly clear example of this effect is shown in Fig.~\ref{fig:stream_on_sky} which we selected for its prominent gaps.
@@
 The use of $\tau$ is advantageous because the growth of the stream is approximately linear in $\tau$.
-On the other hand, in physical space streams on eccentric orbits indeed grow in time yet the growth rate is modulated with periodic expansion and contraction depending on the orbital phase ( \citep[see the top panel of Fig.~5.][for an example]{2016MNRAS.457.3817S}).
+On the other hand, in physical space streams on eccentric orbits indeed grow in time yet the growth rate is modulated with periodic expansion and contraction depending on the orbital phase \citep[see the top panel of Fig.~5.][for an example]{2016MNRAS.457.3817S}.
 Adopting this time space, and reporting the gravitational acceleration along Palomar~5 orbit in this space, the identification of the globular clusters which produced the perturbation, and the time at which it occurred becomes straight forward.
@@
     To further probe that the three clusters above are responsible of producing the gaps observed in this simulation, we conducted additional experiments by including only the perturbation of each of these three clusters at a time, and neglecting the gravitational perturbations of all the other clusters in the Galaxy.
-The results of these simulations are reported in the middle panels of Fig.~\ref{fig:decomposition} and clearly shows that NGC~2808, NGC~7078 and NGC~104 are the clusters responsible for creating the under-dense regions observed in the leading tails of Palomar~5.
+The results of these simulations are reported in the middle panels of Fig.~\ref{fig:decomposition} and clearly show that NGC~2808, NGC~7078 and NGC~104 are the clusters responsible for creating the under-dense regions observed in the leading tails of Palomar~5.
 It is worth noting that the times at which the passages of these clusters occurred, according to our analysis, are in agreement with the observed width of the corresponding gaps: the encounter with NGC~2808 being very recent (only 200~Myr ago) its induced gap is still very thin, because it takes time for a perturbation to grow into an extended gap, as it is the case for those induced by the passages of  NGC~7078 and NGC~104, which occurred in earlier times.
@@
 
-    From this analysis we can derive a statistical view of: 1) the number of gaps generated on Pal~5 stream by the whole system of Galactic globular clusters; 2) the clusters which generated these gaps and the time history of these perturbations.
-From this, we can then quantify: 3) the properties of the perturbers (their masses, sizes and their orbital parameters) as well as 4) the impact geometry of the encounters, which allows us to understand which encounters are more favorable to generate gaps in Pal~5 stream.
+    From this analysis we can derive a statistical view of: 1) the number of gaps generated on the Palomar~5 stream by the whole system of Galactic globular clusters; 2) the clusters which generated these gaps and the time history of these perturbations.
+From this, we can then quantify: 3) the properties of the perturbers (their masses, sizes and their orbital parameters) as well as 4) the impact geometry of the encounters, which allows us to understand which encounters are more favorable to generate gaps in the Palomar~5 stream.
 In the following, we will present results of this analysis, by starting from addressing points 1) and 2).
@@
 
-    By applying the methodology described above to the whole set of simulations, we can reconstruct the history of close passages of Galactic globular clusters to Palomar~5's stream, in the last 5~Gyr, which -- we recall the reader -- is the time interval investigated in our simulations.
+    By applying the methodology described above to the whole set of simulations, we can reconstruct the history of close passages of Galactic globular clusters to Palomar~5's stream, in the last 5~Gyr, which -- we remind the reader -- is the time interval investigated in our simulations.
 The results of this analysis are reported in Fig.~\ref{fig:histogram_impact_time} and in Table~\ref{tab:gaps_per_perturber}.
 NGC~2808 impacts Palomar~5's stream about 200 Myr ago in 44 out of 50 simulations, creating a small gap at a similar position to the one reported in Fig.~\ref{fig:decomposition} in each case.
-Since this interaction occurred less than one orbital period ago, the uncertainties that lead to orbital solutions remain that similar and thus producing consistent results across all simulations.
+Since this interaction occurred less than one orbital period ago, despite the uncertainties the orbital solutions remain nontheless similar and thus produce consistent results across all simulations.
 However, as we continue to turn back time further, the uncertainties in the initial conditions allow the for the various orbital solutions to diverge from one another.
@@
 
-    We need a more sophisticated statistic that can be compared to other simulations and turn to the \textit{gap creation rate} developed by \citet{2012ApJ...748...20C}.
+    We need a more sophisticated statistic that can be compared to other simulations, so we turn to the \textit{gap creation rate} developed by \citet{2012ApJ...748...20C}.
 The gap creation rate is the number of gaps that appear per time and normalized by the length of the stream.
@@
 The encounters that cause gaps are labeled with the same colors as from Fig.~
-      \ref{fig:histogram_impact_time} in the scatter plot, and binned in red for the histograms. The close encounters that did not lead to gaps are shown in grey. }
+      \ref{fig:histogram_impact_time} in the scatter plot, and binned in red for the histograms. The close encounters that did not lead to gaps are shown in black. }
       \label{fig:impact_geometry_statistics}    
@@
 
-    With the perturbers identified, we perform statistics in the pursuit of understanding what conditions are necessary for a globular cluster in order to form a gap on Palomar~5 stream.
+    With the perturbers identified, we perform statistics in the pursuit of understanding what conditions are necessary for a globular cluster to induce a gap on the Palomar~5 stream.
 We turn to impact theory, which in its simplest form is presented in works such as \citet{2008gady.book.....B}.
@@
 A series of perturbers at roughly $\sim8 \times 10^5 M_\odot$ highlights NGC~2808's flybys, where all encounters with impact parameters under 200~pc result in gaps, while those beyond this distance do not.
-In other words, even the most massive globulars with masses greater than about $10^6 \rm{M}_\odot$ cannot cause gaps if there impact parameters are greater than about $300$~pc.
+In other words, even the most massive globulars with masses greater than about $10^6 M_\odot$ cannot cause gaps if their impact parameters are greater than about $300$~pc.
 It is interesting to note that for values of the impact parameter below this threshold, even fast encounters ($\delta v > 300$~km/s) can produce gaps.
@@
       \caption{\textit{Characteristics of gap causing clusters} \textit{Top Left:} Energy-angular momentum space of the globular clusters in the simulations, with 50~$\times$~165 data points representing all sampled initial conditions.
-Clusters impacting Palomar~5 are shown with colored markers; non-gap-causing clusters are shown in gray.
-Palomar~5 is marked with a white symbol and black edge.
+Clusters impacting Palomar~5 are shown with colored markers; large for the samplins that induce a gap and small for those that do not.
+Non-gap-causing clusters are shown in small gray dots.
+The 50 white dots indicate Palomar~5's sampled initial conditions for the current day position.
 The circular velocity-curve is drawn with a light-pink dashed curve.
@@
 The gray-scale represents all 50 stacked orbits, with $\mathcal{P}$ indicating the probability of Palomar~5's position, normalized to $\mathcal{P}_\textrm{max}=1$.
-Markers show the perturbers' positions during each impact.
-The 50 white dots indicates Palomar~5's sampled initial conditions for the current day position.
-The colored markers indicate where the perturber was located when it impact the stream and \textit{not} its present day position.
+The colored markers indicate where the perturber was located when it impacted the stream and \textit{not} its present day position.
 \textit{Bottom Right:} same as bottom left but in the meridional plane.
@@
     Once all the key impact parameters estimated, we can use Eq.~\ref{eq:change_in_momentum} and calculate $\Delta p$, the change in momentum (per unit mass) imparted by a cluster flyby on Palomar~5 stream.
-The result  is shown in Fig.~\ref{fig:deltap}, where the distribution of imparted in change in momentum is shown for all impacts which produce a gap, and compared to those that do not produce one.
+The result  is shown in Fig.~\ref{fig:deltap}, where the distribution of imparted change in momentum is shown for all impacts which produce a gap, and compared to those that do not produce one.
 On average, encounters which lead to gap creations impart a change in momentum on stream particles which is a factor of 10 higher than that of encounters which do not form gaps (but with some overlap in the low-velocity tail).
@@
 half-mass radius) of the clusters that cause gaps on the tails of Palomar~5, dividing them from those that do not.
-As can be seen, no cluster with mass below $10^5 \rm{M}_\odot$ causes gaps on Palomar~5 stream, and all perturbers, except FSR~1758 have a half-mass radius below 10~pc.
+As can be seen, no cluster with mass below $10^5 M_\odot$ causes gaps on the Palomar~5 stream, and all perturbers, except FSR~1758 have a half-mass radius below 10~pc.
 Even more interesting is their distribution in the E-Lz plane, which shows that the clusters that cause gaps are on both direct and retrograde orbits (respectively, negative and positive values of $L_z$), but all confined to an energy interval between $-2$ and $-1 \times10^5~\textrm{km}^2\textrm{s}^{-2}$.
@@
 It should be noted that these results are for Palomar~5 only.
-Streams along various orbits needs to be studied in order for this conclusion to be generalized.
+Streams along various orbits need to be studied in order for this conclusion to be generalized.
 
@@
     We observe that there are no gaps near within 3~kpc to the cluster within our simulation.
-There may be a few reasons for the absence of gaps in portion of the tails closer to the cluster center.
+There may be a few reasons for the absence of gaps in the portion of the tails closer to the cluster center.
 First, as \citet{2016MNRAS.457.3817S} demonstrated, the dispersion of action-frequencies in the stream plays a role.
@@
 This last point has also a consequence in the gap creation rate, which we derived at the end of Sect.~\ref{sect:geometry}: with only 70\% of the tail  ($\sim$ 14~kpc over 20~kpc, excluding the innermost $\pm 3$~kpc from the cluster center) really suitable for forming gaps, the gap creation rates are about 50$\%$ higher than those estimated in the aforementioned section, where values have been derived taking into account the full tail extent.
-This gap creation rate would be still not high enough to reproduce the  number of gaps in Palomar~5 tails, as reported by \citet{2012ApJ...760...75C}.
+This gap creation rate would be still not high enough to reproduce the  number of gaps in Palomar~5's tails, as reported by \citet{2012ApJ...760...75C}.
 In their study, they indeed suggested the presence of  five gaps with a 99~\% detection confidence, leading to a gap creation rate of 0.17~Gyr$^{-1}$~kpc$^{-1}$.
@@
 Moreover, the present-day masses used in our simulations are likely lower than the historical masses of these clusters.
-For example, \citet{2024ApJ...976...54P} conducted a study simulating the dissolution of a realistic globular cluster population, in the aims of identifying how many stellar streams we should expect in the Milky Way and used a mock catalog of globular clusters with an initial amount with masses above 10~$^4 M_\odot$ totaling at about 10,000 clusters.
+For example, \citet{2024ApJ...976...54P} conducted a study simulating the dissolution of a realistic globular cluster population, in the aims of identifying how many stellar streams we should expect in the Milky Way and used a mock catalog of globular clusters with an initial amount with masses above 10$^4 M_\odot$ totaling at about 10,000 clusters.
 This implies that more perturbers could have been present in the past, potentially increasing the frequency and number of gaps in Palomar~5's stream.
@@
 
-  As already suggested in previous works which have studied the impact of baryonic structures on Palomar~5 tails \citep{2017NatAs...1..633P, 2019MNRAS.484.2009B}, it is possible that this cluster lies in a region of the phase-space which is not favorable to distinguish gaps created by dark matter sub-halos from gaps created by baryonic structures, as the Galactic bar and giant molecular clouds.
+  As already suggested in previous works which have studied the impact of baryonic structures on Palomar~5 tails \citep{2017NatAs...1..633P, 2019MNRAS.484.2009B}, it is possible that this cluster lies in a region of the phase-space which is not favorable to distinguish gaps created by dark matter sub-halos from gaps created by baryonic structures, such as the Galactic bar and giant molecular clouds.
 With our study, we show that close encounters with globular clusters constitute a further element that confuses a simple interpretation of the observed Palomar~5 stream gaps (if any).
@@
 Going to lower orbital energies make the situation even worse, because the effect of the Galactic bar, giant molecular clouds and interactions with globular clusters becomes even more efficient.
-In the innermost regions of the Galaxy, where the density of dark matter sub-halos is expected to be maximal, globular cluster streams are intrinsically more difficult to be found, in addition to the fact that the dynamic times in these regions become so small that the stars lost from globular clusters do not redistribute themselves for the most part into thin structures \citep[see][]{2023A&A...673A..44F}.
-It is thus probably only at larger distances from the Galactic center than those spanned by Palomar~5 orbit that the impact of dark matter sub-halos may become dominant, but it should be borne in mind that at these distances the number density of sub-halos also decreases.
+In the innermost regions of the Galaxy, where the density of dark matter sub-halos is expected to be maximal, globular cluster streams are intrinsically more difficult to find, in addition to the fact that the dynamic times in these regions become so small that the stars lost from globular clusters do not redistribute themselves for the most part into thin structures \citep[see][]{2023A&A...673A..44F}.
+It is thus probably only at larger distances from the Galactic center than those spanned by Palomar~5's orbit that the impact of dark matter sub-halos may become dominant, but it should be borne in mind that at these distances the number density of sub-halos also decreases.
 In the future, it will be interesting to apply the type of study conducted here to the whole set of clusters for which streams have been found to quantify the regime in which they are found, whether favorable to the creation of gaps from baryonic structures or not.
@@
     For each of the 50 simulations, we compare the final snapshots between the \texttt{reference} and \texttt{full} simulations using the tail coordinate system, as shown in Fig.~\ref{fig:TailCoordinates}.
-The \texttt{reference} simulations were only saved at the final timestamps, thus from Eq.~\ref{eq:data_volume_estimate} where $N_ts$=1, leading to a data volume of about one hundred messily megabytes.
+The \texttt{reference} simulations were only saved at the final timestamps, thus from Eq.~\ref{eq:data_volume_estimate} where $N_ts$=1, leading to a data volume of about one hundred measly megabytes.
 We inspect these differences by generating 2D density maps of the streams.
@@
 The bin is then replaced with the median.
-We choose to use 10 adjacent points per side, which corresponding to a smoothing length of approximately 1~kpc.
+We choose to use 10 adjacent points per side, which correspond to a smoothing length of approximately 1~kpc.
 This procedure reduces high-frequency noise and smooths the profiles.
@@
 Additionally, for the impacts of NGC~104 and NGC~C7078, wakes can be observed in the density map.
-Another important aspect, is that the strongest peak in gravitational force does not necessarily create a gap.
+Another important aspect is that the strongest peak in gravitational force does not necessarily create a gap.
 Notice how NGC~5272, which was labeled with 0 to indicate that it has the greatest local maxima does not have a gap.
 The reason for this is manifold, for instance, the force needs to be modulated with the time since the change in momentum is the determining factor and not the peak magnitude of the force.
-Additionally, there is an offset of about 200~pc between the stream at the orbit, as seen when viewing the stream in tail coordinates, so peaks upon the orbit are good proxy's for the stream but are not definitive.
+Additionally, there is an offset of about 200~pc between the stream and the orbit, as seen when viewing the stream in tail coordinates, so peaks upon the orbit are good proxies for the stream but are not definitive.
 We found that the top five greatest impacts accounted for all gaps, except for Sampling~014 as shown in Fig.~\ref{fig:gallery3}, whose gap from NGC~6584 corresponded to the 7th peak.
@@
 For instance, if the impact candidate was in the leading tail, only the star particles with $\tau > 0$ are used to constrain the stream track.
-The polynomial coefficients were determined using the through a minimization method using Nelder-Mead algorithm from \texttt{scipy}'s optimization package.
+The polynomial coefficients were determined through a minimization method using the Nelder-Mead algorithm from \texttt{scipy}'s optimization package.
 
@@
 For each of the 50 simulations, this information was compute for the strongest 5 flybys of a perturber with the stream.
-Thus, we compute a sample of 250 impacts, and flag those that give way to gaps.
+Thus, we compute a sample of 250 impacts, and flag those that give rise to gaps.
 We refer the reader to Sect.~\ref{sect:geometry} for the presentation and discussion of the results.

### 2025-01-13 paper/main.tex
added misha's comments

@@
 This method identified the set of potential parameters in an axisymmetric model of the Milky Way that best reproduces all observed stellar streams.
-This work represents a seminal achievement and a significant milestone in the field.
 
@@
 Consequently the system is modeled as being less dense than it should have been.
-In turn, our simulations have a strong initial mass which adds to the amplitudes of the profile density peaks of Fig.~\ref{fig:stream_on_sky}.
+In turn, our simulations have a strong initial mass loss which adds to the amplitudes of the profile density peaks of Fig.~\ref{fig:stream_on_sky}.
 This inaccuracy is acceptable for the scope of this work.
@@
 This is a surprising result.
-Apriori, since the number of star-particles escape at similar rates from the L1 and L2 Lagrange points, each tail is of similar length and density.
+A priori, since the number of star-particles escape at similar rates from the L1 and L2 Lagrange points, each tail is of similar length and density.
 The main difference between the two tails is that the leading tail is closer to the Galactic center than the trailing tail by about 400~pc.
 Since the lengths are equal, and the offset between the tails is small compared to the galactocentric distance of about 10 kpc, we expected the number of gaps in each tail to be more or less the same.
-We can compute the probability of observing the unequal occurrences though the binomial distribution.
+We can compute the probability of observing the unequal occurrences through the binomial distribution.
 First, the since the 44 gaps linked to NGC~2808 are the same fly-by, they are not independent events.
@@
 Since the stream is not zero-dimensional and has length, both the parallel and perpendicular components of the perturber velocity must be considered.
-Consequently, five parameters determine the change in velocity of a given stream particle: $M$, $r_p$, $b$, $W_\parallel$, and $W_\perp$.
+Consequently, five parameters determine the change in velocity of a given stream particle: $M$, $r_p$, $b$, $W_\parallel$, and $W_\perp$, which are the: mass of the perturber, size of the perturber, impact parameter, parallel and perpendicular components of the relative velocity.
 As detailed in Appendix~\ref{sec:reconstruction}, we calculated these parameters for all our \texttt{full} simulations, by selecting -- for each of them --  the strongest 5 flybys of a perturber with the Palomar~5 stream.
@@
 The data set includes the top 5 strongest flybys from each simulation.
-Those that cause gaps are colored while those that do not are left in black.
+Those that cause gaps are colored while those that do not are outlined in grey.
 Note that the meaning of the colors is the same as in Fig.~\ref{fig:histogram_impact_time}.}
@@
 
-  We note that our results seem to be in tension with the conclusions of \citet{2019MNRAS.484.2009B}, who presented a numerical study of Palomar5 tails, orbiting a Milky Way-like potential, where both dark matter sub-halos and baryonic sub-structures (Galactic bar, spiral arms, giant molecular clouds, globular clusters) were taken into account to quantify the importance of these latter in density variations in Pal5 streams.
+  We note that our results seem to be in tension with the conclusions of \citet{2019MNRAS.484.2009B}, who presented a numerical study of Palomar~5 tails, orbiting a Milky Way-like potential, where both dark matter sub-halos and baryonic sub-structures (Galactic bar, spiral arms, giant molecular clouds, globular clusters) were taken into account to quantify the importance of these latter in density variations in Pal5 streams.
 While their methods are extremely similar to ours, their analysis diverges significantly.
@@
     For each of the 50 simulations, we compare the final snapshots between the \texttt{reference} and \texttt{full} simulations using the tail coordinate system, as shown in Fig.~\ref{fig:TailCoordinates}.
-The \texttt{reference} simulations were only saved at the final timestamps, thus from Eq.~\ref{eq:data_volume_estimate} where $N_ts$=1, leading to a data volume of about one hundred measly megabytes.
+The \texttt{reference} simulations were only saved at the final time-stamps, thus from Eq.~\ref{eq:data_volume_estimate} where $N_ts$=1, leading to a data volume of about one hundred measly megabytes.
 We inspect these differences by generating 2D density maps of the streams.
@@
       where $\vec{v}_p$ and $\vec{v}_s$ are the velocities of the perturber and the stream, respectively.
-For each of the 50 simulations, this information was compute for the strongest 5 flybys of a perturber with the stream.
+For each of the 50 simulations, this information was computed for the strongest 5 flybys of a perturber with the stream.
 Thus, we compute a sample of 250 impacts, and flag those that give rise to gaps.
@@
 For the gaps that are not easily visible, we show the 1D profiles  showing the significant under-densities compared to the \texttt{reference} simulations.
-For instance, the second and last panels of Fig.~\ref{fig:gallery0}
+For instance, the second and last panels of Fig.~\ref{fig:gallery0}.
+However, maginalizing the 2D distribution to 1D can erase the signal of oblique gaps.
+Thus, not all gaps that are visible in the 2D space are flagged in the density profiles.
+For example, in \texttt{Sampling 000}, the gaps from NGC~2808 and NGC~5634 are not flagged present in the marginalized density profile.
 

### 2025-01-13 paper/main.tex
full compilation
updated the caption for tail coordinates

@@
       \centering
-      \includegraphics[width=\linewidth]{along_orbit_coordinate_system-approved.png}
-      \caption{Tail coordinates. \textbf{I want to update this}. }
+      \includegraphics[width=\linewidth]{tailcoordinates.png}
+      \caption{A demonstration of the transformation from galactic coordinates to tail coordinates.
+The black dot represents the globular cluster, while the thick black line shows a segment of its orbit.
+A star particle is marked with a star symbol.
+The red vector points from the nearest point on the cluster's orbit to the star particle.
+Its projection onto the galactocentric unit vector is shown in green and labeled $y'$.
+The red vector is also projected onto the vector perpendicular to both the orbital tangent and the galactocentric vector, which defines the local orbital plane and is labeled $z'$.
+The path length along the orbit from the cluster to the nearest point is denoted as $x'$, with $x'$ being positive if the particle is ahead of the cluster.}
       \label{fig:TailCoordinates}

### 2025-01-24 paper/main.tex
paola's pre sub final comments

@@
        \and 
+       Rodrigo Ibata \inst{4}
+       \and 
+       Paolo Bianchini \inst{4}
+       \and
+       Sergey Khoperskov \inst{5}
+       \and
+       Nicolas Leclerc \inst{1}
+       \and 
+       Clement Hottier \inst{1}
+       \and 
+       Eliot Stein  \inst{6}
+       \and
        Misha Haywood \inst{1}
        \and
-       Rodrigo Ibata \inst{4}
-       \and 
-       Paolo Bianchini \inst{4}
-       \and
-       Sergey Khoperskov \inst{5}
-       \and
-       Nicolas Leclerc \inst{1}
-       \and 
-       Clement Hottier \inst{1}
-       \and 
-       Eliot Stein  \inst{1}
-       \and
        David Valls-Gabaud \inst{1}
        \and
-       Owain N. Snaith \inst{6}
+       Owain N. Snaith \inst{7}
        }
@@
 \institute{
-  Paris Observatory. Paris Sciences et Lettres\\
+  LIRA, Observatoire de Paris, Université PSL, Sorbonne Université, Université Paris Cité, CY Cergy Paris Université, CNRS,75014 Paris, France\\
   \email{salvatore.ferrone@obspm.fr}
@@
   \and
+    institute des sciences spatial
+  \and
     Exeter
-    % \thanks{}
   }
@@
   These predictions occurred well before the first detections of Galactic globular cluster tidal tails \citep{1995AJ....109.2553G}.
-Interestingly, \citep{1995AJ....109.2553G}'s detections were made nearly contemporary  with the discovery of the Sagittarius stellar stream by \citet{1994Natur.370..194I} which is the closest example of a stream emerging from a dwarf satellite currently interacting with the Milky Way.
+Interestingly, \citet{1995AJ....109.2553G}'s detections were made nearly contemporary  with the discovery of the Sagittarius stellar stream by \citet{1994Natur.370..194I} which is the closest example of a stream emerging from a dwarf satellite currently interacting with the Milky Way.
 
   Subsequent studies\footnote{For more subsequent observation detections of tidal debris, i.e.
-globular cluster stars beyond the tidal radius see the works of: \citet{1997A&A...320..776L, 2000A&A...356..127T, 2000A&A...359..907L, 2001AAS...19910906S, 2003AJ....126..815L}} extended Grillmair's findings to other globular cluster streams, but were often limited to the detections of stars still close to the cluster tidal radius, until the discovery made by \citet{2001ApJ...548L.165O,2002AAS...200.1001O, 2003AJ....126.2385O} of long and thin tails outside the Palomar~5 globular cluster.
-With a mass of $1.34\pm 0.24 \times 10^4 M_{\sun}$, Palomar~5 is one of the least massive globular clusters in the Galaxy.
+globular cluster stars beyond the tidal radius see the works of: \citet{1997A&A...320..776L, 2000A&A...356..127T, 2000A&A...359..907L, 2001AAS...19910906S, 2003AJ....126..815L}.} extended Grillmair's findings to other globular cluster streams, but were often limited to the detections of stars still close to the cluster tidal radius, until the discovery made by \citet{2001ApJ...548L.165O,2002AAS...200.1001O, 2003AJ....126.2385O} of long and thin tails outside the Palomar~5 globular cluster.
+With a mass of $1.34\pm 0.24 \times 10^4 M_{\sun}$ \citep{2019MNRAS.482.5138B}, Palomar~5 is one of the least massive globular clusters in the Galaxy.
 \citet{2003AJ....126.2385O} showed that its tails contain more mass than the cluster itself.
@@
 The possibility given by Gaia to track stars with coherent movements, over the entirety of the sky, has led to the discovery of dozens of new streams.
-For example, \citet{2018MNRAS.477.4063M} developed the \texttt{streamfinder} algorithm and applied it across a series works of \citep{2018MNRAS.481.3442M,  2018ApJ...865...85I, 2019ApJ...872..152I} to discover the series of streams.%: Slidr, Sylgr, Ylgr, Fimbulthul, Svol, Gjoll, Leiptr, Phlegeton and Fjorm.
+For example, \citet{2018MNRAS.477.4063M} developed the \texttt{streamfinder} algorithm and applied it across a series works \citep{2018MNRAS.481.3442M,  2018ApJ...865...85I, 2019ApJ...872..152I} to discover a multitude of streams.%: Slidr, Sylgr, Ylgr, Fimbulthul, Svol, Gjoll, Leiptr, Phlegeton and Fjorm.
 According to the most recent estimates provided by \citet{2020MNRAS.493.4978S} Fjorm in particular has an extension of more than 150 degrees on the sky which has a projected length more than five times greater than that of Palomar~5.
 
-  The possibility to couple Gaia data with spectroscopic surveys, has extended the study of stellar streams beyond the quantification of their orbital properties, to a full chemical characterization.
+  The possibility to couple Gaia data with spectroscopic surveys has extended the study of stellar streams beyond the quantification of their orbital properties, to a full chemical characterization.
 For example \citet{2019MNRAS.490.3508L} presents $S^5$, and is one example of a cross-catalog among other works \citep{2020AJ....160..181J, 2021ApJ...911..149L, 2022ApJ...928...30L, 2024MNRAS.529.2413U}.
-Currently, there are about a hundred stellar streams are currently known in our Galaxy.
+Currently, about a hundred stellar streams are known in our Galaxy.
 \citet{2023MNRAS.520.5225M} provides a compilation of their tracks on the sky.
@@
   The most accurate model of the Palomar~5 stream would involve full modeling of the internal dynamics of the cluster.
-This would mean computing $N$-Body interactions with a $\mathcal{O}(N^2)$ computation time, stellar evolution, supernovae, an initial mass distribution, treatment of binary star systems, etc.
+This would mean computing $N$-body interactions with a $\mathcal{O}(N^2)$ computation time, stellar evolution, supernovae, an initial mass distribution, treatment of binary star systems, etc.
 \citep[for such an example, see][]{2021NatAs...5..957G, 2016MNRAS.458.1450W}.
@@
 Similarly, \citet{2024MNRAS.528.5189G} generated a globular cluster mass-loss catalog using seven different potential models and found their debris distributions to be rather model-independent, similarly to \citet{2023A&A...673A..44F}.
-While the clusters' positions may depend on the model, we assert that interaction rates and stream formation are unaffected by the choice of Galactic potential model.
+While the clusters' exact positions in time may depend on the model, we assert that interaction rates and stream formation are largely independent the choice of Galactic potential model.
 
@@
     We briefly compare our simulated gaps to the literature on Palomar~5.
-In the bottom panel of Fig.~\ref{fig:decomposition}, we compare the tracks of Palomar~5 that were compiled in Galstreams by \citet{2023MNRAS.520.5225M}.
+In the bottom panel of Fig.~\ref{fig:decomposition}, we compare the tracks of Palomar~5 that were compiled in \texttt{galstreams} by \citet{2023MNRAS.520.5225M}.
 The mid-point positions of the streams do not have the same galacto-centric positions as Palomar~5 from Baumgardt's catalog and are thus offset when projected into tail coordinates.
@@
 If we naively use the same radial dependence of gaps from Palomar~5 for GD-1 by swapping a pericenter from 6~kpc to 16~kpc, we would reduced the number of gap-favorable impacts of more than a factor of 2.
-However, we note that \citet{2022ApJ...941..129D} pre-select the globular clusters which could have experienced a close encounter with the GD-1 stream, by selecting only clusters which pass at a distance less than 0.5 kpc from the stream, having a relative velocity smaller than  300 km/s.
+Moreover, we note that \citet{2022ApJ...941..129D} pre-select the globular clusters which could have experienced a close encounter with the GD-1 stream, by selecting only clusters which pass at a distance less than 0.5 kpc from the stream, having a relative velocity smaller than  300 km/s.
 As shown in our Fig.~\ref{fig:impact_geometry_statistics}, bottom-left panel, in the case of Palomar~5 this choice would lead to exclude most of the encounters favorable to gap creation, which turn out to have relative velocities above 300 km/s.
@@
 This would help identify regimes where baryonic and dark matter-induced substructures can be disentangled, shedding light on the elusive influence of dark matter sub-halos on stellar streams.
-Additionally, we can continue to add realistic physics to the Galactic environment such as the merger of the Sagittarius dwarf galaxy and the passages of the large and small Magellanic clouds.
+Additionally, we can continue to add realistic physics to the Galactic environment such as the merger of the Sagittarius dwarf galaxy and the passages of the Large and Small Magellanic clouds.
 
@@
 The simulations will also be made publicly available at etidal-project.obspm.fr.
+
+\begin{acknowledgements}
+  This work has made use of the computational resources available at the Paris Observatory, as well as those obtained through the DARI grant A1020410154 (PI: P.
+Di Matteo), and of the Astropy, Numpy and Matplotlib libraries (Astropy Collaboration 2013, 2018; Harris et al.
+2020; Hunter 2007).
+SF, PDM and MM would like to thank the Ecole Doctorale 127 for its financial support.
+MM thanks the CNR STM program, thanks to which the collaboration on this research topic began.
+\end{acknowledgements}
 

### 2025-01-27 paper/main.tex
added a tab

@@
       \caption{\textit{Characteristics of gap causing clusters} \textit{Top Left:} Energy-angular momentum space of the globular clusters in the simulations, with 50~$\times$~165 data points representing all sampled initial conditions.
-Clusters impacting Palomar~5 are shown with colored markers; large for the samplins that induce a gap and small for those that do not.
+Clusters impacting Palomar~5 are shown with colored markers; large for the samplings that induce a gap and small for those that do not.
 Non-gap-causing clusters are shown in small gray dots.
@@
 In their study, they indeed suggested the presence of  five gaps with a 99~\% detection confidence, leading to a gap creation rate of 0.17~Gyr$^{-1}$~kpc$^{-1}$.
-If true, this rate would be too high to be explained by Galactic globular clusters only.
+If correct, this rate would be too high to be explained by Galactic globular clusters only.
 

### 2025-01-27 paper/main.tex
Added Sollima references

@@
   Subsequent studies\footnote{For more subsequent observation detections of tidal debris, i.e.
-globular cluster stars beyond the tidal radius see the works of: \citet{1997A&A...320..776L, 2000A&A...356..127T, 2000A&A...359..907L, 2001AAS...19910906S, 2003AJ....126..815L}.} extended Grillmair's findings to other globular cluster streams, but were often limited to the detections of stars still close to the cluster tidal radius, until the discovery made by \citet{2001ApJ...548L.165O,2002AAS...200.1001O, 2003AJ....126.2385O} of long and thin tails outside the Palomar~5 globular cluster.
+globular cluster stars beyond the tidal radius see the works of: \citet{1997A&A...320..776L, 2000A&A...356..127T, 2000A&A...359..907L, 2001AAS...19910906S, 2003AJ....126..815L,2011ApJ...726...47S,2018MNRAS.476.4814S,2020MNRAS.495.2222S}.} extended Grillmair's findings to other globular cluster streams, but were often limited to the detections of stars still close to the cluster tidal radius, until the discovery made by \citet{2001ApJ...548L.165O,2002AAS...200.1001O, 2003AJ....126.2385O} of long and thin tails outside the Palomar~5 globular cluster.
 With a mass of $1.34\pm 0.24 \times 10^4 M_{\sun}$ \citep{2019MNRAS.482.5138B}, Palomar~5 is one of the least massive globular clusters in the Galaxy.

### 2025-01-27 paper/main.tex
$\Lambda$~CDM to $\Lambda$CDM

@@
  % methods heading (mandatory)
-  {For this purpose, we used a particle-test code to simulate the dynamic evolution of Palomar~5 and its tidal tails when Palomar~5 is subject to the gravitational action of the Galaxy and its whole system of globular clusters.}
+  {For this purpose, we used a particle-test code to simulate Palomar~5's tidal tails when subjected to the Galaxy's gravitational field plus its whole system of globular clusters.}
  % results heading (mandatory)
@@
 Notably, the colder regions of the CMB temperature map correlate with whole-sky gravitational lensing maps \citep{2020A&A...641A...8P}--since the photons lost energy as they escaped the gravitational wells.
-The $\Lambda$~Cold Dark Matter ($\Lambda$~CDM) model accounts for these observations almost perfectly, proposing that dark matter exists and that its distribution originated from stochastic quantum fluctuations within the first second of the universe led to density variations.
+The $\Lambda$~Cold Dark Matter ($\Lambda$CDM) model accounts for these observations almost perfectly, proposing that dark matter exists and that its distribution originated from stochastic quantum fluctuations within the first second of the universe led to density variations.
 These variations propagated as the universe expanded, shaping the large-scale structure and distribution of galaxies we observe today.
@@
   % \citet{2019Galax...7...81Z} provide a review for the different proposed mechanisms of forming dark matter haloes theoretically and how to model them practically, showcasing how different assumptions for the fundamental particle behavior of dark matter propogates to the mass and size distribution of sub haloes on the galactic scale.
-The authors rightly state that until particles are discovered $\Lambda$~CDM remains a hypothesis.
+The authors rightly state that until particles are discovered $\Lambda$CDM remains a hypothesis.
 Since the true nature of dark matter is elusive the possibilities for sub-galactic behaviour and structure of dark matter are vast.
@@
 Later, \citet{2012ApJ...748...20C} expanded this idea, proposing that sub-halos could create gaps in stellar streams during flyby encounters, where a sub-halo approaches closely enough to a segment of a stream and significantly changes the orbits of the closest stars.
-\citet{2012ApJ...760...75C, 2013ApJ...768..171C} applied this idea and looked for the presence of gaps in the known Palomar~5 and GD-1 streams, concluding that the density variations found in their streams were consistent with expectations from $\Lambda$~CDM models.
+\citet{2012ApJ...760...75C, 2013ApJ...768..171C} applied this idea and looked for the presence of gaps in the known Palomar~5 and GD-1 streams, concluding that the density variations found in their streams were consistent with expectations from $\Lambda$CDM models.
 \citet{2019ApJ...880...38B} provided further observational evidence for this idea, identifying a gap and a spur in the GD-1 stream that they could not explain by known objects, such as globular clusters, and which they suggested was due to the close passage of a dark matter sub-halo.
-Interestingly enough, as stated in \citet{2019ApJ...880...38B}, the recovered properties of this sub-halo (mass and size) were inconsistent with the $\Lambda$~CDM mass-size relationship presented in \cite{2017MNRAS.466.4974M}.
+Interestingly enough, as stated in \citet{2019ApJ...880...38B}, the recovered properties of this sub-halo (mass and size) were inconsistent with the $\Lambda$CDM mass-size relationship presented in \cite{2017MNRAS.466.4974M}.
 The above-mentioned works are only a few of a much more extensive literature that has explored the impact of dark matter sub-halos in simulated streams \citep{2016ApJ...828L..10H, 2021MNRAS.507.1999H, 2021JCAP...10..043B, 2024arXiv240402953H, 2024arXiv241021174N}, or searched for their traces in observed streams \citep{2016MNRAS.460.2711T, 2017MNRAS.470...60E, 2020ApJ...889...70B, 2020ApJ...892L..37B}.
@@
 Going to lower orbital energies make the situation even worse, because the effect of the Galactic bar, giant molecular clouds and interactions with globular clusters becomes even more efficient.
-In the innermost regions of the Galaxy, where the density of dark matter sub-halos is expected to be maximal, globular cluster streams are intrinsically more difficult to find, in addition to the fact that the dynamic times in these regions become so small that the stars lost from globular clusters do not redistribute themselves for the most part into thin structures \citep[see][]{2023A&A...673A..44F}.
+In the innermost regions of the Galaxy, where the density of dark matter sub-halos is expected to be maximal, globular cluster streams are intrinsically more difficult to find, in addition to the fact that the dynamical times in these regions become so small that the stars lost from globular clusters do not redistribute themselves for the most part into thin structures \citep[see][]{2023A&A...673A..44F}.
 It is thus probably only at larger distances from the Galactic center than those spanned by Palomar~5's orbit that the impact of dark matter sub-halos may become dominant, but it should be borne in mind that at these distances the number density of sub-halos also decreases.

### 2025-01-27 paper/main.tex
small corrections from DVG

@@
 \citet{2021MNRAS.505.5978V} found that only a few outer globular clusters are strongly affected by different potential models.
-Generally, kinematic uncertainties are the dominant factor in orbital solution differences per cluster.
+Generally, kinematic uncertainties are the dominant factor in differences between orbital solutions per cluster.
 Similarly, \citet{2024MNRAS.528.5189G} generated a globular cluster mass-loss catalog using seven different potential models and found their debris distributions to be rather model-independent, similarly to \citet{2023A&A...673A..44F}.
@@
     There is a primary methodological departure from \citet{2023A&A...673A..44F}.
-In that paper, globular clusters evolved under the gravitational effect of the Galaxy only, while now we include also the effect of all the other Galactic globular clusters, by taking into account the direct $N$-body interactions among them.
-To quantify these interactions, all clusters are represented by Plummer spheres, each with mass and half-mass radius as reported in the Baumgardt catalog \citep{2021MNRAS.505.5957B}.
+In that work, globular clusters evolved under the gravitational effect of the Galaxy only, while now we include also the effect of all the other Galactic globular clusters, by taking into account the direct $N$-body interactions among them.
+To quantify these interactions, all clusters are represented by Plummer spheres, each with mass and half-mass radii as reported in the Baumgardt catalog \citep{2021MNRAS.505.5957B}.
 For the remainder of this paper, we refer to these simulations, where the globular clusters interactions are taken into account as \texttt{full} simulations.
@@
       \end{tabular}
-      \caption{The number of gaps created by each perturber across all 50 simulations of Palomar~5's tidal tail. This is the same data Fig.~\ref{fig:histogram_impact_time} but tabulated.}
+      \caption{The number of gaps created by each perturber across all 50 simulations of Palomar~5's tidal tail. These data are color-coded and illustrated in Fig.~\ref{fig:histogram_impact_time}.}
       \label{tab:gaps_per_perturber}

### 2025-01-27 paper/main.tex
added citation to L-BFGS method

@@
       \caption{The distribution and relationship between the impact variables from Eq.~\ref{eq:change_in_momentum} for all close fly-bys considered.
-The encounters that cause gaps are labeled with the same colors as from Fig.~
-      \ref{fig:histogram_impact_time} in the scatter plot, and binned in red for the histograms. The close encounters that did not lead to gaps are shown in black. }
+The encounters that cause gaps are labeled with the same colors as from Fig.~\ref{fig:histogram_impact_time} in the scatter plot, with white rings edges for visibility, and binned in red for the histograms.
+The close encounters that did not lead to gaps are shown in black.
+Indeed, no obvious trend delimiting these planes into gap producing or not emerges.}
       \label{fig:impact_geometry_statistics}    
@@
 As can be seen, no cluster with mass below $10^5 M_\odot$ causes gaps on the Palomar~5 stream, and all perturbers, except FSR~1758 have a half-mass radius below 10~pc.
-Even more interesting is their distribution in the E-Lz plane, which shows that the clusters that cause gaps are on both direct and retrograde orbits (respectively, negative and positive values of $L_z$), but all confined to an energy interval between $-2$ and $-1 \times10^5~\textrm{km}^2\textrm{s}^{-2}$.
+Even more interesting is their distribution in the E-L$_z$ plane, which shows that the clusters that cause gaps are on both direct and retrograde orbits (respectively, negative and positive values of $L_z$), but all confined to an energy interval between $-2$ and $-1 \times10^5~\textrm{km}^2\textrm{s}^{-2}$.
 This is because only clusters within Palomar~5's orbital space can interact with the stream: clusters with higher orbital energies tend to have larger apocenter than that of Palomar~5 and thus spend most of their time away from Palomar~5's orbital volume.
@@
 In contrast, observational data are inherently more challenging due to magnitude limits, contamination from field stars, and the lack of a ground truth for comparison.
-These factors suggest that the number of gaps identified in this study could represent an upper limit.
-
-  However, this conclusion is not necessarily accurate.
+
+  The aforementioned factors could lead one to the conclusion that the number of gaps identified in this study could represent an upper limit.
+However, this conclusion is incomplete.
 Globular clusters lose mass and evaporate over time, leading to an incomplete catalog of perturbers.
@@
 The minimum value of $b$, denoted as $\text{min}(b)$, represents the impact parameter.
-The minimization is carried out using \texttt{scipy}'s optimization package with the \textit{L-BFGS-B} method, which allows us to place bounds on $t$ and $\tau$, ensuring no extrapolation occurs.
+The minimization is carried out using \texttt{scipy}'s optimization package with the \textit{L-BFGS-B} method, which allows us to place bounds on $t$ and $\tau$, ensuring no extrapolation occurs \citep{davidon1991variable}.
 

### 2025-01-27 paper/main.tex
done with DVG comments

@@
 This would help identify regimes where baryonic and dark matter-induced substructures can be disentangled, shedding light on the elusive influence of dark matter sub-halos on stellar streams.
-Additionally, we can continue to add realistic physics to the Galactic environment such as the merger of the Sagittarius dwarf galaxy and the passages of the Large and Small Magellanic clouds.
+Additionally, we can continue to add realistic physics to the Galactic environment such as: the merger of the Sagittarius dwarf galaxy, the passages of the Large and Small Magellanic clouds, as well as a time evolving globular cluster population that compensates the current incomplete globular cluster population, mass loss, and evaporation.
 

### 2025-01-27 paper/main.tex
added point about incomplete census and evaporation of population

--- 2025-01-27 paper/main.tex
@@
 This would help identify regimes where baryonic and dark matter-induced substructures can be disentangled, shedding light on the elusive influence of dark matter sub-halos on stellar streams.
-Additionally, we can continue to add realistic physics to the Galactic environment such as the merger of the Sagittarius dwarf galaxy and the passages of the Large and Small Magellanic clouds.
+Additionally, we can continue to add realistic physics to the Galactic environment such as: the merger of the Sagittarius dwarf galaxy, the passages of the Large and Small Magellanic clouds, as well as a time evolving globular cluster population that compensates for the clusters' mass loss, evaporation, and similarly add members from the current incomplete census.
 

### 2025-01-27 paper/main.tex
owain's comments

@@
  % results heading (mandatory)
-  {Our study shows that the tails of Palomar~5 are strongly perturbed by the close passage of other clusters, in particular of NGC~2808, NGC~7078, NGC~104, and that these perturbations induce the formation of gaps in the tails.}
+  {Our study shows that the tails of Palomar~5 can be strongly perturbed by the close passage of other clusters, in particular of NGC~2808, NGC~7078, NGC~104, and that these perturbations induce the formation of gaps in the tails.}
  % conclusions heading (optional), leave it empty if necessary
-  {This result suggests that baryonic dense stellar systems such as globular clusters are capable of causing gaps in streams, as already demonstrated in the case of galaxy bar-induced perturbations and giant molecular clouds, and reinforce the idea that the effect of these components should be fully addressed when interpreting the morphology and density structures of observed stellar streams.}
+  {These results show that globular clusters are capable of inducing gaps in streams--as other baryonic components such as giant molecular clouds and the galactic bar have been shown to do in other works.
+Therefore, when searching to construct the distribution function of dark matter sub halos within the Milky Way, the gap contribution from globular clusters must be included.}
 
@@
 The two contrasting results could also show that at one distance is the halo is triaxial and at another spherical--highlighting the need for streams at different distances to map the halo shape.
-Recently, \citet{2024ApJ...967...89I} constructed the most accurate model of the Milky Way to date by applying a Monte Carlo Markov Chain (MCMC) fitting procedure.
+Recently, \citet{2024ApJ...967...89I} constructed a Milky Way model by applying a Monte Carlo Markov Chain (MCMC) fitting procedure.
 This method identified the set of potential parameters in an axisymmetric model of the Milky Way that best reproduces all observed stellar streams.
@@
 Later, \citet{2012ApJ...748...20C} expanded this idea, proposing that sub-halos could create gaps in stellar streams during flyby encounters, where a sub-halo approaches closely enough to a segment of a stream and significantly changes the orbits of the closest stars.
-\citet{2012ApJ...760...75C, 2013ApJ...768..171C} applied this idea and looked for the presence of gaps in the known Palomar~5 and GD-1 streams, concluding that the density variations found in their streams were consistent with expectations from $\Lambda$CDM models.
+\citet{2012ApJ...760...75C, 2013ApJ...768..171C} applied this idea and looked for the presence of gaps in the well-known Palomar~5 and GD-1 streams, concluding that the density variations found in their streams were consistent with expectations from $\Lambda$CDM models.
 \citet{2019ApJ...880...38B} provided further observational evidence for this idea, identifying a gap and a spur in the GD-1 stream that they could not explain by known objects, such as globular clusters, and which they suggested was due to the close passage of a dark matter sub-halo.
-Interestingly enough, as stated in \citet{2019ApJ...880...38B}, the recovered properties of this sub-halo (mass and size) were inconsistent with the $\Lambda$CDM mass-size relationship presented in \cite{2017MNRAS.466.4974M}.
+Interestingly enough, as stated in \citet{2019ApJ...880...38B}, the recovered properties of this sub-halo (mass and size) were denser than those with the $\Lambda$CDM mass-size relationship presented in \cite{2017MNRAS.466.4974M}.
 The above-mentioned works are only a few of a much more extensive literature that has explored the impact of dark matter sub-halos in simulated streams \citep{2016ApJ...828L..10H, 2021MNRAS.507.1999H, 2021JCAP...10..043B, 2024arXiv240402953H, 2024arXiv241021174N}, or searched for their traces in observed streams \citep{2016MNRAS.460.2711T, 2017MNRAS.470...60E, 2020ApJ...889...70B, 2020ApJ...892L..37B}.
@@
 
-  Few studies have specifically investigated the effect of globular clusters on stellar streams.
-The work of \citet{2017MNRAS.470...60E} concluded that the observed density variations found in Palomar~5 tails could not be due to the passage of globular clusters.
-However, note that this conclusion was not based on a numerical modeling of this effect.
+  Few studies have specifically investigated the effect of globular clusters on stellar streams. Here's an improved version with clearer phrasing and better flow:
+
+  \citet{2017MNRAS.470...60E} concluded that the observed density variations in the tails of Palomar~5 could not be attributed to the passage of globular clusters.
+Their analysis focused on the characteristics of the observed gaps and involved constraining properties of the progenitor using reconstructive modeling.
+By trial and error, they identified a specific configuration of masses, sizes, impact parameters, times of impact, and relative velocities for two perturbers that successfully reproduced the observed density distribution.
+However, their method does not perform full forward modeling of the entire globular cluster system on Palomar5's stream, as we do in this work.
+While they suggest that the impact rate of globular clusters is likely less significant—given their lower abundance compared to the expected population of dark matter subhalos—they do not explore this aspect in detail.
+However, they do state that it is indeed an avenue for future investigation.
+
   More recently \citet{2022ApJ...941..129D} have examined the possibility that the gaps found in the GD-1 stellar stream could be due to the close passage of globular clusters, concluding that this scenario is highly unlikely.
@@
 Instead, we opt for solving the \textit{restricted-three body problem}, or as known as the \texttt{particle-test method} as we did for \citet{2023A&A...673A..44F}, which we describe here for completeness.
+As demonstrated by \citet{2012A&A...546L...7M}, although the restricted three-body problem neglects the internal evolution of the cluster, it still reproduces very similar stream properties.
+This is because the model captures key extra-cluster physics, such as disk shocking and epicyclic stripping.
 
@@
       \includegraphics[width=\linewidth]{histogram_impact_time.png}
-      \caption{The time of impact for all gap causing flybys summed over all 50 simulations.
+      \caption{The simulation time at which the impacts occurred for all gap causing flybys summed over all 50 simulations.
 Each perturbing cluster is label and is color consistent.
@@
 The data set includes the top 5 strongest flybys from each simulation.
-Those that cause gaps are colored while those that do not are outlined in grey.
+Those that cause gaps are colored, stacked, and overlain atop of those that do--the distribution of which is outlined in grey.
 Note that the meaning of the colors is the same as in Fig.~\ref{fig:histogram_impact_time}.}
@@
 \textit{Top Right:} Mass-size plane of the globular clusters in the simulations, with uncertainties on the masses indicated as vertical lines.
-We recall the reader that the globular cluster catalog at this time does not provide uncertainties for clusters' characteristic radii.
+We remind the reader that the globular cluster catalog at this time does not provide uncertainties for clusters' characteristic radii.
 \textit{Bottom Left:} Palomar~5's orbit in the Galactocentric xy plane.

### 2025-01-27 paper/main.tex
subhalos->sub-halos

@@
 By trial and error, they identified a specific configuration of masses, sizes, impact parameters, times of impact, and relative velocities for two perturbers that successfully reproduced the observed density distribution.
-However, their method does not perform full forward modeling of the entire globular cluster system on Palomar5's stream, as we do in this work.
-While they suggest that the impact rate of globular clusters is likely less significant—given their lower abundance compared to the expected population of dark matter subhalos—they do not explore this aspect in detail.
+However, their method does not perform full forward modeling of the entire globular cluster system on Palomar~5's stream, as we do in this work.
+While they suggest that the impact rate of globular clusters is likely less significant—given their lower abundance compared to the expected population of dark matter sub-halos—-they do not explore this aspect in detail.
 However, they do state that it is indeed an avenue for future investigation.

### 2025-01-28 paper/main.tex
updated affilations
deleted no longer used files
updated DVG affiliation
added preprint and updated affiliations

@@
 %\title{Globular clusters fly-bys and gap creations in stellar streams.}
-\title{Gaps in stellar streams as a result of globular cluster fly-bys}
-
-\subtitle{The case of Palomar~5}
+\title{PREPRINT- Gaps in stellar streams as a result of globular cluster fly-bys}
+
+\subtitle{The case of Palomar~5 -- SUBMITTED TO A\&A}
 
@@
        \and
-       Nicolas Leclerc \inst{1}
-       \and 
-       Clement Hottier \inst{1}
-       \and 
-       Eliot Stein  \inst{6}
+       Nicolas Leclerc \inst{6}
+       \and 
+       Clement Hottier \inst{6}
+       \and 
+       Eliot Stein  \inst{7}
+       \and
+       David Valls-Gabaud \inst{8}
+       \and
+       Owain N. Snaith \inst{9}
        \and
        Misha Haywood \inst{1}
-       \and
-       David Valls-Gabaud \inst{1}
-       \and
-       Owain N. Snaith \inst{7}
        }
@@
   \and
-    institute des sciences spatial
-  \and
-    Exeter
+    UNIDIA, Observatoire de Paris, Université PSL, CNRS, 92190 Meudon, France
+  \and
+    DTIS, ONERA, Université Paris Saclay, 91123 Palaiseau, France 
+  \and
+    LUX, CNRS UMR 8252, Observatoire de Paris, PSL, 61 Avenue de l'Observatoire, 75014 Paris, France
+  \and
+    Department of Physics and Astronomy, University of Exeter, Stocker Rd, Exeter EX4 4QL, United Kingdom
   }
 
-  \date{Born in 1996; Accepted in 2025}
+  \date{SUBMITTED TO A\&A 27 JAN 2025; Accepted XXX}
 

### 2025-02-06 paper/main.tex
removed to A&A
mistakes found by eliot!

@@
 %\title{Globular clusters fly-bys and gap creations in stellar streams.}
-\title{PREPRINT- Gaps in stellar streams as a result of globular cluster fly-bys}
-
-\subtitle{The case of Palomar~5 -- SUBMITTED TO A\&A}
+\title{Gaps in stellar streams as a result of globular cluster fly-bys}
+
+\subtitle{The case of Palomar~5--PREPRINT}
 
@@
 To date, the smallest observed dark matter halo was detected through gravitational lensing in an Einstein ring with a mass of $10^8$ solar masses by \citet{2012Natur.481..341V}.
-However, some models predict that dark matter clumps could exist down to at least the mass of Earth-like planets at least \citep[see][and discussion in \citet{2021arXiv211101148A}]{2005JCAP...08..003G}.
+However, some models predict that dark matter clumps could exist down to at least the mass of Earth-like planets \citep[see][and discussion in \citet{2021arXiv211101148A}]{2005JCAP...08..003G}.
 
@@
 This model is time-independent throughout our simulations.
-This model satisfies a series of observational constraints such as local solar stellar density, the Galactic rotation curve, etc, similarly to other Galactic models such as \texttt{MWpotential2014} from \citet{2015ApJS..216...29B} and \texttt{McMillian2017} from \citet{2017MNRAS.465...76M}.
+This model satisfies a series of observational constraints such as local solar stellar density, the Galactic rotation curve, etc., similarly to other Galactic models such as \texttt{MWpotential2014} from \citet{2015ApJS..216...29B} and \texttt{McMillian2017} from \citet{2017MNRAS.465...76M}.
 However, to balance data volume and computation time, we use only one Galactic potential, which should suffice.
@@
           \end{equation} \label{eq:equation_of_motion_particle} where the index $p$ represents one of the 100,000 particles of interest, $\vec{r_p}$ being its position, and $j$ indexes over the globular clusters as in Eq.~\ref{eq:GCNBody}.
-Note in that Eq.~\ref{eq:equation_of_motion_particle} the positions of the globular clusters are time-dependent since they are being loaded during this step and not computed, unlike Eq.~\ref{eq:GCNBody}.
+Note that in Eq.~\ref{eq:equation_of_motion_particle} the positions of the globular clusters are time-dependent since they are being loaded during this step and not computed, unlike Eq.~\ref{eq:GCNBody}.
 
@@
 The results of these simulations are reported in the middle panels of Fig.~\ref{fig:decomposition} and clearly show that NGC~2808, NGC~7078 and NGC~104 are the clusters responsible for creating the under-dense regions observed in the leading tails of Palomar~5.
-It is worth noting that the times at which the passages of these clusters occurred, according to our analysis, are in agreement with the observed width of the corresponding gaps: the encounter with NGC~2808 being very recent (only 200~Myr ago) its induced gap is still very thin, because it takes time for a perturbation to grow into an extended gap, as it is the case for those induced by the passages of  NGC~7078 and NGC~104, which occurred in earlier times.
+It is worth noting that the times at which the passages of these clusters occurred, according to our analysis, are in agreement with the observed width of the corresponding gaps: the encounter with NGC~2808 being very recent (only 200~Myr ago), its induced gap is still very thin, because it takes time for a perturbation to grow into an extended gap, as it is the case for those induced by the passages of  NGC~7078 and NGC~104, which occurred in earlier times.
 It is also interesting to emphasize that gaps as thin as those generated by the passage of NGC~2808, 200~Myr ago, can be detected by working in the tail coordinate system and by making a comparative analysis (\texttt{full} versus \texttt{reference} simulations): they are so thin that they cannot be directly identified in Galactic coordinates (see Fig.~\ref{fig:stream_on_sky}).\\
@@
       \includegraphics[width=0.45\linewidth]{impact_stats_phase_space.png}
-      \caption{\textit{Characteristics of gap causing clusters} \textit{Top Left:} Energy-angular momentum space of the globular clusters in the simulations, with 50~$\times$~165 data points representing all sampled initial conditions.
+      \caption{\textit{Characteristics of gap causing clusters}.
+\textit{Top Left:} Energy-angular momentum space of the globular clusters in the simulations, with 50~$\times$~165 data points representing all sampled initial conditions.
 Clusters impacting Palomar~5 are shown with colored markers; large for the samplings that induce a gap and small for those that do not.
@@
 
-    We observe that there are no gaps near within 3~kpc to the cluster within our simulation.
+    We observe that there are no gaps near within 3~kpc to the cluster in our simulation.
 There may be a few reasons for the absence of gaps in the portion of the tails closer to the cluster center.
@@
 As we have discussed in Sect.~\ref{sect:geometry}, the number, $N$, of close encounters which lead to gap creation is anti-correlated with the distance $r$ to the Galactic center.
-If we naively use the same radial dependence of gaps from Palomar~5 for GD-1 by swapping a pericenter from 6~kpc to 16~kpc, we would reduced the number of gap-favorable impacts of more than a factor of 2.
+If we naively use the same radial dependence of gaps from Palomar~5 for GD-1 by swapping a pericenter from 6~kpc to 16~kpc, we would reduce the number of gap-favorable impacts of more than a factor of 2.
 Moreover, we note that \citet{2022ApJ...941..129D} pre-select the globular clusters which could have experienced a close encounter with the GD-1 stream, by selecting only clusters which pass at a distance less than 0.5 kpc from the stream, having a relative velocity smaller than  300 km/s.
@@
     For each of the 50 simulations, we compare the final snapshots between the \texttt{reference} and \texttt{full} simulations using the tail coordinate system, as shown in Fig.~\ref{fig:TailCoordinates}.
-The \texttt{reference} simulations were only saved at the final time-stamps, thus from Eq.~\ref{eq:data_volume_estimate} where $N_ts$=1, leading to a data volume of about one hundred measly megabytes.
+The \texttt{reference} simulations were only saved at the final time-stamps, thus from Eq.~\ref{eq:data_volume_estimate} where $N_{ts}$=1, leading to a data volume of about one hundred measly megabytes.
 We inspect these differences by generating 2D density maps of the streams.
@@
 
-    This method, however, has its limitations, especially when detecting smaller gaps.
+    However, this method has its limitations, especially when detecting smaller gaps.
 As outlined by \citet{2015MNRAS.450.1136E}, a small gap is not indicative of a weak impact, but a recent one.
@@
 This is done by first smoothing the image by taking a 5 point moving average kernel.
-Secondly, we use maximum filter to locate coordinates in the $t,\tau$ data plane who are local maxima to at least 10-adjacent data points.
+Secondly, we use maximum filter to locate coordinates in the ($t,\tau$) data plane who are local maxima to at least 10-adjacent data points.
 These locations are then ordered and the top five strongest interactions are saved.
@@
 Another important aspect is that the strongest peak in gravitational force does not necessarily create a gap.
-Notice how NGC~5272, which was labeled with 0 to indicate that it has the greatest local maxima does not have a gap.
+Notice how NGC~5272, which was labeled with 0 to indicate that it has the greatest local maxima, does not have a gap.
 The reason for this is manifold, for instance, the force needs to be modulated with the time since the change in momentum is the determining factor and not the peak magnitude of the force.
@@
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-041.png}      
-      \caption{Tenth gap gallery.}
+      \caption{Tenth gap gallery}
       \label{fig:gallery9}

### 2025-04-03 jobs/christmasTree/tree_in_time.py
remade script to handle other file name

@@ -19,6 +19,15 @@ import sys
 from mpl_toolkits.axes_grid1.inset_locator import inset_axes
 from multiprocessing import Pool
 
+
+# MASS GRID AND RADIUS GRID
+SIZE_GRID = 5
+N_MASS_SAMPLING = SIZE_GRID
+N_RADIUS_SAMPLING = SIZE_GRID # square grid
+MASS_GRID = np.logspace(4,4.8, N_MASS_SAMPLING) # in Msun
+RADIUS_GRID = np.logspace(np.log10(2),np.log10(30),N_RADIUS_SAMPLING)/1000 # in kpc
+
+
 plt.rcParams.update({
     'font.size': 12,          # Basic font size
     'axes.labelsize': 12,     # Label font size
@@ -34,20 +43,86 @@ plt.rcParams.update({
     'lines.markersize': 6,    # Size of markers
 })
 
-def grab_valid_fnames(GCname, montecarlokey, NPs, MASS_INDEX, RADIUS_INDEX, potential_env, internal_dynamics):
+
+def get_streamSnapShotsFileName(GCname, NP, potential_env, internal_dynamics, montecarlokey, MASS, RADIUS, mytype='physical'):
+    """
+    Get the filename for the StreamSnapShots file based on the GC name, Monte Carlo key, mass index, and radius index.
+    
+    Parameters:
+    ----------
+    GCname : str
+        Name of the globular cluster.
+    montecarlokey : str
+        Key for the Monte Carlo simulation.
+    MASS_INDEX : int
+        Index for the mass.
+    RADIUS : int
+        Index for the radius.
+    Returns:
+    -------
+    str
+        The formatted filename.
+    """ 
+    valid_types = ['index', "physical"]
+    if mytype not in valid_types:
+        raise ValueError("type must be one of the following: {:s}".format(str(valid_types)))
+    if mytype == 'index':
+        assert isinstance(MASS, int), "MASS must be an integer but was {:}".format(type(MASS))
+        assert isinstance(RADIUS, int), "RADIUS must be an integer but was {:}".format(type(RADIUS))
+        path=gcs.path_handler._StreamSnapShots(GCname=GCname,NP=NP,potential_env=potential_env,internal_dynamics=internal_dynamics)
[…]

### 2025-04-03 jobs/christmasTree/tree_in_time.py
floor to properly call the file name

@@ -463,7 +463,7 @@ if __name__ == "__main__":
     MASS=1
     RADIUS=3
 
-    MASS = int(np.ceil(MASS_GRID[MASS_INDEX]))
+    MASS = int(np.floor(MASS_GRID[MASS_INDEX]))
     print("RADIUS_GRID[RADIUS_INDEX]",RADIUS_GRID[RADIUS_INDEX])
     RADIUS = int(np.floor(1000*RADIUS_GRID[RADIUS_INDEX]))
     print("Radius", RADIUS)

### 2025-04-03 jobs/christmasTree/main.sh
need to coordinate ncpu's per task with the python script

@@ -6,7 +6,7 @@
 #SBATCH --time=29
 #SBATCH --array=[0-24]
 #SBATCH --ntasks=1
-#SBATCH --cpus-per-task=1
+#SBATCH --cpus-per-task=10
 
 
 srun python tree_in_time.py $SLURM_ARRAY_TASK_ID

### 2025-04-03 jobs/christmasTree/main.sh
dynamically obtaining cpus in python script

@@ -2,11 +2,14 @@
 #SBATCH --output=img_%A_%a.out  # Output file for each array job
 #SBATCH --error=img%A_%a.err   # Error file for each array job
 #SBATCH --job-name=christmas
-#SBATCH --partition=short
-#SBATCH --time=29
+#SBATCH --partition=medium
+#SBATCH --time=300
 #SBATCH --array=[0-24]
 #SBATCH --ntasks=1
-#SBATCH --cpus-per-task=10
+#SBATCH --cpus-per-task=10 
+
+
+### I'm trying to dynamically allocate the number of CPUs to the number of tasks based on the node 
 
 
 srun python tree_in_time.py $SLURM_ARRAY_TASK_ID

### 2025-04-03 jobs/christmasTree/tree_in_time.py

@@ -442,7 +442,10 @@ def main(
         # print("arg",arg)
     
     # Use multiprocessing to generate frames
-    with Pool(processes=10) as pool:
+    cpus = int(os.getenv("SLURM_CPUS_PER_TASK", 1))  # Default to 1 if not set
+    print("CPUS", cpus) # should be 10
+    # Use multiprocessing to generate frames
+    with Pool(processes=cpus) as pool:
         pool.starmap(generate_frame, args)
 
 
@@ -468,12 +471,6 @@ if __name__ == "__main__":
     RADIUS = int(np.floor(1000*RADIUS_GRID[RADIUS_INDEX]))
     print("Radius", RADIUS)
 
-    fname= get_streamSnapShotsFileName(GCname, NPs[0], MWpotential, internal_dynamics, "monte-carlo-009", MASS, RADIUS, mytype='physical')
-    print("fname",fname)
-    print("file exists", os.path.exists(fname))
-    fname = get_streamSnapShotsFileName(GCname, NPs[0], MWpotential, internal_dynamics, "monte-carlo-009", MASS_INDEX, RADIUS_INDEX, mytype='index')
-    print("fname",fname)
-    print("file exists", os.path.exists(fname))
     # Call the main function with the specified parameters
     main(MASS=MASS, RADIUS=RADIUS)
 

### 2025-04-03 jobs/christmasTree/tree_in_time.py
added function to compute the missing frames

@@ -18,6 +18,7 @@ from scipy import signal
 import sys 
 from mpl_toolkits.axes_grid1.inset_locator import inset_axes
 from multiprocessing import Pool
+import re
 
 
 # MASS GRID AND RADIUS GRID
@@ -354,7 +355,29 @@ def generate_frame(i, valid_time_stamps, fnames, valid_NPs, torbit, xorbit, yorb
     print(f"Saved frame {i} to {baseout+figname}")
 
 
+def get_present_frame_numbers(directory):
+    """
+    Get a list of integers present in the file names in the given directory.
     
+    Parameters:
+    directory (str): Path to the directory containing the files.
+    
+    Returns:
+    list: A sorted list of integers extracted from the file names.
+    """
+    frame_numbers = []
+    # Regular expression to match the pattern "christmas_tree_<number>.png"
+    pattern = re.compile(r'christmas_tree_(\d+)\.png')
+    
+    # Iterate over all files in the directory
+    for filename in os.listdir(directory):
+        match = pattern.match(filename)
+        if match:
+            # Extract the integer and add it to the list
+            frame_numbers.append(int(match.group(1)))
+    
+    # Return the sorted list of integers
+    return sorted(frame_numbers)    
 
 def main(
     GCname="Pal5",
@@ -366,6 +389,7 @@ def main(
     MASS=1,
     RADIUS=3,
     fnametype='physical',
+    ONLYUNPROCESSED=True
 ):
     # Pick the color map for the tree
     mycmap = mpl.cm.hsv
@@ -430,23 +454,34 @@ def main(
     
     # Use multiprocessing to generate frames
     ntimestamps = len(valid_time_stamps)
+    timestampindexes= np.arange(ntimestamps)
     fignamebase = "{:s}-StreamSnapShots-{:s}_mass_{:s}_radius_{:s}".format(GCname, montecarlokey, str(MASS).zfill(3), str(RADIUS).zfill(3))
     # make some directories for this....
+    if ONLYUNPROCESSED:
+        # Get the present frame numbers
+        present_frame_numbers = get_present_frame_numbers(baseout)
+        # Remove the present frame numbers from the list of all timestamps
+        absent_frame_numbers = np.delete(timestampindexes, present_frame_numbers)
+
+        print("Number of missing frames", len(absent_frame_numbers))
+        # Remove the present frame numbers from the list of all timestamp
+    else:
[…]

### 2025-04-04 jobs/christmasTree/tree_in_time.py
executed

@@ -5,7 +5,7 @@ It will color coat the tree according to when pericenter passages
 it will display the 1d profile of the tree beneath it for each each group
 
 """
-
+import datetime
 import stream_analysis as sa
 import gcs 
 import tstrippy
@@ -454,19 +454,19 @@ def main(
     
     # Use multiprocessing to generate frames
     ntimestamps = len(valid_time_stamps)
-    timestampindexes= np.arange(ntimestamps)
+    timestampindexes= np.arange(1,ntimestamps)
     fignamebase = "{:s}-StreamSnapShots-{:s}_mass_{:s}_radius_{:s}".format(GCname, montecarlokey, str(MASS).zfill(3), str(RADIUS).zfill(3))
     # make some directories for this....
     if ONLYUNPROCESSED:
         # Get the present frame numbers
         present_frame_numbers = get_present_frame_numbers(baseout)
         # Remove the present frame numbers from the list of all timestamps
-        absent_frame_numbers = np.delete(timestampindexes, present_frame_numbers)
+        absent_frame_numbers = np.setdiff1d(timestampindexes, present_frame_numbers)
 
         print("Number of missing frames", len(absent_frame_numbers))
         # Remove the present frame numbers from the list of all timestamp
     else:
-        absent_frame_numbers= timestampindexes
+        absent_frame_numbers = timestampindexes
     
     args = [
         (i, valid_time_stamps, fnames, valid_NPs, torbit, xorbit, yorbit, zorbit, vxorbit, vyorbit, vzorbit, 
@@ -476,12 +476,17 @@ def main(
     # for arg  in args:
         # print("arg",arg)
     
+    # Use multiprocessing to generate frames
+    cpus = int(os.getenv("SLURM_CPUS_PER_TASK", 1))  # Default to 1 if not set
+    print("CPUS", cpus) # should be 10
     # # Use multiprocessing to generate frames
-    # cpus = int(os.getenv("SLURM_CPUS_PER_TASK", 1))  # Default to 1 if not set
-    # print("CPUS", cpus) # should be 10
-    # # Use multiprocessing to generate frames
-    # with Pool(processes=cpus) as pool:
-    #     pool.starmap(generate_frame, args)
+    starttime=datetime.datetime.now()
+    with Pool(processes=cpus) as pool:
+        pool.starmap(generate_frame, args)
+    endtime=datetime.datetime.now()
+    print("Time taken to generate frames", endtime-starttime)
+    # # Generate frames in parallel
+    
 
 
 if __name__ == "__main__":

### 2025-04-07 jobs/christmasTree/tree_in_time.py
spell correction of comment

@@ -1,5 +1,5 @@
 """
-This module with make the frames for the Christmas tree animation.
+This module will make the frames for the Christmas tree animation.
 It will color coat the tree according to when pericenter passages
 
 it will display the 1d profile of the tree beneath it for each each group

### 2025-04-07 jobs/tailCoordinateAnimations/mp_tauGammaProfileFrames.py
not needed, delete

@@ -1,68 +1,0 @@
-"""
-The secrete sauce is to extract the stream particles within each process!
-the copying really must've added a lot of overhead. 
-
-"""
-
-import gcs
-from gcs import path_handler as ph
-import os 
-import h5py
-import multiprocessing as mp
-import tauGammaProfileFrames as TGPF #type: ignore
-import datetime
-
-def main(dataparams):
-
-    GCname,MWpotential,montecarlokey,NP,internal_dynamics=dataparams
-    outdir=ph.paths['temporary'] + "stream_analysis/tauGammaProfileFrames/" + MWpotential  + GCname + "/" + montecarlokey  + "/"
-    os.makedirs(outdir,exist_ok=True)  
-
-    # get the host horbit 
-    orbfname=ph.GC_orbits(MWpotential,GCname)
-    fullhostorbit=gcs.extractors.GCOrbits.extract_whole_orbit(orbfname,montecarlokey)  
-
-    # get the stream at each snap shot
-    streamFname=ph.StreamSnapShots(GCname=GCname,NP=NP,potential_env=MWpotential,internal_dynamics=internal_dynamics,montecarlokey=montecarlokey)
-    with h5py.File(streamFname,"r") as myfile:
-        timestamps=myfile['time_stamps'][:]
-    nFrames = len(timestamps)
-
-    starttime=datetime.datetime.now()
-
-
-    ncpu = mp.cpu_count()
-    pool = mp.Pool(ncpu)
-    print("ncpu:",ncpu)
-    results = []
-    for i in range(42, nFrames):
-        result = pool.apply_async(TGPF.make_frame, args=(outdir, i, timestamps, fullhostorbit, streamFname))
-    pool.close()
-    pool.join()
-    
-    # Optionally, handle results or check for errors
-    for result in results:
-        try:
-            result.get()  # This will raise any exceptions that occurred in the worker process
-        except Exception as e:
-            print(f"Error in worker process: {e}")
-
-    endtime=datetime.datetime.now()
-    print("Time taken: ",endtime-starttime)
-
-
-
-
-
-if __name__=="__main__":
-    # data params
-    GCname = "Pal5"
-    MWpotential = "pouliasis2017pii-GCNBody"
-    montecarlokey = "monte-carlo-000"
-    NP = int(1e5)
-    internal_dynamics = "isotropic-plummer"
-
-    dataparams = (GCname,MWpotential,montecarlokey,NP,internal_dynamics)
-    main(dataparams)
-
-    

### 2025-04-07 jobs/tailCoordinateAnimations/tauGammaProfileFrames.py

@@ -1,112 +1,0 @@
-import gcs
-from gcs import path_handler as ph
-import numpy as np 
-import matplotlib.pyplot as plt
-import stream_analysis as sa 
-import os 
-import matplotlib as mpl
-import h5py
-
-
-    
-
-def gspec_tau_gamma_and_profile(FIG={"figsize":(10,5)},GSPEC={
-    'width_ratios':[1,0.01],
-    'height_ratios':[1,1],
-    'hspace':0.001,
-    'wspace':0.01
-}):
-    fig=plt.figure(**FIG)
-    gspec=mpl.gridspec.GridSpec(2,2,**GSPEC)
-    axis0 = fig.add_subplot(gspec[0,0])
-    axis1 = fig.add_subplot(gspec[1,0])
-    cbar_ax=fig.add_subplot(gspec[0,1])
-
-    # Adjust the position and size of the colorbar axis to match axis0
-    pos0 = axis0.get_position()
-    pos1 = axis1.get_position()
-    cbar_ax.set_position([pos0.x1 + 0.01, pos0.y0 + pos0.height/4 , 0.01, pos0.height/2 ])
-
-    return fig,axis0,axis1,cbar_ax
-
-
-
-
-def properties(NP,xlims=[-0.1,0.1],ylims=[-0.1,0.1]):
-    PROF={'yscale':"log","xlabel":r"$\tau$ [s kpc / km]","ylabel":"counts"}
-    AXIS={}
-    AXIS['xlim']=xlims
-    AXIS['ylim']=ylims
-    AXIS['ylabel']=r"$\gamma$ [s kpc / km]"
-    AXIS['xlabel']=""
-    AXIS['xticks']=[]
-    AXIS['aspect']="equal"
-    SCAT=sa.plotters.tailcoordinates.density_SCAT_properties(NP)
-    return AXIS,PROF,SCAT
-
-
-
-
-def grab_map_and_profile(NP,tau,gamma,xlims,ylims):
-    X,Y,H=sa.plotters.binned_density.short_cut(NP,tau,gamma,xlims,ylims)
-    H=sa.plotters.binned_density.normalize_density_by_particle_number(H,NP)
-    X,Y,H=sa.plotters.binned_density.order_by_density(X,Y,H)
-    xedges,yedges   =   sa.plotters.binned_density.get_edges(NP,xlims,ylims)
[…]

### 2025-04-07 jobs/densityProfileEvolution/stackedTauDensityMaps.py
made script to create new density maps

@@ -1,0 +1,329 @@
+"""
+
+This script extracts the stream and makes the 1D density maps in time (which is a faster computation)
+It projects it on to the orbit. 
+It follows the evolution of the stream density in time.
+It uses multiprocessing to speed up the computation.
+
+It has to open many different files since I have them spaced apart in different data files.
+
+
+"""
+
+import gcs 
+import numpy as np
+import h5py
+import stream_analysis as sa 
+import os 
+import multiprocessing as mp
+import datetime
+import sys
+
+# MASS GRID AND RADIUS GRID
+SIZE_GRID = 5
+N_MASS_SAMPLING = SIZE_GRID
+N_RADIUS_SAMPLING = SIZE_GRID # square grid
+MASS_GRID = np.logspace(4,4.8, N_MASS_SAMPLING) # in Msun
+RADIUS_GRID = np.logspace(np.log10(2),np.log10(30),N_RADIUS_SAMPLING)/1000 # in kpc
+
+
+ATTRS = {
+    "Author": "Salvatore Ferrone",
+    "Date": datetime.datetime.now().strftime("%Y-%m-%d"),
+    "email":"salvatore.ferrone.1996@gmail.com",
+    "Institution": "l'Università di Roma, La Sapienza",
+    "Description": "a series of 1D density maps of the stream projected onto the orbit for each snapshot",
+}
+
+
+def taudensity_fname(GCname,NP,MWpotential,internal_dynamics,montecarlokey,MASS,RADIUS):
+    fname = "{:s}-tauDensity-{:s}_mass_{:s}_radius_{:s}.hdf5".format(GCname, montecarlokey, str(MASS).zfill(3), str(RADIUS).zfill(3))
+    pathname=gcs.path_handler._tauDensityMaps(GCname=GCname,NP=NP, MWpotential=MWpotential, internal_dynamics=internal_dynamics)
+    if not os.path.exists(pathname):
+        os.makedirs(pathname, exist_ok=True)
+    return pathname+fname
+
+
+def get_streamSnapShotsFileName(GCname, NP, potential_env, internal_dynamics, montecarlokey, MASS, RADIUS, mytype='physical'):
+    """
+    Get the filename for the StreamSnapShots file based on the GC name, Monte Carlo key, mass index, and radius index.
+    
+    Parameters:
+    ----------
+    GCname : str
+        Name of the globular cluster.
+    montecarlokey : str
+        Key for the Monte Carlo simulation.
+    MASS_INDEX : int
+        Index for the mass.
+    RADIUS : int
[…]

### 2025-04-07 jobs/densityProfileEvolution/stackedTauDensityMaps.sh

@@ -1,0 +1,15 @@
+#!/bin/sh
+#SBATCH --output=out/stacked_%A_%a.out  # Output file for each array job
+#SBATCH --error=err/stacked_%A_%a.err   # Error file for each array job
+#SBATCH --job-name=stacked
+#SBATCH --partition=short
+#SBATCH --time=30
+## NOT YET SBATCH --array=[0-24]
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=10 
+
+
+### I'm trying to dynamically allocate the number of CPUs to the number of tasks based on the node 
+
+
+srun python stackedTauDensityMaps.py 1

### 2025-04-07 jobs/densityProfileEvolution/stackedTauDensityMaps.py
white space

@@ -1,13 +1,9 @@
 """
-
 This script extracts the stream and makes the 1D density maps in time (which is a faster computation)
 It projects it on to the orbit. 
 It follows the evolution of the stream density in time.
 It uses multiprocessing to speed up the computation.
-
 It has to open many different files since I have them spaced apart in different data files.
-
-
 """
 
 import gcs 

### 2025-04-07 jobs/densityProfileEvolution/stackedTauDensityMaps.py
fixed calling sequence

@@ -230,7 +230,6 @@ def main(dataparams,hyperparams):
 
     # see if the output is already made:
     outfname=taudensity_fname(GCname, NP, MWpotential, internal_dynamics, montecarlokey, MASS, RADIUS)
-
     if os.path.exists(outfname):
         print(outfname, "already exists. \n Delete if you want to recompute.")
         return
@@ -238,20 +237,22 @@ def main(dataparams,hyperparams):
     # extract the time stamps 
     with h5py.File(fnames[0],"r") as myfile:
         time_stamps=myfile['time_stamps'][:]
+    
+    # extract the host orbit 
+    hostorbit       =   gcs.extractors.GCOrbits.extract_whole_orbit(gcs.path_handler.GC_orbits(MWpotential=MWpotential,GCname=GCname),montecarlokey=montecarlokey)
+    
     # initialize the output arrays
     tdynamical=median_dynamical_time(hostorbit)
     tau_edges=initialize_tau_edges(NP,tdynamical,ndyn)
     tau_centers, counts=initialize_2D_counts(tau_edges,len(time_stamps))
 
-    # extract the host orbit 
-    hostorbit       =   gcs.extractors.GCOrbits.extract_whole_orbit(gcs.path_handler.GC_orbits(MWpotential=MWpotential,GCname=GCname),montecarlokey=montecarlokey)
     
-
+    END_FRAMES = 30
 
     pool = mp.Pool(processes=ncpu)
     # do this in parallel because its slow
     start_time = datetime.datetime.now()
-    results = [pool.apply_async(loop_extract_density_profile, args=(i, time_stamps, fnames, NPs, hostorbit, ndyn, tau_edges))for i in range(start_index,len(time_stamps))]
+    results = [pool.apply_async(loop_extract_density_profile, args=(i, time_stamps, fnames, NPs, hostorbit, ndyn, tau_edges))for i in range(start_index,END_FRAMES)]
     end_time = datetime.datetime.now()
     comp_time = end_time - start_time
 
@@ -289,6 +290,36 @@ def main(dataparams,hyperparams):
         f.create_dataset('fnames', data=fnames)
         f.create_dataset("NPs", data=NPs)
 
+    print("done with {:s}".format(outfname))
+    # clean up
+    del counts
+    del tau_centers
+    del tau_edges
+    del time_stamps
+    del fnames
+    del NPs
+    del hostorbit   
+    del my_phase_space
[…]

### 2025-05-12 paper/main.tex
can't seem to turn off the line numbering
added \nolinenumbers after make title. duh
removed chatGPT's comment on better flow
post-review paper version

@@
 
-%\title{Gaps in stellar streams as result of globular cluster close passages: the case of Palomar~5}
-%\title{The impact of globular clusters close passages on stellar streams: the case of Pal5}
+%\title{Gaps in stellar streams as a result of globular cluster close passages: the case of Palomar~5}
+%\title{The impact of globular clusters' close passages on stellar streams: the case of Pal5}
 %\title{Globular clusters fly-bys and gap creations in stellar streams.}
@@
 
-\subtitle{The case of Palomar~5--PREPRINT}
+\subtitle{The case of Palomar~5}
 
@@
 \abstract
-  %  menntion that only one Galactic model should be sufficient to avoid the criticism
+  %  mention that only one Galactic model should be sufficient to avoid criticism
   {Thin stellar streams, such as those resulting from the tidal disruption of globular clusters, have long been known and used as probes of the gravitational potential of our Galaxy, both its visible and dark contents.
-In particular, the presence of under-density regions, or gaps, along these streams is commonly interpreted as being due to the close passage of dark matter sub-halos.
+In particular, in the literature, we commonly interpret the presence of under-density regions, or gaps, along these streams as being due to the close passage of dark matter sub-halos.
 }
@@
   {In this work, we investigate the perturbations induced on streams by the passage of dense stellar systems, such as globular clusters themselves, to test the possibility that they may cause the formation of gaps as well.
-In particular, we focus on the study of the stream of Palomar~5, a well-known globular cluster in the Galactic halo, which has particularly long tidal tails.
+In particular, we focus on the study of the stream \textbf{\paola{generated by a cluster with characteristics (mass, size, and orbit) similar to the current characteristics}} of Palomar~5, a well-known globular cluster in the Galactic halo, which has particularly long tidal tails.
 }
  % methods heading (mandatory)
-  {For this purpose, we used a particle-test code to simulate Palomar~5's tidal tails when subjected to the Galaxy's gravitational field plus its whole system of globular clusters.}
+  {For this purpose, we used a particle-test code to simulate \textbf{\paola{the formation and evolution of the stream}} when subjected to the Galaxy's gravitational field plus its whole system of globular clusters.}
  % results heading (mandatory)
-  {Our study shows that the tails of Palomar~5 can be strongly perturbed by the close passage of other clusters, in particular of NGC~2808, NGC~7078, NGC~104, and that these perturbations induce the formation of gaps in the tails.}
+  {Our study shows that the \textbf{\paola{such a stream}} can be strongly perturbed by the close passage of other clusters, in particular of NGC~2808, NGC~7078, and NGC~104, and that these perturbations induce the formation of gaps in the tails.}
  % conclusions heading (optional), leave it empty if necessary
-  {These results show that globular clusters are capable of inducing gaps in streams--as other baryonic components such as giant molecular clouds and the galactic bar have been shown to do in other works.
-Therefore, when searching to construct the distribution function of dark matter sub halos within the Milky Way, the gap contribution from globular clusters must be included.}
+  {These results show that globular clusters can induce gaps in \textbf{\paola{cold}} streams--just as others have shown that other baryonic components such as giant molecular clouds and the galactic bar may as well.
+Therefore, a future work that attempts to infer the dark matter sub-halo distribution from stellar stream gaps must include the contributions from globular clusters.}
 
 \maketitle
+% \nolinenumbers
 \section{Introduction}
@@
 Since then, mounting evidence has been assembled to support this claim.
-\citet{1970IAUS...38..351F,1970ApJ...161..802F} elaborated on the dynamics of disk galaxies and briefly acknowledged in the appendix of his paper that the rotation curves of NGC~300 and M~33 cannot be fully explained by disk models and that there must be the same amount of mass or more that is undetected and beyond the extend of the galaxies to explain their kinematics.
-Concurrently \citet{1970ApJ...159..379R} studied the rotation of Andromeda and observed it to be flat---not decreasing as one would expect from a gravitational potential with more mass concentration in the center.
+\citet{1970IAUS...38..351F,1970ApJ...161..802F} elaborated on the dynamics of disk galaxies and briefly acknowledged in the appendix of his paper that disk models cannot fully explain the rotation curves of NGC~300 and M~33 and that there must be the same amount of mass or more that is undetected and beyond the extend of the galaxies to explain their kinematics.
+Concurrently, \citet{1970ApJ...159..379R} studied the rotation of Andromeda and observed it to be flat---not decreasing as one would expect from a gravitational potential with more mass concentration in the center.
 Eventually, astronomers returned to Zwicky's non-luminous matter to explain these fast-rotating disks, as discussed in \citet{1983SciAm.248f..96R}'s review.
@@
 Notably, the colder regions of the CMB temperature map correlate with whole-sky gravitational lensing maps \citep{2020A&A...641A...8P}--since the photons lost energy as they escaped the gravitational wells.
-The $\Lambda$~Cold Dark Matter ($\Lambda$CDM) model accounts for these observations almost perfectly, proposing that dark matter exists and that its distribution originated from stochastic quantum fluctuations within the first second of the universe led to density variations.
+The $\Lambda$~Cold Dark Matter ($\Lambda$CDM) model accounts for these observations almost perfectly, proposing that dark matter exists and that its distribution originated from stochastic quantum fluctuations within the first second of the universe, which led to density variations.
 These variations propagated as the universe expanded, shaping the large-scale structure and distribution of galaxies we observe today.
 
-  % As laid out in \citet{2021PrPNP.11903865A}'s review, despite significant progress, the true nature of dark matter remains elusive and scientists continue to search for evidence of its existence beyond its gravitational effects on cosmic and Galactic scales.
-A key focus today is detecting dark matter on sub-Galactic scales, because at these scales different cosmological scenarios predict different lumpiness of the halos.
-
-  % \citet{2019Galax...7...81Z} provide a review for the different proposed mechanisms of forming dark matter haloes theoretically and how to model them practically, showcasing how different assumptions for the fundamental particle behavior of dark matter propogates to the mass and size distribution of sub haloes on the galactic scale.
-The authors rightly state that until particles are discovered $\Lambda$CDM remains a hypothesis.
-Since the true nature of dark matter is elusive the possibilities for sub-galactic behaviour and structure of dark matter are vast.
-Obtaining constraints on the distirbution of dark matter subhaloes in turn will place constraints on the particle nature of dark matter.
-As proposed by \citet{2012ApJ...748...20C}, a most promising tool to set limits on the number and mass density of the sub-halos, are stellar streams.
-
-  Stellar streams are several-kpc long structures formed by the tidal disruption of globular clusters or dwarf galaxies orbiting a host galaxy.
-These tidal forces arise due to differential gravitational pulls across extended objects, causing stars farther from the galactic center to lag behind while those closer are pulled away.
-This stretching creates two tidal tails that trace the cluster's orbit, unless in the closest vicinity to the object \citep{2007ApJ...659.1212M}.
+  % As laid out in \citet{2021PrPNP.11903865A}'s review, despite significant progress, the true nature of dark matter remains elusive, and scientists continue to search for evidence of its existence beyond its gravitational effects on cosmic and Galactic scales.
+A key focus today is detecting dark matter on sub-galactic scales because, at these scales, different cosmological scenarios predict different lumpiness of the halos.
+
+  % \citet{2019Galax...7...81Z} provides a review of the different proposed mechanisms of forming dark matter haloes theoretically and how to model them practically, showcasing how different assumptions for the fundamental particle behavior of dark matter propagate to the mass and size distribution of sub-haloes on the galactic scale.
+The authors rightly state that until particles are discovered, $\Lambda$CDM remains a hypothesis.
+Since the true nature of dark matter is elusive, the possibilities for sub-galactic behavior and structure of dark matter are vast.
+In turn, obtaining constraints on the distribution of dark matter subhaloes will constrain the particle nature of dark matter.
+As proposed by \citet{2012ApJ...748...20C}, stellar streams are the most promising tool to limit the number and mass density of the sub-haloes.
+
+  Stellar streams are several-kpc-long structures formed by the tidal disruption of globular clusters or dwarf galaxies orbiting a host galaxy.
+These tidal forces arise due to differential gravitational pulls across extended objects, causing stars farther from the galactic center to lag while those closer speed ahead.
+This stretching creates two tidal tails that trace the cluster's orbit unless in the closest vicinity to the object \citep{2007ApJ...659.1212M}.
 Numerical predictions of this phenomenon exist since the 1970s \citep[see, for example][]{1975AJ.....80..290K}.
@@
   These predictions occurred well before the first detections of Galactic globular cluster tidal tails \citep{1995AJ....109.2553G}.
-Interestingly, \citet{1995AJ....109.2553G}'s detections were made nearly contemporary  with the discovery of the Sagittarius stellar stream by \citet{1994Natur.370..194I} which is the closest example of a stream emerging from a dwarf satellite currently interacting with the Milky Way.
-
-  Subsequent studies\footnote{For more subsequent observation detections of tidal debris, i.e.
-globular cluster stars beyond the tidal radius see the works of: \citet{1997A&A...320..776L, 2000A&A...356..127T, 2000A&A...359..907L, 2001AAS...19910906S, 2003AJ....126..815L,2011ApJ...726...47S,2018MNRAS.476.4814S,2020MNRAS.495.2222S}.} extended Grillmair's findings to other globular cluster streams, but were often limited to the detections of stars still close to the cluster tidal radius, until the discovery made by \citet{2001ApJ...548L.165O,2002AAS...200.1001O, 2003AJ....126.2385O} of long and thin tails outside the Palomar~5 globular cluster.
+Interestingly, \citet{1995AJ....109.2553G}'s detections were made nearly contemporary with the discovery of the Sagittarius stellar stream by \citet{1994Natur.370..194I}, which is the closest example of a stream emerging from a dwarf satellite currently interacting with the Milky Way.
+
+  Subsequent studies\footnote{For more subsequent observation detections of tidal debris, i.e., globular cluster stars beyond the tidal radius see the works of: \citet{1997A&A...320..776L, 2000A&A...356..127T, 2000A&A...359..907L, 2001AAS...19910906S, 2003AJ....126..815L,2011ApJ...726...47S,2018MNRAS.476.4814S,2020MNRAS.495.2222S}.} extended Grillmair's findings to other globular cluster streams, but were often limited to the detections of stars still close to the cluster tidal radius, until the discovery made by \citet{2001ApJ...548L.165O,2002AAS...200.1001O, 2003AJ....126.2385O} of long and thin tails outside the Palomar~5 globular cluster.
 With a mass of $1.34\pm 0.24 \times 10^4 M_{\sun}$ \citep{2019MNRAS.482.5138B}, Palomar~5 is one of the least massive globular clusters in the Galaxy.
-\citet{2003AJ....126.2385O} showed that its tails contain more mass than the cluster itself.
-The works of \citet{2006ApJ...641L..37G} and \citet{2015MNRAS.446.3297K} showed the tails to be extended for more than $20^\circ$ degrees in the sky.
+\citet{2003AJ....126.2385O} showed that its tails contain more mass than the cluster.
+The works of \citet{2006ApJ...641L..37G} and \citet{2015MNRAS.446.3297K} showed that the tails have an extent of more than $20^\circ$ degrees in the sky.
 The discovery of its prominent tails stimulated a vigorous and successful search in the following years.
@@
   \citet{2025NewAR.10001713B} provides a review of stellar stream astronomy, which has entered a new era since the publication of the data from the Gaia astrometric mission \citep{2016A&A...595A...1G}.
-Gaia's characterization of billions of stars in the Milky Way allows us to look for these structures coupling photometry, astrometry and, for the brightest stars, spectroscopy.
-The possibility given by Gaia to track stars with coherent movements, over the entirety of the sky, has led to the discovery of dozens of new streams.
-For example, \citet{2018MNRAS.477.4063M} developed the \texttt{streamfinder} algorithm and applied it across a series works \citep{2018MNRAS.481.3442M,  2018ApJ...865...85I, 2019ApJ...872..152I} to discover a multitude of streams.%: Slidr, Sylgr, Ylgr, Fimbulthul, Svol, Gjoll, Leiptr, Phlegeton and Fjorm.
-According to the most recent estimates provided by \citet{2020MNRAS.493.4978S} Fjorm in particular has an extension of more than 150 degrees on the sky which has a projected length more than five times greater than that of Palomar~5.
-
-  The possibility to couple Gaia data with spectroscopic surveys has extended the study of stellar streams beyond the quantification of their orbital properties, to a full chemical characterization.
+Gaia's characterization of billions of stars in the Milky Way allows us to look for these structures, coupling photometry, astrometry, and spectroscopy for the brightest stars.
+The possibility given by Gaia to track stars with coherent movements over the entirety of the sky has led to the discovery of dozens of new streams.
+For example, \citet{2018MNRAS.477.4063M} developed the \texttt{streamfinder} algorithm and applied it across a series of works \citep{2018MNRAS.481.3442M,  2018ApJ...865...85I, 2019ApJ...872..152I} to discover a multitude of streams.%: Slidr, Sylgr, Ylgr, Fimbulthul, Svol, Gjoll, Leiptr, Phlegeton and Fjorm.
+According to the most recent estimates provided by \citet{2020MNRAS.493.4978S}, Fjorm, in particular, has an extension of more than 150 degrees on the sky, which has a projected length more than five times greater than that of Palomar~5.
+
+  The possibility of combining Gaia data with spectroscopic surveys has extended the study of stellar streams beyond the quantification of their orbital properties to a full chemical characterization.
 For example \citet{2019MNRAS.490.3508L} presents $S^5$, and is one example of a cross-catalog among other works \citep{2020AJ....160..181J, 2021ApJ...911..149L, 2022ApJ...928...30L, 2024MNRAS.529.2413U}.
 Currently, about a hundred stellar streams are known in our Galaxy.
-\citet{2023MNRAS.520.5225M} provides a compilation of their tracks on the sky.
-Interestingly, only about 20 streams are associated to known Galactic globular clusters.
+\citet{2023MNRAS.520.5225M} compiled their tracks on the sky into a catalog.
+Interestingly, only about 20 streams are associated with known Galactic globular clusters.
 
@@
 \citet{2018ApJ...867..101B} reviewed this concept from an information-theoretic point of view, identifying which orbits and configurations of stellar streams yield the most information about the Galactic potential.
-However, the use of single streams for constraining the potential led to some ambiguous and not converging results over time.
+However, using single streams to constrain the potential led to ambiguous and non-converging results.
 For example, \citet{2010ApJ...718.1128L} made use of the Sagittarius stream to infer that the dark matter halo of our Galaxy has a triaxial shape \citep[but see also][]{2004MNRAS.351..643H, 2005ApJ...619..800J, 2005ApJ...619..807L}; \citet{2016ApJ...833...31B} concluded that the dark matter halo of our Galaxy is nearly spherical at the distances of the Palomar~5 and GD-1 streams.
-The two contrasting results could also show that at one distance is the halo is triaxial and at another spherical--highlighting the need for streams at different distances to map the halo shape.
+The two contrasting results could also show that the halo is triaxial at one distance yet spherical at another--highlighting the need for streams at different distances to map the halo shape.
 Recently, \citet{2024ApJ...967...89I} constructed a Milky Way model by applying a Monte Carlo Markov Chain (MCMC) fitting procedure.
@@
 
-  % \citet{2022MNRAS.516.5331M} showed that many very metal poor streams exist in the Milky Way.
-On the other hand, it is possible that some globular clusters do not loose a significant number of stars to create tidal tails.
-For instance, \citet{2022A&A...667A.112V} demonstrated that clusters existing in dark matter subhaloes can place the tidal radius far beyond their half mass radii and sheild stars from escape.
+  % \citet{2022MNRAS.516.5331M} showed that many metal-poor streams exist in the Milky Way.
+On the other hand, it is possible that some globular clusters do not lose a significant number of stars to create tidal tails.
+For instance, \citet{2022A&A...667A.112V} demonstrated that clusters existing in dark matter subhaloes could place the tidal radius far beyond their half-mass radii and shield stars from escape.
 Another factor is indeed detectability.
-Additionally, there are many cluster in the internal regional of the galaxy, a region overwhich \texttt{streamfinder} was not applied since the high stellar density implicates a computational cost that is currently far to expensive.
-Also, \citet{2018MNRAS.474.2479B} demonstrates that internal dynamics preferentially kicks out small stars first, which have lower luminosities.
-Earlier escape times places larger distances to the cluster and make detection more difficult.
-
-  Beyond the global visible and dark mass distribution streams can be used also to infer the granularity of the dark matter, that is the mass and density of the sub-halos populating our Galaxy.
+Additionally, there are many clusters in the internal region of the galaxy, a region over which \texttt{streamfinder} was not applied since the high stellar density implies a computational cost that is far too expensive.
+Also, \citet{2018MNRAS.474.2479B} demonstrates that internal dynamics preferentially kick out small stars first, which have lower luminosities.
+Earlier escape times place larger distances to the cluster, making detection more difficult.
+
+  Beyond the global visible and dark mass distribution streams, they can also be used to infer the granularity of the dark matter, that is, the mass and density of the sub-halos populating our galaxy.
 According to simulations by \citet{2008MNRAS.391.1685S}, $\Lambda$~Cold Dark Matter ($\Lambda$CDM) predicts that galaxies grow hierarchically, with dark matter clumps forming at a wide range of masses and sizes.
 These clumps, or \textit{sub-halos}, are predicted to follow a mass distribution with a power-law slope slightly shallower than -2.0.
-To date, the smallest observed dark matter halo was detected through gravitational lensing in an Einstein ring with a mass of $10^8$ solar masses by \citet{2012Natur.481..341V}.
+\citet{2012Natur.481..341V} detected the smallest observed dark matter halo through gravitational lensing in an Einstein ring with a mass of $10^8$ solar masses.
 However, some models predict that dark matter clumps could exist down to at least the mass of Earth-like planets \citep[see][and discussion in \citet{2021arXiv211101148A}]{2005JCAP...08..003G}.
@@
 Interestingly enough, as stated in \citet{2019ApJ...880...38B}, the recovered properties of this sub-halo (mass and size) were denser than those with the $\Lambda$CDM mass-size relationship presented in \cite{2017MNRAS.466.4974M}.
-The above-mentioned works are only a few of a much more extensive literature that has explored the impact of dark matter sub-halos in simulated streams \citep{2016ApJ...828L..10H, 2021MNRAS.507.1999H, 2021JCAP...10..043B, 2024arXiv240402953H, 2024arXiv241021174N}, or searched for their traces in observed streams \citep{2016MNRAS.460.2711T, 2017MNRAS.470...60E, 2020ApJ...889...70B, 2020ApJ...892L..37B}.
-
-  In contrast, a more limited number of works have been interested in understanding whether other structures, e.g.
-from baryonic matter, can cause variations in the density of streams and gaps that can be assimilated and confused with those produced by dark matter sub-halos.
-Among these works, it is worth mentioning the results of \citet{2017NatAs...1..633P} that suggests how the presence of the bar at the center of the Galaxy can perturb the characteristics of a stream such as Palomar~5 and generate gaps along its tail.
-That the Galactic bar could have an influence on streams morphology was also discussed by \citet{2016MNRAS.460..497H} and \citet{2016ApJ...824..104P}, in the case of the Ophiuchus stream \citep{2014MNRAS.443L..84B}.
-Besides the Galactic bar, also giant molecular clouds can produce gaps in stellar streams, as shown by \citet{2016MNRAS.463L..17A}.
-All these works thus seem to indicate that baryonic structures can play an important role in tail morphology.
-In this context,  an extensive numerical study, specifically focused on modeling the tails of Palomar~5, under the influence of the Galactic bar, spiral arms, giant molecular clouds and globular clusters, has been realized by \citet{2019MNRAS.484.2009B}, who concluded that both the influence of the bar and that of the giant molecular clouds can leave imprints in Palomar~5 tidal tails similar to those left by dark matter sub-halos, while the effect of globular clusters was found to be negligible.
-
-  Few studies have specifically investigated the effect of globular clusters on stellar streams. Here's an improved version with clearer phrasing and better flow:
-
-  \citet{2017MNRAS.470...60E} concluded that the observed density variations in the tails of Palomar~5 could not be attributed to the passage of globular clusters.
-Their analysis focused on the characteristics of the observed gaps and involved constraining properties of the progenitor using reconstructive modeling.
+The works above are only a few of a much more extensive literature that has explored the impact of dark matter sub-halos in simulated streams \citep{2016ApJ...828L..10H, 2021MNRAS.507.1999H, 2021JCAP...10..043B, 2024arXiv240402953H, 2024arXiv241021174N}, or searched for their traces in observed streams \citep{2016MNRAS.460.2711T, 2017MNRAS.470...60E, 2020ApJ...889...70B, 2020ApJ...892L..37B}.
+
+  In contrast, a more limited number of works have been interested in understanding whether other structures, e.g., from baryonic matter, can cause variations in the density of streams and gaps that can be assimilated and confused with those produced by dark matter sub-halos.
+Among these works, it is worth mentioning the results of \citet{2017NatAs...1..633P} that suggest how the presence of the bar at the center of the Galaxy can perturb the characteristics of a stream such as Palomar~5 and generate gaps along its tail.
+That the Galactic bar could have an influence on stream morphology was also discussed by \citet{2016MNRAS.460..497H} and \citet{2016ApJ...824..104P}, in the case of the Ophiuchus stream \citep{2014MNRAS.443L..84B}.
+Besides the Galactic bar, giant molecular clouds can also produce gaps in stellar streams, as shown by \citet{2016MNRAS.463L..17A}.
+All these works thus indicate that baryonic structures can play an important role in tail morphology.
+In this context,  an extensive numerical study, specifically focused on modeling the tails of Palomar~5, under the influence of the Galactic bar, spiral arms, giant molecular clouds, and globular clusters, has been realized by \citet{2019MNRAS.484.2009B}, who concluded that both the influence of the bar and that of the giant molecular clouds can leave imprints in Palomar~5 tidal tails similar to those left by dark matter sub-halos.
+In contrast, they found the effect of globular clusters to be negligible.
+
+  Few studies have specifically investigated the effect of globular clusters on stellar streams.
+\citet{2017MNRAS.470...60E} concluded that globular clusters could not be responsible for the observed density variations in the tails of Palomar~5.
+Their analysis focused on the observed gaps' characteristics and involved constraining progenitor properties using reconstructive modeling.
 By trial and error, they identified a specific configuration of masses, sizes, impact parameters, times of impact, and relative velocities for two perturbers that successfully reproduced the observed density distribution.
-However, their method does not perform full forward modeling of the entire globular cluster system on Palomar~5's stream, as we do in this work.
-While they suggest that the impact rate of globular clusters is likely less significant—given their lower abundance compared to the expected population of dark matter sub-halos—-they do not explore this aspect in detail.
-However, they do state that it is indeed an avenue for future investigation.
-
-  More recently \citet{2022ApJ...941..129D} have examined the possibility that the gaps found in the GD-1 stellar stream could be due to the close passage of globular clusters, concluding that this scenario is highly unlikely.
-These first works seem to suggest that the impact of globular clusters on stellar streams is negligible.
-This does not necessarily need to be the general case, especially for streams of clusters as Palomar~5, which are confined in the inner 20~kpc of the Galaxy, where many other globular clusters do also orbit.
+However, as we do in this work, their method does not perform full forward modeling of the entire globular cluster system on Palomar~5's stream.
+While they suggest that the impact rate of globular clusters is likely less significant—given their lower abundance compared to the expected dark matter sub-halos population-they do not explore this aspect in detail.
+However, they state that it is an avenue for future investigation.
+
+  More recently, \citet{2022ApJ...941..129D} have examined the possibility that gaps in the GD-1 stellar stream could be due to the close passage of globular clusters, concluding that this scenario is improbable.
+These first works suggest that the impact of globular clusters on stellar streams is negligible.
+This result does not necessarily need to be the general case, especially for streams of clusters such as Palomar~5, which live in the inner 20~kpc of the Galaxy, where many other globular clusters also orbit.
 For example, \citet{2023A&A...678A..69I} showed that a globular cluster can even collide with other clusters, which implies that cluster stream collisions should happen much more frequently since streams are far more extended than clusters.
 
-  In the present study we wish to fill this gap in the literature of numerically modeling cluster-stream interactions.
-We seek to quantify the importance of the effects of passing globular clusters in the vicinity of streams, to understand whether these systems can also be effective - and how often - in altering the distribution of stars in the tails and producing under-dense regions or gaps.
-To this end, on the following pages we will present the results of simulations of the streams of Palomar~5, subject to the gravitational interaction with the set of 165 Galactic globular clusters for which positions and velocities are known to date, and for which orbits can therefore be reconstructed \citep{2021MNRAS.505.5957B}.
-We chose to simulate Palomar~5 because it is a typical halo cluster with extended tails, and because it is a cluster for which the effect of baryonic structures on its stream has already been studied.
-As we will show, and in tension with previous claims, the close passage of other clusters with the stream of Palomar~5 is not a rare event, and indeed, in the 50 simulations we ran, we find the formation of numerous gaps, for an average of 1.5 gaps per simulation, generated by 18 different clusters among the whole system of Galactic globular clusters.
+  In the present study, we wish to fill this gap in the literature on numerically modeling cluster-stream interactions.
+We seek to quantify the importance of the effects of passing globular clusters in the vicinity of streams to understand whether these systems can also be effective, and how often they alter the distribution of stars in the tails and produce underdense regions or gaps.
+To this end, on the following pages, we will present the results of simulations of the streams of Palomar~5, subject to the gravitational interaction with the set of 165 Galactic globular clusters for which positions and velocities are known to date and for which orbits can therefore be reconstructed \citep{2021MNRAS.505.5957B}.
+We chose to simulate \textbf{\paola{streams formed from a cluster with the current characteristics of}} Palomar~5 because \textbf{\paola{this latter}} is a halo cluster with extended tails and because it is a cluster for which the effect of baryonic structures on its stream has already been studied.
+As we will show, and in tension with previous claims, the close passage of other clusters with \textbf{\paola{such a stream}} is not rare.
+Indeed, in the 50 simulations we ran, we find the formation of numerous gaps, for an average of 1.5 gaps per simulation, generated by 18 different clusters among the whole system of Galactic globular clusters.
 
@@
     \caption{Simulated Palomar~5 stream created by modeling the host cluster as a Plummer sphere disrupting within an axis-symmetric Galactic potential plus the gravitational effect of 164 other galactic globular clusters.
-The top panel shows the distribution of star-particles that escaped the cluster due to tidal forces.
+The top panel shows the distribution of star particles that escaped the cluster due to tidal forces.
 The bottom panel shows the 1D density profile marginalized over longitude.
-The gray fill shows a reference simulation where the same conditions to produce the stream are used but the effect of other globular clusters are not included.
-The large central peak in density are of the particles still bound to Palomar~5.
+The gray fill shows a reference simulation that uses the same conditions to produce the stream but does not include the other globular clusters.
+The large central peak in density is of the particles still bound to Palomar~5.
 $\ell_0,b_0$ are located Palomar~5's cluster's center of mass.
-Two large gaps are present and are due to the passage of two globular clusters.
-$N_p$ indicates the number of particles in a bin while $N$ is the total number of particles which is 100,000.}
+Two large gaps are present due to the passage of two globular clusters.
+$N_p$ indicates the number of particles in a bin, while $N$ is the total number of particles, which is 100,000.}
     \label{fig:stream_on_sky}
@@
 \section{Methods}
-  The most accurate model of the Palomar~5 stream would involve full modeling of the internal dynamics of the cluster.
-This would mean computing $N$-body interactions with a $\mathcal{O}(N^2)$ computation time, stellar evolution, supernovae, an initial mass distribution, treatment of binary star systems, etc.
+  The most accurate model of the Palomar~5 stream would involve full modeling of the internal dynamics of the cluster, which would mean computing $N$-body interactions with a $\mathcal{O}(N^2)$ computation time, stellar evolution, supernovae, an initial mass distribution, treatment of binary star systems, etc.
 \citep[for such an example, see][]{2021NatAs...5..957G, 2016MNRAS.458.1450W}.
 Instead, we opt for solving the \textit{restricted-three body problem}, or as known as the \texttt{particle-test method} as we did for \citet{2023A&A...673A..44F}, which we describe here for completeness.
-As demonstrated by \citet{2012A&A...546L...7M}, although the restricted three-body problem neglects the internal evolution of the cluster, it still reproduces very similar stream properties.
-This is because the model captures key extra-cluster physics, such as disk shocking and epicyclic stripping.
-
-  \subsection{Numerical Methodology}
-    We begin by extracting positions in the sky, proper motions, line-of-sight velocities and distances, as well as masses and half-mass radii, of 165 globular clusters from the Galactic globular cluster catalog by \cite{2021MNRAS.505.5957B}.\footnote{The Baumgardt catalog has been assembled across a series of works, see: \cite{2020PASA...37...46B,2019MNRAS.482.5138B,2018MNRAS.478.1520B}.
+As demonstrated by \citet{2012A&A...546L...7M}, although the restricted three-body problem neglects the internal evolution of the cluster, it still reproduces very similar stream properties since the model captures key extra-cluster physics, such as disk shocking and epicyclic stripping.\\
+  \textbf{\paola{Below, we first present the approach used to include the Galactic globular cluster system in our modeling (Sect.~\ref{numerical_meth}), highlighting the similarities and differences with respect to what we did in \citet{2023A&A...673A..44F}; we then summarize the method used to model the mass loss from the cluster (Sect.~\ref{sec:mass_loss}) and finally discuss the quality of the numerical integration (Sect.~\ref{num_quality}).
+}}
+
+  \subsection{Numerical Methodology}\label{numerical_meth}
+    We begin by extracting positions in the sky, proper motions, line-of-sight velocities, and distances, as well as masses and half-mass radii, of 165 globular clusters from the Galactic globular cluster catalog by \cite{2021MNRAS.505.5957B}.\footnote{The Baumgardt catalog has been assembled across a series of works, see: \cite{2020PASA...37...46B,2019MNRAS.482.5138B,2018MNRAS.478.1520B}.
 The catalog can be found on the World Wide Web at \href{https://people.smp.uq.edu.au/HolgerBaumgardt/globular/}{https://people.smp.uq.edu.au/HolgerBaumgardt/globular/}.} We then convert the initial conditions from sky coordinates into a Galactocentric reference frame, by adopting a velocity for the local standard of rest of $v_{\text{LSR}} = 240$~km~s$^{-1}$ and a peculiar velocity of the Sun equal to $(U_\odot, V_\odot, W_\odot)=(11.1, 12.24, 7.25)$~km~s$^{-1}$, as reported by \citet{2012MNRAS.427..274S}.
-The position of the Sun was set to $(x_\odot,y_\odot,z_\odot) = (-8.34,0,0.027)$~kpc, vertical position above the disk was taken from \citet{2001ApJ...553..184C} and the distance of the Sun to the Galactic center was taken from \citet{2014ApJ...783..130R}.
+We set the Sun's position to $(x_\odot,y_\odot,z_\odot) = (-8.34,0,0.027)$~kpc.
+We took the vertical position above the disk from \citet{2001ApJ...553..184C} and the distance of the Sun to the Galactic center from \citet{2014ApJ...783..130R}.
 These transformations were performed using \texttt{astropy} \citep{2013A&A...558A..33A}.
@@
 This model is time-independent throughout our simulations.
-This model satisfies a series of observational constraints such as local solar stellar density, the Galactic rotation curve, etc., similarly to other Galactic models such as \texttt{MWpotential2014} from \citet{2015ApJS..216...29B} and \texttt{McMillian2017} from \citet{2017MNRAS.465...76M}.
-However, to balance data volume and computation time, we use only one Galactic potential, which should suffice.
+This model satisfies a series of observational constraints such as local solar, stellar density, the Galactic rotation curve, etc., similarly to other Galactic models such as \texttt{MWpotential2014} from \citet{2015ApJS..216...29B} and \texttt{McMillian2017} from \citet{2017MNRAS.465...76M}.
+However, we use only one Galactic potential to balance data volume and computation time, which should suffice.
 \citet{2021MNRAS.505.5978V} found that only a few outer globular clusters are strongly affected by different potential models.
 Generally, kinematic uncertainties are the dominant factor in differences between orbital solutions per cluster.
-Similarly, \citet{2024MNRAS.528.5189G} generated a globular cluster mass-loss catalog using seven different potential models and found their debris distributions to be rather model-independent, similarly to \citet{2023A&A...673A..44F}.
-While the clusters' exact positions in time may depend on the model, we assert that interaction rates and stream formation are largely independent the choice of Galactic potential model.
-
-    Lastly, we select an integration time of 5~Gyr as a compromise between maximizing interaction statistics and adhering to the assumption that the galaxy can be modeled as a time-independent, constant mass distribution.
-\citet{2023A&A...673A.152I} analyzed the orbits of the Galactic cluster population using the same initial conditions as in this work, within five live Milky Way-like potentials from IllustrisTNG \citep{2018MNRAS.473.4077P}.
-They found that in all sampled potentials, orbital changes remain minimal over 5~Gyr, becoming significant only at earlier look-back times when the host galaxy had either significantly less mass or was undergoing a merger event.
+Similarly, \citet{2024MNRAS.528.5189G} generated a globular cluster mass-loss catalog using seven different potential models and found their debris distributions to be rather model-independent, similar to \citet{2023A&A...673A..44F}.
+While the clusters' exact positions in time may depend on the model, we assert that interaction rates and stream formation are largely independent of the choice of the Galactic potential model.
+
+    Lastly, we select an integration time of 5~Gyr as a compromise between maximizing interaction statistics and modeling the galaxy as a time-independent, constant mass distribution.
+\citet{2023A&A...673A.152I} analyzed the orbits of the Galactic cluster population using the same initial conditions as in this work within five live Milky Way-like potentials from IllustrisTNG \citep{2018MNRAS.473.4077P}.
+They found that in all sampled potentials, orbital changes remain minimal over 5~Gyr, becoming significant only at earlier look-back times when the host galaxy had significantly less mass or was undergoing a merger event.
 
@@
     There is a primary methodological departure from \citet{2023A&A...673A..44F}.
-In that work, globular clusters evolved under the gravitational effect of the Galaxy only, while now we include also the effect of all the other Galactic globular clusters, by taking into account the direct $N$-body interactions among them.
-To quantify these interactions, all clusters are represented by Plummer spheres, each with mass and half-mass radii as reported in the Baumgardt catalog \citep{2021MNRAS.505.5957B}.
-For the remainder of this paper, we refer to these simulations, where the globular clusters interactions are taken into account as \texttt{full} simulations.
+In that work, globular clusters evolved under the gravitational effect of the Galaxy only, while now we also include the effect of all the other Galactic globular clusters by considering the direct $N$-body interactions between them.
+First, all clusters are represented by Plummer spheres, each with its own mass and half-mass radius as reported in the Baumgardt catalog \citep{2021MNRAS.505.5957B}.
+For the remainder of this paper, the \texttt{full} simulations consider the gravitational forces from the globular cluster interactions.
 \\
-    For these simulations, we proceed into two steps:
+    For these simulations, we proceed in two steps:
       \begin{enumerate}
           \item First, starting from the Galactocentric positions and velocities of all 165 Galactic globular clusters, we integrate their orbits back in time for 5~Gyr under the influence of the Galaxy itself and their mutual influence.
-In the backward integration, the system of equations of motion for the globular clusters are thus:
+In the backward integration, the system of equations of motion for the globular clusters is thus:
           \begin{equation}
-          \ddot{\vec{r}}_i = -\nabla \Phi(R_i,z_i) + \left.\sum_{j\neq i}^{N_{GC}} \frac{Gm_j}{\left(|\vec{r}_j - \vec{r}_i|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j - \vec{r}_i\right),
+          \ddot{\vec{r}}_i = -\nabla \Phi + \left.\sum_{j\neq i}^{N_{GC}} \frac{Gm_j}{\left(|\vec{r}_j - \vec{r}_i|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j - \vec{r}_i\right),
           \end{equation}\label{eq:GCNBody} 
-          \noindent where $\vec{r}$ indicates the galactocentric position vector, the index $i$ indicates the globular cluster of interest with $R_i,z_i$ being its galactocentric positions in cylindrical coordinates; the index $j$ indicates the other globular clusters that are summed over.
-$N_{GC}$ is the total number of globular clusters which in this study is 165, $m_j$ is the mass of the j-th cluster in the sample, $b_j$ is its Plummer scale radius, and $\vec{r_j}$ is its galactocentric position.
+          \noindent where $\vec{r}$ indicates the galactocentric position vector, the index $i$ indicates the globular cluster of interest; the index $j$ indicates the other globular clusters that are summed over.
+$N_{GC}$ is the total number of globular clusters, which in this study is 165, $m_j$ is the mass of the j-th cluster in the sample, $b_j$ is its Plummer scale radius, and $\vec{r_j}$ is its galactocentric position.
 $\Phi$ represents the same Galactic smooth potential that we discussed previously\citep[][Model~II, in the present case]{2017A&A...598A..66P}.
-Note that the masses and sizes of the globular clusters are kept constant in these simulations and not allowed to vary with time.
-This equally implies that we do not take into account the internal evolution of the globular clusters in these models.
-We discuss the implications of the modeling limitations in the Discussion.
-
-        \item Once the positions and velocities of the entire globular cluster system 5~Gyr ago are found, we sample Palomar~5 with 100,000 particles from a Plummer distribution taking the mass and half-mass radius from the Baumgardt catalog.
-We then integrate the evolution of these particles forward in time to the present day, by taking into account that each particle feels the gravitational potential of the Galaxy, that of its host cluster, and that of all the other clusters in the Galaxy.
-Note that we do not take into account self-gravity among particles.
+Note that the masses and sizes of the globular clusters are kept constant in these simulations and are not allowed to vary with time, which means that we do not consider their internal evolution.
+In the Discussion section, we discuss the implications of the modeling limitations.
+
+        \item Once the we find the positions and velocities of the entire globular cluster system 5~Gyr ago, we sample Palomar~5 with 100,000 particles from a Plummer distribution, taking the mass and half-mass radius from the Baumgardt catalog \paola{can you add here the current values of mass and size of Pal 5 in Baumgardt?}.
+We then integrate the evolution of these particles forward in time to the present day by taking into account that each particle feels the gravitational potential of the Galaxy, that of its host cluster, and that of all the other clusters in the Galaxy.
+Note that we do not take self-gravity among particles into account.
 The particles experience the gravitational field yet do not contribute to it, which is a common assumption made in Galactic dynamics since the mass of an individual star is negligible compared to the mass of the larger dynamical system.
-The equation of motion of a generic particle among the 100,000 which populate Palomar~5 is thus:
+The equation of motion of a generic particle among the 100,000 that populate Palomar~5 is thus:
         \begin{equation}
-          \ddot{\vec{r}}_p = -\nabla \Phi(R_p,z_p) + \left.\sum_{j}^{N_{GC}} \frac{Gm_j}{\left(|\vec{r}_j(t) - \vec{r}_p|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j(t)- \vec{r}_p\right),
+          \ddot{\vec{r}}_p = -\nabla \Phi + \left.\sum_{j}^{N_{GC}} \frac{Gm_j}{\left(|\vec{r}_j(t) - \vec{r}_p|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j(t)- \vec{r}_p\right),
           \end{equation} \label{eq:equation_of_motion_particle} where the index $p$ represents one of the 100,000 particles of interest, $\vec{r_p}$ being its position, and $j$ indexes over the globular clusters as in Eq.~\ref{eq:GCNBody}.
-Note that in Eq.~\ref{eq:equation_of_motion_particle} the positions of the globular clusters are time-dependent since they are being loaded during this step and not computed, unlike Eq.~\ref{eq:GCNBody}.
+Note that in Eq.~\ref{eq:equation_of_motion_particle}, the positions of the globular clusters are time-dependent since they are being loaded during this step and not computed, unlike Eq.~\ref{eq:GCNBody}.
 
@@
 
-        The procedure described so far has been repeated 50 times, each time generating a new set of initial conditions, given the uncertainties on proper motions, line-of-sight velocities, distances to the Sun, and masses, of all clusters, as reported in the Baumgardt catalog.
-We handle these uncertainties through a Monte-Carlo approach, by sampling them with a Gaussian distribution, and by taking into account the covariance term between the proper motions.
-For the first simulation, we use the most probable values for the initial conditions.
-The uncertainties are sampled for all globular clusters.
-Additionally, for each re-sampling of Palomar~5's mass, we also resample the distribution of the 100,000 star-particles.
-
-      During the integration, intermediate snapshots were saved to facilitate the analysis of stellar streams and the effects of cluster impacts.
-Specifically, for each of the 50 realizations of the Palomar~5 stream, we saved $5000$ intermediate time-steps which is a temporal resolution of 1 million years.
+        The procedure described so far has been repeated 50 times, generating a new set of initial conditions each time, given the uncertainties on proper motions, line-of-sight velocities, distances to the Sun, and masses of all clusters, as reported in the Baumgardt catalog.
+We handle these uncertainties through a Monte-Carlo approach by sampling them with a Gaussian distribution and considering the covariance term between the proper motions.
+We use the most probable values for the initial conditions in the first simulation.
+We sample the uncertainties for all globular clusters.
+Additionally, for each resampling of Palomar~5's mass, we also resample the distribution of the 100,000 star particles.
+
+      During the integration, we save intermediate snapshots to facilitate the analysis of stellar streams and the effects of cluster impacts.
+Specifically, for each realization of the Palomar~5 stream, we saved $5000$ in snapshots, equivalent to a temporal resolution of 1 million years.
 We provide the parameters that specify our data volume in Table~\ref{tab:data_volume}.
@@
 
-    To quantify the impact of globular cluster passages to the density of Palomar~5 stream, we performed a second set of simulations, which we refer to as the \texttt{reference} simulations in this paper.
-These \texttt{reference} simulations use the same 50 sets of initial conditions as the \texttt{full} simulations, the same Galactic potential but exclude mutual interactions between globular clusters.
+    To quantify the impact of globular cluster passages on the density of the Palomar~5 stream, we performed a second set of simulations, which we refer to as the \texttt{reference} simulations in this paper.
+These \texttt{reference} simulations use the same 50 sets of initial conditions as the \texttt{full} simulations, the same Galactic potential, but exclude mutual interactions between globular clusters.
 The approach adopted for this second set of simulations is thus equivalent to that adopted already in \citet{2023A&A...673A..44F}.
-In Eq.~\ref{eq:GCNBody} only the gradient of the Galactic potential  is considered.
-In Eq.~\ref{eq:equation_of_motion_particle}, of the second term on the right-side of the equation, only the influence of Palomar~5's Plummer sphere on Palomar~5's particles is considered.
-In other words, sum iterates over only one globular cluster, the host.
-All the interactions with the other clusters in the Galaxy are indeed omitted.
-
-  \subsubsection*{Numerical Stability}
-
-    The integration was conducted using a leapfrog integrator, chosen for its ability to preserve phase-space volume and conserve the Hamiltonian with each integration step.
-This method is preferable to, for instance, a Runge-Kutta scheme, which can introduce non-physical and significant numerical errors in systems that require long-term stability and energy conservation.
-One drawback of the leapfrog integrator is that it requires a uniform time-step across the entire computation, leading to unnecessary computations for a particle after it escapes form the host cluster.
+In Eq.~\ref{eq:GCNBody}, only the gradient of the Galactic potential is considered.
+In Eq.~\ref{eq:equation_of_motion_particle}, of the second term on the right side of the equation, only the influence of Palomar~5's Plummer sphere on Palomar~5's particles is considered.
+In other words, the sum iterates over only one globular cluster, the host.
+We omit all interactions with the other clusters.
+
+\textbf{\paola{\subsection{Mass loss}}}\label{sec:mass_loss}
+
+\textbf{\paola{Each of the 100,000 particles that initially populate the cluster undergoes, over time, acceleration due to the Pal~5 potential, whose mass and half-mass radius we assume to be fixed over time, and to the Galactic potential.
+At each time step, a certain number of particles will therefore acquire sufficient energy to no longer be gravitationally bound to the cluster itself and thus go on to populate the streams, whose mass and spatial extent grow over time.
+It is important to note that in the approach used:}}
+\textbf{\paola{
+\begin{enumerate}
+    \item The masses and sizes of the clusters (and therefore the parameters of the Plummer potentials) do not change over time, which is a blatant simplification, because in a self-consistent approach, these parameters would vary.
+    \item We use the same initial conditions for Pal~5 progenitor as it has today, and this is also a simplification, since Pal~5 - 5 Gyr years ago - must have contained at least part of the mass estimated today in its tails\footnote{\textbf{\paola{We use the same approach (i.e.
+time-independent masses and sizes) to model the whole set of globular clusters.}}}.
+\end{enumerate}}}
+
+\textbf{\paola{The assumption in point 2 is a direct consequence of the approach described in point 1.
+Starting from a cluster with mass and size similar to the current ones can lead to streams with lower velocity dispersions than those we would have obtained if we had used a self-consistent approach.
+In a future article, we will report on the study of gap survival times depending on the masses and sizes of progenitor clusters (Ferrone et al, in prep).
+We note, however, that simplifications of this kind are not uncommon in literature.
+\citet{2017NatAs...1..633P} discussed the formation of gaps in the Pal~5 tails and assumed a time-independent mass of $50,000 M_\odot$ for Pal~5, over the last 4~Gyr; \citet{2017MNRAS.470...60E} adopted a N-body approach to simulate Pal~5 stream, but used Pal~5 current conditions as their progenitor's initial conditions; \citet{2019MNRAS.484.2009B} simulated the Pal~5 stream as emerging from a stellar system with a velocity dispersion of 0.5~km/s (similar to that of particles escaping from our cluster, as we have verified).
+\\The characteristics of the streams modeled in this paper may be considered more representative of those of clusters that are now completely dispersed, i.e.
+it is conceivable that completely dispersed globular clusters that left behind a population of 'orphan' streams passed through characteristics similar to those of Pal~5 today (small masses and extended radii).
+In this sense, the initial conditions chosen (in terms of internal parameters) could be considered more representative of those of streams found in the Galaxy and for which the progenitor is now dissolved \citep[see, for example, the population of streams without progenitors described by][]{2024ApJ...967...89I} than those currently typical of Galactic globular clusters\footnote{\textbf{\paola{In this regard, we recall that \citet{2014ApJ...795...95B} modeled the GD-1 stream as the result of the dissolution of a cluster with a mass of $2 \times 10^4 M_\odot$, and a tidal radius of $0.07$~kpc.}}}.
+}}
+
+  \subsection{Numerical Stability}\label{num_quality}
+
+    We used a leapfrog integrator because of its ability to preserve phase-space volume and conserve the Hamiltonian with each integration step.
+For instance, this method is preferable to a Runge-Kutta scheme, which can introduce non-physical and significant numerical errors in systems that require long-term stability and energy conservation.
+One drawback of the leapfrog integrator is that it requires a uniform time step across the entire computation, leading to unnecessary computations for a particle after it escapes from the host cluster.
 However, energy conservation and phase-space volume preservation are paramount when modeling stellar streams.
-The time step was therefore set to be small enough to conserve energy for the most interior particles within the cluster---ensuring that a higher mass loss did not arise from numerical error.
-For these simulations, a time-step of 10,000 years was selected to maintain adequate energy conservation, with a median variation of $10^{-12} \frac{\Delta E}{E_0}$, where $E_0$ is a particle's initial energy, and $\Delta E$ is the difference between its final and initial energy.
+The time step was, therefore, set to be small enough to conserve energy for the most interior particles within the cluster---ensuring that a higher mass loss did not arise from numerical error.
+We found that a time-step of 10,000 years was adequate to maintain energy conservation, with a median variation of $10^{-12} \frac{\Delta E}{E_0}$, where $E_0$ is a particle's initial energy, and $\Delta E$ is the difference between its final and initial energy.
 
@@
 Therefore, any drift would arise from purely numerical error.
-We selected a time stamp for which the drift in the final position after forward integration, compared to the initial position from the backward integration, was consistently at least two orders of magnitude smaller than the Plummer scale radius used for Palomar~5.
-This high precision ensures that no fictitious numerical forces influenced the system, preventing any artificial mass loss or retention of star particles.
+We selected a timestamp for which the drift in the final position after forward integration, compared to the initial position from the backward integration, was consistently at least two orders of magnitude smaller than the Plummer scale radius used for Palomar~5.
+This high precision ensures that no fictitious numerical forces influence the system, preventing any artificial mass loss or retention of star particles.
 
@@
       \includegraphics[width=\linewidth]{histogram_impact_time.png}
-      \caption{The simulation time at which the impacts occurred for all gap causing flybys summed over all 50 simulations.
-Each perturbing cluster is label and is color consistent.
-The time axis is given in simulation units, with 1~s~kpc~km~$^{-1}$ corresponding to roughly 1~Gyr.
-Note that the y-axis is broken to accommodate the large number of encounters from NGC2808, without overshadowing the other interactions.}
+      \caption{The simulation time when the impacts occurred for all gap-causing flybys summed over all 50 simulations.
+Each perturbing cluster is labeled and color-consistent.
+We report the time axis in simulation units, with 1~s~kpc~km~$^{-1}$ corresponding to roughly 1~Gyr.
+Note that the plot breaks the y-axis to accommodate the large number of encounters from NGC~2808 without overshadowing the other interactions.}
       \label{fig:histogram_impact_time}
@@
 
-      % To our knowledge, this is the first study to investigate the gravitational interactions between all Galactic globular clusters.
-While we did not  observe significant changes to orbital properties of the clusters, when subject to their mutual interactions -- a result which is  expected given the small sizes and vast distances between the clusters -- we do
-    % This is no longer true after Sergey showed me the papers from Ischenko et al for the globular cluster ineractions
-
-    The presence of the other globular clusters has a clear effect on the properties of Palomar~5 stream.
-A particularly clear example of this effect is shown in Fig.~\ref{fig:stream_on_sky} which we selected for its prominent gaps.
-Two of these gaps are clearly visible in Galactic coordinates and become even more apparent when marginalizing over latitude to reconstruct the 1D density profile of the stream, as a function of longitude.
+    The presence of the other globular clusters affects the properties of the Palomar~5 stream.
+Fig.~\ref{fig:stream_on_sky} presents an obvious example of this effect, which we selected for its prominent gaps.
+Two of these gaps are visible in Galactic coordinates and become even more apparent when marginalizing over latitude to reconstruct the 1D density profile of the stream as a function of longitude.
 
     Regarding the shape of the density distribution, the central peak corresponds to the still-intact globular cluster whose stars have not yet escaped.
-The density peaks in the stream are of the same order as the cluster itself.
-This is inconsistent with reality, the cluster's peak density should be higher than the stream's peak density.
+The stream's density peaks are the same order as the cluster itself, which is inconsistent with reality; the cluster's peak density should be higher than the stream's.
 Of course, this discrepancy is a result of our modeling choices.
-We use the present day mass and radius for Palomar~5 for the whole duration of the simulation.
-Consequently the system is modeled as being less dense than it should have been.
-In turn, our simulations have a strong initial mass loss which adds to the amplitudes of the profile density peaks of Fig.~\ref{fig:stream_on_sky}.
+Since we use the present-day mass and radius for Palomar~5 for the whole simulation duration, the system is less dense than it should have been.
+In turn, our simulations have a strong initial mass loss, which adds to the amplitudes of the profile density peaks of Fig.~\ref{fig:stream_on_sky}.
 This inaccuracy is acceptable for the scope of this work.
@@
 
-    To make a more quantitative comparison between the \texttt{reference} and \texttt{full} simulations, we work in the tail coordinate system, in which the stream is aligned with the cluster's orbit.
-The definition of this coordinate system is based on the work of \citet{2004AJ....127.2753D} and is shown in Fig.~\ref{fig:TailCoordinates}.
-Briefly, in this system the $x_{\textrm{tail}}$ coordinate represents the position of a particle along the orbit relative to the globular cluster.
-Positive values of $x_{\textrm{tail}}$ are ahead of the cluster while negative $x_{\textrm{tail}}$ is behind the cluster.
+    To compare the \texttt{reference} and \texttt{full} simulations more quantitatively, we work in the tail coordinate system, in which the stream's central axis aligns with the cluster's orbit.
+We based this coordinate system on the work of \citet{2004AJ....127.2753D} and present it in Fig.~\ref{fig:TailCoordinates}.
+Briefly, in this system, the $x_{\textrm{tail}}$ coordinate represents the position of a particle along the orbit relative to the globular cluster.
+Positive values of $x_{\textrm{tail}}$ are ahead of the cluster, while negative $x_{\textrm{tail}}$ is behind the cluster.
 The $y_{\textrm{tail}}$ coordinate measures the particle's distance within the orbital plane, where positive values indicate that the particle is farther from the Galactic center and negative values indicate that it is closer.
-Fig.~\ref{fig:decomposition} shows a comparison between one of the 50 realizations of Palomar~5 stream taking into account the gravitational interactions with all other globular clusters in the Galaxy (top panel) and omitting them (bottom panel).
+Fig.~\ref{fig:decomposition} shows a comparison between one of the 50 realizations of the Palomar~5 stream, taking into account the gravitational interactions with all other globular clusters in the Galaxy (top panel) and omitting them (bottom panel).
 This comparison clearly shows the presence of two wide ($\sim$100~pc and $\sim$1~kpc) gaps in the leading tail and of a more subtle under-density at $x_{\textrm{tail}}\sim 5$~kpc (we refer the reader to Appendix~\ref{sec:gap_detection} for a detailed description of the under-densities and gaps detection method).
 
-    To find out which globular clusters were responsible of creating these gaps, and when close passages took place, we estimated the gravitational acceleration along the orbit of Palomar~5 and represented it in the ($t$, $\tau$) space.
-$t$ is the simulation time and $\tau$ is indicates how long it will take for Palomar~5 to reach, or how  long ago it passed, a given point on its orbit.
+    To find out which globular clusters were responsible for creating these gaps and when close passages took place, we estimated the gravitational acceleration along the orbit of Palomar~5.
+We represented it in the ($t$, $\tau$) space.
+$t$ is the simulation time, and $\tau$ indicates how long it will take for Palomar~5 to reach a given point in its orbit or how long ago it passed.
 The use of $\tau$ is advantageous because the growth of the stream is approximately linear in $\tau$.
-On the other hand, in physical space streams on eccentric orbits indeed grow in time yet the growth rate is modulated with periodic expansion and contraction depending on the orbital phase \citep[see the top panel of Fig.~5.][for an example]{2016MNRAS.457.3817S}.
-Adopting this time space, and reporting the gravitational acceleration along Palomar~5 orbit in this space, the identification of the globular clusters which produced the perturbation, and the time at which it occurred becomes straight forward.
-We refer the reader to Appendix~\ref{sec:Perturber_Identification} for an in-depth description of the procedure.
-In this way, we could identify that the clusters responsible for creating gaps in the simulation of Palomar~5 stream reported in Fig.~\ref{fig:decomposition} are NGC~2808, NGC~7078 and NGC~104, and that their close passages occurred 200~Myr, -1.9 and -2.1~Gyr ago, respectively.
-
-    To further probe that the three clusters above are responsible of producing the gaps observed in this simulation, we conducted additional experiments by including only the perturbation of each of these three clusters at a time, and neglecting the gravitational perturbations of all the other clusters in the Galaxy.
-The results of these simulations are reported in the middle panels of Fig.~\ref{fig:decomposition} and clearly show that NGC~2808, NGC~7078 and NGC~104 are the clusters responsible for creating the under-dense regions observed in the leading tails of Palomar~5.
+    On the other hand, streams in physical space grow linearly.
+Yet, their eccentricity modulates this growth with periodic expansion and contraction depending on the orbital phase \citep[see the top panel of Fig.~5.][for an example]{2016MNRAS.457.3817S}.
+Adopting this time-space and reporting the gravitational acceleration along the Palomar~5 orbit in this space, identifying the globular clusters that produced the perturbation and the time it occurred becomes straightforward.
+We refer the reader to Appendix~\ref{sec:Perturber_Identification} for an in-depth procedure description.
+In this way, we could identify that the clusters responsible for creating gaps in the simulation of the Palomar~5 stream reported in Fig.~\ref{fig:decomposition} are NGC~2808, NGC~7078, and NGC~104 and that their close passages occurred 200~Myr, -1.9, and -2.1~Gyr ago, respectively.
+
+    To further probe whether the three clusters above are responsible for producing the gaps observed in this simulation, we conducted additional experiments by including only the perturbation of each cluster at a time and neglecting the gravitational perturbations of all the other clusters in the Galaxy.
+
+    To verify that the three suspected clusters are responsible for producing the gaps, we conducted experiments by including one perturbed at a time and excluding all others.
+We present these results in the middle panels of Fig.~\ref{fig:decomposition} and clearly show that NGC~2808, NGC~7078, and NGC~104 are the clusters responsible for creating the under-dense regions observed in the leading tails of Palomar~5.
 It is worth noting that the times at which the passages of these clusters occurred, according to our analysis, are in agreement with the observed width of the corresponding gaps: the encounter with NGC~2808 being very recent (only 200~Myr ago), its induced gap is still very thin, because it takes time for a perturbation to grow into an extended gap, as it is the case for those induced by the passages of  NGC~7078 and NGC~104, which occurred in earlier times.
@@
 
-    The analysis presented in Fig.~\ref{fig:decomposition} has been repeated for the whole set of simulations, and is reported in Appendix~\ref{sec:gallery_of_gaps} for completeness.
+    The analysis presented in Fig.~\ref{fig:decomposition} has been repeated for the whole set of simulations, and Appendix~\ref{sec:gallery_of_gaps} reports all of them for completeness.
 For all of the streams, morphologically speaking, the only changes appear to be the existence of gaps or not.
@@
 
-    From this analysis we can derive a statistical view of: 1) the number of gaps generated on the Palomar~5 stream by the whole system of Galactic globular clusters; 2) the clusters which generated these gaps and the time history of these perturbations.
-From this, we can then quantify: 3) the properties of the perturbers (their masses, sizes and their orbital parameters) as well as 4) the impact geometry of the encounters, which allows us to understand which encounters are more favorable to generate gaps in the Palomar~5 stream.
-In the following, we will present results of this analysis, by starting from addressing points 1) and 2).
+    From this analysis, we can derive a statistical view of 1) the number of gaps generated on the Palomar~5 stream by the system of Galactic globular clusters, 2) the clusters that generated these gaps, and the time history of these perturbations.
+From this, we can then quantify 3) the properties of the perturbers (their masses, sizes, and orbital parameters) as well as 4) the impact geometry of the encounters, which allows us to understand which encounters are more favorable to generate gaps in the Palomar~5 stream.
+In the following, we will present the results of this analysis, addressing points 1) and 2).
 
@@
 
-    By applying the methodology described above to the whole set of simulations, we can reconstruct the history of close passages of Galactic globular clusters to Palomar~5's stream, in the last 5~Gyr, which -- we remind the reader -- is the time interval investigated in our simulations.
-The results of this analysis are reported in Fig.~\ref{fig:histogram_impact_time} and in Table~\ref{tab:gaps_per_perturber}.
-NGC~2808 impacts Palomar~5's stream about 200 Myr ago in 44 out of 50 simulations, creating a small gap at a similar position to the one reported in Fig.~\ref{fig:decomposition} in each case.
-Since this interaction occurred less than one orbital period ago, despite the uncertainties the orbital solutions remain nontheless similar and thus produce consistent results across all simulations.
-However, as we continue to turn back time further, the uncertainties in the initial conditions allow the for the various orbital solutions to diverge from one another.
-Thus, in one configuration a cluster can impact the stream at a given time and yet at the same moment in a different set of initial conditions, it could be on the other side of the galaxy.
+    By applying the methodology described above to the whole set of simulations, we can reconstruct the history of close passages of Galactic globular clusters to Palomar~5's stream in the last 5~Gyr, which -- we remind the reader -- is the time interval investigated in our simulations.
+Fig.~\ref{fig:histogram_impact_time} and Table~\ref{tab:gaps_per_perturber} present results of this analysis.
+NGC~2808 impacted Palomar~5's stream about 200 Myr ago in 44 out of 50 simulations, creating a small gap at a similar position to the one reported in Fig.~\ref{fig:decomposition} in each case.
+Since this interaction occurred less than one orbital period ago, despite the uncertainties, the orbital solutions remain similar and thus produce consistent results across all simulations.
+However, as we continue to turn back time further, the uncertainties in the initial conditions allow the various orbital solutions to diverge from one another.
+Thus, in one configuration, a cluster can impact the stream at a given time, and yet at the same moment, in a different set of initial conditions, it could be on the other side of the galaxy.
+\textbf{\paola{Interestingly, some gaps, formed by close passages occurred early on, can persist for several billions years in our models and be visible until the end of the simulation -- as it is the same for NGC~104, which produces the gap visible in Fig.~2 at about 8~kpc from the cluster.
+We will discuss how these gaps can survive for so long in Appendix~\ref{sec:gaps_vs_gcmass}.
+}}
 
@@
 
-    In total, we report the finding of 73 gaps in our 50 simulations which is on average 1.5 gaps/simulation and were created by 18 different perturbers.
-The distribution of number of gaps appearing per simulation is presented in Table~\ref{table:gap_distribution}.
-If we consider NGC~2808 to be an outlier and exclude it from the experiment, we observe an average of 0.6 gaps per simulation.
+    In total, we report the finding of 73 gaps in our 50 simulations, which is, on average, 1.5 gaps/simulation.
+Eighteen different perturbers provoke the gaps.
+Table~\ref{table:gap_distribution} presents the distribution of the number of gaps appearing per simulation.
+If we consider NGC~2808 an outlier and exclude it from the experiment, we observe an average of 0.6 gaps per simulation.
 
@@
       \caption{Occurrence of gaps in Palomar~5 streams,  in our simulations.
-More specifically, the table reports the number of simulations (second row) for which a given number of gaps (first row) is found.
+More specifically, the table reports the number of simulations (second row) for a given number of gaps (first row).
 }\label{table:gap_distribution}
@@
 
-    We need a more sophisticated statistic that can be compared to other simulations, so we turn to the \textit{gap creation rate} developed by \citet{2012ApJ...748...20C}.
-The gap creation rate is the number of gaps that appear per time and normalized by the length of the stream.
+    We need a more sophisticated statistic to compare our results to other simulations, so we turn to the \textit{gap creation rate} developed by \citet{2012ApJ...748...20C}.
+The gap creation rate is the number of gaps that appear per unit time and is normalized by the stream's length.
 For our simulations, this rate is given by: \begin{equation} \label{eq:gap_creation_rate} \mathcal{R}_{\textrm{Pal 5}} =  \frac{1}{T}\int_{0}^T l^{-1}(t) \sum_i \delta(t-t_i) dt,\end{equation}where $T$ is the total integration time, $l(t)$ is the length of the stream, and $\sum_i \delta(t-t_i)$ sums over the gap occurrences, with $i$ indexing over the number of gaps in a given simulation.
@@
     This computation allows us to analyze the distribution of gap creation rates across all simulations.
-Notice that since the gap creation rate \textit{adds in parallel}, naturally gaps that occur at earlier times when the stream was shorter are weighted higher than those that occur when the stream is longer.
-The results are shown in Fig.~\ref{fig:gapcreationrate}.
-This is roughly consistent with a simple estimate of the average gap creation rate: with 73 gaps over 5~Gyr of integration time for a stream about 20 kpc in length, the naive estimate is approximately 0.015~km~s$^{-1}$~kpc$^{-2} $ (which is roughly equivalent to $0.015~\rm{Gyr^{-1}kpc^{-1}}$).
-This is about double the weighted mean gap creation rate of 0.009~km~s$^{-1}$~kpc$^{-2}$.
-The naive estimate is higher because it does not account for the growth of the stream over time, unlike Eq.~\ref{eq:gap_creation_rate}.
+Notice that since the gap creation rate \textit{adds in parallel}, naturally, gaps that occur at earlier times when the stream was shorter are weighted higher than those that occur when the stream is longer.
+Fig.~\ref{fig:gapcreationrate} presents these results which are is roughly consistent with a simple estimate of the average gap creation rate: with 73 gaps over 5~Gyr of integration time for a stream about 20 kpc in length, the naive estimate is approximately 0.015~km~s$^{-1}$~kpc$^{-2} $ (which is roughly equivalent to $0.015~\rm{Gyr^{-1}kpc^{-1}}$).
+This naive estimate is about double the weighted mean gap creation rate of 0.009~km~s$^{-1}$~kpc$^{-2}$ and is higher because it does not account for the growth of the stream over time, unlike Eq.~\ref{eq:gap_creation_rate}.
 
@@
 
-    Lastly, we note that of the 73 observed gaps only 8 are in the trailing tail and the rest are in the leading tail.
-This is a surprising result.
-A priori, since the number of star-particles escape at similar rates from the L1 and L2 Lagrange points, each tail is of similar length and density.
+    Lastly, we note that of the 73 observed gaps, only eight are in the trailing tail, and the rest are in the leading tail, which is a surprising result.
+A priori, since the star particles escape at similar rates from the L$_1$ and L$_2$ Lagrange points, each tail is of similar length and density.
 The main difference between the two tails is that the leading tail is closer to the Galactic center than the trailing tail by about 400~pc.
-Since the lengths are equal, and the offset between the tails is small compared to the galactocentric distance of about 10 kpc, we expected the number of gaps in each tail to be more or less the same.
+Since the lengths are equal, and the offset between the tails is slight compared to the galactocentric distance of about 10 kpc, we expected the gaps in each tail to be more or less the same.
 We can compute the probability of observing the unequal occurrences through the binomial distribution.
-First, the since the 44 gaps linked to NGC~2808 are the same fly-by, they are not independent events.
-We remove them from this consideration.
-This leaves 21 gaps in the leading tail and 8 in the trailing.
+First, since the 44 gaps linked to NGC~2808 are the same fly-by, they are not independent events.
+We remove them from this consideration, which leaves 21 gaps in the leading tail and 8 in the trailing.
 The probability of observing up to 8 successes in 29 trials, given a 50\% chance of success, is 1.2\%--unlikely, but possible.
-Additionally, other perturbers impact at consistent times, which may also violate the assumption of independent events, as seen with NGC~2808.
+Additionally, other perturbers impact at consistent times, which may violate the assumption of independent events, as seen with NGC~2808.
 
@@
       \caption{The distribution and relationship between the impact variables from Eq.~\ref{eq:change_in_momentum} for all close fly-bys considered.
-The encounters that cause gaps are labeled with the same colors as from Fig.~\ref{fig:histogram_impact_time} in the scatter plot, with white rings edges for visibility, and binned in red for the histograms.
-The close encounters that did not lead to gaps are shown in black.
-Indeed, no obvious trend delimiting these planes into gap producing or not emerges.}
+The colors indicate the encounters that cause gaps and are the same as Fig.~\ref{fig:histogram_impact_time}, with white edges for visibility and binned in red for the histograms.
+The black histogram shows the close encounters that did not lead to gaps.
+Indeed, no obvious trend has emerged delimiting these planes into gap-producing or not.}
       \label{fig:impact_geometry_statistics}    
@@
 
-    With the perturbers identified, we perform statistics in the pursuit of understanding what conditions are necessary for a globular cluster to induce a gap on the Palomar~5 stream.
+    With the perturbers identified, we perform statistics to understand what conditions are necessary for a globular cluster to induce a gap on the Palomar~5 stream.
 We turn to impact theory, which in its simplest form is presented in works such as \citet{2008gady.book.....B}.
 Consider two particles: one stationary and the other moving past it.
-The distance between the two particles at their closest approach is known as the impact parameter.
-To simplify the analysis, the \textit{impulse approximation} is employed, which assumes that the velocity of the perturber remains unchanged during the interaction.
+The impact parameter is the distance between the two particles at their closest approach.
+The \textit{impulse approximation} is employed, which assumes that the velocity of the perturber remains unchanged during the interaction.
 This assumption simplifies the computation.
 
-    To understand how the impacted particle is perturbed, one needs to compute its change in momentum, which is determined by integrating the force acting on the particle over the duration of the interaction.
+    To understand how the impacted particle is perturbed, one needs to compute its change in momentum, which is determined by integrating the force acting on the particle throughout the interaction.
 A useful approximation for this change in momentum, per unit mass, is the force at the closest approach multiplied by an estimate of the interaction time:
@@
 
-    This equation asserts that a more massive perturber, passing closer to the particle, and moving more slowly, will have a greater impact.
+    This equation asserts that a more massive perturber, passing closer to the particle and moving more slowly, will have a greater impact.
 It is important to note that the momentum change is inversely proportional to the velocity of the perturber.
-Note that this contrasts with the intuition from elastic collisions such as those between billiard balls where higher velocities result in greater impacts.
-
-    \citet{2015MNRAS.450.1136E} extended this impact theory from one point mass impacting another to studying how an extended body impacts a stream.
-This quantifies the change in momentum of a given particle as a function of its distance from the point of greatest impact along the stream.
-In their analysis, the perturber was modeled as a Plummer sphere, like in our simulations.
-Since the stream is not zero-dimensional and has length, both the parallel and perpendicular components of the perturber velocity must be considered.
-Consequently, five parameters determine the change in velocity of a given stream particle: $M$, $r_p$, $b$, $W_\parallel$, and $W_\perp$, which are the: mass of the perturber, size of the perturber, impact parameter, parallel and perpendicular components of the relative velocity.
-As detailed in Appendix~\ref{sec:reconstruction}, we calculated these parameters for all our \texttt{full} simulations, by selecting -- for each of them --  the strongest 5 flybys of a perturber with the Palomar~5 stream.
-Thus, we compute a total of  250 impacts, and flag those that give way to gaps.
+Note that this contrasts with the intuition from elastic collisions, such as those between billiard balls, where higher velocities result in greater impacts.
+
+    \citet{2015MNRAS.450.1136E} extended this impact theory from one point mass impacting another to studying how an extended body impacts a stream by quantifying the change in momentum of a given particle as a function of its distance from the point of greatest impact along the stream.
+\citet{2015MNRAS.450.1136E} models their perturber as a Plummer sphere, like in our simulations.
+Since the stream has length, one needs to consider the parallel and perpendicular components of the velocity to describe the impact fully.
+Consequently, five parameters determine the change in velocity of a given stream particle: $M$, $r_p$, $b$, $W_\parallel$, and $W_\perp$, which are the mass of the perturber, size of the perturber, impact parameter, parallel and perpendicular components of the relative velocity.
+As detailed in Appendix~\ref{sec:reconstruction}, we calculated these parameters for all our \texttt{full} simulations by selecting -- for each of them --  the strongest five flybys of a perturber with the Palomar~5 stream.
+Thus, we compute 250 impacts and flag those that give way to gaps.
 
@@
 Therefore, we only present the quantities from Eq.~\ref{eq:change_in_momentum} in Fig.~\ref{fig:impact_geometry_statistics}.
-    Note that in this figure, we show the total relative velocity rather than separating parallel and perpendicular components, because from a visual inspection, we could not find any specific trends when the two velocity components were plotted separately.
-We also exclude the characteristic cluster radius, as it showed little correlation with the results, likely due to the narrow range of globular cluster radii (see Fig.~\ref{fig:mass_size_plane}).
+    Note that this figure shows the total relative velocity rather than separating parallel and perpendicular components because, from a visual inspection, we could not find any specific trends when we plotted the two velocity components separately.
+We also excluded the characteristic cluster radius, which showed little correlation with the results, likely due to the narrow range of globular cluster radii (see Fig.~\ref{fig:mass_size_plane}).
 This factor might be more significant for dark matter sub-halos, where size variation is greater.
 
-    While Fig.~\ref{fig:impact_geometry_statistics} demonstrates that mass, relative velocity, or impact parameter alone cannot predict gap formation, one interesting result emerges: no gaps are created from interactions with impact parameters greater than 300~pc.
+    While Fig.~\ref{fig:impact_geometry_statistics} demonstrates that mass, relative velocity, or impact parameter alone cannot predict gap formation, one interesting result emerges: impact parameters greater than 300~pc do not create gaps.
 The stream widths are roughly 200~pc, as seen in Appendix~\ref{sec:gallery_of_gaps}.
@@
 A series of perturbers at roughly $\sim8 \times 10^5 M_\odot$ highlights NGC~2808's flybys, where all encounters with impact parameters under 200~pc result in gaps, while those beyond this distance do not.
-In other words, even the most massive globulars with masses greater than about $10^6 M_\odot$ cannot cause gaps if their impact parameters are greater than about $300$~pc.
-It is interesting to note that for values of the impact parameter below this threshold, even fast encounters ($\delta v > 300$~km/s) can produce gaps.
-Perhaps, this is not surprising since the range of possible relative velocities is much less than mass and impact parameter which vary two and three orders, respectively while the relative velocities only vary by about a factor of three.
+In other words, even the most massive globular clusters with masses greater than about $10^6 M_\odot$ cannot cause gaps if their impact parameters are greater than roughly $300$~pc.
+Interestingly, even fast encounters ($\delta v > 300$~km/s) can produce gaps for impact parameter values below this threshold.
+Perhaps this is not surprising since the range of possible relative velocities is much less than mass and impact parameters, which vary by two and three orders of magnitude, respectively.
+In contrast, the relative velocities only vary by about a factor of three.
 
@@
 The data set includes the top 5 strongest flybys from each simulation.
-Those that cause gaps are colored, stacked, and overlain atop of those that do--the distribution of which is outlined in grey.
+Those that cause gaps are colored, stacked, and overlain atop those that do not--which is the gray distribution.
 Note that the meaning of the colors is the same as in Fig.~\ref{fig:histogram_impact_time}.}
@@
 \textit{Top Left:} Energy-angular momentum space of the globular clusters in the simulations, with 50~$\times$~165 data points representing all sampled initial conditions.
-Clusters impacting Palomar~5 are shown with colored markers; large for the samplings that induce a gap and small for those that do not.
-Non-gap-causing clusters are shown in small gray dots.
+Clusters impacting Palomar~5 are shown with colored markers, large for the samplings that induce a gap and small for those that do not.
+The small gray dots represent the non-gap-causing clusters.
 The 50 white dots indicate Palomar~5's sampled initial conditions for the current day position.
-The circular velocity-curve is drawn with a light-pink dashed curve.
+The light-pink dashed curve shows the circular velocity curve.
 \textit{Top Right:} Mass-size plane of the globular clusters in the simulations, with uncertainties on the masses indicated as vertical lines.
-We remind the reader that the globular cluster catalog at this time does not provide uncertainties for clusters' characteristic radii.
+We remind the reader that the globular cluster catalog currently does not provide uncertainties for clusters' characteristic radii.
 \textit{Bottom Left:} Palomar~5's orbit in the Galactocentric xy plane.
-The gray-scale represents all 50 stacked orbits, with $\mathcal{P}$ indicating the probability of Palomar~5's position, normalized to $\mathcal{P}_\textrm{max}=1$.
-The colored markers indicate where the perturber was located when it impacted the stream and \textit{not} its present day position.
-\textit{Bottom Right:} same as bottom left but in the meridional plane.
-In all panels the colors of the markers and histogram bars correspond to specific perturber as specified in Fig.~\ref{fig:histogram_impact_time}.}
+The gray scale represents all 50 stacked orbits, with $\mathcal{P}$ indicating the probability of Palomar~5's position, normalized to $\mathcal{P}_\textrm{max}=1$.
+The colored markers indicate the position of the perturber when it impacted the stream and \textit{not} its present-day position.
+\textit{Bottom Right:} same as the bottom left but in the meridional plane.
+In all panels, the colors of the markers and histogram bars correspond to specific perturbers as specified in Fig.~\ref{fig:histogram_impact_time}.}
       \label{fig:mass_size_plane}
@@
 
-    Once all the key impact parameters estimated, we can use Eq.~\ref{eq:change_in_momentum} and calculate $\Delta p$, the change in momentum (per unit mass) imparted by a cluster flyby on Palomar~5 stream.
-The result  is shown in Fig.~\ref{fig:deltap}, where the distribution of imparted change in momentum is shown for all impacts which produce a gap, and compared to those that do not produce one.
-On average, encounters which lead to gap creations impart a change in momentum on stream particles which is a factor of 10 higher than that of encounters which do not form gaps (but with some overlap in the low-velocity tail).
-Interestingly, changes in momentum which lead to gap creations extend over a large range in velocities.
-There is a factor of about 100 between the smallest and largest changes with NGC~2808 (orange color in the histogram) imparting changes in the velocity of stream particles which redistribute over the whole range of $\Delta p$.
+    Once all the key impact parameters are estimated, we can use Eq.~\ref{eq:change_in_momentum} and calculate $\Delta p$, the change in momentum (per unit mass) imparted by a cluster flyby on Palomar~5 stream.
+Fig.~\ref{fig:deltap} shows the distribution of imparted change in momentum for all impacts that produce a gap compared to those that do not.
+On average, encounters that lead to gap creation impart a change in momentum on stream particles, which is a factor of 10 higher than encounters that do not form gaps (but with some overlap in the low-velocity tail).
+Interestingly, changes in momentum, which lead to gap creations, extend over a large range in velocities.
+There is a factor of about 100 between the smallest and largest changes with NGC~2808 (orange color in the histogram) imparting changes in the velocity of stream particles, which redistribute over the whole range of $\Delta p$.
 
     In addition to characterizing the parameters governing cluster encounters with the stream, since we know which clusters have produced gaps on the tail of Palomar~5, we can also verify their orbital and structural properties.
-This is shown in Fig~\ref{fig:mass_size_plane}, where we first show the mass and size (i.e.
-half-mass radius) of the clusters that cause gaps on the tails of Palomar~5, dividing them from those that do not.
-As can be seen, no cluster with mass below $10^5 M_\odot$ causes gaps on the Palomar~5 stream, and all perturbers, except FSR~1758 have a half-mass radius below 10~pc.
-Even more interesting is their distribution in the E-L$_z$ plane, which shows that the clusters that cause gaps are on both direct and retrograde orbits (respectively, negative and positive values of $L_z$), but all confined to an energy interval between $-2$ and $-1 \times10^5~\textrm{km}^2\textrm{s}^{-2}$.
-This is because only clusters within Palomar~5's orbital space can interact with the stream: clusters with higher orbital energies tend to have larger apocenter than that of Palomar~5 and thus spend most of their time away from Palomar~5's orbital volume.
-
-    Finally, it is worth commenting on the location of impacts with the stream, whether they occur when Palomar~5 is close to its pericenter or not.
-As it can be seen from Fig.~\ref{fig:mass_size_plane} (bottom panel) encounters can occur at all orbital phases of Palomar~5, when it is close to pericenter, but also very far away from it, at the outskirts of its orbital space.
-However, when taken all together, the location of gap creating impacts shows a strong negative correlation with the galactocentric radius $r$, with the number $N$ of encounters favorable to gap formation  going as $N = -2.5r + 50$ (with a Pearson coefficient of -0.86).
+Fig~\ref{fig:mass_size_plane} shows just this, where we first show the clusters' mass and size (i.e., half-mass radius) that cause gaps on the tails of Palomar~5, dividing them from those that do not.
+As can be seen, no cluster with mass below $10^5 M_\odot$ causes gaps on the Palomar~5 stream, and all perturbers, except FSR~1758, have a half-mass radius below 10~pc.
+Even more interesting is their distribution in the E-L$_z$ plane, which shows that the clusters that cause gaps are on both direct and retrograde orbits (negative and positive values of $L_z$).
+However, all of the perturbers exist in an energy interval between $-2$ and $-1 \times10^5~\textrm{km}^2\textrm{s}^{-2}$, which is because only clusters within Palomar~5's orbital space can interact with the stream: clusters with higher orbital energies tend to have larger apocenters than that of Palomar~5 and thus spend most of their time away from Palomar~5's orbital volume.
+
+    Finally, it is worth commenting on the location of impacts with the stream, whether they occur when Palomar~5 is close to its pericenter.
+Fig.~\ref{fig:mass_size_plane}'s bottom panel shows that encounters can occur at all orbital phases of Palomar~5, when it is close to the pericenter, but also very far away from it, at the outskirts of its orbital space.
+However, when taken all together, the location of gap-creating impacts shows a strong negative correlation with the galactocentric radius $r$, with the number $N$ of encounters favorable to gap formation going as $N = -2.5r + 50$ (with a Pearson coefficient of -0.86).
 While there are more clusters near the Galactic center, clusters naturally spend more time near their apocenters, and the lower relative velocities in these regions should favor gap creation.
-However, this result suggests that cluster Galactic number density outweighs these factors in determining the number of gaps.
-It should be noted that these results are for Palomar~5 only.
-Streams along various orbits need to be studied in order for this conclusion to be generalized.
-
-    %Our results are consistent with the findings from \citet{2021MNRAS.505.5978V}, who identified a wide range of orbital solutions for the least bound clusters, including both retrograde and prograde orbits.
-In our potential model, only three clusters samplings had positive orbital energies.
+However, this result suggests that the cluster Galactic number density outweighs these factors when determining the number of gaps.
+The results are for Palomar~5 only, and a future study would need to investigate streams along various orbits before generalizing this conclusion.
+
+    %Our results are consistent with the findings from \citet{2021MNRAS.505.5978V}, who identified a wide range of orbital solutions for the least bound clusters, including retrograde and prograde orbits.
+In our potential model, only three cluster samplings had positive orbital energies.
 
@@
 In the bottom panel of Fig.~\ref{fig:decomposition}, we compare the tracks of Palomar~5 that were compiled in \texttt{galstreams} by \citet{2023MNRAS.520.5225M}.
-The mid-point positions of the streams do not have the same galacto-centric positions as Palomar~5 from Baumgardt's catalog and are thus offset when projected into tail coordinates.
+The mid-point positions of the streams do not have the same galactocentric positions as Palomar~5 from Baumgardt's catalog, and the difference creates an offset when projected into tail coordinates.
 Moreover, since we sample the distances to Palomar~5, the galactocentric position within the Baumgardt catalog varies.
-To combat this, we position the mid-point of the Galstream tracks at the center of mass of the cluster, which allows us to compare the length of the observed tracks to our simulated streams.
+To combat this, we position the mid-point of the \texttt{Galstream} tracks at the cluster's center of mass, allowing us to compare the length of the observed tracks to our simulated streams.
 We use the three tracks Pal5-PW19, Pal5-S20, and Pal5-I21 from \citet{2019AJ....158..223P}, \citet{2020MNRAS.495.2222S}, and \citet{2021ApJ...914..123I}, respectively.
-For each track, we find the distance from the cluster in either direction and count the number of gaps that fall within this range.
-This is presented in Fig.~\ref{fig:GapsWithinSight}.
-In the maximum limit many gaps could appear, at a rate of about one per realization.
-However, at the shortest reported stream-length only a few gaps occur.
+For each track, we find the distance from the cluster in either direction, count the number of gaps within this range, and present this in Fig.~\ref{fig:GapsWithinSight}.
+In the maximum limit, many gaps could appear at a rate of about one per realization.
+However, only a few gaps occur at the shortest reported stream length.
+\textbf{\paola{We note, in particular, that the gap generated by the recent perturbation induced on the stream by the passage of NGC~2808 (whose occurrence is very likely according to our models) sits in a portion of the simulated tail which is at the edge of the observed one (see bottom panel in Fig.~2).
+This element and the fact that this gap is skinny, because it is very recent, probably make its detection difficult.
+}}
       \begin{figure}
@@
         \caption{The distribution of gaps as a function of absolute distance from the center of mass of Palomar 5's globular cluster.
-The number of gaps within the observable range is evaluated on an individual track basis, with track lengths determined by the minimum and maximum $x'$-coordinates from the bottom panel of Fig.~\ref{fig:decomposition}.
+We count the number of gaps within the observable range on an individual track basis, with track lengths determined by the minimum and maximum $x'$-coordinates from the bottom panel of Fig.~\ref{fig:decomposition}.
 The ranges for \citet{2019AJ....158..223P}, \citet{2020MNRAS.495.2222S}, and \citet{2021ApJ...914..123I} are shown for their respective dimensions, and the number of gaps within these ranges is indicated in the figure.
-Vertical bars mark the $x' > 0$ limit of the stream--we only show the right hand side instead of both for clarity, as most gaps are located on the right-hand side of the stream in any case.
+Vertical bars mark the $x' > 0$ limit of the stream--we only show the right-hand side instead of both for clarity, as most gaps are located on the right-hand side of the stream in any case.
 }
@@
 
-    We observe that there are no gaps near within 3~kpc to the cluster in our simulation.
+    We observe no gaps within 3~kpc of the cluster in our simulation.
 There may be a few reasons for the absence of gaps in the portion of the tails closer to the cluster center.
@@
 For instance, the frequency corresponding to the azimuthal action, $J_\phi$, is $\dot\theta_\phi = - \frac{\mathcal{\partial H}}{\partial J_\phi}$ and where $\theta_\phi$ is the angle describing the position of the particle phase space between momentum ($p_\phi$) and position ($\phi$) axes---where $\phi$ is the azimuthal angle between the x-y axes in physical space.
-Regions of the stream with well-separated frequencies are more susceptible to gap formation, while those with a wide frequency range (near the cluster) tend to erase the history of impacts.
+Stream regions with well-separated frequencies are more susceptible to gap formation, while those with a wide frequency range (near the cluster) tend to erase the history of impacts.
 Thus, for a gap to form, the imparted change in frequency must exceed the range of frequencies in the impacted region.
-Thus, the strong flybys that occurred close to the cluster were inconsequential for gap formation.
-
-    Another possible explanation is the different thickness of the simulated tails at different distances from the center of the cluster.
-At distances of less than about 3~kpc from Palomar~5, the tails are very thin, with a typical thickness being less than 100~pc.
-In order to cause gaps in these regions, the clusters would have to pass close to the stream, with impact parameters similar to the thickness itself.
+Therefore, the strong flybys close to the cluster were inconsequential for gap formation.
+
+    Another possible explanation is the different thicknesses of the simulated tails at different distances from the cluster's center.
+At distances less than 3~kpc from Palomar~5, the tails are skinny, with a typical thickness of less than 100~pc.
+To cause gaps in these regions, the clusters would have to pass close to the stream, with impact parameters similar to the thickness itself.
 Note that the thinness of the tails at these distances is probably a direct consequence of the initial parameters we chose for the simulation.
-We have assumed for Palomar~5, 5~Gyr years ago, the same internal parameters (mass and size) that the cluster has today.
-With such parameters, after 5~Gyr of evolution, our system has lost most of its mass, and therefore the part of the tails closest to the cluster itself, whose density depends mainly on the most recent mass loss \citep[see, for example, Fig.~A.3 in][]{2012A&A...546L...7M}, is necessarily very thin because the simulated cluster has essentially no more mass to lose.
-This last point has also a consequence in the gap creation rate, which we derived at the end of Sect.~\ref{sect:geometry}: with only 70\% of the tail  ($\sim$ 14~kpc over 20~kpc, excluding the innermost $\pm 3$~kpc from the cluster center) really suitable for forming gaps, the gap creation rates are about 50$\%$ higher than those estimated in the aforementioned section, where values have been derived taking into account the full tail extent.
-This gap creation rate would be still not high enough to reproduce the  number of gaps in Palomar~5's tails, as reported by \citet{2012ApJ...760...75C}.
-In their study, they indeed suggested the presence of  five gaps with a 99~\% detection confidence, leading to a gap creation rate of 0.17~Gyr$^{-1}$~kpc$^{-1}$.
+For Palomar~5, 5~Gyr years ago, we assumed the same internal parameters (mass and size) as the cluster today.
+With such parameters, after 5~Gyr of evolution, our system has lost most of its mass, and therefore, the part of the tails closest to the cluster itself, whose density depends mainly on the most recent mass loss \citep[see, for example, Fig.~A.3 in][]{2012A&A...546L...7M}, is necessarily very thin because the simulated cluster has essentially no more mass to lose.
+This last point also has a consequence in the gap creation rate, which we derived at the end of Sect.~\ref{sect:geometry}: with only 70\% of the tail  ($\sim$ 14~kpc over 20~kpc, excluding the innermost $\pm 3$~kpc from the cluster center) suitable for forming gaps, the gap creation rates are about 50$\%$ higher than those estimated in the previous section, where values have been derived taking into account the full tail extent.
+This gap creation rate would still not be high enough to reproduce the number of gaps in Palomar~5's tails, as reported by \citet{2012ApJ...760...75C}.
+Their study suggested the presence of five gaps with a 99~\% detection confidence, leading to a gap creation rate of 0.17~Gyr$^{-1}$~kpc$^{-1}$.
 If correct, this rate would be too high to be explained by Galactic globular clusters only.
 
+\textbf{\paola{An additional explanation for the absence of gaps in the inner regions of a stream is given in App.xx.
+If a gap forms in a stream region where stars from the cluster coexist at different times, all stars lost at times before the impact acquire momentum and acceleration due to the perturber.
+Stars lost at different times move, however, in a given region of the stream with different velocities, and the same is true for the respective gaps, which will therefore lose spatial coherence over a time that is proportional to the size of the gap and inversely proportional to the velocity dispersion in that section of the stream.
+Since the inner regions of streams are those where stars lost at different times coexist more easily, this element must also be taken into account when discussing the visibility (or persistence) of gaps, and this independently of the nature of the perturber.}}
+
 \section{Discussion}
 
-  The simulations presented in this paper suggest that, in the last 5~Gyr of evolution, Palomar~5's stream could have experienced multiple close encounters with other Galactic globular clusters, some of which are able to create gaps -- even a few kpc wide -- in its tails.
-Currently, the presence -- or not -- of gaps in the observed portion of Palomar~5 tails is debated: \citet{2016ApJ...819....1I} found no statistically significant gap in Palomar~5 tails, while \citet{2017MNRAS.470...60E}, analyzing the same data-set as \citet{2016ApJ...819....1I}, suggested the presence of a few dips and gaps in the tails, at angular distances between $2^\circ$ and $9^\circ$ from the cluster center \citep[see also][]{ 2020ApJ...889...70B}.
-While our simulations produced 73 gaps across 50 realizations, only about 22 are beyond the current length of the observed portion of the stream.
-This means on average, we obtain at least one gap from a globular cluster within the past 5~Gyr.
+  The simulations presented in this paper suggest that, in the last 5~Gyr of evolution, Palomar~5's stream could have experienced multiple close encounters with other Galactic globular clusters, some of which can create gaps -- even a few kpc wide -- in its tails.
+Currently, the literature debates whether or not a gap exists in the observed portion of Palomar~5 tails.
+\citet{2016ApJ...819....1I} found no statistically significant gap in Palomar~5 tails, while \citet{2017MNRAS.470...60E}, analyzing the same data-set as \citet{2016ApJ...819....1I}, suggested the presence of a few dips and gaps in the tails, at angular distances between $2^\circ$ and $9^\circ$ from the cluster center \citep[see also][]{ 2020ApJ...889...70B}.
+While our simulations produced 73 gaps across 50 realizations, only about 22 are beyond the current length of the observed portion of the stream, which means, on average, we obtain at least one gap from a globular cluster within the past 5~Gyr.
 However, we did not attempt to create mock observations or simulate a full detection process accounting for the challenges of disentangling field stars from stream stars.
@@
 
-  The aforementioned factors could lead one to the conclusion that the number of gaps identified in this study could represent an upper limit.
+  The factors mentioned above could lead one to conclude that the number of gaps identified in this study could represent an upper limit.
 However, this conclusion is incomplete.
@@
 Moreover, the present-day masses used in our simulations are likely lower than the historical masses of these clusters.
-For example, \citet{2024ApJ...976...54P} conducted a study simulating the dissolution of a realistic globular cluster population, in the aims of identifying how many stellar streams we should expect in the Milky Way and used a mock catalog of globular clusters with an initial amount with masses above 10$^4 M_\odot$ totaling at about 10,000 clusters.
-This implies that more perturbers could have been present in the past, potentially increasing the frequency and number of gaps in Palomar~5's stream.
+For example, \citet{2024ApJ...976...54P} conducted a study simulating the dissolution of a realistic globular cluster population to identify how many stellar streams we should expect in the Milky Way and used a mock catalog of globular clusters with masses above 10$^4 M_\odot$ totaling about 10,000 clusters.
+Their setup implies that more perturbers could have been present in the past, potentially increasing the frequency and number of gaps in Palomar~5's stream.
 
@@
 Specifically, \citet{2019MNRAS.484.2009B} focused on examining power spectra, analyzing variations in the one-dimensional stream density in stellar counts along the length of the stream.
-
-  Upon comparison with our simulations, we observe that the power spectra from our \texttt{full} simulations that include globular clusters and the \texttt{reference} simulations that do not, are not significantly different.
+Upon comparison with our simulations, we observe that the power spectra from our \texttt{full} simulations that include globular clusters and the \texttt{reference} simulations that do not significantly differ.
 We suggest that this may be due to the signal from only one or two gaps caused by globular clusters not being sufficient to produce notable differences in the overall power spectrum.
-Additionally, \citet{2019MNRAS.484.2009B} may not have inspected the profiles for individual gaps, which could explain why these features were not reported in their study.
-Further investigations would be needed to confirm this interpretation and to thoroughly assess the impact of globular clusters on stream density profiles.
+Additionally, \citet{2019MNRAS.484.2009B} may not have inspected the profiles for individual gaps, which may be why they did not report them.
+Further investigations would be needed to confirm this interpretation and thoroughly assess the impact of globular clusters on stream density profiles.
 \\
 
-  The impact of globular clusters close interactions with stellar streams has also been the object of another recent paper, by \citet{2022ApJ...941..129D}, who concluded that the chance for GD-1 gaps, which have been reported in a number of works \citep[see, for example, ][]{2019ApJ...880...38B,2018MNRAS.477.1893D,2020AAS...23533607D} to be produced by globular clusters is very low.
-This result is not necessarily in contradiction with ours, since GD-1 has a pericenter which is almost twice that of Palomar~5 \citep[see, for example][]{2019MNRAS.486.2995M}.
-As we have discussed in Sect.~\ref{sect:geometry}, the number, $N$, of close encounters which lead to gap creation is anti-correlated with the distance $r$ to the Galactic center.
-If we naively use the same radial dependence of gaps from Palomar~5 for GD-1 by swapping a pericenter from 6~kpc to 16~kpc, we would reduce the number of gap-favorable impacts of more than a factor of 2.
-Moreover, we note that \citet{2022ApJ...941..129D} pre-select the globular clusters which could have experienced a close encounter with the GD-1 stream, by selecting only clusters which pass at a distance less than 0.5 kpc from the stream, having a relative velocity smaller than  300 km/s.
-As shown in our Fig.~\ref{fig:impact_geometry_statistics}, bottom-left panel, in the case of Palomar~5 this choice would lead to exclude most of the encounters favorable to gap creation, which turn out to have relative velocities above 300 km/s.
-It would be interesting to repeat a similar study as the one made by \citet{2022ApJ...941..129D} for GD-1, this time imposing no selection on possible candidate clusters.\\
-
-  As already suggested in previous works which have studied the impact of baryonic structures on Palomar~5 tails \citep{2017NatAs...1..633P, 2019MNRAS.484.2009B}, it is possible that this cluster lies in a region of the phase-space which is not favorable to distinguish gaps created by dark matter sub-halos from gaps created by baryonic structures, such as the Galactic bar and giant molecular clouds.
-With our study, we show that close encounters with globular clusters constitute a further element that confuses a simple interpretation of the observed Palomar~5 stream gaps (if any).
-Other clusters and streams in the orbital energy range of Palomar~5 may suffer of the same difficulty.
-Going to lower orbital energies make the situation even worse, because the effect of the Galactic bar, giant molecular clouds and interactions with globular clusters becomes even more efficient.
-In the innermost regions of the Galaxy, where the density of dark matter sub-halos is expected to be maximal, globular cluster streams are intrinsically more difficult to find, in addition to the fact that the dynamical times in these regions become so small that the stars lost from globular clusters do not redistribute themselves for the most part into thin structures \citep[see][]{2023A&A...673A..44F}.
-It is thus probably only at larger distances from the Galactic center than those spanned by Palomar~5's orbit that the impact of dark matter sub-halos may become dominant, but it should be borne in mind that at these distances the number density of sub-halos also decreases.
-In the future, it will be interesting to apply the type of study conducted here to the whole set of clusters for which streams have been found to quantify the regime in which they are found, whether favorable to the creation of gaps from baryonic structures or not.
+  \textbf{ \paola{\citet{2017MNRAS.470...60E} is another world that discusses the impact of globular clusters in producing gaps on the Pal~5 stream.
+In section~6.3 of their article, \citet{2017MNRAS.470...60E} discuss the fact that globular clusters with masses greater than $10^6 M_\odot$ are rare in the Galaxy, especially at distances compatible with the orbit of Pal~5, while -- based on previous works -- they estimate that sub-halos of dark matter of similar mass are at least three times greater in number.
+This difference leads them to conclude that the gaps they report in the Pal~5 streams are more likely to be induced by dark matter sub-halos than by globular clusters.
+  However, as we show in this paper, even clusters with masses below $10^6 M_\odot$ can produce gaps.
+The clusters perturbing the Pal~5 stream are clusters with masses up to 10 times lower than those considered by \citet{2017MNRAS.470...60E}.
+It is thus possible that \citet{2017MNRAS.470...60E} have underestimated the impact of globular clusters' close passages on a stream such as Pal~5.
+It is more difficult, however, to make a comparative analysis between the role of clusters and sub-halos at equal stellar mass.
+Being less dense than globular clusters, sub-halos should produce less intense perturbations on streams (for fixed impact parameters and relative velocities).
+A systematic comparative analysis to quantify the role of clusters and sub-halos in producing gaps and perturbations in stellar streams still needs to be done.}}
+
+  The impact of globular clusters' close interactions with stellar streams has also been the object of another recent paper by \citet{2022ApJ...941..129D}, who concluded that the chance for GD-1 gaps, which have been reported in several works \citep[see, for example, ][]{2019ApJ...880...38B,2018MNRAS.477.1893D,2020AAS...23533607D} to be produced by globular clusters is very low.
+This result is not necessarily in contradiction with ours since GD-1 has a pericenter which is almost twice that of Palomar~5 \citep[see, for example][]{2019MNRAS.486.2995M}.
+As discussed in Sect.~\ref{sect:geometry}, the number, $N$, of close encounters that lead to gap creation is anti-correlated with the distance $r$ to the Galactic center.
+If we naively use the same radial dependence of gaps from Palomar~5 for GD-1 by swapping a pericenter from 6~kpc to 16~kpc, we would reduce the number of gap-favorable impacts by more than a factor of 2.
+Moreover, we note that \citet{2022ApJ...941..129D} pre-selects the globular clusters that could have experienced a close encounter with the GD-1 stream by selecting only clusters that pass at a distance less than 0.5 kpc from the stream, having a relative velocity smaller than  300 km/s.
+As shown in our Fig.~\ref{fig:impact_geometry_statistics}, bottom-left panel, in the case of Palomar~5, this choice would lead to excluding most of the encounters favorable to gap creation, which turn out to have relative velocities above 300 km/s.
+It would be interesting to repeat a similar study as the one made by \citet{2022ApJ...941..129D} for GD-1, imposing no selection on possible candidate clusters.\\
+
+  As already suggested in previous works which have studied the impact of baryonic structures on Palomar~5 tails \citep{2017NatAs...1..633P, 2019MNRAS.484.2009B}, this cluster may lie in a region of the phase-space which is not favorable to distinguish gaps created by dark matter sub-halos from gaps created by baryonic structures, such as the Galactic bar and giant molecular clouds.
+Our study shows that close encounters with globular clusters constitute a further element that confuses a simple interpretation of the observed Palomar~5 stream gaps (if any).
+Other clusters and streams in the orbital energy range of Palomar~5 may suffer from the same difficulty.
+Going to lower orbital energies worsens the situation because the effect of the Galactic bar, giant molecular clouds, and interactions with globular clusters becomes even more efficient.
+In the innermost regions of the Galaxy, where the density of dark matter sub-halos is maximal, globular cluster streams are intrinsically more difficult to find, in addition to the fact that the dynamical times in these regions become so small that the stars lost from globular clusters do not redistribute themselves for the most part into thin structures \citep[see][]{2023A&A...673A..44F}.
+It is thus probably only at larger distances from the Galactic center than those spanned by Palomar~5's orbit that the impact of dark matter sub-halos may become dominant, but bear in mind that at these distances, the number density of sub-halos also decreases.
+In the future, it will be interesting to apply the type of study conducted here to the whole set of streams to quantify the regime in which they are favorable to the creation of gaps from baryonic structures.
 
@@
   Our study demonstrates that globular cluster flybys can produce density gaps in the stellar streams of Palomar~5.
-The occurrence and characteristics of these gaps depend on the dynamical and structural properties of the stream, the mass of the perturber, and the impact parameter.
+The occurrence and characteristics of these gaps depend on the stream's dynamical and structural properties, the perturber's mass, and the impact parameter.
 While our simulations predict the formation of gaps in Palomar~5's tails, the predicted gaps do not align with the regions of the stream currently observed.
-Several factors contribute to the absence of simulated gaps in the observed portions of Palomar~5's tails, including the stream's varying thickness, the dispersion of action-frequencies near the cluster, and the initial conditions adopted in our simulations.
+Several factors contribute to the absence of simulated gaps in the observed portions of Palomar~5's tails, including the stream's varying thickness, the dispersion of action frequencies near the cluster, and the initial conditions adopted in our simulations.
 
@@
   Future work should involve a systematic comparison of gap creation rates across various Galactic potentials and extend the scope to a wider range of streams and globular clusters.
-This would help identify regimes where baryonic and dark matter-induced substructures can be disentangled, shedding light on the elusive influence of dark matter sub-halos on stellar streams.
-Additionally, we can continue to add realistic physics to the Galactic environment such as: the merger of the Sagittarius dwarf galaxy, the passages of the Large and Small Magellanic clouds, as well as a time evolving globular cluster population that compensates for the clusters' mass loss, evaporation, and similarly add members from the current incomplete census.
+Such a work could disentangle regimes where baryonic and dark matter-induced substructures can dominate gap creation, shedding light on the elusive influence of dark matter sub-halos on stellar streams.
+Additionally, we can continue to add realistic physics to the Galactic environment, such as the merger of the Sagittarius dwarf galaxy, the passages of the Large and Small Magellanic clouds, as well as a time-evolving globular cluster population that compensates for the clusters' mass loss, evaporation, and similarly add members from the current incomplete census.
 
@@
   The code used to produce the data in these simulations is publicly available at https://github.com/salvatore-ferrone/tstrippy.
-Examples on the codes installation and use are provided at https://tstrippy.readthedocs.io.
+Examples of the installation and use of the code are provided at https://tstrippy.readthedocs.io.
 The simulations will also be made publicly available at etidal-project.obspm.fr.
@@
   This work has made use of the computational resources available at the Paris Observatory, as well as those obtained through the DARI grant A1020410154 (PI: P.
-Di Matteo), and of the Astropy, Numpy and Matplotlib libraries (Astropy Collaboration 2013, 2018; Harris et al.
+Di Matteo) and of the Astropy, Numpy, and Matplotlib libraries (Astropy Collaboration 2013, 2018; Harris et al.
 2020; Hunter 2007).
-SF, PDM and MM would like to thank the Ecole Doctorale 127 for its financial support.
+SF, PDM, and MM thank the École Doctorale 127 for its financial support.
 MM thanks the CNR STM program, thanks to which the collaboration on this research topic began.
@@
     For each of the 50 simulations, we compare the final snapshots between the \texttt{reference} and \texttt{full} simulations using the tail coordinate system, as shown in Fig.~\ref{fig:TailCoordinates}.
-The \texttt{reference} simulations were only saved at the final time-stamps, thus from Eq.~\ref{eq:data_volume_estimate} where $N_{ts}$=1, leading to a data volume of about one hundred measly megabytes.
+We only saved the \texttt{reference} simulations at the final time-stamps, thus from Eq.~\ref{eq:data_volume_estimate} where $N_{ts}$=1, leading to a data volume of about one hundred measly megabytes.
 We inspect these differences by generating 2D density maps of the streams.
 Additionally, we marginalize over the $y'$-coordinate to produce 1D density profiles along the $x'$-axis.
-These comparisons are shown in the results section (see Fig.~\ref{fig:profiles}).
+We show these comparisons in the results section (see Fig.~\ref{fig:profiles}).
     \begin{figure}
@@
 The black dot represents the globular cluster, while the thick black line shows a segment of its orbit.
-A star particle is marked with a star symbol.
+The star symbol indicates a star particle.
 The red vector points from the nearest point on the cluster's orbit to the star particle.
-Its projection onto the galactocentric unit vector is shown in green and labeled $y'$.
-The red vector is also projected onto the vector perpendicular to both the orbital tangent and the galactocentric vector, which defines the local orbital plane and is labeled $z'$.
-The path length along the orbit from the cluster to the nearest point is denoted as $x'$, with $x'$ being positive if the particle is ahead of the cluster.}
+The galactocentric position vector is the $y'$-axis.
+The local orbital plane, the cross-product between the orbit's local position and velocity vector, defines the $z'$-axis.
+We project the red vector on each of these axes.
+We measure the x' coordinate as the path length along the orbit from the cluster to the nearest point.
+Positive $x'$, $y'$, $z'$ indicate that a star is ahead of the cluster, farther from the galactic center, and above the orbital plane, respectively.}
       \label{fig:TailCoordinates}
@@
       \caption{A comparison between the density maps and profiles of the \texttt{full} and \texttt{reference} simulations is presented.
-The locations where the \texttt{full} simulation is less dense than the \texttt{reference} simulation by 2-$\sigma$ is highlighted by blue vertical bars.
-Bins below the noise threshold are not considered when measuring the differences between the two 1D profiles.
-The 1D profiles have been smoothed with a median box-car filter.}
+The blue vertical bars highlight the locations where the \texttt{full} simulation is less dense than the \texttt{reference} simulation by 2-$\sigma$.
+We do not consider bins below the noise threshold when measuring the differences between the two 1D profiles.
+We smoothed the 1D profiles with a median box-car filter.}
       \label{fig:profiles}
@@
     To construct the 1D density profiles, we bin the data using the $\sqrt{N}$ rule, where $N$ is the number of data points ($N_p = 10^5$).
-After binning the 1D profiles, we apply a boxcar smoothing technique.
-This raw profile density curve is smoothed with the median box-car filter.
-At each bin, a selected number of adjacent data points from both sides are placed in a list.
-The bin is then replaced with the median.
-We choose to use 10 adjacent points per side, which correspond to a smoothing length of approximately 1~kpc.
+After binning the 1D profiles, we apply a median boxcar smoothing technique.
+At each bin, we select a number of adjacent data points from both sides, place them in a list, and replace the central value of the bin with the median.
+We use 10 adjacent points per side, corresponding to a smoothing length of approximately 1~kpc.
 This procedure reduces high-frequency noise and smooths the profiles.
-For instance, notice the absence of a high mass peak indicating for the center of mass in the bottom panel of Fig.~\ref{fig:profiles}.
-
-    With the smoothed 1D density profiles in hand, we search for regions where the \texttt{full} simulations are significantly under-dense compared to \texttt{reference} simulations, surpassing stochastic fluctuations.
-To accomplish this, we first impose a signal-to-noise ratio threshold, $\mathcal{SNR}$.
-The signal is defined as the log of the counts per bin from the \texttt{reference} 1D density profile, and errors are propagated assuming a Poisson distribution.
+For instance, notice the absence of a high mass peak indicating the center of mass in the bottom panel of Fig.~\ref{fig:profiles}.
+
+    With the smoothed 1D density profiles in hand, we search regions where the \texttt{full} simulations are significantly under-dense compared to \texttt{reference} simulations, surpassing stochastic fluctuations.
+We first impose a signal-to-noise ratio threshold, $\mathcal{SNR}$.
+The signal is the log of the counts per bin from the \texttt{reference} 1D density profile; we propagate errors assuming a Poisson distribution.
 We then compute a threshold for the number of counts in the \texttt{reference} simulation, $N_v$, using the transcendental equation:
@@
     where $\mathcal{R}_i$ is the log ratio, $N_{f,i}$ are the counts from the \texttt{full} simulation, and $N_{v,i}$ are the counts from the \texttt{reference} simulation for each bin $i$.
-We then analyze the distribution of $\mathcal{R}_i$.
-If the differences between the density profiles are primarily due to stochastic processes of similar magnitude, this distribution should resemble a Gaussian as expected from the central limit theorem.
-Thus, we flag all regions where the density is under-dense by more than two standard deviations, which should highlight regions whose under-density is unlikely to be the result of the sum of stochastic processes but rather the passage of another globular cluster.
+We then analyze the $\mathcal{R}_i$ distribution.
+If the differences between the density profiles are primarily due to stochastic processes of similar magnitude, this distribution should resemble a Gaussian, as expected from the central limit theorem.
+Thus, we flag all regions where the density is underdense by more than two standard deviations, which should highlight regions whose underdensity is unlikely to be the result of the sum of stochastic processes but rather the passage of another globular cluster.
 
     However, this method has its limitations, especially when detecting smaller gaps.
-As outlined by \citet{2015MNRAS.450.1136E}, a small gap is not indicative of a weak impact, but a recent one.
-This is because gap growth is a dispersion phenomenon.
+As outlined by \citet{2015MNRAS.450.1136E}, since gap growth is a dispersion phenomenon, a small gap is not indicative of a weak impact but a recent one.
 Additionally, since our streams have finite width, some gaps are oblique with respect to the stream axis.
@@
 Therefore, this quantitative analysis serves as an \textit{aid} to visual inspection rather than a complete substitute for it.
-This method helps particularly with large subtle gaps that the eye does not notice in the 2D maps.
-These profiles are included in the Appendix~\ref{sec:gallery_of_gaps}.
+This method helps with significant, subtle gaps the eye does not notice in the 2D maps.
+Appendix~\ref{sec:gallery_of_gaps} presents these profiles.
 
@@
 
-    A key question we seek to answer is: from the gaps present at the end of the simulation, who caused them?
+    We seek to answer a key question: From the gaps present at the end of the simulation, who caused them?
 To address this, we examine the evolution of stream density over time.
 Instead of using the x'-coordinate, we introduce $\tau$, which represents time rather than distance.
-Specifically, $\tau$ indicates how long it will take for a cluster to reach, or how long ago it passed, a given point on its orbit.
-This is advantageous because the growth of the stream is approximately linear in $\tau$ whereas in physical space, streams on eccentric orbits expand and contract depending on the orbital phase.
-
-    \citet{2016MNRAS.457.3817S} extended the analysis of \citet{2015MNRAS.450.1136E}, demonstrating that action-angle variables provide a useful coordinate system for analyzing stream evolution, as actions are conserved and the angles associated with the stream's growth evolve linearly over time.
-Although we became aware of this work only after completing our analysis, we note that $\tau$ is a suitable approximation and behaves similarly to the angle variable corresponding to the azimuthal action: $\tau \approx \theta_{\phi,i} - \theta_{\phi,\text{GC}}$.
+Specifically, $\tau$ indicates how long a cluster will take to reach a given point in its orbit or how long ago it passed.
+This choice of coordinates is advantageous because the growth of the stream is approximately linear in $\tau$.
+In contrast, in physical space, streams on eccentric orbits expand and contract depending on the orbital phase.
+
+    \citet{2016MNRAS.457.3817S} extended the analysis of \citet{2015MNRAS.450.1136E}, demonstrating that action-angle variables provide a useful coordinate system for analyzing stream evolution, as actions are conserved quantities and their associated angles grow linearly over time.
+Although we became aware of this work only after completing our analysis, we note that $\tau$ is a suitable approximation and behaves similarly to the angle variable corresponding to the azimuthal action: $\tau \approx \theta_{\phi, i} - \theta_{\phi,\text{GC}}$.
 
@@
 The bottom panel shows the evolution of the stream density over time.
-To avoid extreme low-density regions at the stream's edges, we applied the same density threshold as from Eq.~\ref{eq:density_threshold} to focus on the more significant areas of the stream.
+To avoid extremely low-density regions at the stream's edges, we applied the same density threshold as from Eq.~\ref{eq:density_threshold} to focus on the more significant areas of the stream.
 Next, we modeled Palomar~5's orbit as a proxy for its stream and sampled points along the orbit to measure the gravitational force exerted by other globular clusters.
-This force data is displayed in the top panel of Fig.~\ref{fig:force-on-orbit}, showing how the total gravitational acceleration on Palomar~5's stream evolves over its length throughout the simulation.
+The top panel of Fig.~\ref{fig:force-on-orbit} shows how the total gravitational acceleration on Palomar~5's stream evolves over its length throughout the simulation.
 
@@
       \caption{This figure demonstrates how we determined which globular clusters were responsible for the gaps.
-The y-axis is $\tau$, which is a coordinate in units of time that indicates how far ahead or behind it is from a globular cluster.
-The x-axis is the simulation time, $t$, where 0~s km~kpc$^{-1}$ indicates present time.
-The bottom plot showcases the evolution of the stream density in simulation time.
-The density was used to determine a suitable length of the stream at a given function.
-This length was then used to extract a piece of Palomar~5's orbit.
-This orbital segment is used to approximate the stream.
-Then, the gravitational force from all other clusters was computed on the orbit.
-This is shown in the top plot, where the gravitational force is in measured in acceleration and is given in integration units: km$^2$~kpc$^{-1}$~s$^{-2}$.
+The y-axis is $\tau$, a coordinate in units of time that indicates how far ahead or behind it is from a globular cluster.
+The x-axis is the simulation time, $t$, where 0~s km~kpc$^{-1}$ indicates the present time.
+The bottom plot showcases the evolution of the stream density over simulation time.
+We used the density to determine a suitable stream length.
+Subsequently, we used this length to extract a piece of Palomar~5's orbit.
+This orbital segment approximates the stream.
+Then, we compute the gravitational force from all other stream stand-in clusters, shown in the top plot.
+The gravitational force is measured in acceleration and is given in integration units: km$^2$~kpc$^{-1}$~s$^{-2}$.
 Moments of high acceleration indicate the passage of another cluster.
-The top 5 strongest passages are labeled with red circles as well as the name of the clusters.
+The red circles indicate the top 5 strongest passages, as well as the names of the clusters.
 The example shown in this plot is the same simulation as Fig.~\ref{fig:stream_on_sky}.}
@@
     We then used \texttt{scipy}'s \texttt{ndimage} \citep{2020NatMe..17..261V} package to identify the top five local maxima in the data space of gravitational acceleration $\vec{g}$ as a function of time $t$ and the stream coordinate $\tau$.
-This is done by first smoothing the image by taking a 5 point moving average kernel.
-Secondly, we use maximum filter to locate coordinates in the ($t,\tau$) data plane who are local maxima to at least 10-adjacent data points.
-These locations are then ordered and the top five strongest interactions are saved.
-Once these local maxima were identified, we iterated over the contributions of individual globular clusters to determine which cluster contributed the most to each peak in $\vec{g}$.
-Each significant peak was labeled with the corresponding globular cluster.
+First, we smooth the data space by taking a 5-point moving average kernel.
+Secondly, we use a maximum filter to locate coordinates in the ($t,\tau$) data plane that are local maxima to at least 10 adjacent data points.
+We order these locations and save the top five strongest interactions.
+Then, we iterated over the contributions of individual globular clusters to determine which cluster contributed the most to each peak in $\vec{g}$.
+We label each significant peak with the corresponding globular cluster.
 
     Afterward, we cross-referenced these peaks with the locations of the gaps identified by studying the density maps and profiles from Fig.~\ref{fig:profiles}.
-For large gaps resulting from strong interactions, we observed that after an impact, a low-density wake is left behind in the ($t,\tau$) plane, which can be seen corresponding to the impacts of NGC~104 and NGC~7808 in the bottom panel of Fig.~\ref{fig:force-on-orbit}.
+The impact leaves a low-density wake in the ($t,\tau$) plane for large gaps resulting from strong interactions.
+the bottom panel of Fig.~\ref{fig:force-on-orbit} shows the wakes corresponding to the impacts of NGC~104 and NGC~7808.
 
@@
 Notice the periodic ribbons of force in the $(t,\tau)$ plane.
-This is due to pericenter passages where Palomar~5 is getting closer to the center of mass of the globular cluster system.
-Additionally, for the impacts of NGC~104 and NGC~C7078, wakes can be observed in the density map.
+The ribbons are due to pericenter passages where Palomar~5 is getting closer to the center of mass of the globular cluster system.
+Additionally, for the impacts of NGC~104 and NGC~7078, wakes can be observed in the density map.
 Another important aspect is that the strongest peak in gravitational force does not necessarily create a gap.
 Notice how NGC~5272, which was labeled with 0 to indicate that it has the greatest local maxima, does not have a gap.
-The reason for this is manifold, for instance, the force needs to be modulated with the time since the change in momentum is the determining factor and not the peak magnitude of the force.
+The reason for this is manifold.
+For instance, the force needs to be modulated with time since the change in momentum is the determining factor and not the peak magnitude of the force.
 Additionally, there is an offset of about 200~pc between the stream and the orbit, as seen when viewing the stream in tail coordinates, so peaks upon the orbit are good proxies for the stream but are not definitive.
@@
 
-    The results of this analysis were compiled into a table.
-If a gap was attributed to a particular globular cluster, it was labeled as \texttt{TRUE}; otherwise, it was marked as \texttt{FALSE}.
-For a handful of simulations, to double check that the verdict made correctly when passing from suspect to culprit, we re-compute the simulations yet individually adding one globular cluster at a time.
-As a result, we are able to confirm that singular gaps arise from the suspected clusters, an example of which is shown in Fig.~\ref{fig:decomposition}.
+    We compile the results of this analysis into a table.
+Each row of the table corresponds to a gap; its corresponding suspects are the columns.
+The culprit is labeled as \texttt{TRUE}.
+For a handful of simulations, to double-check that we make the correct verdict, we recompute the simulations, yet individually adding one globular cluster at a time.
+As a result, we can confirm that singular gaps arise from the suspected clusters.
+Fig.~\ref{fig:decomposition} shows one such example.
 
@@
 In the following, we describe how we estimated these parameters during impacts in our simulations.
-    To achieve this, we begin by identifying the approximate moments of impact from the most significant clusters, as determined in the previous analysis in Sec.~\ref{sec:Perturber_Identification}.
-Then, we refine these estimates in order to to pinpoint both the exact location of the impact along the stream and the precise moment it occurred.
-To do so, we fit a third-order parametric polynomial to the stream, using the saved snapshots from our simulations:
+    To achieve this, we identify the impact of the most significant clusters, as determined in the previous analysis in Sec.~\ref{sec:Perturber_Identification}.
+Then, we refine these estimates to pinpoint the exact location of the impact along the stream and the precise moment it occurred.
+To do so, we fit a third-order parametric polynomial to the stream using the saved snapshots from our simulations:
       \begin{equation}
@@
 We found that a second-order polynomial was insufficient to capture the curvature along the full length of the stream, with divergence at the ends of the tails.
-A third-order polynomial was sufficient and desirable, as it is the lowest order that adequately captures the path of the stream over the entire length under consideration.
-
-      In this analysis, only one side of the stream is considered.
+A third-order polynomial was sufficient and desirable, as the lowest order adequately captures the stream's path over the entire length under consideration.
+
+      In this analysis, we only consider one side of the stream.
 For instance, if the impact candidate was in the leading tail, only the star particles with $\tau > 0$ are used to constrain the stream track.
@@
 
-      Since the simulation snapshots were saved at a temporal resolution of 1 Myr--—rather than at the integration time-step, which would have generated an excessive amount of data—--we interpolate between snapshots to more precisely estimate the impact geometry.
-This is achieved by fitting the stream at five time-steps surrounding the approximate impact time, a period of 5 Myr, which sufficiently covers the interaction time.
+      Since we saved the simulation snapshots at a temporal resolution of 1~Myr--rather than at the integration time-step, which would have generated excessive data, we must interpolate between snapshots to more precisely estimate the impact geometry.
+We fit the stream's shape with a parametric third-order polynomial at the five time steps surrounding the approximate impact time.
+This time is a period of 5~Myr, sufficiently covering the interaction time.
 The interaction time can be estimated as $t \approx \frac{100~\text{pc}}{300~\frac{\text{km}}{\text{s}}} \approx 0.3~\text{Myr}$.
 
-      As a result, each polynomial coefficient can now be expressed as a function of time.
+      Then, we use a cubic spline interpolation for the coefficients describing the stream's shape, which allows us to describe each polynomial coefficient as a function of time.
 Consequently, we can parameterize the stream as a function of both simulation time and position along the stream:
@@
         \end{equation}
-      where $\vec{s}(t, \tau)$ is the galactocentric position of a point on the stream, $\vec{p}(t)$ is the position of the perturber, and $b$ is the distance between the two.
+      Where $\vec{s}(t, \tau)$ is the galactocentric position of a point on the stream, $\vec{p}(t)$ is the position of the perturber, and $b$ is the distance between the two.
 The minimum value of $b$, denoted as $\text{min}(b)$, represents the impact parameter.
-The minimization is carried out using \texttt{scipy}'s optimization package with the \textit{L-BFGS-B} method, which allows us to place bounds on $t$ and $\tau$, ensuring no extrapolation occurs \citep{davidon1991variable}.
-
-      Once this minimization is performed, determining the relative velocity becomes straightforward.
-Since the minimization provides the impact parameter, time of impact, and corresponding value of $\tau$, we can compute the derivatives of the parametric equations at $t_{\text{min}}$ and $\tau_{\text{min}}$.
+We perform the minimization using \texttt{scipy}'s optimization package with the \textit{L-BFGS-B} method, which allows us to place bounds on $t$ and $\tau$, ensuring no extrapolation occurs \citep{davidon1991variable}.
+
+      Once this minimization is done, determining the relative velocity becomes straightforward.
+Since the minimization provides the impact parameter, time of impact, and the corresponding value of $\tau$, we can compute the derivatives of the parametric equations at $t_{\text{min}}$ and $\tau_{\text{min}}$.
 The parallel and perpendicular components of the perturber's velocity relative to the stream are given by:
@@
       where $\vec{v}_p$ and $\vec{v}_s$ are the velocities of the perturber and the stream, respectively.
-For each of the 50 simulations, this information was computed for the strongest 5 flybys of a perturber with the stream.
-Thus, we compute a sample of 250 impacts, and flag those that give rise to gaps.
+We compute the strongest five flybys for each of the 50 simulations.
+Thus, we created a sample of 250 impacts and flagged those that give rise to gaps.
 We refer the reader to Sect.~\ref{sect:geometry} for the presentation and discussion of the results.
+
+\section{Survival of gaps in streams emerging from more massive globular clusters}\label{sec:gaps_vs_gcmass}
+
+\textbf{\paola{In Section~\ref{sec:mass_loss}, we emphasized that the conditions chosen to simulate the formation and evolution of streams around Pal~5 are probably unrealistic.
+By choosing, as we did, the same values for the mass and size of the cluster 5 billion years ago as the cluster has today, we did not take into account the loss of mass and internal evolution of the cluster over time.
+In particular, the velocity dispersion of the stars in a cluster and the velocity dispersion of the stars leaving the cluster and entering the stream depend on the cluster's mass.
+It is therefore legitimate to ask whether, using a progenitor with a different mass and/or size, the gaps survive and whether, in particular, gaps formed several billion years ago can still be visible today.
+}}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=1\linewidth]{gap_no_more.png}
+    \caption{When the mass of Palomar~5 is increased, the gap is lost.}
+    \label{fig:gap_no_more.png}
+\end{figure*}
+
+\textbf{\paola{While we anticipate that a specific study on this aspect is in preparation (Ferrone et al, in prep), we would like to preview a few results here.
+Fig.~\ref{fig:gap_no_more} (top and bottom panels) shows, respectively, a stream generated in one of our simulations, using as initial mass (i.e., 5 billion years ago) for Pal~5 one similar to the current one or a larger mass ($xx M_\odot$).
+For simplicity, in both experiments, we left the cluster size unchanged.
+In both cases, a gap forms due to an encounter (specifically with the cluster NGCxxx) that occurred xx billion years ago.
+However, while in Fig.~\ref{fig:gap_no_more}, top panel, the gap persists for xx billion years, in Fig.
+xx (bottom panel), the gap is no longer observable at the end of the simulation.
+Looking more closely at the reasons for this difference, we see (Fig.\ref{fig:decoherence_shocked_outflow}) that the gap formed xx billion years ago in Fig.
+xx (bottom) survives in all the populations lost from the cluster before the impact and which had time to reach the part of the tail affected by the passage of NGCxxx.
+What happens, however, is that after a coherent gap has formed in the tail, and in particular in the populations of stars in this part of the tail that escaped from the cluster at different times, in each of these populations the respective gap tends to travel along the streams at different speeds, which leads to a phase shift of the gap and its loss of visibility.
+In other words, the gap can survive for billions of years in each population of stars that left the cluster at different times, but it is no longer.
+Still, initially coherent and therefore visible because each of these populations travels along the tail (and drags a gap with it) at different speeds.
+The loss of coherence occurs at a time $t_{visibility}$ which is directly proportional to the gap size and inversely proportional to the velocity dispersion in that part of the stream.
+So, if populations of stars lost at different times co-exist in the portion of the stream where the impact occurs, the gap will remain visible for a time that depends on the mass of the progenitor system at the time these stars escaped.
+From this, we can understand that various factors determine the visibility of a gap in a stream.
+Mass affects the velocity dispersion in a given region of the stream, but this velocity dispersion, in turn, depends on the history of mass loss of the cluster.
+Suppose stars dominate a given region of the stream, all lost simultaneously from the cluster.
+In that case, the resulting velocity dispersion is lower than if various populations coexist in the same region, which has implications for these structures' persistence.
+}}
+
+\begin{figure*}
+    \includegraphics[width=\linewidth]{decoherence_shocked_outflow.png}
+    \caption{Nice}
+    \label{fig:decoherence_shocked_outflow}
+\end{figure*}
 
@@
     Figures~\ref{fig:gallery0}-\ref{fig:gallery11} include all of the realizations of the \texttt{full} simulations of Palomar~5 in tail coordinates.
-If a gap is present, it is indicated by a red arrow with the responsible globular cluster.
+A red arrow aids the reader in identifying the gap, accompanied by the responsible globular cluster.
 For the gaps that are not easily visible, we show the 1D profiles showing the significant under-densities compared to the \texttt{reference} simulations.
 For instance, the second and last panels of Fig.~\ref{fig:gallery0}.
-However, maginalizing the 2D distribution to 1D can erase the signal of oblique gaps.
-Thus, not all gaps that are visible in the 2D space are flagged in the density profiles.
-For example, in \texttt{Sampling 000}, the gaps from NGC~2808 and NGC~5634 are not flagged present in the marginalized density profile.
+However, marginalizing the 2D distribution to 1D can erase the signal of oblique gaps.
+Thus, not all gaps in the 2D space are visible in the density profiles.
+For example, in \texttt{Sampling 000}, the gaps from NGC~2808 and NGC~5634 are not flagged as present in the marginalized density profile.
 
@@
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-000.png}
-      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-000.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-001.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-002.png}
-      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-002.png}
+      % \includegraphics[width=\linewidth]{tau-profile-monte-carlo-000.png}
+      % \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-001.png}
+      % \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-002.png}
+      % \includegraphics[width=\linewidth]{tau-profile-monte-carlo-002.png}
       \caption{First gap gallery. See the text in Appendix~\ref{sec:gallery_of_gaps} for a description.}
@@
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-003.png}      
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-004.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-005.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-006.png}
-      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-006.png}
+      % \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-004.png}
+      % \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-005.png}
+      % \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-006.png}
+      % \includegraphics[width=\linewidth]{tau-profile-monte-carlo-006.png}
       \caption{Second gap gallery}
@@
 
-    \begin{figure*}
-      \centering
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-007.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-008.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-009.png}      
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-010.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-011.png}
-      \caption{Third gap gallery}
-      \label{fig:gallery2}
-      \end{figure*}
-
-    \begin{figure*}
-      \centering
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-012.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-013.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-014.png}      
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-015.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-016.png}
-      \caption{Fourth gap gallery}
-      \label{fig:gallery3}
-      \end{figure*}
-
-    \begin{figure*}
-      \centering
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-017.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-018.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-019.png}      
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-020.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-021.png}
-      \caption{Fifth gap gallery}
-      \label{fig:gallery4}
-      \end{figure*}
-
-    \begin{figure*}
-      \centering
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-022.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-023.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-024.png}   
-      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-024.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-025.png}
-      \caption{Sixth gap gallery}
-      \label{fig:gallery5}
-      \end{figure*}
-
-    \begin{figure*}
-      \centering
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-026.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-027.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-028.png}
-      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-028.png}
-      \caption{Seventh gap gallery}
-      \label{fig:gallery6}
-      \end{figure*}
-
-    \begin{figure*}
-      \centering      
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-029.png}    
-      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-029.png}  
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-030.png}
-      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-030.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-031.png}
-      \caption{Eighth gap gallery}
-      \label{fig:gallery7}
-      \end{figure*}
-
-    \begin{figure*}
-      \centering
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-032.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-033.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-034.png}      
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-035.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-036.png}      
-      \caption{Ninth gap gallery}
-      \label{fig:gallery8}
-      \end{figure*}
-
-    \begin{figure*}
-      \centering
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-037.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-038.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-039.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-040.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-041.png}      
-      \caption{Tenth gap gallery}
-      \label{fig:gallery9}
-    \end{figure*}
-
-    \begin{figure*}
-      \centering
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-042.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-043.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-044.png}
-      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-044.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-045.png}
-      \caption{Eleventh gap gallery}
-      \label{fig:gallery010}
-    \end{figure*}
-
-    \begin{figure*}
-      \centering
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-046.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-047.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-048.png}
-      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-048.png}
-      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-049.png}
-      \caption{Twelfth gap gallery}
-      \label{fig:gallery11}
-    \end{figure*}
+    % \begin{figure*}
+    %   \centering
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-007.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-008.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-009.png}      
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-010.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-011.png}
+    %   \caption{Third gap gallery}
+    %   \label{fig:gallery2}
+    %   \end{figure*}
+
+    % \begin{figure*}
+    %   \centering
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-012.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-013.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-014.png}      
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-015.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-016.png}
+    %   \caption{Fourth gap gallery}
+    %   \label{fig:gallery3}
+    %   \end{figure*}
+
+    % \begin{figure*}
+    %   \centering
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-017.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-018.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-019.png}      
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-020.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-021.png}
+    %   \caption{Fifth gap gallery}
+    %   \label{fig:gallery4}
+    %   \end{figure*}
+
+    % \begin{figure*}
+    %   \centering
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-022.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-023.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-024.png}   
+    %   \includegraphics[width=\linewidth]{tau-profile-monte-carlo-024.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-025.png}
+    %   \caption{Sixth gap gallery}
+    %   \label{fig:gallery5}
+    %   \end{figure*}
+
+    % \begin{figure*}
+    %   \centering
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-026.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-027.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-028.png}
+    %   \includegraphics[width=\linewidth]{tau-profile-monte-carlo-028.png}
+    %   \caption{Seventh gap gallery}
+    %   \label{fig:gallery6}
+    %   \end{figure*}
+
+    % \begin{figure*}
+    %   \centering      
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-029.png}    
+    %   \includegraphics[width=\linewidth]{tau-profile-monte-carlo-029.png}  
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-030.png}
+    %   \includegraphics[width=\linewidth]{tau-profile-monte-carlo-030.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-031.png}
+    %   \caption{Eighth gap gallery}
+    %   \label{fig:gallery7}
+    %   \end{figure*}
+
+    % \begin{figure*}
+    %   \centering
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-032.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-033.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-034.png}      
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-035.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-036.png}      
+    %   \caption{Ninth gap gallery}
+    %   \label{fig:gallery8}
+    %   \end{figure*}
+
+    % \begin{figure*}
+    %   \centering
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-037.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-038.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-039.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-040.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-041.png}      
+    %   \caption{Tenth gap gallery}
+    %   \label{fig:gallery9}
+    % \end{figure*}
+
+    % \begin{figure*}
+    %   \centering
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-042.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-043.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-044.png}
+    %   \includegraphics[width=\linewidth]{tau-profile-monte-carlo-044.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-045.png}
+    %   \caption{Eleventh gap gallery}
+    %   \label{fig:gallery010}
+    % \end{figure*}
+
+    % \begin{figure*}
+    %   \centering
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-046.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-047.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-048.png}
+    %   \includegraphics[width=\linewidth]{tau-profile-monte-carlo-048.png}
+    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-049.png}
+    %   \caption{Twelfth gap gallery}
+    %   \label{fig:gallery11}
+    % \end{figure*}
 

### 2025-05-13 paper/main.tex
editing the figure to have the right dimensions
explaining experiment in the appendix

@@
 
-\textbf{\paola{In Section~\ref{sec:mass_loss}, we emphasized that the conditions chosen to simulate the formation and evolution of streams around Pal~5 are probably unrealistic.
+In Section~\ref{sec:mass_loss}, we emphasized that the conditions chosen to simulate the formation and evolution of streams around Pal~5 are probably unrealistic.
 By choosing, as we did, the same values for the mass and size of the cluster 5 billion years ago as the cluster has today, we did not take into account the loss of mass and internal evolution of the cluster over time.
@@
 It is therefore legitimate to ask whether, using a progenitor with a different mass and/or size, the gaps survive and whether, in particular, gaps formed several billion years ago can still be visible today.
-}}
 
@@
     \caption{When the mass of Palomar~5 is increased, the gap is lost.}
-    \label{fig:gap_no_more.png}
+    \label{fig:gap_no_more}
 \end{figure*}
 
-\textbf{\paola{While we anticipate that a specific study on this aspect is in preparation (Ferrone et al, in prep), we would like to preview a few results here.
-Fig.~\ref{fig:gap_no_more} (top and bottom panels) shows, respectively, a stream generated in one of our simulations, using as initial mass (i.e., 5 billion years ago) for Pal~5 one similar to the current one or a larger mass ($xx M_\odot$).
-For simplicity, in both experiments, we left the cluster size unchanged.
-In both cases, a gap forms due to an encounter (specifically with the cluster NGCxxx) that occurred xx billion years ago.
-However, while in Fig.~\ref{fig:gap_no_more}, top panel, the gap persists for xx billion years, in Fig.
-xx (bottom panel), the gap is no longer observable at the end of the simulation.
-Looking more closely at the reasons for this difference, we see (Fig.\ref{fig:decoherence_shocked_outflow}) that the gap formed xx billion years ago in Fig.
-xx (bottom) survives in all the populations lost from the cluster before the impact and which had time to reach the part of the tail affected by the passage of NGCxxx.
-What happens, however, is that after a coherent gap has formed in the tail, and in particular in the populations of stars in this part of the tail that escaped from the cluster at different times, in each of these populations the respective gap tends to travel along the streams at different speeds, which leads to a phase shift of the gap and its loss of visibility.
+While we anticipate that a specific study on this aspect is in preparation (Ferrone et al, in prep), we would like to preview a few results here.
+Fig.~\ref{fig:gap_no_more} shows two streams with the same orbital parameters that were impacted by NGC~7078.
+The top panel shows a stream that comes from a host mass of $10,000$~$M_\odot$, while in the second panel we increased the mass to $\sim40,000$~$M_\odot$.
+For simplicity, both clusters have the same half-mass radius of 27~pc.
+The encounter occurred about two and a half billion years ago by NGC~7078, which has a mass of XX.
+The gap persists for the lighter host, yet is barely visible in the case of the more massive host.
+
+\salvatore{Fig.~\ref{fig:decoherence_shocked_outflow} dissects the more massive stream of Fig.~\ref{fig:gap_no_more} and instructs us why the gap does not persist.
+The bottom panel shows the total density distribution, and dissects it into two profiles: stars that either are \textit{leaked} or \textit{shocked} out of the cluster.
+By shocked, we mean the stars that are lost during pericenter passages.
+Tidal forces generally scale with $r^{-3}$, thus significantly strengthening through their pericenter passages.
+See Fig.~3 of \citet{2004AJ....127.2753D} for an illustration of the magnitude of the tidal forces along an orbit.
+A consequence of the pericenter passages is that many more stars escape, fleeing with a higher mean velocity and velocity dispersion, which can be visualized in the \textit{Christmas tree} illustration in Fig.~A.3 of \citet{2012A&A...546L...7M}.
+We consider stars that leave between pericenter passages to be \textit{leaked} stars.
+We identify pericenter passages by computing each local minimum of the cluster's galactic-centric 3D radius with \texttt{scipy.signal.find\_peaks}.
+The duration of each shock is: $t_{\textrm{shock}}$ = 2$\sqrt{r/|\nabla\Phi|}$, where $r$ is the galactocentric radius and $|\nabla\Phi|$ is the magnitude of the acceleration from the galactic potential.
+Stars within this range about a given pericenter passage are all given the same color and shown as individual groups in the first, second, and third panels of Fig.~\ref{fig:decoherence_shocked_outflow}.}
+
+\salvatore{The situation can be simplified as one-dimensional collisionless dispersion.
+These packets of stars can be described as a population with the same initial position that drifts about a mean velocity, $\langle v \rangle$, with velocity dispersion, $\sigma_v$.
+The density evolution is a Gaussian whose amplitude decreases inversely with time, and whose standard deviation grows linearly with time:}
+\begin{equation}    
+        \rho(x,t) = \frac{1}{\sigma_v t \sqrt{2\pi} }\textrm{exp}\left(-\frac{1}{2}\left(\frac{x-\langle v \rangle t}{\sigma_v t}\right)^2\right).
+\end{equation}
+
+\salvatore{It is important to note that within a single packet, the stars self-segregate.
+The position of the density peak is at $\langle v \rangle t$, while the faster stars are ahead and the slower ones are behind.
+The packets are separated by a distance of $\langle  v\rangle T$, where $T$ is the orbital period.
+Regarding our simulations, since the mass and radius of the host are held constant, the mean velocity and velocity dispersion are the same across all packets.
+The density of each packet decreases since the easier-to-remove stars are stripped early on.
+}
+
+\salvatore{Consider an impact that happens at $x\prime$.
+The distance from the impact site to each packet is: $\delta x= x\prime - \langle v\rangle \left(t-n T\right)$, where $n$ is the packet index, with $n=0$ being the first packet escaped at the first pericenter passage in the simulation.
+Note that $\delta x\propto n$, and velocity segregation means $v\propto x$.
+Therefore, the different parts of the separate packets are hit at a given impact site.
+The earlier packets are hit in regions where the stars have lower velocities than the later packets.
+As a result, the gaps in the later packets quickly overtake those from the earlier packets.
+As a result, the gaps across the packets go out of phase and interfere destructively.}
+
+\salvatore{In the simplest consideration, of two gaps of the same dimension $\sigma_x$ and an offset of $\delta x$, the gaps are out of phase when $1<\sigma_x/\delta x$.
+In the temporal analysis of the stream presented in Fig.~\ref{fig:decoherence_shocked_outflow}, we did not observe at any point of the simulation that the gap initially persisted to then go out of phase.
+Drawing from the wisdom in \citet{2015MNRAS.450.1136E} \& \citet{2016MNRAS.457.3817S}, we see the contrast between the density peaks and the density minimum of a gap, as well as the width, grow in time, somewhere between $\propto \sqrt{t}$ and linear.
+In this simplified 1D collisionless dispersion analysis, the gaps between the different packets space apart linearly in time with $\delta x\propto \delta vt$, where $\delta v$ is the mean velocity of the about the impact site for two different packets.
+This inclines us to believe that the gaps space apart before they become significant.
+}
+
+\salvatore{Nonetheless, we believe this is the first consideration of decomposing the stream into packets of tidal bursts to understand the effect of the internal dynamics on gap persistence.
+Since gap persistence depends on the mass and internal evolution of the cluster, characterizing the gap can constrain our knowledge of the internal evolution.
+}
+% \paola{What happens is that the different and in particular in the populations of stars in this part of the tail that escaped from the cluster at different times, in each of these populations the respective gap tends to travel along the streams at different speeds, which leads to a phase shift of the gap and its loss of visibility.
 In other words, the gap can survive for billions of years in each population of stars that left the cluster at different times, but it is no longer.
@@
 Mass affects the velocity dispersion in a given region of the stream, but this velocity dispersion, in turn, depends on the history of mass loss of the cluster.
-Suppose stars dominate a given region of the stream, all lost simultaneously from the cluster.
+Suppose stars dominate a given stream region, and all are lost simultaneously from the cluster.
 In that case, the resulting velocity dispersion is lower than if various populations coexist in the same region, which has implications for these structures' persistence.
-}}
+}
 
@@
     \includegraphics[width=\linewidth]{decoherence_shocked_outflow.png}
-    \caption{Nice}
+    \caption{We dissect the density of the stream from the bottom panel of Fig.~\ref{fig:gap_no_more} into stars that are either shocked or leaked out, see text for details.
+The bottom panel shows the total 1D profile with the two different populations.
+The second from the bottom dissects the total shocked outflow into individual pericenter passages.
+Each of which is color-coated by the escape time.
+The top two panels separate the first four shocked outflow populations to aid visualization of the gap placement from the impact of NGC~7078.
+}
     \label{fig:decoherence_shocked_outflow}

### 2025-05-14 paper/main.tex
responded to all of Paola's comments

@@
   {In this work, we investigate the perturbations induced on streams by the passage of dense stellar systems, such as globular clusters themselves, to test the possibility that they may cause the formation of gaps as well.
-In particular, we focus on the study of the stream \textbf{\paola{generated by a cluster with characteristics (mass, size, and orbit) similar to the current characteristics}} of Palomar~5, a well-known globular cluster in the Galactic halo, which has particularly long tidal tails.
+In particular, we focus on the study of the stream generated by a cluster with characteristics (mass, size, and orbit) similar to the current characteristics of Palomar~5, a well-known globular cluster in the Galactic halo, which has particularly long tidal tails.
 }
  % methods heading (mandatory)
-  {For this purpose, we used a particle-test code to simulate \textbf{\paola{the formation and evolution of the stream}} when subjected to the Galaxy's gravitational field plus its whole system of globular clusters.}
+  {For this purpose, we used a particle-test code to simulate the formation and evolution of the stream when subjected to the Galaxy's gravitational field plus its whole system of globular clusters.}
  % results heading (mandatory)
-  {Our study shows that the \textbf{\paola{such a stream}} can be strongly perturbed by the close passage of other clusters, in particular of NGC~2808, NGC~7078, and NGC~104, and that these perturbations induce the formation of gaps in the tails.}
+  {Our study shows that such a stream can be strongly perturbed by the close passage of other clusters, in particular of NGC~2808, NGC~7078, and NGC~104, and that these perturbations induce the formation of gaps in the tails.}
  % conclusions heading (optional), leave it empty if necessary
-  {These results show that globular clusters can induce gaps in \textbf{\paola{cold}} streams--just as others have shown that other baryonic components such as giant molecular clouds and the galactic bar may as well.
+  {These results show that globular clusters can induce gaps in cold streams--just as others have shown that other baryonic components, such as giant molecular clouds and the galactic bar, may as well.
 Therefore, a future work that attempts to infer the dark matter sub-halo distribution from stellar stream gaps must include the contributions from globular clusters.}
@@
 To this end, on the following pages, we will present the results of simulations of the streams of Palomar~5, subject to the gravitational interaction with the set of 165 Galactic globular clusters for which positions and velocities are known to date and for which orbits can therefore be reconstructed \citep{2021MNRAS.505.5957B}.
-We chose to simulate \textbf{\paola{streams formed from a cluster with the current characteristics of}} Palomar~5 because \textbf{\paola{this latter}} is a halo cluster with extended tails and because it is a cluster for which the effect of baryonic structures on its stream has already been studied.
-As we will show, and in tension with previous claims, the close passage of other clusters with \textbf{\paola{such a stream}} is not rare.
+We chose to simulate streams formed from a cluster with the current characteristics of Palomar~5 because this latter is a halo cluster with extended tails, and because it is a cluster for which the effect of baryonic structures on its stream has already been studied.
+As we will show, and in tension with previous claims, the close passage of other clusters with such a stream is not rare.
 Indeed, in the 50 simulations we ran, we find the formation of numerous gaps, for an average of 1.5 gaps per simulation, generated by 18 different clusters among the whole system of Galactic globular clusters.
@@
 As demonstrated by \citet{2012A&A...546L...7M}, although the restricted three-body problem neglects the internal evolution of the cluster, it still reproduces very similar stream properties since the model captures key extra-cluster physics, such as disk shocking and epicyclic stripping.\\
-  \textbf{\paola{Below, we first present the approach used to include the Galactic globular cluster system in our modeling (Sect.~\ref{numerical_meth}), highlighting the similarities and differences with respect to what we did in \citet{2023A&A...673A..44F}; we then summarize the method used to model the mass loss from the cluster (Sect.~\ref{sec:mass_loss}) and finally discuss the quality of the numerical integration (Sect.~\ref{num_quality}).
-}}
+  Below, we first present the approach used to include the Galactic globular cluster system in our modeling (Sect.~\ref{numerical_meth}), highlighting the similarities and differences with respect to what we did in \citet{2023A&A...673A..44F}; we then summarize the method used to model the mass loss from the cluster (Sect.~\ref{sec:mass_loss}) and finally discuss the quality of the numerical integration (Sect.~\ref{num_quality}).
 
@@
 
-        \item Once the we find the positions and velocities of the entire globular cluster system 5~Gyr ago, we sample Palomar~5 with 100,000 particles from a Plummer distribution, taking the mass and half-mass radius from the Baumgardt catalog \paola{can you add here the current values of mass and size of Pal 5 in Baumgardt?}.
+        \item Once the we find the positions and velocities of the entire globular cluster system 5~Gyr ago, we sample Palomar~5 with 100,000 particles from a Plummer distribution, taking the mass and half-mass radius from the Baumgardt catalog: $1.3\times10^{4}~\textrm{M}_\odot$ \& $27.6~\textrm{pc}$.
 We then integrate the evolution of these particles forward in time to the present day by taking into account that each particle feels the gravitational potential of the Galaxy, that of its host cluster, and that of all the other clusters in the Galaxy.
@@
 
-\textbf{\paola{\subsection{Mass loss}}}\label{sec:mass_loss}
-
-\textbf{\paola{Each of the 100,000 particles that initially populate the cluster undergoes, over time, acceleration due to the Pal~5 potential, whose mass and half-mass radius we assume to be fixed over time, and to the Galactic potential.
+\subsection{Mass loss}
+
+Each of the 100,000 particles that initially populate the cluster undergoes, over time, acceleration due to the Pal~5 potential, whose mass and half-mass radius we assume to be fixed over time, and to the Galactic potential.
 At each time step, a certain number of particles will therefore acquire sufficient energy to no longer be gravitationally bound to the cluster itself and thus go on to populate the streams, whose mass and spatial extent grow over time.
-It is important to note that in the approach used:}}
-\textbf{\paola{
+It is important to note that in the approach used:
+
 \begin{enumerate}
     \item The masses and sizes of the clusters (and therefore the parameters of the Plummer potentials) do not change over time, which is a blatant simplification, because in a self-consistent approach, these parameters would vary.
-    \item We use the same initial conditions for Pal~5 progenitor as it has today, and this is also a simplification, since Pal~5 - 5 Gyr years ago - must have contained at least part of the mass estimated today in its tails\footnote{\textbf{\paola{We use the same approach (i.e.
-time-independent masses and sizes) to model the whole set of globular clusters.}}}.
-\end{enumerate}}}
-
-\textbf{\paola{The assumption in point 2 is a direct consequence of the approach described in point 1.
+    \item We use the same initial conditions for Pal~5 progenitor as it has today, and this is also a simplification, since Pal~5 - 5 Gyr years ago - must have contained at least part of the mass estimated today in its tails\footnote{We use the same approach (i.e.
+time-independent masses and sizes) to model the whole set of globular clusters.}.
+\end{enumerate}
+
+The assumption in point 2 is a direct consequence of the approach described in point 1.
 Starting from a cluster with mass and size similar to the current ones can lead to streams with lower velocity dispersions than those we would have obtained if we had used a self-consistent approach.
@@
 We note, however, that simplifications of this kind are not uncommon in literature.
-\citet{2017NatAs...1..633P} discussed the formation of gaps in the Pal~5 tails and assumed a time-independent mass of $50,000 M_\odot$ for Pal~5, over the last 4~Gyr; \citet{2017MNRAS.470...60E} adopted a N-body approach to simulate Pal~5 stream, but used Pal~5 current conditions as their progenitor's initial conditions; \citet{2019MNRAS.484.2009B} simulated the Pal~5 stream as emerging from a stellar system with a velocity dispersion of 0.5~km/s (similar to that of particles escaping from our cluster, as we have verified).
-\\The characteristics of the streams modeled in this paper may be considered more representative of those of clusters that are now completely dispersed, i.e.
-it is conceivable that completely dispersed globular clusters that left behind a population of 'orphan' streams passed through characteristics similar to those of Pal~5 today (small masses and extended radii).
-In this sense, the initial conditions chosen (in terms of internal parameters) could be considered more representative of those of streams found in the Galaxy and for which the progenitor is now dissolved \citep[see, for example, the population of streams without progenitors described by][]{2024ApJ...967...89I} than those currently typical of Galactic globular clusters\footnote{\textbf{\paola{In this regard, we recall that \citet{2014ApJ...795...95B} modeled the GD-1 stream as the result of the dissolution of a cluster with a mass of $2 \times 10^4 M_\odot$, and a tidal radius of $0.07$~kpc.}}}.
-}}
+\citet{2017NatAs...1..633P} discussed the formation of gaps in the Pal~5 tails and assumed a time-independent mass of $50,000 ~\textrm{M}_\odot$ for Pal~5, over the last 4~Gyr; \citet{2017MNRAS.470...60E} adopted a N-body approach to simulate Pal~5 stream, but used Pal~5 current conditions as their progenitor's initial conditions; \citet{2019MNRAS.484.2009B} simulated the Pal~5 stream as emerging from a stellar system with a velocity dispersion of 0.5~km/s (similar to that of particles escaping from our cluster, as we have verified).
+
+The characteristics of the streams modeled in this paper may be considered more representative of those of clusters that are now completely dispersed, i.e.
+it is conceivable that completely dispersed globular clusters that left behind a population of `orphan' streams passed through characteristics similar to those of Pal~5 today (small masses and extended radii).
+In this sense, the initial conditions chosen (in terms of internal parameters) could be considered more representative of those of streams found in the Galaxy and for which the progenitor is now dissolved \citep[see, for example, the population of streams without progenitors described by][]{2024ApJ...967...89I} than those currently typical of Galactic globular clusters\footnote{In this regard, we recall that \citet{2014ApJ...795...95B} modeled the GD-1 stream as the result of the dissolution of a cluster with a mass of $2 \times 10^4 M_\odot$, and a tidal radius of $0.07$~kpc.}
 
@@
 Thus, in one configuration, a cluster can impact the stream at a given time, and yet at the same moment, in a different set of initial conditions, it could be on the other side of the galaxy.
-\textbf{\paola{Interestingly, some gaps, formed by close passages occurred early on, can persist for several billions years in our models and be visible until the end of the simulation -- as it is the same for NGC~104, which produces the gap visible in Fig.~2 at about 8~kpc from the cluster.
-We will discuss how these gaps can survive for so long in Appendix~\ref{sec:gaps_vs_gcmass}.
+\textbf{\salvatore{We will discuss the necessary conditions for creating a gap in Appendix~\ref{sec:gaps_vs_gcmass}.
 }}
@@
 However, only a few gaps occur at the shortest reported stream length.
-\textbf{\paola{We note, in particular, that the gap generated by the recent perturbation induced on the stream by the passage of NGC~2808 (whose occurrence is very likely according to our models) sits in a portion of the simulated tail which is at the edge of the observed one (see bottom panel in Fig.~2).
+We note, in particular, that the gap generated by the recent perturbation induced on the stream by the passage of NGC~2808 (whose occurrence is very likely according to our models) sits in a portion of the simulated tail which is at the edge of the observed one (see bottom panel in Fig.~2).
 This element and the fact that this gap is skinny, because it is very recent, probably make its detection difficult.
-}}
       \begin{figure}
@@
 
-\textbf{\paola{An additional explanation for the absence of gaps in the inner regions of a stream is given in App.xx.
-If a gap forms in a stream region where stars from the cluster coexist at different times, all stars lost at times before the impact acquire momentum and acceleration due to the perturber.
-Stars lost at different times move, however, in a given region of the stream with different velocities, and the same is true for the respective gaps, which will therefore lose spatial coherence over a time that is proportional to the size of the gap and inversely proportional to the velocity dispersion in that section of the stream.
-Since the inner regions of streams are those where stars lost at different times coexist more easily, this element must also be taken into account when discussing the visibility (or persistence) of gaps, and this independently of the nature of the perturber.}}
+\textbf{\salvatore{An additional explanation for the absence of gaps in the inner regions of a stream is given in App.~\ref{sec:gaps_vs_gcmass}.
+Briefly, the eccentricity of an orbit induces tidal shocking at the pericenter passages that cause episodes of increased mass loss where the escaped stars leave with a higher mean velocity and greater velocity dispersion.
+In essence, streams from progenitors on eccentric orbits are made of two components: a continuous flow of stars plus many packets of stars that burst out from pericenter passages.
+The gap impact occurs at a single position, creating a gap in each subpopulation.
+However, the gaps have different drift rates across the groups, and they eventually go out of phase, erasing the impact's signature.
+The packets of stars disperse with time.
+However, bursts contribute to more escaped stars as the simulation evolves than the continuous outflow.
+As a result, at late stages of the simulation, the region nearest to the globular cluster is made of distinct yet overlapping populations.
+After an impact in this location, the impact site between the groups quickly goes out of phase.
+}}
 
@@
 
-  \textbf{ \paola{\citet{2017MNRAS.470...60E} is another world that discusses the impact of globular clusters in producing gaps on the Pal~5 stream.
-In section~6.3 of their article, \citet{2017MNRAS.470...60E} discuss the fact that globular clusters with masses greater than $10^6 M_\odot$ are rare in the Galaxy, especially at distances compatible with the orbit of Pal~5, while -- based on previous works -- they estimate that sub-halos of dark matter of similar mass are at least three times greater in number.
+  \textbf{ \salvatore{\citet{2017MNRAS.470...60E} is another work that discusses the impact of globular clusters in producing gaps on the Pal~5 stream.
+In section~6.3 of their article, \citet{2017MNRAS.470...60E} discuss the fact that globular clusters with masses greater than $10^6~\textrm{M}_\odot$ are rare in the Galaxy, especially at distances compatible with the orbit of Pal~5, while -- based on previous works -- they estimate that sub-halos of dark matter of similar mass are at least three times greater in number.
 This difference leads them to conclude that the gaps they report in the Pal~5 streams are more likely to be induced by dark matter sub-halos than by globular clusters.
-  However, as we show in this paper, even clusters with masses below $10^6 M_\odot$ can produce gaps.
+  However, as we show in this paper, even clusters with masses below $10^6~\textrm{M}\odot$ can produce gaps.
 The clusters perturbing the Pal~5 stream are clusters with masses up to 10 times lower than those considered by \citet{2017MNRAS.470...60E}.
 It is thus possible that \citet{2017MNRAS.470...60E} have underestimated the impact of globular clusters' close passages on a stream such as Pal~5.
+We can derive from \citet{2016MNRAS.457.3817S}'s third figure that a more concentrated system delivers a higher velocity kick but over a shorter distance, while a more spread out system will affect more stars yet perturb them less.
 It is more difficult, however, to make a comparative analysis between the role of clusters and sub-halos at equal stellar mass.
@@
 For simplicity, both clusters have the same half-mass radius of 27~pc.
-The encounter occurred about two and a half billion years ago by NGC~7078, which has a mass of XX.
+The encounter occurred about two and a half billion years ago by NGC~7078, which has a mass of $5.16\times10^{5}~M_\odot$.
 The gap persists for the lighter host, yet is barely visible in the case of the more massive host.
@@
 
-\salvatore{The situation can be simplified as one-dimensional collisionless dispersion.
+\salvatore{Each group can be simplified as one-dimensional collisionless dispersion.
 These packets of stars can be described as a population with the same initial position that drifts about a mean velocity, $\langle v \rangle$, with velocity dispersion, $\sigma_v$.
@@
 Regarding our simulations, since the mass and radius of the host are held constant, the mean velocity and velocity dispersion are the same across all packets.
-The density of each packet decreases since the easier-to-remove stars are stripped early on.
+Secondly, the density of each packet decreases since the easier-to-remove stars are stripped early on.
 }
@@
 
-\salvatore{In the simplest consideration, of two gaps of the same dimension $\sigma_x$ and an offset of $\delta x$, the gaps are out of phase when $1<\sigma_x/\delta x$.
+\salvatore{In the simplest consideration, of two gaps of the same dimension $\sigma_x$ and an offset of $\delta x$, the gaps are out of phase when the separation is greater than the width: $1<\sigma_x/\delta x$.
 In the temporal analysis of the stream presented in Fig.~\ref{fig:decoherence_shocked_outflow}, we did not observe at any point of the simulation that the gap initially persisted to then go out of phase.

### 2025-05-15 paper/main.tex
updated caption for new figure

@@
 A consequence of the pericenter passages is that many more stars escape, fleeing with a higher mean velocity and velocity dispersion, which can be visualized in the \textit{Christmas tree} illustration in Fig.~A.3 of \citet{2012A&A...546L...7M}.
-We consider stars that leave between pericenter passages to be \textit{leaked} stars.
+We consider the continous outflow of stars between pericenter passages to be \textit{leaked} stars.
 We identify pericenter passages by computing each local minimum of the cluster's galactic-centric 3D radius with \texttt{scipy.signal.find\_peaks}.
 The duration of each shock is: $t_{\textrm{shock}}$ = 2$\sqrt{r/|\nabla\Phi|}$, where $r$ is the galactocentric radius and $|\nabla\Phi|$ is the magnitude of the acceleration from the galactic potential.
-Stars within this range about a given pericenter passage are all given the same color and shown as individual groups in the first, second, and third panels of Fig.~\ref{fig:decoherence_shocked_outflow}.}
+Stars within this range about a given pericenter passage are all given the same color and shown as individual groups in the first, second, third, and fourth panels of Fig.~\ref{fig:decoherence_shocked_outflow}.}
 
@@
 In this simplified 1D collisionless dispersion analysis, the gaps between the different packets space apart linearly in time with $\delta x\propto \delta vt$, where $\delta v$ is the mean velocity of the about the impact site for two different packets.
-This inclines us to believe that the gaps space apart before they become significant.
+This inclines us to believe that the gaps space apart before they become significantly deep or wide, however this need not be the general case.
 }
@@
 
-\begin{figure*}
+  \begin{figure}
       \includegraphics[width=\linewidth]{decoherence_shocked_outflow.png}
-    \caption{We dissect the density of the stream from the bottom panel of Fig.~\ref{fig:gap_no_more} into stars that are either shocked or leaked out, see text for details.
-The bottom panel shows the total 1D profile with the two different populations.
-The second from the bottom dissects the total shocked outflow into individual pericenter passages.
-Each of which is color-coated by the escape time.
-The top two panels separate the first four shocked outflow populations to aid visualization of the gap placement from the impact of NGC~7078.
-}
+      \caption{We dissect the one-dimensional density of the stream from the bottom panel of Fig.~\ref{fig:gap_no_more}.
+We divide the stars into two categories: the multiple groups that are shocked out at different pericenter passages, or those that leak out of the cluster continuously.
+The bottom panel shows the leaked population in blue, and each group that was shocked out at each pericenter passage is color-coded by the escape time.
+The sum of the leaked and shocked outflow gives the total profile density, which is shown in black.
+The top four panels display the first shocked groups from different pericenter passages, shown individually.
+Each group shows the gap resulting from the impact of NGC~7078 $\sim$2.5~Gyr ago, yet they are progressively offset as explained in the text.}
       \label{fig:decoherence_shocked_outflow}
-\end{figure*}
+  \end{figure}
 

### 2025-05-15 paper/main.tex
uncommented plots, added label, fixed ??

@@
 
-  \date{SUBMITTED TO A\&A 27 JAN 2025; Accepted XXX}
+  \date{SUBMITTED TO A\&A XXX; Accepted XXX}
 
@@
 
-\subsection{Mass loss}
+\subsection{Mass loss}\label{sec:mass_loss}
 
@@
       \includegraphics[width=\linewidth]{decomposition-monte-carlo-009-with-3-gaps-domidpoint-shift.png}
-      \caption{Density maps of Palomar~5's stream in the tail coordinate system.
+      \caption{Density maps of Palomar~5's stream in the tail coordinate system, where $x\prime$ is the integrate arclength along the cluster's orbit and $y\prime$ is the distance within the orbital plane.
 The color scale represents normalized particle counts (total: 100,000).
@@
 Moreover, we note that \citet{2022ApJ...941..129D} pre-selects the globular clusters that could have experienced a close encounter with the GD-1 stream by selecting only clusters that pass at a distance less than 0.5 kpc from the stream, having a relative velocity smaller than  300 km/s.
-As shown in our Fig.~\ref{fig:impact_geometry_statistics}, bottom-left panel, in the case of Palomar~5, this choice would lead to excluding most of the encounters favorable to gap creation, which turn out to have relative velocities above 300 km/s.
+As shown in our Fig.~\ref{fig:impact_geometry_statistics}, bottom-left panel, in the case of Palomar~5, this choice would lead to excluding most of the encounters favorable to gap creation, which turn out to have relative velocities above 300 km/s, \textbf{reducing the total to just 19 gaps}.
 It would be interesting to repeat a similar study as the one made by \citet{2022ApJ...941..129D} for GD-1, imposing no selection on possible candidate clusters.\\
@@
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-000.png}
-      % \includegraphics[width=\linewidth]{tau-profile-monte-carlo-000.png}
-      % \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-001.png}
-      % \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-002.png}
-      % \includegraphics[width=\linewidth]{tau-profile-monte-carlo-002.png}
+      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-000.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-001.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-002.png}
+      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-002.png}
       \caption{First gap gallery. See the text in Appendix~\ref{sec:gallery_of_gaps} for a description.}
@@
       \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-003.png}      
-      % \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-004.png}
-      % \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-005.png}
-      % \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-006.png}
-      % \includegraphics[width=\linewidth]{tau-profile-monte-carlo-006.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-004.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-005.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-006.png}
+      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-006.png}
       \caption{Second gap gallery}
@@
 
-    % \begin{figure*}
-    %   \centering
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-007.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-008.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-009.png}      
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-010.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-011.png}
-    %   \caption{Third gap gallery}
-    %   \label{fig:gallery2}
-    %   \end{figure*}
-
-    % \begin{figure*}
-    %   \centering
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-012.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-013.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-014.png}      
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-015.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-016.png}
-    %   \caption{Fourth gap gallery}
-    %   \label{fig:gallery3}
-    %   \end{figure*}
-
-    % \begin{figure*}
-    %   \centering
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-017.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-018.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-019.png}      
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-020.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-021.png}
-    %   \caption{Fifth gap gallery}
-    %   \label{fig:gallery4}
-    %   \end{figure*}
-
-    % \begin{figure*}
-    %   \centering
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-022.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-023.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-024.png}   
-    %   \includegraphics[width=\linewidth]{tau-profile-monte-carlo-024.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-025.png}
-    %   \caption{Sixth gap gallery}
-    %   \label{fig:gallery5}
-    %   \end{figure*}
-
-    % \begin{figure*}
-    %   \centering
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-026.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-027.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-028.png}
-    %   \includegraphics[width=\linewidth]{tau-profile-monte-carlo-028.png}
-    %   \caption{Seventh gap gallery}
-    %   \label{fig:gallery6}
-    %   \end{figure*}
-
-    % \begin{figure*}
-    %   \centering      
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-029.png}    
-    %   \includegraphics[width=\linewidth]{tau-profile-monte-carlo-029.png}  
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-030.png}
-    %   \includegraphics[width=\linewidth]{tau-profile-monte-carlo-030.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-031.png}
-    %   \caption{Eighth gap gallery}
-    %   \label{fig:gallery7}
-    %   \end{figure*}
-
-    % \begin{figure*}
-    %   \centering
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-032.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-033.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-034.png}      
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-035.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-036.png}      
-    %   \caption{Ninth gap gallery}
-    %   \label{fig:gallery8}
-    %   \end{figure*}
-
-    % \begin{figure*}
-    %   \centering
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-037.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-038.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-039.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-040.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-041.png}      
-    %   \caption{Tenth gap gallery}
-    %   \label{fig:gallery9}
-    % \end{figure*}
-
-    % \begin{figure*}
-    %   \centering
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-042.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-043.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-044.png}
-    %   \includegraphics[width=\linewidth]{tau-profile-monte-carlo-044.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-045.png}
-    %   \caption{Eleventh gap gallery}
-    %   \label{fig:gallery010}
-    % \end{figure*}
-
-    % \begin{figure*}
-    %   \centering
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-046.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-047.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-048.png}
-    %   \includegraphics[width=\linewidth]{tau-profile-monte-carlo-048.png}
-    %   \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-049.png}
-    %   \caption{Twelfth gap gallery}
-    %   \label{fig:gallery11}
-    % \end{figure*}
+    \begin{figure*}
+      \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-007.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-008.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-009.png}      
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-010.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-011.png}
+      \caption{Third gap gallery}
+      \label{fig:gallery2}
+      \end{figure*}
+
+    \begin{figure*}
+      \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-012.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-013.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-014.png}      
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-015.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-016.png}
+      \caption{Fourth gap gallery}
+      \label{fig:gallery3}
+      \end{figure*}
+
+    \begin{figure*}
+      \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-017.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-018.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-019.png}      
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-020.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-021.png}
+      \caption{Fifth gap gallery}
+      \label{fig:gallery4}
+      \end{figure*}
+
+    \begin{figure*}
+      \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-022.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-023.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-024.png}   
+      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-024.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-025.png}
+      \caption{Sixth gap gallery}
+      \label{fig:gallery5}
+      \end{figure*}
+
+    \begin{figure*}
+      \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-026.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-027.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-028.png}
+      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-028.png}
+      \caption{Seventh gap gallery}
+      \label{fig:gallery6}
+      \end{figure*}
+
+    \begin{figure*}
+      \centering      
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-029.png}    
+      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-029.png}  
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-030.png}
+      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-030.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-031.png}
+      \caption{Eighth gap gallery}
+      \label{fig:gallery7}
+      \end{figure*}
+
+    \begin{figure*}
+      \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-032.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-033.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-034.png}      
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-035.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-036.png}      
+      \caption{Ninth gap gallery}
+      \label{fig:gallery8}
+      \end{figure*}
+
+    \begin{figure*}
+      \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-037.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-038.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-039.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-040.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-041.png}      
+      \caption{Tenth gap gallery}
+      \label{fig:gallery9}
+    \end{figure*}
+
+    \begin{figure*}
+      \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-042.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-043.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-044.png}
+      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-044.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-045.png}
+      \caption{Eleventh gap gallery}
+      \label{fig:gallery010}
+    \end{figure*}
+
+    \begin{figure*}
+      \centering
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-046.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-047.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-048.png}
+      \includegraphics[width=\linewidth]{tau-profile-monte-carlo-048.png}
+      \includegraphics[width=\linewidth]{gallery_of_gaps_monte-carlo-049.png}
+      \caption{Twelfth gap gallery}
+      \label{fig:gallery11}
+    \end{figure*}
 

### 2025-05-15 paper/main.tex
deleted my comments in the article

@@
 Thus, in one configuration, a cluster can impact the stream at a given time, and yet at the same moment, in a different set of initial conditions, it could be on the other side of the galaxy.
-\textbf{\salvatore{We will discuss the necessary conditions for creating a gap in Appendix~\ref{sec:gaps_vs_gcmass}.
-}}
+We will discuss the necessary conditions for creating a gap in Appendix~\ref{sec:gaps_vs_gcmass}.
 
@@
 
-\textbf{\salvatore{An additional explanation for the absence of gaps in the inner regions of a stream is given in App.~\ref{sec:gaps_vs_gcmass}.
+An additional explanation for the absence of gaps in the inner regions of a stream is given in App.~\ref{sec:gaps_vs_gcmass}.
 Briefly, the eccentricity of an orbit induces tidal shocking at the pericenter passages that cause episodes of increased mass loss where the escaped stars leave with a higher mean velocity and greater velocity dispersion.
@@
 After an impact in this location, the impact site between the groups quickly goes out of phase.
-}}
 
@@
 
-  \textbf{ \salvatore{\citet{2017MNRAS.470...60E} is another work that discusses the impact of globular clusters in producing gaps on the Pal~5 stream.
+  \citet{2017MNRAS.470...60E} is another work that discusses the impact of globular clusters in producing gaps on the Pal~5 stream.
 In section~6.3 of their article, \citet{2017MNRAS.470...60E} discuss the fact that globular clusters with masses greater than $10^6~\textrm{M}_\odot$ are rare in the Galaxy, especially at distances compatible with the orbit of Pal~5, while -- based on previous works -- they estimate that sub-halos of dark matter of similar mass are at least three times greater in number.
@@
 Being less dense than globular clusters, sub-halos should produce less intense perturbations on streams (for fixed impact parameters and relative velocities).
-A systematic comparative analysis to quantify the role of clusters and sub-halos in producing gaps and perturbations in stellar streams still needs to be done.}}
+A systematic comparative analysis to quantify the role of clusters and sub-halos in producing gaps and perturbations in stellar streams still needs to be done.
 
@@
 
-  \salvatore{Fig.~\ref{fig:decoherence_shocked_outflow} dissects the more massive stream of Fig.~\ref{fig:gap_no_more} and instructs us why the gap does not persist.
+  Fig.~\ref{fig:decoherence_shocked_outflow} dissects the more massive stream of Fig.~\ref{fig:gap_no_more} and instructs us why the gap does not persist.
 The bottom panel shows the total density distribution, and dissects it into two profiles: stars that either are \textit{leaked} or \textit{shocked} out of the cluster.
@@
 The duration of each shock is: $t_{\textrm{shock}}$ = 2$\sqrt{r/|\nabla\Phi|}$, where $r$ is the galactocentric radius and $|\nabla\Phi|$ is the magnitude of the acceleration from the galactic potential.
-Stars within this range about a given pericenter passage are all given the same color and shown as individual groups in the first, second, third, and fourth panels of Fig.~\ref{fig:decoherence_shocked_outflow}.}
-
-  \salvatore{Each group can be simplified as one-dimensional collisionless dispersion.
+Stars within this range about a given pericenter passage are all given the same color and shown as individual groups in the first, second, third, and fourth panels of Fig.~\ref{fig:decoherence_shocked_outflow}.
+
+  Each group can be simplified as one-dimensional collisionless dispersion.
 These packets of stars can be described as a population with the same initial position that drifts about a mean velocity, $\langle v \rangle$, with velocity dispersion, $\sigma_v$.
-The density evolution is a Gaussian whose amplitude decreases inversely with time, and whose standard deviation grows linearly with time:}
+The density evolution is a Gaussian whose amplitude decreases inversely with time, and whose standard deviation grows linearly with time:
   \begin{equation}    
@@
 
-  \salvatore{It is important to note that within a single packet, the stars self-segregate.
+  It is important to note that within a single packet, the stars self-segregate.
 The position of the density peak is at $\langle v \rangle t$, while the faster stars are ahead and the slower ones are behind.
@@
 Secondly, the density of each packet decreases since the easier-to-remove stars are stripped early on.
-}
-
-  \salvatore{Consider an impact that happens at $x\prime$.
+
+  Consider an impact that happens at $x\prime$.
 The distance from the impact site to each packet is: $\delta x= x\prime - \langle v\rangle \left(t-n T\right)$, where $n$ is the packet index, with $n=0$ being the first packet escaped at the first pericenter passage in the simulation.
@@
 As a result, the gaps in the later packets quickly overtake those from the earlier packets.
-As a result, the gaps across the packets go out of phase and interfere destructively.}
-
-  \salvatore{In the simplest consideration, of two gaps of the same dimension $\sigma_x$ and an offset of $\delta x$, the gaps are out of phase when the separation is greater than the width: $1<\sigma_x/\delta x$.
+As a result, the gaps across the packets go out of phase and interfere destructively.
+
+  In the simplest consideration, of two gaps of the same dimension $\sigma_x$ and an offset of $\delta x$, the gaps are out of phase when the separation is greater than the width: $1<\sigma_x/\delta x$.
 In the temporal analysis of the stream presented in Fig.~\ref{fig:decoherence_shocked_outflow}, we did not observe at any point of the simulation that the gap initially persisted to then go out of phase.
@@
 This inclines us to believe that the gaps space apart before they become significantly deep or wide, however this need not be the general case.
-}
-
-  \salvatore{Nonetheless, we believe this is the first consideration of decomposing the stream into packets of tidal bursts to understand the effect of the internal dynamics on gap persistence.
+
+  Nonetheless, we believe this is the first consideration of decomposing the stream into packets of tidal bursts to understand the effect of the internal dynamics on gap persistence.
 Since gap persistence depends on the mass and internal evolution of the cluster, characterizing the gap can constrain our knowledge of the internal evolution.
-}
-  % \paola{What happens is that the different and in particular in the populations of stars in this part of the tail that escaped from the cluster at different times, in each of these populations the respective gap tends to travel along the streams at different speeds, which leads to a phase shift of the gap and its loss of visibility.
-In other words, the gap can survive for billions of years in each population of stars that left the cluster at different times, but it is no longer.
-Still, initially coherent and therefore visible because each of these populations travels along the tail (and drags a gap with it) at different speeds.
-The loss of coherence occurs at a time $t_{visibility}$ which is directly proportional to the gap size and inversely proportional to the velocity dispersion in that part of the stream.
-So, if populations of stars lost at different times co-exist in the portion of the stream where the impact occurs, the gap will remain visible for a time that depends on the mass of the progenitor system at the time these stars escaped.
-From this, we can understand that various factors determine the visibility of a gap in a stream.
-Mass affects the velocity dispersion in a given region of the stream, but this velocity dispersion, in turn, depends on the history of mass loss of the cluster.
-Suppose stars dominate a given stream region, and all are lost simultaneously from the cluster.
-In that case, the resulting velocity dispersion is lower than if various populations coexist in the same region, which has implications for these structures' persistence.
-}
 

### 2025-05-15 paper/main.tex
reintroduced green text

--- 2025-05-15 paper/main.tex
@@
 
-    By applying the methodology described above to the whole set of simulations, we can reconstruct the history of close passages of Galactic globular clusters to Palomar~5's stream in the last 5~Gyr, which -- we remind the reader -- is the time interval investigated in our simulations.
-Fig.~\ref{fig:histogram_impact_time} and Table~\ref{tab:gaps_per_perturber} present results of this analysis.
-NGC~2808 impacted Palomar~5's stream about 200 Myr ago in 44 out of 50 simulations, creating a small gap at a similar position to the one reported in Fig.~\ref{fig:decomposition} in each case.
-Since this interaction occurred less than one orbital period ago, despite the uncertainties, the orbital solutions remain similar and thus produce consistent results across all simulations.
-However, as we continue to turn back time further, the uncertainties in the initial conditions allow the various orbital solutions to diverge from one another.
-Thus, in one configuration, a cluster can impact the stream at a given time, and yet at the same moment, in a different set of initial conditions, it could be on the other side of the galaxy.
-\textbf{\salvatore{We will discuss the necessary conditions for creating a gap in Appendix~\ref{sec:gaps_vs_gcmass}.
+    \textbf{\salvatore{An additional explanation for the absence of gaps in the inner regions of a stream is given in App.~\ref{sec:gaps_vs_gcmass}.
+Briefly, the eccentricity of an orbit induces tidal shocking at the pericenter passages that cause episodes of increased mass loss where the escaped stars leave with a higher mean velocity and greater velocity dispersion.
+In essence, streams from progenitors on eccentric orbits are made of two components: a continuous flow of stars plus many packets of stars that burst out from pericenter passages.
+The gap impact occurs at a single position, creating a gap in each subpopulation.
+However, the gaps have different drift rates across the groups, and they eventually go out of phase, erasing the impact's signature.
+The packets of stars disperse with time.
+However, bursts contribute to more escaped stars as the simulation evolves than the continuous outflow.
+As a result, at late stages of the simulation, the region nearest to the globular cluster is made of distinct yet overlapping populations.
+After an impact in this location, the impact site between the groups quickly goes out of phase.
 }}
@@
 
-  \salvatore{Each group can be simplified as one-dimensional collisionless dispersion.
+  Each group can be simplified as one-dimensional collisionless dispersion.
 These packets of stars can be described as a population with the same initial position that drifts about a mean velocity, $\langle v \rangle$, with velocity dispersion, $\sigma_v$.
-The density evolution is a Gaussian whose amplitude decreases inversely with time, and whose standard deviation grows linearly with time:}
+The density evolution is a Gaussian whose amplitude decreases inversely with time, and whose standard deviation grows linearly with time:
   \begin{equation}    
@@
 Since gap persistence depends on the mass and internal evolution of the cluster, characterizing the gap can constrain our knowledge of the internal evolution.
-}
-  % \paola{What happens is that the different and in particular in the populations of stars in this part of the tail that escaped from the cluster at different times, in each of these populations the respective gap tends to travel along the streams at different speeds, which leads to a phase shift of the gap and its loss of visibility.
-In other words, the gap can survive for billions of years in each population of stars that left the cluster at different times, but it is no longer.
-Still, initially coherent and therefore visible because each of these populations travels along the tail (and drags a gap with it) at different speeds.
-The loss of coherence occurs at a time $t_{visibility}$ which is directly proportional to the gap size and inversely proportional to the velocity dispersion in that part of the stream.
-So, if populations of stars lost at different times co-exist in the portion of the stream where the impact occurs, the gap will remain visible for a time that depends on the mass of the progenitor system at the time these stars escaped.
-From this, we can understand that various factors determine the visibility of a gap in a stream.
-Mass affects the velocity dispersion in a given region of the stream, but this velocity dispersion, in turn, depends on the history of mass loss of the cluster.
-Suppose stars dominate a given stream region, and all are lost simultaneously from the cluster.
-In that case, the resulting velocity dispersion is lower than if various populations coexist in the same region, which has implications for these structures' persistence.
 }

### 2025-05-23 paper/main.tex
added back the paragraph in "The history and ..."

@@
 
-    \textbf{\salvatore{An additional explanation for the absence of gaps in the inner regions of a stream is given in App.~\ref{sec:gaps_vs_gcmass}.
-Briefly, the eccentricity of an orbit induces tidal shocking at the pericenter passages that cause episodes of increased mass loss where the escaped stars leave with a higher mean velocity and greater velocity dispersion.
-In essence, streams from progenitors on eccentric orbits are made of two components: a continuous flow of stars plus many packets of stars that burst out from pericenter passages.
-The gap impact occurs at a single position, creating a gap in each subpopulation.
-However, the gaps have different drift rates across the groups, and they eventually go out of phase, erasing the impact's signature.
-The packets of stars disperse with time.
-However, bursts contribute to more escaped stars as the simulation evolves than the continuous outflow.
-As a result, at late stages of the simulation, the region nearest to the globular cluster is made of distinct yet overlapping populations.
-After an impact in this location, the impact site between the groups quickly goes out of phase.
-}}
+    By applying the methodology described above to the whole set of simulations, we can reconstruct the history of close passages of Galactic globular clusters to Palomar~5's stream in the last 5~Gyr, which -- we remind the reader -- is the time interval investigated in our simulations.
+Fig.~\ref{fig:histogram_impact_time} and Table~\ref{tab:gaps_per_perturber} present results of this analysis.
+NGC~2808 impacted Palomar~5's stream about 200 Myr ago in 44 out of 50 simulations, creating a small gap at a similar position to the one reported in Fig.~\ref{fig:decomposition} in each case.
+Since this interaction occurred less than one orbital period ago, despite the uncertainties, the orbital solutions remain similar and thus produce consistent results across all simulations.
+However, as we continue to turn back time further, the uncertainties in the initial conditions allow the various orbital solutions to diverge from one another.
+Thus, in one configuration, a cluster can impact the stream at a given time, and yet at the same moment, in a different set of initial conditions, it could be on the other side of the galaxy.
+We will discuss the necessary conditions for creating a gap in Appendix~\ref{sec:gaps_vs_gcmass}.
 
@@
 
-  \salvatore{In the simplest consideration, of two gaps of the same dimension $\sigma_x$ and an offset of $\delta x$, the gaps are out of phase when the separation is greater than the width: $1<\sigma_x/\delta x$.
-In the temporal analysis of the stream presented in Fig.~\ref{fig:decoherence_shocked_outflow}, we did not observe at any point of the simulation that the gap initially persisted to then go out of phase.
+  \salvatore{In the simplest consideration, of two packets of the same dimension $\sigma_x$ and an offset of $\delta x$, the gaps are out of phase when the separation is greater than the width: $1\sim<\sigma_x/\delta x$.
 Drawing from the wisdom in \citet{2015MNRAS.450.1136E} \& \citet{2016MNRAS.457.3817S}, we see the contrast between the density peaks and the density minimum of a gap, as well as the width, grow in time, somewhere between $\propto \sqrt{t}$ and linear.

### 2025-05-23 paper/main.tex
tried to make the appendix less confusing

@@
 
-  Each group can be simplified as one-dimensional collisionless dispersion.
-These packets of stars can be described as a population with the same initial position that drifts about a mean velocity, $\langle v \rangle$, with velocity dispersion, $\sigma_v$.
-The density evolution is a Gaussian whose amplitude decreases inversely with time, and whose standard deviation grows linearly with time:
-  \begin{equation}    
-          \rho(x,t) = \frac{1}{\sigma_v t \sqrt{2\pi} }\textrm{exp}\left(-\frac{1}{2}\left(\frac{x-\langle v \rangle t}{\sigma_v t}\right)^2\right).
-  \end{equation}
-
-    \salvatore{It is important to note that within a single packet, the stars self-segregate.
-The position of the density peak is at $\langle v \rangle t$, while the faster stars are ahead and the slower ones are behind.
-The packets are separated by a distance of $\langle  v\rangle T$, where $T$ is the orbital period.
-Regarding our simulations, since the mass and radius of the host are held constant, the mean velocity and velocity dispersion are the same across all packets.
-Secondly, the density of each packet decreases since the easier-to-remove stars are stripped early on.
-}
-
-  \salvatore{Consider an impact that happens at $x\prime$.
-The distance from the impact site to each packet is: $\delta x= x\prime - \langle v\rangle \left(t-n T\right)$, where $n$ is the packet index, with $n=0$ being the first packet escaped at the first pericenter passage in the simulation.
-Note that $\delta x\propto n$, and velocity segregation means $v\propto x$.
-Therefore, the different parts of the separate packets are hit at a given impact site.
-The earlier packets are hit in regions where the stars have lower velocities than the later packets.
-As a result, the gaps in the later packets quickly overtake those from the earlier packets.
-As a result, the gaps across the packets go out of phase and interfere destructively.}
-
-  \salvatore{In the simplest consideration, of two packets of the same dimension $\sigma_x$ and an offset of $\delta x$, the gaps are out of phase when the separation is greater than the width: $1\sim<\sigma_x/\delta x$.
-Drawing from the wisdom in \citet{2015MNRAS.450.1136E} \& \citet{2016MNRAS.457.3817S}, we see the contrast between the density peaks and the density minimum of a gap, as well as the width, grow in time, somewhere between $\propto \sqrt{t}$ and linear.
-In this simplified 1D collisionless dispersion analysis, the gaps between the different packets space apart linearly in time with $\delta x\propto \delta vt$, where $\delta v$ is the mean velocity of the about the impact site for two different packets.
-This inclines us to believe that the gaps space apart before they become significantly deep or wide, however this need not be the general case.
-}
-
-    \salvatore{Nonetheless, we believe this is the first consideration of decomposing the stream into packets of tidal bursts to understand the effect of the internal dynamics on gap persistence.
-Since gap persistence depends on the mass and internal evolution of the cluster, characterizing the gap can constrain our knowledge of the internal evolution.
-}
-
     \begin{figure}
@@
 
+    Each group can be simplified as one-dimensional collisionless dispersion.
+These packets of stars can be described as a population with the same initial position that drifts about a mean velocity, $\langle v \rangle$, with velocity dispersion, $\sigma_v$.
+The density evolution is a Gaussian whose amplitude decreases inversely with time, and whose standard deviation grows linearly with time:
+    \begin{equation}    
+          \rho(x,t) = \frac{1}{\sigma_v t \sqrt{2\pi} }\textrm{exp}\left(-\frac{1}{2}\left(\frac{x-\langle v \rangle t}{\sigma_v t}\right)^2\right).
+    \end{equation}
+
+    \salvatore{It is important to note that within a single packet, the stars self-segregate.
+The position of the density peak is at $\langle v \rangle t$, while the faster stars are ahead and the slower ones are behind.
+The packets are separated by a distance of $\langle  v\rangle T$, where $T$ is the orbital period.
+Regarding our simulations, since the mass and radius of the host are held constant, the mean velocity and velocity dispersion are the same across all packets.
+Secondly, the density of each packet decreases since the easier-to-remove stars are stripped early on.
+}
+
+    \salvatore{Consider an impact that happens at $x\prime$.
+The distance from the impact site to each packet is: $\delta x= x\prime - \langle v\rangle \left(t-n T\right)$, where $n$ is the packet index, with $n=0$ being the first packet escaped at the first pericenter passage in the simulation.
+Note that $\delta x\propto n$, and velocity segregation means $v\propto x$.
+Therefore, the different parts of the separate packets are hit at a given impact site.
+The earlier packets are hit in regions where the stars have lower velocities than the later packets.
+As a result, the gaps in the later packets quickly overtake those from the earlier packets.
+As a result, the gaps across the packets go out of phase and interfere destructively.}
+
+	  \salvatore{The aforementioned analysis highlights the complexity of the conditions on the stream that are necessary for allowing gap formation.
+Despite this, in the attempts to gain further preliminary insight, we can create an even simpler consideration of a gap created in two different pockets, considering the gaps' drift rates and width, ignoring their density contrasts.
+If we consider two packets with the same dimensions $\sigma_x$ and an offset of $\delta x$, the gaps are out of phase when the separation is greater than the width: $1\sim<\sigma_x/\delta x$.
+If $\delta x$ grows much faster than the gap can develop, then the gap is never visible.
+If the gaps grow in size much faster than their drift rate, then they persist and continue to grow in significance.
+In the simulations presented in this analysis, we only observe the two extremes and do not produce a case where a significant gap is created to blur out.
+However, this is not necessarily the general case.}
+
+    \salvatore{Nonetheless, this is a first look at considering the distribution of stream stars originating from shocked outflow, and they allow or disallow gap formation.
+Fuller considerations would include mass loss from internal dynamics, such as three-body encounters, and a non-constant mass loss rate due to changes in the density distribution resulting from internal mass evolution.}
+
   \section{Gallery of Gaps} \label{sec:gallery_of_gaps}

### 2025-05-23 paper/main.tex
removed name for highlighting green text

@@
 
-    \textbf{\salvatore{An additional explanation for the absence of gaps in the inner regions of a stream is given in App.~\ref{sec:gaps_vs_gcmass}.
+    An additional explanation for the absence of gaps in the inner regions of a stream is given in App.~\ref{sec:gaps_vs_gcmass}.
 Briefly, the eccentricity of an orbit induces tidal shocking at the pericenter passages that cause episodes of increased mass loss where the escaped stars leave with a higher mean velocity and greater velocity dispersion.
@@
 
-  \textbf{ \salvatore{\citet{2017MNRAS.470...60E} is another work that discusses the impact of globular clusters in producing gaps on the Pal~5 stream.
+  \citet{2017MNRAS.470...60E} is another work that discusses the impact of globular clusters in producing gaps on the Pal~5 stream.
 In section~6.3 of their article, \citet{2017MNRAS.470...60E} discuss the fact that globular clusters with masses greater than $10^6~\textrm{M}_\odot$ are rare in the Galaxy, especially at distances compatible with the orbit of Pal~5, while -- based on previous works -- they estimate that sub-halos of dark matter of similar mass are at least three times greater in number.
 This difference leads them to conclude that the gaps they report in the Pal~5 streams are more likely to be induced by dark matter sub-halos than by globular clusters.
+
   However, as we show in this paper, even clusters with masses below $10^6~\textrm{M}\odot$ can produce gaps.
@@
 
-    \salvatore{Fig.~\ref{fig:decoherence_shocked_outflow} dissects the more massive stream of Fig.~\ref{fig:gap_no_more} and instructs us why the gap does not persist.
+    Fig.~\ref{fig:decoherence_shocked_outflow} dissects the more massive stream of Fig.~\ref{fig:gap_no_more} and instructs us why the gap does not persist.
 The bottom panel shows the total density distribution, and dissects it into two profiles: stars that either are \textit{leaked} or \textit{shocked} out of the cluster.
@@
 The duration of each shock is: $t_{\textrm{shock}}$ = 2$\sqrt{r/|\nabla\Phi|}$, where $r$ is the galactocentric radius and $|\nabla\Phi|$ is the magnitude of the acceleration from the galactic potential.
-Stars within this range about a given pericenter passage are all given the same color and shown as individual groups in the first, second, third, and fourth panels of Fig.~\ref{fig:decoherence_shocked_outflow}.}
+Stars within this range about a given pericenter passage are all given the same color and shown as individual groups in the first, second, third, and fourth panels of Fig.~\ref{fig:decoherence_shocked_outflow}.
 
@@
 
-    \salvatore{It is important to note that within a single packet, the stars self-segregate.
+    It is important to note that within a single packet, the stars self-segregate.
 The position of the density peak is at $\langle v \rangle t$, while the faster stars are ahead and the slower ones are behind.
@@
 Secondly, the density of each packet decreases since the easier-to-remove stars are stripped early on.
-}
-
-    \salvatore{Consider an impact that happens at $x\prime$.
+
+    Consider an impact that happens at $x\prime$.
 The distance from the impact site to each packet is: $\delta x= x\prime - \langle v\rangle \left(t-n T\right)$, where $n$ is the packet index, with $n=0$ being the first packet escaped at the first pericenter passage in the simulation.
@@
 As a result, the gaps in the later packets quickly overtake those from the earlier packets.
-As a result, the gaps across the packets go out of phase and interfere destructively.}
-
-	  \salvatore{The aforementioned analysis highlights the complexity of the conditions on the stream that are necessary for allowing gap formation.
+As a result, the gaps across the packets go out of phase and interfere destructively.
+
+	  The aforementioned analysis highlights the complexity of the conditions on the stream that are necessary for allowing gap formation.
 Despite this, in the attempts to gain further preliminary insight, we can create an even simpler consideration of a gap created in two different pockets, considering the gaps' drift rates and width, ignoring their density contrasts.
@@
 In the simulations presented in this analysis, we only observe the two extremes and do not produce a case where a significant gap is created to blur out.
-However, this is not necessarily the general case.}
-
-    \salvatore{Nonetheless, this is a first look at considering the distribution of stream stars originating from shocked outflow, and they allow or disallow gap formation.
-Fuller considerations would include mass loss from internal dynamics, such as three-body encounters, and a non-constant mass loss rate due to changes in the density distribution resulting from internal mass evolution.}
+However, this is not necessarily the general case.
+
+    Nonetheless, this is a first look at considering the distribution of stream stars originating from shocked outflow, and they allow or disallow gap formation.
+Fuller considerations would include mass loss from internal dynamics, such as three-body encounters, and a non-constant mass loss rate due to changes in the density distribution resulting from internal mass evolution.
 
